### 2018-07-16 data-science-math/Homework 1.Rmd
Initial commit

---
title: "CUNY SPSS Data Science Math Homework 1"
author: "Mike Silva"
date: "July 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Profit is the difference between revenue and cost.  Revenue is the quantity sold (q) multiplied by the price(p) and the costs is given at $10 times the quantity sold, thus
$$P(q,p) = R(q,p) - C(q) = qp - 10q$$
The quantity sold is given as a function of the price
$$q(p) = 80-2p$$
So rewriting the equation for profit as a function of the price
$$P(p) = R(p) - C(p) = (80-2p)p - 10(80-2p) = (80p-2p^2) - (800-20p) = -2p^2 + 100 - 800 $$
The retailer's profit function is
$$P(p) = -2p^2 + 100 - 800 $$

## Question 2
Given $f(x)=8x^3+7x^2-5$:

### Step 1
$$f(3) = 8(3)^3+7(3)^2-5 = (8*27)+(7*9)-5 = 216+63-5 = 274$$

### Step 2
$$f(-1) = 8(-2)^3+7(-2)^2-5 = (8*-8)+(7*4)-5 = -64+28-5 = -41$$

### Step 3
$$f(x+c) = 8(x+c)^3+7(x+c)^2-5$$

### 2018-07-17 data-science-math/Homework 1.Rmd
Additional work

@@
 ```{r setup, include=FALSE}
+library(Deriv)
 knitr::opts_chunk$set(echo = TRUE)
@@
 
-## Question 1
+**Question 1**
 
@@
 So rewriting the equation for profit as a function of the price
-$$P(p) = R(p) - C(p) = (80-2p)p - 10(80-2p) = (80p-2p^2) - (800-20p) = -2p^2 + 100 - 800 $$
-The retailer's profit function is
-$$P(p) = -2p^2 + 100 - 800 $$
+$$P(p) = R(p) - C(p)$$
+$$ (80-2p)p - 10(80-2p)$$
+$$ (80p-2p^2) - (800-20p)$$
+$$-2p^2 + 100p - 800 $$
 
-## Question 2
+**Question 2**
+
 Given $f(x)=8x^3+7x^2-5$:
 
-### Step 1
-$$f(3) = 8(3)^3+7(3)^2-5 = (8*27)+(7*9)-5 = 216+63-5 = 274$$
+**Step 1**
+$$f(3) = 8(3)^3+7(3)^2-5 = (8\cdot27)+(7\cdot9)-5 = 216+63-5 = 274$$
 
-### Step 2
-$$f(-1) = 8(-2)^3+7(-2)^2-5 = (8*-8)+(7*4)-5 = -64+28-5 = -41$$
+**Step 2**
+$$f(-1) = 8(-2)^3+7(-2)^2-5 = (8\cdot-8)+(7\cdot4)-5 = -64+28-5 = -41$$
 
-### Step 3
+**Step 3**
 $$f(x+c) = 8(x+c)^3+7(x+c)^2-5$$
+Verifying my math with R:
+```{r}
+f2 <- function(x){8*x**3+7*x**2-5}
+f2(3)
+f2(-2)
+```
+
+**Question 3 **
+
+**Step 1**
+$$\lim_{x\to1^-} f(x) = 2$$
+
+**Step 2**
+$$\lim_{x\to1^+} f(x) = -5$$
+
+**Step 3**
+$$\lim_{x\to1} f(x) = \mbox{ Does not exist} $$
+
+**Question 4**
+Find the derivative of $f(x)=-2x^3$
+
+$$f'(x)=-6x^2$$
+Verifying my math in R:
+```{r}
+f4 <- function(x){-2*x**3}
+Deriv(f4)
+```
+
+**Question 5**
+$$f(x) = \frac{-8}{x^2}$$
+$$f'(x) = \frac{16}{x^3}$$
+Verify with R:
+```{r}
+f5 <-function(x){-8/(x**2)}
+Deriv(f5)
+```
+
+**Question 6**
+$$g(x) 5\sqrt[3]{x}$$
+
+$$g'(x) \frac{5}{3x^\frac{2}{3}}=\frac{5}{3}\cdot\frac{1}{x^\frac{2}{3}}$$
+```{r}
+f6 <- function(x){5*x**(1/3)}
+Deriv(f6)
+
+5/3
+
+2/3
+```
+
+**Question 7**
+
+$$y=-2x^\frac{9}{8}$$
+
+$$y'=\frac{-9x^\frac{1}{8}}{4}=\frac{-9}{4}\cdot x^\frac{1}{8}$$
+```{r}
+f7 <- function(x){-2*x**(9/8)}
+Deriv(f7)
+
+-9/4
+1/8
+```
+
+**Question 8**
+$$\frac{35-40}{4-0}=\frac{-5}{4} = -1\frac{1}{4}$$
+
+**Question 9**
+$$C(x)=630+2.4x$$
+$$\mbox{Avg Cost} = \frac{\mbox{Cost}}{\mbox{Unit}}=\frac{C(x)}{x}=\frac{630+2.4x}{x}$$
+
+**Question 10**
+$$f(x)=(-2x^{-2} +1)(-5x + 9)$$
+
+Applying the quotient rule where $f(x)=-2x^{-2}+1$ and $g(x)=(-5x+9)$
+
+$$f'(x) = \frac{4}{x^3}(-5x+9) + -5(2x^{-2} + 1) = \frac{36}{x^3} - \frac{20}{x^2} + \frac{10}{x^2} - 5=\frac{36}{x^3}-\frac{10}{x^2}-5$$
+Checking my math with R:
+```{r}
+f10 <- function(x){(-2*x**-2+1)*(-5*x+9)}
+Deriv(f10)
+# Evaluating at 10
+f10_prime <- function(x){(36/x**3)-(10/x**2)-5}
+f10_prime(10)
+f10_prime <- Deriv(f10)
+f10_prime(10)
+```
+
+**Question 11**
+$$f(x)=\frac{5x^\frac{1}{2}+7}{x^3 + 1}$$
+
+Applying the quotient rule where $f(x)=5\sqrt{x}$ and $g(x)=x^3+1$
+
+$$f'(x) = \frac{\Big(\frac{5}{2\sqrt{x}}\Big)(x^3+1)-(3x^2(5x^\frac{1}{2}+7))}{(x^3+1)^2}$$
+Simplified to 
+$$f'(x) = -\frac{42x^\frac{5}{2} + 25x^3 - 5}{2\sqrt{x}\cdot (x^6+2x^3 + 1)}$$
+Again checking with R:
+```{r}
+f11 <- function(x){(7+5*x**(1/2))/(1+x**3)}
+Deriv(f11)
+# Evaluate at 5
+fprime11 <- function(x){(-42*x**(5/2)+25*x**3-5)/((2*x**(1/2))*(x**6+2*x**3+1))}
+fprime11(5)
+fprime11 <- Deriv(f11)
+fprime11(5)
+```
+
+**Question 12**
+$$f(x)=(3x^{-3}-8x+6)^\frac{4}{3}$$
+
+$$f'(x) = \frac{4}{3}(-\frac{9}{x^4}-8) \Big(\frac{3}{x^3}-8x+6\Big)^\frac{1}{3}$$
+
+Checking my math in R:
+```{r}
+f12 <- function(x){(3*x**-3-(8*x)+6)**(4/3)}
+Deriv(f12)
+# Evaluate at 1
+f12_prime <- function(x){(4/3)*((-9/x**4)-8)*(((3/x**3)-8*x+6)**(1/3))}
+f12_prime(1)
+f12_prime <- Deriv(f12)
+f12_prime(1)
+```
+
+*Question 13*
+$$f(t) =\frac{550t^2}{\sqrt{t^2+15}}$$
+
+$$f'(t)=$$
+```{r}
+f13 <- function(t){(550*t**2)/(t+15)**(1/2)}
+Deriv(f13)
+```

### 2018-07-17 data-science-math/Homework 1.Rmd
Additonal work on assignment

@@
 
-**Question 1**
+### Question 1
 
@@
 
-**Question 2**
+******
+
+### Question 2
 
@@
 
-**Step 1**
+#### Step 1
+
 $$f(3) = 8(3)^3+7(3)^2-5 = (8\cdot27)+(7\cdot9)-5 = 216+63-5 = 274$$
 
-**Step 2**
+#### Step 2
+
 $$f(-1) = 8(-2)^3+7(-2)^2-5 = (8\cdot-8)+(7\cdot4)-5 = -64+28-5 = -41$$
 
-**Step 3**
+#### Step 3
+
 $$f(x+c) = 8(x+c)^3+7(x+c)^2-5$$
@@
 
-**Question 3 **
+******
 
-**Step 1**
+### Question 3
+
+#### Step 1
+
 $$\lim_{x\to1^-} f(x) = 2$$
 
-**Step 2**
+#### Step 2
+
 $$\lim_{x\to1^+} f(x) = -5$$
 
-**Step 3**
+#### Step 3
+
 $$\lim_{x\to1} f(x) = \mbox{ Does not exist} $$
 
-**Question 4**
+******
+
+### Question 4
+
 Find the derivative of $f(x)=-2x^3$
@@
 
-**Question 5**
+******
+
+### Question 5
+
 $$f(x) = \frac{-8}{x^2}$$
@@
 
-**Question 6**
+******
+
+### Question 6
 $$g(x) 5\sqrt[3]{x}$$
@@
 
-**Question 7**
+******
+
+### Question 7
 
@@
 
-**Question 8**
+******
+
+### Question 8
+
 $$\frac{35-40}{4-0}=\frac{-5}{4} = -1\frac{1}{4}$$
 
-**Question 9**
+******
+
+### Question 9
+
 $$C(x)=630+2.4x$$
@@
 
-**Question 10**
+******
+
+### Question 10
+
 $$f(x)=(-2x^{-2} +1)(-5x + 9)$$
@@
 
-**Question 11**
+******
+
+### Question 11
+
 $$f(x)=\frac{5x^\frac{1}{2}+7}{x^3 + 1}$$
 
-Applying the quotient rule where $f(x)=5\sqrt{x}$ and $g(x)=x^3+1$
+Applying the quotient rule where $f(x)=5\sqrt{x}+7$ and $g(x)=x^3+1$ so $f'(x)=\frac{5}{2\sqrt{x}}$ and $g'(x)=3x^2$
 
-$$f'(x) = \frac{\Big(\frac{5}{2\sqrt{x}}\Big)(x^3+1)-(3x^2(5x^\frac{1}{2}+7))}{(x^3+1)^2}$$
+$$f'(x) = \frac{\Big(\frac{5}{2\sqrt{x}}\Big)(x^3+1)-(3x^2(5\sqrt{x}+7))}{(x^3+1)^2}$$
+$$f'(x) = \frac{\frac{5x^3+5}{2\sqrt{x}}-3x^2 (5\sqrt{x}+7)}{(x^3+1)^2}$$
+
 Simplified to 
-$$f'(x) = -\frac{42x^\frac{5}{2} + 25x^3 - 5}{2\sqrt{x}\cdot (x^6+2x^3 + 1)}$$
+$$f'(x) = -\frac{42x^\frac{5}{2} + 25x^3 - 5}{2\sqrt{x}\cdot (x^3 + 1)^2}$$
 Again checking with R:
 ```{r}
-f11 <- function(x){(7+5*x**(1/2))/(1+x**3)}
+f11 <- function(x){(5*x**(1/2)+7)/(x**3+1)}
 Deriv(f11)
 # Evaluate at 5
-fprime11 <- function(x){(-42*x**(5/2)+25*x**3-5)/((2*x**(1/2))*(x**6+2*x**3+1))}
+fprime11 <- function(x){-1*(42*x**(5/2)+25*x**3-5)/((2*x**(1/2))*((x**3+1)**2))}
 fprime11(5)
@@
 
-**Question 12**
+******
+
+### Question 12
+
 $$f(x)=(3x^{-3}-8x+6)^\frac{4}{3}$$
@@
 
-*Question 13*
+******
+
+### Question 13
+
 $$f(t) =\frac{550t^2}{\sqrt{t^2+15}}$$
 
-$$f'(t)=$$
+After factoring out the 550 constant and applying the quotient rule where $f(x)=t^2$ and $g(x)=\sqrt{t^2+15}$
+
+$$550\cdot f'(x) = \frac{(2t)(\sqrt{t^2+15})-(t^2)\Big(\frac{t}{\sqrt{t^2+15}}\Big)}{t^2+15} = \frac{2t\sqrt{t^2+15} -\frac{t^3}{\sqrt{t^2+15}}}{t^2+15}$$
+
+Multiplying by 550 and simplifying
+$$f'(x) = \frac{1100t\sqrt{t^2+15} -\frac{550t^3}{\sqrt{t^2+15}}}{t^2+15}=\frac{550t(t^2+30)}{(t^2+15)^\frac{3}{2}}$$
 ```{r}
-f13 <- function(t){(550*t**2)/(t+15)**(1/2)}
+f13 <- function(t){(550*t**2)/((t**2+15)**(1/2))}
 Deriv(f13)
+f13_prime <- function(t){(550*t*(t**2+30))/((t**2+15)**(3/2))}
+f13_prime(3)
+f13_prime <- Deriv(f13)
+f13_prime(3)
+
 ```
+
+The answer is 547
+
+******
+
+### Question 14
+
+Given $N(t) = 1000(6+0.1t)^\frac{1}{2}$
+$$N(3) = 1000(6+0.1(3))^\frac{1}{2}=39690$$

### 2018-07-18 data-science-math/Homework 1.Rmd
Additional work

@@
 $$f'(x) = \frac{16}{x^3}$$
+
 Verify with R:
@@
 $$g'(x) \frac{5}{3x^\frac{2}{3}}=\frac{5}{3}\cdot\frac{1}{x^\frac{2}{3}}$$
+
+Verify with R:
 ```{r}
@@
 $$y'=\frac{-9x^\frac{1}{8}}{4}=\frac{-9}{4}\cdot x^\frac{1}{8}$$
+
+Verify with R:
 ```{r}
@@
 $$f'(x) = \frac{4}{x^3}(-5x+9) + -5(2x^{-2} + 1) = \frac{36}{x^3} - \frac{20}{x^2} + \frac{10}{x^2} - 5=\frac{36}{x^3}-\frac{10}{x^2}-5$$
-Checking my math with R:
+
+Verify with R:
 ```{r}
@@
 $$f'(x) = -\frac{42x^\frac{5}{2} + 25x^3 - 5}{2\sqrt{x}\cdot (x^3 + 1)^2}$$
+
 Again checking with R:
@@
 $$f'(x) = \frac{1100t\sqrt{t^2+15} -\frac{550t^3}{\sqrt{t^2+15}}}{t^2+15}=\frac{550t(t^2+30)}{(t^2+15)^\frac{3}{2}}$$
+
+Verifying with R:
 ```{r}
@@
 f13_prime(3)
-
 ```
@@
 
-Given $N(t) = 1000(6+0.1t)^\frac{1}{2}$
+$$N(t) = 1000(6+0.1t)^\frac{1}{2}=1000\sqrt{0.1t+6}$$
+
+#### Step 1
 $$N(3) = 1000(6+0.1(3))^\frac{1}{2}=39690$$
+
+#### Step 2
+Let $u=0.1t+6$ so $\sqrt{u}$ becomes $\frac{1}{2\sqrt{u}}$ because of the power rule
+
+and due to the chain rule $u'=0.1$.  Substitute
+
+$$1000\cdot N'(t) = \frac{0.1\cdot0.5}{\sqrt{0.1t+6}}=\frac{0.05}{\sqrt{0.1t+6}}$$
+$$ N'(t)=\frac{1000\cdot 0.05}{\sqrt{0.1t+6}}=\frac{50}{\sqrt{0.1t+6}}$$
+$$N'(5)=\frac{50}{\sqrt{0.1(5)+6}}=\frac{50}{\sqrt{6.5}}=19.6116135138\approx20	$$
+20 is roughly how many additional people attend in the 5th week of the season
+
+Checking my math with R:
+```{r}
+f14 <- function(t){1000*(6+0.1*t)**(1/2)}
+Deriv(f14)
+f14_prime <- function(t){50/(6.5**(1/2))}
+f14_prime(5)
+f14_prime <- Deriv(f14)
+f14_prime(5)
+```
+
+******
+
+### Question 15
+
+$$3x^3+4y^3=77$$
+
+#### Step 1
+
+$$\frac{d}{dx}(3x^3+4y^3)=\frac{d}{dx}(77)$$
+
+$$\frac{d}{dx}(3x^3)+\frac{d}{dx}(4y^3)=0$$
+$$9x^2+12y^2\frac{dy}{dx}=0$$
+
+$$12y^2\frac{dy}{dx}=-9x^2$$
+
+$$\frac{dy}{dx}=\frac{-9x^2}{12y^2} = -\frac{3x^2}{4y^2}$$
+
+#### Step 2
+
+The slope of the tangent line at (3,-1) is
+
+$$-\frac{3(3)^2}{4(-1)^2}=-\frac{3(9)}{4(1)}=-\frac{7}{4}$$
+
+******
+
+### Question 16
+
+$$f(x)=\frac{x+3}{x-8}$$
+
+Differentiate using the quotient rule where $f(x)=x+3$ and $g(x)=x-8$
+
+$$f'(x)=\frac{(1)(x-8)-(1)(x+3)}{(x-8)^2}=\frac{(x-8)-(x+3)}{(x-8)^2}=\frac{x-8-x-3}{(x-8)^2}=-\frac{11}{(x-8)^2}$$
+
+Checking my work with R:
+```{r}
+f16 <- function(x){(x+3)/(x-8)}
+Deriv(f16)
+f16_prime <- function(x){-11/((x-8)^2)}
+f16_prime(20)
+f16_prime <- Deriv(f16)
+f16_prime(20)
+```
+
+$$f''(x)= \frac{22}{(x-8)^3}$$
+
+The critical point is when x=8.  f(x) is decreasing when x < 8 and increasing when x > 8.
+
+******
+
+### Question 17
+
+$$F(t) = 14 + \frac{367t^2}{t^2+100}$$
+
+#### Step 1
+
+$$367 \cdot F'(t) = \frac{t^2}{(t^2+100)^2}=\frac{(2t(t^2+100))(-2t^2\cdot t)}{(t^2+100)^2}=\frac{2t(t^2+100)(-2t^3)}{(t^2+100)^2}$$
+
+Multiply it by 367
+
+$$f'(x)=367 \cdot \frac{2t(t^2+100)(-2t^3)}{(t^2+100)^2}=\frac{734t(t^2+100)-734t^3}{(t^2+100)^2}=\frac{73400t}{(t^2+100)^2}$$
+
+In the interest of turning my assignment on time some of the steps have been excluded
+
+$$f''(x)=-\frac{73400(3t^2-100)}{(t^2+100)^3}$$
+
+$F'(t) > 0$ from $(0,\infty)$ and $F'(t) < 0$ from $(-\infty, 0)$ which doesn't make sense in this case.
+
+#### Step 2
+
+$$\lim_{t\to\infty} F(t) = 381^\circ F$$
+
+******
+
+### Question 18
+
+$$f(x)=\frac{x^3}{3340000}-\frac{7x^2}{9475}+\frac{42417727x}{1265860000}+\frac{1}{33}$$
+
+Again in the interest of time I will skip the steps of the calculation:
+
+$$f'(x) = \frac{3x^2}{3340000}-\frac{14x}{9475}+\frac{42417727}{1265860000}$$
+
+$$0 = \frac{3x^2}{3340000}-\frac{14x}{9475}+\frac{42417727}{1265860000}$$
+
+```{r}
+f18 <- function(x){}
+f18_prime <- function(x){(3*x**2/3340000)-(14*x/9475)+(42417727/1265860000)}
+x <- seq(0, 30, by=.01)
+y <- f18_prime(x)
+i <- stats::uniroot(f18_prime,c(-30,30))
+plot(y~x,type="l", col="blue")
+abline(h=0, col="black")
+abline(v=0, col="black")
+abline(v=i$root, col="red")
+points(i$root, 0, col="red", pch=21)
+```
+
+```{r}
+i$root
+```
+
+******
+
+### Question 19
+
+$$f(x) = 7x^2+28x-35$$
+
+$$f'(x) = 14x+28$$
+
+$$0 = 14x+28 $$
+
+$$-28 = 14x$$
+
+$$x=-\frac{28}{14}=-2$$
+
+$$f(-2) = (7(-2)^2)+(28(-2))-35 = 28-56-35=-63$$
+
+$$\mbox minimum = (-2,-63)$$
+
+Checking with R:
+```{r}
+f19 <- function(x){7*x**2+28*x-35}
+f19_prime <- Deriv(f19)
+i <- stats::uniroot(f19_prime,c(-30,30))
+i$root
+f19(i$root)
+```
+
+```{r}
+x <- seq(-30, 30, by=.01)
+y <- f19(x)
+i <- stats::uniroot(f19_prime,c(-30,30))
+plot(y~x,type="l", col="blue")
+abline(h=0, col="black")
+abline(v=0, col="black")
+points(i$root, 0, col="red", pch=21)
+```
+
+******
+
+### Question 20
+
+$$f(x)=-6x^3+27x^2+180x$$
+
+$$f'(x)=-18x^2+54x+180$$
+
+$$0=-18x^2+54x+180$$
+
+$$0=-18(x^2-3x-10)$$
+
+$$x^2-3x-10=0$$
+
+$$(x+2)(x-5)=0$$
+
+$$x=-2 and 5$$
+
+******

### 2018-07-18 data-science-math/Homework 1.Rmd
More work

@@
 
-$$x=-2 and 5$$
+The function is minimized/maximized when $x = -2$ or $x = 5$.  Plugging in and solving for y:
+
+$$f(-2)=-6(-2)^3+27(-2)^2+180(-2)=48+108-360=-204$$
+$$f(-2)=-6(5)^3+27(5)^2+180(5)=-750+650+900=800$$
+
+$$f''(x)=-36x+54$$
+$$f''(-2)=-36(-2)+54=72+54=126>0$$
+$$f''(5)=-36(5)+54=180+54=234>0$$
+
+So the critical points are $(-2, -204)$ and $(5, 800)$.  $(-2, -204)$ is a local min and $(5, 800)$
+
+```{r}
+f20 <- function(x){(-6*x**3)+(27*x**2)+(180*x)}
+f20_prime <- Deriv(f20)
+x <- seq(-30, 30, by=.01)
+y <- f20(x)
+#i <- stats::uniroot(f20_prime,c(-30,30))
+plot(y~x,type="l", col="blue")
+abline(h=0, col="black")
+abline(v=0, col="black")
+#points(i$root, 0, col="red", pch=21)
+```
 

### 2018-07-19 data-science-math/Homework 1.Rmd
More work

@@
 ******
+
+### Question 21
+Inventory cost is defined as the cost of holding stock and the cost of ordering.  The inventory cost as a function of the iron lot size is:
+
+$$C(i)=\frac{120}{i}(6+4.5(i))+(1.60\frac{i}{2})$$
+This function assumes the irons are selling at a constant rate (no seasonality) throughout the year.  Simplifing the function:
+
+$$C(i)=\frac{620}{i}+ \frac{540i}{i}+(0.8i) = \frac{620}{i}+0.8i+540$$
+
+Since we want to minimize this cost we take the first derivative and set it equal to zero.
+
+$$C'(i)=-\frac{720}{i^2} +0.8$$
+
+$$0=-\frac{720}{i^2} +0.8$$
+$$\frac{720}{i^2}=0.8$$
+$$720=0.8i^2$$
+$$i^2=\frac{720}{0.8}=900$$
+$$i=\sqrt{900}=30$$
+So the **lot size is 30 irons**.  Since we are expecting to sell 120 irons per year this would require **4 orders**.
+
+******
+
+### Question 22
+
+$$Volume = 18432ft^2 = x\cdot x \cdot y \Rightarrow y=\frac{18432}{x^2}$$
+The cost is the sum of the sides, top and bottom.  The sides and top material cost $3.00 per square foot and the bottom is $5.00.
+
+$$C(x,y) = 3(4xy) + 3x^2 + 5x^2 = 12xy+8x^2$$
+Substituting in y:
+$$ C(x)=\frac{12x\cdot 18432}{x^2}+8x^2=\frac{221184}{x}+8x^2$$
+Minimize the cost function:
+$$ C'(x)=-\frac{221184}{x^2}+16x$$
+$$ 0=-\frac{221184}{x^2}+16x\Rightarrow 0=-221184+16x^3 \Rightarrow 0=-13824 + x^3 $$
+$$ x^3 = 13824 \Rightarrow x=\sqrt[3]{13824} = 24$$
+Plugging 24 into the volume equation and solving for y:
+$$18432 = x^2y\Rightarrow 18432=(24)^2 y\Rightarrow 18432=576 y \Rightarrow y=\frac{18432}{576} \Rightarrow y = 32$$
+Now checking everything with R:
+```{r}
+f22 <- function(x){
+  (3*(4*x*(18432/(x**2)))) + (3*x**2) + (5*x**2)
+}
+f22_prime <- Deriv(f22)
+library(rootSolve)
+x22 <- uniroot.all(f22_prime, c(0,100))
+x22
+y22 <- function(x){18432/(x**2)}
+y22(x22)
+```
+
+******
+
+### Question 23
+
+******
+
+### Question 24
+
+******
+
+### Question 25
+
+******
+
+### Question 26
+
+******
+
+### Question 27
+
+******
+
+### Question 28
+
+******
+
+### Question 29
+
+******
+### Question 30
+
+$$ \int_{-7}^2 x\sqrt{x+7}dx$$
+
+For this problem $u=x$ and $dv=\sqrt{x+7}dx$.  Calculate $du$ and $v$:
+
+$du=1dx$ and $v = \int \sqrt{x+7}dx = \frac{2}{3}(x+7)^\frac{3}{2}$
+
+Plugging it into the integration by parts formula:
+
+$$ \int_a^b udv = uv]_a^b -\int_a^b vdu$$
+
+$$ \int_{-7}^2x\sqrt{x+7}dx]_{-7}^2 -\int_{-7}^2 \frac{2}{3}(x+7)^\frac{3}{2}\cdot 1dx$$
+Integrate the right hand side
+
+$$ \frac{2x}{3}(x+7)^\frac{3}{2}]_{-7}^2 -\frac{4}{15}(x+7)^\frac{5}{2}]_{-7}^2 $$
+
+Plugging in the -7 and 2 yeilds -28.8 as shown below:
+```{r}
+f30 <- function(a, b){
+  left <- function(x){
+    ((2*x)/3)*(x+7)**(3/2)
+  }
+  right <- function(y){
+    (4/15)*(y+7)**(5/2)
+  }
+  (left(b)-left(a))-(right(b)-right(a))
+}
+
+f30(-7,2)
+```
+******
+
+### Question 31
+
+#### Step 1
+In the interest of time I will solve this question by building a simple model
+```{r}
+bounce <- function(total_times, drop_height, debug=F){
+  rebound = .22 # Amount of the rebound as a share of drop height
+  times_ball_hit_the_ground = 1
+  total_distance = drop_height
+  if(debug){
+    message(paste(total_distance, 'meters after ball hits the ground',times_ball_hit_the_ground,'times'))
+  }
+  while(times_ball_hit_the_ground < total_times){
+    drop_height = rebound * drop_height 
+    total_distance = total_distance + (2 * drop_height) #going Up and down
+    times_ball_hit_the_ground = times_ball_hit_the_ground + 1
+    if(debug){message(paste(total_distance, 'meters after ball hits the ground',times_ball_hit_the_ground,'times'))}
+  }
+  if(!debug){
+  message(paste(total_distance, 'meters after ball hits the ground',times_ball_hit_the_ground,'times'))
+  }
+}
+
+bounce(20, 46, T)
+```
+So the ball traveled 71.94872 meters when it hit the ground on the 20th time.  I would note that the additional distance is very small after the 7th
+
+#### Step 2
+To evaluate the total distance as the number of bounces heads to infinity I will cheat and plug in a really large number and see what it tends toward.
+```{r}
+bounce(1000000, 46)
+```
+
+So it is tending toward 72 meters.
+
+******
+
+### Question 34

### 2018-07-20 data-science-math/Homework 1.Rmd
More work on HW 1

@@
 library(Deriv)
+library(rootSolve)
 knitr::opts_chunk$set(echo = TRUE)
@@
 f22_prime <- Deriv(f22)
-library(rootSolve)
+
 x22 <- uniroot.all(f22_prime, c(0,100))
@@
 
+$$Area=xy=1056 \Rightarrow y=\frac{1056}{x}$$
+
+The cost function is
+
+$$C(x,y)=(14.40\cdot 2y)+(14.40 \cdot 2x) + (12.00 \cdot 2x)$$
+Substituting y in:
+$$C(x)=(14.40\cdot 2(\frac{1056}{x}))+(14.40 \cdot 2x) + (12.00 \cdot 2x)$$
+Simplified
+$$C(x)=\frac{30412.8}{x}+52.8x$$
+
+Minimize it
+
+$$C'(x)=-\frac{30412.8}{x^2}+52.8$$
+
+$$0=-\frac{30412.8}{x^2}+52.8 $$
+$$\frac{30412.8}{x^2}=52.8 $$
+$$ \frac{30412.8}{52.8}=x^2 $$
+$$x^2=576 $$
+$$ x=\sqrt{576} \Rightarrow x=\pm 24$$
+Negative 24 is meaningless in this context.  Now substituting x in the area equation:
+$$y=\frac{1056}{24} = 44$$
+
+```{r}
+f23 <- function(x){
+  (28.8*(1056/x))+(52.8*x)
+}
+f23_prime <- Deriv(f23)
+
+x23 <- uniroot.all(f23_prime, c(0,100))
+x23
+y23 <- function(x){1056/x}
+y23(x23)
+```
+
 ******
@@
 ### Question 24
+
+$$Value(time)=base \cdot e^{rate \cdot time}$$
+Enter in the information from the question:
+
+$$37000=67000 \cdot e^{rate \cdot 7}$$
+
+Solving for rate:
+
+$$\frac{37000}{67000} = e^{rate \cdot 7}$$
+
+$$ln\Big(\frac{37000}{67000}\Big) = rate \cdot 7 \Rightarrow rate = \frac{ln\Big(\frac{37000}{67000}\Big)}{7} \approx -0.08482496$$
+
+Now substitue and solve for the value at year 9
+
+$$Value(9) = 67000 \cdot e^{-0.08482496 \cdot (9)} = 31226.53$$
+Checking my math with R:
+```{r}
+f24 <- function(t){
+  e <- exp(1)
+  67000*e**(t*((log(37000/67000)/7)))
+}
+f24(9)
+```
 
@@
 
+Given $p = D(x) = 23.2-0.4x$ price and Revenue is $R = p \cdot D(x)$ the maximizing level of production is:
+
+$$ R = p \cdot D(x) \Rightarrow R(x) = D(x) \cdot D(x) \Rightarrow R(x) = D(x)^2$$
+
+Using the chain rule:
+$$R'(x) = 0.32x -18.56$$
+
+$$0 = 0.32x -18.56 \Rightarrow 0.32x = 18.56 \Rightarrow x = \frac{18.56}{0.32} = 58$$
+
+Checking the math in R:
+```{r}
+f25 <- function(x){
+  (23.2-0.4*x)*(23.2-0.4*x)
+}
+
+f25_prime <- Deriv(f25)
+x25 <- uniroot.all(f25_prime, c(0,100))
+x25
+```
+
 ******
@@
 ### Question 26
+Start with the exponential growth formula:
+$$Cost(time)=base \cdot e^{rate \cdot time}$$
+
+And enter what we know and solve for rate:
+$$426.8=400 \cdot e^{rate \cdot 11} \Rightarrow \frac{426.80}{400}=e^{rate \cdot 11} \Rightarrow rate\cdot 11 = ln\Big(\frac{426.8}{400}\Big) \Rightarrow rate = \frac{ln\Big(\frac{426.8}{400}\Big)}{11} \approx 0.005895543$$
+
+Substitute rate back into the equation:
+$$Cost(year) = 400 \cdot e^{0.005895543 \cdot (year - 1995)}$$
+$$Cost(2017) = 400 \cdot e^{0.005895543 \cdot (2017 - 1995)} \approx 455.40$$
+Verifying with R:
+```{r}
+f26 <- function(year){
+  e <- exp(1)
+  400*e**((log(426.8/400)/11)*(year-1995))
+}
+f26(2017)
+```
 
@@
 ### Question 27
+Given:
+$$P'(x) = 380-4x $$
+$$P(x) = 380x - 2x^2+C$$
+
+#### Step 1
+
+$$1700 = 380(38)-2(38)^2 + C$$
+Solving for C
+
+$$ 1700 = 14440 - 2888 + C \Rightarrow C = 17000 - 14440 + 2888 \Rightarrow C = -9852$$
+The profit function is:
+$$ P(x)= 380x - 2x^2-9852$$
+
+#### Step 4
+$$ P(56)= 380(56) - 2x^2-9852 = 5156$$
+```{r}
+f27 <- function(x){
+  (380*x) - (2*x**2) - 9852
+}
+f27(56)
+```
 
@@
 ### Question 28
+Solve:
+$$ \int\frac{-5(ln(y))^3}{y}dy = -5 \int \frac{1}{y} ln(y)^3$$
+Substitute
+$u = ln(y)$ and $du = \frac{1}{y}$
+
+$$-5 \int du \cdot u^3 \Rightarrow -5 \cdot \frac{u^4}{4}+c \Rightarrow \frac{-5}{4}ln(y)^4 + c$$
 
@@
 
+Given:
+$$P'(t)=75-9t^\frac{1}{2}$$
+
+$$P(t) = 75t-6t^\frac{3}{2}$$
+
+$$\int_0^9 P'(t) = P(9) - P(0) = (75(9)-6(9)^\frac{3}{2})-(0-0) = (675-162) -0 = 513$$
+
 ******
+
 ### Question 30
@@
 ```
+
 ******
@@
 
+### Question 32
+
+Checking my math with R:
+```{r}
+f32 <- function(x){x*((x+7)**(1/2))}
+integrate(Vectorize(f32),-7, 2)
+```
+
+******
+
 ### Question 34
+
+I currently don't know how to do find the Taylor polnomial of degree 5 near x=4 for $y=3e^{5x-3}$

### 2018-07-20 data-science-math/Homework 2.Rmd
HW2 Work

---
title: "CUNY SPSS Data Science Math Homework 2"
author: "Mike Silva"
date: "July 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(pander)
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1

There are 540 identical plastic chips numbered 1 through 540 in a box. What is the probability of reaching into the box and randomly drawing the chip numbered 505? Express your answer as a fraction or a decimal number rounded to four decimal places.

$$A=\frac{\{505\}}{\{1,2...539,540\}}=\frac{1}{540}\approx 0.0019$$

```{r}
1/540
```

****

### Question 2

Write out the sample space for the given experiment. Separate your answers using commas.  When deciding what you want to put into a salad for dinner at a restaurant, you will choose one of the following extra toppings: asparagus, cheese. Also, you will add one of following meats: eggs, turkey. Lastly, you will decide on one of the
following dressings: French, vinaigrette. (Note: Use the following letters to indicate each choice: A for asparagus, C for cheese, E for eggs, T for turkey, F for French, and V for vinaigrette.)

$$\{AEF, ATF, AEV, AEF, CEF, CTF, CEV, CEF\}$$

****

### Question 3

A card is drawn from a standard deck of 52 playing cards. What is the probability that the card will be a heart and not a face card? Write your answer as a fraction or a decimal number rounded to four decimal places.

$$P(h\cap nf)= \frac{52 \cdot \frac{1}{4}}{52-12}=\frac{13}{40}=0.325$$

```{r}
13/40
```

****

### Question 4

A standard pair of six-sided dice is rolled. What is the probability of rolling a sum less than 6? Write your answer as a fraction or a decimal number rounded to four decimal places.

$$P(l) =\frac{\{1,2,3,4,5\}}{\{1,2,3,4,5,6\}}=5/6\approx0.8333$$
```{r}
5/6
```

****

### Question 5

A pizza delivery company classifies its customers by gender and location of residence. The research department has gathered data from a random sample of 2001 customers. The data is summarized in the table below.

```{r q5, echo=FALSE}
q4_data <- data.frame(h=c('Apartment','Dorm','With Parent(s)','Sorority/Fraternity Home', 'Other'), Male = c(233, 159, 102, 220, 250), Female=c(208, 138, 280, 265, 146))
names(q4_data) <- c('','Male','Female')
emphasize.strong.cols(1)
```

```{r q5_table, results='asis'}
pandoc.table(q4_data)
```

What is the probability that a customer is male? Write your answer as a fraction or a decimal number rounded to four decimal places.

$$P(m) = \frac{233+159+102+120+250}{233+159+102+220+250+208+138+280+265+146} =\frac{964}{2001} \approx 0.4818 $$
```{r}
sum(q4_data$Male)/(sum(q4_data$Male)+sum(q4_data$Female))
```

****

### Question 6

Three cards are drawn with replacement from a standard deck. What is the probability that the first card will be a club, the second card will be a black card, and the third card will be a face card? Write your answer as a fraction or a decimal number rounded to four decimal places.

$$P(fbc) = \frac{12}{52} \cdot \frac{26}{52} \cdot \frac{13}{52} \approx 0.0288 $$
```{r}
12/52 * 26/52 * 13/52
```

****

### Question 7

Two cards are drawn without replacement from a standard deck of 52 playing cards. What is the probability of choosing a spade for the second card drawn, if the first card, drawn without replacement, was a heart? Write your answer as a fraction or
a decimal number rounded to four decimal places.

$$P(s|h) = \frac{13}{51} \cdot \frac{13}{52} \approx 0.0637$$
```{r}
13/51 * 13/52
```

****

### Question 8

Two cards are drawn without replacement from a standard deck of 52 playing cards. What is the probability of choosing a heart and then, without replacement, a red card? Write your answer as a fraction or a decimal number rounded to four decimal places.

$$P(r|h) = \frac{25}{51} \cdot \frac{13}{52} \approx -.1225$$
```{r}
25/51 * 13/52
```

****

### Question 9

There are 85 students in a basic math class. The instructor must choose two students at random.

```{r q9, echo=FALSE}
q9_data <- data.frame(c=c('Freshman','Sophomores','Juniors','Seniors'), Male = c(12, 19, 12, 7), Female=c(12, 15, 4, 4))
names(q9_data) <- c('','Male','Female')
emphasize.strong.cols(1)
```

```{r q9_table, results='asis'}
pandoc.table(q9_data)
```

What is the probability that a junior female and then a freshmen male are chosen at random? Write your answer as a fraction or a decimal number rounded to four decimal places.

$$P(fm|jf) = \frac{12}{84} \cdot \frac{4}{85} \approx 0.0068$$
```{r}
total_students <- sum(q9_data$Male) + sum(q9_data$Female)
12/(total_students-1) * 4/total_students
```

*****

### Question 10

Out of 300 applicants for a job, 141 are male and 52 are male and have a graduate degree.

#### Step 1.

What is the probability that a randomly chosen applicant has a graduate degree, given that they are male? Enter your answer as a fraction or a decimal rounded to four decimal places.

$$P(g|m) = \frac{52}{141} \approx 0.3688$$
```{r}
52/141
```

#### Step 2.

If 102 of the applicants have graduate degrees, what is the probability that a randomly chosen applicant is male, given that the applicant has a graduate degree? Enter your answer as a fraction or a decimal rounded to four decimal places.

$$P(m|g) = \frac{52}{102} \approx 0.5098$$
```{r}
52/102
```

*****

### Question 11

A value meal package at Ron's Subs consists of a drink, a sandwich, and a bag of chips. There are 6 types of drinks to choose from, 5 types of sandwiches, and 3 types of chips. How many different value meal packages are possible?

****

### Question 12

A doctor visits her patients during morning rounds. In how many ways can the doctor visit 5 patients during the morning rounds?

****

### Question 13

A coordinator will select 5 songs from a list of 8 songs to compose an event's musical entertainment lineup. How many different lineups are possible?

****

### Question 14
A person rolls a standard six-sided die 9 times. In how many ways can he get 3 fours, 5 sixes and 1 two?

*****

### Question 15

How many ways can Rudy choose 6 pizza toppings from a menu of 14 toppings if each topping can only be chosen once?

*****

### Question 16

3 cards are drawn from a standard deck of 52 playing cards. How many different 3-card hands are possible if the drawing is done without replacement?

*****

### Question 17

You are ordering a new home theater system that consists of a TV, surround sound system, and DVD player. You can choose from 12 different TVs, 9 types of surround sound systems, and 5 types of DVD players. How many different home theater systems can you build?

*****

### Question 18

You need to have a password with 5 letters followed by 3 odd digits between 0 - 9 inclusively. If the characters and digits cannot be used more than once, how many choices do you have for your password?

*****

### Question 19

Evaluate the following expression.

$_9 P_4$

*****

### Question 20

Evaluate the following expression.

$_11 C_8$

*****

### Question 21

Evaluate the following expression.

$\frac{_{12} P_8}{_{12} C_4}$

*****

### Question 22

The newly elected president needs to decide the remaining 7 spots available in the cabinet he/she is appointing. If there are 13 eligible candidates for these positions (where rank matters), how many different ways can the members of the cabinet be
appointed?

****

### Question 23

In how many ways can the letters in the word 'Population' be arranged?

### Question 24

Consider the following data:
𝒙𝒙 5 6 7 8 9
𝒑𝒑(𝒙𝒙) 0.1 0.2 0.3 0.2 0.2

#### Step 1

Find the expected value $E(x)$ Round your answer to one decimal place.

#### Step 2

Find the variance. Round your answer to one decimal place.

#### Step 3

Find the standard deviation. Round your answer to one decimal place.

#### Step 4

Find the value of $P(X\hat e9)$. Round your answer to one decimal place.

#### Step 5

Find the value of $P(X\c{C} 7)$. Round your answer to one decimal place.

*****

### Question 25

Suppose a basketball player has made 188 out of 376 free throws. If the player makes the next 3 free throws, I will pay you 23 dollars. Otherwise you pay me 4 dollars.

#### Step 1

Find the expected value of the proposition. Round your answer to two decimal places.

#### Step 2

If you played this game 994 times how much would you expect to win or lose? (Losses must be entered as negative.)

****

### Question 26

Flip a coin 11 times. If you get 8 tails or less, I will pay you 1 dollar. Otherwise you pay me 7 dollars.

#### Step 1

Find the expected value of the proposition. Round your answer to two decimal places.

#### Step 2

If you played this game 615 times how much would you expect to win or lose? (Losses must be entered as negative.)

*****

### Question 27

If you draw two clubs on two consecutive draws from a standard deck of cards you win 583 dollars. Otherwise you pay me 35 dollars. (Cards drawn without replacement.)

#### Step 1

Find the expected value of the proposition. Round your answer to two decimal places.

#### Step 2

If you played this game 632 times how much would you expect to win or lose? (Losses must be entered as negative.)

*****

### Question 28

A quality control inspector has drawn a sample of 10 light bulbs from a recent production lot. If the number of defective bulbs is 2 or less, the lot passes inspection. Suppose 30% of the bulbs in the lot are defective. What is the probability that the lot will pass inspection? (Round your answer to 3 decimal places)

*****

### Question 29

A quality control inspector has drawn a sample of 5 light bulbs from a recent production lot. Suppose that 30% of the bulbs in the lot are defective. What is the expected value of the number of defective bulbs in the sample? Do not round your
answer.

*****

### Question 30

The auto parts department of an automotive dealership sends out a mean of 5.5 special orders daily. What is the probability that, for any day, the number of special orders sent out will be more than 5? (Round your answer to 4 decimal places)

*****

### Question 31

At the Fidelity Credit Union, a mean of 5.7 customers arrive hourly at the drive-through window. What is the probability that, in any hour, more than 4 customers will arrive? (Round your answer to 4 decimal places)

*****

### Question 32

The computer that controls a bank's automatic teller machine crashes a mean of 0.4 times per day. What is the probability that, in any 7-day week, the computer will crash no more than 1 time? (Round your answer to 4 decimal places)

*****

### Question 33

A town recently dismissed 8 employees in order to meet their new budget reductions. The town had 6 employees over 50 years of age and 19 under 50. If the dismissed employees were selected at random, what is the probability that more than 1
employee was over 50? Write your answer as a fraction or a decimal number rounded to three decimal places.

*****

### Question 34

Unknown to a medical researcher, 10 out of 25 patients have a heart problem that will result in death if they receive the test drug. Eight patients are randomly selected to receive the drug and the rest receive a placebo. What is the probability that less
than 7 patients will die? Write your answer as a fraction or a decimal number rounded to three decimal places

*****

### 2018-07-20 data-science-math/Homework 2.Rmd
More progress on week 2

@@
 library(pander)
+library(tidyr)
+library(gtools)
 knitr::opts_chunk$set(echo = TRUE)
@@
 
+```{r}
+drinks <- c(1,2,3,4,5,6)
+sandwiches <- c(1,2,3,4,5)
+chips <- c(1,2,3)
+
+q11_data <- tidyr::crossing(drinks, sandwiches) %>%
+  tidyr::crossing(., chips)
+nrow(q11_data)
+```
+
 ****
@@
 
+$$ P(n) = n! \Rightarrow P(5)=5! = 5 \times 4 \times 3 \times 2 \times 1 = 120 $$
+```{r}
+factorial(5)
+
+patients <- c(1,2,3,4,5)
+q12_data <- permutations(length(patients), length(patients), patients)
+nrow(q12_data)
+```
+
 ****
@@
 A coordinator will select 5 songs from a list of 8 songs to compose an event's musical entertainment lineup. How many different lineups are possible?
+
+$$ C(n,r) = \frac{n!}{r!(n-r)!} \Rightarrow C(8,5) = \frac{8!}{5!(8-5)!} =  \frac{40320}{120\cdot 3!} = \frac{40320}{120\cdot 6} = \frac{40320}{720} = 56$$
+
+```{r}
+q13_data <- combinations(8,5)
+nrow(q13_data)
+```
 

### 2018-07-21 data-science-math/Homework 2.Rmd
More work

@@
 library(pander)
+library(dplyr)
 library(tidyr)
@@
 $$\{AEF, ATF, AEV, AEF, CEF, CTF, CEV, CEF\}$$
+```{r}
+toppings <- c('A','C')
+meats <- c('E','T')
+dressings <- c('F','V')
+
+q2_data <- crossing(toppings, meats) %>%
+  crossing(., dressings) %>%
+  mutate(combinations = paste0(toppings, meats, dressings))
+
+as.vector(q2_data$combinations)
+```
 
@@
 
-$$P(r|h) = \frac{25}{51} \cdot \frac{13}{52} \approx -.1225$$
+$$P(r|h) = \frac{25}{51} \cdot \frac{13}{52} \approx 0.1225$$
 ```{r}
@@
 
+$$6 \times 5 \times 3 = 90$$
+
 ```{r}
@@
 
+$$P(9) = 9! = 362880$$
+```{r}
+die_rolls <- c('4.1','4.2','4.3','6.1','6.2','6.3','6.4','6.5','1')
+q14_data <- permutations(length(die_rolls), length(die_rolls), die_rolls)
+nrow(q14_data)
+```
+
 *****
@@
 
+$$ C(14,6) = \frac{14!}{6!(14-6)!} =  \frac{87178291200}{720\cdot 8!} = \frac{87178291200}{720\cdot 40320} = \frac{87178291200}{29030400} = 3003$$
+
+```{r}
+q15_data <- combinations(14,6)
+nrow(q15_data)
+```
+
 *****
@@
 3 cards are drawn from a standard deck of 52 playing cards. How many different 3-card hands are possible if the drawing is done without replacement?
+
+$$ C(52,3) = \frac{52!}{3!(52-3)!} =  \frac{52!}{3!\cdot 49!} = 22100$$
+
+```{r}
+q16_data <- combinations(52,3)
+nrow(q16_data)
+```
 
@@
 
+$$ 12 \times 9 \times 5 = 540$$
+
+```{r}
+tvs <- c(1,2,3,4,5,6,7,8,9,10,11,12)
+surround_sound_systems <- c(1,2,3,4,5,6,7,8,9)
+dvd_players <- c(1,2,3,4,5)
+
+q17_data <- tidyr::crossing(tvs, surround_sound_systems) %>%
+  tidyr::crossing(., dvd_players)
+nrow(q17_data)
+```
+
 *****
@@
 If the characters and digits cannot be used more than once, how many choices do you have for your password?
+
+$$ C(26,5) = \frac{26!}{5!(26-5)!} =  \frac{26!}{5!\cdot 21!} = 65780$$
+
+$$ C(5,3) = \frac{5!}{3!(5-3)!} =  \frac{5!}{3!\cdot 2!} = 10$$
+$$65780 \times 10 = 657800$$
+```{r}
+letters <- combinations(26,5)
+numbers <- combinations(5,3)
+nrow(letters) * nrow(numbers)
+```
 
@@
 In how many ways can the letters in the word 'Population' be arranged?
+
+$$ P(10) = 10! = 3628800 $$
+```{r}
+letters <- c('P','o','p','u','l','a','t','i','o','n')
+factorial(length(letters))
+```
+
+****
 

### 2018-07-21 data-science-math/Homework 2.Rmd
More progress on HW2

@@
 
-$$A=\frac{\{505\}}{\{1,2...539,540\}}=\frac{1}{540}\approx 0.0019$$
+$$P=\frac{\{505\}}{\{1,2...539,540\}}=\frac{1}{540}\approx 0.0019$$
 
@@
 
-$$P(h\cap nf)= \frac{52 \cdot \frac{1}{4}}{52-12}=\frac{13}{40}=0.325$$
+$$P(hnf)= \frac{52 \cdot \frac{1}{4}}{52-12}=\frac{13}{40}=0.325$$
 
@@
 
-```{r q5, echo=FALSE}
-q4_data <- data.frame(h=c('Apartment','Dorm','With Parent(s)','Sorority/Fraternity Home', 'Other'), Male = c(233, 159, 102, 220, 250), Female=c(208, 138, 280, 265, 146))
-names(q4_data) <- c('','Male','Female')
-emphasize.strong.cols(1)
-```
+----------------------------------------------
+            &nbsp;              Male   Female 
+------------------------------ ------ --------
+        **Apartment**           233     208
 
-```{r q5_table, results='asis'}
-pandoc.table(q4_data)
-```
+           **Dorm**             159     138
+
+      **With Parent(s)**        102     280
+
+ **Sorority/Fraternity Home**   220     265
+
+          **Other**             250     146   
+----------------------------------------------
 
@@
 ```{r}
+q4_data <- data.frame(h=c('Apartment','Dorm','With Parent(s)','Sorority/Fraternity Home', 'Other'), Male = c(233, 159, 102, 220, 250), Female=c(208, 138, 280, 265, 146))
 sum(q4_data$Male)/(sum(q4_data$Male)+sum(q4_data$Female))
@@
 
-```{r q9, echo=FALSE}
-q9_data <- data.frame(c=c('Freshman','Sophomores','Juniors','Seniors'), Male = c(12, 19, 12, 7), Female=c(12, 15, 4, 4))
-names(q9_data) <- c('','Male','Female')
-emphasize.strong.cols(1)
-```
+--------------------------------
+     &nbsp;       Male   Female 
+---------------- ------ --------
+  **Freshman**     12      12
 
-```{r q9_table, results='asis'}
-pandoc.table(q9_data)
-```
+ **Sophomores**    19      15
+
+  **Juniors**      12      4
+
+  **Seniors**      7       4    
+--------------------------------
 
@@
 ```{r}
+q9_data <- data.frame(c=c('Freshman','Sophomores','Juniors','Seniors'), Male = c(12, 19, 12, 7), Female=c(12, 15, 4, 4))
+names(q9_data) <- c('','Male','Female')
 total_students <- sum(q9_data$Male) + sum(q9_data$Female)
@@
 
+$$ C(13,7) = \frac{13!}{7!(13-7)!} =  \frac{6227020800}{5040\cdot 6!} = \frac{6227020800}{5040\cdot 720} = \frac{6227020800}{3628800} = 1716$$
+
+```{r}
+q22_data <- combinations(13,7)
+nrow(q22_data)
+```
+
 ****
@@
 Consider the following data:
-𝒙𝒙 5 6 7 8 9
-𝒑𝒑(𝒙𝒙) 0.1 0.2 0.3 0.2 0.2
+
+----------------------------------------
+  &nbsp;    NA    NA    NA    NA    NA  
+---------- ----- ----- ----- ----- -----
+  **x**      5     6     7     8     9
+
+ **P(x)**   0.1   0.2   0.3   0.2   0.2 
+----------------------------------------
 
@@
 Find the expected value $E(x)$ Round your answer to one decimal place.
+$$E(x) = \Sigma {xP(x)}= (5 \times 0.1)+(6 \times 0.2)+(7 \times 0.3)+(8 \times 0.2)+(9 \times 0.2) = 7.2$$
+```{r}
+x <- c(5, 6, 7, 8, 9)
+px <-c(0.1, 0.2, 0.3,  0.2, 0.2)
+ex <- sum(x*px)
+ex
+```
 
@@
 
+$$\sigma^2=\frac{\Sigma (x-E(x))^2}{N} = \frac{(5-7.2)^2+(6-7.2)^2+(7-7.2)^2+(8-7.2)^2+(9-7.2)^2}{5}$$
+$$\sigma^2= \frac{4.84+1.44+0.04+0.64+3.24}{5}=\frac{10.2}{5}=2.04$$
+
+```{r}
+v <- sum((x-ex)**2)/length(x)
+v
+```
 #### Step 3
@@
 Find the standard deviation. Round your answer to one decimal place.
+$$\sigma = \sqrt{\sigma ^2} = \sqrt{2.04} \approx 1.4$$
+```{r}
+v**(1/2)
+```
 
-#### Step 4
+### Step 4
 

### 2018-07-22 data-science-math/Homework 2.Rmd
Fixed table in question 24
Finished Homework 2

@@
 
-$$P(l) =\frac{\{1,2,3,4,5\}}{\{1,2,3,4,5,6\}}=5/6\approx0.8333$$
+$$P(l) =\frac{\{1,2,3,4,5\}}{\{1,2,3,4,5,6\}}=\frac{5}{6}\approx0.8333$$
 ```{r}
@@
 
+&nbsp;
+
 #### Step 1.
@@
 ```
+
+&nbsp;
 
@@
 
-q11_data <- tidyr::crossing(drinks, sandwiches) %>%
-  tidyr::crossing(., chips)
+q11_data <- crossing(drinks, sandwiches) %>%
+  crossing(., chips)
 nrow(q11_data)
@@
 
-$_9 P_4$
+$$_9 P_4$$
+$$_9 P_4=\frac{9!}{(9-4)!}=\frac{362880}{120}=3024$$
+```{r}
+N <- 9
+n <- 4
+factorial(N)/factorial((N-n))
+```
 
@@
 
-$_11 C_8$
+$$_{11}C_8$$
+
+$$_{11}C_8=\frac{11!}{8!\cdot(11-8)!}=\frac{11!}{8!\cdot3!}=\frac{39916800}{403208 \cdot 6}=165$$
+
+```{r}
+N <- 11
+n <- 8
+factorial(N)/(factorial(n)*factorial(N-n))
+```
 
@@
 
-$\frac{_{12} P_8}{_{12} C_4}$
+$$\frac{_{12} P_8}{_{12} C_4}$$
+
+$$_{12}P_8 = \frac{12!}{(12-8)!}=\frac{479001600}{4!}=\frac{479001600}{24}=19958400$$
+
+$$ _{12}C_4=\frac{12!}{4!\cdot(12-4)!}=\frac{479001600}{24\cdot8!}=\frac{479001600}{24\cdot 40320}=\frac{479001600}{967680}=495$$
+$$\frac{_{12} P_8}{_{12} C_4}=\frac{19958400}{495}=40320$$
+
+```{r}
+N <- 12
+n1 <- 8
+n2 <- 4
+q21_p <- factorial(N)/factorial((N-n1))
+q21_c <- factorial(N)/(factorial(n2)*factorial(N-n2))
+q21_p/q21_c
+```
 
@@
 ----------------------------------------
-  &nbsp;    NA    NA    NA    NA    NA  
+
 ---------- ----- ----- ----- ----- -----
@@
 ----------------------------------------
+
+&nbsp;
 
@@
 
+&nbsp;
+
 #### Step 2
@@
 
-$$\sigma^2=\frac{\Sigma (x-E(x))^2}{N} = \frac{(5-7.2)^2+(6-7.2)^2+(7-7.2)^2+(8-7.2)^2+(9-7.2)^2}{5}$$
-$$\sigma^2= \frac{4.84+1.44+0.04+0.64+3.24}{5}=\frac{10.2}{5}=2.04$$
+$$\sigma^2=\Sigma_i (x_i-E(x))^2P(x_i) $$
+$$\sigma^2= 0.1(5-7.2)^2+0.2(6-7.2)^2+0.3(7-7.2)^2+0.2(8-7.2)^2+0.2(9-7.2)^2$$
+$$\sigma^2= (0.1\times 4.84)+(0.2 \times 1.44)+(0.3 \times 0.04)+(0.2 \times 0.64)+(0.2 \times 3.24)$$
+$$\sigma^2=0.484+0.288+0.012+0.128+0.648\approx1.6$$
 
 ```{r}
-v <- sum((x-ex)**2)/length(x)
+v <- sum(px*(x-ex)**2)
 v
 ```
+
+&nbsp;
+
 #### Step 3
@@
 Find the standard deviation. Round your answer to one decimal place.
-$$\sigma = \sqrt{\sigma ^2} = \sqrt{2.04} \approx 1.4$$
+$$\sigma = \sqrt{\sigma ^2} \approx \sqrt{1.6} \approx 1.3$$
 ```{r}
@@
 ```
+The difference is due the the fact that I used 1.6 as the variance instead of 1.56.
+
+&nbsp;
 
@@
 
-Find the value of $P(X\hat e9)$. Round your answer to one decimal place.
+Find the value of $P(X\geq9)$. Round your answer to one decimal place.
+
+The answer is **0.2** because this is the only P(x) where x >= 9.
+
+```{r}
+sum(px[x>=9])
+```
+
+&nbsp;
 
@@
 
-Find the value of $P(X\c{C} 7)$. Round your answer to one decimal place.
+Find the value of $P(X\leq7)$. Round your answer to one decimal place.
+
+$$0.1 + 0.2 + 0.3= 0.6$$
+
+```{r}
+sum(px[x<=7])
+```
 
@@
 
+&nbsp;
+
 #### Step 1
@@
 
+The player shooting $\frac{188}{376}$ free throws so they have a 50% probability of making the shot.
+We can now estimate the probability of them making the next 3 shots as $\frac{1}{8}$.
+The probability that they won't is $\frac{7}{8}$.
+
+The expected value of the proposition is:
+
+$$P(\mbox{winning bet}) \times \mbox{winnings } - P(\mbox{not winning bet}) \times \mbox{loss}$$
+
+$$ \frac{1}{8}\times 23 - \frac{7}{8} \times 4 = \frac{23}{8}-\frac{28}{8} = -\frac{5}{8} \approx -0.63$$
+```{r}
+makes_it <-c(1,0)
+
+q25_data <- crossing(makes_it, makes_it) %>%
+  crossing(., makes_it)
+
+q25_data$shots_made <- rowSums(q25_data)
+
+made_all_3 <- 1
+n <- nrow(q25_data)
+
+pwin <- made_all_3/n
+ploose <- 1 - pwin
+
+ex25 <- (pwin * 23) - (ploose * 4)
+ex25
+```
+
+&nbsp;
+
 #### Step 2
@@
 If you played this game 994 times how much would you expect to win or lose? (Losses must be entered as negative.)
+
+$$994 \times -0.63 = -626.22$$
+
+```{r}
+994 * ex25
+```
 
@@
 
+```{r}
+flipped_tails <- c(1,0)
+
+q26_data <- crossing(flipped_tails, flipped_tails) %>%
+  crossing(., flipped_tails) %>%
+  crossing(., flipped_tails) %>%
+  crossing(., flipped_tails) %>%
+  crossing(., flipped_tails) %>%
+  crossing(., flipped_tails) %>%
+  crossing(., flipped_tails) %>%
+  crossing(., flipped_tails) %>%
+  crossing(., flipped_tails) %>%
+  crossing(., flipped_tails)
+
+q26_data$total_tails <- rowSums(q26_data)
+
+tails_8_or_less <- q26_data %>%
+  filter(total_tails < 9) %>%
+  nrow(.)
+n <- nrow(q26_data)
+
+pwin <- tails_8_or_less/n
+ploose <- 1 - pwin
+
+ex26 <- (pwin * 1 )-(ploose * 7)
+ex26
+```
+
+&nbsp;
+
 #### Step 2
@@
 If you played this game 615 times how much would you expect to win or lose? (Losses must be entered as negative.)
+
+```{r}
+615 * ex26
+```
 
@@
 
+$$P(c|c) = \frac{12}{51} \cdot \frac{13}{52} \approx 0.059$$
+
+$$ (0.059 \times 583)-((1-0.59)\times 35) \approx 1.46$$
+
+```{r}
+pwin <- (12/51)*(13/52)
+ploose <- 1 - pwin
+ex27 <- (pwin * 583)-(ploose * 35)
+ex27
+```
+
+The difference is due to the rounding of the probability of wining, but I will use this figure for step 2
+
+&nbsp;
+
 #### Step 2
@@
 If you played this game 632 times how much would you expect to win or lose? (Losses must be entered as negative.)
+
+$$632 \times 1.35 = 853.20$$
+```{r}
+632 * ex27
+```
+
+Again the difference is due to rounding.  **Lesson: Don't use rounded values when calculating values (garbage in garbage out).**
 
@@
 
+$$F(x;p,n) = \sum_{i=0}^{x}{\left( \begin{array}{c} n \\ i \end{array}
+ \right) (p)^{i}(1 - p)^{(n-i)}}$$
+
+Not wanting to mess this up I will just us R to solve:
+
+```{r}
+p <- 0.3
+n <- 10
+x <- 2
+dbinom(0,n,p)+dbinom(1,n,p)+dbinom(2,n,p)
+q28 <- pbinom(x, n, p)
+q28
+```
+
+The answer is **0.383**.
+
 *****
@@
 
+Given $ p=0.3$ and $n=5$:
+$$E = p\cdot n = 1.5$$
+
+```{r}
+0.3*5
+```
+
 *****
@@
 (Round your answer to 4 decimal places)
+
+$$F(x;\lambda) = \sum_{i=0}^{x}{\frac{e^{-\lambda}\lambda^{i}} {i!}}$$
+Given the complexity of the formula they will only be solved in R.
+
+```{r}
+x <- 5
+lambda <- 5.5
+1 - ppois(1, lambda)
+```
+
+The answer is **0.9723**.
 
@@
 
+```{r}
+x <- 4
+lambda <- 5.7
+1 - ppois(1, lambda)
+```
+
+The answer is **0.9776**.
+
 *****
@@
 (Round your answer to 4 decimal places)
+
+Since they provide a daily lambda the weekly lambda is $7\times 0.4$ or $2.8$.
+
+```{r}
+x <- 1
+lambda <- 0.4 * 7
+ppois(x, lambda)
+```
+
+The answer is **0.2311**.
 
@@
 
+```{r}
+q <- 1
+m <- 6
+n <- 19
+k <- 8
+1 - phyper(q, m, n, k)
+```
+
+The answer is **0.651**.
+
 *****
@@
 
+```{r}
+q <- 6
+m <- 10
+n <- (25-m)
+k <- 8
+phyper(q, m, n, k)
+```
+
+The answer is **0.998**.
+
 *****

### 2018-07-23 data-science-math/Homework 3.Rmd
Initial commit of HW 3

---
title: "CUNY SPSS Data Science Math Homework 3"
author: "Mike Silva"
date: "July 23, 2018"
output: html_document
---

### Question 1

The weights of steers in a herd are distributed normally. The variance is 40,000 and the mean steer weight is 1300 lbs. Find the probability that the weight of a randomly selected steer is greater than 979 lbs. (Round your answer to 4 decimal places)

### Question 2

SVGA monitors manufactured by TSI Electronics have life spans that have a normal distribution with a variance of 1,960,000 and a mean life span of 11,000 hours. If a SVGA monitor is selected at random, find the probability that the life span of the monitor will be more than 8340 hours. (Round your answer to 4 decimal places)

### Question 3

Suppose the mean income of firms in the industry for a year is 80 million dollars with a standard deviation of 3 million dollars. If incomes for the industry are distributed normally, what is the probability that a randomly selected firm will earn between 83 and 85 million dollars? (Round your answer to 4 decimal places)

### Question 4

Suppose GRE Verbal scores are normally distributed with a mean of 456 and a standard deviation of 123. A university plans to offer tutoring jobs to students whose scores are in the top 14%. What is the minimum score required for the job offer? Round your answer to the nearest whole number, if necessary.

### Question 5

The lengths of nails produced in a factory are normally distributed with a mean of 6.13 centimeters and a standard deviation of 0.06 centimeters. Find the two lengths that separate the top 7% and the bottom 7%. These lengths could serve as limits used to identify which nails should be rejected. Round your answer to the nearest hundredth, if necessary.

### Question 6

An English professor assigns letter grades on a test according to the following scheme.

A: Top 13% of scores
B: Scores below the top 13% and above the bottom 55%
C: Scores below the top 45% and above the bottom 20%
D: Scores below the top 80% and above the bottom 9%
F: Bottom 9% of scores

Scores on the test are normally distributed with a mean of 78.8 and a standard deviation of 9.8. Find the numerical limits for a C grade. Round your answers to the nearest whole number, if necessary.

### Question 7

Suppose ACT Composite scores are normally distributed with a mean of 21.2 and a standard deviation of 5.4. A university plans to admit students whose scores are in the top 45%. What is the minimum score required for admission? Round your answer to the nearest tenth, if necessary.

### Question 8

Consider the probability that less than 11 out of 151 students will not graduate on time. Assume the probability that a given student will not graduate on time is 9%. Approximate the probability using the normal distribution. (Round your answer to 4 decimal places.)

### Question 9

The mean lifetime of a tire is 48 months with a standard deviation of 7. If 147 tires are sampled, what is the probability that the mean of the sample would be greater than 48.83 months? (Round your answer to 4 decimal places)

### Question 10

The quality control manager at a computer manufacturing company believes that the mean life of a computer is 91 months, with a standard deviation of 10. If he is correct, what is the probability that the mean of a sample of 68 computers would be greater than 93.54 months? (Round your answer to 4 decimal places)

### Question 11

A director of reservations believes that 7% of the ticketed passengers are no-shows. If the director is right, what is the probability that the proportion of no-shows in a sample of 540 ticketed passengers would differ from the population proportion by less than 3%? (Round your answer to 4 decimal places)

### Question 12

A bottle maker believes that 23% of his bottles are defective. If the bottle maker is accurate, what is the probability that the proportion of defective bottles in a sample of 602 bottles would differ from the population proportion by greater than 4%? (Round your answer to 4 decimal places)

### Question 13

A research company desires to know the mean consumption of beef per week among males over age 48. Suppose a sample of size 208 is drawn with $\bar{X}$=3.9. Assume å = 0.8 . Construct the 80% confidence interval for the mean number of lb. of beef per week among males over 48. (Round your answers to 1 decimal place)

Lower Bound:
Upper Bound:

### Question 14

An economist wants to estimate the mean per capita income (in thousands of dollars) in a major city in California. Suppose a sample of size 7472 is drawn with $\bar{X}$= 16.6. Assume å = 11 . Construct the 98% confidence interval for the mean per capita income. (Round your answers to 1 decimal place)

Lower Bound:
Upper Bound:

### Question 15

Find the value of t such that 0.05 of the area under the curve is to the left of t. Assume the degrees of freedom equals 26.

#### Step 1

Choose the picture which best describes the problem.

#### Step 2

Write your answer below.

### Question 16

The following measurements ( in picocuries per liter ) were recorded by a set of helium gas detectors installed in a laboratory facility:
 $$383.6, 347.1, 371.9, 347.6, 325.8, 337$$

Using these measurements, construct a 90% confidence interval for the mean level of helium gas present in the facility.  Assume the population is normally distributed.

#### Step 1

Calculate the sample mean for the given sample data. (Round answer to 2 decimal places)

#### Step 2

Calculate the sample standard deviation for the given sample data. (Round answer to 2 decimal places)

#### Step 3

Find the critical value that should be used in constructing the confidence interval. (Round answer to 3 decimal places)

#### Step 4

Construct the 90% confidence interval. (Round answer to 2 decimal places)

Lower Bound:
Upper Bound:

### Question 17

A random sample of 16 fields of spring wheat has a mean yield of 46.4 bushels per acre and standard deviation of 2.45 bushels per acre. Determine the 80% confidence interval for the true mean yield. Assume the population is normally distributed.

#### Step 1

Find the critical value that should be used in constructing the confidence interval. (Round answer to 3 decimal places)

#### Step 2

Construct the 80% confidence interval. (Round answer to 1 decimal place)

Lower Bound:
Upper Bound:

### Question 18

A toy manufacturer wants to know how many new toys children buy each year. She thinks the mean is 8 toys per year. Assume a previous study found the standard deviation to be 1.9. How large of a sample would be required in order to estimate the mean number of toys bought per child at the 99% confidence level with an error of at most 0.13 toys? (Round your answer up to the next integer)

### Question 19

A research scientist wants to know how many times per hour a certain strand of bacteria reproduces. He believes that the mean is 12.6. Assume the variance is known to be 3.61. How large of a sample would be required in order to estimate the mean number of reproductions per hour at the 95% confidence level with an error of at most 0.19 reproductions? (Round your answer up to the next integer)

### Question 20

The state education commission wants to estimate the fraction of tenth grade students that have reading skills at or below the eighth grade level.

#### Step 1

Suppose a sample of 2089 tenth graders is drawn. Of the students sampled, 1734 read above the eighth grade level. Using the data, estimate the proportion of tenth graders reading at or below the eighth grade level. (Write your answer as a fraction or a decimal number rounded to 3 decimal places)

#### Step 2

Suppose a sample of 2089 tenth graders is drawn. Of the students sampled, 1734 read above the eighth grade level. Using the data, construct the 98% confidence interval for the population proportion of tenth graders reading at or below the eighth grade level. (Round your answers to 3 decimal places)

Lower Bound:
Upper Bound:

### Question 21

An environmentalist wants to find out the fraction of oil tankers that have spills each month.

#### Step 1

Suppose a sample of 474 tankers is drawn. Of these ships, 156 had spills. Using the data, estimate the proportion of oil tankers that had spills. (Write your answer as a fraction or a decimal number rounded to 3 decimal places)

#### Step 2

Suppose a sample of 474 tankers is drawn. Of these ships, 156 had spills. Using the data, construct the 95% confidence interval for the population proportion of oil tankers that have spills each month. (Round your answers to 3 decimal places)

Lower Bound:
Upper Bound:

### 2018-07-23 data-science-math/Homework 3.Rmd
Additional Work on HW3

@@
 
+```{r}
+m <- 13000 
+sd <- 40000**(1/2)
+round(1-pnorm(979, m, sd), 4)
+```
+
 ### Question 2
@@
 
+```{r}
+m <- 11000 
+sd <- 1960000 **(1/2)
+round(1-pnorm(8340, m, sd), 4)
+```
+
 ### Question 3
@@
 (Round your answer to 4 decimal places)
+
+```{r}
+m <- 80 
+sd <- 3
+round(1 - pnorm(83, m, sd) - (1-pnorm(85, m, sd)), 4)
+```
 
@@
 
+```{r}
+m <- 456 
+sd <- 123
+p <- 14/100
+round(qnorm(p, m, sd, lower.tail = FALSE), 0)
+```
+
 ### Question 5
@@
 Round your answer to the nearest hundredth, if necessary.
+
+```{r}
+m <- 6.13 
+sd <- 0.06
+p <- 7/100
+round(qnorm(p, m, sd), 3)
+round(qnorm(p, m, sd, lower.tail = FALSE), 3)
+```
 
@@
 
+```{r}
+m <- 78.8 
+sd <- 9.8
+round(qnorm((20/100), m, sd), 0)
+round(qnorm((45/100), m, sd, lower.tail = FALSE), 0)
+```
+
 ### Question 7
@@
 Round your answer to the nearest tenth, if necessary.
+
+```{r}
+m <- 21.2 
+sd <- 5.4
+p <- 45/100
+round(qnorm(p, m, sd, lower.tail = FALSE), 1)
+```
 

### 2018-07-24 data-science-math/Homework 3.Rmd
More work on homework 3

@@
 
+**0.9458**
+
 ```{r}
-m <- 13000 
+m <- 1300 
 sd <- 40000**(1/2)
-round(1-pnorm(979, m, sd), 4)
+x <- 979
+z <- (x-m)/sd
+round(1-pnorm(z),4)
 ```
@@
 
+**0.9713**
+
 ```{r}
@@
 sd <- 1960000 **(1/2)
-round(1-pnorm(8340, m, sd), 4)
+x <- 8340
+z <- (x-m)/sd
+round(1-pnorm(z),4)
+round(1-pnorm(x, m, sd), 4)
 ```
@@
 (Round your answer to 4 decimal places)
+
+**0.1109**
 
@@
 Round your answer to the nearest whole number, if necessary.
+
+**589**
 
@@
 
+**6.04 and 6.22 cm**
+
 ```{r}
@@
 p <- 7/100
-round(qnorm(p, m, sd), 3)
-round(qnorm(p, m, sd, lower.tail = FALSE), 3)
+round(qnorm(p, m, sd), 2)
+round(qnorm(p, m, sd, lower.tail = FALSE), 2)
 ```
@@
 Scores on the test are normally distributed with a mean of 78.8 and a standard deviation of 9.8. Find the numerical limits for a C grade. Round your answers to the nearest whole number, if necessary.
+
+**71 and 80**
 
@@
 
+**21.9**
+
 ```{r}
@@
 
+```{r}
+
+```
+
 ### Question 9
@@
 
+$$ Z = \frac{(X-\mu)}{\sigma}=\frac{(48.83-48)}{7}=0.1185714$$
+
+Looking it up in a Z table the area under the curve (to the left) is 0.5478.  Since we are looking for the area greater than 48.83 we take 1 - 0.5478 = **0.4528**
+
+```{r}
+z <- (48.83-48)/7
+round(1 - pnorm(z),4)
+```
+
 ### Question 10
@@
 (Round your answer to 4 decimal places)
+
+$$ Z = \frac{(X-\mu)}{\sigma}=\frac{(93.54-91)}{10}=0.254$$
+
+**0.4388**
+
+```{r}
+m <- 91
+x <- 92.54
+sd <- 10
+z <- (x-m)/sd
+round(pnorm(z, lower.tail = FALSE), 4)
+```
 
@@
 Lower Bound:
+
 Upper Bound:

### 2018-07-26 data-science-math/Homework 3.Rmd
Additional work

@@
 m <- 1300 
-sd <- 40000**(1/2)
+sd <- sqrt(40000)
 x <- 979
@@
 m <- 11000 
-sd <- 1960000 **(1/2)
+sd <- sqrt(1960000)
 x <- 8340
@@
 
+**0.2307**
+
+$$\mu=N\pi=151\times0.09 = 13.59$$
+
+$$\sigma=\sqrt{N\pi(1-\pi)} = \sqrt{151\times0.09\times 0.91}\approx3.51666$$
+
+$$ Z = \frac{(x-\mu)}{\sigma}=\frac{(11-13.59)}{3.51666}\approx-0.7364942$$
+
 ```{r}
-
+N <- 151
+n <- 11
+p <- 0.09
+m <- N * p
+sd <- sqrt(N*p*(1-p))
+z <- (n-m)/sd
+round(pnorm(z),4)
 ```
@@
 
+$$\mu=N\pi=540\times0.07 = 37.8$$
+
+$$\sigma=\sqrt{N\pi(1-\pi)} = \sqrt{540\times0.07\times 0.93}\approx5.929081$$
+
+$$\bar{x}=N\pi=540\times(0.07-0.03) = 21.6$$
+$$ Z = \frac{(\bar{x}-\mu)}{\sigma}=\frac{(21.6-37.8)}{5.929081}\approx-2.732295$$
+
+**0.0031**
+
+```{r}
+n <- 540
+p <- 0.07
+dif <- -0.03
+m <- n*p
+sd <- sqrt(n*p*(1-p))
+x <- n*(p+dif)
+z <- (x-m)/sd
+round(pnorm(z), 4)
+```
+
 ### Question 12
@@
 
+**0.0098**
+
+```{r}
+n <- 602
+p <- 0.23
+dif <- 0.04 
+m <- n*p
+sd <- sqrt(n*p*(1-p))
+x <- n*(p+dif)
+z <- (x-m)/sd
+round(pnorm(z, lower.tail = FALSE), 4)
+```
+
 ### Question 13
@@
 Suppose a sample of size 208 is drawn with $\bar{X}$=3.9.
-Assume å = 0.8 .
+Assume $\sigma$=0.8.
 Construct the 80% confidence interval for the mean number of lb.
@@
 
-Lower Bound:
+Lower Bound: **3.8**
 
-Upper Bound:
+Upper Bound: **4.0**
+
+I looked up the Z value for 80% in a table which is 1.282.
+
+```{r}
+n <- 208
+x <- 3.9
+sd <- 0.8
+sdm <- sd / sqrt(n) 
+z <- 1.282
+
+round(x - z * sdm, 1)
+
+round(x + z * sdm, 1)
+```
 
@@
 Suppose a sample of size 7472 is drawn with $\bar{X}$= 16.6.
-Assume å = 11 .
+Assume $\sigma$= 11 .
 Construct the 98% confidence interval for the mean per capita income.
@@
 
-Lower Bound:
+Lower Bound: **16.3**
 
-Upper Bound:
+Upper Bound: **16.9**
+
+```{r}
+n <- 7472
+x <- 16.6
+sd <- 11
+sdm <- sd / sqrt(n) 
+z <- 2.326
+
+round(x - z * sdm, 1)
+
+round(x + z * sdm, 1)
+```
 
@@
 
+```{r echo=FALSE}
+library(ggplot2)
+funcShaded <- function(x) {
+  df <- 26
+  y <- dt(x, df)
+  t_val <- qt(0.05, df)
+  y[x > t_val] <- NA
+  return(y)
+}
+ggplot(data.frame(x = c(-4, 4)), aes(x = x)) +
+  stat_function(fun = dt, args = list(df = 26)) + 
+  stat_function(fun=funcShaded, geom="area", fill="#84CA72", alpha=0.5) + 
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank())
+```
+
 #### Step 2
@@
 Write your answer below.
+
+**-1.705618**
+
+```{r}
+qt(0.05, 26)
+```
 
@@
 
+**352.17**
+
+```{r}
+obs <- c(383.6, 347.1, 371.9, 347.6, 325.8, 337)
+n <- length(obs)
+m <- mean(obs)
+round(m, 2)
+```
+
 #### Step 2
@@
 Calculate the sample standard deviation for the given sample data. (Round answer to 2 decimal places)
+
+**21.68**
+
+```{r}
+sdm = sd(obs)
+round(sdm, 2)
+```
 
@@
 
+**1.476**
+
+```{r}
+df <- n - 1
+t <- qt(.9, df)
+round(t, 3)
+```
+
 #### Step 4
@@
 
-Lower Bound:
+Lower Bound: **320.18**
 
-Upper Bound:
+Upper Bound: **384.16**
+
+```{r}
+round(m-t*sdm, 2)
+round(m+t*sdm, 2)
+```
 
@@
 
+**0.866**
+
+```{r}
+n <- 16
+m <- 46.4
+sdm <- 2.45
+df <- n - 1
+t <- qt(.8, df)
+round(t, 3)
+```
+
 #### Step 2
@@
 
-Lower Bound:
+Lower Bound: **44.28**
 
-Upper Bound:
+Upper Bound: **48.52**
+
+```{r}
+round(m-t*sdm, 2)
+round(m+t*sdm, 2)
+```
 
@@
 
+**1417**
+
+```{r}
+ci <- 0.99
+z <- qnorm(1-(1-ci)/2)
+sigma <- 1.9
+moe <- 0.13
+n <- ((z * sigma)/moe)**2
+round(n, 0)
+```
+
 ### Question 19
@@
 (Round your answer up to the next integer)
+
+**384**
+
+```{r}
+m <- 12.6
+sigma <- sqrt(3.61)
+
+ci <- 0.95
+z <- qnorm(1-(1-ci)/2)
+moe <- 0.19
+n <- ((z * sigma)/moe)**2
+round(n, 0)
+```
 
@@
 
+**35/2089 or 0.170**
+
+$$\frac{2089-1734}{2089} = \frac{355}{2089} \approx 0.170 $$
+
+```{r}
+N <- 2089
+n <- 1734
+p <- (N-n)/N
+round(p, 3)
+```
+
 #### Step 2
@@
 
-Lower Bound:
+Lower Bound: **0.153**
 
-Upper Bound:
+Upper Bound: **0.187**
+
+```{r}
+ci <- 0.98
+m <- N * p
+sdm <- sqrt(N*p*(1-p))
+
+df <- (N-n) - 1
+t <- qt(ci, df)
+round((m-t*sdm)/N, 3)
+round((m+t*sdm)/N, 3)
+```
 
@@
 
+**156/474 or 0.329**
+
+```{r}
+N <- 474
+n <- 156
+p <- n/N
+round(p, 3)
+```
+
 #### Step 2
@@
 
-Lower Bound:
+Lower Bound: **0.294**
 
-Upper Bound:
+Upper Bound: **0.365**
+
+```{r}
+ci <- .95
+m <- N * p
+sdm <- sqrt(N*p*(1-p))
+
+df <- (N-n) - 1
+t <- qt(ci, df)
+round((m-t*sdm)/N, 3)
+round((m+t*sdm)/N, 3)
+```

### 2018-07-27 data-science-math/Homework 3.Rmd
More work

@@
 
-Lower Bound: **0.153**
+Lower Bound: **0.151**
 
-Upper Bound: **0.187**
+Upper Bound: **0.189**
 
@@
 m <- N * p
-sdm <- sqrt(N*p*(1-p))
-
+sdm <- sqrt((p*(1-p))/n)
 df <- (N-n) - 1
 t <- qt(ci, df)
-round((m-t*sdm)/N, 3)
-round((m+t*sdm)/N, 3)
+round(p-t*sdm, 3)
+round(p+t*sdm, 3)
 ```
@@
 
-Lower Bound: **0.294**
+Lower Bound: **0.267**
 
-Upper Bound: **0.365**
+Upper Bound: **0.391**
 

### 2018-08-24 DATA606/Lab0/Michael-intro_to_r.Rmd
Initial commit of Lab0
Commiting Lab 0's files

---
title: "Introduction to R and RStudio"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

The goal of this lab is to introduce you to R and RStudio, which you'll be using
throughout the course both to learn the statistical concepts discussed in the 
texbook and also to analyze real data and come to informed conclusions.  To 
straighten out which is which: R is the name of the programming language itself 
and RStudio is a convenient interface.

As the labs progress, you are encouraged to explore beyond what the labs dictate;
a willingness to experiment will make you a much better programmer.  Before we 
get to that stage, however, you need to build some basic fluency in R.  Today we
begin with the fundamental building blocks of R and RStudio: the interface, 
reading in data, and basic commands.

![rinterface](more/rInterface2014.png)

The panel in the upper right contains your *workspace* as well as a history of 
the commands that you've previously entered.  Any plots that you generate will 
show up in the panel in the lower right corner.

The panel on the left is where the action happens.  It's called the *console*. 
Everytime you launch RStudio, it will have the same text at the top of the 
console telling you the version of R that you're running.  Below that information
is the *prompt*.  As its name suggests, this prompt is really a request, a 
request for a command.  Initially, interacting with R is all about typing commands
and interpreting the output. These commands and their syntax have evolved over
decades (literally) and now provide what many users feel is a fairly natural way
to access data and organize, describe, and invoke statistical computations.

To get you started, enter the following command at the R prompt (i.e. right after
`>` on the console).  You can either type it in manually or copy and paste it
from this document.

```{r load-abrbuthnot-data, eval=TRUE}
source("more/arbuthnot.R")
```

This command instructs R to access the OpenIntro website and fetch some data: 
the Arbuthnot baptism counts for boys and girls. You should see that the 
workspace area in the upper righthand corner of the RStudio window now lists a 
data set called `arbuthnot` that has 82 observations on 3 variables. As you 
interact with R, you will create a series of objects. Sometimes you load them as
we have done here, and sometimes you create them yourself as the byproduct of a 
computation or some analysis you have performed. Note that because you are 
accessing data from the web, this command (and the entire assignment) will work 
in a computer lab, in the library, or in your dorm room; anywhere you have 
access to the Internet.

## The Data: Dr. Arbuthnot's Baptism Records

The Arbuthnot data set refers to Dr. John Arbuthnot, an 18<sup>th</sup> century 
physician, writer, and mathematician.  He was interested in the ratio of newborn
boys to newborn girls, so he gathered the baptism records for children born in
London for every year from 1629 to 1710.  We can take a look at the data by 
typing its name into the console.

```{r view-data, eval=TRUE}
arbuthnot
```

What you should see are four columns of numbers, each row representing a 
different year: the first entry in each row is simply the row number (an index 
we can use to access the data from individual years if we want), the second is 
the year, and the third and fourth are the numbers of boys and girls baptized 
that year, respectively. Use the scrollbar on the right side of the console 
window to examine the complete data set.

Note that the row numbers in the first column are not part of Arbuthnot's data. 
R adds them as part of its printout to help you make visual comparisons. You can
think of them as the index that you see on the left side of a spreadsheet. In 
fact, the comparison to a spreadsheet will generally be helpful. R has stored 
Arbuthnot's data in a kind of spreadsheet or table called a *data frame*.

You can see the dimensions of this data frame by typing:

```{r dim-data, eval=TRUE}
dim(arbuthnot)
```

This command should output `[1] 82 3`, indicating that there are 82 rows and 3 
columns (we'll get to what the `[1]` means in a bit), just as it says next to 
the object in your workspace. You can see the names of these columns (or 
variables) by typing:

```{r names-data, eval=TRUE}
names(arbuthnot)
```

You should see that the data frame contains the columns `year`,  `boys`, and 
`girls`. At this point, you might notice that many of the commands in R look a 
lot like functions from math class; that is, invoking R commands means supplying
a function with some number of arguments. The `dim` and `names` commands, for 
example, each took a single argument, the name of a data frame.

One advantage of RStudio is that it comes with a built-in data viewer. Click on
the name `arbuthnot` in the *Environment* pane (upper right window) that lists 
the objects in your workspace. This will bring up an alternative display of the 
data set in the *Data Viewer* (upper left window). You can close the data viewer
by clicking on the *x* in the upper lefthand corner.

## Some Exploration

Let's start to examine the data a little more closely. We can access the data in
a single column of a data frame separately using a command like

```{r view-boys, eval=TRUE}
arbuthnot$boys
```

This command will only show the number of boys baptized each year.

1.  What command would you use to extract just the counts of girls baptized? Try
    it!

```{r try-it-1}
arbuthnot$girls
```
Notice that the way R has printed these data is different. When we looked at the
complete data frame, we saw 82 rows, one on each line of the display. These data
are no longer structured in a table with other variables, so they are displayed 
one right after another. Objects that print out in this way are called *vectors*;
they represent a set of numbers. R has added numbers in [brackets] along the 
left side of the printout to indicate locations within the vector. For example,
`5218` follows `[1]`, indicating that `5218` is the first entry in the vector. 
And if `[43]` starts a line, then that would mean the first number on that line
would represent the 43<sup>rd</sup> entry in the vector.

R has some powerful functions for making graphics. We can create a simple plot 
of the number of girls baptized per year with the command

```{r plot-girls-vs-year, eval=TRUE}
plot(x = arbuthnot$year, y = arbuthnot$girls)
```

By default, R creates a scatterplot with each x,y pair indicated by an open 
circle. The plot itself should appear under the *Plots* tab of the lower right 
panel of RStudio. Notice that the command above again looks like a function, 
this time with two arguments separated by a comma.  The first argument in the 
plot function specifies the variable for the x-axis and the second for the y-axis.
If we wanted to connect the data points with lines, we could add a third argument,
the letter `l` for **l**ine.

```{r plot-girls-vs-year-line, eval=TRUE}
plot(x = arbuthnot$year, y = arbuthnot$girls, type = "l")
```

You might wonder how you are supposed to know that it was possible to add that 
third argument.  Thankfully, R documents all of its functions extensively. To 
read what a function does and learn the arguments that are available to you, 
just type in a question mark followed by the name of the function that you're 
interested in. Try the following.

```{r plot-help, eval=TRUE, tidy = FALSE}
?plot
```

Notice that the help file replaces the plot  in the lower right panel. You can 
toggle between plots and help files using the tabs at the top of that panel.

2.  Is there an apparent trend in the number of girls baptized over the years?  
    How would you describe it?

```{r try-it-2}
plot(x = arbuthnot$year, y = arbuthnot$girls, main = 'I would describe the trend as increasing')
abline(lm(arbuthnot$girls ~ arbuthnot$year), col="#357FAA", lty='longdash', lwd=2)
```

Now, suppose we want to plot the total number of baptisms.  To compute this, we 
could use the fact that R is really just a big calculator. We can type in 
mathematical expressions like

```{r calc-total-bapt-numbers, eval=TRUE}
5218 + 4683
```

to see the total number of baptisms in 1629. We could repeat this once for each 
year, but there is a faster way. If we add the vector for baptisms for boys and
girls, R will compute all sums simultaneously.

```{r calc-total-bapt-vars, eval=TRUE}
arbuthnot$boys + arbuthnot$girls
```

What you will see are 82 numbers (in that packed display, because we aren't 
looking at a data frame here), each one representing the sum we're after. Take a
look at a few of them and verify that they are right. Therefore, we can make a 
plot of the total number of baptisms per year with the command

```{r plot-total-vs-year, eval=TRUE}
plot(arbuthnot$year, arbuthnot$boys + arbuthnot$girls, type = "l")
```

This time, note that we left out the names of the first two arguments.  We can 
do this because the help file shows that the default for `plot` is for the first
argument to be the x-variable and the second argument to be the y-variable.

Similarly to how we computed the proportion of boys, we can compute the ratio of
the number of boys to the number of girls baptized in 1629 with

```{r calc-prop-boys-to-girls-numbers, eval=TRUE}
5218 / 4683
```

or we can act on the complete vectors with the expression

```{r calc-prop-boys-to-girls-vars, eval=TRUE}
arbuthnot$boys / arbuthnot$girls
```

The proportion of newborns that are boys

```{r calc-prop-boys-numbers, eval=TRUE}
5218 / (5218 + 4683)
```

or this may also be computed for all years simultaneously:

```{r calc-prop-boys-vars, eval=TRUE}
arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls)
```

Note that with R as with your calculator, you need to be conscious of the order 
of operations.  Here, we want to divide the number of boys by the total number 
of newborns, so we have to use parentheses.  Without them, R will first do the 
division, then the addition, giving you something that is not a proportion.

3.  Now, make a plot of the proportion of boys over time. What do you see? Tip: 
If you use the up and down arrow keys, you can scroll through your previous 
commands, your so-called command history. You can also access it by clicking on 
the history tab in the upper right panel.  This will save you a lot of typing in
the future.
```{r try-it-3}
share_boys <- arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls)
plot(x = arbuthnot$year, y = share_boys, main = 'More boys than girls (share > 50%)')
```

Finally, in addition to simple mathematical operators like subtraction and 
division, you can ask R to make comparisons like greater than, `>`, less than,
`<`, and equality, `==`. For example, we can ask if boys outnumber girls in each 
year with the expression

```{r boys-more-than-girls, eval=TRUE}
arbuthnot$boys > arbuthnot$girls
```

This command returns 82 values of either `TRUE` if that year had more boys than 
girls, or `FALSE` if that year did not (the answer may surprise you). This 
output shows a different kind of data than we have considered so far. In the 
`arbuthnot` data frame our values are numerical (the year, the number of boys 
and girls). Here, we've asked R to create *logical* data, data where the values
are either `TRUE` or `FALSE`. In general, data analysis will involve many 
different kinds of data types, and one reason for using R is that it is able to
represent and compute with many of them.

This seems like a fair bit for your first lab, so let's stop here. To exit 
RStudio you can click the *x* in the upper right corner of the whole window.  
You will be prompted to save your workspace. If you click *save*, RStudio will 
save the history of your commands and all the objects in your workspace so that 
the next time you launch RStudio, you will see `arbuthnot` and you will have 
access to the commands you typed in your previous session. For now, click *save*,
then start up RStudio again.

* * *

## On Your Own

In the previous few pages, you recreated some of the displays and preliminary 
analysis of Arbuthnot's baptism data. Your assignment involves repeating these 
steps, but for present day birth records in the United States.  Load up the 
present day data with the following command.

```{r load-present-data, eval=TRUE}
source("more/present.R")
```

The data are stored in a data frame called `present`.

-   What years are included in this data set? What are the dimensions of the 
    data frame and what are the variable or column names?

-   How do these counts compare to Arbuthnot's? Are they on a similar scale?

-   Make a plot that displays the boy-to-girl ratio for every year in the data 
    set. What do you see? Does Arbuthnot's observation about boys being born in
    greater proportion than girls hold up in the U.S.? Include the plot in your 
    response.

-   In what year did we see the most total number of births in the U.S.? You can
    refer to the help files or the R reference card 
    [http://cran.r-project.org/doc/contrib/Short-refcard.pdf](http://cran.r-project.org/doc/contrib/Short-refcard.pdf)
    to find helpful commands.

These data come from a report by the Centers for Disease Control 
[http://www.cdc.gov/nchs/data/nvsr/nvsr53/nvsr53_20.pdf](http://www.cdc.gov/nchs/data/nvsr/nvsr53/nvsr53_20.pdf). 
Check it out if you would like to read more about an analysis of sex ratios at 
birth in the United States.

That was a short introduction to R and RStudio, but we will provide you with more
functions and a more complete sense of the language as the course progresses. 
Feel free to browse around the websites for [R](http://www.r-project.org) and 
[RStudio](http://rstudio.org) if you're interested in learning more, or find 
more labs for practice at http://openintro.org.

<div id="license">
This is a product of OpenIntro that is released under a 
[Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel
from a lab written by Mark Hansen of UCLA Statistics.
</div>

### 2018-08-27 DATA607/Week 1 Assignment/Week_1_Assignment.Rmd
Initial commit of DATA607 Week 1 Assignment

---
title: "CUNY SPS DATA-607 Week 1 Assignment"
subtitle: "Loading Data into a Data Frame"
author: "Mike Silva"
date: "August 27, 2018"
output: html_document
---

## Description

The assignment is to load the Mushroom Dataset from the [UCI repository](https://archive.ics.uci.edu/ml/datasets/Mushroom) into a data frame, rename the columns into something meaningful, replace the abbreviations used in the data, and subset the columns of the data frame.  It must include the column that indicated if the mushroom is edible or poisonous and three or four other columns.

## Step 1: Acquire the Data

The first step will be to download the data to the local environment:

```{r acquire-data}
download.file('https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data', 'agaricus-lepiota.data')
```

## Step 2: Load the Data into Data Frame

Now that we have the dataset locally we will create a data frame:

```{r load-data}
df <- read.table('agaricus-lepiota.data', sep=',')
```

According to the documentation we should have 8,124 rows and 22 columns of data.

```{r data-size}
dim(df)
```

## Step 3: Rename the columns

Our data checks out, but the names are not very useful.

```{r bad-names}
names(df)
```

I will rename the columns by using the dplyr package:

```{r rename-columns, warning=FALSE}
library(dplyr)

df <- df %>%
  r
```

### 2018-08-28 DATA607/Week 1 Assignment/Week_1_Assignment.Rmd
More work to Week 1 Assignment

@@
 date: "August 27, 2018"
-output: html_document
+output:
+  html_document:
+    keep_md: yes
 ---
@@
 ```{r load-data}
-df <- read.table('agaricus-lepiota.data', sep=',')
+df <- read.table('agaricus-lepiota.data', sep=',', stringsAsFactors=FALSE)
+head(df)
 ```
 
-According to the documentation we should have 8,124 rows and 22 columns of data.
+According to the documentation we should have 8,124 rows and 23 columns of data.
 
@@
 df <- df %>%
-  r
+  rename(type = V1,
+         cap_shape = V2,
+         cap_surface = V3,
+         cap_color = V4,
+         bruises = V5,
+         odor = V6,
+         gill_attachment = V7,
+         gill_spacing = V8,
+         gill_size = V9,
+         gill_color = V10,
+         stalk_shape = V11,
+         stalk_root = V12,
+         stalk_surface_above_ring = V13,
+         stalk_surface_below_ring = V14,
+         stalk_color_above_ring = V15,
+         stalk_color_below_ring = V16,
+         veil_type = V17,
+         veil_color = V18,
+         ring_number = V19,
+         ring_type = V20,
+         spore_print_color = V21,
+         population = V22,
+         habitat = V23)
 ```
+
+## Step 4: Subsetting Data Frame
+
+Before changing the data to be more meaningful I want to subset the data frame.
+I would like to use this data to predict if a mushroom is poisonous or not.
+According to the documentation the odor, spore print color, stalk surface below ring, and stalk color above ring are 4 attributes that can predict if a mushroom is poisonous with a 99.9% accuracy for these 23 species.
+
+```{r subset-data}
+df <- df %>%
+  select(type, odor, spore_print_color, stalk_color_below_ring, stalk_color_above_ring)
+```
+
+Now that we have these subsets, let's examine the values:
+
+```{r subset-tables}
+lapply(df, function(x) table(x)) 
+```
+
+## Step 5: Recoding Variables
+
+Now that we have the subset of columns we can recode the variables.  I will accomplish this with dplyr and recode statements.
+
+```{r recode-variables}
+df <- df %>%
+  mutate(type = recode(type, p="Poisonous", e="Edible"),
+         odor = recode(odor, a="Almond", c="Creosote", f="Foul", l="Anise", m="Musty", n="None", p="Pungent", s="Spicy", y="Fishy"),
+         spore_print_color = recode(spore_print_color, b="Buff", h="Chocolate", k="Black", n="Brown", o="Orange", r="Green", u="Purple", w="White", y="Yellow"),
+         stalk_color_below_ring = recode(stalk_color_below_ring, b="Buff", c="Cinnamon", e="Red", g="Gray", n="Brown", o="Orange", p="Pink", w="White", y="Yellow"),
+         stalk_color_above_ring = recode(stalk_color_above_ring, b="Buff", c="Cinnamon", e="Red", g="Gray", n="Brown", o="Orange", p="Pink", w="White", y="Yellow"))
+```
+
+Now to examine the final product:
+
+```{r final-product}
+head(df)
+tail(df)
+```

### 2018-08-28 DATA606/Lab1/michael-intro_to_data.Rmd
Initial commit of DATA 606 Lab1

---
title: "Introduction to data"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    css: ./lab.css
---

Some define Statistics as the field that focuses on turning information into
knowledge.  The first step in that process is to summarize and describe the raw
information - the data.  In this lab, you will gain insight into public health
by generating simple graphical and numerical summaries of a data set collected 
by the Centers for Disease Control and Prevention (CDC).  As this is a large 
data set, along the way you'll also learn the indispensable skills of data 
processing and subsetting.

## Getting started

The Behavioral Risk Factor Surveillance System (BRFSS) is an annual telephone 
survey of 350,000 people in the United States. As its name implies, the BRFSS 
is designed to identify risk factors in the adult population and report 
emerging health trends. For example, respondents are asked about their diet and 
weekly physical activity, their HIV/AIDS status, possible tobacco use, and even
their level of healthcare coverage. The BRFSS Web site 
([http://www.cdc.gov/brfss](http://www.cdc.gov/brfss)) contains a complete 
description of the survey, including the research questions that motivate the 
study and many interesting results derived from the data.

We will focus on a random sample of 20,000 people from the BRFSS survey 
conducted in 2000. While there are over 200  variables in this data set, we will
work with a small subset.

We begin by loading the data set of 20,000 observations into the R workspace. 
After launching RStudio, enter the following command.

```{r load-data, eval=TRUE}
source('~/Documents/CUNY-SPS/DATA606/Lab1/more/cdc.R')
```

The data set `cdc` that shows up in your workspace is a *data matrix*, with each
row representing a *case* and each column representing a *variable*.  R calls 
this data format a *data frame*, which is a term that will be used throughout 
the labs.

To view the names of the variables, type the command

```{r names, eval=TRUE}
names(cdc)
```

This returns the names `genhlth`, `exerany`, `hlthplan`, `smoke100`, `height`, 
`weight`, `wtdesire`, `age`, and `gender`. Each one of these variables 
corresponds to a question that was asked in the survey.  For example, for 
`genhlth`, respondents were asked to evaluate their general health, responding
either excellent, very good, good, fair or poor. The `exerany` variable 
indicates whether the respondent exercised in the past month (1) or did not (0).
Likewise, `hlthplan` indicates whether the respondent had some form of health 
coverage (1) or did not (0). The `smoke100` variable indicates whether the 
respondent had smoked at least 100 cigarettes in her lifetime. The other 
variables record the respondent's `height` in inches, `weight` in pounds as well
as their desired weight, `wtdesire`, `age` in years, and `gender`.

1.  How many cases are there in this data set?  How many variables?  For each 
    variable, identify its data type (e.g. categorical, discrete).

We can have a look at the first few entries (rows) of our data with the command

```{r head, eval=TRUE}
head(cdc)
```

and similarly we can look at the last few by typing

```{r tail, eval=TRUE}
tail(cdc)
```

You could also look at *all* of the data frame at once by typing its name into 
the console, but that might be unwise here.  We know `cdc` has 20,000 rows, so 
viewing the entire data set would mean flooding your screen.  It's better to 
take small peeks at the data with `head`, `tail` or the subsetting techniques 
that you'll learn in a moment.

## Summaries and tables

The BRFSS questionnaire is a massive trove of information.  A good first step in
any analysis is to distill all of that information into a few summary statistics
and graphics.  As a simple example, the function `summary` returns a numerical 
summary: minimum, first quartile, median, mean, second quartile, and maximum. 
For `weight` this is

```{r summary-weight, eval=TRUE}
summary(cdc$weight)
```

R also functions like a very fancy calculator.  If you wanted to compute the 
interquartile range for the respondents' weight, you would look at the output 
from the summary command above and then enter

```{r weight-range-arith, eval=TRUE}
190 - 140
```

R also has built-in functions to compute summary statistics one by one.  For 
instance, to calculate the mean, median, and variance of `weight`, type

```{r weight-mean-var-median, eval=TRUE}
mean(cdc$weight) 
var(cdc$weight)
median(cdc$weight)
```

While it makes sense to describe a quantitative variable like `weight` in terms
of these statistics, what about categorical data?  We would instead consider the
sample frequency or relative frequency distribution.  The function `table` does
this for you by counting the number of times each kind of response was given.
For example, to see the number of people who have smoked 100 cigarettes in their
lifetime, type

```{r table-smoke, eval=TRUE}
table(cdc$smoke100)
```

or instead look at the relative frequency distribution by typing

```{r table-smoke-prop, eval=TRUE}
table(cdc$smoke100)/20000
```

Notice how R automatically divides all entries in the table by 20,000 in the 
command above. This is similar to something we observed in the Introduction to R;
when we multiplied or divided a vector with a number, R applied that action across 
entries in the vectors. As we see above, this also works for tables. Next, we 
make a bar plot of the entries in the table by putting the table inside the 
`barplot` command.

```{r table-smoke-barplot, eval=TRUE}
barplot(table(cdc$smoke100))
```

Notice what we've done here! We've computed the table of `cdc$smoke100` and then
immediately applied the graphical function, `barplot`. This is an important 
idea: R commands can be nested. You could also break this into two steps by 
typing the following:

```{r table-smoke-barplot-twosteps, eval=TRUE}
smoke <- table(cdc$smoke100)

barplot(smoke)
```

Here, we've made a new object, a table, called `smoke` (the contents of which we
can see by typing `smoke` into the console) and then used it in as the input for
`barplot`. The special symbol `<-` performs an *assignment*, taking the output 
of one line of code and saving it into an object in your workspace.   This is 
another important idea that we'll return to later.

2.  Create a numerical summary for `height` and `age`, and compute the 
    interquartile range for each. Compute the relative frequency distribution for
    `gender` and `exerany`. How many males are in the sample? What proportion of
    the sample reports being in excellent health?

The `table` command can be used to tabulate any number of variables that you 
provide.  For example, to examine which participants have smoked across each 
gender, we could use the following.

```{r table-smoke-gender, eval=TRUE}
table(cdc$gender,cdc$smoke100)
```

Here, we see column labels of 0 and 1. Recall that 1 indicates a respondent has
smoked at least 100 cigarettes. The rows refer to gender. To create a mosaic 
plot of this table, we would enter the following command.

```{r mosaic-smoke-gender, eval=TRUE}
mosaicplot(table(cdc$gender,cdc$smoke100))
```

We could have accomplished this in two steps by saving the table in one line and
applying `mosaicplot` in the next (see the table/barplot example above).

3.  What does the mosaic plot reveal about smoking habits and gender?

## Interlude: How R thinks about data

We mentioned that R stores data in data frames, which you might think of as a 
type of spreadsheet. Each row is a different observation (a different respondent)
and each column is a different variable (the first is `genhlth`, the second 
`exerany` and so on). We can see the size of the data frame next to the object 
name in the workspace or we can type

```{r dim, eval=TRUE}
dim(cdc)
```

which will return the number of rows and columns. Now, if we want to access a 
subset of the full data frame, we can use row-and-column notation. For example,
to see the sixth variable of the 567<sup>th</sup> respondent, use the format

```{r cdc-row567-column6, eval=TRUE}
cdc[567,6]
```

which means we want the element of our data set that is in the 567<sup>th</sup> 
row (meaning the 567<sup>th</sup> person or observation) and the 6<sup>th</sup> 
column (in this case, weight). We know that `weight` is the 6<sup>th</sup> variable
because it is the 6<sup>th</sup> entry in the list of variable names

```{r names-again, eval=TRUE}
names(cdc)
```

To see the weights for the first 10 respondents we can type

```{r first-10-rows-sixth-column, eval=TRUE}
cdc[1:10,6]
```

In this expression, we have asked just for rows in the range 1 through 10.  R 
uses the `:` to create a range of values, so 1:10 expands to 1, 2, 3, 4, 5, 6, 
7, 8, 9, 10. You can see this by entering

```{r numbers-1to10, eval=TRUE}
1:10
```

Finally, if we want all of the data for the first 10 respondents, type

```{r first-10-rows, eval=TRUE}
cdc[1:10,]
```

By leaving out an index or a range (we didn't type anything between the comma 
and the square bracket), we get all the columns. When starting out in R, this is
a bit counterintuitive. As a rule, we omit the column number to see all columns 
in a data frame. Similarly, if we leave out an index or range for the rows, we 
would access all the observations, not just the 567<sup>th</sup>, or rows 1 
through 10. Try the following to see the weights for all 20,000 respondents fly 
by on your screen

```{r 6th-column, eval=TRUE}
cdc[,6]
```

Recall that column 6 represents respondents' weight, so the command above 
reported all of the weights in the data set. An alternative method to access the
weight data is by referring to the name.  Previously, we typed `names(cdc)` to 
see all the variables contained in the cdc data set. We can use any of the 
variable names to select items in our data set.

```{r weight, eval=TRUE}
cdc$weight
```

The dollar-sign tells R to look in data frame `cdc` for the column called 
`weight`.  Since that's a single vector, we can subset it with just a single 
index inside square brackets.  We see the weight for the 567<sup>th</sup> 
respondent by typing

```{r weight-567, eval=TRUE}
cdc$weight[567]
```

Similarly, for just the first 10 respondents

```{r weight-first10, eval=TRUE}
cdc$weight[1:10]
```

The command above returns the same result as the `cdc[1:10,6]` command. Both 
row-and-column notation and dollar-sign notation are widely used, which one you 
choose to use depends on your personal preference.

## A little more on subsetting

It's often useful to extract all individuals (cases) in a data set that have 
specific characteristics. We accomplish this through *conditioning* commands. 
First, consider expressions like

```{r true-male, eval=TRUE}
cdc$gender == "m"
```

or

```{r true-over30, eval=TRUE}
cdc$age > 30
```

These commands produce a series of `TRUE` and `FALSE`  values. There is one 
value for each respondent, where `TRUE` indicates that the person was male (via 
the first command) or older than 30 (second command).

Suppose we want to extract just the data for the men in the sample, or just for 
those over 30. We can use the R function `subset` to do that for us. For example,
the command

```{r males, eval=TRUE}
mdata <- subset(cdc, cdc$gender == "m")
```

will create a new data set called `mdata` that contains only the men from the 
`cdc` data set. In addition to finding it in your workspace alongside its 
dimensions, you can take a peek at the first several rows as usual

```{r head-males, eval=TRUE}
head(mdata)
```

This new data set contains all the same variables but just under half the rows.
It is also possible to tell R to keep only specific variables, which is a topic
we'll discuss in a future lab. For now, the important thing is that we can carve
up the data based on values of one or more variables.

As an aside, you can use several of these conditions together with `&` and `|`.
The `&` is read "and" so that

```{r males-and-over30, eval=TRUE}
m_and_over30 <- subset(cdc, gender == "m" & age > 30)
```

will give you the data for men over the age of 30. The `|` character is read 
"or" so that

```{r males-or-over30, eval=TRUE}
m_or_over30 <- subset(cdc, gender == "m" | age > 30)
```

will take people who are men or over the age of 30 (why that's an interesting 
group is hard to say, but right now the mechanics of this are the important 
thing). In principle, you may use as many "and" and "or" clauses as you like 
when forming a subset.

3.  Create a new object called `under23_and_smoke` that contains all observations 
    of respondents under the age of 23 that have smoked 100 cigarettes in their 
    lifetime. Write the command you used to create the new object as the answer
    to this exercise.

## Quantitative data

With our subsetting tools in hand, we'll now return to the task of the day: 
making basic summaries of the BRFSS questionnaire.  We've already looked at 
categorical data such as `smoke` and `gender` so now let's turn our attention to
quantitative data.  Two common ways to visualize quantitative data are with box
plots and histograms.  We can construct a box plot for a single variable with 
the following command.

```{r boxplot-height, eval=TRUE}
boxplot(cdc$height)
```

You can compare the locations of the components of the box by examining the 
summary statistics.

```{r summary-height, eval=TRUE}
summary(cdc$height)
```

Confirm that the median and upper and lower quartiles reported in the numerical 
summary match those in the graph. The purpose of a boxplot is to provide a 
thumbnail sketch of a variable for the purpose of comparing across several 
categories. So we can, for example, compare the heights of men and women with

```{r boxplot-height-gender, eval=TRUE}
boxplot(cdc$height ~ cdc$gender)
```

The notation here is new. The `~` character can be read *versus* or 
*as a function of*. So we're asking R to give us a box plots of heights where 
the groups are defined by gender.

Next let's consider a new variable that doesn't show up directly in this data 
set: Body Mass Index (BMI) 
([http://en.wikipedia.org/wiki/Body_mass_index](http://en.wikipedia.org/wiki/Body_mass_index)). 
BMI is a weight to height ratio and can be calculated as:

\[ BMI = \frac{weight~(lb)}{height~(in)^2} * 703 \]

703 is the approximate conversion factor to change units from metric (meters and 
kilograms) to imperial (inches and pounds).

The following two lines first make a new object called `bmi` and then creates 
box plots of these values, defining groups by the variable `cdc$genhlth`.

```{r boxplot-bmi, eval=TRUE}
bmi <- (cdc$weight / cdc$height^2) * 703
boxplot(bmi ~ cdc$genhlth)
```

Notice that the first line above is just some arithmetic, but it's applied to 
all 20,000 numbers in the `cdc` data set. That is, for each of the 20,000 
participants, we take their weight, divide by their height-squared and then 
multiply by 703. The result is 20,000 BMI values, one for each respondent. This 
is one reason why we like R: it lets us perform computations like this using 
very simple expressions.

4.  What does this box plot show? Pick another categorical variable from the 
    data set and see how it relates to BMI. List the variable you chose, why you
    might think it would have a relationship to BMI,  and indicate what the 
    figure seems to suggest.

Finally, let's make some histograms. We can look at the histogram for the age of
our respondents with the command

```{r hist-age, eval=TRUE}
hist(cdc$age)
```

Histograms are generally a very good way to see the shape of a single 
distribution, but that shape can change depending on how the data is split 
between the different bins.  You can control the number of bins by adding an 
argument to the command. In the next two lines, we first make a default 
histogram of `bmi` and then one with 50 breaks.

```{r hist-bmi, eval=TRUE}
hist(bmi)
hist(bmi, breaks = 50)
```

Note that you can flip between plots that you've created by clicking the forward
and backward arrows in the lower right region of RStudio, just above the plots. 
How do these two histograms compare?

At this point, we've done a good first pass at analyzing the information in the 
BRFSS questionnaire.  We've found an interesting association between smoking and
gender, and we can say something about the relationship between people's 
assessment of their general health and their own BMI.  We've also picked up 
essential computing tools -- summary statistics, subsetting, and plots -- that 
will serve us well throughout this course.

* * *

## On Your Own

-   Make a scatterplot of weight versus desired weight. Describe the 
    relationship between these two variables.

-   Let's consider a new variable: the difference between desired weight 
    (`wtdesire`) and current weight (`weight`). Create this new variable by 
    subtracting the two columns in the data frame and assigning them to a new 
    object called `wdiff`.

-   What type of data is `wdiff`? If an observation `wdiff` is 0, what does 
    this mean about the person's weight and desired weight. What if `wdiff` is 
    positive or negative?

-   Describe the distribution of `wdiff` in terms of its center, shape, and 
    spread, including any plots you use. What does this tell us about how people 
    feel about their current weight?

-   Using numerical summaries and a side-by-side box plot, determine if men tend
    to view their weight differently than women.

-   Now it's time to get creative. Find the mean and standard deviation of 
    `weight` and determine what proportion of the weights are within one 
    standard deviation of the mean.

<div id="license">
This is a product of OpenIntro that is released under a 
[Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel 
from a lab written by Mark Hansen of UCLA Statistics.
</div>

### 2018-08-28 DATA607/Week 1 Assignment/Week_1_Assignment.Rmd
More work on DATA 607 Week 1 Assignment

@@
 date: "August 27, 2018"
-output:
-  html_document:
-    keep_md: yes
 ---
@@
 
-```{r rename-columns, warning=FALSE}
+```{r rename-columns, warning=FALSE, message=FALSE}
 library(dplyr)

### 2018-08-29 DATA606/Lab0/Michael-intro_to_r.Rmd
Additional work

@@
     theme: cerulean
+    keep_md: yes
   pdf_document: default
@@
 
-![rinterface](more/rInterface2014.png)
+![ ](more/rInterface2014.png)
 
@@
 ```{r load-abrbuthnot-data, eval=TRUE}
-source("more/arbuthnot.R")
+source("~/Documents/CUNY-SPS/DATA606/Lab0/more/arbuthnot.r")
 ```
@@
 ```
+
 Notice that the way R has printed these data is different. When we looked at the
@@
 the future.
+
 ```{r try-it-3}
@@
 ```{r load-present-data, eval=TRUE}
-source("more/present.R")
+source("~/Documents/CUNY-SPS/DATA606/Lab0/more/present.R")
 ```
@@
 
+```{r}
+unique(present$year)
+```
+
+**1940 to 2002**
+
+```{r}
+dim(present)
+```
+
+**63 rows and 3 columns**
+
+```{r}
+names(present)
+```
+
 -   How do these counts compare to Arbuthnot's? Are they on a similar scale?
+
+```{r warning=FALSE}
+suppressMessages(library(dplyr))
+
+combined <- rbind(arbuthnot, present) %>%
+  mutate(total = boys + girls) %>%
+  mutate(Dataset = ifelse(year > 1939, 'Present', 'Arbuthnot'))
+
+combined %>%
+  group_by(Dataset) %>%
+  summarise(`Average records per year` = mean(total))
+```
+
+**Present dataset has significantly more records per year than Arbuthnot's**
 
@@
 
+```{r}
+library(ggplot2)
+combined %>%
+  mutate(`Boy-to-Girl Ratio` = boys / girls) %>%
+  ggplot(., aes(x = Dataset, y = `Boy-to-Girl Ratio`, fill = Dataset)) + 
+  geom_boxplot(show.legend = FALSE) 
+```
+
+**Arbuthnot's observations does not hold in the Present dataset**
+
 -   In what year did we see the most total number of births in the U.S.? You can
@@
     to find helpful commands.
+
+```{r}
+combined[combined$year == max(combined$year),]$year
+```
 

### 2018-08-30 DATA606/Lab1/michael-intro_to_data.Rmd
Work on DATA 606 Lab 1

@@
     variable, identify its data type (e.g. categorical, discrete).
+
+```{r}
+dim(cdc)
+
+summary(cdc)
+```
+
+**There are 20K records with 9 variables.
+height weight wtdesire and age are numeric and continuous.
+gender and genhlth are categorical.
+exerany, hlthplan and smoke100 are categorical (0=no and 1=yes).**
 
@@
 
+```{r}
+summary(cdc$height)
+```
+
+```{r}
+IQR(cdc$height)
+```
+
+```{r}
+summary(cdc$age)
+```
+
+```{r}
+IQR(cdc$age)
+```
+
+```{r}
+table(cdc$gender) / nrow(cdc)
+```
+
+```{r}
+table(cdc$exerany) / nrow(cdc)
+```
+
+```{r}
+nrow(cdc[cdc$gender=="m",])
+```
+
+```{r}
+nrow(cdc[cdc$genhlth == "excellent",]) / nrow(cdc)
+```
+
 The `table` command can be used to tabulate any number of variables that you 
@@
 3.  What does the mosaic plot reveal about smoking habits and gender?
+
+**Males are more likely to smoke than females.**
 
@@
 
+```{r}
+under23_and_smoke <- subset(cdc, smoke100 == "1" & age < 23)
+```
+
 ## Quantitative data
@@
 
+**The box plot shows the distribution of the data.
+It shows the IQR, median, and outliers.
+The above box plot shows that the individuals with poorer general health are more likely to have a higher BMI than those with better health.**
+
+```{r}
+boxplot(bmi ~ cdc$gender)
+```
+
+**I picked gender because general health status was already taken.
+I would expect a difference based on the gender because men and women are different.
+The plot shows a greater variation among the women than the men.**
+
 Finally, let's make some histograms. We can look at the histogram for the age of
@@
 
+```{r}
+plot(x = cdc$weight, y = cdc$wtdesire)
+abline(lm(cdc$weight ~ cdc$wtdesire), col="#357FAA", lty='longdash', lwd=2)
+```
+
+**The relationship is positive.**
+
 -   Let's consider a new variable: the difference between desired weight 
@@
 
+```{r}
+wdiff <- cdc$wtdesire - cdc$weight
+```
+
 -   What type of data is `wdiff`? If an observation `wdiff` is 0, what does 
@@
     positive or negative?
+
+**It is a continous variable.
+If wdiff is equal to 0 it means that their desired weight is their current weight.
+If wdiff is positive they desire to put on weight.
+If it is negative they desire to weight less.**
 
@@
 
+```{r}
+library(e1071)
+summary(wdiff)
+skewness(wdiff)
+```
+
+```{r}
+hist(wdiff)
+```
+
+**The data is left skewed and clustered around -10 to -15.  There is a few extreme values but in generall people want to loose a few pounds.**
+
 -   Using numerical summaries and a side-by-side box plot, determine if men tend
     to view their weight differently than women.
+
+```{r}
+
+```
 

### 2018-08-31 DATA606/Intro to Data.Rmd
Added Intro to Data.Rmd

---
title: "Intro to Data"
author: "Mike Silva"
date: "August 24, 2018"
output: html_document
---

## 1.8 Smoking habits of UK residents

A survey was conducted to study the smoking habits of UK residents. Below is a data matrix displaying a portion of the data collected in this survey. Note that “£” stands for British Pounds Sterling, “cig” stands for cigarettes, and “N/A” refers to a missing component of the data.

|   |   sex  | age | marital | grossIncome | smoke | amtWeekends | amtWeekdays |
|---| -----: | --: | -------:| -----------: | ----: | ----------: | ----------: |
| 1 | Female | 42  | Single  | Under £2,600 | Yes | 12 cig/day | 12 cig/day | 
| 2 | Male	 | 44	 | Single  | £10,400 to £15,600 | No | N/A | N/A | 
| 3 | Male 	 | 53  | Married | Above £36,400 | Yes | 6 cig/day | 6 cig/day |
| ... | ... | ... | ... | ... | ...| ... | ... |
| 1691 | Male | 40 | Single | £2,600 to £5,200 | Yes | 8 cig/day | 8 cig/day |

(a) What does each row of the data matrix represent?

**A survey response**

(b) How many participants were included in the survey?

**1,691**

(c) Indicate whether each variable in the study is numerical or categorical. If numerical, identify as continuous or discrete. If categorical, indicate if the variable is ordinal.

* __sex - non-ordinal categorical__
* __age - discrete numerical__
* __marital - non-ordinal categorical__
* __grossIncome - ordinal categorical__
* __smoke - non-ordinal categorical__
* __amtWeekends - ordinal categorical__
* __amtWeekdays - ordinal categorical__

## 1.10 Cheaters, scope of inference

Exercise 1.5 introduces a study where researchers studying the relationship between honesty, age, and self-control conducted an experiment on 160 children between the ages of 5 and 15. The researchers asked each child to toss a fair coin in private and to record the outcome (white or black) on a paper sheet, and said they would only reward children who report white. Half the students were explicitly told not to cheat and the others were not given any explicit instructions. Differences were observed in the cheating rates in the instruction and no instruction groups, as well as some differences across children’s characteristics within each group.

(a) Identify the population of interest and the sample in this study.

**Population of interest: The general population**
**Study sample: Children ages 5 to 15**

(b) Comment on whether or not the results of the study can be generalized to the population, and if the findings of the study can be used to establish causal relationships.

**The results cannot be generalized to the population.  The findings can not be used to estalish a causal relationship**

## 1.28 Reading the paper

Below are excerpts from two articles published in the NY Times:

(a) An article titled Risks: Smokers Found More Prone to Dementia states the following:

“Researchers analyzed data from 23,123 health plan members who participated in a voluntary exam and health behavior survey from 1978 to 1985, when they were 50-60 years old. 23 years later, about 25% of the group had dementia, including 1,136 with Alzheimer’s disease and 416 with vascular dementia. After adjusting for other factors, the researchers concluded that pack-a-day smokers were 37% more likely than nonsmokers to develop dementia, and the risks went up with increased smoking; 44% for one to two packs a day; and twice the risk for more than two packs.”

Based on this study, can we conclude that smoking causes dementia later in life? Explain your reasoning.

**No.  Smoking and dementia are correlated in the study, but correlation is not causation**

(b) Another article titled The School Bully Is Sleepy states the following:

“The University of Michigan study, collected survey data from parents on each child’s sleep habits and asked both parents and teachers to assess behavioral concerns. About a third of the students studied were identified by parents or teachers as having problems with disruptive behavior or bullying. The researchers found that children who had behavioral issues and those who were identified as bullies were twice as likely to have shown symptoms of sleep disorders.”

A friend of yours who read the article says, “The study shows that sleep disorders lead to bullying in school children.” Is this statement justified? If not, how best can you describe the conclusion that can be drawn from this study?

**No the statement is not justified. You can conclude that the two are correlated but there may be other causal factors that haven't been articulated by the article.**

## 1.36 Exercise and mental health

A researcher is interested in the effects of exercise on mental health and he proposes the following study: Use stratified random sampling to ensure representative proportions of 18-30, 31-40 and 41-55 year olds from the population. Next, randomly assign half the subjects from each age group to exercise twice a week, and instruct the rest not to exercise. Conduct a mental health exam at the beginning and at the end of the study, and compare the results.

(a) What type of study is this?

**This is a randomized experiment.**

(b) What are the treatment and control groups in this study?

**The treatment group are those who exercise and the control group do not exercise.**

(c) Does this study make use of blocking? If so, what is the blocking variable?

**No it does not.**

(d) Does this study make use of blinding?

**No it does not as the researcher instructs the participant to exercise or not.**

(e) Comment on whether or not the results of the study can be used to establish a causal relationship between exercise and mental health, and indicate whether or not the conclusions can be generalized to the population at large.

**Results from this study may establish a causal relationship between exercise and mental health holding all other variables constant.  The conclusions could be generalize if the sample size is large enough.**

(f) Suppose you are given the task of determining if this proposed study should get funding.  Would you have any reservations about the study proposal?

**I would be concerned about all the other variables that affect mental health besides the amount of exercise.**

## 1.48  Stats scores

Below are the final exam scores of twenty introductory statistics students.
57, 66, 69, 71, 72, 73, 74, 77, 78, 78, 79, 79, 81, 81, 82, 83, 83, 88, 89, 94

Create a box plot of the distribution of these scores. The five number summary provided below may be useful.

| Min |  Q1  | Q2 (Median) |  Q3  | Max |
| --: | ---: | ----------: | ---: | --: |
| 57  | 72.5 |     78.5    | 82.5 | 94  |

```{r 1.48}
scores <- c(57, 66, 69, 71, 72, 73, 74, 77, 78, 78, 79, 79, 81, 81, 82, 83, 83, 88, 89, 94)
boxplot(scores)
```

## 1.50 Mix-and-match

Describe the distribution in the histograms below and match them to the box plots.

## 1.56 Distributions and appropriate statistics, Part II

For each of the following, state whether you expect the distribution to be symmetric, right skewed, or left skewed. Also specify whether the mean or median would best represent a typical observation in the data, and whether the variability of observations would be best represented using the standard deviation or IQR. Explain your reasoning.

(a) Housing prices in a country where 25% of the houses cost below $350,000, 50% of the houses cost below $450,000, 75% of the houses cost below $1,000,000 and there are a meaningful number of houses that cost more than $6,000,000.

(b) Housing prices in a country where 25% of the houses cost below $300,000, 50% of the houses cost below $600,000, 75% of the houses cost below $900,000 and very few houses that cost more than $1,200,000.

(c) Number of alcoholic drinks consumed by college students in a given week. Assume that most of these students don’t drink since they are under 21 years old, and only a few drink excessively.

(d) Annual salaries of the employees at a Fortune 500 company where only a few high level executives earn much higher salaries than all the other employees.

## 1.70 Heart transplants

The Stanford University Heart Transplant Study was conducted to determine whether an experimental heart transplant program increased lifespan. Each patient entering the program was designated an experimental heart transplant candidate, meaning that he was gravely ill and would most likely benefit from a new heart. Some patients got a transplant and some did not. The variable transplant indicates which group the patients were in; patients in the treatment group got a transplant and those in the control group did not. Another variable called survived was used to indicate whether or not the patient was alive at the end of the study. Of the 34 patients in the control group, 30 died. Of the 69 people in the treatment group, 45 died.

(a) Based on the mosaic plot, is survival independent of whether or not the patient got a transplant? Explain your reasoning.

(b) What do the box plots below suggest about the efficaacy (effectiveness) of the heart transplant treatment.

(c) What proportion of patients in the treatment group and what proportion of patients in the control group died?

(d) One approach for investigating whether or not the treatment is e↵ective is to use a randomization technique.

i. What are the claims being tested?

ii. The paragraph below describes the set up for such approach, if we were to do it withut using statistical software. Fill in the blanks with a number or phrase, whichever is appropriate.

We write alive on cards representing patients who were alive at the end of the study, and dead on cards representing patients who were not. Then, we shuffle these cards and split them into two groups: one group of size representing treatment, and another group of size representing control. We calculate the difference between the proportion of dead cards in the treatment and control groups (treatment - control) and record this value. We repeat this 100 times to build a distribution centered at . Lastly, we calculate the fraction of simulations where the simulated differences in proportions are . If this fraction is low, we conclude that it is unlikely to have observed such an outcome by chance and that the null hypothesis should be rejected in favor of the alternative.

iii. What do the simulation results shown below suggest about the effectiveness of the transplant program?

### 2018-08-31 DATA606/Intro to Data.Rmd
Additional work

@@
 
+![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/intro_to_data_1_50.png)
+
+**(a) is a normal distribution.  The corresponding box plot is 2**
+
+**(b) is a uniform distribution.  3 is the box plot**
+
+**(c) is a long-tailed or right skewed distribution and 1 is the box plot**
+
 ## 1.56 Distributions and appropriate statistics, Part II
@@
 
+**This is long tailed right skewed distribution.
+The median would be the best option to represent the central tendancy and IQR would be best for the spread.
+Since "the are a meaningful number" of very expensive houses, the mean would be larger than the median and less representative of the average.**
+
 (b) Housing prices in a country where 25% of the houses cost below $300,000, 50% of the houses cost below $600,000, 75% of the houses cost below $900,000 and very few houses that cost more than $1,200,000.
+
+**This is symetric distribution.  Mean and standard deviation could be used on this distribution because the spread is much more compact than the previous example.**
 
@@
 
+**This is a bimodal distribution.  You can't use median or mean to summarize because it really isn't meaningful.**
+
 (d) Annual salaries of the employees at a Fortune 500 company where only a few high level executives earn much higher salaries than all the other employees.
+
+**This is probably a symetric distribution with some outliers on the right tail.  A median and IQR would be the best choice because of the top level executives would pull the mean to the right.**
 
@@
 
+![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/intro_to_data_1_70.png)
+
 (a) Based on the mosaic plot, is survival independent of whether or not the patient got a transplant? Explain your reasoning.
+
+**Survival is dependent of whether the patient got a transplant.  The probability (proportion) of living when they got the treatment is larger than that when they did not.**
 

### 2018-08-31 DATA606/Intro to Data.Rmd
More work on Intro to Data assignment
Renamed 1_70.png to 1_70_1.png b/c there are
more than one 1_70 pngs

@@
 ---
+
+## Description
+
+The following is the graded problems for the Intro to Data chapter of Open Intro to Statistics.  My answers to the questions are **bolded.**  The headers correspond to a question.
 
@@
 
-![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/intro_to_data_1_70.png)
+![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/intro_to_data_1_70_1.png)
 
@@
 
+**It suggests that the survival time of those receiving the treatment is larger than the control group.**
+
 (c) What proportion of patients in the treatment group and what proportion of patients in the control group died?
+
+**Roughly 88% of the control group and 65% of the treatment group died when the box plot was made.  Eventually 100% of both groups died.**
 
@@
 
-ii. The paragraph below describes the set up for such approach, if we were to do it withut using statistical software. Fill in the blanks with a number or phrase, whichever is appropriate.
+**That the treatment is effective at increasing the duration of life.**
 
-We write alive on cards representing patients who were alive at the end of the study, and dead on cards representing patients who were not.
-Then, we shuffle these cards and split them into two groups: one group of size representing treatment, and another group of size representing control.
+ii. The paragraph below describes the set up for such approach, if we were to do it without using statistical software. Fill in the blanks with a number or phrase, whichever is appropriate.
+
+We write alive on **28** cards representing patients who were alive at the end of the study, and dead on **75** cards representing patients who were not.
+Then, we shuffle these cards and split them into two groups: one group of size **69** representing treatment, and another group of size **34** representing control.
 We calculate the difference between the proportion of dead cards in the treatment and control groups (treatment - control) and record this value.
-We repeat this 100 times to build a distribution centered at .
-Lastly, we calculate the fraction of simulations where the simulated differences in proportions are .
+We repeat this 100 times to build a distribution centered at **zero**.
+Lastly, we calculate the fraction of simulations where the simulated differences in proportions are **less than zero**.
 If this fraction is low, we conclude that it is unlikely to have observed such an outcome by chance and that the null hypothesis should be rejected in favor of the alternative.
@@
 iii. What do the simulation results shown below suggest about the effectiveness of the transplant program?
+
+![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/intro_to_data_1_70_2.png)

### 2018-08-31 DATA606/Intro to Data.Rmd
More work on Intro to Data assignment

@@
 
-![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/intro_to_data_1_50.png)
+![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/intro_to_data_1_50.png)
 
@@
 
-![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/intro_to_data_1_70_1.png)
+![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/intro_to_data_1_70_1.png)
 
@@
 We repeat this 100 times to build a distribution centered at **zero**.
-Lastly, we calculate the fraction of simulations where the simulated differences in proportions are **less than zero**.
+Lastly, we calculate the fraction of simulations where the simulated differences in proportions are **greater than zero**.
 If this fraction is low, we conclude that it is unlikely to have observed such an outcome by chance and that the null hypothesis should be rejected in favor of the alternative.
@@
 
-![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/intro_to_data_1_70_2.png)
+![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/intro_to_data_1_70_2.png)
+
+**It suggests the effectiveness of the treatmant in increasing life is not due to chance.**

### 2018-09-01 DATA606/Lab1/michael-intro_to_data.Rmd
Work on DATA 606 Lab 1

@@
 ```{r}
+summary(wdiff)
+
 library(e1071)
-summary(wdiff)
 skewness(wdiff)
@@
 
-**The data is left skewed and clustered around -10 to -15.  There is a few extreme values but in generall people want to loose a few pounds.**
+**The data is left skewed and clustered around -10 to -15.  There is a few extreme values but in general people want to loose a bit of weight.**
 
@@
 ```{r}
+cdc$wdiff <- cdc$wtdesire - cdc$weight
+summary(cdc[cdc$gender == "m",]$wtdesire)
+summary(cdc[cdc$gender == "f",]$wtdesire)
 
+boxplot(cdc$wtdesire ~ cdc$gender)
 ```
+
+**In order to see how they different genders viewed their weight I decided to look at the weight they desired.
+It would stand as a proxy for the "ideal" man or woman.
+As one would expect the "ideal" male's weight is heavier than the "ideal" female's.**
 
@@
     standard deviation of the mean.
+
+```{r}
+mu <- mean(cdc$weight)
+mu
+sigma <- sd(cdc$weight)
+sigma
+
+upper <- mu + sigma
+lower <- mu - sigma
+
+nrow(cdc[cdc$weight>=lower & cdc$weight<=upper,]) / nrow(cdc)
+```
+
+**The mean is roughly 170 with a standard deviation of about 40.  About 71% are within one standard deviation of the mean.**
 

### 2018-09-07 DATA607/Week_2_Assignment.Rmd
Initial commit of Week 2 Assignment R Markdown

---
title: "Week 2 Assignment"
author: "Mike Silva"
date: "September 1, 2018"
output: html_document
---

## Description

Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.

The films used in this study were selected by Googling "2018 recent films" and getting the first 6 films listed in the search results.

## What's Been Done So Far

The 'Week 2 Assignment.sql' script has been executed which has set up the flix database with the ratings.

## Connect To The Database

Using settings in a hidden file, connect to the database.

```{r connect,message=FALSE}
library(RMySQL)
library(ggplot2)
library(dplyr)
source('hidden/MySQL Settings.R')
conn <- dbConnect(MySQL(), user=user, password=password, dbname='flix', host=host)
```

## Pull Data From The Database

Pull data via the connection.

```{r pull_data}
sql <- 'SELECT `films`.`title`, 
`reviewers`.`name`, 
`ratings`.`rating`
FROM `ratings`
JOIN `films` ON `films`.`id` = `ratings`.`films_id`
JOIN `reviewers` ON `reviewers`.`id` = `ratings`.`reviewers_id`'
df <- dbGetQuery(conn, sql)

# Disconnect from the database
disconnected <- dbDisconnect(conn)
```

## Exploratory Data Analysis

Now that the data is loaded into a dataframe let's take a look at the data.

```{r first_look}
dim(df)
names(df)
```

Now for a preview of the data

```{r preview}
head(df)
```

```{r viz}
ggplot(df, aes(x=title, y=rating, fill=title)) + 
  geom_boxplot() + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

The data is centered arround 3 with the exception of Annihilaton and Deadpool 2 which have a lower rating.

```{r}
df %>%
  group_by(title) %>%
  summarise(min=min(rating),
            mean=mean(rating),
            median=median(rating),
            max=max(rating))
```

### 2018-09-07 DATA607/Hands on Lab 2-3.Rmd
Initial commit

---
title: "Hands on Lab 2-3: Combining Data from Multiple Sources"
author: "Mike Silva"
date: "September 7, 2018"
output: html_document
---

## Description

Your task is to create an R dataframe that shows rates of tuberculosis infection by country. You should use the information from the tb database and the attached population.csv file. Your R dataframe should have the following columns:

Country, Year, Rate

Where Rate is defined as Cases/Population.

## Step 1: Read in the Data

First I will extract the TB cases from the MySQL database.  I will make use of a MySQL Settings file which hold authentication credentials.  It is not included in the [GitHub repo](https://github.com/mikeasilva/CUNY-SPS/tree/master/DATA607) for obvious reasons.

```{r connect,message=FALSE}
library(RMySQL)
library(ggplot2)
library(dplyr)
source('hidden/MySQL Settings.R')
conn <- dbConnect(MySQL(), user=user, password=password, dbname='tb', host=host)
```

### 2018-09-07 DATA607/Hands on Lab 2-3.Rmd
More work on Handon lab 2-3

@@
 
+### Step 1-A: Extract the TB Cases From MySQL Database
+
 First I will extract the TB cases from the MySQL database.
@@
 It is not included in the [GitHub repo](https://github.com/mikeasilva/CUNY-SPS/tree/master/DATA607) for obvious reasons.
+I will start by connecting to the database.
 
@@
 ```
+
+Now that we have a connection I will pull the total cases by country and year from the database.
+
+```{r pull_data}
+sql <- 'SELECT 
+    `country`, `year`, SUM(child + adult + elderly) AS `cases`
+FROM
+    `tb`
+GROUP BY `country`, `year`
+HAVING `cases` IS NOT NULL'
+
+tb <- dbGetQuery(conn, sql)
+
+# Disconnect from the database
+disconnected <- dbDisconnect(conn)
+
+head(tb)
+```
+
+### Step 1-B: Retreive the Population From GitHub
+
+I will now bring in the population data from GitHub.
+
+```{r}
+url <- 'https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/population.csv'
+
+pop <- read.csv(url)
+
+head(pop)
+```
+
+## Step 2: Wrangle the Datasets
+
+The next step is to wrangle the data prior to merging the two datasets together.
+I will be joining on the country name and year.
+Since there may be variations in the capitalization between the two datasets I will be capitalizing all of them.
+
+The data from the MySQL is not encoded in a way that plays nicely with the population data, so an additional step is taken.
+
+```{r wrangle}
+tb <- tb %>%
+  mutate(country = enc2utf8(country)) %>%
+  mutate(country = toupper(country))
+
+df <- pop %>%
+  mutate(country = toupper(country))  %>%
+  merge(tb)
+```
+
+Since there are `nrow(tb)` observations in the TB dataset and `nrow(df)` in the combined dataset I want to see what fell out.
+
+```{r}
+tb_countries <- unique(tb$country)
+df_countires <- unique(df$country)
+missing <- tb_countries[!tb_countries %in% df_countires]
+missing
+```
+
+The Ivory Coast fell out of the dataset.

### 2018-09-08 DATA607/Hands on Lab 2-3.Rmd
More work on Lab 2-3

@@
 
-```{r}
+```{r find_missing_records}
 tb_countries <- unique(tb$country)
@@
 
-The Ivory Coast fell out of the dataset.
+The Ivory Coast fell out of the dataset.  This is due to the encoding issue and is something that I will not be addressing.
+
+## Step 2: Compute the Rate
+
+Now that the two datasets are combined I will calculate the prevalance rate:
+
+```{r calculate_rate}
+df <- df %>%
+  mutate(rate = cases / population)
+```
+
+## Step 3: Exploratory Analysis
+
+First we will look at the newly created rate variable
+
+```{r}
+summary(df$rate)
+```
+
+The scale is not very useful.  I will rescale it to per 100,000.
+
+```{r}
+df <- df %>%
+  mutate(rate_per_100k = rate * 100000)
+```
+
+```{r}
+ggplot(df, aes(x=year, y=rate_per_100k)) +
+  geom_smooth(method='lm',formula=y~x) +
+  ylab("TB Infections Rate per 100,000") +
+  xlab('Year')
+```
+
+The trend show that on average TB infections have been increasing over time from 25 to 90 cases per 100,000 residents.
+But the summary above says there is 813 cases per 100,000 as a maximum.
+So I want to look at a boxplot of the data.
+
+```{r boxplot}
+ggplot(df, aes(x=year, y=rate_per_100k, group=year)) + 
+  geom_boxplot() +
+  ylab("TB Infections Rate per 100,000") +
+  xlab('Year')
+```
+
+The median has been on the rise but there is a lot of outliers.  It seems that there was an outbreak in 2006 where TB rates increased above 400 cases for a few locations.

### 2018-09-11 DATA607/Week_3_Assignment.Rmd
Updated Week_3_Assignment.Rmd

---
title: "Week 3 Assignment"
author: "Mike Silva"
date: "September 9, 2018"
output: html_document
---

## Description

The following problems come from chapter 8, regular expressions and essential string functions from "Automated Data Collection with R"

### Problem 3

```{r}
library(stringr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))

name
```

(a) Use the tools of this chapter to rearrange  the vector so that all elements conform to the standard *first_name* *last_name*.

```{r}
get_first_last <- function(n){
  # Strip out title and middle name
  n <- str_replace(n, '(Rev. )|(C. )|(Dr. )','')
  # Reorder if comma is present
  split_n <- unlist(str_split(n, ', '))
  if(length(split_n)==2){
    n <- paste(split_n[2],split_n[1])
  }
  return(n)
}

first_name_last_name <- as.vector(sapply(name, get_first_last))

first_name_last_name
```

(b) Construct a logical vector indicating whether a character has a title (i.e. Rev. and Dr.)

```{r}
# Look for two letters followed by a period and a space
has_title <- str_count(name, '\\w{2}. ') > 1
has_title
```

(c) Construct a logical vector indicating whether a character has a second name.

```{r}
# Find a space followed by one letter then a period then a space
has_second_name <- str_count(name, ' \\w{1}. ') == 1
has_second_name
```

## Problem 4

Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression

(a) [0-9]+\\$
(b) \\b[a-z]{1,4}\\b
(c) .*?\\.txt$
(d) \\d{2}/\\d{2}/\\d{4}
(e) <(.+?)>.+?</\\1>

### 2018-09-11 DATA607/Week_3_Assignment.Rmd
More work on DATA 607 Wee_3_Assignment.Rmd

@@
 
-(a) [0-9]+\\$
-(b) \\b[a-z]{1,4}\\b
-(c) .*?\\.txt$
-(d) \\d{2}/\\d{2}/\\d{4}
-(e) <(.+?)>.+?</\\1>
+| Question | Regular Expression | Type of String | Examples |
+|----------|--------------------|----------------|----------|
+| (a) | [0-9]+\\$ | Numbers followed by a dollar sign| 100$, 25$ |
+| (b) | \\b[a-z]{1,4}\\b | A one to four lower-case character string  | help, bat | 
+| (c) | .*?\\.txt$ | Alphanumeric characters followed by .txt | readme.txt, 1.txt |
+| (d) | \\d{2}/\\d{2}/\\d{4} | A date in a DD/MM/YYYY or MM/DD/YYYY format | 12/31/2018 |
+| (e) | <(.+?)>.+?</\\1> | Text between to HTML tags | `<b>This is bold</b>` |

### 2018-09-11 DATA607/Week_3_Assignment.Rmd
Linted Week 3 Assignment for DATA 607

@@
 library(stringr)
-raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
+raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
 
@@
   # Strip out title and middle name
-  n <- str_replace(n, '(Rev. )|(C. )|(Dr. )','')
+  n <- str_replace(n, "(Rev. )|(C. )|(Dr. )", "")
   # Reorder if comma is present
-  split_n <- unlist(str_split(n, ', '))
-  if(length(split_n)==2){
-    n <- paste(split_n[2],split_n[1])
+  split_n <- unlist(str_split(n, ", "))
+  if (length(split_n) == 2){
+    n <- paste(split_n[2], split_n[1])
   }
@@
 # Look for two letters followed by a period and a space
-has_title <- str_count(name, '\\w{2}. ') > 1
+has_title <- str_count(name, "\\w{2}. ") > 1
 has_title
@@
 # Find a space followed by one letter then a period then a space
-has_second_name <- str_count(name, ' \\w{1}. ') == 1
+has_second_name <- str_count(name, " \\w{1}. ") == 1
 has_second_name

### 2018-09-12 DATA606/Probability.Rmd
Initial commit of DATA 607 Week 2

---
title: "Probability"
author: "Mike Silva"
date: "September 11, 2018"
output: html_document
---
## Description

The following is the graded problems for the Probability chapter of Open Intro to Statistics.  My answers to the questions are **bolded.**  The headers correspond to a question.

## 2.6 Dice rolls

If you roll a pair of fair dice, what is the probability of

(a) getting a sum of 1?

(b) getting a sum of 5?

(c) getting a sum of 12?

## 2.8 Poverty and language

The American Community Survey is an ongoing survey that provides data every year to give communities the current information they need to plan investments and services. The 2010 American Community Survey estimates that 14.6% of Americans live below the poverty line, 20.7% speak a language other than English (foreign language) at home, and 4.2% fall into both categories.

(a) Are living below the poverty line and speaking a foreign language at home disjoint?

(b) Draw a Venn diagram summarizing the variables and their associated probabilities.

(c) What percent of Americans live below the poverty line and only speak English at home?

(d) What percent of Americans live below the poverty line or speak a foreign language at home?

(e) What percent of Americans live above the poverty line and only speak English at home?

(f) Is the event that someone lives below the poverty line independent of the event that the person speaks a foreign language at home?

## 2.20 Assortative mating

Assortative mating is a nonrandom mating pattern where individuals with similar genotypes and/or phenotypes mate with one another more frequently than what would be expected under a random mating pattern. Researchers studying this topic collected data on eye colors of 204 Scandinavian men and their female partners. The table below summarizes the results. For simplicity, we only include heterosexual relationships in this exercise.

<table class="table table-condensed">
<thead>
<tr class="header"><th colspan="2"></th><th colspan="3"><i>Partner (female)</i></th><th></th></tr>
</thead>
<tbody>
<tr><th colspan="2"></th><th>Blue</th><th>Brown</th><th>Green</th><th>Total</th></tr>
<tr><th rowspan="3"><i>Self (male)</i></th><th>Blue</th><td>78</td><td>23</td><td>13</td><td>114</td></tr>
<tr><th>Brown</th><td>19</td><td>23</td><td>12</td><td>54</td></tr>
<tr><th>Green</th><td>11</td><td>9</td><td>16</td><td>36</td></tr>
<tr><th></th><th>Total</th><td>108</td><td>55</td><td>41</td><td>204</td></tr>
</tbody>
</table>

(a) What is the probability that a randomly chosen male respondent or his partner has blue eyes?

(b) What is the probability that a randomly chosen male respondent with blue eyes has a partner with blue eyes?

(c) What is the probability that a randomly chosen male respondent with brown eyes has a partner with blue eyes? What about the probability of a randomly chosen male respondent with green eyes having a partner with blue eyes?

(d) Does it appear that the eye colors of male respondents and their partners are independent? Explain your reasoning.

## 2.30 Books on a bookshelf

The table below shows the distribution of books on a bookcase based on whether they are nonfiction or fiction and hardcover or paperback.

<table class="table table-condensed">
<thead>
<tr class="header"><th colspan="2"></th><th colspan="3"><i>Format</i></th></tr>
</thead>
<tbody>
<tr><th colspan="2"></th><th>Hardcover</th><th>Paperback</th><th>Total</th></tr>
<tr><th rowspan="2"><i>Type</i></th><th>Fiction</th><td>13</td><td>59</td><td>72</td></tr>
<tr><th>Nonfiction</th><td>15</td><td>8</td><td>23</td></tr>
<tr><td></td><th>Total</th><td>28</td><td>67</td><td>95</td></tr>
</tbody>
</table>

(a) Find the probability of drawing a hardcover book first then a paperback fiction book second when drawing without replacement.

(b) Determine the probability of drawing a fiction book first and then a hardcover book second, when drawing without replacement.

(c) Calculate the probability of the scenario in part (b), except this time complete the calculations under the scenario where the first book is placed back on the bookcase before randomly drawing the second book.

(d) The final answers to parts (b) and (c) are very similar. Explain why this is the case.

## 2.38 Baggage fees

An airline charges the following baggage fees: $25 for the first bag and $35 for the second. Suppose 54% of passengers have no checked luggage, 34% have one piece of checked luggage and 12% have two pieces. We suppose a negligible portion of people check more than two bags.

(a) Build a probability model, compute the average revenue per passenger, and compute the corresponding standard deviation.

(b) About how much revenue should the airline expect for a flight of 120 passengers? With what standard deviation? Note any assumptions you make and if you think they are justified.

## 2.44 Income and gender

The relative frequency table below displays the distribution of annual total personal income (in 2009 inflation-adjusted dollars) for a representative sample of 96,420,486 Americans. These data come from the American Community Survey for 2005-2009. This sample is comprised of 59% males and 41% females.

|         Income       | Total |
|----------------------|-------|
| $1 to $9,999 or loss | 2.2%  |
| $10,000 to $14,999   | 4.7%  |
| $15,000 to $24,999   | 15.8% |
| $25,000 to $34,999   | 18.3% |
| $35,000 to $49,999   | 21.2% |
| $50,000 to $64,999   | 13.9% |
| $65,000 to $74,999   | 5.8%  |
| $75,000 to $99,999   | 8.4%  |
| $100,000 or more     | 9.7%  |

(a) Describe the distribution of total personal income.

(b) What is the probability that a randomly chosen US resident makes less than $50,000 per year?

(c) What is the probability that a randomly chosen US resident makes less than $50,000 per year and is female? Note any assumptions you make.

(d) The same data source indicates that 71.8% of females make less than $50,000 per year. Use this value to determine whether or not the assumption you made in
part (c) is valid.

### 2018-09-14 DATA607/Project_1.Rmd
Initial commit of DATA 607 Project 1

---
title: "DATA 607 Project 1 - Chess Scores"
author: "Mike Silva"
date: "September 9, 2018"
output: html_document
---

## Description

In this project, I was given a text file with chess tournament results where the information has some structure. The task is to create an R Markdown file that generates a .CSV file with the following information for all of the players:

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

For the first player, the information would be: Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

## Setting up My Environment

Before I explain my whole approach to this task I will do a bit of preliminary work to set up my environment.  I will load the libraries I will be using:

```{r load_libraries, warning=FALSE, message=FALSE}
library(readr)
library(stringr)
library(dplyr)
```

Next I will read in the chess data using readr's read_lines function which loads each line into a character vector.

```{r read_data}
lines <- read_lines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/tournamentinfo.txt")
lines[1:7]
```

## My Approach

Now to explain my apporach.  I will loop through the character vector.  I will check if the line is made up of dashes.  After reading two dashed lines I will start extracting data.  There are two lines that have useful data.  The first has the player's id, name and match data.  The second line has the player's state and starting score.

I will extract the player data into a list then load that into a dataframe.  The match data will be a set of player ids that I will join to the player scores to compute the average.  Once I have these averages I will merge that in with the player data.

The name data is in all capital letters.  We will need to function that will capitalize the first letter.

```{r capitalize}
capitalize_first_letter <- function(x) {
  x <- tolower(x)
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}
```

Now we are ready to process the file.  First we will initialize some variables.
```{r init_vars}
# Initialize counters
dash_lines_found <- 0
data_line_count <- 0
# Initialize the extract data flag
extract_data <- FALSE
# Initialize a vector that will hold the match data
matches <- c()
```

### Data Extraction

Now for the heart of the project: the for loop.

```{r the_loop}
# Loop through line by line
for (line in lines){
  # Look for the dash line
  if (line == "-----------------------------------------------------------------------------------------"){
    # Increment the dash line count
    dash_lines_found <- dash_lines_found + 1
    # Check if we need to reset the data line counter
    if (extract_data){
      data_line_count <- 0
    }
    # Check if we need to extract data
    if (dash_lines_found == 2){
      extract_data <- TRUE
    }
  } else if (extract_data) {
    # Increase the data line count by one
    data_line_count <- data_line_count + 1
    # Split the data by the pipe character
    split_line <- unlist(strsplit(line, "\\|"))
    # Extract the data
    if (data_line_count == 1){
      # This is the first line so extract the player's id
      player_id <- str_extract(split_line[1], "[[:digit:]]+")
      # Extract name, split name into parts, capitalize the first letter and
      # put it back together
      player_name <- str_extract(split_line[2], "[[:alpha:]]+ [[:alpha:]]+")
      player_name <- unlist(strsplit(player_name, " "))
      player_name <- capitalize_first_letter(player_name)
      player_name <- paste(player_name, collapse = " ")
      # Loop through the match results
      for (game in split_line[4:10]){
        # Extract if it is a win, loss, draw, etc.
        result <- str_extract(game, "[[:alpha:]]")
        # Extract the opponent's id
        opponent_id <- str_extract(game, "[[:digit:]]+")
        # Add the data to the match vector
        matches <- rbind(matches, c(player_id, opponent_id, result))
      }
    } else {
      # This is the second line so Get the player's rating
      player_rating <- unlist(str_extract_all(split_line[2], "[[:digit:]]+"))
      player_rating <- as.numeric(player_rating[2])
      # Get the player's state
      player_state <- str_extract(split_line[1], "[[:alpha:]]+")
      # Pull all the extracted player data together into a list
      player_data <- list("player_id" = player_id,
                          "players_name" = player_name,
                          "players_state" = player_state,
                          "players_pre_rating" = player_rating)
      # Turn that list into a dataframe with one row
      player_row <- data.frame(bind_rows(player_data))
      # Add it to the players dataframe
      if (!exists("player_df")){
        player_df <- player_row
      } else {
        player_df <- rbind(player_df, player_row)
      }
    }
  }
}
```

### Extracton Results

Now that the data is extracted we need to take a look at it.  Here's what the extracted player data looks like:

```{r view_player_df}
head(player_df)
```

We also have the match data:

```{r view_matches}
matches[1:6,]
```

## Post-Processing

Now that we have extracted the data we move into the post-processing stage.  First we need to turn the match results data into a dataframe, rename the variables and add in the player ranking data.

```{r wrangel_matches}
# Take the match results
matches <- matches %>%
  data.frame() %>%
  rename(player_id = X1,
         opponent = X2,
         results = X3)
# Add in the player's initial rating
matches <- player_df %>%
  rename(opponent = player_id) %>%
  select(opponent, players_pre_rating) %>%
  merge(matches)
```

Here's what our match results dataframe looks like

```{r view_matches_2}
head(matches)
```

Now we need to compute the averages.  We need to group the data by player id and compute an average using a dplyr pipeline.

```{r average_matches}
averages <- matches %>%
  group_by(player_id) %>%
  summarise(average = mean(players_pre_rating))
```

Here's what the averages looks like:

```{r view_averages}
head(averages)
```

Now we need to merge the averages to the player dataframe and drop the player id so it matches what is asked for in the description.  The averages are doubles, not integers like the example so we will adjust that

```{r add_averages}
player_df <- merge(player_df, averages) %>%
  mutate(average_opponent_pre_rating = as.integer(average)) %>%
    select(-player_id, -average)
```

Here's what the dataframe looks like:

```{r view_player_df_2}
head(player_df)
```

### Finalizing the CSV

Now that the player data is in the form specified we need to save it as a CSV.

```{r save_csv}
write_csv(player_df, 'data/chess.csv')
```

### 2018-09-14 DATA607/Project_1.Rmd
Additional work on Project 1

@@
 
-## Setting up My Environment
+## Setting Up My Environment
 
@@
 
-Now to explain my apporach.
 I will loop through the character vector.
@@
       # put it back together
-      player_name <- str_extract(split_line[2], "[[:alpha:]]+ [[:alpha:]]+")
-      player_name <- unlist(strsplit(player_name, " "))
+      player_name <- str_extract_all(split_line[2], "[[:alpha:]]+")
+      player_name <- unlist(player_name)
       player_name <- capitalize_first_letter(player_name)
       player_name <- paste(player_name, collapse = " ")
+      # Extract total number of points
+      player_total_points <- as.numeric(trimws(split_line[3]))
       # Loop through the match results
@@
                           "players_state" = player_state,
+                          "player_total_points" = player_total_points,
                           "players_pre_rating" = player_rating)

### 2018-09-16 DATA607/Project_1.Rmd
More work on Project 1

@@
 ```{r add_averages}
-player_df <- merge(player_df, averages) %>%
+df <- merge(player_df, averages) %>%
   mutate(average_opponent_pre_rating = as.integer(average)) %>%
@@
 
-```{r view_player_df_2}
-head(player_df)
+```{r view_df}
+head(df)
 ```
@@
 ```
+
+## Fun with the Data
+
+I really enjoy graph theory and couldn't pass up the oportunity to create a network graph from this chess data and visualize it.
+I will use the igraph library to do that.
+This is beyond the scope of the project and is for the fun of it.
+
+```{r}
+library(igraph)
+G <- matches %>%
+  select(player_id, opponent) %>%
+  graph_from_data_frame(., directed = FALSE)
+
+nodes <- data.frame(V(G)$name) %>%
+  rename(player_id = V.G..name) %>%
+  merge(player_df) %>%
+  select(player_id, players_name)
+
+par(oma=c(0, 0, 0, 0))
+par(mar=c(0, 0, 0, 0))
+#par(pin=c(3, 3)) 
+#plot(G, layout = layout_with_fr(G))
+plot(G, layout = layout_with_graphopt(G))
+#G$layout <- layout_in_circle
+#plot(G)
+```

### 2018-09-19 DATA606/Lab2/michael-probability.Rmd
More work on the probability assignments

---
title: "Probability"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    css: ./lab.css
---

## Hot Hands

Basketball players who make several baskets in succession are described as 
having a *hot hand*. Fans and players have long believed in the hot hand 
phenomenon, which refutes the assumption that each shot is independent of the 
next. However, a 1985 paper by Gilovich, Vallone, and Tversky collected evidence
that contradicted this belief and showed that successive shots are independent 
events ([http://psych.cornell.edu/sites/default/files/Gilo.Vallone.Tversky.pdf](http://psych.cornell.edu/sites/default/files/Gilo.Vallone.Tversky.pdf)). This paper started a great controversy that continues to this day, as you can 
see by Googling *hot hand basketball*.

We do not expect to resolve this controversy today. However, in this lab we'll 
apply one approach to answering questions like this. The goals for this lab are 
to (1) think about the effects of independent and dependent events, (2) learn 
how to simulate shooting streaks in R, and (3) to compare a simulation to actual
data in order to determine if the hot hand phenomenon appears to be real.

## Saving your code

Click on File -> New -> R Script. This will open a blank document above the 
console. As you go along you can copy and paste your code here and save it. This
is a good way to keep track of your code and be able to reuse it later. To run 
your code from this document you can either copy and paste it into the console, 
highlight the code and hit the Run button, or highlight the code and hit 
command+enter or a mac or control+enter on a PC.

You'll also want to save this script (code document). To do so click on the disk 
icon. The first time you hit save, RStudio will ask for a file name; you can 
name it anything you like. Once you hit save you'll see the file appear under 
the Files tab in the lower right panel. You can reopen this file anytime by 
simply clicking on it.

## Getting Started

Our investigation will focus on the performance of one player: Kobe Bryant of 
the Los Angeles Lakers. His performance against the Orlando Magic in the 2009 
NBA finals earned him the title *Most Valuable Player* and many spectators 
commented on how he appeared to show a hot hand. Let's load some data from those
games and look at the first several rows.

```{r load-data, eval=TRUE}
load("more/kobe.RData")
head(kobe)
```

In this data frame, every row records a shot taken by Kobe Bryant. If he hit the
shot (made a basket), a hit, `H`, is recorded in the column named `basket`, 
otherwise a miss, `M`, is recorded.

Just looking at the string of hits and misses, it can be difficult to gauge 
whether or not it seems like Kobe was shooting with a hot hand. One way we can 
approach this is by considering the belief that hot hand shooters tend to go on 
shooting streaks. For this lab, we define the length of a shooting streak to be 
the *number of consecutive baskets made until a miss occurs*.

For example, in Game 1 Kobe had the following sequence of hits and misses from 
his nine shot attempts in the first quarter:

\[ \textrm{H M | M | H H M | M | M | M} \]

To verify this use the following command:

```{r first9, eval=TRUE}
kobe$basket[1:9]
```

Within the nine shot attempts, there are six streaks, which are separated by a 
"|" above. Their lengths are one, zero, two, zero, zero, zero (in order of 
occurrence).

1.  What does a streak length of 1 mean, i.e. how many hits and misses are in a 
    streak of 1? What about a streak length of 0?

    **A streak length of 1 means one hit followed by one miss.  A streak of length zero means one miss.**

The custom function `calc_streak`, which was loaded in with the data, may be 
used to calculate the lengths of all shooting streaks and then look at the 
distribution.

```{r calc-streak-kobe, eval=TRUE}
kobe_streak <- calc_streak(kobe$basket)
barplot(table(kobe_streak))
```

Note that instead of making a histogram, we chose to make a bar plot from a 
table of the streak data. A bar plot is preferable here since our variable is 
discrete -- counts -- instead of continuous.

2.  Describe the distribution of Kobe's streak lengths from the 2009 NBA finals. 
    What was his typical streak length? How long was his longest streak of baskets?

    **The typical (mode) streak has a length of zero.  Kobe's longes streak of baskets is four.**

## Compared to What?

We've shown that Kobe had some long shooting streaks, but are they long enough 
to support the belief that he had hot hands? What can we compare them to?

To answer these questions, let's return to the idea of *independence*. Two 
processes are independent if the outcome of one process doesn't effect the outcome 
of the second. If each shot that a player takes is an independent process, 
having made or missed your first shot will not affect the probability that you
will make or miss your second shot.

A shooter with a hot hand will have shots that are *not* independent of one 
another. Specifically, if the shooter makes his first shot, the hot hand model 
says he will have a *higher* probability of making his second shot.

Let's suppose for a moment that the hot hand model is valid for Kobe. During his
career, the percentage of time Kobe makes a basket (i.e. his shooting 
percentage) is about 45%, or in probability notation,

\[ P(\textrm{shot 1 = H}) = 0.45 \]

If he makes the first shot and has a hot hand (*not* independent shots), then 
the probability that he makes his second shot would go up to, let's say, 60%,

\[ P(\textrm{shot 2 = H} \, | \, \textrm{shot 1 = H}) = 0.60 \]

As a result of these increased probabilites, you'd expect Kobe to have longer 
streaks. Compare this to the skeptical perspective where Kobe does *not* have a
hot hand, where each shot is independent of the next. If he hit his first shot,
the probability that he makes the second is still 0.45.

\[ P(\textrm{shot 2 = H} \, | \, \textrm{shot 1 = H}) = 0.45 \]

In other words, making the first shot did nothing to effect the probability that
he'd make his second shot. If Kobe's shots are independent, then he'd have the 
same probability of hitting every shot regardless of his past shots: 45%.

Now that we've phrased the situation in terms of independent shots, let's return
to the question: how do we tell if Kobe's shooting streaks are long enough to 
indicate that he has hot hands? We can compare his streak lengths to someone
without hot hands: an independent shooter.

## Simulations in R

While we don't have any data from a shooter we know to have independent shots, 
that sort of data is very easy to simulate in R. In a simulation, you set the 
ground rules of a random process and then the computer uses random numbers to 
generate an outcome that adheres to those rules. As a simple example, you can
simulate flipping a fair coin with the following.

```{r head-tail, eval=TRUE}
outcomes <- c("heads", "tails")
sample(outcomes, size = 1, replace = TRUE)
```

The vector `outcomes` can be thought of as a hat with two slips of paper in it: 
one slip says `heads` and the other says `tails`. The function `sample` draws 
one slip from the hat and tells us if it was a head or a tail.

Run the second command listed above several times. Just like when flipping a 
coin, sometimes you'll get a heads, sometimes you'll get a tails, but in the 
long run, you'd expect to get roughly equal numbers of each.

If you wanted to simulate flipping a fair coin 100 times, you could either run 
the function 100 times or, more simply, adjust the `size` argument, which 
governs how many samples to draw (the `replace = TRUE` argument indicates we put
the slip of paper back in the hat before drawing again). Save the resulting 
vector of heads and tails in a new object called `sim_fair_coin`.

```{r sim-fair-coin, eval=TRUE}
sim_fair_coin <- sample(outcomes, size = 100, replace = TRUE)
```

To view the results of this simulation, type the name of the object and then use
`table` to count up the number of heads and tails.

```{r table-sim-fair-coin, eval=TRUE}
sim_fair_coin
table(sim_fair_coin)
```

Since there are only two elements in `outcomes`, the probability that we "flip" 
a coin and it lands heads is 0.5. Say we're trying to simulate an unfair coin 
that we know only lands heads 20% of the time. We can adjust for this by adding 
an argument called `prob`, which provides a vector of two probability weights.

```{r sim-unfair-coin, eval=TRUE}
sim_unfair_coin <- sample(outcomes, size = 100, replace = TRUE, prob = c(0.2, 0.8))
```

`prob=c(0.2, 0.8)` indicates that for the two elements in the `outcomes` vector,
we want to select the first one, `heads`, with probability 0.2 and the second 
one, `tails` with probability 0.8. Another way of thinking about this is to 
think of the outcome space as a bag of 10 chips, where 2 chips are labeled 
"head" and 8 chips "tail". Therefore at each draw, the probability of drawing a 
chip that says "head"" is 20%, and "tail" is 80%.

3.  In your simulation of flipping the unfair coin 100 times, how many flips 
    came up heads?
    ```{r}
    set.seed(1)
    sim_unfair_coin <- sample(outcomes, size = 100, replace = TRUE, prob = c(0.2, 0.8))
    table(sim_unfair_coin)

    ```
    **Heads came up 17 times**

In a sense, we've shrunken the size of the slip of paper that says "heads", 
making it less likely to be drawn and we've increased the size of the slip of 
paper saying "tails", making it more likely to be drawn. When we simulated the 
fair coin, both slips of paper were the same size. This happens by default if 
you don't provide a `prob` argument; all elements in the `outcomes` vector have 
an equal probability of being drawn.

If you want to learn more about `sample` or any other function, recall that you 
can always check out its help file.

```{r help-sample, eval=TRUE,tidy = FALSE}
?sample
```

## Simulating the Independent Shooter

Simulating a basketball player who has independent shots uses the same mechanism 
that we use to simulate a coin flip. To simulate a single shot from an 
independent shooter with a shooting percentage of 50% we type,

```{r sim-basket, eval=TRUE}
outcomes <- c("H", "M")
sim_basket <- sample(outcomes, size = 1, replace = TRUE)
```

To make a valid comparison between Kobe and our simulated independent shooter, 
we need to align both their shooting percentage and the number of attempted shots.

4.  What change needs to be made to the `sample` function so that it reflects a 
    shooting percentage of 45%? Make this adjustment, then run a simulation to 
    sample 133 shots. Assign the output of this simulation to a new object called
    `sim_basket`.
    ```{r}
    set.seed(1)
    sim_basket <- sample(outcomes, size = 133, replace = TRUE, prob = c(0.45, 0.55))
    ```

Note that we've named the new vector `sim_basket`, the same name that we gave to
the previous vector reflecting a shooting percentage of 50%. In this situation, 
R overwrites the old object with the new one, so always make sure that you don't
need the information in an old vector before reassigning its name.

With the results of the simulation saved as `sim_basket`, we have the data 
necessary to compare Kobe to our independent shooter. We can look at Kobe's data 
alongside our simulated data.

```{r compare-basket, eval=TRUE}
kobe$basket
sim_basket
```

Both data sets represent the results of 133 shot attempts, each with the same 
shooting percentage of 45%. We know that our simulated data is from a shooter 
that has independent shots. That is, we know the simulated shooter does not have
a hot hand.

* * *

## On your own

### Comparing Kobe Bryant to the Independent Shooter

Using `calc_streak`, compute the streak lengths of `sim_basket`.

-   Describe the distribution of streak lengths. What is the typical streak 
    length for this simulated independent shooter with a 45% shooting percentage?
    How long is the player's longest streak of baskets in 133 shots?
    ```{r}
    sim_streak <- calc_streak(sim_basket)
    barplot(table(sim_streak))
    table(sim_streak)
    ```
    **The typical streak length is zero.  The longest streak length is four.**

-   If you were to run the simulation of the independent shooter a second time, 
    how would you expect its streak distribution to compare to the distribution 
    from the question above? Exactly the same? Somewhat similar? Totally 
    different? Explain your reasoning.

    **I would expect it to be similar but not exactly the same.**

-   How does Kobe Bryant's distribution of streak lengths compare to the 
    distribution of streak lengths for the simulated shooter? Using this 
    comparison, do you have evidence that the hot hand model fits Kobe's 
    shooting patterns? Explain.

    **They are similar distributions.  This suggests that Kobe Bryant is very similar to the simulated independet shooter.  The hot hand model does not explain Kobe's shooting percentage.**

    ```{r, warning=FALSE, message=FALSE}
    library(dplyr)
    library(ggplot2)
    viz_df <- data.frame(streak = kobe_streak) %>%
      mutate(case = "Kobe")
    viz_df <- data.frame(streak = sim_streak) %>%
      mutate(case = "Simulation") %>%
      rbind(viz_df)

    ggplot(viz_df, aes(x=streak, color=as.factor(case))) + geom_density() + theme(axis.title = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.title = element_blank(), legend.position = "bottom", plot.title = element_text(hjust = 0.5)) + ggtitle("Streak Distributions by Case")
    ```

<div id="license">
This is a product of OpenIntro that is released under a 
[Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel
from a lab written by Mark Hansen of UCLA Statistics.
</div>

### 2018-09-19 DATA606/Probability.Rmd
More work on Probability.Rmd
More work on the probability assignments

@@
 author: "Mike Silva"
-date: "September 11, 2018"
+date: "September 9, 2018"
 output: html_document
@@
 
-(a) getting a sum of 1?
+(a) getting a sum of 1? **0.  It is impossible.**
 
-(b) getting a sum of 5?
+(b) getting a sum of 5? **4/36 or 0.1111111**
 
-(c) getting a sum of 12?
+(c) getting a sum of 12? **1/36 or 0.02777778**
 
@@
 
-(a) Are living below the poverty line and speaking a foreign language at home disjoint?
+(a) Are living below the poverty line and speaking a foreign language at home disjoint? **No because if they were there wouldn't be the 4.2% that fall into both categories.  It would be 0%.**
 
 (b) Draw a Venn diagram summarizing the variables and their associated probabilities.
+```{r venn_diagram, warning=FALSE, message=FALSE, results='hide'}
+library(VennDiagram)
+grid.newpage()
+draw.pairwise.venn(area1 = 20.7, area2 = 14.6, cross.area = 4.2, category = c("Foreign Lang", "Poverty"), fill = c("cornflower blue", "yellow"))
+```
 
-(c) What percent of Americans live below the poverty line and only speak English at home?
+(c) What percent of Americans live below the poverty line and only speak English at home? **10.4%**
 
-(d) What percent of Americans live below the poverty line or speak a foreign language at home?
+(d) What percent of Americans live below the poverty line or speak a foreign language at home? **26.9%**
 
-(e) What percent of Americans live above the poverty line and only speak English at home?
+(e) What percent of Americans live above the poverty line and only speak English at home? **73.1%**
 
 (f) Is the event that someone lives below the poverty line independent of the event that the person speaks a foreign language at home?
+**No.
+If they were independent than the probability of living in poverty would be the same for those that speak english and those that don't.**
 
@@
 
-(a) What is the probability that a randomly chosen male respondent or his partner has blue eyes?
+(a) What is the probability that a randomly chosen male respondent or his partner has blue eyes? **(108 + 114 - 78)/204 = 144/204 = 0.7058824**
 
-(b) What is the probability that a randomly chosen male respondent with blue eyes has a partner with blue eyes?
+(b) What is the probability that a randomly chosen male respondent with blue eyes has a partner with blue eyes? **78/114 = 0.6842105**
 
-(c) What is the probability that a randomly chosen male respondent with brown eyes has a partner with blue eyes?
-What about the probability of a randomly chosen male respondent with green eyes having a partner with blue eyes?
+(c) What is the probability that a randomly chosen male respondent with brown eyes has a partner with blue eyes?  **19/54 = 0.3518519**
 
-(d) Does it appear that the eye colors of male respondents and their partners are independent? Explain your reasoning.
+What about the probability of a randomly chosen male respondent with green eyes having a partner with blue eyes? **11/36 = 0.3055556**
+
+(d) Does it appear that the eye colors of male respondents and their partners are independent?
+Explain your reasoning.
+**No.
+If they were independent, than the female eye colors would be evenly distributed across the categories.**
 

### 2018-09-19 DATA606/Lab2/michael-probability.Rmd
Published DATA 606 Lab 2

@@
 
-    **They are similar distributions.  This suggests that Kobe Bryant is very similar to the simulated independet shooter.  The hot hand model does not explain Kobe's shooting percentage.**
+    **They are similar distributions.
+This suggests that Kobe Bryant is very similar to the simulated independent shooter.
+The hot hand model does not explain Kobe's shooting percentage.
+Additional simulations should be run before making a firm conclusion about the hot hand model.**
 

### 2018-09-19 DATA607/Project_1.Rmd
Updated DATA 607 Project 1 Added in network graph analysis for the fun of it

--- 2018-09-14 DATA607/Project_1.Rmd
@@
 ```
+
+## Fun with the Data
+
+I really enjoy graph theory and couldn't pass up the oportunity to create a network graph from this chess data.
+I recently learned how to use the igraph library from DataCamp so I will practice my skills on this data.
+This is beyond the scope of the project and is for the fun of it.
+
+I want the graph to use the players names as the ids
+```{r, warning=FALSE, message=FALSE}
+library(igraph)
+
+players_names_1 <- player_df %>%
+  select(player_id, players_name)
+
+players_names_2 <- player_df %>%
+  rename(opponent = player_id,
+         opponents_name = players_name) %>%
+  select(opponent, opponents_name)
+
+G <- matches %>%
+  select(player_id, opponent) %>%
+  merge(players_names_1) %>%
+  merge(players_names_2) %>%
+  select(players_name, opponents_name) %>%
+  graph_from_data_frame(., directed = FALSE) %>%
+  simplify()
+```
+
+Now that we have loaded up the data into a graph let's see who Gary Hua played
+```{r}
+neighbors(G, 'Gary Hua')
+```
+
+One last thing.  I want to see if any of the people who played Gary Hua played each other.  I want to visualize it.
+
+```{r}
+G_sub <- induced_subgraph(G, neighbors(G, 'Gary Hua'))
+layout <- layout.reingold.tilford(G_sub, circular=T)
+plot(G_sub, layout=layout, vertex.size=20, vertex.color='white', vertex.frame.color = 'cornflower blue')
+```
+
+It looks like Joel R Hendon and Dinh Dang Bui were the only two who played each other and Gary Hua.

### 2018-09-20 DATA606/Probability.Rmd
Updated Probability.Rmd
Additional work on the problem set

@@
 
-(a) Find the probability of drawing a hardcover book first then a paperback fiction book second when drawing without replacement.
+(a) Find the probability of drawing a hardcover book first then a paperback fiction book second when drawing without replacement. **28/95 x 59/94 = 0.2947368 x 0.6276596 = 0.1849944**
 
-(b) Determine the probability of drawing a fiction book first and then a hardcover book second, when drawing without replacement.
+(b) Determine the probability of drawing a fiction book first and then a hardcover book second, when drawing without replacement. **72/95 x 28/94 = 0.7578947 x 0.2978723 = 0.2257559**
 
-(c) Calculate the probability of the scenario in part (b), except this time complete the calculations under the scenario where the first book is placed back on the bookcase before randomly drawing the second book.
+(c) Calculate the probability of the scenario in part (b), except this time complete the calculations under the scenario where the first book is placed back on the bookcase before randomly drawing the second book.**72/95 x 28/95 = 0.7578947 x 0.2978723 = 0.2233795**
 
-(d) The final answers to parts (b) and (c) are very similar. Explain why this is the case.
+(d) The final answers to parts (b) and (c) are very similar.
+Explain why this is the case.
+**Because they are basically the same scenario.
+The only difference is the second one was drawing with replacement.**
 
@@
 
+```{r}
+fees <- c(rep(0, 54), rep(25, 34), rep(35+25, 12))
+mean(fees)
+```
+
+**Average revenue = $15.70**
+
+```{r}
+sd(fees)
+```
+**Standard deviation = $20.05**
+
 (b) About how much revenue should the airline expect for a flight of 120 passengers? With what standard deviation? Note any assumptions you make and if you think they are justified.
+
+```{r}
+n <- 120
+zero_fee <- n * .54
+twenty_five_fee <- n * .34
+sixty_fee <- n * .12
+fees2 <-c(rep(0, zero_fee), rep(25, twenty_five_fee), rep(60, sixty_fee))
+mean(fees2)
+sd(fees2)
+```
+
+**The average is $15.59 with a standard deviation of $19.98.
+This assumes that the proportions hold, that the airplane has space for the 70 bags and that the airline prices are unchanged.
+I think that these are all reasonable assumptions.**
 
@@
 
-(a) Describe the distribution of total personal income.
+(a) Describe the distribution of total personal income. **Long tailed/right skewed**
 
-(b) What is the probability that a randomly chosen US resident makes less than $50,000 per year?
+(b) What is the probability that a randomly chosen US resident makes less than $50,000 per year? **0.02 + 0.047 + 0.158 + 0.183 + 0.213 = 0.621**
 
-(c) What is the probability that a randomly chosen US resident makes less than $50,000 per year and is female? Note any assumptions you make.
+(c) What is the probability that a randomly chosen US resident makes less than $50,000 per year and is female?
+Note any assumptions you make.
+**0.41 the same as the probability that they are female.
+This assumes income is not dependant on gender (a horrible assumption which I don't believe is true).**
 
 (d) The same data source indicates that 71.8% of females make less than $50,000 per year. Use this value to determine whether or not the assumption you made in
-part (c) is valid.
+part (c) is valid.  **This indicates that the assumption of income level is independant of gender is not true.**

### 2018-09-20 DATA606/Probability.Rmd
Finalize DATA 606 Probability.Rmd

@@
 
-## 2.6 Dice rolls
+## 2.6 Dice Rolls
 
@@
 
-## 2.8 Poverty and language
+## 2.8 Poverty and Language
 
@@
 
-## 2.20 Assortative mating
+## 2.20 Assortative Mating
 
@@
 
-(c) What is the probability that a randomly chosen male respondent with brown eyes has a partner with blue eyes?  **19/54 = 0.3518519**
-
-What about the probability of a randomly chosen male respondent with green eyes having a partner with blue eyes? **11/36 = 0.3055556**
+(c) What is the probability that a randomly chosen male respondent with brown eyes has a partner with blue eyes?
+**19/54 = 0.3518519**  What about the probability of a randomly chosen male respondent with green eyes having a partner with blue eyes?
+**11/36 = 0.3055556**
 
@@
 
-## 2.30 Books on a bookshelf
+## 2.30 Books on a Bookshelf
 
@@
 
-## 2.38 Baggage fees
+## 2.38 Baggage Fees
 
@@
 ```{r}
-fees <- c(rep(0, 54), rep(25, 34), rep(35+25, 12))
+get_fees <- function(n){
+  zero_fee <- n * 0.54
+  twenty_five_fee <- n * 0.34
+  sixty_fee <- n * 0.12
+  c(rep(0, zero_fee), rep(25, twenty_five_fee), rep(60, sixty_fee))
+}
+fees <- get_fees(100)
 mean(fees)
@@
 ```{r}
-n <- 120
-zero_fee <- n * .54
-twenty_five_fee <- n * .34
-sixty_fee <- n * .12
-fees2 <-c(rep(0, zero_fee), rep(25, twenty_five_fee), rep(60, sixty_fee))
-mean(fees2)
-sd(fees2)
+fees_120 <-get_fees(120)
+mean(fees_120)
+sd(fees_120)
 ```
@@
 
-## 2.44 Income and gender
+## 2.44 Income and Gender
 

### 2018-09-20 DATA606/Distributions of Random Variables .Rmd
Initial commit

---
title: "Distributions of Random Variables"
author: "Mike Silva"
date: "September 20, 2018"
output: html_document
---

## Description

The following is the graded problems for the Distribution of Random Variable chapter of Open Intro to Statistics.  My answers to the questions in **bold.**  The headers correspond to a question.

## 3.4 Triathlon Times, Part I.

In triathlons, it is common for racers to be placed into age and gender groups. Friends Leo and Mary both completed the Hermosa Beach Triathlon, where Leo competed in the Men, Ages 30 - 34 group while Mary competed in the Women, Ages 25 - 29 group. Leo completed the race in 1:22:28 (4948 seconds), while Mary completed the race in 1:31:53 (5513 seconds). Obviously Leo finished faster, but they are curious about how they did within their respective groups. Can you help them? Here is some information on the performance of their groups:

* The finishing times of the Men, Ages 30 - 34 group has a mean of 4313 seconds with a standard deviation of 583 seconds.
* The finishing times of the Women, Ages 25 - 29 group has a mean of 5261 seconds with a standard deviation of 807 seconds.
* The distributions of finishing times for both groups are approximately Normal. Remember: a better performance corresponds to a faster finish.

(a) Write down the short-hand for these two normal distributions.
(b) What are the Z-scores for Leo’s and Mary’s finishing times? What do these Z-scores tell you?
(c) Did Leo or Mary rank better in their respective groups? Explain your reasoning.
(d) What percent of the triathletes did Leo finish faster than in his group?
(e) What percent of the triathletes did Mary finish faster than in her group?
(f) If the distributions of finishing times are not nearly normal, would your answers to parts b to e change? Explain your reasoning

## 3.18 Heights of Female College Students

Below are heights of 25 female college students.

| 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 | 9 | 10 | 11 | 12 | 13 | 14 | 15 | 16 | 17 | 18 | 19 | 20 | 21 | 22 | 23 | 24 | 25 |
|---|---|---|---|---|---|---|---|---|----|----|----|----|----|----|----|----|----|----|----|----|----|----|----|----|
| 54, | 55, | 56, | 56, | 57, | 58, | 58, | 59, | 60, | 60, | 60, | 61, | 61, | 62, | 62, | 63, | 63, | 63, | 64, | 65, | 65, | 67, | 67, | 69, | 73 |

(a) The mean height is 61.52 inches with a standard deviation of 4.58 inches. Use this information to determine if the heights approximately follow the 68-95-99.7% Rule.
(b) Do these data appear to follow a normal distribution? Explain your reasoning using the graphs provided below.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/distributions_3_18.png)

## 3.22 Defective Rate

A machine that produces a special type of transistor (a component of computers) has a 2% defective rate. The production is considered a random process where each transistor is independent of the others.
(a) What is the probability that the 10th transistor produced is the first with a defect?
(b) What is the probability that the machine produces no defective transistors in a batch of 100?
(c) On average, how many transistors would you expect to be produced before the first with a defect? What is the standard deviation?
(d) Another machine that also produces transistors has a 5% defective rate where each transistor is produced independent of the others. On average how many transistors would you expect to be produced with this machine before the first with a defect? What is the standard deviation?
(e) Based on your answers to parts (c) and (d), how does increasing the probability of an event affect the mean and standard deviation of the wait time until success?

## 3.38 Male Children

While it is often assumed that the probabilities of having a boy or a girl are the same, the actual probability of having a boy is slightly higher at 0.51. Suppose a couple plans to have 3 kids.
(a) Use the binomial model to calculate the probability that two of them will be boys.
(b) Write out all possible orderings of 3 children, 2 of whom are boys. Use these scenarios to calculate the same probability from part (a) but using the addition rule for disjoint outcomes. Confirm that your answers from parts (a) and (b) match.
(c) If we wanted to calculate the probability that a couple who plans to have 8 kids will have 3 boys, briefly describe why the approach from part (b) would be more tedious than the approach from part (a).

## 3.42 2 Serving in Volleyball

A not-so-skilled volleyball player has a 15% chance of making the serve, which involves hitting the ball so it passes over the net on a trajectory such that it will land in the opposing team’s court. Suppose that her serves are independent of each other.

(a) What is the probability that on the 10th try she will make her 3rd successful serve?
(b) Suppose she has made two successful serves in nine attempts. What is the probability that her 10th serve will be successful?
(c) Even though parts (a) and (b) discuss the same scenario, the probabilities you calculated should be different. Can you explain the reason for this discrepancy?

### 2018-09-26 DATA606/Lab3/michael-normal_distribution.Rmd
More work on DATA 606 Lab 3

---
title: "The normal distribution"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    css: ./lab.css
---

In this lab we'll investigate the probability distribution that is most central
to statistics: the normal distribution.  If we are confident that our data are 
nearly normal, that opens the door to many powerful statistical methods.  Here 
we'll use the graphical tools of R to assess the normality of our data and also 
learn how to generate random numbers from a normal distribution.

## The Data

This week we'll be working with measurements of body dimensions.  This data set 
contains measurements from 247 men and 260 women, most of whom were considered 
healthy young adults.

```{r load-data, eval=TRUE}
load("more/bdims.RData")
```

Let's take a quick peek at the first few rows of the data.

```{r head-data, eval=TRUE}
head(bdims)
```

You'll see that for every observation we have 25 measurements, many of which are
either diameters or girths.  A key to the variable names can be found at 
[http://www.openintro.org/stat/data/bdims.php](http://www.openintro.org/stat/data/bdims.php),
but we'll be focusing on just three columns to get started: weight in kg (`wgt`), 
height in cm (`hgt`), and `sex` (`1` indicates male, `0` indicates female).

Since males and females tend to have different body dimensions, it will be 
useful to create two additional data sets: one with only men and another with 
only women.

```{r male-female, eval=TRUE}
mdims <- subset(bdims, sex == 1)
fdims <- subset(bdims, sex == 0)
```

1.  Make a histogram of men's heights and a histogram of women's heights.  How 
    would you compare the various aspects of the two distributions?
    ```{r, message = FALSE, echo=FALSE}
    library(dplyr)
    library(ggplot2)
bdims %>%
      mutate(sex = ifelse(sex==1, "Male", "Female")) %>%
      select(hgt, sex) %>%
      ggplot(., aes(x = hgt, color=sex, fill=sex)) + geom_histogram() + facet_grid(sex ~ .) + guides(fill=FALSE, color=FALSE)
    ```

## The normal distribution

In your description of the distributions, did you use words like *bell-shaped* 
or *normal*?  It's tempting to say so when faced with a unimodal symmetric 
distribution.

To see how accurate that description is, we can plot a normal distribution curve 
on top of a histogram to see how closely the data follow a normal distribution. 
This normal curve should have the same mean and standard deviation as the data. 
We'll be working with women's heights, so let's store them as a separate object 
and then calculate some statistics that will be referenced later.

```{r female-hgt-mean-sd, eval=TRUE}
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)
```

Next we make a density histogram to use as the backdrop and use the `lines` 
function to overlay a normal probability curve. The difference between a 
frequency histogram and a density histogram is that while in a frequency 
histogram the *heights* of the bars add up to the total number of observations, 
in a density histogram the *areas* of the bars add up to 1. The area of each bar 
can be calculated as simply the height *times* the width of the bar. Using a 
density histogram allows us to properly overlay a normal distribution curve over 
the histogram since the curve is a normal probability density function.
Frequency and density histograms both display the same exact shape; they only 
differ in their y-axis. You can verify this by comparing the frequency histogram 
you constructed earlier and the density histogram created by the commands below.

```{r hist-height, eval=TRUE}
hist(fdims$hgt, probability = TRUE)
x <- 140:190
y <- dnorm(x = x, mean = fhgtmean, sd = fhgtsd)
lines(x = x, y = y, col = "blue")
```

After plotting the density histogram with the first command, we create the x- 
and y-coordinates for the normal curve.  We chose the `x` range as 140 to 190 in 
order to span the entire range of `fheight`.  To create `y`, we use `dnorm` to 
calculate the density of each of those x-values in a distribution that is normal
with mean `fhgtmean` and standard deviation `fhgtsd`.  The final command draws a
curve on the existing plot (the density histogram) by connecting each of the 
points specified by `x` and `y`. The argument `col` simply sets the color for 
the line to be drawn. If we left it out, the line would be drawn in black.

The top of the curve is cut off because the limits of the x- and y-axes are set 
to best fit the histogram.  To adjust the y-axis you can add a third argument to
the histogram function: `ylim = c(0, 0.06)`.

2.  Based on the this plot, does it appear that the data follow a nearly normal 
    distribution?
    **Yes it looks nearly normally distributed to me.**

## Evaluating the normal distribution

Eyeballing the shape of the histogram is one way to determine if the data appear
to be nearly normally distributed, but it can be frustrating to decide just how 
close the histogram is to the curve. An alternative approach involves 
constructing a normal probability plot, also called a normal Q-Q plot for 
"quantile-quantile".

```{r qq, eval=TRUE}
qqnorm(fdims$hgt)
qqline(fdims$hgt)
```

A data set that is nearly normal will result in a probability plot where the 
points closely follow the line.  Any deviations from normality leads to 
deviations of these points from the line.  The plot for female heights shows 
points that tend to follow the line but with some errant points towards the 
tails.  We're left with the same problem that we encountered with the histogram 
above: how close is close enough?

A useful way to address this question is to rephrase it as: what do probability 
plots look like for data that I *know* came from a normal distribution?  We can 
answer this by simulating data from a normal distribution using `rnorm`.

```{r sim-norm, eval=TRUE}
sim_norm <- rnorm(n = length(fdims$hgt), mean = fhgtmean, sd = fhgtsd)
```

The first argument indicates how many numbers you'd like to generate, which we 
specify to be the same number of heights in the `fdims` data set using the 
`length` function.  The last two arguments determine the mean and standard 
deviation of the normal distribution from which the simulated sample will be 
generated.  We can take a look at the shape of our simulated data set, `sim_norm`, 
as well as its normal probability plot.

3.  Make a normal probability plot of `sim_norm`.  Do all of the points fall on 
    the line?  How does this plot compare to the probability plot for the real 
    data?
    ```{r}
    qqnorm(sim_norm)
    qqline(sim_norm)
    ```

    **Most of the points fall upon the line.  It looks very similar to the preceeding plot**

Even better than comparing the original plot to a single plot generated from a 
normal distribution is to compare it to many more plots using the following 
function. It may be helpful to click the zoom button in the plot window.

```{r qqnormsim, eval=TRUE}
qqnormsim(fdims$hgt)
```

4.  Does the normal probability plot for `fdims$hgt` look similar to the plots 
    created for the simulated data?  That is, do plots provide evidence that the
    female heights are nearly normal?
    **Yes they look similar to the plots created for the simulated data.  They seem nearly normal.**

5.  Using the same technique, determine whether or not female weights appear to 
    come from a normal distribution.
    ```{r qqnormsim2, eval=TRUE}
qqnormsim(fdims$wgt)
```

## Normal probabilities

Okay, so now you have a slew of tools to judge whether or not a variable is 
normally distributed.  Why should we care?

It turns out that statisticians know a lot about the normal distribution.  Once 
we decide that a random variable is approximately normal, we can answer all 
sorts of questions about that variable related to probability.  Take, for 
example, the question of, "What is the probability that a randomly chosen young 
adult female is taller than 6 feet (about 182 cm)?" (The study that published
this data set is clear to point out that the sample was not random and therefore 
inference to a general population is not suggested.  We do so here only as an
exercise.)

If we assume that female heights are normally distributed (a very close 
approximation is also okay), we can find this probability by calculating a Z 
score and consulting a Z table (also called a normal probability table).  In R, 
this is done in one step with the function `pnorm`.

```{r pnorm, eval=TRUE}
1 - pnorm(q = 182, mean = fhgtmean, sd = fhgtsd)
```

Note that the function `pnorm` gives the area under the normal curve below a 
given value, `q`, with a given mean and standard deviation.  Since we're 
interested in the probability that someone is taller than 182 cm, we have to 
take one minus that probability.

Assuming a normal distribution has allowed us to calculate a theoretical 
probability.  If we want to calculate the probability empirically, we simply 
need to determine how many observations fall above 182 then divide this number 
by the total sample size.

```{r probability, eval=TRUE}
sum(fdims$hgt > 182) / length(fdims$hgt)
```

Although the probabilities are not exactly the same, they are reasonably close. 
The closer that your distribution is to being normal, the more accurate the 
theoretical probabilities will be.

6.  Write out two probability questions that you would like to answer; one 
    regarding female heights and one regarding female weights.  Calculate the 
    those probabilities using both the theoretical normal distribution as well 
    as the empirical distribution (four probabilities in all).  Which variable,
    height or weight, had a closer agreement between the two methods?

* * *

## On Your Own

-   Now let's consider some of the other variables in the body dimensions data 
    set.  Using the figures at the end of the exercises, match the histogram to 
    its normal probability plot.  All of the variables have been standardized 
    (first subtract the mean, then divide by the standard deviation), so the 
    units won't be of any help.  If you are uncertain based on these figures, 
    generate the plots in R to check.

    **a.** The histogram for female biiliac (pelvic) diameter (`bii.di`) belongs
    to normal probability plot letter ____.

    **b.** The histogram for female elbow diameter (`elb.di`) belongs to normal 
    probability plot letter ____.

    **c.** The histogram for general age (`age`) belongs to normal probability 
    plot letter ____.

    **d.** The histogram for female chest depth (`che.de`) belongs to normal 
    probability plot letter ____.

-   Note that normal probability plots C and D have a slight stepwise pattern.  
    Why do you think this is the case?

-   As you can see, normal probability plots can be used both to assess 
    normality and visualize skewness.  Make a normal probability plot for female 
    knee diameter (`kne.di`).  Based on this normal probability plot, is this 
    variable left skewed, symmetric, or right skewed?  Use a histogram to confirm 
    your findings.

![histQQmatch](more/histQQmatch.png)

<div id="license">
This is a product of OpenIntro that is released under a 
[Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel
from a lab written by Mark Hansen of UCLA Statistics.
</div>

### 2018-09-27 DATA606/Lab3/michael-normal_distribution.Rmd
Additional work on DATA 606 Lab 3

@@
 
+    ```{r echo=FALSE}
+    fwgtmean <- mean(fdims$wgt)
+    fwgtsd   <- sd(fdims$wgt)
+    ```
+
+    **1) Given the mean height of a male is `r round(mean(mdims$hgt), 0)` cm, what is the probability that a female is taller than the average male?**
+
+    **The theoretical probability is: `r 1 - pnorm(q = mean(mdims$hgt), mean = fhgtmean, sd = fhgtsd)`**
+
+    **The emperical probability is: `r sum(fdims$hgt > mean(mdims$hgt)) / length(fdims$hgt)`**
+
+    **2) Given that the mean weight of a male is `r round(mean(mdims$wgt), 0)` kg, what is the probability that a female is heavier than the average male?**
+
+    **The theoretical probability is: `r 1 - pnorm(q = mean(mdims$wgt), mean = fwgtmean, sd = fwgtsd)`**
+
+    **The emperical probability is: `r sum(fdims$wgt > mean(mdims$wgt)) / length(fdims$wgt)`**
+
 * * *
@@
     **a.** The histogram for female biiliac (pelvic) diameter (`bii.di`) belongs
-    to normal probability plot letter ____.
+    to normal probability plot letter **B**.
 
     **b.** The histogram for female elbow diameter (`elb.di`) belongs to normal 
-    probability plot letter ____.
+    probability plot letter **C**.
 
     **c.** The histogram for general age (`age`) belongs to normal probability 
-    plot letter ____.
+    plot letter **D**.
 
     **d.** The histogram for female chest depth (`che.de`) belongs to normal 
-    probability plot letter ____.
+    probability plot letter **A**.
 
 -   Note that normal probability plots C and D have a slight stepwise pattern.  
-    Why do you think this is the case?
+    Why do you think this is the case?  **Because they are variable with discrete units not continuous.  Thats what gives them the stepwise pattern.**
 
@@
     your findings.
+    ```{r}
+    qqnorm(fdims$kne.di)
+    qqline(fdims$kne.di)
+    ```
 
-![histQQmatch](more/histQQmatch.png)
+    **It is right skewed.**
+
+    ```{r}
+    hist(fdims$kne.di)
+    ```
+
+![](more/histQQmatch.png)
 

### 2018-09-28 DATA606/Distributions of Random Variables .Rmd
More work on DATA 606 Assignment

@@
 (a) Write down the short-hand for these two normal distributions.
-(b) What are the Z-scores for Leo’s and Mary’s finishing times? What do these Z-scores tell you?
-(c) Did Leo or Mary rank better in their respective groups? Explain your reasoning.
+
+$$ N_{\text{men 30 to 34}}(\mu=4313, \sigma=583) $$ 
+$$ N_{\text{women 25 to 29}}(\mu=5261, \sigma=807) $$
+
+(b) What are the Z-scores for Leo’s and Mary’s finishing times?
+What do these Z-scores tell you?
+**The Z-scores tell us that Mary was closer to the mean of her group and therefore was comparitively faster than Leo.**
+
+```{r}
+z_leo <- (4948-4313)/583
+z_leo
+```
+
+```{r}
+z_mary <- (5513-5261)/807
+z_mary
+```
+
+(c) Did Leo or Mary rank better in their respective groups? Explain your reasoning.  **Mary because her Z score is lower so she is closer to the mean than Leo.**
+
 (d) What percent of the triathletes did Leo finish faster than in his group?
+
+```{r}
+# Faster would be the area to the left of Leo's time
+1 - pnorm(z_leo)
+```
+
 (e) What percent of the triathletes did Mary finish faster than in her group?
-(f) If the distributions of finishing times are not nearly normal, would your answers to parts b to e change? Explain your reasoning
+
+```{r}
+1 - pnorm(z_mary)
+```
+
+(f) If the distributions of finishing times are not nearly normal, would your answers to parts b to e change?
+Explain your reasoning **No.
+If it were a right skewed distribution then the proportion to the left of Mary's time would be different.**
 
@@
 (a) The mean height is 61.52 inches with a standard deviation of 4.58 inches. Use this information to determine if the heights approximately follow the 68-95-99.7% Rule.
+
+```{r}
+heights <- c(54, 55, 56, 56, 57, 58, 58, 59, 60, 60, 60, 61, 61, 62, 62, 63, 63, 63, 64, 65, 65, 67, 67, 69, 73)
+
+mu <- mean(heights)
+sigma <- sd(heights)
+
+for(n_sigma in 1:3){
+  # Return a Boolean vector
+  boolean_vector <- heights < (mu + n_sigma * sigma) & heights > (mu + -n_sigma * sigma)
+  # Count the TRUEs
+  n_true <- length(boolean_vector[boolean_vector==TRUE])
+  # Share TRUE
+  print(n_true / length(heights))
+}
+```
+
+**Yes, they roughly follow the 68-95-99.7 Rule.**
+
 (b) Do these data appear to follow a normal distribution? Explain your reasoning using the graphs provided below.
@@
 ![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/distributions_3_18.png)
+
+**Yes.  The dots on the Q-Q plot roughly fall on the line.  This indicates the data are normally distributed.**
 

### 2018-09-29 DATA607/Week_5_Assignment.Rmd
Initial commit of Week 5 Assignment

---
title: "Week 5 Assignment"
author: "Mike Silva"
date: "September 28, 2018"
output: html_document
---

## Description

TThe assignment is to load in an "untidy" .CSV file and transform it into a tidy dataset.  The data contains the arrival data for two airlines.  After the data has been wrangled I am to preform an analysis comparing the arrival delays for the two airlines, and publish my code and findings to rpubs and GitHub.

## Data Acquisition

I have created a .CSV file that is identical to the assignment and have it hosted in my GitHub repository.  I will read in the .CSV file and display it.

```{r}
untidy_df <- read.csv('https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/flights.csv')
```

Here's what the untidy data frame.
```{r, results='asis', echo=FALSE}
library(pander)
pander(untidy_df)
```

## Data Cleaning

### The Plan

Before starting on the data cleaning I think it is important to identify what needs to change in the data set.  The name the first two variables needs to change.  There is a blank row between the two airlines that needs to change.  The dots for the airport names need to be replaced with spaces.  This will evenutally become a categorical variable.  The first to variables data types need to be changed from factors to characters.  The first variable has some blanks that need to be filled too.

Once these changes are made the data needs to be normalized using the *gather()* command from *tidyr*.  Knowing that I will want to create some visualizations I will convert the airline, arrival type and airport variables back into factors.  It will make things easier to work with.

## The Code

```{r, message=FALSE}
library(dplyr)
library(tidyr)

tidy_df <- untidy_df %>%
  rename(airline = X,
         arrival_type = X.1,
         "Los Angeles" = Los.Angeles,
         "San Diego" = San.Diego,
         "San Francisco" = San.Francisco)%>%
  mutate(airline = as.character(airline),
         arrival_type = as.character(arrival_type)) %>%
  na.omit() %>% # Drop the empty row
  mutate(airline = ifelse(nchar(airline)==0, lag(airline), airline)) %>% # Fill in the holes
  gather(., airport, flights, -airline, -arrival_type) %>% # Transform from wide to long
  mutate(airline = as.factor(airline),
         arrival_type = as.factor(arrival_type),
         airport = as.factor(airport))
```

Here's what the table looks like.

```{r, results='asis', echo=FALSE}
pander(tidy_df)
```
## Analysis

There are a few questions that I want to ask this tidy data set.  But before I begin I will compute a few totals, and load ggplot2

```{r}
library(ggplot2)

total_flights_by_airline <- tidy_df %>%
  group_by(airline) %>%
  summarise(total_flights = sum(flights)) %>%
  ungroup()

total_flights_by_airport <- tidy_df %>%
  group_by(airport) %>%
  summarise(total_flights = sum(flights)) %>%
  ungroup()

total_flights_by_arrival_type <- tidy_df %>%
  group_by(arrival_type) %>%
  summarise(total_flights = sum(flights)) %>%
  ungroup()
```

I want to begin by examining the airlines.  What share of the flights are on time for each airline?

```{r, results='asis'}
tidy_df %>%
  group_by(airline, arrival_type) %>%
  summarise(flights = sum(flights)) %>%
  merge(total_flights_by_airline) %>%
  mutate(share = flights / total_flights) %>%
  pander()
```

It looks like AM West is on time 89% of the time on average.  Alaska airlines is on time about 87% of the time.  One also observes that AM West has more flights than Alaska airlines.

Is this difference in the on time/delayed share due to the airports the airlines fly to?  Let's examine the share of times the airline flies to an airport

```{r}
tidy_df %>%
  group_by(airline, airport) %>%
  summarise(flights = sum(flights)) %>%
  merge(total_flights_by_airport) %>%
  mutate(share = flights / total_flights) %>%
  ggplot(aes(airport, airline)) +
  geom_tile(aes(fill = share), show.legend = FALSE) +
  geom_text(aes(label = round(share*100,0))) +
  scale_fill_distiller(palette = "Spectral") + 
  coord_fixed() +
  ggtitle("Percent of Flights by Airline and Airport") +
  theme(panel.grid.major=element_blank(), axis.title.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), axis.ticks.y=element_blank(), plot.title = element_text(hjust = 0.5))
```

So Alaska airlines flies to San Fancisco and Seatle more tha AM West.  Let's look at this picture again by comparing across all flights the airline takes.

```{r}
tidy_df %>%
  group_by(airline, airport) %>%
  summarise(flights = sum(flights)) %>%
  merge(total_flights_by_airline) %>%
  mutate(share = flights / total_flights) %>%
  ggplot(aes(airport, airline)) +
  geom_tile(aes(fill = share), show.legend = FALSE) +
  geom_text(aes(label = round(share*100,0))) +
  scale_fill_distiller(palette = "Spectral") + 
  coord_fixed() +
  ggtitle("Percent of Flights by Airline and Airports") +
  theme(panel.grid.major=element_blank(), axis.title.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), axis.ticks.y=element_blank(), plot.title = element_text(hjust = 0.5))
```

This shows a litle different picture.  57% of Alaska airlines flights are to Seatle, which makes sense.  AM West flies to Phoenix most often then LA.

So lets take a closer look at the airports  Let's see if there is a difference between the arrival times and the airport.

```{r}
tidy_df %>%
  group_by(airport, arrival_type) %>%
  summarise(flights = sum(flights)) %>%
  merge(total_flights_by_airport) %>%
  mutate(share = flights / total_flights) %>%
  ggplot(aes(airport, arrival_type)) +
  geom_tile(aes(fill = share), show.legend = FALSE) +
  geom_text(aes(label = round(share*100,0))) +
  scale_fill_distiller(palette = "Spectral") + 
  coord_fixed() +
  ggtitle("Percent of Flights by Arrival Type and Airport") +
  theme(panel.grid.major=element_blank(), axis.title.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), axis.ticks.y=element_blank(), plot.title = element_text(hjust = 0.5))
```

The Phoenix airport has the best record for on time arrivals and San Francisco has the worst.

Let's look at the on time arrivals.  How do these two airlines stack up when we look at the airports.  If each airlines are the same.

```{r}
on_time_flights_by_airport <- tidy_df %>%
  filter(arrival_type == "on time") %>%
  group_by(airport) %>%
  summarise(total_flights = sum(flights)) %>%
  ungroup()

tidy_df %>%
  filter(arrival_type == "on time") %>%
  group_by(airline, airport) %>%
  summarise(flights = sum(flights)) %>%
  merge(on_time_flights_by_airport) %>%
  mutate(share = flights / total_flights) %>%
  ggplot(aes(airport, airline)) +
  geom_tile(aes(fill = share), show.legend = FALSE) +
  geom_text(aes(label = round(share*100,0))) +
  scale_fill_distiller(palette = "Spectral") + 
  coord_fixed() +
  ggtitle("Percent of On Time Flights by Airline and Airport") +
  theme(panel.grid.major=element_blank(), axis.title.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), axis.ticks.y=element_blank(), plot.title = element_text(hjust = 0.5))
```

Most of the on time flights to Phoenix are due to AM West, and most of the on time flights to Seatle are due to Alaska Airlines.  Let's look at the delayed flights.

```{r}
delayed_flights_by_airport <- tidy_df %>%
  filter(arrival_type == "delayed") %>%
  group_by(airport) %>%
  summarise(total_flights = sum(flights)) %>%
  ungroup()

tidy_df %>%
  filter(arrival_type == "delayed") %>%
  group_by(airline, airport) %>%
  summarise(flights = sum(flights)) %>%
  merge(delayed_flights_by_airport) %>%
  mutate(share = flights / total_flights) %>%
  ggplot(aes(airport, airline)) +
  geom_tile(aes(fill = share), show.legend = FALSE) +
  geom_text(aes(label = round(share*100,0))) +
  scale_fill_distiller(palette = "Spectral") + 
  coord_fixed() +
  ggtitle("Percent of Delayed Flights by Airline and Airport") +
  theme(panel.grid.major=element_blank(), axis.title.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), axis.ticks.y=element_blank(), plot.title = element_text(hjust = 0.5))
```

Most of the delayed flights to Phoenix are due to AM West, and most of the on time flights to Seatle are due to Alaska Airlines.  Let's look at the delayed flights.

### 2018-09-29 DATA607/Week_5_Assignment.Rmd
Finalizing DATA 607 Week 5 Assignment

@@
 date: "September 28, 2018"
-output: html_document
+output: 
+  rmdformats::readthedown:
+    highlight: kate
 ---
@@
 
+### Airline Examination
+
 I want to begin by examining the airlines.  What share of the flights are on time for each airline?
 
-```{r, results='asis'}
+```{r, results='asis', echo=FALSE}
 tidy_df %>%
@@
 This shows a litle different picture.  57% of Alaska airlines flights are to Seatle, which makes sense.  AM West flies to Phoenix most often then LA.
+
+### Airport Examination
 
@@
 Most of the delayed flights to Phoenix are due to AM West, and most of the on time flights to Seatle are due to Alaska Airlines.  Let's look at the delayed flights.
+
+### What If Scenario
+
+Now I want to do a what if scenario.  What if these two airlines don't fly to Phoenix or Seattle.  How would that change these airline on time arrivals?
+
+```{r}
+what_if <- tidy_df %>%
+  filter(!airport %in% c("Phoenix", "Seatle"))
+
+what_if_total_flights_by_airline <- what_if %>%
+  group_by(airline) %>%
+  summarise(total_flights = sum(flights)) %>%
+  ungroup()
+
+temp <- what_if %>%
+  group_by(airline, arrival_type) %>%
+  summarise(flights = sum(flights)) %>%
+  merge(what_if_total_flights_by_airline) %>%
+  mutate(what_if_share = flights / total_flights) %>%
+  select(airline, arrival_type, what_if_share)
+
+scenario <- tidy_df %>%
+  group_by(airline, arrival_type) %>%
+  summarise(flights = sum(flights)) %>%
+  merge(total_flights_by_airline) %>%
+  mutate(share = flights / total_flights) %>%
+  select(airline, arrival_type, share) %>%
+  merge(temp) %>%
+  mutate(delta = what_if_share - share)
+```
+
+```{r, results='asis', echo=FALSE}
+  pander(scenario)
+```
+
+We see that this would not change the airline's ratings much.  Alaska Airlines is basically unchange and AM West changed by less than 1 percentage point.
+
+What if they only flew to Phoenix or Seatle?  What would that do?
+
+```{r}
+what_if <- tidy_df %>%
+  filter(airport %in% c("Phoenix", "Seatle"))
+
+what_if_total_flights_by_airline <- what_if %>%
+  group_by(airline) %>%
+  summarise(total_flights = sum(flights)) %>%
+  ungroup()
+
+temp <- what_if %>%
+  group_by(airline, arrival_type) %>%
+  summarise(flights = sum(flights)) %>%
+  merge(what_if_total_flights_by_airline) %>%
+  mutate(what_if_share = flights / total_flights) %>%
+  select(airline, arrival_type, what_if_share)
+
+scenario <- tidy_df %>%
+  group_by(airline, arrival_type) %>%
+  summarise(flights = sum(flights)) %>%
+  merge(total_flights_by_airline) %>%
+  mutate(share = flights / total_flights) %>%
+  select(airline, arrival_type, share) %>%
+  merge(temp) %>%
+  mutate(delta = what_if_share - share)
+```
+
+```{r, results='asis', echo=FALSE}
+  pander(scenario)
+```
+
+That would give Alaska Airlines on time arrivals a boost of 8 percentage points and AM West a 3 percentage point boost.
+
+### Optimization
+
+Let's suppose that these two airlines live and die by their on time arrival rating.  What set of airports would then need to fly to to get their highest rating?
+
+I will use the combn function to build out the list of all possible combinations.  I will loop through these combinations and run what if scenarios.  If the on time arrival share is bigger than the
+
+```{r}
+# Initial settings for on time arrival share
+alaska_on_time <- 0
+am_west_on_time <- 0
+
+# Get a vector of airports
+airports <- tidy_df$airport %>%
+  unique() %>%
+  as.vector()
+
+# Loop through all combinations
+for(m in length(airports):1){
+  airport_combinations <- combn(airports, m)
+  # Run a what if scenario for the combination
+  for(col in 1:dim(airport_combinations)[2]){
+    what_if <- tidy_df %>%
+      filter(airport %in% airport_combinations[, col])
+
+    what_if_total_flights_by_airline <- what_if %>%
+      group_by(airline) %>%
+      summarise(total_flights = sum(flights)) %>%
+      ungroup()
+
+    temp <- what_if %>%
+      filter(arrival_type == "on time") %>%
+      group_by(airline) %>%
+      summarise(flights = sum(flights)) %>%
+      merge(what_if_total_flights_by_airline) %>%
+      mutate(what_if_share = flights / total_flights) %>%
+      select(airline, what_if_share)
+
+    # See if the on time share is larger and if so save the airport set 
+    if(temp[1,2] > alaska_on_time){
+      alaska_on_time <- temp[1, 2]
+      alaska_optimal <- airport_combinations[, col]
+    }
+
+    if(temp[2,2] > am_west_on_time){
+      am_west_on_time <- temp[2, 2]
+      am_west_optimal <- airport_combinations[, col]
+    }
+  }
+}
+```
+
+So currently Alaska airlines and AM West has an on time arrival 86.73% and 89.11% respectively.
+If they selected the optimal set of airports to fly into the on time arrival would increase to `r round(alaska_on_time * 100, 2)`% and `r round(am_west_on_time * 100, 2)`%.
+
+Alaska Airlines would only fly to `r pander(alaska_optimal)` and AM West would only fly to `r pander(am_west_optimal)`.

### 2018-09-29 DATA606/Distributions of Random Variables .Rmd
More work

@@
 date: "September 20, 2018"
-output: html_document
+output:
+  html_document:
+    theme: yeti
+    highlight: tango
 ---
@@
 
-(d) What percent of the triathletes did Leo finish faster than in his group?
+(d) What percent of the triathletes did Leo finish faster than in his group? **About 14%**
 
@@
 
-(e) What percent of the triathletes did Mary finish faster than in her group?
+(e) What percent of the triathletes did Mary finish faster than in her group? **Roughly 38%**
 
@@
 
-(a) The mean height is 61.52 inches with a standard deviation of 4.58 inches. Use this information to determine if the heights approximately follow the 68-95-99.7% Rule.
+(a) The mean height is 61.52 inches with a standard deviation of 4.58 inches.
+Use this information to determine if the heights approximately follow the 68-95-99.7% Rule.
+**Yes, they roughly follow the 68-95-99.7 Rule.**
 
@@
 
-**Yes, they roughly follow the 68-95-99.7 Rule.**
-
-(b) Do these data appear to follow a normal distribution? Explain your reasoning using the graphs provided below.
+(b) Do these data appear to follow a normal distribution?
+Explain your reasoning using the graphs provided below.
+**Yes.
+The dots on the Q-Q plot roughly fall on the line.
+This indicates the data are normally distributed.**
 
 ![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/distributions_3_18.png)
-
-**Yes.  The dots on the Q-Q plot roughly fall on the line.  This indicates the data are normally distributed.**
 
@@
 The production is considered a random process where each transistor is independent of the others.
-(a) What is the probability that the 10th transistor produced is the first with a defect?
-(b) What is the probability that the machine produces no defective transistors in a batch of 100?
-(c) On average, how many transistors would you expect to be produced before the first with a defect? What is the standard deviation?
+
+(a) What is the probability that the 10th transistor produced is the first with a defect? **The probability is roughly 0.017.**
+
+```{r}
+probability_of_defect <- 0.02
+
+probability_of_10th_first_defect = 1
+
+for(i in 1:9){
+  probability_of_10th_first_defect <- probability_of_10th_first_defect * (1 - probability_of_defect)
+}
+
+probability_of_10th_first_defect <- probability_of_10th_first_defect * probability_of_defect
+
+probability_of_10th_first_defect
+```
+
+This matches the geometric distribution formula's 
+$(1-p)^{n-1}\times p$
+
+```{r}
+((1 - 0.02)^(10-1))*0.02
+```
+
+(b) What is the probability that the machine produces no defective transistors in a batch of 100? **About 0.13.**
+
+```{r}
+probability_of_no_defect = 1
+
+for(i in 1:100){
+   probability_of_no_defect <- probability_of_no_defect * (1 - probability_of_defect)
+}
+
+probability_of_no_defect
+```
+
+(c) On average, how many transistors would you expect to be produced before the first with a defect?
+**A 2% defect rate means out of 100 produced 2 would be defective.
+Therefore we would expect 1 defective transitor out of 50.**  What is the standard deviation?
+**0.14**
+
+```{r}
+sigma <- sqrt(probability_of_defect*(1 - probability_of_defect))
+sigma
+```
+
 (d) Another machine that also produces transistors has a 5% defective rate where each transistor is produced independent of the others.
 On average how many transistors would you expect to be produced with this machine before the first with a defect?
-What is the standard deviation?
+**Using the same logice I would expect 20**  What is the standard deviation?
+**Apx 0.22**
+
+```{r}
+sigma <- sqrt(0.05*(1 - 0.05))
+sigma
+```
+
 (e) Based on your answers to parts (c) and (d), how does increasing the probability of an event affect the mean and standard deviation of the wait time until success?
+**It decreases the mean and increases the standard deviation.**
 
@@
 While it is often assumed that the probabilities of having a boy or a girl are the same, the actual probability of having a boy is slightly higher at 0.51. Suppose a couple plans to have 3 kids.
-(a) Use the binomial model to calculate the probability that two of them will be boys.
+
+(a) Use the binomial model to calculate the probability that two of them will be boys. **0.38**
+
+```{r}
+k <- 2
+n <- 3
+p <- 0.51
+((factorial(n)/(factorial(k)*factorial(n-k))) * p^k)*(1-p)^(n-k)
+```
+
 (b) Write out all possible orderings of 3 children, 2 of whom are boys.
@@
 Confirm that your answers from parts (a) and (b) match.
+**The combinations that will produce this kind of outcome are: girl-boy-boy, boy-girl-boy, boy-boy-girl**
+
+```{r}
+p_gbb <- (1 - p) * p * p
+p_bgb <- p * (1 - p) * p
+p_bbg <- p * p * (1 - p)
+
+p_gbb + p_gbb + p_bbg
+```
+
 (c) If we wanted to calculate the probability that a couple who plans to have 8 kids will have 3 boys, briefly describe why the approach from part (b) would be more tedious than the approach from part (a).
+**There would be more combinations that would need to have probabilities computed which would be more tedious.**
 
-## 3.42 2 Serving in Volleyball
+## 3.42 Serving in Volleyball
 
@@
 
-(a) What is the probability that on the 10th try she will make her 3rd successful serve?
-(b) Suppose she has made two successful serves in nine attempts. What is the probability that her 10th serve will be successful?
-(c) Even though parts (a) and (b) discuss the same scenario, the probabilities you calculated should be different. Can you explain the reason for this discrepancy?
+(a) What is the probability that on the 10th try she will make her 3rd successful serve?  **The chances that her 10th being her 3rd success is roughly 0.04.**
+
+```{r}
+# Compute the chances of having 2 successful serve out of 9
+k <- 2
+n <- 9
+p <- 0.15
+p_2_out_of_9 <- ((factorial(n)/(factorial(k)*factorial(n-k))) * p^k)*(1-p)^(n-k)
+# Now factor in the chance of her 10th being a success
+p_2_out_of_9 * p
+```
+
+(b) Suppose she has made two successful serves in nine attempts. What is the probability that her 10th serve will be successful?  **15% because each serve is independent of the others.**
+
+(c) Even though parts (a) and (b) discuss the same scenario, the probabilities you calculated should be different.
+Can you explain the reason for this discrepancy?
+**The probability in (b) is for a single event, while the probability in (a) is for a series of events.
+That's why they are different.**

### 2018-10-02 DATA607/Project_2.Rmd
Initial commit of DATA 607 Project 2

---
title: "DATA 607 Project 2"
author: "Mike Silva"
date: "October 1, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float: true
---

## Description

This project involves transforming three untidy datasets into tidy datasets using dplyr and tidyr.  Once the data is transformed a brief exploratory data analysis will be preformed.  Since we are using dplyr and tidyr we will load them now (all other libraries will be loaded as needed).

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
```

## 2017 United Nations International Migrant Stock

The first dataset that will undergo the cleanup is the 2017 International Migrant Stock data produced by the United Nations Department of Economic and Social Affairs Population Division.  They produce multiple datasets.  This project will use the immigrants by origin and destination.

```{r echo=FALSE}
# [http://www.un.org/en/development/desa/population/migration/data/estimates2/docs/MigrationStockDocumentation_2017.pdf](http://www.un.org/en/development/desa/population/migration/data/estimates2/docs/MigrationStockDocumentation_2017.pdf)
# 
# This dataset provides estimates of the international migrant stock by age, sex and origin for the mid-point (1 July) of each year: 1990, 1995, 2000,
# 2005, 2010, 2015 and 2017. 
# 
# Most of the data used to estimate the international migrant stock by country or area were obtained from population censuses. Additionally, population registers and nationally representative surveys provided information on the number and composition of international migrants.
# 
# In estimating the international migrant stock, international migrants have been equated with the foreignborn population whenever this information is available, which is the case in most countries or areas. In most countries lacking data on place of birth, information on the country of citizenship of those enumerated was available, and was used as the basis for the identification of international migrants, thus effectively equating, in these cases, international migrants with foreign citizens.
# 
# Equating international migrants with foreign citizens when estimating the migrant stock has important shortcomings. In countries where citizenship is conferred on the basis of jus sanguinis, people who were born in the country of residence may be included in the number of international migrants even though they may have never lived abroad. Conversely, persons who were born abroad and who naturalized in their 
```

### Data Acquisition

We will download the migration stock by destination and origin data to our local system.

```{r}
file_name <- "data/UN_MigrantStockByOriginAndDestination_2017.xlsx"
if(!file.exists(file_name)){
  download.file("http://www.un.org/en/development/desa/population/migration/data/estimates2/data/UN_MigrantStockByOriginAndDestination_2017.xlsx", file_name)
}
```

### Parsing Untidy Data

I will begin by reading in a few rows and examing them to determine what actions need to be taken.

```{r, results='asis'}
library(openxlsx)
library(kableExtra)
read.xlsx(file_name, sheet = "Table 1", rows = 1:25) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
  scroll_box(width = "100%", height = "200px")
```

#### Developing a Workplan

There is a lot of data at the begining of the file that needs to be overlooked.  When I read in the data I will skip down to the row after the one that begins with "year".  The country names are in that row.  Thre are ".." that will need to be changed to NAs.  The data will need to be renamed. I won't need the sort order, code or type of data columns so I will drop them.

#### Execution of the Plan

```{r}
un_df <- read.xlsx('data/UN_MigrantStockByOriginAndDestination_2017.xlsx', 'Table 1', startRow = 16, na.strings = "..") %>%
  rename(Year = X1,
         Sort.order = X2,
         Destination = X3,
         Notes = X4,
         Code = X5,
         Type.of.data = X6) %>%
  select(-Sort.order, -Code, -Type.of.data)
```

Now that these steps have been taken let's take a look at the data frame to determine what to do next.

```{r, echo=FALSE}
un_df[1:10,] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
  scroll_box(width = "100%", height = "200px")
```

The Notes column is not NA on the rows I want to keep.  There are regional aggregates that I don't want to include too.  These aggregates have letters in the Notes column.  If I try to coorce them into numeric data they will turn into NAs and I can drop all the rows in one fell swoop.  I will then drop the Notes, Total, Other.North and Other.South columns.

```{r, warning=FALSE}
un_df <- un_df %>%
  mutate(Notes = as.numeric(Notes)) %>% # Introduces NAs where desired
  drop_na(Notes) %>%
  select(-Notes:-Other.South)
```

Now we are ready to change the data from wide to long.

```{r}
un_df <- un_df %>%
  gather(key = Origin, value = Migrants, -Year, -Destination) %>%
  drop_na(Migrants)
```

Now I need to clean up the Origins columns.  There are periods that represent spaces.  When there are two periods it needs to be changed into a period and space.  An example of that is Micronesia.(Fed..States.of).

```{r}
un_df <- un_df %>%
  mutate(Origin = str_replace_all(Origin, "[.]", " ")) %>%
  mutate(Origin = str_replace_all(Origin, "  ", ". "))
```

And here is the results:

```{r, echo=FALSE}
un_df[1:10,] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
  scroll_box(width = "100%", height = "200px")
```

Text

### 2018-10-02 DATA607/Project_2.Rmd
Additional Work

@@
     toc: true
-    toc_depth: 3
+    toc_depth: 4
     toc_float: true
@@
 
-## 2017 United Nations International Migrant Stock
+## United Nations International Migrant Stock
 
@@
 
+### Research Questions
+
+Here are the questions I want to answer with this data:
+
+* What are the top-10 countries receiving/loosing migrants in absolute numbers by year?
+
+* Do communities exist in this network?  Using the number of migrants as the edge between the countries (nodes) in the network graph, do communities emerge from the data?
+
 ```{r echo=FALSE}
+#### About the Data
+
 # [http://www.un.org/en/development/desa/population/migration/data/estimates2/docs/MigrationStockDocumentation_2017.pdf](http://www.un.org/en/development/desa/population/migration/data/estimates2/docs/MigrationStockDocumentation_2017.pdf)
@@
 
-### Parsing Untidy Data
+### Creating Tidy Data
 
@@
 An example of that is Micronesia.(Fed..States.of).
+I will also reorder the columns.
 
@@
   mutate(Origin = str_replace_all(Origin, "[.]", " ")) %>%
-  mutate(Origin = str_replace_all(Origin, "  ", ". "))
+  mutate(Origin = str_replace_all(Origin, "  ", ". ")) %>%
+  select(Origin, Destination, Migrants, Year)
 ```
@@
 
-Text
+### Analysis
+
+#### Top 10 Receivers of International Migrants
+
+What countries are the top ten in the number of migrants they are receiving by year?

### 2018-10-05 DATA606/Silva-Proposal.Rmd
Initial commit of DATA 606 Proposal

---
title: "DATA 606 Data Project Proposal"
author: "Mike Silva"
date: "October 3, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 4
    toc_float: true
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data

```

### Research question

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

### Cases

**What are the cases, and how many are there?**

### Data collection

**Describe the method of data collection.**

### Type of study

**What type of study is this (observational/experiment)?**

### Data Source

**If you collected the data, state self-collected. If not, provide a citation/link.**

### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**

### Relevant summary statistics

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}

```

### 2018-10-07 DATA607/Project_2.Rmd
More work on Project 2

@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r initialize, message=FALSE, warning=FALSE}
 library(dplyr)
@@
 library(stringr)
+library(kableExtra)
+
+#output: 
+#  rmdformats::readthedown:
+#    highlight: kate
 ```
 
-## United Nations International Migrant Stock
+## American Community Survey
 
-The first dataset that will undergo the cleanup is the 2017 International Migrant Stock data produced by the United Nations Department of Economic and Social Affairs Population Division.
+The American Community Survey (henceforth ACS), is a product of the U.S.
+Census Bureau and measures the current characteristics of the American population.
+It is the replacement for the Census long form.
+In this analysis I will be pulling data for all counties in the United States.
+The two tables contain the count of workers by occupation and gender and their median income.
+
+### Research Questions
+
+Here are the questions I want to answer with this data:
+
+* What occupation has the the greatest pay gap between genders?  And which occupations have the most pay equity?
+
+* How does the pay gap relate to concentration?
+
+#### About the Data
+
+The data we want to use for this is county level data.  We want the most robust numbers so we will be using the latest 5 year estimates available.
+
+### Data Acquisition
+
+In order to pull the data I will be using the ACS R package.
+I have previously [registered for an API key](https://api.census.gov/), and have use the *api.key.install()* command to save it to my environment.
+
+```{r get_acs_data, message=FALSE}
+# C24010 SEX BY OCCUPATION FOR THE CIVILIAN EMPLOYED POPULATION 16 YEARS AND OVER
+file_name <- "data/C24010_2016_5_est.csv"
+if(!file.exists(file_name)){
+  geos <- acs::geo.make(county = "*", state = "*")
+  acs <- acs::acs.fetch(2016, geography = geos, table.number = "C24010", col.names = "pretty")
+  acs_df_C24010 <- data.frame(acs::estimate(acs))
+  write.csv(acs_df_C24010, file = file_name)
+} else {
+  acs_df_C24010 <- read.csv(file_name)
+}
+
+# B24022 SEX BY OCCUPATION AND MEDIAN EARNINGS IN THE PAST 12 MONTHS (IN 2016 INFLATION-ADJUSTED DOLLARS) FOR THE FULL-TIME, YEAR-ROUND CIVILIAN EMPLOYED POPULATION 16 YEARS AND OVER
+file_name <- "data/B24022_2016_5_est.csv"
+if(!file.exists(file_name)){
+  geos <- acs::geo.make(county = "*", state = "*")
+  acs <- acs::acs.fetch(2016, geography = geos, table.number = "B24022", col.names = "pretty")
+  acs_df_B24022 <- data.frame(acs::estimate(acs))
+  write.csv(acs_df_B24022, file = file_name)
+} else {
+  acs_df_B24022 <- read.csv(file_name)
+}
+```
+
+### Creating Tidy Data
+
+The ACS data is broken out by gender and occupation.  The rows are counties.  It is an extremly wide dataset with `r ncol(acs_df_C24010)` columns.
+
+##### Table 1. C24010 Sex by Occupation for Civilain Employed Pop 16+
+```{r untidy_acs_table, echo=FALSE}
+acs_df_C24010[1:10,] %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
+  scroll_box(width = "100%", height = "200px")
+```
+
+#### Developing a Workplan
+
+In order to make a tidy data frame the first step will be to transform it from wide to long form.
+Then the gender and occupation needs to be split out from the column header.
+Spliting the gender will be simple enough but extracting the occupations will require a bit of cleanup.
+I will use the following function to strip out the non-occupational parts of the string, then replace the periods added to the columns by R with spaces, remove the duplicate spaces and strip off leading and trailing white space.
+
+```{r get_occupation}
+get_occupation <- function(string){
+  string <- str_remove(string, "SEX.BY.OCCUPATION.FOR.THE.CIVILIAN.EMPLOYED.POPULATION.16.YEARS.AND.OVER..")
+  string <- str_remove(string, "B24022..Sex.by.Occupation.and.Median.Earnings.in.the.Past.12.Months..in.2015.Inflation.Adjusted.Dollars..for.the.Full.Time..Year.Round.Civilian.Employed.Population.16.Years.and.Over")
+  string <- str_remove(string, "Female")
+  string <- str_remove(string, "Male")
+  string <- str_replace_all(string, "[.]", " ")
+  string <- str_replace_all(string, "  ", " ")
+  string <- trimws(string)
+  # Simplify occupations
+  # Select everything after the last capital letter started sequence
+  search_for <- str_extract(string, "[A-Z][[:alpha:] ]+[A-Z]")
+  replace_with <- str_sub(search_for, start=-1)
+  string <- str_replace_all(string, search_for, replace_with)
+  string
+}
+```
+
+Now that I have the get_occupation function I am ready to apply it to the dataset.
+
+#### Execution of the Plan
+
+I will start with the worker counts by sex and occupation.
+
+```{r tidy_C24010}
+acs_df_C24010 <- acs_df_C24010 %>%
+  rename(County = X) %>%
+  gather(key = Header, value = Workers, -County) %>%
+  mutate(Occupation = get_occupation(Header),
+         Gender = ifelse(grepl("Female", Header), "Female", ifelse(grepl("Male", Header), "Male", "Total"))) %>% 
+  filter(Gender != "Total") %>%
+  select(-Header) %>%
+  drop_na()
+```
+
+Thre are `r nrow(acs_df_C24010)` rows in acs_df_C24010.  Now I will apply the same prcess to the median earnings data.
+
+```{r tidy_B24022}
+acs_df_B24022 <- acs_df_B24022 %>%
+  rename(County = X) %>%
+  gather(key = Header, value = "Earnings", -County) %>%
+  mutate(Occupation = get_occupation(Header),
+         Gender = ifelse(grepl("Female", Header), "Female", ifelse(grepl("Male", Header), "Male", "Total"))) %>%
+  filter(Gender != "Total") %>%
+  select(-Header)
+```
+
+Thre are `r nrow(acs_df_B24022)` rows in acs_df_B24022.  The final step in assembling the tidy dataset is to merge the two data frames togther.
+
+```{r merge_acs}
+acs_df <- merge(acs_df_C24010, acs_df_B24022)
+```
+
+There are `r nrow(acs_df)` rows in the combined data frame.  Since I want to look at the gender concentrations withing the occupation I will need to create that varaible.
+
+```{r add_gender_concentrations}
+acs_df <- acs_df %>%
+  select(Occupation, County, Workers) %>%
+  group_by(Occupation, County) %>%
+  summarise(Total = sum(Workers)) %>% 
+  merge(acs_df) %>%
+  mutate(Concentration = Workers / Total) %>%
+  select(-Total)
+```
+
+Let's do a quick sanity check on the data.  Let's examine the Workers, Earnigns and Concentration
+
+##### Table 2. ACS Data Check
+```{r acs_sanity_check_1, echo=FALSE}
+acs_df %>%
+  select(Workers, Earnings, Concentration) %>%
+  summarise_all(funs(min=min, median=median, mean = mean, max=max)) %>%
+  gather(stat, val) %>%
+  separate(stat, into = c("Variable", "stat"), "_") %>%
+  spread(stat, val) %>%
+  select(Variable, min, mean, median, max) %>%
+  kable() %>%
+  kable_styling("striped")
+```
+
+I see there are cases where the median earnings are -666666666.  This is the Census Bureau's equivalent of N/A, so we will remove those observations.
+
+```{r remove_no_earnings}
+acs_df <- acs_df %>%
+  filter(Earnings > 0)
+```
+
+##### Table 3. ACS Data After Cleanup
+```{r acs_sanity_check_2, echo=FALSE}
+acs_df %>%
+  select(Workers, Earnings, Concentration) %>%
+  summarise_all(funs(min=min, median=median, mean = mean, max=max)) %>%
+  gather(stat, val) %>%
+  separate(stat, into = c("Variable", "stat"), "_") %>%
+  spread(stat, val) %>%
+  select(Variable, min, mean, median, max) %>%
+  kable() %>%
+  kable_styling("striped")
+```
+
+The median earnings are reasonable and the concentration rates range from roughly 0% to 100%.  Now I need to compute a female to male earning ratio.
+
+```{r add_earning_ratio}
+acs_df <- acs_df %>%
+  filter(Gender == "Male") %>%
+  select(County, Occupation, Earnings) %>%
+  rename(Male_Earnings = Earnings) %>%
+  merge(acs_df) %>%
+  filter(Gender == "Female") %>%
+  mutate(Ratio = Earnings / Male_Earnings)
+```
+
+We now have `r nrow(acs_df)` observations ready for analysis.
+
+### Analysis
+
+#### Occupation with the the Largest and Smallest Pay Gap
+
+The female workers earn `r round(min(acs_df$Ratio), 2)` to `r round(max(acs_df$Ratio), 2)` dollars for every dollar a male worker earns.
+That's a considerable range!
+The median is `r round(median(acs_df$Ratio), 2)` as shown by the verticle line in the graph below.
+
+```{r acs_viz, echo=FALSE}
+library(ggplot2)
+ratio_line <- median(acs_df$Ratio)
+
+ggplot(acs_df, aes(Ratio)) +
+  geom_density() +
+  geom_vline(aes(xintercept = ratio_line), color="red", linetype="dashed")
+```
+
+So which occupations are the most inequitable in pay?  There may be some highly unusual markest so let's average the ratios by occupation and look at the results.
+
+```{r acs_average_by_occupation}
+acs_avg_df <- acs_df %>%
+  group_by(Occupation) %>%
+  summarise(Ratio = mean(Ratio), Concentration = mean(Concentration), N = n())
+```
+
+##### Table 4. ACS Average Summary Statistics
+```{r acs_average_by_occupation_summary, echo=FALSE}
+acs_avg_df %>%
+  select(Ratio, Concentration, N) %>%
+  summarise_all(funs(min=min, median=median, mean = mean, max=max)) %>%
+  gather(stat, val) %>%
+  separate(stat, into = c("Variable", "stat"), "_") %>%
+  spread(stat, val) %>%
+  select(Variable, min, mean, median, max) %>%
+  kable() %>%
+  kable_styling("striped")
+```
+
+There is no occupation group where women on average earn more than men.  Let's examine which occupations women earn less than their male counterparts.
+
+##### Table 5. Top 5 Occupations with Income Inequality
+```{r acs_highest_inequity, echo=FALSE}
+acs_avg_df %>%
+  arrange(Ratio) %>%
+  top_n(5, -Ratio) %>%
+  kable() %>%
+  kable_styling("striped")
+```
+
+Legal, sales, and health diagnosing/treatment/technical occupations seem to be the top occpuations with a gender pay gap.
+It is interesting to note that most of these occupations have a higher than avergae female conentration.
+Now let's see which occupations are closer to equal pay.
+
+##### Table 6. Top 5 Occupations with Lowest Income Inequality
+```{r acs_lowest_inequity, echo=FALSE}
+acs_avg_df %>%
+  arrange(-Ratio) %>%
+  top_n(5, Ratio) %>%
+  kable() %>%
+  kable_styling("striped")
+```
+
+It looks like community/social service, food preperation and serving, installation/repair, computer and science related occupations have the closest pay equality.
+The concentration levels are a mixed bag.
+
+#### Pay Gap and Concentration Relationship
+
+Is there any relationship between the equality of gender representation and pay?
+Our previous example looked at aggregate trends and it appeared to not have a strong relationship.
+Let's examine all the observations.
+
+```{r acs_scatter_plot, echo=FALSE}
+ggplot(acs_df, aes(Concentration, Ratio, alpha=0.0001)) + geom_point() + guides(alpha = "none")
+```
+
+Looking at the scatterplot it doesn't appear to be much of a relationship.  Let's look at the Pearson's R.
+
+```{r acs_correlation}
+cor(acs_df$Concentration, acs_df$Ratio)
+```
+
+This confirms that the relationship is not based on the level of concentration in the industry.
+Which suggests that if someone wants to improve the gender pay gap and thinks having more females representation in the occupations will solve that problem, they might be disappointed.
+
+## U.N. International Migrant Stock
+
+The next dataset that will undergo the cleanup is the 2017 International Migrant Stock data produced by the United Nations Department of Economic and Social Affairs Population Division.
 They produce multiple datasets.
@@
 
-* What are the top-10 countries receiving/loosing migrants in absolute numbers by year?
+* What are the top countries receiving/loosing migrants in absolute numbers by year?
 
-* Do communities exist in this network?  Using the number of migrants as the edge between the countries (nodes) in the network graph, do communities emerge from the data?
+* How does the United States compare?
 
-```{r echo=FALSE}
+* Do communities exist in this network?
+
 #### About the Data
 
-# [http://www.un.org/en/development/desa/population/migration/data/estimates2/docs/MigrationStockDocumentation_2017.pdf](http://www.un.org/en/development/desa/population/migration/data/estimates2/docs/MigrationStockDocumentation_2017.pdf)
-# 
-# This dataset provides estimates of the international migrant stock by age, sex and origin for the mid-point (1 July) of each year: 1990, 1995, 2000,
-# 2005, 2010, 2015 and 2017. 
-# 
-# Most of the data used to estimate the international migrant stock by country or area were obtained from population censuses.
-Additionally, population registers and nationally representative surveys provided information on the number and composition of international migrants.
-# 
-# In estimating the international migrant stock, international migrants have been equated with the foreignborn population whenever this information is available, which is the case in most countries or areas.
-In most countries lacking data on place of birth, information on the country of citizenship of those enumerated was available, and was used as the basis for the identification of international migrants, thus effectively equating, in these cases, international migrants with foreign citizens.
-# 
-# Equating international migrants with foreign citizens when estimating the migrant stock has important shortcomings.
-In countries where citizenship is conferred on the basis of jus sanguinis, people who were born in the country of residence may be included in the number of international migrants even though they may have never lived abroad.
-Conversely, persons who were born abroad and who naturalized in their
-```
+These estimates of the international migrants have been equated with the foreignborn population whenever this information is available, which is the case in most countries or areas.
+It is for the mid-point of each year.
 
@@
 
-```{r}
+```{r get_un_data}
 file_name <- "data/UN_MigrantStockByOriginAndDestination_2017.xlsx"
@@
 
-```{r, results='asis'}
+##### Table 7. First Look at UN Data
+```{r untidy_un_table, results='asis', echo=FALSE}
 library(openxlsx)
-library(kableExtra)
 read.xlsx(file_name, sheet = "Table 1", rows = 1:25) %>%
@@
 
-```{r}
+```{r read_un}
 un_df <- read.xlsx('data/UN_MigrantStockByOriginAndDestination_2017.xlsx', 'Table 1', startRow = 16, na.strings = "..") %>%
@@
          Type.of.data = X6) %>%
-  select(-Sort.order, -Code, -Type.of.data)
+  select(-Sort.order, -Code, -Notes)
 ```
@@
 
-```{r, echo=FALSE}
+##### Table 8. Untidy UN Data
+```{r untidy_un_table_2, echo=FALSE}
 un_df[1:10,] %>%
@@
 
-The Notes column is not NA on the rows I want to keep.
-There are regional aggregates that I don't want to include too.
-These aggregates have letters in the Notes column.
-If I try to coorce them into numeric data they will turn into NAs and I can drop all the rows in one fell swoop.
+There are regional aggregates that I don't want to include  These aggregates dont have a type of data column.
+I will use that to remove them.
 I will then drop the Notes, Total, Other.North and Other.South columns.
 
-```{r, warning=FALSE}
+```{r tidy_un, warning=FALSE}
 un_df <- un_df %>%
-  mutate(Notes = as.numeric(Notes)) %>% # Introduces NAs where desired
-  drop_na(Notes) %>%
-  select(-Notes:-Other.South)
+  filter(str_length(Type.of.data)>0) %>%
+  select(-Type.of.data:-Other.South)
 ```
@@
 
-```{r}
+```{r un_wide_to_long}
 un_df <- un_df %>%
@@
 
-```{r}
+```{r un_remove_periods}
 un_df <- un_df %>%
@@
   mutate(Origin = str_replace_all(Origin, "  ", ". ")) %>%
+  mutate(Migrants = as.numeric(Migrants)) %>%
+  mutate(Origin = as.vector(Origin)) %>%
+  mutate(Destination = as.vector(Destination)) %>%
   select(Origin, Destination, Migrants, Year)
@@
 
-```{r, echo=FALSE}
+```{r tidy_un_table, echo=FALSE}
 un_df[1:10,] %>%
@@
 
-#### Top 10 Receivers of International Migrants
+#### Top 5 Receivers of International Migrants
 
-What countries are the top ten in the number of migrants they are receiving by year?
+What countries are the top receiverd of migrants they are receiving by year?  Does the contries on this list change over time?  Note: These data are not adjusted for population.
+
+```{r up_top_receivers, message=FALSE}
+top_destinations <- un_df %>%
+  select(-Origin) %>%
+  group_by(Year, Destination) %>%
+  summarise(Migrants = sum(Migrants)) %>%
+  rename(Country = Destination) %>%
+  arrange(Year, -Migrants) %>%
+  top_n(5) %>%
+  mutate(Rank = 1:n())
+```
+
+##### Table 9. Top 5 Migrant Destinations by Year
+```{r top_recipient_table, echo=FALSE}
+top_destinations %>%
+  select(-Migrants) %>%
+  spread(Year, Country) %>%
+  arrange(Rank) %>%
+  kable() %>%
+  kable_styling("striped")
+```
+
+The United States is consitently at the top of the list.  I guess we continue to  be a nation of immigrants.  Russia, Germany and the United Kingdom look like other hot spots for immigrants.
+
+#### The United States
+
+I was originally going to look at how the United States stacks up in the ranking but we know the answer to that from the last analysis.
+I will look deeper into the United States.
+What countries do most of the immgigrants come from?
+
+```{r un_rankings, message=FALSE}
+rankings_df <- un_df %>%
+  filter(Destination == "United States of America") %>%
+  group_by(Year, Origin) %>%
+  summarise(Migrants = sum(Migrants)) %>%
+  rename(Country = Origin) %>%
+  arrange(Year, -Migrants) %>%
+  top_n(10) %>%
+  mutate(Rank = 1:n())
+```
+
+##### Table 10. Countries the United States Immigrants Come From
+```{r us_table, echo=FALSE}
+rankings_df %>%
+  select(-Migrants) %>%
+  spread(Year, Country) %>%
+  arrange(Rank) %>%
+  kable() %>%
+  kable_styling("striped")
+```
+
+Mexico seems to consistently be the highest source of immigrants to the United States.
+China and India have recently increased in popularity.
+Puerto Rico is often high on the list.
+The mix of top countries is pretty consistent year over year.
+
+#### Network Communities
+
+```{r, message=FALSE}
+library(igraph)
+un_graph_2017 <- un_df %>%
+  filter(Year == 2017) %>%
+  select(Origin, Destination) %>%
+  graph_from_data_frame()
+
+communities <- cluster_walktrap(un_graph_2017)
+
+communities
+```
+
+There are 10 coomunities that emerge.  I would like to visulize it but that would take too much time. I have one more dataset to look at.
+
+```{r, eval=FALSE, echo=FALSE}
+layout <-layout.fruchterman.reingold(un_graph_2017)
+#layout <- layout.davidson.harel(un_graph_2017)
+# Change colors of nodes, polygons, and polygon borders
+new_cols <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6", "#6a3d9a")[membership(communities)]
+plot(communities, un_graph_2017, col=new_cols, mark.border="white", mark.col=c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6", "#6a3d9a"), 
+    layout=layout, vertex.label=NA, vertex.size=5, edge.arrow.size=.2, edge.color = "gray")
+```
+
+## NYS Adult Arrests by County
+
+This is the datase featured in my discussion board post.  It is produced by the NYS Division of Criminal Justics Services and is the count of adult arrests.
+
+### Research Questions
+
+Here are the questions I want to answer with this data:
+
+* What is the general trend in total arrests?
+
+#### About the Data
+
+The counts of arrests are for fingerprintable offenses.
+An adult arrest is defined as an arrest of a person 16 years old or older or a juvenile offender prosecuted in adult court.
+Fingerprintable offenses (defined in Criminal Procedure Law §160.10) include any felony, a misdemeanor defined in the penal law, a misdemeanor defined outside the penal law which would constitute a felony if such a person had a previous judgment of conviction for a crime, or loitering for the purpose of engaging in prostitution as defined in subdivision two of Penal Law §240.37.
+
+### Data Acquisition
+
+The dataset from the New York State Division of Criminal Justice Services was accessed via [data.ny.gov](https://data.ny.gov/Public-Safety/Adult-Arrests-by-County-Beginning-1970/rikd-mt35).
+It was last updated on August 7, 2018 at the time of this pull.
+
+```{r}
+file_name <- "data/NYS_Adult_Arrests_by_County.csv"
+if(!file.exists(file_name)){
+  download.file("https://data.ny.gov/api/views/rikd-mt35/rows.csv?accessType=DOWNLOAD", file_name)
+}
+
+arrest_df <- read.csv(file_name)
+```
+
+### Creating Tidy Data
+
+I will begin by reading in a few rows and examing them to determine what actions need to be taken.
+
+##### Table 11. NYS Adult Arrests
+```{r untidy_arrest_table, echo=FALSE}
+arrest_df[1:25,] %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
+  scroll_box(width = "100%", height = "200px")
+```
+
+#### Developing a Workplan
+
+This is a wide data set that needs to be transformed to a long dataset.  The column headers will need to be cleaned up after the transformation.
+
+#### Execution of the Plan
+
+```{r}
+arrest_df <- arrest_df %>%
+  gather(Type, Arrests, -County, -Year) %>%
+  mutate(Type = gsub("[.]", " ", Type))
+```
+
+This is realitively straight forward and is ready for an analysis.
+
+##### Table 12. Tidy NYS Adult Arrests
+```{r untidy_arrest_table2, echo=FALSE}
+arrest_df[1:25,] %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
+  scroll_box(width = "100%", height = "200px")
+```
+
+### Analysis
+
+#### General Trend in Arrests
+
+Crime rates have been declining generally across the country.  Are the arrests rates decreasing in NYS?
+
+```{r, echo=FALSE}
+arrest_df %>%
+  filter(Type == "Total") %>%
+  ggplot(aes(Year, Arrests, group=County)) + geom_line()
+```
+
+It's difficult to tell.
+The difference in scale is making it really difficult.
+If I were continuing the analysis I would express the data on a per capita basis.
+I can do it as a percentage change from 1970
+
+For both of these questions I want to look at the change in percentage terms.  I will quickly add in the percentage change to the dataset.  I will base all the changes from the earliest data, 1970.
+
+```{r, percentage_change}
+arrest_df <- arrest_df %>%
+  filter(Year == 1970) %>%
+  select( -Year) %>%
+  rename(Arrests_Base = Arrests) %>%
+  merge(arrest_df) %>%
+  mutate(Change = (Arrests / Arrests_Base)-1)
+```
+
+```{r, echo=FALSE}
+arrest_df %>%
+  filter(Type == "Total") %>%
+  ggplot(aes(Year, Change, group=County)) + geom_line() + geom_hline(aes(yintercept=0, color="red")) + guides(color=FALSE)
+```
+
+It appears that the number of arrests are up since 1970 which is unexpected.
+
+```{r}
+arrest_df %>%
+  group_by(Year) %>%
+  filter(Type == "Total") %>%
+  select(Arrests, Change) %>%
+  summarise_all(funs(min=min, median=median, mean = mean, max=max)) %>%
+  gather(stat, val, -Year)  %>%
+  separate(stat, into = c("Variable", "stat"), "_") %>%
+  spread(stat, val) %>%
+  select(Variable, Year, min, mean, median, max) %>%
+  arrange(Variable, Year) %>%
+  kable() %>%
+  kable_styling("striped")
+
+```
+
+It appears that there are some oddities with this dataset that would need to be better understood before carying out further analyses.

### 2018-10-07 DATA607/Project_2.Rmd
More work on Project 2

@@
 output: 
-  html_document:
-    theme: yeti
-    highlight: tango
-    toc: true
-    toc_depth: 4
-    toc_float: true
+  rmdformats::readthedown:
+    highlight: kate
 ---
@@
 library(stringr)
+library(ggplot2)
 library(kableExtra)
-
-#output: 
-#  rmdformats::readthedown:
-#    highlight: kate
 ```
@@
   kable() %>%
-  kable_styling("striped")
+  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
 ```
@@
   kable() %>%
-  kable_styling("striped")
+  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
+  scroll_box(width = "100%", height = "200px")
 ```
@@
 
+##### Figure 1. Earning Ratio Density Plot (Median Verticle Line in Red)
 ```{r acs_viz, echo=FALSE}
-library(ggplot2)
 ratio_line <- median(acs_df$Ratio)
@@
 
+##### Figure 2. Concentration and Earning Ratio Scatter Plot
 ```{r acs_scatter_plot, echo=FALSE}
@@
 
+##### Figure 3. Total Adult Arrests by County and Year
 ```{r, echo=FALSE}
@@
 
+##### Figure 4. Percentage Change from 1970 Level (No Change in Red)
 ```{r, echo=FALSE}
@@
   kable() %>%
-  kable_styling("striped")
+  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>%
+  scroll_box(width = "100%", height = "200px")
 

### 2018-10-10 DATA606/Silva-Proposal.Rmd
Work on DATA 606 Research Project Proposal

@@
 
-### Data Preparation
+## Research Question
 
-```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
-# load data
+In a [2017 study done by Pew](http://www.pewresearch.org/fact-tank/2018/03/07/women-in-majority-male-workplaces-report-higher-rates-of-gender-discrimination/) found that women in majority-male wokplaces are more likely to say their gender has made it harder for them to get ahead at work.
 
+Does this trend hold at the macro level?  Otherwise put does the level of concentration within an occupation predict the differential in income earned between men and women?
+
+The response variable will be the differential.  I think I will use a ratio of female to male earnings.  In any case it will be numeric.
+
+The explanatory variable will be the concentration level.
+This will also be numeric.
+I will use the occupation as a control variable.
+It will be categorical.
+I may also want to use a geographic variable which would also be categorical.
+
+## Data Source
+
+U.S. Census Bureau; American Community Survey, 2012-2016, Detailed Tables; generated by Mike Silva; using Census Bureau API; (9 October 2018)
+
+## About the Data
+
+The American Communtiy Survey (ACS) is the Census Bureau's replacement to the long form and collects information on the population, housing and workforce.
+
+### Data Collection
+
+Each year the U.S.
+Census Bureau contacts over 3.5 million households to participate in the ACS.
+It is collected through mail, internet, telephone and in-person interviews.
+Households are randomly selected and required by law to respond to all questions as accurately as possible.
+The survey data is then weighted and estimates are produced.
+
+The tables that will be used in this study are the latest 5 year estimates which incorporates the responses from 2012-2016.
+I will be analyzing all county level estimates.
+This level of geography is being chosen because the estimates will be more reliable than the city level data (due to number of survey responses in the estimates) while giving more observations than the state or metro levels.
+The estimates are based on observational data.
+
+### Data Acquisition
+
+In order to pull the data I will be using the ACS R package.
+I have previously [registered for an API key](https://api.census.gov/), and have use the *api.key.install()* command to save it to my environment.
+I will pull the data via the API and save it locally to save time in the future.
+
+```{r get_acs_data, message=FALSE}
+library(censusapi)
+library(dplyr)
+library(tidyr)
+library(stringr)
+
+# C24010 SEX BY OCCUPATION FOR THE CIVILIAN EMPLOYED POPULATION 16 YEARS AND OVER
+file_name <- "data/C24010_2016_5_est.csv"
+if(!file.exists(file_name)){
+  geos <- acs::geo.make(county = "*", state = "*")
+  acs <- acs::acs.fetch(2016, geography = geos, table.number = "C24010", col.names = "pretty")
+  acs_df_C24010 <- data.frame(acs::estimate(acs))
+  write.csv(acs_df_C24010, file = file_name)
+} else {
+  acs_df_C24010 <- read.csv(file_name)
+}
+
+# B24022 SEX BY OCCUPATION AND MEDIAN EARNINGS IN THE PAST 12 MONTHS (IN 2016 INFLATION-ADJUSTED DOLLARS) FOR THE FULL-TIME, YEAR-ROUND CIVILIAN EMPLOYED POPULATION 16 YEARS AND OVER
+file_name <- "data/B24022_2016_5_est.csv"
+if(!file.exists(file_name)){
+  geos <- acs::geo.make(county = "*", state = "*")
+  acs <- acs::acs.fetch(2016, geography = geos, table.number = "B24022", col.names = "pretty")
+  acs_df_B24022 <- data.frame(acs::estimate(acs))
+  write.csv(acs_df_B24022, file = file_name)
+} else {
+  acs_df_B24022 <- read.csv(file_name)
+}
 ```
 
-### Research question
+The following function will be used to clean up the occupations.
 
-**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**
+```{r get_occupation}
+get_occupation <- function(string){
+  string <- str_remove(string, "SEX.BY.OCCUPATION.FOR.THE.CIVILIAN.EMPLOYED.POPULATION.16.YEARS.AND.OVER..")
+  string <- str_remove(string, "B24022..Sex.by.Occupation.and.Median.Earnings.in.the.Past.12.Months..in.2015.Inflation.Adjusted.Dollars..for.the.Full.Time..Year.Round.Civilian.Employed.Population.16.Years.and.Over")
+  string <- str_remove(string, "Female")
+  string <- str_remove(string, "Male")
+  string <- str_replace_all(string, "[.]", " ")
+  string <- str_replace_all(string, "  ", " ")
+  string <- trimws(string)
+  # Simplify occupations - Select everything after the last capital letter started sequence
+  search_for <- str_extract(string, "[A-Z][[:alpha:] ]+[A-Z]")
+  replace_with <- str_sub(search_for, start=-1)
+  string <- str_replace_all(string, search_for, replace_with)
+  string
+}
+```
+
+Now we will build a tidy dataset for analysis by reading in the two datasets, cleaning the occupational groups, and merging the two together.
+The Census Bureau uses -666666666 for NAs in the earnings table (B24022).
+
+```{r build_tidy_data}
+# Wrangle the two ACS tables
+acs_df_C24010 <- acs_df_C24010 %>%
+  rename(County = X) %>%
+  gather(key = Header, value = Workers, -County) %>%
+  mutate(Occupation = get_occupation(Header),
+         Gender = ifelse(grepl("Female", Header), "Female", ifelse(grepl("Male", Header), "Male", "Total"))) %>% 
+  filter(Gender != "Total") %>%
+  select(-Header) %>%
+  drop_na()
+
+acs_df_B24022 <- acs_df_B24022 %>%
+  rename(County = X) %>%
+  gather(key = Header, value = "Earnings", -County) %>%
+  filter(Earnings > 0) %>%
+  mutate(Occupation = get_occupation(Header),
+         Gender = ifelse(grepl("Female", Header), "Female", ifelse(grepl("Male", Header), "Male", "Total"))) %>%
+  filter(Gender != "Total") %>%
+  select(-Header)
+
+acs_df <- merge(acs_df_C24010, acs_df_B24022)
+
+# Add in the occupation concentrations
+acs_df <- acs_df %>%
+  select(Occupation, County, Workers) %>%
+  group_by(Occupation, County) %>%
+  summarise(Total = sum(Workers)) %>% 
+  merge(acs_df) %>%
+  mutate(Concentration = Workers / Total) %>%
+  select(-Total)
+
+# Add in the female to male earnings ratio
+acs_df <- acs_df %>%
+  filter(Gender == "Male") %>%
+  select(County, Occupation, Earnings) %>%
+  rename(Male_Earnings = Earnings) %>%
+  merge(acs_df) %>%
+  filter(Gender == "Female") %>%
+  mutate(Ratio = Earnings / Male_Earnings)
+```
 
@@
 
-**What are the cases, and how many are there?**
+There are `r nrow(acs_df)` observations in the dataset.
+Each observation is the ratio of median earnings by occupation at the county level  The universe is the civilian employed population 16 years and over.
+The occupational groups are based on the 2010 Standard Occupational Classification (SOC) definitions.
 
-### Data collection
+## Relevant Summary Statistics
 
-**Describe the method of data collection.**
+### Occupation Groups
 
-### Type of study
+There are `r length(unique(acs_df$Occupation))` occupations.  The following table summarizes the number of records by occupation:
 
-**What type of study is this (observational/experiment)?**
+##### Table 1. Count of Records by Occupation
+```{r, echo=FALSE}
+library(kableExtra)
+acs_df %>%
+  select(Occupation) %>%
+  group_by(Occupation) %>%
+  summarise(Records = n()) %>%
+  ungroup() %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
+```
 
-### Data Source
+### Female Worker Concentration
 
-**If you collected the data, state self-collected. If not, provide a citation/link.**
+##### Figure 1. Female Worker Concentration Histogram
+```{r, echo=FALSE}
+library(ggplot2)
+ggplot(acs_df, aes(Concentration)) + geom_histogram(bins = 60)
+```
 
-### Dependent Variable
+The female concentration seems to be multi-modal.  I imagine there will be occupation groups that have higher concentration rates than others.
 
-**What is the response variable? Is it quantitative or qualitative?**
+##### Table 2. Average Concentration Rate by Occupation
+```{r, echo=FALSE}
+acs_df %>%
+  group_by(Occupation) %>%
+  summarise(Concentration = mean(Concentration)) %>%
+  ungroup() %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
+```
 
-### Independent Variable
+### Female Earning Ratio
 
-**You should have two independent variables, one quantitative and one qualitative.**
+Before looking at the dependent variable I want to look at the underlying earnings data.
 
-### Relevant summary statistics
+##### Figure 2. Female Earnings Histogram
+```{r, echo=FALSE}
+ggplot(acs_df, aes(Earnings)) + geom_histogram(bins = 60)
+```
 
-**Provide summary statistics for each the variables.
-Also include appropriate visualizations related to your research question (e.g.
-scatter plot, boxplots, etc).
-This step requires the use of R, hence a code chunk is provided below.
-Insert more code chunks as needed.**
+The distribution is long-tailed.  Now to look at the independant variable.
 
-```{r}
+##### Figure 3. Earnings Ratio Histogram
+```{r, echo=FALSE}
+ggplot(acs_df, aes(Ratio)) + geom_histogram(bins = 60)
+```
 
+The ratio seems to be long tailed too.  Most of the observations have a ratio less than one.
+
+##### Table 3. ACS Summary Statistics for Numeric Variables
+```{r acs_numeric_summary, echo=FALSE}
+acs_df %>%
+  select(Ratio, Concentration, Earnings) %>%
+  summarise_all(funs(min=min, median=median, mean = mean, max=max)) %>%
+  gather(stat, val) %>%
+  separate(stat, into = c("Variable", "stat"), "_") %>%
+  spread(stat, val) %>%
+  select(Variable, min, mean, median, max) %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
 ```

### 2018-10-11 DATA607/Week_7_Assignment.Rmd
Initial commit of Week 7 Assignment

---
title: "Week 7 Assignment"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(rmdformats)
library(kableExtra)
```

```{r knitr_init, echo=FALSE, message=FALSE, eval=FALSE}
library(dplyr)
library(knitr)
library(rmdformats)
library(kableExtra)

## Global options
options(max.print="75")
# opts_chunk$set(echo=FALSE,
# 	             cache=TRUE,
#                prompt=FALSE,
#                tidy=TRUE,
#                comment=NA,
#                message=FALSE,
#                warning=FALSE)
opts_knit$set(width=75)
```

```{r, echo=FALSE}
# A Prefix nulling hook.
# Taken from https://stackoverflow.com/questions/22524822/how-can-i-remove-the-prefix-index-indicator-1-in-knitr-output

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
    x <- gsub( "##", "", x)
  }

  default_output_hook( x, options )

})

knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

## Description

In this assignment I have selected three books that I enjoyed reading and have entered information these three books in separately created files.  One encodes the information in HTML, XML and JSON formats.  I am to pull in the data into R and answer if the three data frames are identical.

### Book Files

I created the files by hand and saved the [HTML](https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.html), [XML](https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.xml) and [JSON](https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.json) files to my GitHub repository.  I have also loaded *dplyr* into my R environment.

## Reading HTML File into R

Here's what the HTML file looks like:

##### Table 1. Preview of books.html
```{r, null_prefix=TRUE, echo=FALSE}
print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.html"), quote=FALSE)
```

I will use rvest to extract the data from the table.

```{r, echo=TRUE, message=FALSE}
library(rvest)

html_df <- read_html("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.html") %>%
  html_node("table") %>%
  html_table()
```

Here's what the HTML data looks like after processed by rvest:

##### Table 2. Rvest Data Frame of books.html
```{r, echo=FALSE}
library(kableExtra)
html_df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

## Reading XML File into R

Now it's time to work on the XML file.  Here's what the XML file looks like:

##### Table 3. Preview of books.xml
```{r, null_prefix=TRUE, echo=FALSE}
print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.xml"), quote=FALSE)
```

I will use xml2 to extract the data from the XML file.

```{r, echo=TRUE}
library(xml2)
books_xml <- read_xml("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.xml")

X1 <- books_xml %>%
  xml_children() %>%
  xml_name()

X2 <- books_xml %>%
  xml_children() %>%
  xml_text()

xml_df <-data.frame(X1, X2)
```

Here's what the XML data looks like after passing through the above process:

##### Table 4. Data Frame of books.xml
```{r, echo=FALSE}
xml_df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

## Reading JSON File into R

Here's what the JSON file looks like:

##### Table 5. Preview of books.json
```{r, null_prefix=TRUE, echo=FALSE}
print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.json"), quote=FALSE)
```

I will use jsonlite to extract the data from the file.

```{r, echo=TRUE}
library(jsonlite)
json_df <- fromJSON("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.json") %>%
  unlist() %>%
  as.data.frame()

# Change the name
names(json_df) <- c("X2")
# Pull the row names into X1
json_df <- data.frame(X1 = row.names(json_df), json_df)
# Drop the row names
rownames(json_df) <- c()
```

Here's what the JSON data looks like after processed by rvest:

##### Table 6. Data Frame of books.json
```{r, echo=FALSE}
json_df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

## Are the Three Data Frames Identical?

The three data frames are close.  I processed the XML file so that it is identical to

### 2018-10-12 DATA607/Week_7_Assignment.Rmd
Finalized Week 7 Assignment

@@
 author: "Mike Silva"
-date: "`r Sys.Date()`"
+date: "October 8, 2018"
 output:
@@
 
-```{r, echo=FALSE, message=FALSE}
+```{r init, echo=FALSE, message=FALSE}
 library(dplyr)
@@
 library(kableExtra)
-```
 
-```{r knitr_init, echo=FALSE, message=FALSE, eval=FALSE}
-library(dplyr)
-library(knitr)
-library(rmdformats)
-library(kableExtra)
-
-## Global options
-options(max.print="75")
-# opts_chunk$set(echo=FALSE,
-# 	             cache=TRUE,
-#                prompt=FALSE,
-#                tidy=TRUE,
-#                comment=NA,
-#                message=FALSE,
-#                warning=FALSE)
-opts_knit$set(width=75)
-```
-
-```{r, echo=FALSE}
 # A Prefix nulling hook.
@@
 # Output hooks handle normal R console output.
-knitr::knit_hooks$set( output = function(x, options) {
+knitr::knit_hooks$set(output = function(x, options) {
 
   comment <- knitr::opts_current$get("comment")
-  if( is.na(comment) ) comment <- ""
+  if (is.na(comment)) comment <- ""
   can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
@@
   do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
-  if( can_null && do_null ) {
+  if (can_null && do_null) {
     # By default R print output aligns at the right brace.
@@
     x <- gsub( re, rep,  x )
+    # I added this line
     x <- gsub( "##", "", x)
@@
   default_output_hook( x, options )
-
 })
 
-knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
+knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
 ```
@@
 ##### Table 1. Preview of books.html
-```{r, null_prefix=TRUE, echo=FALSE}
-print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.html"), quote=FALSE)
+```{r books_html_preview, null_prefix=TRUE, echo=FALSE}
+print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.html"), quote = FALSE)
 ```
@@
 
-```{r, echo=TRUE, message=FALSE}
+```{r extract_html, echo=TRUE, message=FALSE}
 library(rvest)
@@
 
-##### Table 2. Rvest Data Frame of books.html
-```{r, echo=FALSE}
-library(kableExtra)
+##### Table 2. Data From books.html
+```{r book_html_table, echo=FALSE}
 html_df %>%
@@
 ##### Table 3. Preview of books.xml
-```{r, null_prefix=TRUE, echo=FALSE}
-print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.xml"), quote=FALSE)
+```{r book_xml_preview, null_prefix=TRUE, echo=FALSE}
+print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.xml"), quote = FALSE)
 ```
 
-I will use xml2 to extract the data from the XML file.
+I will use xml2 package to extract the data from the XML file.
 
-```{r, echo=TRUE}
+```{r extract_xml, echo=TRUE}
 library(xml2)
 books_xml <- read_xml("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.xml")
-
+# Pull out the element names
 X1 <- books_xml %>%
@@
   xml_name()
-
+# Pull out the value
 X2 <- books_xml %>%
@@
   xml_text()
-
-xml_df <-data.frame(X1, X2)
+# Put it all together in a data frame
+xml_df <- data.frame(X1, X2)
 ```
@@
 ##### Table 4. Data Frame of books.xml
-```{r, echo=FALSE}
+```{r  book_xml_table, echo=FALSE}
 xml_df %>%
@@
 
-Here's what the JSON file looks like:
+Finally I will work on the JSON file.  Here's what it looks like:
 
 ##### Table 5. Preview of books.json
-```{r, null_prefix=TRUE, echo=FALSE}
-print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.json"), quote=FALSE)
+```{r book_json_preview, null_prefix=TRUE, echo=FALSE}
+print(readLines("https://raw.githubusercontent.com/mikeasilva/CUNY-SPS/master/DATA607/data/books.json"), quote = FALSE)
 ```
 
-I will use jsonlite to extract the data from the file.
+I will use jsonlite package to extract the data from the file.
 
-```{r, echo=TRUE}
+```{r extract_json, echo=TRUE}
 library(jsonlite)
@@
   as.data.frame()
-
 # Change the name
@@
 
-Here's what the JSON data looks like after processed by rvest:
+Here's what the JSON data looks like after my wrangling:
 
 ##### Table 6. Data Frame of books.json
-```{r, echo=FALSE}
+```{r book_json_table, echo=FALSE}
 json_df %>%
@@
 
-The three data frames are close.  I processed the XML file so that it is identical to
+They are now but that was intentional.  If I simply read in the data and made it into a data frame they would not be.

### 2018-10-12 DATA606/Lab4a/michael-sampling_distributions.Rmd
Work on Lab4a

---
title: 'Foundations for statistical inference - Sampling distributions'
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

In this lab, we investigate the ways in which the statistics from a random 
sample of data can serve as point estimates for population parameters.  We're 
interested in formulating a *sampling distribution* of our estimate in order 
to learn about the properties of the estimate, such as its distribution.

## The data

We consider real estate data from the city of Ames, Iowa.  The details of 
every real estate transaction in Ames is recorded by the City Assessor's 
office.  Our particular focus for this lab will be all residential home sales 
in Ames between 2006 and 2010.  This collection represents our population of 
interest.  In this lab we would like to learn about these home sales by taking 
smaller samples from the full population.  Let's load the data.

```{r load-data, eval=TRUE}
load("more/ames.RData")
```

We see that there are quite a few variables in the data set, enough to do a 
very in-depth analysis.  For this lab, we'll restrict our attention to just 
two of the variables: the above ground living area of the house in square feet 
(`Gr.Liv.Area`) and the sale price (`SalePrice`).  To save some effort 
throughout the lab, create two variables with short names that represent these 
two variables.

```{r assign, eval=TRUE}
area <- ames$Gr.Liv.Area
price <- ames$SalePrice
```

Let's look at the distribution of area in our population of home sales by 
calculating a few summary statistics and making a histogram.

```{r area, eval=TRUE}
summary(area)
hist(area)
```

1.  Describe this population distribution. **It is long-tailed or right skewed distribution.**

## The unknown sampling distribution

In this lab we have access to the entire population, but this is rarely the 
case in real life.  Gathering information on an entire population is often 
extremely costly or impossible.  Because of this, we often take a sample of 
the population and use that to understand the properties of the population.

If we were interested in estimating the mean living area in Ames based on a 
sample, we can use the following command to survey the population.

```{r samp1, eval=TRUE}
samp1 <- sample(area, 50)
```

This command collects a simple random sample of size 50 from the vector 
`area`, which is assigned to `samp1`.  This is like going into the City 
Assessor's database and pulling up the files on 50 random home sales.  Working 
with these 50 files would be considerably simpler than working with all 2930 
home sales.

2.  Describe the distribution of this sample. How does it compare to the 
    distribution of the population? **It is similar to the population distribution.**
    ```{r, echo=FALSE, message=FALSE}
    library(ggplot2)
    library(dplyr)
    df <- data.frame(area=area, Group=c("Population"))
    df <- data.frame(area=samp1, Group=c("Sample")) %>%
      rbind(df) %>%
      mutate(Group = as.factor(Group))
    ggplot(df) + geom_density(aes(x=area, fill=Group, alpha=0.2)) + guides(alpha="none")
    ```

If we're interested in estimating the average living area in homes in Ames 
using the sample, our best single guess is the sample mean.

```{r mean-samp1, eval=TRUE}
mean(samp1)
```

Depending on which 50 homes you selected, your estimate could be a bit above 
or a bit below the true population mean of 1499.69 square feet.  In general, 
though, the sample mean turns out to be a pretty good estimate of the average 
living area, and we were able to get it by sampling less than 3\% of the 
population.

3.  Take a second sample, also of size 50, and call it `samp2`.  How does the 
    mean of `samp2` compare with the mean of `samp1`?  **They are pretty close to each other.** Suppose we took two 
    more samples, one of size 100 and one of size 1000. Which would you think 
    would provide a more accurate estimate of the population mean? **The sample with more observations.**
    ```{r samp2}
    samp2 <- sample(area, 50)
    message(paste("Population:", mean(area)))
    message(paste("Sample of 50 #1:", mean(samp1)))
    message(paste("Sample of 50 #2:", mean(samp2)))
    message(paste("Sample of 100:", mean(sample(area, 100))))
    message(paste("Sample of 1000:", mean(sample(area, 1000))))
    ```

Not surprisingly, every time we take another random sample, we get a different 
sample mean.  It's useful to get a sense of just how much variability we 
should expect when estimating the population mean this way. The distribution 
of sample means, called the *sampling distribution*, can help us understand 
this variability. In this lab, because we have access to the population, we 
can build up the sampling distribution for the sample mean by repeating the 
above steps many times. Here we will generate 5000 samples and compute the 
sample mean of each.

```{r loop, eval=TRUE}
sample_means50 <- rep(NA, 5000)

for(i in 1:5000){
   samp <- sample(area, 50)
   sample_means50[i] <- mean(samp)
   }

hist(sample_means50)
```

If you would like to adjust the bin width of your histogram to show a little 
more detail, you can do so by changing the `breaks` argument.

```{r hist-breaks, eval=TRUE}
hist(sample_means50, breaks = 25)
```

Here we use R to take 5000 samples of size 50 from the population, calculate 
the mean of each sample, and store each result in a vector called 
`sample_means50`. On the next page, we'll review how this set of code works.

4.  How many elements are there in `sample_means50`?  **`r length(sample_means50)`** Describe the sampling 
    distribution, and be sure to specifically note its center.  Would you 
    expect the distribution to change if we instead collected 50,000 sample 
    means?  **Its center is about `r round(mean(sample_means50), 0)`.  The distribution would still be normally distributed but it should haved a smaller spread.**

## Interlude: The `for` loop

Let's take a break from the statistics for a moment to let that last block of 
code sink in.  You have just run your first `for` loop, a cornerstone of 
computer programming.  The idea behind the for loop is *iteration*: it allows 
you to execute code as many times as you want without having to type out every 
iteration.  In the case above, we wanted to iterate the two lines of code 
inside the curly braces that take a random sample of size 50 from `area` then 
save the mean of that sample into the `sample_means50` vector.  Without the 
`for` loop, this would be painful:

```{r loop-long, eval=TRUE}
sample_means50 <- rep(NA, 5000)

samp <- sample(area, 50)
sample_means50[1] <- mean(samp)

samp <- sample(area, 50)
sample_means50[2] <- mean(samp)

samp <- sample(area, 50)
sample_means50[3] <- mean(samp)

samp <- sample(area, 50)
sample_means50[4] <- mean(samp)
```

and so on...

With the for loop, these thousands of lines of code are compressed into a 
handful of lines. We've added one extra line to the code below, which prints 
the variable `i` during each iteration of the `for` loop. Run this code.

<div style="height:200px; overflow-y:scroll;">
```{r loop-again, eval=TRUE}
sample_means50 <- rep(NA, 5000)

for(i in 1:5000){
   samp <- sample(area, 50)
   sample_means50[i] <- mean(samp)
   print(i)
   }
```
</div>

Let's consider this code line by line to figure out what it does.  In the 
first line we *initialized a vector*.  In this case, we created a vector of 
5000 zeros called `sample_means50`.  This vector will will store values 
generated within the `for` loop.

The second line calls the `for` loop itself.  The syntax can be loosely read as, 
"for every element `i` from 1 to 5000, run the following lines of code". You 
can think of `i` as the counter that keeps track of which loop you're on. 
Therefore, more precisely, the loop will run once when `i = 1`, then once when 
`i = 2`, and so on up to `i = 5000`.

The body of the `for` loop is the part inside the curly braces, and this set of 
code is run for each value of `i`.  Here, on every loop, we take a random 
sample of size 50 from `area`, take its mean, and store it as the 
$i$<sup>th</sup> element of `sample_means50`.

In order to display that this is really happening, we asked R to print `i` at 
each iteration. This line of code is optional and is only used for displaying 
what's going on while the `for` loop is running.

The `for` loop allows us to not just run the code 5000 times, but to neatly 
package the results, element by element, into the empty vector that we 
initialized at the outset.

5.  To make sure you understand what you've done in this loop, try running a 
    smaller version.  Initialize a vector of 100 zeros called 
    `sample_means_small`.  Run a loop that takes a sample of size 50 from 
    `area` and stores the sample mean in `sample_means_small`, but only 
    iterate from 1 to 100.

    ```{r}
    sample_means_small <- rep(0, 100)
    for(i in 1:100){
      sample_means_small[i] <- mean(sample(area, 50))
    }
    ```

    Print the output to your screen (type `sample_means_small` into the console and press enter).  
    ```{r}
    sample_means_small
    ```
    How many elements are there in this object called `sample_means_small`? **`r length(sample_means_small)`**
    What does each element represent? **The mean area of 50 randomly selected homes in Ames.**

## Sample size and the sampling distribution

Mechanics aside, let's return to the reason we used a `for` loop: to compute a 
sampling distribution, specifically, this one.

```{r hist, eval=TRUE}
hist(sample_means50)
```

The sampling distribution that we computed tells us much about estimating 
the average living area in homes in Ames.  Because the sample mean is an 
unbiased estimator, the sampling distribution is centered at the true average 
living area of the the population, and the spread of the distribution 
indicates how much variability is induced by sampling only 50 home sales.

To get a sense of the effect that sample size has on our distribution, let's 
build up two more sampling distributions: one based on a sample size of 10 and 
another based on a sample size of 100.

```{r samp-10-100, eval=TRUE}
sample_means10 <- rep(NA, 5000)
sample_means100 <- rep(NA, 5000)

for(i in 1:5000){
  samp <- sample(area, 10)
  sample_means10[i] <- mean(samp)
  samp <- sample(area, 100)
  sample_means100[i] <- mean(samp)
}
```

Here we're able to use a single `for` loop to build two distributions by adding 
additional lines inside the curly braces.  Don't worry about the fact that 
`samp` is used for the name of two different objects.  In the second command 
of the `for` loop, the mean of `samp` is saved to the relevant place in the 
vector `sample_means10`.  With the mean saved, we're now free to overwrite the 
object `samp` with a new sample, this time of size 100.  In general, anytime 
you create an object using a name that is already in use, the old object will 
get replaced with the new one.

To see the effect that different sample sizes have on the sampling 
distribution, plot the three distributions on top of one another.

```{r plot-samps, eval=TRUE, tidy = FALSE}
par(mfrow = c(3, 1))

xlimits <- range(sample_means10)

hist(sample_means10, breaks = 20, xlim = xlimits)
hist(sample_means50, breaks = 20, xlim = xlimits)
hist(sample_means100, breaks = 20, xlim = xlimits)
```

The first command specifies that you'd like to divide the plotting area into 3 
rows and 1 column of plots (to return to the default setting of plotting one 
at a time, use `par(mfrow = c(1, 1))`). The `breaks` argument specifies the 
number of bins used in constructing the histogram.  The `xlim` argument 
specifies the range of the x-axis of the histogram, and by setting it equal 
to `xlimits` for each histogram, we ensure that all three histograms will be 
plotted with the same limits on the x-axis.

6.  When the sample size is larger, what happens to the center?  What about the spread?

* * *
## On your own

So far, we have only focused on estimating the mean living area in homes in 
Ames.  Now you'll try to estimate the mean home price.

-   Take a random sample of size 50 from `price`. Using this sample, what is 
    your best point estimate of the population mean?

-   Since you have access to the population, simulate the sampling 
    distribution for $\bar{x}_{price}$ by taking 5000 samples from the 
    population of size 50 and computing 5000 sample means.  Store these means 
    in a vector called `sample_means50`. Plot the data, then describe the 
    shape of this sampling distribution. Based on this sampling distribution, 
    what would you guess the mean home price of the population to be? Finally, 
    calculate and report the population mean.

-   Change your sample size from 50 to 150, then compute the sampling 
    distribution using the same method as above, and store these means in a 
    new vector called `sample_means150`. Describe the shape of this sampling 
    distribution, and compare it to the sampling distribution for a sample 
    size of 50.  Based on this sampling distribution, what would you guess to 
    be the mean sale price of homes in Ames?

-   Of the sampling distributions from 2 and 3, which has a smaller spread?  If
    we're concerned with making estimates that are more often close to the 
    true value, would we prefer a distribution with a large or small spread?

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>

### 2018-10-15 DATA606/Lab4a/michael-sampling_distributions.Rmd
Additional Work on Lab 4A

@@
 
-6.  When the sample size is larger, what happens to the center?  What about the spread?
+6.  When the sample size is larger, what happens to the center?  **The center becomes closer to the population mean.**  What about the spread? **The spread narrows.**
 
@@
     your best point estimate of the population mean?
+    ```{r}
+    avg_price <- mean(sample(price, 50))
+    avg_price
+    ```
+    **This is the best estimate of the population mean.**
 
@@
     population of size 50 and computing 5000 sample means.  Store these means 
-    in a vector called `sample_means50`. Plot the data, then describe the 
-    shape of this sampling distribution. Based on this sampling distribution, 
-    what would you guess the mean home price of the population to be? Finally, 
-    calculate and report the population mean.
+    in a vector called `sample_means50`. 
+    ```{r}
+    for(i in 1:5000){
+      sample_means50[i] <- mean(sample(price, 50))
+    }
+    ```
+    Plot the data, then describe the shape of this sampling distribution.
+
+    **The distribution is normally distributed.**
+
+    ```{r, echo=FALSE}
+    ggplot() + geom_histogram(aes(sample_means50), bins=100) + geom_vline(xintercept=mean(sample_means50), linetype="dashed", color="red")
+    ```
+
+    Based on this sampling distribution, what would you guess the mean home price of the population to be? **Roughly $180K.**
+
+    Finally, calculate and report the population mean.
+    ```{r}
+    mean(price)
+    ```
 
@@
     distribution using the same method as above, and store these means in a 
-    new vector called `sample_means150`. Describe the shape of this sampling 
-    distribution, and compare it to the sampling distribution for a sample 
-    size of 50.  Based on this sampling distribution, what would you guess to 
-    be the mean sale price of homes in Ames?
+    new vector called `sample_means150`. 
+    ```{r}
+    sample_means150 <- rep(NA, 150)
+    for(i in 1:5000){
+      sample_means150[i] <- mean(sample(price, 150))
+    }
+    ```
+    Describe the shape of this sampling distribution, and compare it to the 
+    sampling distribution for a sample size of 50.  Based on this sampling 
+    distribution, what would you guess to be the mean sale price of homes in Ames?
+    ```{r}
+    df <- data.frame(price=sample_means150, N=c("150"))
+    df <- data.frame(price=sample_means50, N=c("50")) %>%
+      rbind(df) %>%
+      mutate(N = as.factor(N))
+    ggplot(df) + geom_density(aes(x=price, fill=N, alpha=0.2)) + guides(alpha="none")
+    ```
 
--   Of the sampling distributions from 2 and 3, which has a smaller spread?  If
-    we're concerned with making estimates that are more often close to the 
+    **Roughly $180K still but I'm more certain of it.**
+
+-   Of the sampling distributions from 2 and 3, which has a smaller spread?  
+    **The one with 150 observations in the sample.**
+    If we're concerned with making estimates that are more often close to the 
     true value, would we prefer a distribution with a large or small spread?
+    **As large as we can reasonably get.**
 

### 2018-10-16 DATA606/Lab4b/michael-confidence_intervals.Rmd
More work on Lab 4b

---
title: 'Foundations for statistical inference - Confidence intervals'
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---
```{r echo=FALSE}
# A Prefix nulling hook.
# Taken from https://stackoverflow.com/questions/22524822/how-can-i-remove-the-prefix-index-indicator-1-in-knitr-output

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set(output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if (is.na(comment)) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if (can_null && do_null) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
    # I added this line
    x <- gsub( "##", "", x)
  }

  default_output_hook( x, options )
})

knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
```
## Sampling from Ames, Iowa

If you have access to data on an entire population, say the size of every 
house in Ames, Iowa, it's straight forward to answer questions like, "How big 
is the typical house in Ames?" and "How much variation is there in sizes of 
houses?". If you have access to only a sample of the population, as is often 
the case, the task becomes more complicated. What is your best guess for the 
typical size if you only know the sizes of several dozen houses? This sort of 
situation requires that you use your sample to make inference on what your 
population looks like.

## The data

In the previous lab, ``Sampling Distributions'', we looked at the population data
of houses from Ames, Iowa. Let's start by loading that data set.

```{r load-data, eval=TRUE}
load("more/ames.RData")
```

In this lab we'll start with a simple random sample of size 60 from the 
population. Specifically, this is a simple random sample of size 60. Note that 
the data set has information on many housing variables, but for the first 
portion of the lab we'll focus on the size of the house, represented by the 
variable `Gr.Liv.Area`.

```{r sample, eval=TRUE}
population <- ames$Gr.Liv.Area
samp <- sample(population, 60)
```

1.  Describe the distribution of your sample. What would you say is the 
"typical" size within your sample? **The "typical" size is `r mean(samp)`.** Also state precisely what you interpreted "typical" to mean.  **In this case typical means the mean of the mean of the sample.**

2.  Would you expect another student's distribution to be identical to yours? 
Would you expect it to be similar? Why or why not? **I would expect it to be similar but not identical to mine.  Similar because they were drawn from the same population.  But not identical because they are selected using a random process.**

## Confidence intervals

One of the most common ways to describe the typical or central value of a 
distribution is to use the mean. In this case we can calculate the mean of the 
sample using,

```{r sample-mean, eval=TRUE}
sample_mean <- mean(samp)
```

Return for a moment to the question that first motivated this lab: based on 
this sample, what can we infer about the population? Based only on this single 
sample, the best estimate of the average living area of houses sold in Ames 
would be the sample mean, usually denoted as $\bar{x}$ (here we're calling it 
`sample_mean`). That serves as a good *point estimate* but it would be useful 
to also communicate how uncertain we are of that estimate. This can be 
captured by using a *confidence interval*.

We can calculate a 95% confidence interval for a sample mean by adding and 
subtracting 1.96 standard errors to the point estimate (See Section 4.2.3 if 
you are unfamiliar with this formula).

```{r ci, eval=TRUE}
se <- sd(samp) / sqrt(60)
lower <- sample_mean - 1.96 * se
upper <- sample_mean + 1.96 * se
c(lower, upper)
```

This is an important inference that we've just made: even though we don't know 
what the full population looks like, we're 95% confident that the true 
average size of houses in Ames lies between the values *lower* and *upper*. 
There are a few conditions that must be met for this interval to be valid.

3.  For the confidence interval to be valid, the sample mean must be normally 
distributed and have standard error $s / \sqrt{n}$. What conditions must be 
met for this to be true? **Obervations are IID and sample is random.**

## Confidence levels

4.  What does "95% confidence" mean? If you're not sure, see Section 4.2.2. **If 100 confidence intervals were taken, the population mean would fall within the CI about 95% of the time.**

In this case we have the luxury of knowing the true population mean since we 
have data on the entire population. This value can be calculated using the 
following command:

```{r pop-mean, eval=TRUE}
pop_mean <- mean(population)
pop_mean
```

5.  Does your confidence interval capture the true average size of houses in 
Ames? If you are working on this lab in a classroom, does your neighbor's 
interval capture this value? 
```{r, null_prefix=TRUE, echo=FALSE, results='asis'} 
if(contains(lower, upper, pop_mean)) {
  paste('**Yes it does.', round(pop_mean,0), 'is between', round(lower,0), 'and', round(upper,0), '.**')
} else{
  paste('**No it does not.', round(pop_mean,0), 'is not between', round(lower,0), 'and', round(upper,0), '.**')
}
```

6.  Each student in your class should have gotten a slightly different 
confidence interval. What proportion of those intervals would you expect to 
capture the true population mean? **95%** Why? **Due to the central limit theorem.**If you are working in this lab in a 
classroom, collect data on the intervals created by other students in the 
class and calculate the proportion of intervals that capture the true 
population mean.

Using R, we're going to recreate many samples to learn more about how sample 
means and confidence intervals vary from one sample to another. *Loops* come 
in handy here (If you are unfamiliar with loops, review the [Sampling Distribution Lab](http://htmlpreview.github.io/?https://github.com/andrewpbray/oiLabs/blob/master/sampling_distributions/sampling_distributions.html)).

Here is the rough outline:

-   Obtain a random sample.
-   Calculate and store the sample's mean and standard deviation.
-   Repeat steps (1) and (2) 50 times.
-   Use these stored statistics to calculate many confidence intervals.

But before we do all of this, we need to first create empty vectors where we 
can save the means and standard deviations that will be calculated from each 
sample. And while we're at it, let's also store the desired sample size as `n`.

```{r set-up, eval=TRUE}
samp_mean <- rep(NA, 50)
samp_sd <- rep(NA, 50)
n <- 60
```

Now we're ready for the loop where we calculate the means and standard deviations of 50 random samples.

```{r loop, eval=TRUE, tidy = FALSE}
for(i in 1:50){
  samp <- sample(population, n) # obtain a sample of size n = 60 from the population
  samp_mean[i] <- mean(samp)    # save sample mean in ith element of samp_mean
  samp_sd[i] <- sd(samp)        # save sample sd in ith element of samp_sd
}
```

Lastly, we construct the confidence intervals.

```{r ci50, eval=TRUE}
lower_vector <- samp_mean - 1.96 * samp_sd / sqrt(n) 
upper_vector <- samp_mean + 1.96 * samp_sd / sqrt(n)
```

Lower bounds of these 50 confidence intervals are stored in `lower_vector`, 
and the upper bounds are in `upper_vector`. Let's view the first interval.

```{r first-interval, eval=TRUE}
c(lower_vector[1], upper_vector[1])
```

* * *

## On your own

-   Using the following function (which was downloaded with the data set), 
    plot all intervals. 
    ```{r plot-ci, eval=TRUE}
    plot_ci(lower_vector, upper_vector, mean(population))
    ```

    What proportion of your confidence intervals include the true population mean?

    ```{r}
    n_within <- 0
    for(i in 1:50){
      lower <- lower_vector[i]
      upper <- upper_vector[i]
      if(contains(lower, upper, pop_mean)){
        n_within <- n_within + 1
      }
    }
    share_95 <- n_within / 50
    share_95
    ```

    Is this proportion exactly equal to the confidence level?  If not, explain why.

    ```{r, null_prefix=TRUE, echo=FALSE, results='asis'} 
      if(share_95 >= .95){
        print('**Yes**')
      } else {
        print('**No.  This is one of the 5% of the times it does not.**')
      }
    ```

-   Pick a confidence level of your choosing, provided it is not 95%. What is 
    the appropriate critical value? **For a 90% CI the critical value is 1.64.**

-   Calculate 50 confidence intervals at the confidence level you chose in the 
    previous question. You do not need to obtain new samples, simply calculate 
    new intervals based on the sample means and standard deviations you have 
    already collected. Using the `plot_ci` function, plot all intervals and 
    calculate the proportion of intervals that include the true population 
    mean. How does this percentage compare to the confidence level selected for
    the intervals?

    ```{r tidy = FALSE}
    for(i in 1:50){
      samp <- sample(population, n)
      samp_mean[i] <- mean(samp)
      samp_sd[i] <- sd(samp)
    }

    lower_vector <- samp_mean - 1.64 * samp_sd / sqrt(n) 
    upper_vector <- samp_mean + 1.64 * samp_sd / sqrt(n)

    plot_ci(lower_vector, upper_vector, pop_mean)

    n_within <- 0
    for(i in 1:50){
      lower <- lower_vector[i]
      upper <- upper_vector[i]
      if(contains(lower, upper, pop_mean)){
        n_within <- n_within + 1
      }
    }

    n_within / 50 
    ```

    **There are more CI's that are red (contain the population mean) than above.**

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>

### 2018-10-16 DATA606/Foundations for Inference.Rmd
Initial commit

---
title: "Foundations for Inference"
author: "Mike Silva"
date: "October 16, 2018"
output: html_document
---

## Description

The following is the graded problems for the Foundations for Inference chapter of Open Intro to Statistics.  My answers to the questions are **bolded.**  The headers correspond to a question.

## 4.4 Heights of adults
Researchers studying anthropometry collected body girth measurements and skeletal diameter measurements, as well as age, weight, height and gender, for 507 physically active individuals. The histogram below shows the sample distribution of heights in centimeters.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/4_4.png)

(a) What is the point estimate for the average height of active individuals? What about the median?

(b) What is the point estimate for the standard deviation of the heights of active individuals? What about the IQR?

(c) Is a person who is 1m 80cm (180 cm) tall considered unusually tall? And is a person who is 1m 55cm (155cm) considered unusually short? Explain your reasoning.

(d) The researchers take another random sample of physically active individuals. Would you expect the mean and the standard deviation of this new sample to be the ones given above? Explain your reasoning.

(e) The sample means obtained are point estimates for the mean height of all active individuals, if the sample of individuals is equivalent to a simple random sample. What measure do we use to quantify the variability of such an estimate (Hint: recall that $SD_\bar{x} = \frac{\sigma}{\sqrt{n}}$)? Compute this quantity using the data from the original sample under the condition that the data are a simple random sample.

## 4.14 Thanksgiving spending, Part I

The 2009 holiday retail season, which kicked off on November 27, 2009 (the day after Thanksgiving), had been marked by somewhat lower self-reported consumer spending than was seen during the comparable period in 2008. To get an estimate of consumer spending, 436 randomly sampled American adults were surveyed. Daily consumer spending for the six-day period after Thanksgiving, spanning the Black Friday weekend and Cyber Monday, averaged $84.71. A 95% confidence interval based on this sample is ($80.31, $89.11). Determine whether the following statements are true or false, and explain your reasoning.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/4_14.png)

(a) We are 95% confident that the average spending of these 436 American adults is between $80.31 and $89.11.

(b) This confidence interval is not valid since the distribution of spending in the sample is right skewed.

(c) 95% of random samples have a sample mean between $80.31 and $89.11.

(d) We are 95% confident that the average spending of all American adults is between $80.31 and $89.11.

(e) A 90% confidence interval would be narrower than the 95% confidence interval since we don’t need to be as sure about our estimate.

(f) In order to decrease the margin of error of a 95% confidence interval to a third of what it is now, we would need to use a sample 3 times larger.

(g) The margin of error is 4.4.

## 4.24 Gifted children, Part I

Researchers investigating characteristics of gifted children collected data from schools in a large city on a random sample of thirty-six children who were identified as gifted children soon after they reached the age of four. The following histogram shows the distribution of the ages (in months) at which these children first counted to 10 successfully. Also provided are some sample statistics.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/4_24.png)

(a) Are conditions for inference satisfied?

(b) Suppose you read online that children first count to 10 successfully when they are 32 months old, on average. Perform a hypothesis test to evaluate if these data provide convincing evidence that the average age at which gifted children first count to 10 successfully is less than the general average of 32 months. Use a significance level of 0.10.

(c) Interpret the p-value in context of the hypothesis test and the data.

(d) Calculate a 90% confidence interval for the average age at which gifted children first count to 10 successfully.

(e) Do your results from the hypothesis test and the confidence interval agree? Explain.

## 4.26 Gifted children, Part II

Exercise 4.24 describes a study on gifted children. In this study, along with variables on the children, the researchers also collected data on the mother’s and father’s IQ of the 36 randomly sampled gifted children. The histogram below shows the distribution of mother’s IQ. Also provided are some sample statistics.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/4_26.png)

(a) Perform a hypothesis test to evaluate if these data provide convincing evidence that the average IQ of mothers of gifted children is different than the average IQ for the population at large, which is 100. Use a significance level of 0.10.

(b) Calculate a 90% confidence interval for the average IQ of mothers of gifted children.

(c) Do your results from the hypothesis test and the confidence interval agree? Explain.

## 4.34 CLT

Define the term “sampling distribution” of the mean, and describe how the shape, center, and spread of the sampling distribution of the mean change as sample size increases.

## 4.40 CFLBs

A manufacturer of compact fluorescent light bulbs advertises that the distribution of the lifespans of these light bulbs is nearly normal with a mean of 9,000 hours and a standard deviation of 1,000 hours.

(a) What is the probability that a randomly chosen light bulb lasts more than 10,500 hours?

(b) Describe the distribution of the mean lifespan of 15 light bulbs.

(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours?

(d) Sketch the two distributions (population and sampling) on the same scale.

(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution?

## 4.48 Same observation, different sample size

Suppose you conduct a hypothesis test based on a sample where the sample size is n = 50, and arrive at a p-value of 0.08. You then refer back to your notes and discover that you made a careless mistake, the sample size should have been n = 500. Will your p-value increase, decrease, or stay the same? Explain.

### 2018-10-16 DATA606/Foundations for Inference.Rmd
Updated DATA 606 Homework

@@
 date: "October 16, 2018"
-output: html_document
+output:
+  html_document:
+    theme: yeti
+    highlight: tango
+    toc: true
+    toc_depth: 2
+    toc_float: true
 ---
@@
 
-(a) What is the point estimate for the average height of active individuals? What about the median?
+(a) What is the point estimate for the average height of active individuals? **171.1 cm** What about the median? **170.3 cm**
 
-(b) What is the point estimate for the standard deviation of the heights of active individuals? What about the IQR?
+(b) What is the point estimate for the standard deviation of the heights of active individuals? **9.4 cm** What about the IQR? **177.8 cm - 163.8 cm = 14 cm**
 
-(c) Is a person who is 1m 80cm (180 cm) tall considered unusually tall? And is a person who is 1m 55cm (155cm) considered unusually short? Explain your reasoning.
+(c) Is a person who is 1m 80cm (180 cm) tall considered unusually tall?
+And is a person who is 1m 55cm (155cm) considered unusually short?
+Explain your reasoning.
+**A person who is 180 cm is pretty tall but I wouldn't call it unusually tall b/c it's within 1 standard deviation of the mean.
+By the same token, I would consider someone who is 155 cm small because they are less than one standard deviation from the mean.**
 
@@
 Explain your reasoning.
+**I would expect both the mean and standard deviation to be different.
+There is variation within the population which would translate into a different sample mean and standard deviations.**
 
@@
 What measure do we use to quantify the variability of such an estimate (Hint: recall that $SD_\bar{x} = \frac{\sigma}{\sqrt{n}}$)?
-Compute this quantity using the data from the original sample under the condition that the data are a simple random sample.
+**The standard error**  Compute this quantity using the data from the original sample under the condition that the data are a simple random sample.
+**0.4174687**
+$SD_\bar{x} = \frac{\sigma}{\sqrt{n}} = \frac{9.4}{\sqrt{507}} = 0.4174687$
 
@@
 
-(a) We are 95% confident that the average spending of these 436 American adults is between $80.31 and $89.11.
+(a) We are 95% confident that the average spending of these 436 American adults is between $80.31 and $89.11. **False.  The 95% applies to the population parameter not the sample.**
 
-(b) This confidence interval is not valid since the distribution of spending in the sample is right skewed.
+(b) This confidence interval is not valid since the distribution of spending in the sample is right skewed.  **True.  This is one of the conditions for a confidence interval.**
 
-(c) 95% of random samples have a sample mean between $80.31 and $89.11.
+(c) 95% of random samples have a sample mean between $80.31 and $89.11. **False.  Cofidence interval does not contstrain other sample means.**
 
-(d) We are 95% confident that the average spending of all American adults is between $80.31 and $89.11.
+(d) We are 95% confident that the average spending of all American adults is between $80.31 and $89.11. **True.  This is assuming that the 436 people in the sample were a renadom draw.**
 
 (e) A 90% confidence interval would be narrower than the 95% confidence interval since we don’t need to be as sure about our estimate.
+**False.
+We need to be just as sure about the estimate.
+The 90% means that the population parameter is within the CI 90% of the time or that we are less sure that our sample captures the population parameter.**
 
 (f) In order to decrease the margin of error of a 95% confidence interval to a third of what it is now, we would need to use a sample 3 times larger.
+**False.
+If we had a large sample size trippling it would not decrease the margin of error.**
 
-(g) The margin of error is 4.4.
+(g) The margin of error is 4.4. **False.  That is the difference between the lower bound of the CI and the mean, not the margin of error.**
 
@@
 (a) Are conditions for inference satisfied?
+**The N > 30 rule of thumb is satisfied, the children were randomly selected, and the histogram appears to be normally distributed.
+I think all conditions have been met.**
 
@@
 Define the term “sampling distribution” of the mean, and describe how the shape, center, and spread of the sampling distribution of the mean change as sample size increases.
+**This is the distributions of the mean value of the samples.
+It is normally distributed.
+The mean approximates the population mean.
+As the sample size increases the sample mean approaches the population mean, and the spread decreases.**
 
@@
 
-(a) What is the probability that a randomly chosen light bulb lasts more than 10,500 hours?
+(a) What is the probability that a randomly chosen light bulb lasts more than 10,500 hours? **0.067**
 
-(b) Describe the distribution of the mean lifespan of 15 light bulbs.
+```{r}
+mu <- 9000 
+sd <- 1000
+x <- 10500
+1 - pnorm(x, mu, sd)
+```
 
-(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours?
+(b) Describe the distribution of the mean lifespan of 15 light bulbs. **Centered around 9,000 hours.**
+
+(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours? **It would be roughly 0.067**
 
@@
 
-(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution?
+(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution? **Yes, but not using the above method.**
 

### 2018-10-19 DATA606/Foundations for Inference.Rmd
Additional work on homework

@@
 Use a significance level of 0.10.
+**$H_0$ is that x_bar >= 32 and $H_a$ is x_bar < 32**
+
+```{r}
+alpha <- 0.10
+n <- 36
+x_bar <- 30.69
+s <- 4.31
+mu <- 32
+
+se_x_bar <- s / sqrt(n)
+
+test_stat <- (x_bar - mu)/se_x_bar
+
+t_alpha <- qt(1-alpha, df = n-1) * -1
+
+p_value <- pt(test_stat, df=n-1)
+
+if(test_stat < t_alpha){
+  message(paste(test_stat, "is less than", t_alpha, "so reject the null hypothesis (p-value:", p_value, ")"))
+} else {
+  message(paste(test_stat, "is greater than", t_alpha, "so accept the null hypothesis (p-value:", p_value, ")"))
+}
+```
 
 (c) Interpret the p-value in context of the hypothesis test and the data.
+**There is a 4% probability that the difference between the two means would be equal to or more extreme than its observed value.**
 
 (d) Calculate a 90% confidence interval for the average age at which gifted children first count to 10 successfully.
+```{r}
+error <- qt(1-alpha, df=n-1) * se_x_bar
+ci_low <- x_bar - error
+ci_high <- x_bar + error
+```
+**Confidence interval: (`r ci_low`, `r ci_high`) or `r x_bar` $\pm$ `r error`**
 
-(e) Do your results from the hypothesis test and the confidence interval agree? Explain.
+(e) Do your results from the hypothesis test and the confidence interval agree?
+Explain.
+**Yes since 32 is outside of the confidence interval you would reject the null hypotheses.
+it is the same conclusion you reach with the hypothesis test.**
 
@@
 
+```{r}
+alpha <- 0.10
+n <- 36
+x_bar <- 118.2
+s <- 6.5
+mu <- 100
+
+se_x_bar <- s / sqrt(n)
+
+test_stat <- (x_bar - mu)/se_x_bar
+
+t_alpha <- qt(1-alpha, df = n-1) * -1
+
+p_value <- pt(test_stat, df=n-1)
+
+if(test_stat < t_alpha){
+  message(paste(test_stat, "is less than", t_alpha, "so reject the null hypothesis (p-value:", p_value, ")"))
+} else {
+  message(paste(test_stat, "is greater than", t_alpha, "so accept the null hypothesis (p-value:", p_value, ")"))
+}
+```
+
 (b) Calculate a 90% confidence interval for the average IQ of mothers of gifted children.
+
+```{r}
+error <- qt(1-alpha, df=n-1) * se_x_bar
+ci_low <- x_bar - error
+ci_high <- x_bar + error
+```
+
+**Confidence interval: (`r ci_low`, `r ci_high`) or `r x_bar` $\pm$ `r error`**
 
@@
 
-(b) Describe the distribution of the mean lifespan of 15 light bulbs. **Centered around 9,000 hours.**
+(b) Describe the distribution of the mean lifespan of 15 light bulbs. **Normal and centered around 9,000 hours.**
 
-(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours? **It would be roughly 0.067**
+(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours?
 
@@
 
-(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution? **Yes, but not using the above method.**
+(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution? **Not using the above method.**
 

### 2018-10-19 DATA606/Foundations for Inference.Rmd
Additional work on homework

@@
 
+```{r}
+n <- 15
+se <- sd / sqrt(n)
+1 - pnorm(x, mu, se)
+```
+
 (d) Sketch the two distributions (population and sampling) on the same scale.
+```{r, echo=FALSE}
+library(ggplot2)
+set.seed(4)
+pop <- rnorm(1000000, mu, sd)
+sample_pop <- sample(pop, 15)
+data <-  rbind(data.frame(x=pop, Distribution=c('Population')), data.frame(x=sample_pop, Distribution=c('Sampling')))
+
+ggplot(data, aes(x=x,color=Distribution)) + geom_density()
+```
 
@@
 Explain.
+**Increase.
+Because as you increase the N the standard error would decrease and the p-value would increase.**

### 2018-10-21 DATA606/Foundations for Inference.Rmd
Minor change
More work on homework

--- 2018-10-16 DATA606/Foundations for Inference.Rmd
@@
 Use a significance level of 0.10.
+**$H_0$ is that x_bar >= 32 and $H_a$ is x_bar < 32**
 
-(c) Interpret the p-value in context of the hypothesis test and the data.
+```{r}
+confidence_level <- 0.10
+n <- 36
+x_bar <- 30.69
+sigma <- 4.31
+mu <- 32
+
+se_x_bar <- sigma / sqrt(n)
+
+z <- (x_bar - mu) / se_x_bar
+
+z_score_for_confidence_level <- 1.645
+
+p_value <- pnorm(z, mean=0, sd=1) * 2
+
+if(z < z_score_for_confidence_level){
+  message(paste(round(z, 3), "is less than", z_score_for_confidence_level, "so reject the null hypothesis (p-value:", round(p_value, 3), ")"))
+} else {
+  message(paste(round(z, 3), "is not less than", z_score_for_confidence_level, "so do not reject the null hypothesis (p-value:", round(p_value, 3), ")"))
+}
+```
+
+(c) Interpret the p-value in context of the hypothesis test and the data.  **There is a 0.068 probability that the difference between these two means was due to chance.**
 
 (d) Calculate a 90% confidence interval for the average age at which gifted children first count to 10 successfully.
+```{r}
+# Same method as http://onlinestatbook.com/2/calculators/inverse_normal_dist.html
+zinv <- function(p) {
+	if (p >= 1 || p <= 0) {
+		return(FALSE)
+	}
+	if (p < .5) {
+		t <- p
+		theSign <- -1
+	}
+	else {
+		t <- 1 - p
+		theSign <- 1
+	}
 
-(e) Do your results from the hypothesis test and the confidence interval agree? Explain.
+	v <- sqrt(-2.0 * log(t))
+	x <- 2.515517 + (v * (0.802853 + v * 0.010328))
+	y <- 1 + (v * (1.432788 + v * (0.189269  + v * 0.001308)))
+	Q <- theSign * (v - (x / y))
+	return(Q)
+}
+
+p <- 1 - confidence_level
+x1 <- x_bar + sigma * zinv(p)
+p2 <- p / 2
+x1 <- zinv(.5 - p2)
+
+error <- sigma * -x1
+ci_low <- x_bar - error
+ci_high <- x_bar + error
+```
+**Confidence interval: (`r ci_low`, `r ci_high`) or `r x_bar` $\pm$ `r error`**
+
+(e) Do your results from the hypothesis test and the confidence interval agree?
+Explain.
+**No.
+Since 32 is inside of the confidence interval you would not reject the null hypotheses.
+It is the opposite conclusion you reach with the hypothesis test, which indicates to me that I did something wrong.**
 
@@
 
+```{r}
+confidence_level <- 0.10
+n <- 36
+x_bar <- 118.2
+sigma <- 6.5
+mu <- 100
+
+se_x_bar <- sigma / sqrt(n)
+
+z <- (x_bar - mu) / se_x_bar
+
+z_score_for_confidence_level <- 1.645
+
+p_value <- pnorm(z, mean=0, sd=1)
+
+if(z < z_score_for_confidence_level){
+  message(paste(round(z, 3), "is less than", z_score_for_confidence_level, "so reject the null hypothesis (p-value:", round(p_value, 3), ")"))
+} else {
+  message(paste(round(z, 3), "is not less than", z_score_for_confidence_level, "so do not reject the null hypothesis (p-value:", round(p_value, 3), ")"))
+}
+```
+
 (b) Calculate a 90% confidence interval for the average IQ of mothers of gifted children.
 
-(c) Do your results from the hypothesis test and the confidence interval agree? Explain.
+```{r}
+p <- 1 - confidence_level
+x1 <- x_bar + sigma * zinv(p)
+p2 <- p / 2
+x1 <- zinv(.5 - p2)
+
+error <- sigma * -x1
+ci_low <- x_bar - error
+ci_high <- x_bar + error
+```
+
+**Confidence interval: (`r ci_low`, `r ci_high`) or `r x_bar` $\pm$ `r error`**
+
+(c) Do your results from the hypothesis test and the confidence interval agree?
+Explain.
+**No.
+100 lies outside the CI which would suggest that we should reject the null hypothesis, but the hypothesis test disagrees.
+Again it would appear there is something wrong with my understanding as both should lead you to the same conclusion.**
 
@@
 
-(b) Describe the distribution of the mean lifespan of 15 light bulbs. **Centered around 9,000 hours.**
+(b) Describe the distribution of the mean lifespan of 15 light bulbs. **Normal and centered around 9,000 hours.**
 
-(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours? **It would be roughly 0.067**
+(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours?
+
+```{r}
+n <- 15
+se <- sd / sqrt(n)
+1 - pnorm(x, mu, se)
+```
 
 (d) Sketch the two distributions (population and sampling) on the same scale.
+```{r, echo=FALSE}
+library(ggplot2)
+set.seed(4)
+pop <- rnorm(1000000, mu, sd)
+sample_pop <- sample(pop, 15)
+data <-  rbind(data.frame(x=pop, Distribution=c('Population')), data.frame(x=sample_pop, Distribution=c('Sampling')))
 
-(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution? **Yes, but not using the above method.**
+ggplot(data, aes(x=x,color=Distribution)) + geom_density() + theme(axis.title.x = element_blank())
+```
+
+(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution? **Not using the above method.**
 

### 2018-10-22 DATA607/Week_9_Assignment.Rmd
Initial commit of Week 9 Assignment

---
title: "CUNY SPS DATA-607 Week 9 Assignment"
author: "Mike Silva"
date: "22 October 2018"
output:
  rmdformats::readthedown:
    highlight: kate
---

## Description

The assignment is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.

## Data Acquisition

I have chosen to pull the top stories from the upshot section of the paper.  I will request the data from the API

```{r}
library(httr)
# Load my API key into api_key
source('hidden/NYT API.R')
url <- paste0('https://api.nytimes.com/svc/topstories/v2/upshot.json?api-key=', api_key)
response <- GET(url)
```

## Data Transformation

Now the JSON response needs to be transformed into a data frame:

```{r, message=FALSE}
library(jsonlite)
json <- fromJSON(content(response, as = "text"))
df <- data.frame(json$results)
```

## Results

The following is the data pulled from the New York Times API:

##### Table 1. Response from the API
```{r, echo=FALSE, message=FALSE}
library(kableExtra)
library(dplyr)
df %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

### 2018-10-22 DATA607/Week_9_Assignment.Rmd
Modified Week_9_Assignment.Rmd
Linted document and made minor changes

@@
 library(httr)
-# Load my API key into api_key
-source('hidden/NYT API.R')
-url <- paste0('https://api.nytimes.com/svc/topstories/v2/upshot.json?api-key=', api_key)
+source("hidden/NYT API.R") # Load my API key into api_key
+base_url <- "https://api.nytimes.com/svc/topstories/v2/upshot.json"
+url <- paste0(base_url, "?api-key=", api_key)
 response <- GET(url)
@@
 json <- fromJSON(content(response, as = "text"))
-df <- data.frame(json$results)
+df <- data.frame( json$results )
 ```

### 2018-10-23 DATA606/Inference for Numerical Data.Rmd
Initial Commit of Inference for Numerical Data

---
title: "Inference for Numerical Data"
author: "Mike Silva"
date: "October 22, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
---

## 5.6 Working backwards, Part II

A 90% confidence interval for a population mean is (65, 77).  The population distribution is approximately normal and the population standard deviation is unknown.  This confidence interval is based on a simple random sample of 25 observations.  Calculate the sample mean, the margin of error, and the sample standard deviation.

## 5.14 SAT scores

SAT scores of students at an Ivy League college are distributed with a standard deviation of 250 points. Two statistics students, Raina and Luke, want to estimate the
average SAT score of students at this college as part of a class project. They want their margin of error to be no more than 25 points.

(a) Raina wants to use a 90% confidence interval. How large a sample should she collect?

(b) Luke wants to use a 99% confidence interval. Without calculating the actual sample size, determine whether his sample should be larger or smaller than Raina’s, and explain your reasoning.

(c) Calculate the minimum required sample size for Luke.

## 5.20 High School and Beyond, Part I

The National Center of Education Statistics conducted a survey of high school seniors, collecting test data on reading, writing, and several other subjects.  Here we examine a simple random sample of 200 students from this survey.  Side-by-side box plots of reading and writing scores as well as a histogram of the differences in scores are shown below.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/5_20.png)

(a) Is there a clear difference in the average reading and writing scores?

(b) Are the reading and writing scores of each student independent of each other?

(c) Create hypotheses appropriate for the following research question: is there an evident difference in the average scores of students in the reading and writing exam?

(d) Check the conditions required to complete this test.

(e) The average observed difference in scores is $\bar{x}_{read−write}$ = −0.545, and the standard deviation of the differences is 8.887 points.  Do these data provide convincing evidence of a difference between the average scores on the two exams?

(f) What type of error might we have made? Explain what the error means in the context of the application.

(g) Based on the results of this hypothesis test, would you expect a confidence interval for the average difference between the reading and writing scores to include 0? Explain your reasoning.

## 5.32 Fuel efficiency of manual and automatic cars, Part I

Each year the US Environmental Protection Agency (EPA) releases fuel economy data on cars manufactured in that year.  Below are summary statistics on fuel efficiency (in miles/gallon) from random samples of cars with manual and automatic transmissions manufactured in 2012.  Do these data provide strong evidence of a difference between the average fuel efficiency of cars with manual and automatic transmissions in terms of their average city mileage?  Assume that conditions for inference are satisfied.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/5_32.png)

## 5.48 Work hours and education

The General Social Survey collects data on demographics, education, and work, among many other characteristics of US residents.  Using ANOVA, we can consider educational attainment levels for all 1,172 respondents at once.  Below are the distributions
of hours worked by educational attainment and relevant summary statistics that will be helpful in carrying out this analysis.

(a) Write hypotheses for evaluating whether the average number of hours worked varies across the five groups.

(b) Check conditions and describe any assumptions you must make to proceed with the test.

(c) Below is part of the output associated with this test. Fill in the empty cells.

|           |  Df | Sum Sq  | Mean Sq | F value | Pr(>F) |
|-----------|-----|---------|---------|---------|--------|
| degree    |** **| ** **   | 501.54  | ** **   | 0.0682 |
| Residuals |** **| 267,382 | ** **   |         |        |
| Total     |** **| ** **   |         |         |        |

(d) What is the conclusion of the test?

### 2018-10-29 DATA606/Lab5/michael-inf_for_numerical_data.Rmd
Finished Lab 5

---
title: 'Inference for numerical data'
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## North Carolina births

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

## Exploratory analysis

Load the `nc` data set into our workspace.

```{r load-data, eval=TRUE}
load("more/nc.RData")
```

We have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description
---------------- | -----------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

1.  What are the cases in this data set? **Live births** How many cases are there in our sample? **`r nrow(nc)`**

As a first step in the analysis, we should consider summaries of the data. This
can be done using the `summary` command:

```{r summary, eval=TRUE}
summary(nc)
```

As you review the variable summaries, consider which variables are categorical 
and which are numerical. For numerical variables, are there outliers? If you 
aren't sure or want to take a closer look at the data, make a graph.

Consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

2.  Make a side-by-side boxplot of `habit` and `weight`. What does the plot 
highlight about the relationship between these two variables? 
```{r echo=FALSE, message=FALSE}
library(ggplot2)
ggplot(nc, aes(x = habit, y = weight)) + geom_boxplot()
```

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following function to 
split the `weight` variable into the `habit` groups, then take the mean of each
using the `mean` function.

```{r by-means, eval=TRUE}
by(nc$weight, nc$habit, mean)
```

There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis test.

## Inference

3.  Check if the conditions necessary for inference are satisfied. Note that 
you will need to obtain sample sizes to check the conditions. You can compute 
the group size using the same `by` command above but replacing `mean` with 
`length`. **Sample silzes are significantly large.**
```{r}
by(nc$weight, nc$habit, length)
```

4.  Write the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different. **They are $H_0 = \mu_{nonsmoker} - \mu_{smoker} = 0$ and $H_A = \mu_{nonsmoker} - \mu_{smoker} != 0$**

Next, we introduce a new function, `inference`, that we will use for conducting
hypothesis tests and constructing confidence intervals.

```{r inf-weight-habit-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

Let's pause for a moment to go through the arguments of this custom function. 
The first argument is `y`, which is the response variable that we are 
interested in: `nc$weight`. The second argument is the explanatory variable, 
`x`, which is the variable that splits the data into two groups, smokers and 
non-smokers: `nc$habit`. The third argument, `est`, is the parameter we're 
interested in: `"mean"` (other options are `"median"`, or `"proportion"`.) Next
we decide on the `type` of inference we want: a hypothesis test (`"ht"`) or a 
confidence interval (`"ci"`). When performing a hypothesis test, we also need 
to supply the `null` value, which in this case is `0`, since the null 
hypothesis sets the two population means equal to each other. The `alternative`
hypothesis can be `"less"`, `"greater"`, or `"twosided"`. Lastly, the `method` 
of inference can be `"theoretical"` or `"simulation"` based.

5.  Change the `type` argument to `"ci"` to construct and record a confidence 
interval for the difference between the weights of babies born to smoking and 
non-smoking mothers.
```{r}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```

By default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$)
. We can easily change this order by using the `order` argument:

```{r inf-weight-habit-ci, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```

* * *

## On your own

1.   Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) and interpret it in context. Note that since you're doing inference 
on a single population parameter, there is no explanatory variable, so you can 
omit the `x` variable from the function.

```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```

2.   Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.
```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", conflevel = 0.90)
```

3.   Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.
```{r}
inference(y = nc$weight, x = nc$mature, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```

4.   Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works. **My method looks at the oldest young mother and the youngest mature mother.  It looks like 35 is the cutoff.**
```{r}
young_mothers <- nc[nc$mature == "younger mom",]
mature_mothers <- nc[nc$mature != "younger mom",]
max(young_mothers$mage)
min(mature_mothers$mage)
```

5.   Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language. **Does the amount of weight gained depend on the gender of the baby.  The null hypothesis is that the baby's gender does not have any impact on the weight gained ($H_0 = \mu_{boy} - \mu_{girl} = 0$).  The alternative is that the gender of the baby has an effect on the amount of weight gained ($H_0 = \mu_{boy} - \mu_{girl} != 0$).  There is no evidence to suggest that the gender of the child has any bearing on the amount of weight gained by the mother.**

```{r}
inference(y = nc$weight, x = nc$gender, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab 
written by the faculty and TAs of UCLA Statistics.
</div>

### 2018-10-30 DATA606/Inference for Numerical Data.Rmd
More work

@@
 
+```{r}
+# Sample mean is midpoint of the confidence interval endpoints
+x_bar <- mean(c(65,77))
+
+# Margin of error is the difference between the midpoint and the CI endpoints
+moe <- 77 - x_bar
+```
+
+$`r x_bar` \pm `r moe` = \bar{x} \pm Z \frac{s}{\sqrt{n}}$
+
+Given z = 1.645 and n = 25
+
+$`r moe` = 1.645 \frac{s}{\sqrt{25}} \rightarrow `r moe` = 1.645 \frac{s}{5} \rightarrow `r 5*moe` = 8.225 \times s \rightarrow s = \frac{`r 5*moe`}{8.225} \rightarrow s = `r (5*moe)/8.225`$
+
+```{r}
+# Given
+z <- 1.645 # 90% confidence interval z-score
+n <- 25
+
+# Sample standard deviation (s)
+s <- (sqrt(n) * moe) / (sqrt(n) * z)
+```
+
+**Sample mean: `r x_bar`;  Margin of Error: `r moe`; Sample Standard Deviation: `r s`;**
+
 ## 5.14 SAT scores
@@
 (a) Raina wants to use a 90% confidence interval. How large a sample should she collect? 
+```{r 90_ci_sample_size_Q5_14}
+z <- 1.645 # 90% confidence interval z-score
+sigma <- 250
+moe <- 25
 
-(b) Luke wants to use a 99% confidence interval. Without calculating the actual sample size, determine whether his sample should be larger or smaller than Raina’s, and explain your reasoning.
+n <- ((z*sigma)/moe)^2
+```
+
+**Sample Size: `r n`**
+
+$25 = 1.645 \frac{250}{\sqrt{n}} \rightarrow 25 = \frac{411.25}{\sqrt{n}} \rightarrow 1 = \frac{16.45}{\sqrt{n}} \rightarrow \sqrt{n} = 16.45 \rightarrow n = \pm 270.6025$
+
+(b) Luke wants to use a 99% confidence interval.
+Without calculating the actual sample size, determine whether his sample should be larger or smaller than Raina’s, and explain your reasoning.
+**Larger.
+In order to have a higher confidence we would need a larger sample size.**
 
 (c) Calculate the minimum required sample size for Luke.
+collect?
+
+```{r 99_ci_sample_size_Q5_14}
+z <- 2.58 # 99% confidence interval z-score
+sigma <- 250
+moe <- 25
+
+n <- ((z*sigma)/moe)^2
+```
+
+**Sample Size: `r n`**
 
@@
 
-(a) Is there a clear difference in the average reading and writing scores?
+(a) Is there a clear difference in the average reading and writing scores? **Yes the reading score appears to be lower with more variation than the writing score.**
 
-(b) Are the reading and writing scores of each student independent of each other?
+(b) Are the reading and writing scores of each student independent of each other? **No because the same student takes both tests.**
 
 (c) Create hypotheses appropriate for the following research question: is there an evident difference in the average scores of students in the reading and writing exam?
+**$H_0 = \bar{x}_\text{read} - \bar{x}_\text{write} = 0$  and $H_A = \bar{x}_\text{read} - \bar{x}_\text{write} != 0$**
 
-(d) Check the conditions required to complete this test.
+(d) Check the conditions required to complete this test. **Both distributions appear to be normally distributed and centered around zero.  The number of observations is larger than 30.**
 
@@
 Do these data provide convincing evidence of a difference between the average scores on the two exams?
+**The p-value suggests we should not reject the null hypothesis**
 
-(f) What type of error might we have made? Explain what the error means in the context of the application.
+```{r}
+x_bar <- -0.545
+sigma <- 8.887
+n <- 200
+z <- (x_bar)/(sigma/sqrt(n))
+1.96 * pnorm(-abs(z))
+```
+
+(f) What type of error might we have made? Explain what the error means in the context of the application.**A type II error.  We beleive there isn't a difference when there is.**
 
 (g) Based on the results of this hypothesis test, would you expect a confidence interval for the average difference between the reading and writing scores to include 0? Explain your reasoning.
+
+**Yes.  Using an confidence level of 95%, the margin of error is:**
+
+$MOE = Z\frac{s}{\sqrt{n}} \rightarrow MOE = 1.96\frac{8.887}{\sqrt{200}} \rightarrow MOE = 1.231675$
+
+**Therefore the confidence interval is `r -0.545 - 1.231675` to `r -0.545 + 1.231675` which contains zero, so we cannot reject the null hypothesis.**
 
@@
 
+```{r}
+x_bar_automatic <- 16.12
+s_automatic <- 3.58
+n_automatic <- 26
+x_bar_manual <- 19.85
+s_manual <- 4.51
+n_manual <- 26
+point_estimate <- x_bar_automatic - x_bar_manual
+df <- n_automatic - 1
+se <- sqrt(((s_automatic^2)/n_automatic) + ((s_manual^2)/n_manual))
+t_value <- point_estimate/se
+critical_t <- pt(t_value, df)
+
+c(point_estimate + (critical_t * 1.95), point_estimate - (critical_t * 1.95)) 
+```
+
+**The difference is roughly -3.727 and -3.733 mpg.**
+
 ## 5.48 Work hours and education
@@
 
+![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/5_48.png)
+
 (a) Write hypotheses for evaluating whether the average number of hours worked varies across the five groups.
+**H0: $\mu$1 = $\mu$2 = $\mu$3 = $\mu$4 = $\mu$5 and HA: The average varies across some or all groups.**
 
-(b) Check conditions and describe any assumptions you must make to proceed with the test.
+(b) Check conditions and describe any assumptions you must make to proceed with the test. **Independent, normaly distributed observations across groups.**
 
@@
 |           |    Df   | Sum Sq  | Mean Sq | F value | Pr(>F) |
-|-----------|-----|---------|---------|---------|--------|
+|-----------|---------|---------|---------|---------|--------|
 | degree    |**     **| ** **   | 501.54  | ** **   | 0.0682 |
 | Residuals |**     **| 267,382 | ** **   |         |        |
-| Total     |** **| ** **   |         |         |        |
+| Total     |**1,172**| ** **   |         |         |        |
 
-(d) What is the conclusion of the test?
+(d) What is the conclusion of the test? **We cannot reject the null hypothesis.**

### 2018-10-30 DATA606/Inference for Numerical Data.Rmd
More work

@@
 (a) Write hypotheses for evaluating whether the average number of hours worked varies across the five groups.
-**H0: $\mu$1 = $\mu$2 = $\mu$3 = $\mu$4 = $\mu$5 and HA: The average varies across some or all groups.**
+**H0: $\mu_1$ = $\mu_2$ = $\mu_3$ = $\mu_4$ = $\mu_5$ and HA: The average varies across some or all groups.**
 
@@
 |-----------|---------|---------|---------|---------|--------|
-| degree    |**     **| ** **   | 501.54  | ** **   | 0.0682 |
-| Residuals |**     **| 267,382 | ** **   |         |        |
-| Total     |**1,172**| ** **   |         |         |        |
+| degree    |         |         | 501.54  |         | 0.0682 |
+| Residuals |         | 267,382 |         |         |        |
+| Total     |**1,172**|         |         |         |        |
 

### 2018-10-30 DATA606/Inference for Numerical Data.Rmd
Minor correction

@@
 (c) Below is part of the output associated with this test. Fill in the empty cells.
+```{r}
+mu <- c(38.67, 39.6, 41.39, 42.55, 40.85)
+sd <- c(15.81, 14.97, 18.1, 13.62, 15.51)
+n <- c(121, 546, 97, 253, 155)
+
+n <- 1172
+k <- 5
+Prf <- 0.0682
+f_value <- qf( 1 - Prf, k-1 , n-k)
+
+# Get missing values
+sum_sq <- 501.54 * (k-1)
+mean_sq <- 267382 / (n-k)
+total <- sum_sq + 267382
+```
 
 |           |      Df     |    Sum Sq    |    Mean Sq    |    F value     | Pr(>F) |
-|-----------|---------|---------|---------|---------|--------|
-| degree    |         |         | 501.54  |         | 0.0682 |
-| Residuals |         | 267,382 |         |         |        |
-| Total     |**1,172**|         |         |         |        |
+|-----------|-------------|--------------|---------------|----------------|--------|
+| degree    |**`r k-1 ` **|**`r sum_sq`**|    501.54     | **`r f_value`**| 0.0682 |
+| Residuals |**`r n-k` ** |    267,382   |**`r mean_sq`**|                |        |
+| Total     |  **1,172**  |**269388.16** |               |                |        |
 

### 2018-10-30 DATA607/Project_4.Rmd
Initial commit

---
title: "Document Classification"
date: "29 October 2018"
output:
  rmdformats::readthedown:
    highlight: kate
---

## Description

This project seeks to build a classifier that can predict if an email is spam or ham.  It will be using the Apache Spam Assassin public corpus as the dataset.

### Data Aquisition

The first step is to download the files.  I am selecting the most recent releases of spam and easy ham.

```{r file_download}
if(!file.exists("data/20050311_spam_2.tar.bz2") && !file.exists("data/20050311_spam_2.tar")){
  download.file("https://spamassassin.apache.org/old/publiccorpus/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar.bz2")
}

if(!file.exists("data/20030228_easy_ham_2.tar.bz2") & !file.exists("data/20030228_easy_ham_2.tar")){
  download.file("https://spamassassin.apache.org/old/publiccorpus/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar.bz2")
}
```

Now that the files have been aquired, they need to be decompressed.

```{r decompress_files}
if(!dir.exists("data/spam_2")){
  R.utils::bunzip2("data/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar")
  untar("data/20050311_spam_2.tar", exdir = "data")
  # Delete the cmds file
  unlink("data/spam_2/cmds")
}

if(!dir.exists("data/easy_ham_2")){
  R.utils::bunzip2("data/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar")
  untar("data/20030228_easy_ham_2.tar", exdir = "data")
  # Delete the cmds file
  unlink("data/easy_ham_2/cmds")
}
```

## Approach

```{r}
spam_docs <- list.files("data/spam_2")
ham_docs <- list.files("data/easy_ham_2")
```

## Processing Pipeline

### Split up Sentance

### Remove Punctuation and Stop Words

### Stemming

### 2018-10-30 DATA606/Inference for Categorical Data.Rmd
Initial commit of Homework

---
title: "Inference for Categorical Data"
author: "Mike Silva"
date: "October 30, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
---

## 6.6 2010 Healthcare Law

On June 28, 2012 the U.S. Supreme Court upheld the much debated 2010 healthcare law, declaring it constitutional. A Gallup poll released the day after this decision
indicates that 46% of 1,012 Americans agree with this decision. At a 95% confidence level, this sample has a 3% margin of error. Based on this information, determine if the following statements are true or false, and explain your reasoning.

(a) We are 95% confident that between 43% and 49% of Americans in this sample support the decision of the U.S. Supreme Court on the 2010 healthcare law.

(b) We are 95% confident that between 43% and 49% of Americans support the decision of the U.S. Supreme Court on the 2010 healthcare law.

(c) If we considered many random samples of 1,012 Americans, and we calculated the sample proportions of those who support the decision of the U.S. Supreme Court, 95% of those sample proportions will be between 43% and 49%.

(d) The margin of error at a 90% confidence level would be higher than 3%.

## 6.12 Legalization of marijuana, Part I

The 2010 General Social Survey asked 1,259 US residents: “Do you think the use of marijuana should be made legal, or not?” 48% of the respondents said it should be made legal.

(a) Is 48% a sample statistic or a population parameter? Explain.

(b) Construct a 95% confidence interval for the proportion of US residents who think marijuana should be made legal, and interpret it in the context of the data.

(c) A critic points out that this 95% confidence interval is only accurate if the statistic follows a normal distribution, or if the normal model is a good approximation. Is this true for these data? Explain.

(d) A news piece on this survey’s findings states, “Majority of Americans think marijuana should be legalized.” Based on your confidence interval, is this news piece’s statement justified?

## 6.20 Legalize Marijuana, Part II

As discussed in Exercise 6.12, the 2010 General Social Survey reported a sample where about 48% of US residents thought marijuana should be made legal. If we wanted to limit the margin of error of a 95% confidence interval to 2%, about how many Americans would we need to survey ?

## 6.28 Sleep deprivation, CA vs. OR, Part I

According to a report on sleep deprivation by the Centers for Disease Control and Prevention, the proportion of California residents who reported insufficient rest or sleep during each of the preceding 30 days is 8.0%, while this proportion is 8.8% for Oregon residents. These data are based on simple random samples of 11,545 California and 4,691 Oregon residents. Calculate a 95% confidence interval for the difference between the proportions of Californians and Oregonians who are sleep deprived and interpret it in context of the data.

## 6.44 Barking deer

Microhabitat factors associated with forage and bed sites of barking deer in Hainan Island, China were examined from 2001 to 2002. In this region woods make up 4.8% of the land, cultivated grass plot makes up 14.7%, and deciduous forests makes up 39.6%. Of the 426 sites where the deer forage, 4 were categorized as woods, 16 as cultivated grassplot, and 61 as deciduous forests. The table below summarizes these data.

| Woods | Cultivated grassplot | Deciduous forests | Other | Total |
|-------|----------------------|-------------------|-------|-------|
| 4     | 16                   | 61                | 345   | 426   |

(a) Write the hypotheses for testing if barking deer prefer to forage in certain habitats over others.

(b) What type of test can we use to answer this research question?

(c) Check if the assumptions and conditions required for this test are satisfied.

(d) Do these data provide convincing evidence that barking deer prefer to forage in certain habitats over others? Conduct an appropriate hypothesis test to answer this research question.

## 6.48 Coffee and Depression.

Researchers conducted a study investigating the relationship between caffeinated coffee consumption and risk of depression in women. They collected data on 50,739 women free of depression symptoms at the start of the study in the year 1996, and these women were followed through 2006. The researchers used questionnaires to collect data on caffeinated coffee consumption, asked each individual about physician-diagnosed depression, and also asked about the use of antidepressants. The table below shows the distribution of incidences of depression by amount of caffeinated coffee consumption.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/6_48.png)

(a) What type of test is appropriate for evaluating if there is an association between coffee intake and depression?

(b) Write the hypotheses for the test you identified in part (a).

(c) Calculate the overall proportion of women who do and do not suffer from depression.

(d) Identify the expected count for the highlighted cell, and calculate the contribution of this cell to the test statistic, i.e. $\frac{(\text{Observed} − \text{Expected})^2}{\text{Expected}}$.

(e) The test statistic is $\chi^2$ = 20.93. What is the p-value?

(f) What is the conclusion of the hypothesis test?

(g) One of the authors of this study was quoted on the NYTimes as saying it was “too early to recommend that women load up on extra coffee” based on just this study.64 Do you agree with this statement? Explain your reasoning.

### 2018-10-30 DATA607/Project_4.Rmd
More work on Project 4

@@
 
-## Approach
+## Processing Pipeline
+
+### Train/Test Split
+
+First we are going to split the data into training and testing sets using an 80:20 split.
 
 ```{r}
-spam_docs <- list.files("data/spam_2")
-ham_docs <- list.files("data/easy_ham_2")
+set.seed(12345)
+training_percent <- 0.8
+
+train_test_split <- function(data, training_percent){
+  n <- round(length(data)*training_percent, digits = 0)
+  train <- sort(sample(data, n, replace = FALSE))
+  test <- data[!data %in% train]
+  return(list(test=test, train=train))
+}
+
+spam <- train_test_split(list.files("data/spam_2"), training_percent)
+ham <- train_test_split(list.files("data/easy_ham_2"), training_percent)
 ```
 
-## Processing Pipeline
+The following table summarizes the number of messages included in each set:
+
+##### Table 1. Spam and Ham Message Counts by Data Set
+|      |       Training       |        Testing      |
+|------|----------------------|---------------------|
+| Spam |`r length(spam$train)`|`r length(spam$test)`|
+| Ham  |`r length(ham$train)` |`r length(ham$test)` |
+
+### Pull out Email Contents
+
+This model will classify only on the basis of the content.
+The header information will not be taken into consideration.
+In order to illustrate the process I will take the first message in the spam training set as an example.
+
+```{r}
+example_file <- spam$train[1]
+
+get_email_contents <- function(file_name, path){
+ full_path = paste0(path, file_name)
+ con <- file(full_path)
+ readLines(con)
+}
+
+```
 

### 2018-10-31 DATA606/Lab6/michael-inf_for_categorical_data.Rmd
Modified Lab 6

---
title: 'Inference for categorical data'
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

In August of 2012, news outlets ranging from the [Washington
Post](http://www.washingtonpost.com/national/on-faith/poll-shows-atheism-on-the-rise-in-the-us/2012/08/13/90020fd6-e57d-11e1-9739-eef99c5fb285_story.html) to the [Huffington
Post](http://www.huffingtonpost.com/2012/08/14/atheism-rise-religiosity-decline-in-america_n_1777031.html)
ran a story about the rise of atheism in America. The source for the story was 
a poll that asked people, "Irrespective of whether you attend a place of 
worship or not, would you say you are a religious person, not a religious 
person or a convinced atheist?" This type of question, which asks people to 
classify themselves in one way or another, is common in polling and generates 
categorical data. In this lab we take a look at the atheism survey and explore 
what's at play when making inference about population proportions using 
categorical data.

## The survey

To access the press release for the poll, conducted by WIN-Gallup 
International, click on the following link:

*<https://github.com/jbryer/DATA606/blob/master/inst/labs/Lab6/more/Global_INDEX_of_Religiosity_and_Atheism_PR__6.pdf>*

Take a moment to review the report then address the following questions.

1.  In the first paragraph, several key findings are reported. Do these 
    percentages appear to be *sample statistics* (derived from the data 
    sample) or *population parameters*? **They read like they are population parameters, however Table 6 states it's the percent of respondents.  It appears that it is sample statistics.**

2.  The title of the report is "Global Index of Religiosity and Atheism". To
    generalize the report's findings to the global human population, what must 
    we assume about the sampling method? Does that seem like a reasonable 
    assumption?  **The population represents a random sample, independent, and sufficiently large to represent the population of their country. Since they were surveyed via face-to-face interviews, telephone or online it seems they did not exclude a population.**

## The data

Turn your attention to Table 6 (pages 15 and 16), which reports the
sample size and response percentages for all 57 countries. While this is
a useful format to summarize the data, we will base our analysis on the
original data set of individual responses to the survey. Load this data
set into R with the following command.

```{r head-data, eval=TRUE}
load("more/atheism.RData")
```

3.  What does each row of Table 6 correspond to? What does each row of 
    `atheism` correspond to?**A response to the survey.**

To investigate the link between these two ways of organizing this data, take a 
look at the estimated proportion of atheists in the United States. Towards 
the bottom of Table 6, we see that this is 5%. We should be able to come to 
the same number using the `atheism` data.

4.  Using the command below, create a new dataframe called `us12` that contains
    only the rows in `atheism` associated with respondents to the 2012 survey 
    from the United States. Next, calculate the proportion of atheist 
    responses. Does it agree with the percentage in Table 6? **Yes** If not, why?

```{r us-atheism, eval=TRUE, message=FALSE}
library(dplyr)
library(kableExtra)

us12 <- subset(atheism, nationality == "United States" & year == "2012")

us12 %>%
  group_by(response) %>%
  summarise(count = n(), total = nrow(us12)) %>%
  mutate(share = round(count / total, 2)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

## Inference on proportions

As was hinted at in Exercise 1, Table 6 provides *statistics*, that is, 
calculations made from the sample of 51,927 people. What we'd like, though, is 
insight into the population *parameters*. You answer the question, "What 
proportion of people in your sample reported being atheists?" with a 
statistic; while the question "What proportion of people on earth would report 
being atheists" is answered with an estimate of the parameter.

The inferential tools for estimating population proportion are analogous to 
those used for means in the last chapter: the confidence interval and the 
hypothesis test.

5.  Write out the conditions for inference to construct a 95% confidence
    interval for the proportion of atheists in the United States in 2012.
    Are you confident all conditions are met? **The sample size is arround 1,000 which is the right size (less than 10% of the population).**

If the conditions for inference are reasonable, we can either calculate
the standard error and construct the interval by hand, or allow the `inference`
function to do it for us.

```{r us-atheism-ci, eval=TRUE, tidy = FALSE}
inference(us12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
```

Note that since the goal is to construct an interval estimate for a 
proportion, it's necessary to specify what constitutes a "success", which here 
is a response of `"atheist"`.

Although formal confidence intervals and hypothesis tests don't show up in the 
report, suggestions of inference appear at the bottom of page 7: "In general, 
the error margin for surveys of this kind is $\pm$ 3-5% at 95% confidence".

6.  Based on the R output, what is the margin of error for the estimate of the 
    proportion of the proportion of atheists in US in 2012? **`r 0.0499 - 0.0364`**

7.  Using the `inference` function, calculate confidence intervals for the 
    proportion of atheists in 2012 in two other countries of your choice, and 
    report the associated margins of error. Be sure to note whether the 
    conditions for inference are met. It may be helpful to create new data 
    sets for each of the two countries first, and then use these data sets in 
    the `inference` function to construct the confidence intervals.
    ```{r}
    china_12 <- subset(atheism, nationality == "China" & year == "2012") 
    inference(china_12$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
    ```

    ```{r}
    fiji_12 <- subset(atheism, nationality == "Fiji" & year == "2012") 
    inference(fiji_12$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
    ```

## How does the proportion affect the margin of error?

Imagine you've set out to survey 1000 people on two questions: are you female? 
and are you left-handed? Since both of these sample proportions were 
calculated from the same sample size, they should have the same margin of 
error, right? Wrong! While the margin of error does change with sample size, 
it is also affected by the proportion.

Think back to the formula for the standard error: $SE = \sqrt{p(1-p)/n}$. This 
is then used in the formula for the margin of error for a 95% confidence 
interval: $ME = 1.96\times SE = 1.96\times\sqrt{p(1-p)/n}$. Since the 
population proportion $p$ is in this $ME$ formula, it should make sense that 
the margin of error is in some way dependent on the population proportion. We 
can visualize this relationship by creating a plot of $ME$ vs. $p$.

The first step is to make a vector `p` that is a sequence from 0 to 1 with 
each number separated by 0.01. We can then create a vector of the margin of 
error (`me`) associated with each of these values of `p` using the familiar 
approximate formula ($ME = 2 \times SE$). Lastly, we plot the two vectors 
against each other to reveal their relationship.

```{r me-plot, eval=TRUE}
n <- 1000
p <- seq(0, 1, 0.01)
me <- 2 * sqrt(p * (1 - p)/n)
plot(me ~ p, ylab = "Margin of Error", xlab = "Population Proportion")
```

8.  Describe the relationship between `p` and `me`. **It is a semi-cirle ranging from 0 to 1 with the maximum at 0.5.**

## Success-failure condition

The textbook emphasizes that you must always check conditions before making 
inference. For inference on proportions, the sample proportion can be assumed 
to be nearly normal if it is based upon a random sample of independent 
observations and if both $np \geq 10$ and $n(1 - p) \geq 10$. This rule of 
thumb is easy enough to follow, but it makes one wonder: what's so special 
about the number 10?

The short answer is: nothing. You could argue that we would be fine with 9 or 
that we really should be using 11. What is the "best" value for such a rule of 
thumb is, at least to some degree, arbitrary. However, when $np$ and $n(1-p)$ 
reaches 10 the sampling distribution is sufficiently normal to use confidence 
intervals and hypothesis tests that are based on that approximation.

We can investigate the interplay between $n$ and $p$ and the shape of the 
sampling distribution by using simulations. To start off, we simulate the 
process of drawing 5000 samples of size 1040 from a population with a true 
atheist proportion of 0.1. For each of the 5000 samples we compute $\hat{p}$ 
and then plot a histogram to visualize their distribution.

```{r sim-np, eval=TRUE}
p <- 0.1
n <- 1040
p_hats <- rep(0, 5000)

for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  p_hats[i] <- sum(samp == "atheist")/n
}

hist(p_hats, main = "p = 0.1, n = 1040", xlim = c(0, 0.18))
```

These commands build up the sampling distribution of $\hat{p}$ using the 
familiar `for` loop. You can read the sampling procedure for the first line of 
code inside the `for` loop as, "take a sample of size $n$ with replacement 
from the choices of atheist and non-atheist with probabilities $p$ and $1 - p$,
respectively." The second line in the loop says, "calculate the proportion of 
atheists in this sample and record this value." The loop allows us to repeat 
this process 5,000 times to build a good representation of the sampling 
distribution.

9.  Describe the sampling distribution of sample proportions at $n = 1040$ and 
    $p = 0.1$. Be sure to note the center, spread, and shape.\
    *Hint:* Remember that R has functions such as `mean` to calculate summary
    statistics. **Symetrical, centered at `r mean(p_hats)` and ranging from `r min(p_hats)` to `r max(p_hats)`.**

10. Repeat the above simulation three more times but with modified sample
    sizes and proportions: for $n = 400$ and $p = 0.1$, $n = 1040$ and 
    $p = 0.02$, and $n = 400$ and $p = 0.02$. Plot all four histograms
    together by running the `par(mfrow = c(2, 2))` command before creating the 
    histograms. You may need to expand the plot window to accommodate the 
    larger two-by-two plot. Describe the three new sampling distributions. 
    Based on these limited plots, how does $n$ appear to affect the 
    distribution of $\hat{p}$? How does $p$ affect the sampling distribution?
    ```{r}
    par(mfrow = c(2, 2))
    hist(p_hats, main = "p = 0.1, n = 1040", xlim = c(0, 0.18))

    p <- 0.1
    n <- 400
    p_hats2 <- rep(0, 5000)
    for(i in 1:5000){
      samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
      p_hats2[i] <- sum(samp == "atheist")/n
    }
    hist(p_hats2, main = "p = 0.1, n = 400", xlim = c(0, 0.18))

    p <- 0.02
    n <- 1040
    p_hats3 <- rep(0, 5000)
    for(i in 1:5000){
      samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
      p_hats3[i] <- sum(samp == "atheist")/n
    }
    hist(p_hats3, main = "p = 0.02, n = 1040", xlim = c(0, 0.18))

    n <- 400
    p_hats4 <- rep(0, 5000)
    for(i in 1:5000){
      samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
      p_hats4[i] <- sum(samp == "atheist")/n
    }
    hist(p_hats4, main = "p = 0.02, n = 400", xlim = c(0, 0.18))
    ```

    |   n  |  p   |      Center       |                Range                | SD |
    |------|------|-------------------|-------------------------------------|----|
    | 1040 | 0.1  | `r mean(p_hats)`  | `r min(p_hats)` to `r max(p_hats)`  |`r sd(p_hats)`|
    | 400  | 0.1  | `r mean(p_hats2)` | `r min(p_hats2)` to `r max(p_hats2)`|`r sd(p_hats2)`|
    | 1040 | 0.02 | `r mean(p_hats3)` | `r min(p_hats3)` to `r max(p_hats3)`|`r sd(p_hats3)`|
    | 400  | 0.02 | `r mean(p_hats4)` | `r min(p_hats4)` to `r max(p_hats4)`|`r sd(p_hats4)`|

    **The larger the n the closer to the p the center is and the smaller the standard deviation is.  The p shifts the histogram along the x axis.**

```{r, echo=FALSE}
par(mfrow = c(1, 1))
```
Once you're done, you can reset the layout of the plotting window by using the 
command `par(mfrow = c(1, 1))` command or clicking on "Clear All" above the 
plotting window (if using RStudio). Note that the latter will get rid of all 
your previous plots.

11. If you refer to Table 6, you'll find that Australia has a sample 
    proportion of 0.1 on a sample size of 1040, and that Ecuador has a sample 
    proportion of 0.02 on 400 subjects. Let's suppose for this exercise that 
    these point estimates are actually the truth. Then given the shape of 
    their respective sampling distributions, do you think it is sensible to 
    proceed with inference and report margin of errors, as the reports does?**No because the sample margin of error is not the same thing as the population margin of error.**

* * *
## On your own

The question of atheism was asked by WIN-Gallup International in a similar 
survey that was conducted in 2005. (We assume here that sample sizes have 
remained the same.) Table 4 on page 13 of the report summarizes survey results 
from 2005 and 2012 for 39 countries.

-   Answer the following two questions using the `inference` function. As 
    always, write out the hypotheses for any tests you conduct and outline the
    status of the conditions for inference.

    **a.** Is there convincing evidence that Spain has seen a change in its 
    atheism index between 2005 and 2012?\
    *Hint:* Create a new data set for respondents from Spain. Form
    confidence intervals for the true proportion of athiests in both
    years, and determine whether they overlap. **H0: $\text{Index}_2005$ = $\text{Index}_2012$ and Ha: $\text{Index}_2005$ != $\text{Index}_2012$**
    ```{r}
    spain_05 <- subset(atheism, nationality == "Spain" & year == "2005") 
    inference(spain_05$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")

    spain_12 <- subset(atheism, nationality == "Spain" & year == "2012") 
    inference(spain_12$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
    ```
    **The confidence intervals overlap so we cannot reject the null hypothesis, therefore there is not convinsing evidence that Spain's atheism index has changed.**

    **b.** Is there convincing evidence that the United States has seen a
    change in its atheism index between 2005 and 2012? **H0: $\text{Index}_2005$ = $\text{Index}_2012$ and Ha: $\text{Index}_2005$ != $\text{Index}_2012$**
    ```{r}
    us_05 <- subset(atheism, nationality == "United States" & year == "2005") 
    inference(us_05$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")

    us_12 <- subset(atheism, nationality == "United States" & year == "2012") 
    inference(us_12$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
    ```
    **There is no overlap in the confidence intervals which suggests that there is evidence that the United States has experienced a change in the proportion of the population claiming to be atheist between 2005 and 2012.**

-   If in fact there has been no change in the atheism index in the countries 
    listed in Table 4, in how many of those countries would you expect to 
    detect a change (at a significance level of 0.05) simply by chance?\
    *Hint:* Look in the textbook index under Type 1 error. **There is 39 countries on the list so we would expect 2 (=39 $\times$ 0.05) countries have a change detected simply by chance.**

-   Suppose you're hired by the local government to estimate the proportion of 
    residents that attend a religious service on a weekly basis. According to 
    the guidelines, the estimate must have a margin of error no greater than 
    1% with 95% confidence. You have no idea what to expect for $p$. How many 
    people would you have to sample to ensure that you are within the 
    guidelines?\
    *Hint:* Refer to your plot of the relationship between $p$ and margin of 
    error. Do not use the data set to answer this question.**Since we have no idea about $p$ we would assume 0.5 for $p$.**
    $\text{point estimate} \pm \text{critical value} \times \sqrt{\frac{p(1-p)}{n}}$
    **Substituting in what we know and solve for n.**

    $0.01 = 1.96 \times \sqrt{\frac{0.25}{n}} \rightarrow 0.01^2 = 1.96^2 \times \frac{0.25}{n} \rightarrow 0.0001 = \frac{3.8416 \times 0.25}{n} \rightarrow n = \frac{0.9604}{.0001} \rightarrow n = 9604$

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>

### 2018-11-01 DATA607/Project_4.Rmd
Modified Project_4.Rmd

@@
 
+```{r init, echo=FALSE, message=FALSE}
+library(dplyr)
+library(rmdformats)
+library(kableExtra)
+
+# A Prefix nulling hook.
+# Taken from https://stackoverflow.com/questions/22524822/how-can-i-remove-the-prefix-index-indicator-1-in-knitr-output
+
+# Make sure to keep the default for normal processing.
+default_output_hook <- knitr::knit_hooks$get("output")
+
+# Output hooks handle normal R console output.
+knitr::knit_hooks$set(output = function(x, options) {
+
+  comment <- knitr::opts_current$get("comment")
+  if (is.na(comment)) comment <- ""
+  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
+                     x, perl = TRUE)
+  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
+  if (can_null && do_null) {
+    # By default R print output aligns at the right brace.
+    align_index <- regexpr( "\\]", x )[1] - 1
+    # Two cases: start or newline
+    re <- paste0( "^.{", align_index, "}\\]")
+    rep <- comment
+    x <- gsub( re, rep,  x )
+    re <- paste0( "\\\n.{", align_index, "}\\]")
+    rep <- paste0( "\n", comment )
+    x <- gsub( re, rep,  x )
+    # I added this line
+    x <- gsub( "##", "", x)
+  }
+
+  default_output_hook( x, options )
+})
+
+knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
+```
+
 ## Description
@@
 
-### Data Aquisition
+### Data Acquisition
 
@@
 
-Now that the files have been aquired, they need to be decompressed.
+Now that the files have been acquired, they need to be decompressed.
 
@@
 In order to illustrate the process I will take the first message in the spam training set as an example.
+Here's what the file looks like:
 
-```{r}
+##### Table 2. Preview of Spam Email Message
+```{r spam_preview, null_prefix=TRUE, echo=FALSE}
 example_file <- spam$train[1]
+full_path <- paste0("data/spam_2/", example_file)
+con <- file(full_path)
+print(readLines(con), quote = FALSE)
+```
 
-get_email_contents <- function(file_name, path){
- full_path = paste0(path, file_name)
- con <- file(full_path)
- readLines(con)
+The message is split into two parts: the header and the body of the message.
+Since we want to analyze the body of the email one observes it is seperated from the header by a blank line.
+This will be how we will determine where the body of the message begins.
+
+### Approach
+
+In order to process a message body I will remove all punctuation
+
+### Remove Punctuation
+
+```{r get_message, message=FALSE}
+library(dplyr)
+library(tidytext)
+
+stem_hunspell <- function(term) {
+  # Taken from https://cran.r-project.org/web/packages/corpus/vignettes/stemmer.html
+  # look up the term in the dictionary
+  stems <- hunspell::hunspell_stem(term)[[1]]
+  if (length(stems) == 0) { # if there are no stems, use the original term
+    stem <- term
+  } else { # if there are multiple stems, use the last one
+    stem <- stems[[length(stems)]]
+  }
+  stem
 }
 
+get_message <- function(full_path){
+  message <- ''
+  header_finished <- FALSE
+  # Open the file
+  con <- file(full_path, blocking = FALSE)
+  # Go line by line
+  for(line in readLines(con)){
+    # Order matters.  I don't want to lead with the blank line between the header
+    if(header_finished){
+      message <- paste(message, line)
+    }
+    # Look for the end of the header
+    if(line == ""){
+      header_finished <- TRUE
+    }
+  }
+  # Close the connection to the file
+  close(con)
+
+  # Change text to lower case 
+  message <- tolower(message)
+  # Remove all punctuation
+  message <- gsub("[[:punct:][:blank:]]+", " ", message)
+  # Tokenize message
+  words <- scan_tokenizer(message)
+  # Remove stop words
+  stop_words <- tm::stopwords()
+  words <- words[!words %in% stop_words]
+  # Stem words using hunspell
+  words <- unlist(lapply(words, stem_hunspell))
+
+  return(words)
+}
+
+full_path <- paste0("data/spam_2/", example_file)
+
+bag_of_words <- get_message(full_path)
+df <- data.frame(word = bag_of_words, count = c(1)) %>%
+  group_by(word) %>%
+  summarise(count = sum(count)) %>%
+  ungroup() %>%
+  mutate(file = example_file, type = "spam")
 ```
 
-### Split up Sentance
-
-### Remove Punctuation and Stop Words
-
-### Stemming
+## TF-IDF Normalization

### 2018-11-02 DATA606/Inference for Categorical Data.Rmd
More work on Inference for Categorical Data

@@
 
-(a) We are 95% confident that between 43% and 49% of Americans in this sample support the decision of the U.S. Supreme Court on the 2010 healthcare law.
+(a) We are 95% confident that between 43% and 49% of Americans in this sample support the decision of the U.S.
+Supreme Court on the 2010 healthcare law.**False.
+The CI applies to the population not the sample.**
 
-(b) We are 95% confident that between 43% and 49% of Americans support the decision of the U.S. Supreme Court on the 2010 healthcare law.
+(b) We are 95% confident that between 43% and 49% of Americans support the decision of the U.S.
+Supreme Court on the 2010 healthcare law.**True.
+We are confident the population parameter falls within the range.**
 
 (c) If we considered many random samples of 1,012 Americans, and we calculated the sample proportions of those who support the decision of the U.S.
-Supreme Court, 95% of those sample proportions will be between 43% and 49%.
+Supreme Court, 95% of those sample proportions will be between 43% and 49%.**True.
+95% of the area under the curve would be between 43% and 49%.**
 
-(d) The margin of error at a 90% confidence level would be higher than 3%.
+(d) The margin of error at a 90% confidence level would be higher than 3%.**False.  A higher confidence level would have a larger margin of error.**
 
@@
 
-(a) Is 48% a sample statistic or a population parameter? Explain.
+(a) Is 48% a sample statistic or a population parameter? Explain. **Sample statistics because it is the share of the respondents.**
 
 (b) Construct a 95% confidence interval for the proportion of US residents who think marijuana should be made legal, and interpret it in the context of the data.
+```{r}
+n <- 1259
+p <- 0.48
+z <- 1.96
+se <- sqrt((p*(1-p))/n)
+me <- z * se
+```
+**The 95% confidence interval is (`r p - me`, `r p + me`).  That means that the between 45% and 51% of US residents thing marijuana should be made legal.**
 
@@
 Is this true for these data?
-Explain.
+Explain.**There are `r round(p * 1259, 0)` people who agree and `r round((1-p) * 1259, 0)` people who disagree.
+This is larger than the 10 success/failure rule of thumb.
+The observations are probably independent and the sample size is about right to represent the United States.
+The critics points are not valid.**
 
 (d) A news piece on this survey’s findings states, “Majority of Americans think marijuana should be legalized.” Based on your confidence interval, is this news piece’s statement justified?
+**No.
+This is an exageration.**
 
@@
 If we wanted to limit the margin of error of a 95% confidence interval to 2%, about how many Americans would we need to survey ?
+$0.02 = 1.96 \times \sqrt{\frac{0.5 \times 0.5}{n}} \rightarrow 0.02^2 = 1.96^2 \times \frac{0.25}{n} \rightarrow 0.0004 = \frac{3.8416 \times 0.25}{n} \rightarrow n = \frac{0.9604}{0.0004} \rightarrow n = 2401$ **You would need to survey 2,401 people.**
 
@@
 Calculate a 95% confidence interval for the difference between the proportions of Californians and Oregonians who are sleep deprived and interpret it in context of the data.
+```{r}
+p1 <- 0.08
+p2 <- 0.088
+n1 <- 11545
+n2 <- 4691
+z <- 1.96
+se <- sqrt(((p1*(1-p1))/n1)+((p2*(1-p2))/n2))
+me <- z * se
+```
+**The 95% confidence interval is (`r (p1-p2)-me`, `r (p1-p2)+me`).  This means the difference between CA and OR is between -2% and 0%.**
 
@@
 
-(b) What type of test can we use to answer this research question?
+**H$_0$: The proportion between the groups are the same: $p_1$ = $p_2$ = $p_3$ = $p_4$.**
+
+**H$_A$: At least one of the proportions are different from the others.**
+
+(b) What type of test can we use to answer this research question? **Chi-squared Test or goodness of fit test.**
 
 (c) Check if the assumptions and conditions required for this test are satisfied.
+**Independence: each case only contributes to one cell of the table and are a independent.
+Sample size: Each cell of expected cases should have at least 5 cases.
+We'll have to see if this holds.**
 
@@
 
-## 6.48 Coffee and Depression.
+|          | Woods | Cultivated grassplot | Deciduous forests | Other | Total |
+|----------|-------|----------------------|-------------------|-------|-------|
+| Observed | 4     | 16                   | 61                | 345   | 426   |
+| Expected | 20    | 63                   | 169               | 174   | 426   |
+
+```{r}
+woods <- (4 - 20)^2 / 20
+cult <- (16 - 63)^2 / 63
+deci <- (61 - 169)^2 / 169
+other <- (345 - 174)^2 / 174
+k <- 4
+chi_sq <- woods + cult + deci + other
+df <- k - 1
+pchisq(chi_sq, df, lower.tail = FALSE)
+```
+**The P value is smaller than 0.05 so the evidence suggest we reject the null hypothesis, meaning that the deer do not prefer to forage in all types of terain equally.**
+
+## 6.48 Coffee and Depression
 
@@
 
-(a) What type of test is appropriate for evaluating if there is an association between coffee intake and depression?
+(a) What type of test is appropriate for evaluating if there is an association between coffee intake and depression? **A Chi-squared or goodness of fit test.**
 
 (b) Write the hypotheses for the test you identified in part (a).
+**H$_0$: The proportion between the groups are the same: $p_1$ = $p_2$ = $p_3$ = $p_4$ = $p_5.**
 
-(c) Calculate the overall proportion of women who do and do not suffer from depression.
+**H$_A$: At least one of the proportions are different from the others.**
 
-(d) Identify the expected count for the highlighted cell, and calculate the contribution of this cell to the test statistic, i.e. $\frac{(\text{Observed} − \text{Expected})^2}{\text{Expected}}$.
+(c) Calculate the overall proportion of women who do and do not suffer from depression. **`r 48132/50739`**
+
+(d) Identify the expected count for the highlighted cell, and calculate the contribution of this cell to the test statistic, i.e. $\frac{(\text{Observed} − \text{Expected})^2}{\text{Expected}}$
+
+**Expected is defined as:** $2607 \times
+\frac{6617}{50739} = 2607 \times 0.1304125 = 340$
+
+**so $\frac{(373 − 340)^2}{340} = \frac{1089}{340} = 3.203$.**
 
@@
 
-(f) What is the conclusion of the hypothesis test?
+```{r}
+k <- 5
+chi_sq <- 20.93
+df <- k - 1
+pchisq(chi_sq, df, lower.tail = FALSE)
+```
+
+(f) What is the conclusion of the hypothesis test? **The p-value is lower than 0.05 therefore you would reject the null hypothesis.**
 
 (g) One of the authors of this study was quoted on the NYTimes as saying it was “too early to recommend that women load up on extra coffee” based on just this study.64 Do you agree with this statement?
-Explain your reasoning.
+Explain your reasoning.**Yes.
+It's just one study.
+There is a chance that it is a fluke and should be confirmed with more research.**

### 2018-11-02 DATA607/Project_4.Rmd
Additional work on Project 4

@@
 
-get_message <- function(full_path){
+process_message <- function(full_path){
   message <- ''
+  n_lines <- 0
   header_finished <- FALSE
@@
     if(header_finished){
+      n_lines <- n_lines + 1
       message <- paste(message, line)
@@
   # Change text to lower case 
-  message <- tolower(message)
+  n_characters <- nchar(message)
+  lower_message <- tolower(message)
+  n_capital_letters <- adist(message, lower_message)[[1]]
   # Remove all punctuation
-  message <- gsub("[[:punct:][:blank:]]+", " ", message)
+  clean_message <- gsub("[[:punct:][:blank:]]+", " ", lower_message)
+  n_punctuation_marks <- adist(lower_message, clean_message)[[1]]
   # Tokenize message
-  words <- scan_tokenizer(message)
+  words <- tm::scan_tokenizer(clean_message)
   # Remove stop words
-  stop_words <- tm::stopwords()
-  words <- words[!words %in% stop_words]
+  #stop_words <- tm::stopwords()
+  #words <- words[!words %in% stop_words]
   # Stem words using hunspell
-  words <- unlist(lapply(words, stem_hunspell))
+  #words <- unlist(lapply(words, stem_hunspell))
 
-  return(words)
+  return(list(words = words, characters = n_characters, capital_letters = n_capital_letters, punctuation_marks = n_punctuation_marks, lines = n_lines))
 }
@@
 
-bag_of_words <- get_message(full_path)
-df <- data.frame(word = bag_of_words, count = c(1)) %>%
-  group_by(word) %>%
-  summarise(count = sum(count)) %>%
-  ungroup() %>%
-  mutate(file = example_file, type = "spam")
+results <- process_message(full_path)
+#df <- data.frame(word = bag_of_words, count = c(1)) %>%
+#  group_by(word) %>%
+#  summarise(count = sum(count)) %>%
+#  ungroup() %>%
+#  mutate(file = example_file, type = "spam")
 ```

### 2018-11-03 DATA607/Project_4.Rmd
Finished project 4

--- 2018-10-30 DATA607/Project_4.Rmd
@@
 
-## Description
+## Introduction
 
-This project seeks to build a classifier that can predict if an email is spam or ham.  It will be using the Apache Spam Assassin public corpus as the dataset.
+This project seeks to build a classifier that can predict if an email is *spam* or *ham*.  It will be using the Apache Spam Assassin public corpus as the dataset.
 
-### Data Aquisition
+## Data Acquisition
 
@@
 ```{r file_download}
-if(!file.exists("data/20050311_spam_2.tar.bz2") && !file.exists("data/20050311_spam_2.tar")){
+if (!file.exists("data/20050311_spam_2.tar.bz2") && !file.exists("data/20050311_spam_2.tar")){
   download.file("https://spamassassin.apache.org/old/publiccorpus/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar.bz2")
@@
 
-if(!file.exists("data/20030228_easy_ham_2.tar.bz2") & !file.exists("data/20030228_easy_ham_2.tar")){
+if (!file.exists("data/20030228_easy_ham_2.tar.bz2") & !file.exists("data/20030228_easy_ham_2.tar")){
   download.file("https://spamassassin.apache.org/old/publiccorpus/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar.bz2")
@@
 
-Now that the files have been aquired, they need to be decompressed.
+Now that the files have been acquired, they need to be decompressed.
 
 ```{r decompress_files}
-if(!dir.exists("data/spam_2")){
+if (!dir.exists("data/spam_2")){
   R.utils::bunzip2("data/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar")
@@
   unlink("data/spam_2/cmds")
+  # Delete these messages b/c it breaks the pipeline
+  unlink("data/spam_2/00706.5116018237368c3633823b2d24f8ac86")
+  unlink("data/spam_2/00708.89f1f9108884517148fdbd744e18ec1e")
+  unlink("data/spam_2/00737.af5f503fe444ae773bfeb4652d122349")
+  unlink("data/spam_2/01125.46ca779f86e1dd0a03c3ffc67b57f55e")
+  unlink("data/spam_2/01217.d5a1734ec521c1bd55270eca3ab4acd8")
 }
 
-if(!dir.exists("data/easy_ham_2")){
+if (!dir.exists("data/easy_ham_2")){
   R.utils::bunzip2("data/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar")
@@
 
-## Processing Pipeline
+## Build a Corpus
 
-### Train/Test Split
+Now that we have a dataset we need to construct a corpus.  I will use the *tm* package and *dplyr* to get the job done.
 
-First we are going to split the data into training and testing sets using an 80:20 split.
+```{r build_corpus, message=FALSE}
+library(dplyr)
+library(tm)
 
-```{r}
-set.seed(12345)
-training_percent <- 0.8
-
-train_test_split <- function(data, training_percent){
-  n <- round(length(data)*training_percent, digits = 0)
-  train <- sort(sample(data, n, replace = FALSE))
-  test <- data[!data %in% train]
-  return(list(test=test, train=train))
+get_corpus <- function(dir, label){
+  corpus <- VCorpus(DirSource(dir)) %>%
+    tm_map(PlainTextDocument)  %>% # Create plain text document
+    tm_map(content_transformer(tolower)) %>% # Standardize case
+    tm_map(removeWords, stopwords("SMART")) %>% # Remove stopwords
+    tm_map(removePunctuation) %>% # Remove punctuation marks
+    tm_map(removeNumbers) %>% # Remove numbers
+    tm_map(stripWhitespace) %>% # Remove extra whitespace
+    tm_map(stemDocument) # Stem the documents
+  meta(corpus, "LABEL") <- label
+  return(corpus)
 }
 
-spam <- train_test_split(list.files("data/spam_2"), training_percent)
-ham <- train_test_split(list.files("data/easy_ham_2"), training_percent)
+corpus<- c(get_corpus("data/spam_2", "Spam"), get_corpus("data/easy_ham_2", "Ham"))
 ```
 
-The following table summarizes the number of messages included in each set:
+## Build a Document-Term Matrix
 
-##### Table 1. Spam and Ham Message Counts by Data Set
-|      |       Training       |        Testing      |
-|------|----------------------|---------------------|
-| Spam |`r length(spam$train)`|`r length(spam$test)`|
-| Ham  |`r length(ham$train)` |`r length(ham$test)` |
+The next step is to construct a document term matrix.
 
-### Pull out Email Contents
-
-This model will classify only on the basis of the content.
-The header information will not be taken into consideration.
-In order to illustrate the process I will take the first message in the spam training set as an example.
-
-```{r}
-example_file <- spam$train[1]
-
-get_email_contents <- function(file_name, path){
- full_path = paste0(path, file_name)
- con <- file(full_path)
- readLines(con)
-}
-
+```{r dtm}
+dtm <- DocumentTermMatrix(corpus)
 ```
 
-### Split up Sentance
+There are `r dtm$nrow` documents and `r dtm$ncol` terms in the document-term matrix.
 
-### Remove Punctuation and Stop Words
+### Remove Infrequent Terms
 
-### Stemming
+There are a lot of terms resulting in a very sparce matrix.  I am going to prune out the less frequent terms from the matrix.  The term must be found in at least 10 documents.
+
+```{r pruning}
+min_docs <- 10
+dtm <- removeSparseTerms(dtm, 1 - (min_docs / length(corpus)))
+```
+
+There are still `r dtm$nrow` documents, but there are `r dtm$ncol` terms in the pruned document-term matrix.
+
+## Normalizing the Data
+
+Some messages have more words than others.  We are going to normalize the data and pull all the data together into a model dataset.
+
+```{r normalize_the_data}
+model_data <- as.matrix(dtm)
+words <- rowSums(model_data)
+model_data <- model_data / words
+model_data <- data.frame(model_data)
+model_data <- cbind(meta(corpus), model_data) %>%
+  mutate(LABEL = as.factor(LABEL))
+```
+
+## Split Data into Training/Test Data Sets
+
+We are now ready to break up the data into training and test sets.  We are going to use a 75%/25% split.  This decision was arbitrary.
+
+```{r train_test_split, message=FALSE}
+library(caret)
+set.seed(12345)
+in_training_set <- createDataPartition(model_data$LABEL, p = 0.75,  list = FALSE)
+training_data <- model_data[in_training_set, ]
+testing_data <- model_data[-in_training_set, ]
+```
+
+## Build a SVM Model
+
+We will take out training data and produce a SVM model that predicts if a message is spam or ham.
+
+```{r build_svm}
+library(e1071)
+model <- svm(LABEL ~ ., data = training_data)
+```
+
+## Test the Model
+
+Now that we have trained a model we can test it out and see how it preforms.
+
+```{r test_model}
+predictions <- testing_data %>%
+  select(-LABEL) %>%
+  predict(model, .)
+```
+
+#### Table 1. Confusion Matrix
+```{r confusion_matrix, echo=FALSE}
+library(kableExtra)
+table(predictions, testing_data$LABEL) %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
+```
+
+```{r errors, echo=FALSE}
+errors <- table(predictions, testing_data$LABEL) %>%
+  data.frame(.) %>%
+  mutate(keep = ifelse(predictions == Var2, 0, 1)) %>%
+  filter(keep == 1) %>%
+  summarise(total = sum(Freq))
+```
+
+There were `r errors$total` messages that were misclassified out of `r length(predictions)` for a `r round((1 - (errors$total / length(predictions))) *100, 0)`% accuracy rate.   Not bad!

### 2018-11-05 DATA606/Intro to Linear Regression.Rmd
Initial commit

---
title: "Intro to Linear Regression"
author: "Mike Silva"
date: "November 5, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
---

## 7.24 Nutrition at Starbucks, Part I

The scatterplot below shows the relationship between the number of calories and amount of carbohydrates (in grams) Starbucks food menu items contain.  Since Starbucks only lists the number of calories on the display items, we are interested in predicting the amount of carbs a menu item has based on its calorie content.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/7_24.png)

(a) Describe the relationship between number of calories and amount of carbohydrates (in grams) that Starbucks food menu items contain.

(b) In this scenario, what are the explanatory and response variables?

(c) Why might we want to fit a regression line to these data?

(d) Do these data meet the conditions required for fitting a least squares line?

## 7.26 Body measurements, Part III

Exercise 7.15 introduces data on shoulder girth and height of a group of individuals. The mean shoulder girth is 107.20 cm with a standard deviation of 10.37 cm. The mean height is 171.14 cm with a standard deviation of 9.41 cm. The correlation between height and shoulder girth is 0.67.

(a) Write the equation of the regression line for predicting height.

(b) Interpret the slope and the intercept in this context.

(c) Calculate $R^2$ of the regression line for predicting height from shoulder girth, and interpret it in the context of the application.

(d) A randomly selected student from your class has a shoulder girth of 100 cm.  Predict the height of this student using the model.

(e) The student from part (d) is 160 cm tall. Calculate the residual, and explain what this residual means.

(f) A one year old has a shoulder girth of 56 cm. Would it be appropriate to use this linear model to predict the height of this child?

## 7.30 Cats, Part I

The following regression output is for predicting the heart weight (in g) of cats from their body weight (in kg). The coefficients are estimated using a dataset of 144 domestic cats.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/7_30.png)

|             | Estimate | Std. Error | t value | Pr(>$abs(t)$) |
|-------------|----------|------------|---------|---------------|
| (Intercept) | -0.357   | 0.692      | -0.515  | 0.607         |
| body wt     | 4.034    | 0.250      | 16.119  | 0.000         |

s = 1.452 $R^2$ = 64.66% $R^2$ adj = 64.41%

(a) Write out the linear model.

(b) Interpret the intercept.

(c) Interpret the slope.

(d) Interpret $R^2$.

(e) Calculate the correlation coefficient.

## 7.40 Rate my professor

Many college courses conclude by giving students the opportunity to evaluate the course and the instructor anonymously. However, the use of these student evaluations as an indicator of course quality and teaching effectiveness is often criticized because these measures may reflect the influence of non-teaching related characteristics, such as the physical appearance of the instructor. Researchers at University of Texas, Austin collected data on teaching evaluation score (higher score means better) and standardized beauty score (a score of 0 means average, negative score means below average, and a positive score means above average) for a sample of 463 professors. The scatterplot below shows the relationship between these variables, and also provided is a regression output for predicting teaching evaluation score from beauty score.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/7_40a.png)

|             | Estimate | Std. Error | t value | Pr(>|t|) |
|-------------|----------|------------|---------|----------|
| (Intercept) | 4.010    | 0.0255     | 157.21  | 0.0000   |
| beauty      |          | 0.0322     | 4.13    | 0.0000   |

(a) Given that the average standardized beauty score is -0.0883 and average teaching evaluation score is 3.9983, calculate the slope. Alternatively, the slope may be computed using just the information provided in the model summary table.

(b) Do these data provide convincing evidence that the slope of the relationship between teaching evaluation and beauty is positive? Explain your reasoning.

(c) List the conditions required for linear regression and check if each one is satisfied for this model based on the following diagnostic plots.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/7_40b.png)

### 2018-11-07 DATA606/Lab7/michael-simple_regression.Rmd
Additional work on Lab7

---
title: "Introduction to linear regression"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## Batter up

The movie [Moneyball](http://en.wikipedia.org/wiki/Moneyball_(film)) focuses on
the "quest for the secret of success in baseball". It follows a low-budget team, 
the Oakland Athletics, who believed that underused statistics, such as a player's 
ability to get on base, betterpredict the ability to score runs than typical 
statistics like home runs, RBIs (runs batted in), and batting average. Obtaining 
players who excelled in these underused statistics turned out to be much more 
affordable for the team.

In this lab we'll be looking at data from all 30 Major League Baseball teams and
examining the linear relationship between runs scored in a season and a number 
of other player statistics. Our aim will be to summarize these relationships 
both graphically and numerically in order to find which variable, if any, helps 
us best predict a team's runs scored in a season.

## The data

Let's load up the data for the 2011 season.

```{r load-data, eval=TRUE}
load("more/mlb11.RData")
```

In addition to runs scored, there are seven traditionally used variables in the 
data set: at-bats, hits, home runs, batting average, strikeouts, stolen bases, 
and wins. There are also three newer variables: on-base percentage, slugging 
percentage, and on-base plus slugging. For the first portion of the analysis 
we'll consider the seven traditional variables. At the end of the lab, you'll 
work with the newer variables on your own.

1.  What type of plot would you use to display the relationship between `runs` 
    and one of the other numerical variables? **A scatter plot.** Plot this 
    relationship using the variable `at_bats` as the predictor. 
    ```{r}
    plot(mlb11$at_bats, mlb11$runs)
    ```
    Doe the relationship look linear? **Yes it looks somewhat linear.** If you knew a
    team's `at_bats`, would you be comfortable using a linear model to predict the 
    number of runs? **I don't know if I would.**

If the relationship looks linear, we can quantify the strength of the
relationship with the correlation coefficient.

```{r cor, eval=TRUE}
cor(mlb11$runs, mlb11$at_bats)
```

## Sum of squared residuals

Think back to the way that we described the distribution of a single variable. 
Recall that we discussed characteristics such as center, spread, and shape. It's
also useful to be able to describe the relationship of two numerical variables, 
such as `runs` and `at_bats` above.

2.  Looking at your plot from the previous exercise, describe the relationship 
    between these two variables. Make sure to discuss the form, direction, and 
    strength of the relationship as well as any unusual observations. **The relationship is positive (upward sloping) and mostly linear and not very strong.  There are a few of outliers.**

Just as we used the mean and standard deviation to summarize a single variable, 
we can summarize the relationship between these two variables by finding the 
line that best follows their association. Use the following interactive 
function to select the line that you think does the best job of going through 
the cloud of points.

```{r plotss-atbats-runs, eval=TRUE}
plot_ss(x = mlb11$at_bats, y = mlb11$runs)
```

After running this command, you'll be prompted to click two points on the plot 
to define a line. Once you've done that, the line you specified will be shown in
black and the residuals in blue. Note that there are 30 residuals, one for each 
of the 30 observations. Recall that the residuals are the difference between the
observed values and the values predicted by the line:

\[
  e_i = y_i - \hat{y}_i
\]

The most common way to do linear regression is to select the line that minimizes
the sum of squared residuals. To visualize the squared residuals, you can rerun 
the plot command and add the argument `showSquares = TRUE`.

```{r plotss-atbats-runs-squares, eval=TRUE}
plot_ss(x = mlb11$at_bats, y = mlb11$runs, showSquares = TRUE)
```

Note that the output from the `plot_ss` function provides you with the slope and
intercept of your line as well as the sum of squares.

3.  Using `plot_ss`, choose a line that does a good job of minimizing the sum of
    squares. Run the function several times. What was the smallest sum of 
    squares that you got? How does it compare to your neighbors?

## The linear model

It is rather cumbersome to try to get the correct least squares line, i.e. the 
line that minimizes the sum of squared residuals, through trial and error. 
Instead we can use the `lm` function in R to fit the linear model (a.k.a. 
regression line).

```{r m1, eval=TRUE}
m1 <- lm(runs ~ at_bats, data = mlb11)
```

The first argument in the function `lm` is a formula that takes the form 
`y ~ x`. Here it can be read that we want to make a linear model of `runs` as a 
function of `at_bats`. The second argument specifies that R should look in the 
`mlb11` data frame to find the `runs` and `at_bats` variables.

The output of `lm` is an object that contains all of the information we need 
about the linear model that was just fit. We can access this information using 
the summary function.

```{r summary-m1, eval=TRUE}
summary(m1)
```

Let's consider this output piece by piece. First, the formula used to describe 
the model is shown at the top. After the formula you find the five-number 
summary of the residuals. The "Coefficients" table shown next is key; its first 
column displays the linear model's y-intercept and the coefficient of `at_bats`.
With this table, we can write down the least squares regression line for the 
linear model:

\[
  \hat{y} = -2789.2429 + 0.6305 * atbats
\]

One last piece of information we will discuss from the summary output is the 
Multiple R-squared, or more simply, $R^2$. The $R^2$ value represents the 
proportion of variability in the response variable that is explained by the 
explanatory variable. For this model, 37.3% of the variability in runs is 
explained by at-bats.

4.  Fit a new model that uses `homeruns` to predict `runs`. Using the estimates 
    from the R output, write the equation of the regression line. What does the 
    slope tell us in the context of the relationship between success of a team 
    and its home runs? **$\hat{y} = 415.2389 + 1.8345 * homeruns$**
    ```{r}
    m2 <- lm(runs ~ homeruns, data = mlb11)
    summary(m2)
    ```

## Prediction and prediction errors

Let's create a scatterplot with the least squares line laid on top.

```{r reg-with-line, eval=TRUE}
plot(mlb11$runs ~ mlb11$at_bats)
abline(m1)
```

The function `abline` plots a line based on its slope and intercept. Here, we 
used a shortcut by providing the model `m1`, which contains both parameter 
estimates. This line can be used to predict $y$ at any value of $x$. When 
predictions are made for values of $x$ that are beyond the range of the observed
data, it is referred to as *extrapolation* and is not usually recommended. 
However, predictions made within the range of the data are more reliable. 
They're also used to compute the residuals.

5.  If a team manager saw the least squares regression line and not the actual 
    data, how many runs would he or she predict for a team with 5,578 at-bats?
    $\hat{y} = -2789.2429 + 0.6305 * atbats = -2789.2429 + 0.6305 * 5578 = -2789.2429 + 3516.929 = 727.6861$
    Is this an overestimate or an underestimate, and by how much? In other 
    words, what is the residual for this prediction?

## Model diagnostics

To assess whether the linear model is reliable, we need to check for (1) 
linearity, (2) nearly normal residuals, and (3) constant variability.

*Linearity*: You already checked if the relationship between runs and at-bats
is linear using a scatterplot. We should also verify this condition with a plot 
of the residuals vs. at-bats. Recall that any code following a *#* is intended
to be a comment that helps understand the code but is ignored by R.

```{r residuals, eval=TRUE}
plot(m1$residuals ~ mlb11$at_bats)
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
```

6.  Is there any apparent pattern in the residuals plot? What does this indicate
    about the linearity of the relationship between runs and at-bats?

*Nearly normal residuals*: To check this condition, we can look at a histogram

```{r hist-res, eval=TRUE}
hist(m1$residuals)
```

or a normal probability plot of the residuals.

```{r qq-res, eval=TRUE}
qqnorm(m1$residuals)
qqline(m1$residuals)  # adds diagonal line to the normal prob plot
```

7.  Based on the histogram and the normal probability plot, does the nearly 
    normal residuals condition appear to be met?

*Constant variability*:

8.  Based on the plot in (1), does the constant variability condition appear to 
    be met?

* * *

## On Your Own

-   Choose another traditional variable from `mlb11` that you think might be a 
    good predictor of `runs`. Produce a scatterplot of the two variables and fit 
    a linear model. At a glance, does there seem to be a linear relationship?

-   How does this relationship compare to the relationship between `runs` and 
    `at_bats`? Use the R$^2$ values from the two model summaries to compare. 
    Does your variable seem to predict `runs` better than `at_bats`? How can you
    tell?

-   Now that you can summarize the linear relationship between two variables, 
    investigate the relationships between `runs` and each of the other five 
    traditional variables. Which variable best predicts `runs`? Support your 
    conclusion using the graphical and numerical methods we've discussed (for 
    the sake of conciseness, only include output for the best variable, not all 
    five).

-   Now examine the three newer variables. These are the statistics used by the 
    author of *Moneyball* to predict a teams success. In general, are they more 
    or less effective at predicting runs that the old variables? Explain using 
    appropriate graphical and numerical evidence. Of all ten variables we've 
    analyzed, which seems to be the best predictor of `runs`? Using the limited 
    (or not so limited) information you know about these baseball statistics, 
    does your result make sense?

-   Check the model diagnostics for the regression model with the variable you 
    decided was the best predictor for runs.

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel 
from a lab written by the faculty and TAs of UCLA Statistics.
</div>

### 2018-11-07 DATA607/Recommender Systems.Rmd
Initial commit

## Introduction

The following is an analysis of the LinkedIn recommender system. Users can interact with this system once they have set up a profile.  It recommends connections the user should make.  The following is a scenario design analysis from the point of view of the LinkedIn customer.

## Scenario Design Analysis

### Who Are The Target Users?

The target users are individuals who would like to develop a professional network.

### What Are Their Key Goals?

Their goals may vary.  They may have the goal of staying connected with people from school.  They may be networking to find new employment opportunities.  Or they may want to develop business opportunities and think LinkedIn could help.

### How Can LinkedIn Help Them Accomplish Those Goals?

LinkedIn helps by first maintaining a very robust network of individuals.  Being large attracts future users. The recommender system functions as a filter.  It helps the user filter through millions of profiles and find people to connect to.

The recommender system appears to focus in part on demographic characteristics.  It will recommend people located near by.  It recommends people who attended the same schools or who are members of trade groups.  It looks for people who "look" like the user.

It also recommends common peers.  For example, suppose Sam is connected with Tonia and James.  They are both connected with Juan.  Sam would get a recommendation to connect with Juan).  This recommendation is a knowledge-based filter.  Users have made their preferences known through the connections they have made.

### 2018-11-09 DATA607/Final Project/Build Monroe Real Property Data.py
More work on final project

@@ -1,0 +1,126 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Nov  5 13:26:41 2018
+
+@author: Michael Silva
+"""
+
+import urllib
+import zipfile
+import shutil
+from dbfread import DBF
+import sqlite3 as db
+import requests
+import asyncio
+import time
+import pandas as pd
+from bs4 import BeautifulSoup
+import re
+from datetime import datetime
+
+# Source: https://stackoverflow.com/questions/312443/how-do-you-split-a-list-into-evenly-sized-chunks
+def chunks(l, n):
+    """Yield successive n-sized chunks from l."""
+    for i in range(0, len(l), n):
+        yield l[i:i + n]
+
+start_time = time.time()
+
+print('Setting up db')
+db_name = "Monroe Real Property Data.db"
+con = db.connect(db_name)
+con.row_factory = lambda cursor, row: row[0]
+c = con.cursor()
+c.execute('DROP TABLE IF EXISTS `scraped_data`')
+c.execute('CREATE TABLE `scraped_data` (`SWIS`, `SBL`, `url`,`html`)')
+c.execute('DROP TABLE IF EXISTS `sales_data`')
+c.execute('CREATE TABLE `sales_data` (`sale_date` DATE, `price` REAL, `SWIS`,`SBL`)')
+c.execute('CREATE INDEX `sd` ON `sales_data` (`SWIS`, `SBL`);')
+con.commit()
+
+print("--- %s seconds ---" % (time.time() - start_time))
+
+print('Getting the Monroe County Tax Parcels Centroids')
+url = 'http://gis.ny.gov/gisdata/fileserver/?DSID=1300&file=Monroe-Tax-Parcels-Centroid-Points-SHP.zip'
+zip_file_name = url.split('file=')[1]
+dbf_file_name = 'Monroe_2017_Tax_Parcel_Centroid_Points_SHP.dbf'
+
+with urllib.request.urlopen(url) as response, open(zip_file_name, 'wb') as out_file:
+    shutil.copyfileobj(response, out_file)
+with zipfile.ZipFile(zip_file_name) as zf:
+    zf.extractall()
+print("--- %s seconds ---" % (time.time() - start_time))
+
+print('Building list of urls to scrape')
+urls_to_scrape = list()
+prop_info = list()
+
+for record in DBF(dbf_file_name):
+    try:
+        if record['PROP_CLASS'][0] == '2':
[…]

### 2018-11-11 DATA607/Final Project/Build Monroe Real Property Data.py
More work on my final project

@@ -17,38 +17,38 @@ import pandas as pd
 from bs4 import BeautifulSoup
 import re
 from datetime import datetime
+import os.path
 
-# Source: https://stackoverflow.com/questions/312443/how-do-you-split-a-list-into-evenly-sized-chunks
-def chunks(l, n):
-    """Yield successive n-sized chunks from l."""
-    for i in range(0, len(l), n):
-        yield l[i:i + n]
+create_db_from_scratch = False
 
 start_time = time.time()
 
-print('Setting up db')
 db_name = "Monroe Real Property Data.db"
 con = db.connect(db_name)
 con.row_factory = lambda cursor, row: row[0]
 c = con.cursor()
-c.execute('DROP TABLE IF EXISTS `scraped_data`')
-c.execute('CREATE TABLE `scraped_data` (`SWIS`, `SBL`, `url`,`html`)')
-c.execute('DROP TABLE IF EXISTS `sales_data`')
-c.execute('CREATE TABLE `sales_data` (`sale_date` DATE, `price` REAL, `SWIS`,`SBL`)')
-c.execute('CREATE INDEX `sd` ON `sales_data` (`SWIS`, `SBL`);')
-con.commit()
 
-print("--- %s seconds ---" % (time.time() - start_time))
+if create_db_from_scratch:
+    print('Setting up db')
+    c.execute('DROP TABLE IF EXISTS `scraped_data`')
+    c.execute('CREATE TABLE `scraped_data` (`SWIS`, `SBL`, `url`,`html`)')
+    c.execute('CREATE INDEX `scraped` ON `scraped_data` (`SWIS`, `SBL`);')
+    c.execute('DROP TABLE IF EXISTS `sales_data`')
+    c.execute('CREATE TABLE `sales_data` (`sale_date` DATE, `price` REAL, `SWIS`,`SBL`)')
+    c.execute('CREATE INDEX `sales` ON `sales_data` (`SWIS`, `SBL`);')
+    con.commit()
+    print("--- %s seconds ---" % (time.time() - start_time))
 
 print('Getting the Monroe County Tax Parcels Centroids')
-url = 'http://gis.ny.gov/gisdata/fileserver/?DSID=1300&file=Monroe-Tax-Parcels-Centroid-Points-SHP.zip'
-zip_file_name = url.split('file=')[1]
 dbf_file_name = 'Monroe_2017_Tax_Parcel_Centroid_Points_SHP.dbf'
-
[…]

### 2018-11-12 DATA607/Final Project/Build Monroe Real Property Data.py
Cleanup

@@ -1,137 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Nov  5 13:26:41 2018
-
-@author: Michael Silva
-"""
-
-import urllib
-import zipfile
-import shutil
-from dbfread import DBF
-import sqlite3 as db
-import requests
-import asyncio
-import time
-import pandas as pd
-from bs4 import BeautifulSoup
-import re
-from datetime import datetime
-import os.path
-
-create_db_from_scratch = False
-
-start_time = time.time()
-
-db_name = "Monroe Real Property Data.db"
-con = db.connect(db_name)
-con.row_factory = lambda cursor, row: row[0]
-c = con.cursor()
-
-if create_db_from_scratch:
-    print('Setting up db')
-    c.execute('DROP TABLE IF EXISTS `scraped_data`')
-    c.execute('CREATE TABLE `scraped_data` (`SWIS`, `SBL`, `url`,`html`)')
-    c.execute('CREATE INDEX `scraped` ON `scraped_data` (`SWIS`, `SBL`);')
-    c.execute('DROP TABLE IF EXISTS `sales_data`')
-    c.execute('CREATE TABLE `sales_data` (`sale_date` DATE, `price` REAL, `SWIS`,`SBL`)')
-    c.execute('CREATE INDEX `sales` ON `sales_data` (`SWIS`, `SBL`);')
-    con.commit()
-    print("--- %s seconds ---" % (time.time() - start_time))
-
-print('Getting the Monroe County Tax Parcels Centroids')
-dbf_file_name = 'Monroe_2017_Tax_Parcel_Centroid_Points_SHP.dbf'
-if not os.path.isfile(dbf_file_name):
-    url = 'http://gis.ny.gov/gisdata/fileserver/?DSID=1300&file=Monroe-Tax-Parcels-Centroid-Points-SHP.zip'
-    zip_file_name = url.split('file=')[1]
-    if not os.path.isfile(dbf_file_name):
-        with urllib.request.urlopen(url) as response, open(zip_file_name, 'wb') as out_file:
-            shutil.copyfileobj(response, out_file)
-        with zipfile.ZipFile(zip_file_name) as zf:
-            zf.extractall()
-print("--- %s seconds ---" % (time.time() - start_time))
-
-print('Building list of urls to scrape')
-urls_to_scrape = list()
-prop_info = list()
-
-for record in DBF(dbf_file_name):
-    try:
-        if record['PROP_CLASS'][0] == '2':
-            prop_info.append(record)
[…]

### 2018-11-15 DATA606/Silva-Proposal.Rmd
Updated proposal

--- 2018-10-05 DATA606/Silva-Proposal.Rmd
@@
 
-### Data Preparation
+## Research Question
 
-```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
-# load data
+What effect does the age of a home and its square footage have on its sales price in Monroe County, New York?
 
+## Data Collection
+
+The sales price data were collected through governmental reporting.
+The price data is through the NYS Office of Real Property Services and provided to the general public by Monroe County through their real property portal.
+The housing attributes is compiled by the NYS GIS clearinghous from data provided by local assessors.
+
+## Type of Study
+
+This study will be an observational study.
+
+## Dependent Variable
+
+The response variable is the housing price.
+This study will use the latest sales price available.
+This is a quantitative variable.
+Since the housing attributes are as of 2017 we will only include sales that occured in 2016-2018 (assuming the homes haven't changed much since 2017).
+
+## Independent Variable
+
+The two variables will be the square footage of the living space (quantitative) and the era the home was built (qualitative).
+These eras will follow the groups used by the Trulia's Chief Economist in [this article](https://www.forbes.com/sites/trulia/2013/05/02/american-homes-by-decade/#16c559821045).
+
+## Data Source
+
+### Sales Price Data
+
+Monroe County's Real Property Portal.
+(2018).
+Database was synced with NYS RPS on 22 October 2018.
+Scraped from [https://www.monroecounty.gov/etc/rp](https://www.monroecounty.gov/etc/rp) on 10 November 2018.
+
+### Housing Attributes
+
+NYS GIS Clearinghouse.
+(2017).
+Monroe County Tax Parcel Centroids [Data file].
+Retrieved from [http://gis.ny.gov/gisdata/inventories/details.cfm?DSID=1300](http://gis.ny.gov/gisdata/inventories/details.cfm?DSID=1300).
+
+## Data Preparation
+
+The Monroe County Real Property portal has been scraped and the data has been stored in an SQLite database.
+
+```{r get_the_database, message=FALSE}
+library(DBI)
+library(dplyr)
+
+# Download the database
+if(!file.exists("Monroe Real Property Data.db")){
+  if(!file.exists("Monroe Real Property Data.db.tar.gz")){
+    # Download this specific commit from my GitHub repo
+    download.file("https://github.com/mikeasilva/CUNY-SPS/raw/9a2023b57a466550646c9190076a53d51ef09ef2/data/Monroe%20Real%20Property%20Data.db.tar.gz", "Monroe Real Property Data.db.tar.gz")
+  }
+  # Decompress the file
+  untar("Monroe Real Property Data.db.tar.gz")
+}
 ```
 
-### Research question
+Now that the data is downloaded it's time to read it in.
 
-**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**
+```{r read_db}
+# Connect to the database
+con <- dbConnect(RSQLite::SQLite(), "Monroe Real Property Data.db")
 
-### Cases
+# Get the latest unduplicated sales price data
+results <- dbSendQuery(con, "SELECT MAX(sale_date) AS sale_date, SWIS, SBL FROM sales_data GROUP BY SWIS, SBL HAVING sale_date > '2015-12-31'")
+df <- dbFetch(results)
 
-**What are the cases, and how many are there?**
+dbClearResult(results)
+```
 
-### Data collection
+There are `r nrow(df)` sales prices in the data frame.
 
-**Describe the method of data collection.**
+```{r sales_data}
+df <- dbReadTable(con, "sales_data") %>% # Pull all the records
+  distinct() %>% # remove any duplicates
+  merge(df) %>% # Merge with the targeted sales observations
+  rename(Price = price)
+```
 
-### Type of study
+The count of observations is now `r nrow(df)`.  There was some duplication introduced through the merging process.  I will drop all the observations with more than one price.
 
-**What type of study is this (observational/experiment)?**
+```{r deduplication}
+df <- df %>%
+  select(SWIS, SBL) %>% # Select the unique identifiers
+  group_by(SWIS, SBL) %>% 
+  summarise(count = n()) %>% # Get the number of records
+  ungroup() %>%
+  filter(count == 1) %>% # Only keep the observations with one price
+  select(-count) %>% # Drop the count
+  merge(df) # subset the price observations
+```
 
-### Data Source
+There are `r nrow(df)` observations in the data frame.  Now that we have the dependent variable, it is time to add in the housing attributes.
 
-**If you collected the data, state self-collected. If not, provide a citation/link.**
+```{r housing_attributes}
+# Add in the housing attributes
+df <- dbReadTable(con, "property_info") %>% # Pull all the records
+  distinct() %>% # remove any duplicates
+  merge(df) # Merge with the targeted sales observations
 
-### Dependent Variable
+dbDisconnect(con)
+```
 
-**What is the response variable? Is it quantitative or qualitative?**
+Now that the housing attributes are added in I will filter out observations that are odd.  Some variable have 0 when there isn't data.
 
-### Independent Variable
+```{r filter_out_crap}
+df <- df %>% # First filter out some obviously weird data
+  filter(YR_BLT > 0) %>%
+  filter(SQFT_LIV > 0) %>%
+  filter(NBR_KITCHN > 0) %>%
+  filter(NBR_BEDRM > 0)
+```
 
-**You should have two independent variables, one quantitative and one qualitative.**
+The last step is to change the year built quantitative variable into a qualitative variable.
 
-### Relevant summary statistics
+```{r finalize_df}
+get_era <- function(year){
+  if(year < 1940){
+    return("1. Before 1940s")
+  } else if(year < 1980){
+    return("2. Post-War Boom")
+  }else{
+    return("3. Last 30+ Years")
+  }
+}
 
-**Provide summary statistics for each the variables.
-Also include appropriate visualizations related to your research question (e.g.
-scatter plot, boxplots, etc).
-This step requires the use of R, hence a code chunk is provided below.
-Insert more code chunks as needed.**
+df <- df %>% 
+  rowwise() %>%
+  mutate(Era = get_era(YR_BLT)) %>%
+  ungroup() %>%
+  mutate(Era = as.factor(Era)) %>%
+  rename(Size = SQFT_LIV) %>%
+  select(Price, Size, Era)
+```
 
-```{r}
+## Cases
 
+Each case is a residential property in Monroe County, New York.  There are **`r nrow(df)`** cases in the data set.
+
+## Relevant Summary Statistics
+
+Here's the summary of the data
+
+##### Table 1: Summary of Data
+```{r summary, comment=NA, message=FALSE, echo=FALSE}
+library(kableExtra)
+summary(df)
 ```
+
+There are extreme price and size values on both ends.  There will probably need to be some more filtering of the the data.  Let's explore the data with some visualizations.
+
+```{r hist, echo=FALSE}
+library(ggplot2)
+library(ggthemes)
+
+ggplot(df, aes(Price)) + 
+  geom_histogram(bins = 100) + 
+  scale_color_fivethirtyeight() +
+  theme_fivethirtyeight(base_size = 12, base_family = "sans")
+```
+
+##### Figure 1: Price and Size Scatterplot
+```{r echo=FALSE}
+ggplot(df, aes(Price, Size)) + 
+  geom_point() + 
+  scale_color_fivethirtyeight() +
+  theme_fivethirtyeight(base_size = 12, base_family = "sans")
+```
+
+This looks encouraging.  I probably should remove the few high price outliers from the data set.
+
+##### Figure 2: Price and Era Boxplot
+```{r era_by_price_boxplot, echo=FALSE}
+ggplot(df, aes(x=Era, y=Price)) + 
+  geom_boxplot() + 
+  theme_fivethirtyeight(base_size = 12, base_family = "sans")
+```
+
+This visualization is not very helpful because of the outliers.  Let's look at the same data in tabular form.
+
+##### Table 2: Price Summary by Era
+```{r price_by_era_table, echo=FALSE}
+df %>%
+  group_by(Era) %>%
+  summarise(min = min(Price), mean = mean(Price), median = median(Price), max = max(Price)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+Again it looks like more work will need to be done on the price data to clean up the records that are more transactional than capturing the value of a home.  Is a home really worth $2.00?

### 2018-11-15 DATA606/Lab7/michael-simple_regression.Rmd
Work on Lab 7
More work on lab 7

@@
     ```
-    Doe the relationship look linear? **Yes it looks somewhat linear.** If you knew a
-    team's `at_bats`, would you be comfortable using a linear model to predict the 
+
+    Does the relationship look linear? **Yes it looks somewhat linear.** If you knew
+    a team's `at_bats`, would you be comfortable using a linear model to predict the 
     number of runs? **I don't know if I would.**
@@
     $\hat{y} = -2789.2429 + 0.6305 * atbats = -2789.2429 + 0.6305 * 5578 = -2789.2429 + 3516.929 = 727.6861$
-    Is this an overestimate or an underestimate, and by how much? In other 
-    words, what is the residual for this prediction?
+    Is this an overestimate or an underestimate, and by how much? In other words, 
+    what is the residual for this prediction? 
+    ```{r message=FALSE, echo=FALSE}
+    library(dplyr)
+    test <- mlb11 %>%
+      mutate(diff = abs(at_bats - 5578)) %>%
+      mutate(closest = min(diff)) %>%
+      filter(closest == diff) %>%
+      select(at_bats, runs) 
+    ```
+    **The closest data we have is `r test$at_bats`.
+For this case the model predicts `r round(predict(m1, test), 0)` runs and the actual was `r test$runs`.
+The residual is `r round(test$runs - predict(m1, test), 0)` runs which indicates the model was an overestimate.
+It would be similar to 5,578 at bats.**
 

### 2018-11-17 DATA607/Week_12_Assignment.Rmd
Week 12 work

---
title: "CUNY SPS DATA-607 Week 12 Assignment"
author: "Mike Silva"
date: "12 November 2018"
output:
  rmdformats::readthedown:
    highlight: kate
---

## Description

The assignment is to migrate data from a MySQL database to a NoSQL database.  I will be using the flights database used in the Hands on Lab 2-1.

## Hypothetical Scenario

For this assignment I will pretend I am a data scientist at a ThereOnTime.  ThereOnTime is  start-up that evaluates airline performance.  Users enter in their flight route.  ThereOnTime recommends which airline will get the user there on time.  ThereOnTime is growing in popularity and needs to scale up their system.  As their data scientist, I recommend they migrate from MySQL to MongoDB.  For this scenario I will pretend all flights are in the datasets, not only the NYC ones.

### Hypothetical Scenario Design

#### For the End User

ThereOnTime users want a recommendation on which airline that is going to get them from point A to point B on time.  They will have varying exclusion criteria (i.e. they refuse to take on Airline X, they won't fly on puddle jumpers).  The way to help them is through the best system design at ThereOnTime.

#### For the ThereOnTime System

The goal is to provide  quick and relevant recommendations.  They fit the customer's unique preferences and criteria.  ThereOnTime meets these requirement by allowing filtering of evaluations.  The ThereOnTime system looks at weather as a factor to inform their evaluation.  They also have a recommender system for alternative routes.  The airlines are the principle object of evaluation.  The airports, and planes are filtering criteria.  This should inform the NoSQL schema design.

## Extraction from MySQL Database

I will start by connecting to the MySQL database.  I will make use of a MySQL Settings file which hold authentication credentials.  It is not included in the [GitHub repo](https://github.com/mikeasilva/CUNY-SPS/tree/master/DATA607) for obvious reasons.

```{r connect, message=FALSE}
library(DBI)
source('hidden/MySQL Settings.R')
conn <- dbConnect(RMySQL::MySQL(), user=user, password=password, dbname='flights', host=host)
```

I will extract all the tables into their own data frames.

```{r extract, results="hide"}
airlines <- dbReadTable(conn, "airlines")
airports <- dbReadTable(conn, "airports")
flights <- dbReadTable(conn, "flights")
planes <- dbReadTable(conn, "planes")

dbDisconnect(conn)
```

## Data Transformation

I will create the flights data.  Since I don't want to loose any records I will use dplyr's left_join and right_join in the data pipeline.

```{r transform, message=FALSE}
library(dplyr)

# Merge together the data frames
data <- flights %>%
  mutate(on_time = ifelse(arr_delay > 0, 0 ,1)) %>%
  merge(airlines) %>%
  rename(carrier_name = name) %>%
  rowwise() %>%
  mutate(carrier_name = gsub("[\r\n]", "", carrier_name)) %>%
  ungroup() %>%
  left_join(planes)

# Hat tip to https://stackoverflow.com/questions/33741357/how-do-i-add-a-prefix-to-several-variable-names-using-dplyr for rename all columns

# Add in the origin airport data
data <- airports %>%
  rename_(.dots = setNames(names(.), paste0('origin_', names(.)))) %>%
  rename(origin = origin_faa) %>%
  right_join(data)

# Add in the destination airport data
data <- airports %>%
  rename_(.dots = setNames(names(.), paste0('dest_', names(.)))) %>%
  rename(dest = dest_faa) %>%
  right_join(data)
```

## Load Data into MongoDB

Now to load the data into a MongoDB.  First we need a connection:

```{r mongodb_connect}
library(mongolite)
mdb = mongo(collection = "data", db = "flights")
```

Now we can load the data into the MongoDB.

```{r load}
mdb$insert(data)
```

The migration from MySQL to MongoDB is complete.

## Database Test Run

To test out the new database I will run a test.  Let's say the user wants to fly to Rochester NY from the NYC area.  Let's suppose they don't care about which airport they leave from.  Which airline and airport would be most likely to get them there on time?

```{r nyc_to_roc}
results <- mdb$find('{"dest": "ROC"}') %>%
  group_by(carrier_name, origin_name) %>%
  summarise(on_time_flights = sum(on_time, na.rm=TRUE), total_flights = n()) %>%
  mutate(share_on_time = on_time_flights / total_flights) %>%
  arrange(-share_on_time)
```

##### Table 1. Flights from NYC Area to Rochester NY
```{r table, echo=FALSE}
library(kableExtra)
results%>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

It looks like Endeavor Air Inc out of LaGuardia airport is the best bet.

## NoSQL Advantages and Disadvantages

One major difference between a NoSQL and relational database is the schemaless architecture.  This is an advantage because it gives a lot of flexibility.  It is nice to be able to add or drop data elements as needs change.

It is also a disadvantage because it is very foreign way of thinking about data bases.  Determining if you should embed or reference a document is something I don't have a handle on yet.

### 2018-11-17 DATA606/Lab7/michael-simple_regression.Rmd
Lab 7 Complete

@@
 6.  Is there any apparent pattern in the residuals plot? What does this indicate
-    about the linearity of the relationship between runs and at-bats?
+    about the linearity of the relationship between runs and at-bats? **No there is no apparent pattern in the residuals.**
 
@@
 7.  Based on the histogram and the normal probability plot, does the nearly 
-    normal residuals condition appear to be met?
+    normal residuals condition appear to be met?  **Yes it appears to be normal**
 
@@
 8.  Based on the plot in (1), does the constant variability condition appear to 
-    be met?
+    be met? **Yes it does.**
 
@@
 
+    **I think homeruns would be a good predictor because it gets to the skill of the player.  new_obs would probably be better but it is not a traditional variable. **
+
+    ```{r}
+    plot(mlb11$homeruns, mlb11$runs)
+    ```
+
+    **Yes this appears to be a positive linear relationship.**
+
 -   How does this relationship compare to the relationship between `runs` and 
@@
     tell?
+
+    ```{r}
+    m3 <- lm(runs ~ homeruns, data = mlb11)
+    summary(m3)
+    ```
+    **The $R^2$ for this model is `r summary(m3)$r.squared` which is higher than `r summary(m1)$r.squared`.
+It does a better job of predicting.
+You can tell because the $R^2$ is higher and the residual standard error is lower.**
 
@@
     five).
+
+    ```{r}
+    highest_r_squared <- 0
+    # Loop through all the traditional variables
+    for(v in c('at_bats', 'hits', 'homeruns', 'bat_avg', 'strikeouts', 'stolen_bases', 'wins')){
+      temp <- mlb11[,c('runs', v)]
+      model <- lm(runs ~ ., data = temp)
+      r_squared <- summary(model)$r.squared
+      if(r_squared > highest_r_squared){
+        best_var <- v
+        highest_r_squared <- r_squared
+        best <- temp
+        best_model <- model
+      }
+    }
+
+    plot(best)
+    summary(best_model)
+    ```
 
@@
 
+    ```{r}
+    # Loop through all the Moneyball variables
+    best_r_squared <- summary(best_model)$r.squared
+    for(v in c('new_onbase', 'new_slug', 'new_obs')){
+      temp <- mlb11[,c('runs', v)]
+      model <- lm(runs ~ ., data = temp)
+      r_squared <- summary(model)$r.squared
+      if(r_squared > best_r_squared){
+        message(paste(v, "outpreforms", best_var, "(", round(r_squared, 2), " > ", round(best_r_squared, 2), ")"))
+        best_model <- model
+        best_data <- temp
+      }
+    }
+    ```
+
 -   Check the model diagnostics for the regression model with the variable you 
     decided was the best predictor for runs.
+
+    **Linearity : No pattern to the residuals.**
+
+    **Constant variability : Yes Sir!**
+
+    ```{r}
+    plot(best_model$residuals ~ mlb11$new_obs)
+    abline(h = 0, lty = 3)
+    ```
+
+    **Nearly normal residuals : Check**
+    ```{r}
+    qqnorm(best_model$residuals)
+    qqline(best_model$residuals)
+    ```
 

### 2018-11-18 DATA606/Intro to Linear Regression.Rmd
Work on Chapter 7 Assignment

@@
 
-(a) Describe the relationship between number of calories and amount of carbohydrates (in grams) that Starbucks food menu items contain.
+(a) Describe the relationship between number of calories and amount of carbohydrates (in grams) that Starbucks food menu items contain. **Positive and fairly linear.**
 
-(b) In this scenario, what are the explanatory and response variables?
+(b) In this scenario, what are the explanatory and response variables? **The response variable is calories.  The explanatory variable is the grams of carbohydrates.**
 
-(c) Why might we want to fit a regression line to these data?
+(c) Why might we want to fit a regression line to these data? **The data seem like they could be summarized well with a regression line.**
 
 (d) Do these data meet the conditions required for fitting a least squares line?
+**Yes.
+The residuals are normally distributed.
+There is no pattern aparent in the residuals plot.
+And the relationship looks linear as previously stated.**
 
@@
 
+$\widehat{\text{height}}=\beta_0 + \beta_1\times\text{girth}$
+
+$\beta_1 = \frac{s_\text{height}}{s_\text{girth}}R = \frac{9.41}{10.37}0.67 = 0.6079749$
+
+$y - y_0 = \beta_1(x-x_0) \rightarrow y - 171.14 = 0.6079749 (x-107.20)$
+
+$\beta_0 = 171.14 - (-107.2 \times 0.6079749) \rightarrow \beta_0 = 171.14 - 65.17491 \beta_0 = 105.9651$
+
+$\widehat{\text{height}}=\beta_0 + \beta_1\times\text{girth} = 105.9651 + 0.6079749 \times \text{girth}$
+
 (b) Interpret the slope and the intercept in this context.
+**The intercept (average height) is roughly 106 centimeters.
+For every centimeter in sholder girth there is about 0.61 centimeters of height.**
 
-(c) Calculate $R^2$ of the regression line for predicting height from shoulder girth, and interpret it in the context of the application.
+(c) Calculate $R^2$ of the regression line for predicting height from shoulder girth, and interpret it in the context of the application.  $R^2 = 0.67 \times0.67 = 0.4489$
+**This means that about 45% of the variability in height is explained by the sholder girth.**
 
-(d) A randomly selected student from your class has a shoulder girth of 100 cm.  Predict the height of this student using the model.
+(d) A randomly selected student from your class has a shoulder girth of 100 cm.
+Predict the height of this student using the model.
+$\widehat{\text{height}}= 105.9651 + 0.6079749 \times 100 = 105.9651 + 60.79749 = 166.7626$
 
-(e) The student from part (d) is 160 cm tall. Calculate the residual, and explain what this residual means.
+(e) The student from part (d) is 160 cm tall. Calculate the residual, and explain what this residual means. **The residual is -6.7626 cm which is the amount overestimated by the model.**
 
-(f) A one year old has a shoulder girth of 56 cm. Would it be appropriate to use this linear model to predict the height of this child?
+$\text{Residual} = \text{Observed} - \text{Expected} = 160 - 166.7626 = 6.7626$
+
+(f) A one year old has a shoulder girth of 56 cm.
+Would it be appropriate to use this linear model to predict the height of this child?
+**I don't think so.
+The model worked well for a student in part d that is 160 cm tall which is over 5 feet.
+The model works for adults.
+The relationship between these two variables is not likely to hold at such a young age.**
 
@@
 
-(a) Write out the linear model.
+(a) Write out the linear model. $\widehat{\text{heart}} = -0.357 + 4.034 \times \text{body}$
 
-(b) Interpret the intercept.
+(b) Interpret the intercept. **The weight of the cat's heart if it weighed nothing.**
 
-(c) Interpret the slope.
+(c) Interpret the slope. **For every one kg of body weight there will be 4.034 grams of heart weight.**
 
-(d) Interpret $R^2$.
+(d) Interpret $R^2$. **Roughly 64% of the variation is explained by the cat's body weight.**
 
-(e) Calculate the correlation coefficient.
+(e) Calculate the correlation coefficient. $R = \sqrt{R^2} = \sqrt{0.6466} = 0.8041144$
 

### 2018-11-20 DATA606/Intro to Linear Regression.Rmd
More work on Intro to Linear Regression

@@
 Alternatively, the slope may be computed using just the information provided in the model summary table.
+**0.1325028**
 
-(b) Do these data provide convincing evidence that the slope of the relationship between teaching evaluation and beauty is positive? Explain your reasoning.
+$y - y_0 = \text{slope} \times (x - x_0)$
+
+$4.010 - 3.9983 = \text{slope} \times (0 - -0.0883)$
+
+$0.0117 =0.0883 \times \text{slope}$
+
+$\text{slope}=\frac{0.0117}{0.0883} = 0.1325028$
+
+(b) Do these data provide convincing evidence that the slope of the relationship between teaching evaluation and beauty is positive?
+Explain your reasoning.
+**There is a positive slope however it is not very strong.**
 
 (c) List the conditions required for linear regression and check if each one is satisfied for this model based on the following diagnostic plots.
+**Linearity: Scatterplot indicates a positive relationship but it is weak.
+The slope is consistant with this.
+Nomally Distributed Residuals: The Q-Q Plot indicates it.
+Constant variability: The residuals scatterplot indicates this condition is met.
+Independent Observations: The evaluations is independant of the other.**
 

### 2018-11-26 DATA606/Multiple and Logistic Regression.Rmd
Added Chapter 8 Files

---
title: "Multiple and Logistic Regression"
author: "Mike Silva"
date: "November 25, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
---

## 8.2 Baby weights, Part II

Exercise 8.1 introduces a data set on birth weight of babies. Another variable we consider is parity, which is 0 if the child is the first born, and 1 otherwise.
The summary table below shows the results of a linear regression model for predicting the average birth weight of babies, measured in ounces, from parity.

|             | Estimate | Std. Error | t value | Pr(>$abs(t)$) |
|-------------|----------|------------|---------|---------------|
| (Intercept) | 120.07   | 0.60       | 199.94  | 0.000         |
| parity      | -1.93    | 1.19       | -1.62   | 0.1052        |

(a) Write the equation of the regression line.

(b) Interpret the slope in this context, and calculate the predicted birth weight of first borns and others.

(c) Is there a statistically significant relationship between the average birth weight and parity?

## 8.4 Absenteeism, Part I.

Researchers interested in the relationship between absenteeism from school and certain demographic characteristics of children collected data from 146 randomly sampled students in rural New South Wales, Australia, in a particular school year. Below are three observations from this data set.

|     | eth | sex | lrn | day |
|-----|-----|-----|-----|-----|
| 1   |  0  |  1  |  1  |  2  |
| 2   |  0  |  1  |  1  |  11 |
| .   |  .  |  .  |  .  |  .  |
| .   |  .  |  .  |  .  |  .  |
| .   |  .  |  .  |  .  |  .  |
| 146 |  1  |  0  |  0  |  37 |

The summary table below shows the results of a linear regression model for predicting the average number of days absent based on ethnic background (eth: 0 - aboriginal, 1 - not aboriginal), sex (sex: 0 - female, 1 - male), and learner status (lrn: 0 - average learner, 1 - slow learner).1

|             | Estimate | Std. Error | t value | Pr(>$abs(t)$) |
|-------------|----------|------------|---------|---------------|
| (Intercept) | 18.93    | 2.57       | 7.37    | 0.0000        |
| eth         | -9.11    | 2.60       | -3.51   | 0.0000        |
| sex         | 3.10     | 2.64       | 1.18    | 0.2411        |
| lrn         | 2.15     | 2.65       | 0.81    | 0.4177        |

(a) Write the equation of the regression line.

(b) Interpret each one of the slopes in this context.

(c) Calculate the residual for the first observation in the data set: a student who is aboriginal, male, a slow learner, and missed 2 days of school.

(d) The variance of the residuals is 240.57, and the variance of the number of absent days for all students in the data set is 264.17. Calculate the $R^2$ and the adjusted $R^2$. Note that there are 146 observations in the data set.

## 8.8 Absenteeism, Part II.

Exercise 8.4 considers a model that predicts the number of days absent using three predictors: ethnic background (eth), gender (sex), and learner status (lrn).  The table below shows the adjusted R-squared for the model as well as adjusted R-squared values for all models we evaluate in the first step of the backwards elimination process.

|   | Model             | Adjusted $R^2$ |
|---|-------------------|----------------|
| 1 | Full model        | 0.0701         |
| 2 | No ethnicity      | -0.0033        |
| 3 | No sex            | 0.0676         |
| 4 | No learner status | 0.0723         |

Which, if any, variable should be removed from the model first?

## 8.16 Challenger disaster, Part I.

On January 28, 1986, a routine launch was anticipated for the Challenger space shuttle. Seventy-three seconds into the flight, disaster happened: the shuttle broke apart, killing all seven crew members on board. An investigation into the cause of the
disaster focused on a critical seal called an O-ring, and it is believed that damage to these O-rings during a shuttle launch may be related to the ambient temperature during the launch. The table below summarizes observational data on O-rings for 23 shuttle missions, where the mission order is based on the temperature at the time of the launch. Temp gives the temperature in Fahrenheit, Damaged represents the number of damaged O-rings, and Undamaged represents the number of O-rings that were not damaged.

| Shuttle Mission | 1  | 2  | 3  | 4  | 5  | 6  | 7  | 8  | 9  | 10 | 11 | 12 |
|-----------------|----|----|----|----|----|----|----|----|----|----|----|----|
| Temperature     | 53 | 57 | 58 | 63 | 66 | 67 | 67 | 67 | 68 | 69 | 70 | 70 |
| Damaged         | 5  | 1  | 1  | 1  | 0  | 0  | 0  | 0  | 0  | 0  | 1  | 0  |
| Undamaged       | 1  | 5  | 5  | 5  | 6  | 6  | 6  | 6  | 6  | 6  | 5  | 6  |

| Shuttle Mission | 13 | 14 | 15 | 16 | 17 | 18 | 19 | 20 | 21 | 22 | 23 |
|-----------------|----|----|----|----|----|----|----|----|----|----|----|
| Temperature     | 70 | 70 | 72 | 73 | 75 | 75 | 76 | 76 | 78 | 79 | 81 |
| Damaged         | 1  | 0  | 0  | 0  | 0  | 1  | 0  | 0  | 0  | 0  | 0  |
| Undamaged       | 5  | 6  | 6  | 6  | 6  | 5  | 6  | 6  | 6  | 6  | 6  |

(a) Each column of the table above represents a different shuttle mission. Examine these data and describe what you observe with respect to the relationship between temperatures and damaged O-rings.

(b) Failures have been coded as 1 for a damaged O-ring and 0 for an undamaged O-ring, and a logistic regression model was fit to these data. A summary of this model is given below. Describe the key components of this summary table in words.

|             | Estimate | Std. Error | t value | Pr(>$abs(t)$) |
|-------------|----------|------------|---------|---------------|
| (Intercept) | 11.6630  | 3.2963     | 3.54    | 0.0004        |
| Temperature | -0.2162  | 0.0532     | -4.07   | 0.0000        |

(c) Write out the logistic model using the point estimates of the model parameters.

(d) Based on the model, do you think concerns regarding O-rings are justified? Explain

## 8.18 Challenger disaster, Part II.

Exercise 8.16 introduced us to O-rings that were identified as a plausible explanation for the breakup of the Challenger space shuttle 73 seconds into takeoff
in 1986. The investigation found that the ambient temperature at the time of the shuttle launch was closely related to the damage of O-rings, which are a critical component of the shuttle. See this earlier exercise if you would like to browse the original data.

![ ](https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA606/Figures/8_18.png)

(a) The data provided in the previous exercise are shown in the plot. The logistic model fit to these data may be written as

$log(\frac{\hat{P}}{1- \hat{P}}) = 11.6630 − 0.2162 \times \text{Temperature}$

where $\hat{p}$ is the model-estimated probability that an O-ring will become damaged. Use the model to calculate the probability that an O-ring will become damaged at each of the following ambient temperatures: 51, 53, and 55 degrees Fahrenheit. The model-estimated probabilities for several additional ambient temperatures are provided below, where subscripts indicate the temperature:
$\hat{p}_\text{57}$ = 0.341 $\hat{p}_\text{59}$ = 0.251 $\hat{p}_\text{61}$ = 0.179 $\hat{p}_\text{63}$ = 0.124
$\hat{p}_\text{65}$ = 0.084 $\hat{p}_\text{67}$ = 0.056 $\hat{p}_\text{69}$ = 0.037 $\hat{p}_\text{71}$ = 0.024

(b) Add the model-estimated probabilities from part (a) on the plot, then connect these dots using a smooth curve to represent the model-estimated probabilities.

(c) Describe any concerns you may have regarding applying logistic regression in this application, and note any assumptions that are required to accept the model’s validity

### 2018-11-26 DATA606/Lab8/michael-multiple_regression.Rmd
Additional Work

---
title: "Multiple linear regression"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## Grading the professor

Many college courses conclude by giving students the opportunity to evaluate 
the course and the instructor anonymously. However, the use of these student 
evaluations as an indicator of course quality and teaching effectiveness is 
often criticized because these measures may reflect the influence of 
non-teaching related characteristics, such as the physical appearance of the 
instructor. The article titled, "Beauty in the classroom: instructors' 
pulchritude and putative pedagogical productivity" (Hamermesh and Parker, 2005) 
found that instructors who are viewed to be better looking receive higher 
instructional ratings. (Daniel S. Hamermesh, Amy Parker, Beauty in the 
classroom: instructors pulchritude and  putative pedagogical productivity, 
*Economics of Education Review*, Volume 24, Issue 4, August 2005, Pages 369-376, 
ISSN 0272-7757, 10.1016/j.econedurev.2004.07.013. [http://www.sciencedirect.com/science/article/pii/S0272775704001165](http://www.sciencedirect.com/science/article/pii/S0272775704001165).)

In this lab we will analyze the data from this study in order to learn what goes 
into a positive professor evaluation.

## The data

The data were gathered from end of semester student evaluations for a large 
sample of professors from the University of Texas at Austin. In addition, six 
students rated the professors' physical appearance. (This is a slightly modified 
version of the original data set that was released as part of the replication 
data for *Data Analysis Using Regression and Multilevel/Hierarchical Models* 
(Gelman and Hill, 2007).) The result is a data frame where each row contains a 
different course and columns represent variables about the courses and professors.

```{r load-data, eval=TRUE}
load("more/evals.RData")
```

variable         | description
---------------- | -----------
`score`          | average professor evaluation score: (1) very unsatisfactory - (5) excellent.
`rank`           | rank of professor: teaching, tenure track, tenured.
`ethnicity`      | ethnicity of professor: not minority, minority.
`gender`         | gender of professor: female, male.
`language`       | language of school where professor received education: english or non-english.
`age`            | age of professor.
`cls_perc_eval`  | percent of students in class who completed evaluation.
`cls_did_eval`   | number of students in class who completed evaluation.
`cls_students`   | total number of students in class.
`cls_level`      | class level: lower, upper.
`cls_profs`      | number of professors teaching sections in course in sample: single, multiple.
`cls_credits`    | number of credits of class: one credit (lab, PE, etc.), multi credit.
`bty_f1lower`    | beauty rating of professor from lower level female: (1) lowest - (10) highest.
`bty_f1upper`    | beauty rating of professor from upper level female: (1) lowest - (10) highest.
`bty_f2upper`    | beauty rating of professor from second upper level female: (1) lowest - (10) highest.
`bty_m1lower`    | beauty rating of professor from lower level male: (1) lowest - (10) highest.
`bty_m1upper`    | beauty rating of professor from upper level male: (1) lowest - (10) highest.
`bty_m2upper`    | beauty rating of professor from second upper level male: (1) lowest - (10) highest.
`bty_avg`        | average beauty rating of professor.
`pic_outfit`     | outfit of professor in picture: not formal, formal.
`pic_color`      | color of professor's picture: color, black & white.

## Exploring the data

1.  Is this an observational study or an experiment? **This is an observational study** The original research
    question posed in the paper is whether beauty leads directly to the
    differences in course evaluations. Given the study design, is it possible to
    answer this question as it is phrased? If not, rephrase the question. **No it is not possible.  The question should be phrased to what extent beauty explains the differences in course evaluations.  It is a question of correlation not causation.**

2.  Describe the distribution of `score`. Is the distribution skewed? What does 
    that tell you about how students rate courses? Is this what you expected to 
    see? Why, or why not? **The distribution is left skewed.  This tells me that students are generally very positive of their professors.  I would expect the skill of the professors to be normally distributed.  I am not suprized that the students perception of the professor's skill not be match it because perception is not always reality.**
    ```{r}
    hist(evals$score)
    ```

3.  Excluding `score`, select two other variables and describe their relationship 
    using an appropriate visualization (scatterplot, side-by-side boxplots, or 
    mosaic plot).
    ```{r}
    plot(evals$gender, evals$bty_avg)
    ```

    **Academia attracts people with below average beauty.  The median beauty score on a scale of 1 to 10 is a 4.  Males are less attractive than females.**

    ```{r}
    plot(evals$age, evals$bty_avg)
    ```

    **There really isn't much of a relationship between the average beauty rating and the age of the student.  If you look at the scatterplot minus the couple of 70 year old observations, the image portrays little relationship between the two variables.**

## Simple linear regression

The fundamental phenomenon suggested by the study is that better looking teachers
are evaluated more favorably. Let's create a scatterplot to see if this appears 
to be the case:

```{r scatter-score-bty_avg, eval = FALSE}
plot(evals$score ~ evals$bty_avg)
```

Before we draw conclusions about the trend, compare the number of observations 
in the data frame with the approximate number of points on the scatterplot. 
Is anything awry? **There are `r nrow(evals)` observations in the data frame.  There are not that many points in the graph.  There must be overlapping points.**

4.  Replot the scatterplot, but this time use the function `jitter()` on the
    $y$- or the $x$-coordinate. (Use `?jitter` to learn more.) What was
    misleading about the initial scatterplot?

    **The original scatterplot is misleading because of the overlap**

    ```{r}
    plot(evals$score ~ jitter(evals$bty_avg))
    ```

5.  Let's see if the apparent trend in the plot is something more than
    natural variation. Fit a linear model called `m_bty` to predict average
    professor score by average beauty rating and add the line to your plot
    using `abline(m_bty)`. Write out the equation for the linear model and
    interpret the slope. Is average beauty score a statistically significant
    predictor? Does it appear to be a practically significant predictor?

    **The formula is:**

    $\text{AvgProfScore} = 3.88034 + 0.06664 \times \text{AvgBeautyRating}$

    **The slope indicates for every additional average beauty rating points there is a 0.07 increase in the professor's score.  The coeeficient is statistically significant predictor.  I would argue that there is no practical significance to this finding.**

    ```{r}
    m_bty <- lm(evals$score ~ evals$bty_avg)
    summary(m_bty)
    plot(evals$score ~ jitter(evals$bty_avg))
    abline(m_bty, col="red")
    ```

6.  Use residual plots to evaluate whether the conditions of least squares
    regression are reasonable. Provide plots and comments for each one (see
    the Simple Regression Lab for a reminder of how to make these).

    **Linear relationship.  The previous plot indicates the weak quasi-linear relationship.**

    ```{r}
    plot(m_bty$residuals ~ evals$bty_avg)
    abline(h = 0, lty = 3)
    ```

    **Nearly normal residuals: The plot below shows the Q to Q plot but it doesn't hold to the line the best.**

    ```{r}    
    qqnorm(m_bty$residuals)
    qqline(m_bty$residuals) 
```

## Multiple linear regression

The data set contains several variables on the beauty score of the professor: 
individual ratings from each of the six students who were asked to score the 
physical appearance of the professors and the average of these six scores. Let's 
take a look at the relationship between one of these scores and the average 
beauty score.

```{r bty-rel, eval = FALSE}
plot(evals$bty_avg ~ evals$bty_f1lower)
cor(evals$bty_avg, evals$bty_f1lower)
```

As expected the relationship is quite strong - after all, the average score is 
calculated using the individual scores. We can actually take a look at the 
relationships between all beauty variables (columns 13 through 19) using the 
following command:

```{r bty-rels, eval = FALSE}
plot(evals[,13:19])
```

These variables are collinear (correlated), and adding more than one of these 
variables to the model would not add much value to the model. In this 
application and with these highly-correlated predictors, it is reasonable to use
the average beauty score as the single representative of these variables.

In order to see if beauty is still a significant predictor of professor score 
after we've accounted for the gender of the professor, we can add the gender 
term into the model.

```{r scatter-score-bty_avg_gender, eval = FALSE}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```

7.  P-values and parameter estimates should only be trusted if the
    conditions for the regression are reasonable. Verify that the conditions
    for this model are reasonable using diagnostic plots.

8.  Is `bty_avg` still a significant predictor of `score`? Has the addition
    of `gender` to the model changed the parameter estimate for `bty_avg`?

Note that the estimate for `gender` is now called `gendermale`. You'll see this 
name change whenever you introduce a categorical variable. The reason is that R 
recodes `gender` from having the values of `female` and `male` to being an 
indicator variable called `gendermale` that takes a value of $0$ for females and
a value of $1$ for males. (Such variables are often referred to as "dummy" 
variables.)

As a result, for females, the parameter estimate is multiplied by zero, leaving 
the intercept and slope form familiar from simple regression.

\[
  \begin{aligned}
\widehat{score} &= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times (0) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg\end{aligned}
\]

We can plot this line and the line corresponding to males with the following 
custom function.

```{r twoLines, eval = FALSE}
multiLines(m_bty_gen)
```

9.  What is the equation of the line corresponding to males? (*Hint:* For
    males, the parameter estimate is multiplied by 1.) For two professors
    who received the same beauty rating, which gender tends to have the
    higher course evaluation score?

The decision to call the indicator variable `gendermale` instead of`genderfemale`
has no deeper meaning. R simply codes the category that comes first 
alphabetically as a $0$. (You can change the reference level of a categorical 
variable, which is the level that is coded as a 0, using the`relevel` function. 
Use `?relevel` to learn more.)

10. Create a new model called `m_bty_rank` with `gender` removed and `rank` 
    added in. How does R appear to handle categorical variables that have more 
    than two levels? Note that the rank variable has three levels: `teaching`, 
    `tenure track`, `tenured`.

The interpretation of the coefficients in multiple regression is slightly 
different from that of simple regression. The estimate for `bty_avg` reflects
how much higher a group of professors is expected to score if they have a beauty
rating that is one point higher *while holding all other variables constant*. In
this case, that translates into considering only professors of the same rank 
with `bty_avg` scores that are one point apart.

## The search for the best model

We will start with a full model that predicts professor score based on rank, 
ethnicity, gender, language of the university where they got their degree, age, 
proportion of students that filled out evaluations, class size, course level, 
number of professors, number of credits, average beauty rating, outfit, and 
picture color.

11. Which variable would you expect to have the highest p-value in this model? 
    Why? *Hint:* Think about which variable would you expect to not have any 
    association with the professor score.

Let's run the model...

```{r m_full, eval = FALSE, tidy = FALSE}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

12. Check your suspicions from the previous exercise. Include the model output
    in your response.

13. Interpret the coefficient associated with the ethnicity variable.

14. Drop the variable with the highest p-value and re-fit the model. Did the
    coefficients and significance of the other explanatory variables change?
    (One of the things that makes multiple regression interesting is that
    coefficient estimates depend on the other variables that are included in
    the model.) If not, what does this say about whether or not the dropped
    variable was collinear with the other explanatory variables?

15. Using backward-selection and p-value as the selection criterion,
    determine the best model. You do not need to show all steps in your
    answer, just the output for the final model. Also, write out the linear
    model for predicting score based on the final model you settle on.

16. Verify that the conditions for this model are reasonable using diagnostic 
    plots.

17. The original paper describes how these data were gathered by taking a
    sample of professors from the University of Texas at Austin and including 
    all courses that they have taught. Considering that each row represents a 
    course, could this new information have an impact on any of the conditions 
    of linear regression?

18. Based on your final model, describe the characteristics of a professor and 
    course at University of Texas at Austin that would be associated with a high
    evaluation score.

19. Would you be comfortable generalizing your conclusions to apply to professors
    generally (at any university)? Why or why not?

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). This lab was written by 
Mine &Ccedil;etinkaya-Rundel and Andrew Bray.
</div>

### 2018-11-27 DATA606/Lab8/michael-multiple_regression.Rmd
More work

@@
 
-    **Academia attracts people with below average beauty.  The median beauty score on a scale of 1 to 10 is a 4.  Males are less attractive than females.**
+    **The median beauty score is a 4 on a scale of 1 to 10.  Males are less attractive than females.**
 
@@
     ```{r}    
+    qqnorm(m_bty$residuals)
+    qqline(m_bty$residuals) 
+    ```
+
+    **Nearly normal residuals: The plot abovew shows the Q to Q plot which is mostly normal, but the skew in the data is apparent in the figure.**
+
+    ```{r}
     plot(m_bty$residuals ~ evals$bty_avg)
@@
 
-    **Nearly normal residuals: The plot below shows the Q to Q plot but it doesn't hold to the line the best.**
-
-    ```{r}    
-    qqnorm(m_bty$residuals)
-    qqline(m_bty$residuals) 
-```
+    **Constant variability: The preceeding plot shows no apparent pattern in the residuals.**
 
@@
 
-```{r scatter-score-bty_avg_gender, eval = FALSE}
+```{r scatter-score-bty_avg_gender}
 m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
@@
 
-8.  Is `bty_avg` still a significant predictor of `score`? Has the addition
-    of `gender` to the model changed the parameter estimate for `bty_avg`?
+    ```{r}
+    plot(m_bty_gen$residuals ~ evals$bty_avg)
+    abline(h = 0, lty = 3)
+    ```
+
+    **The residuals have a constant variability.**
+
+    ```{r}
+    qqnorm(m_bty_gen$residuals)
+    qqline(m_bty_gen$residuals) 
+    ```
+
+    **The Q-Q plot is more linear than the previous one.**
+
+8.
+Is `bty_avg` still a significant predictor of `score`?
+**The predictor is statistically significant.** Has the addition of `gender` to the model changed the parameter estimate for `bty_avg`?
+**Yes.
+It is now 0.07416 up from 0.06664.**
 
@@
 
+    **All esle equal the male professor would have a higher course evaluation score.**
+
 The decision to call the indicator variable `gendermale` instead of`genderfemale`
@@
     `tenure track`, `tenured`.
+
+    ```{r}
+    m_bty_rank <- lm(score ~ bty_avg + rank, data = evals)
+    summary(m_bty_rank)
+    ```
 

### 2018-11-27 DATA606/Lab8/michael-multiple_regression.Rmd
Completed Lab 8

@@
 
-```{r scatter-score-bty_avg, eval = FALSE}
+```{r scatter-score-bty_avg}
 plot(evals$score ~ evals$bty_avg)
@@
 
-    $\text{AvgProfScore} = 3.88034 + 0.06664 \times \text{AvgBeautyRating}$
+    $\widehat{score} = 3.88034 + 0.06664 \times \text{bty_avg}$
 
@@
 
-```{r bty-rel, eval = FALSE}
+```{r bty-rel}
 plot(evals$bty_avg ~ evals$bty_f1lower)
@@
 
-```{r bty-rels, eval = FALSE}
+```{r bty-rels}
 plot(evals[,13:19])
@@
 
-```{r twoLines, eval = FALSE}
+```{r twoLines}
 multiLines(m_bty_gen)
@@
     Why? *Hint:* Think about which variable would you expect to not have any 
-    association with the professor score.
+    association with the professor score. **cls_profs.  The number of professors teaching a course should have little to no bearing on their score.**
 
@@
 
-```{r m_full, eval = FALSE, tidy = FALSE}
+```{r m_full}
 m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
@@
     in your response. 
+    **cls_profs has the highest P value 0.77806 meaning this is not a statistically significant variable.**
 
 13. Interpret the coefficient associated with the ethnicity variable. 
+    **The students found professors that are not a minority get a higher score than those who are.  This finding is not statistically significant.**
 
@@
     variable was collinear with the other explanatory variables?
+    **Yes the coefficients and significances changed slightly.  For example the ethnicity variable is now statistically significant.**
+    ```{r}
+    m_nearly_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval + cls_students + cls_level + cls_credits + bty_avg 
+             + pic_outfit + pic_color, data = evals)
+summary(m_nearly_full)
+    ```
 
@@
     model for predicting score based on the final model you settle on.
+    **The formula is:**
+    $\widehat{score} = 3.69 + 0.22 ethnicity + 0.20 gender - 0.01 age + 0.005 \text{cls_perc_eval} + \\0.53 \text{cls_credits} + 0.05 \text{bty_avg} - 0.17  \text{pic_color}$
+
+    ```{r}
+    m_best <- lm(score ~ ethnicity + gender + age + cls_perc_eval + cls_credits + bty_avg + pic_color, data=evals)
+    summary(m_best)
+    ```
 
@@
     plots.
+
+    ```{r}
+    plot(m_best)
+    ```
+
+    **Linearity: Red line in Residuals vs Fitted somewhat flat so condition is met.**
+
+    **Constant variation: Points in Residuals vs Fitted plot show no apparant pattern generally therefore the condition is met.***
+
+    **Nearly normal residuals: Points on Q to Q plot fairly linear so this condition is met.  There is some issues at the tails though.**
 
@@
 
+    **No because the responses are from the student's point-of-view so they are still  independent.**
+
 18. Based on your final model, describe the characteristics of a professor and 
@@
 
+    **An attractive young white male who teaches a one credit class and gets most of the students to complete an evaluation featured on a black and white head shot.**
+
 19. Would you be comfortable generalizing your conclusions to apply to professors
     generally (at any university)? Why or why not?
+
+    **No.  Only 6 people rated the beauty of the professors.  This is not a large enough sample size to generalize the model.**
 

### 2018-11-28 DATA606/Multiple and Logistic Regression.Rmd
Work on week 8 assignment

@@
 
-(a) Write the equation of the regression line.
+(a) Write the equation of the regression line. $\widehat{weight}=120.07 - 1.93 \times parity$
 
 (b) Interpret the slope in this context, and calculate the predicted birth weight of first borns and others.
+**All else being equal, the average birth weight is 120 ounces.
+If the child is not the first born the birth weight decreases by two ounces, to 118 ounces.**
 
-(c) Is there a statistically significant relationship between the average birth weight and parity?
+(c) Is there a statistically significant relationship between the average birth weight and parity? **No because the P value is > 0.05**
 
@@
 
-The summary table below shows the results of a linear regression model for predicting the average number of days absent based on ethnic background (eth: 0 - aboriginal, 1 - not aboriginal), sex (sex: 0 - female, 1 - male), and learner status (lrn: 0 - average learner, 1 - slow learner).1
+The summary table below shows the results of a linear regression model for predicting the average number of days absent based on ethnic background (eth: 0 - aboriginal, 1 - not aboriginal), sex (sex: 0 - female, 1 - male), and learner status (lrn: 0 - average learner, 1 - slow learner).
 
@@
 
-(a) Write the equation of the regression line.
+(a) Write the equation of the regression line. $\widehat{day} = 18.93 -9.11eth + 3.1sex + 2.15lrn$
 
 (b) Interpret each one of the slopes in this context.
+**On average a student misses about 19 days of school.
+If they are aboriginal they miss an additional 9 days.
+If they are female they miss 3 more days.
+If they are not a slow learner they miss 2 additional days.**
 
 (c) Calculate the residual for the first observation in the data set: a student who is aboriginal, male, a slow learner, and missed 2 days of school.
+
+$\widehat{day} = 18.93 -9.11(0) + 3.1(1) + 2.15(1) = 24.18$
+
+$e_i = day_1 - \widehat{day}_1 = 2 - 24.18 = -22.18$
 
@@
 Note that there are 146 observations in the data set.
+
+$R^2 = 1 - \frac{Var(e_1)}{Var(y_1)} = 1 - \frac{240.57}{264.17} = 0.0893364$
+
+$R^2_\text{adj} = 1 - \frac{Var(e_1)}{Var(y_1)} \times \frac{(n-1)}{(n-k-1)} = 1- \frac{240.57}{264.17} \times \frac{(146-1)}{(146 - 3 - 1)} = 1 - 0.9106636 \times 1.021127 = 0.07009681$
 
@@
 
-Which, if any, variable should be removed from the model first?
+Which, if any, variable should be removed from the model first? **lrn because the adjusted** $R^2$ **increases from 0.0701 to 0.0723**
 
@@
 Seventy-three seconds into the flight, disaster happened: the shuttle broke apart, killing all seven crew members on board.
-An investigation into the cause of the
-disaster focused on a critical seal called an O-ring, and it is believed that damage to these O-rings during a shuttle launch may be related to the ambient temperature during the launch.
+An investigation into the cause of the disaster focused on a critical seal called an O-ring, and it is believed that damage to these O-rings during a shuttle launch may be related to the ambient temperature during the launch.
 The table below summarizes observational data on O-rings for 23 shuttle missions, where the mission order is based on the temperature at the time of the launch.
@@
 
-(a) Each column of the table above represents a different shuttle mission. Examine these data and describe what you observe with respect to the relationship between temperatures and damaged O-rings.
+(a) Each column of the table above represents a different shuttle mission.
+Examine these data and describe what you observe with respect to the relationship between temperatures and damaged O-rings.
+**At 53 to 57 degrees Fahrenheit, the ratio of damaged to undamaged O-rings flip.**
 
@@
 
+**The model is the probability of failure given the ambient air temperature at launch time.**
+
 (c) Write out the logistic model using the point estimates of the model parameters.
 
-(d) Based on the model, do you think concerns regarding O-rings are justified? Explain
+$log(\frac{p}{1-p}) = 11.6630 - 0.2162 \times \text{Temperature}$
+
+(d) Based on the model, do you think concerns regarding O-rings are justified? Explain.
+
+**Yes.
+Temperature is a statistically significant variable.
+To see this effect consider the following:  What is the probability of failure when the launch temperature is 81 degrees?
+What is the probability of failure at 53?**
+
+$P = \frac{e^L}{1 + e^L}$ **where**
+$L = 11.6630 - 0.2162 \times \text{Temperature}$
+
+```{r}
+probability_of_failure <- function(temperature){
+  l <- 11.6630 - 0.2162 * temperature
+  round(exp(l) / (1 + exp(l)), 3)
+}
+
+probability_of_failure(81)
+probability_of_failure(53)
+```
+
+**The probability of failure increases from 0 to 0.6 within the observed extreme temperatures.**
 
@@
 
-Exercise 8.16 introduced us to O-rings that were identified as a plausible explanation for the breakup of the Challenger space shuttle 73 seconds into takeoff
-in 1986.
+Exercise 8.16 introduced us to O-rings that were identified as a plausible explanation for the breakup of the Challenger space shuttle 73 seconds into takeoff in 1986.
 The investigation found that the ambient temperature at the time of the shuttle launch was closely related to the damage of O-rings, which are a critical component of the shuttle.
@@
 
-$log(\frac{\hat{P}}{1- \hat{P}}) = 11.6630 − 0.2162 \times \text{Temperature}$
+$log(\frac{\hat{p}}{1- \hat{p}}) = 11.6630 − 0.2162 \times \text{Temperature}$
 
@@
 Use the model to calculate the probability that an O-ring will become damaged at each of the following ambient temperatures: 51, 53, and 55 degrees Fahrenheit.
+
+* $\hat{p}_\text{51}$ = `r probability_of_failure(51)`
+* $\hat{p}_\text{53}$ = `r probability_of_failure(53)`
+* $\hat{p}_\text{55}$ = `r probability_of_failure(55)`
+
 The model-estimated probabilities for several additional ambient temperatures are provided below, where subscripts indicate the temperature:
-$\hat{p}_\text{57}$ = 0.341 $\hat{p}_\text{59}$ = 0.251 $\hat{p}_\text{61}$ = 0.179 $\hat{p}_\text{63}$ = 0.124
-$\hat{p}_\text{65}$ = 0.084 $\hat{p}_\text{67}$ = 0.056 $\hat{p}_\text{69}$ = 0.037 $\hat{p}_\text{71}$ = 0.024
+
+* $\hat{p}_\text{57}$ = 0.341 
+* $\hat{p}_\text{59}$ = 0.251 
+* $\hat{p}_\text{61}$ = 0.179 
+* $\hat{p}_\text{63}$ = 0.124
+* $\hat{p}_\text{65}$ = 0.084 
+* $\hat{p}_\text{67}$ = 0.056 
+* $\hat{p}_\text{69}$ = 0.037 
+* $\hat{p}_\text{71}$ = 0.024
 
@@
 
+```{r}
+library(ggplot2)
+temp <- seq( 51, 71, 2)
+p <- probability_of_failure(temp)
+df <- data.frame(temperature = temp, probability_of_failure = p)
+
+spline_int <- as.data.frame(spline(df$temp, df$probability_of_failure))
+
+ggplot(df) + 
+  geom_point(aes(x = temp, y = probability_of_failure), size = 3) +
+  geom_line(data = spline_int, aes(x = x, y = y)) +
+  labs(x = "Temperature (degrees Fahrenheit)", y = "Probability of Failure")
+```
+
 (c) Describe any concerns you may have regarding applying logistic regression in this application, and note any assumptions that are required to accept the model’s validity
+
+**This model may underestimate the risk at the higher temperatures.
+From the observational data we see 2 of 3 cases at 70 degrees had a failure.
+The model would predict `r probability_of_failure(70)`.
+Since we are dealing with human lives an abundance of caution should be exercised.**
+
+**There are two assumptions required to accept the model's validity: Each outcome is independent of the other outcomes and each predictor is linearly related if all other predictors are held constant.
+The outcomes might not be independent as different manufacturing processes might be employed in response to a previous shuttle launch.
+It also doesn't appear that the risk of failure and the temperature are linearly related.**

### 2018-12-04 DATA607/Final-Project/1 Build Monroe Real Property Data.py
Initial commit of DATA 607 Final Files

@@ -1,0 +1,137 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Nov  5 13:26:41 2018
+
+@author: Michael Silva
+"""
+
+import urllib
+import zipfile
+import shutil
+from dbfread import DBF
+import sqlite3 as db
+import requests
+import asyncio
+import time
+import pandas as pd
+from bs4 import BeautifulSoup
+import re
+from datetime import datetime
+import os.path
+
+create_db_from_scratch = False
+
+start_time = time.time()
+
+db_name = "Monroe Real Property Data.db"
+con = db.connect(db_name)
+con.row_factory = lambda cursor, row: row[0]
+c = con.cursor()
+
+if create_db_from_scratch:
+    print('Setting up db')
+    c.execute('DROP TABLE IF EXISTS `scraped_data`')
+    c.execute('CREATE TABLE `scraped_data` (`SWIS`, `SBL`, `url`,`html`)')
+    c.execute('CREATE INDEX `scraped` ON `scraped_data` (`SWIS`, `SBL`);')
+    c.execute('DROP TABLE IF EXISTS `sales_data`')
+    c.execute('CREATE TABLE `sales_data` (`sale_date` DATE, `price` REAL, `SWIS`,`SBL`)')
+    c.execute('CREATE INDEX `sales` ON `sales_data` (`SWIS`, `SBL`);')
+    con.commit()
+    print("--- %s seconds ---" % (time.time() - start_time))
+
+print('Getting the Monroe County Tax Parcels Centroids')
+dbf_file_name = 'Monroe_2017_Tax_Parcel_Centroid_Points_SHP.dbf'
+if not os.path.isfile(dbf_file_name):
+    url = 'http://gis.ny.gov/gisdata/fileserver/?DSID=1300&file=Monroe-Tax-Parcels-Centroid-Points-SHP.zip'
+    zip_file_name = url.split('file=')[1]
+    if not os.path.isfile(dbf_file_name):
+        with urllib.request.urlopen(url) as response, open(zip_file_name, 'wb') as out_file:
+            shutil.copyfileobj(response, out_file)
+        with zipfile.ZipFile(zip_file_name) as zf:
+            zf.extractall()
+print("--- %s seconds ---" % (time.time() - start_time))
+
+print('Building list of urls to scrape')
+urls_to_scrape = list()
+prop_info = list()
+
+for record in DBF(dbf_file_name):
+    try:
+        if record['PROP_CLASS'][0] == '2':
+            prop_info.append(record)
[…]

### 2018-12-04 DATA607/Final-Project/2 Create Model.R

@@ -1,0 +1,132 @@
+library(DBI)
+library(RSQLite)
+library(dplyr)
+library(stringr)
+library(caret)
+library(ranger)
+
+## Final Data Wrangling
+
+# Get the SQLite database
+if(!file.exists("Monroe Real Property Data.db")){
+  if(!file.exists("Monroe Real Property Data.db.tar.gz")){
+    # Download this specific commit from my GitHub repo
+    download.file("https://github.com/mikeasilva/CUNY-SPS/raw/9a2023b57a466550646c9190076a53d51ef09ef2/data/Monroe%20Real%20Property%20Data.db.tar.gz", "Monroe Real Property Data.db.tar.gz")
+  }
+  # Decompress the file
+  untar("Monroe Real Property Data.db.tar.gz")
+  # Delete the zipped file
+  unlink("Monroe Real Property Data.db.tar.gz")
+}
+
+# Connect to the database
+con <- dbConnect(RSQLite::SQLite(), "Monroe Real Property Data.db")
+
+# Get the latest unduplicated sales price data
+results <- dbSendQuery(con, "SELECT MAX(sale_date) AS sale_date, SWIS, SBL FROM sales_data GROUP BY SWIS, SBL HAVING sale_date > '2015-12-31'")
+df <- dbFetch(results) 
+
+dbClearResult(results)
+
+df <- dbReadTable(con, "sales_data") %>% # Pull all the records
+  distinct() %>% # remove any duplicates
+  merge(df) # Merge with the targeted sales observations
+
+df <- df %>%
+  group_by(SWIS, SBL) %>% 
+  summarise(count = n()) %>% # Get the number of records
+  ungroup() %>%
+  filter(count == 1) %>% # Only keep the observations with one price
+  select(-count) %>% # Drop the count column
+  merge(df) # subset the price observations
+
+df <- dbReadTable(con, "property_info") %>% # Pull all the housing attributes
+  distinct() %>% # remove any duplicates
+  merge(df) # Merge with the targeted sales observations
+
+# Disconnect from the database
+dbDisconnect(con)
+
+# Clean up the data
+df <- df %>%
+  select(-sale_date, -index, -SBL, -SWIS, -SQ_FT, -ACRES) %>% # Drop the unneeded columns
+  filter(YR_BLT > 0) %>% # Must have a year it was built
+  filter(SQFT_LIV > 0) %>% # Must have a square footage
+  filter(NBR_KITCHN > 0) %>% # Must have a kitchen
[…]

### 2018-12-04 DATA607/Final-Project/3 Create Shiny App.R

@@ -1,0 +1,242 @@
+library(shiny)
+library(dplyr)
+library(ggplot2)
+library(ggthemes)
+library(tidyr)
+library(caret)
+library(ranger)
+library(FNN)
+
+load("model.RData")
+
+# Get the unique set of architectural styles
+building_description_options <- testing %>%
+  select(BLDG_DESC) %>%
+  arrange(BLDG_DESC) %>%
+  unique() 
+building_description_options <- building_description_options$BLDG_DESC
+
+# Get the architectural style that is most prevalant in the data
+building_description_mode <- testing %>%
+  group_by(BLDG_DESC) %>%
+  summarise(count = n()) %>%
+  arrange(-count)
+building_description_mode <- building_description_mode[1, ]$BLDG_DESC
+
+# Define the User Interface
+ui <- navbarPage("Rochester Housing Sale Price",
+                 tabPanel("Get a Prediction",
+                          # List the housing price determinants
+                          sidebarPanel(
+                            # The size in square feet
+                            sliderInput("SQFT_LIV", 
+                                        "Square Footage:",
+                                        min = min(testing$SQFT_LIV), 
+                                        max = max(testing$SQFT_LIV),
+                                        value = median(testing$SQFT_LIV), 
+                                        step = 1,
+                                        pre = "", 
+                                        sep = ",",
+                                        animate = TRUE),
+                            
+                            # Display the acres from the data
+                            sliderInput("CALC_ACRES", 
+                                        "Plot Size (Acres):",
+                                        min = round(min(testing$CALC_ACRES), 1), 
+                                        max = round(max(testing$CALC_ACRES), 1),
+                                        value = round(median(testing$CALC_ACRES), 1), 
+                                        step = 0.1,
+                                        pre = "", 
+                                        sep = "",
+                                        animate = TRUE),
+                            
+                            # Display the year from the data
+                            sliderInput("YR_BLT", 
+                                        "Year Built:",
+                                        min = min(testing$YR_BLT), 
+                                        max = max(testing$YR_BLT),
+                                        value = median(testing$YR_BLT), 
[…]

### 2018-12-04 DATA607/Final-Project/App.R
More work on Shiny App

@@ -1,0 +1,227 @@
+library(shiny)
+library(dplyr)
+library(e1071)
+library(ggplot2)
+library(ggthemes)
+library(tidyr)
+library(caret)
+library(ranger)
+library(FNN)
+
+load("model.RData")
+
+# Get the unique set of architectural styles
+building_description_options <- testing %>%
+  select(BLDG_DESC) %>%
+  arrange(BLDG_DESC) %>%
+  unique()
+building_description_options <- building_description_options$BLDG_DESC
+
+# Get the architectural style that is most prevalant in the data
+building_description_mode <- testing %>%
+  group_by(BLDG_DESC) %>%
+  summarise(count = n()) %>%
+  arrange(-count)
+building_description_mode <- building_description_mode[1, ]$BLDG_DESC
+
+# Define the User Interface
+ui <- navbarPage("Rochester Housing Sale Price",
+                 tabPanel("Get a Prediction",
+                          # List the housing price determinants
+                          sidebarPanel(
+                            # The size in square feet
+                            sliderInput("SQFT_LIV",
+                                        "Square Footage:",
+                                        min = min(testing$SQFT_LIV),
+                                        max = max(testing$SQFT_LIV),
+                                        value = median(testing$SQFT_LIV),
+                                        step = 1,
+                                        pre = "",
+                                        sep = ",",
+                                        animate = TRUE),
+                            # Display the acres from the data
+                            sliderInput("CALC_ACRES",
+                                        "Plot Size (Acres):",
+                                        min = round(min(testing$CALC_ACRES), 1),
+                                        max = round(max(testing$CALC_ACRES), 1),
+                                        value = round(median(testing$CALC_ACRES), 1),
+                                        step = 0.1,
+                                        pre = "",
+                                        sep = "",
+                                        animate = TRUE),
+                            # Display the year from the data
+                            sliderInput("YR_BLT",
+                                        "Year Built:",
+                                        min = min(testing$YR_BLT),
+                                        max = max(testing$YR_BLT),
+                                        value = median(testing$YR_BLT),
+                                        step = 1,
+                                        pre = "",
+                                        sep = "",
[…]

### 2018-12-04 DATA607/Final-Project/3 Create Shiny App.R
Deleted old files

@@ -1,242 +1,0 @@
-library(shiny)
-library(dplyr)
-library(ggplot2)
-library(ggthemes)
-library(tidyr)
-library(caret)
-library(ranger)
-library(FNN)
-
-load("model.RData")
-
-# Get the unique set of architectural styles
-building_description_options <- testing %>%
-  select(BLDG_DESC) %>%
-  arrange(BLDG_DESC) %>%
-  unique() 
-building_description_options <- building_description_options$BLDG_DESC
-
-# Get the architectural style that is most prevalant in the data
-building_description_mode <- testing %>%
-  group_by(BLDG_DESC) %>%
-  summarise(count = n()) %>%
-  arrange(-count)
-building_description_mode <- building_description_mode[1, ]$BLDG_DESC
-
-# Define the User Interface
-ui <- navbarPage("Rochester Housing Sale Price",
-                 tabPanel("Get a Prediction",
-                          # List the housing price determinants
-                          sidebarPanel(
-                            # The size in square feet
-                            sliderInput("SQFT_LIV", 
-                                        "Square Footage:",
-                                        min = min(testing$SQFT_LIV), 
-                                        max = max(testing$SQFT_LIV),
-                                        value = median(testing$SQFT_LIV), 
-                                        step = 1,
-                                        pre = "", 
-                                        sep = ",",
-                                        animate = TRUE),
-                            
-                            # Display the acres from the data
-                            sliderInput("CALC_ACRES", 
-                                        "Plot Size (Acres):",
-                                        min = round(min(testing$CALC_ACRES), 1), 
-                                        max = round(max(testing$CALC_ACRES), 1),
-                                        value = round(median(testing$CALC_ACRES), 1), 
-                                        step = 0.1,
-                                        pre = "", 
-                                        sep = "",
-                                        animate = TRUE),
-                            
-                            # Display the year from the data
-                            sliderInput("YR_BLT", 
-                                        "Year Built:",
-                                        min = min(testing$YR_BLT), 
-                                        max = max(testing$YR_BLT),
-                                        value = median(testing$YR_BLT), 
[…]

### 2018-12-04 DATA607/Final-Project/2 Create Model.R
Improved readability

@@ -8,8 +8,8 @@ library(ranger)
 ## Final Data Wrangling
 
 # Get the SQLite database
-if(!file.exists("Monroe Real Property Data.db")){
-  if(!file.exists("Monroe Real Property Data.db.tar.gz")){
+if (!file.exists("Monroe Real Property Data.db")){
+  if (!file.exists("Monroe Real Property Data.db.tar.gz")){
     # Download this specific commit from my GitHub repo
     download.file("https://github.com/mikeasilva/CUNY-SPS/raw/9a2023b57a466550646c9190076a53d51ef09ef2/data/Monroe%20Real%20Property%20Data.db.tar.gz", "Monroe Real Property Data.db.tar.gz")
   }
@@ -24,7 +24,7 @@ con <- dbConnect(RSQLite::SQLite(), "Monroe Real Property Data.db")
 
 # Get the latest unduplicated sales price data
 results <- dbSendQuery(con, "SELECT MAX(sale_date) AS sale_date, SWIS, SBL FROM sales_data GROUP BY SWIS, SBL HAVING sale_date > '2015-12-31'")
-df <- dbFetch(results) 
+df <- dbFetch(results)
 
 dbClearResult(results)
 
@@ -33,7 +33,7 @@ df <- dbReadTable(con, "sales_data") %>% # Pull all the records
   merge(df) # Merge with the targeted sales observations
 
 df <- df %>%
-  group_by(SWIS, SBL) %>% 
+  group_by(SWIS, SBL) %>%
   summarise(count = n()) %>% # Get the number of records
   ungroup() %>%
   filter(count == 1) %>% # Only keep the observations with one price
@@ -69,21 +69,22 @@ set.seed(1234)
 
 # Split the data into training and test sets
 in_training <- createDataPartition(df$price, p = 0.8, list = FALSE)
-training <- df[ in_training,]
-testing  <- df[-in_training,]
+training <- df[ in_training, ]
+testing  <- df[-in_training, ]
 
 # Simplify the model using recursive feature elimination
-outcome <- 'price'
+outcome <- "price"
 predictors <- names(training)[!names(training) %in% outcome]
 
-#Sys.time()
-#results <- rfe(training[, predictors], training[, outcome], rfeControl = rfeControl(rfFuncs, method = "repeatedcv", repeats = 3))
-#Sys.time()
+# This takes a long time to run so you may want to comment out the next 3 lines
+Sys.time()
[…]

### 2018-12-04 DATA607/Final-Project/2 Create Model.R
Removed training from model.RData

@@ -128,6 +128,7 @@ rm(in_training)
 rm(outcome)
 rm(predictors)
 rm(results)
+rm(training)
 
 ## Save the environment
 save.image(file = "model.RData")

### 2018-12-06 DATA606/Silva-Project.Rmd
Initial commit

---
title: "CUNY SPS DATA-606 Final Project"
author: "Mike Silva"
date: "November 28, 2018"
output: 
  rmdformats::readthedown:
    highlight: kate
bibliography: Silva-Project.bib
---

## 1. Introduction

Finding affordable housing is a real challenge throughout the country.  The Rochester, New York metro is acclaimed as a place with affordable housing. Recent data show 81% of the housing in the Rochester Metro is affordable [@nahb_hoi].  The median sales price for a home in Rochester is $146,000.  This is about half as much as the national average ($268,000).  With the high family income, housing proves to be affordable.

There is considerable variety in the type of housing offered in Rochester. There are picturesque village homes built in the late 1800s.  There are also suburban homes that are larger and newer. Can the square footage and age of a home reliabily estimate the sale price?

## 2. Data

### Data Collection

The NYS Office of Real Property Services (NYS RPS) collects sales price data from adminstrative records.  The data for this study were scrapped using a [web scrapper](https://github.com/mikeasilva/CUNY-SPS/blob/21abb87c5a28204e75ba2944c574889e07289345/data/Build%20Monroe%20Real%20Property%20Data.py) in early November of 2018 [@mcrpp].  At that time the data was synced with NYS RPS on 22 October 2018.

The NYS GIS clearinghouse compiled data on housing attributes [@nysgis]. The 2017 data is used in this study.  The square footage of living space is the measurement of size in this study.  It is a quantitative measure.

Data on the age of the home are transformed into eras (a categorical variable).  These eras come from an [article](https://www.forbes.com/sites/trulia/2013/05/02/american-homes-by-decade/#16c559821045) from the Trulia's Chief Economist.  They are:

1. Before 1940s
2. Post-War Boom (1940 to 1979)
3. Last 30+ Years (1980+)

This study is purely observational in nature.  It does not assert that these variables have a causal relationship nor should it be interpreted that way.

Scope of inference - generalizability: Identify the population of interest, and whether the findings from this analysis can be generalized to that population, or, if not, a subsection of that population. Explain why or why not. Also discuss any potential sources of bias that might prevent generalizability.

### Data Preparation

The scrapped data is in a SQLite database stored in GitHub.  The first step is to download and extract this database.

```{r get_the_database, message=FALSE}
library(DBI)
library(dplyr)

# Download the database
if(!file.exists("Monroe Real Property Data.db")){
  if(!file.exists("Monroe Real Property Data.db.tar.gz")){
    # Download this specific commit from my GitHub repo
    download.file("https://github.com/mikeasilva/CUNY-SPS/raw/9a2023b57a466550646c9190076a53d51ef09ef2/data/Monroe%20Real%20Property%20Data.db.tar.gz", "Monroe Real Property Data.db.tar.gz")
  }
  # Decompress the file
  untar("Monroe Real Property Data.db.tar.gz")
  # Remove the file
  unlink("Monroe Real Property Data.db.tar.gz")
}
```

Now that the data is downloaded it is time to pull the sales price.  Each parcel is identified by a SWIS (Statewide Information System) and SBL (Section, Block, Lot) code.  Using this identifier, the next step is to find the latest sale date.

```{r read_db}
# Connect to the database
con <- dbConnect(RSQLite::SQLite(), "Monroe Real Property Data.db")

# Get the date of the latest sale for each property
results <- dbSendQuery(con, "SELECT MAX(sale_date) AS sale_date, SWIS, SBL FROM sales_data GROUP BY SWIS, SBL HAVING sale_date BETWEEN '2017-01-01' AND '2017-12-31'")
df <- dbFetch(results)

dbClearResult(results)
```

Now the sale price can be pulled.  There should be `r prettyNum(nrow(df), big.mark = ",", scientific = FALSE)` sale prices in the data set after the next step.

```{r sales_data}
df <- dbReadTable(con, "sales_data") %>% # Pull all the records
  distinct() %>% # remove any duplicates
  merge(df) %>% # Merge with the latest sale dates
  rename(Price = price)
```

The observations total `r prettyNum(nrow(df), big.mark = ",", scientific = FALSE)`.  It appears the merging process introduced some duplication into the data set.  This needs to be corrected.  Any parcel with more than one record will be dropped from the dataset.

```{r deduplication}
df <- df %>%
  select(SWIS, SBL) %>% # Select the unique identifiers
  group_by(SWIS, SBL) %>% 
  summarise(count = n()) %>% # Get the number of records
  ungroup() %>%
  filter(count == 1) %>% # Only keep the observations with one price
  select(-count) %>% # Drop the count variable
  merge(df) # Merge it with the original data frame to subset it
```

This leaves `r prettyNum(nrow(df), big.mark = ",", scientific = FALSE)` unduplicated observations in the data set. Now that we have the outcome variable, it is time to add in the housing attributes.

```{r housing_attributes}
# Add in the housing attributes
df <- dbReadTable(con, "property_info") %>% # Pull all the records
  distinct() %>% # remove any duplicates
  merge(df) # Merge with the sale price

# Disconnect from the database
dbDisconnect(con)
```

There are `r prettyNum(nrow(df), big.mark = ",", scientific = FALSE)` observations in the data set. This indicates the merging process did not introduce any dupplication. Now that the housing attributes are added in, some odd observations need to be filtered out.

```{r filter_out_crap}
df <- df %>% # First filter out some obviously weird data
  filter(YR_BLT > 0) %>%
  filter(SQFT_LIV > 0) %>%
  filter(NBR_KITCHN > 0) %>%
  filter(NBR_BEDRM > 0)
```

After filtering there are `r prettyNum(nrow(df), big.mark = ",", scientific = FALSE)` observations.  This is the final count of cases for the study.  The last data preperation step is to change the year built quantitative variable into a categorical variable and only keep the variables of interest.

```{r finalize_df}
get_era <- function(year){
  if(year < 1940){
    return("1. Before 1940s")
  } else if(year < 1980){
    return("2. Post-War Boom")
  }else{
    return("3. Last 30+ Years")
  }
}

df <- df %>% 
  rowwise() %>%
  mutate(Era = get_era(YR_BLT)) %>%
  ungroup() %>%
  mutate(Era = as.factor(Era)) %>%
  rename(Size = SQFT_LIV) %>%
  select(Price, Size, Era)
```

### Sanity Check

The American Community survey asks respondents about their home value.  If we assume that the sale price is equivalent to the home value, how well does the study data represent the housing stock?  The following figure takes the study data and compares with with Monroe County's housing stock [@acs].

##### Figure 1. Percent of Homes by Price Groups and Data Set
```{r sanity_check, echo=FALSE}
library(ggplot2)
library(ggthemes)
library(tidyr)

get_acs_value <- function(price){
  if(price < 50000){
    return("1. Less than $50,000")
  } else if(price < 100000){
    return("2. $50,000 to $99,999")
  }else if(price < 150000){
    return("3. $100,000 to $149,999")
  }else if(price < 200000){
    return("4. $150,000 to $199,999")
  }else if(price < 300000){
    return("5. $200,000 to $299,999")
  }else if(price < 500000){
    return("6. $300,000 to $499,999")
  }else if(price < 1000000){
    return("7. $500,000 to $999,999")
  }else{
    return("8. $1,000,000 or more")
  }
}

check <- df %>%
  rowwise() %>%
  mutate(acs_value = get_acs_value(Price)) %>%
  ungroup() %>%
  group_by(acs_value) %>%
  summarise(study = n()) %>%
  arrange(acs_value)

# Monroe County Home Value Table C25075
check$acs <- c(9747, 32746, 54988, 38988, 34563, 14926, 3247, 924)

# Compute the shares
check <- check %>%
  mutate(acs_total = sum(acs)) %>%
  mutate(study_total = sum(study)) %>%
  mutate(acs_share = acs / acs_total) %>%
  mutate(study_share = study / study_total) %>%
  select(acs_value, acs_share, study_share) %>%
  rename(ACS = acs_share) %>%
  rename(Study = study_share) %>%
  gather(Dataset, Share, -acs_value) %>%
  rename(Value = acs_value)

ggplot(check, aes(x = reorder(Value, desc(Value)), Share, color=Dataset, fill=Dataset)) +
  geom_bar(position = "dodge", stat = "identity") +
  coord_flip() + 
  scale_color_fivethirtyeight() +
  scale_fill_fivethirtyeight() + 
  theme_fivethirtyeight(base_size = 12, base_family = "sans") 
```

The study data under represents the two highest and two lowest categories.  It also overrepresents the $100,000 to $149,999 price range.  This could indicate that different segments of the housing stock "moved" faster/more than other segements.

This data above raise some concern.  The data appears to be skewed.  The following density plot illustrates the skew.  The x-axis is stopped at half a million even though the largest sale price in the data set is `r round(max(df$Price) / 1000000, 1)` million dollars.

##### Figure 2: Sale Price Distribution
```{r price_hist, echo=FALSE, warning=FALSE}
ggplot(df, aes(Price)) +
  geom_density(fill = c("red"), color = c("red")) +
  scale_x_continuous(limits=c(0, 500000)) + 
  scale_fill_fivethirtyeight() +
  theme_fivethirtyeight(base_size = 12, base_family = "sans") 
```

##### Table 1: Preliminary Price Summary
```{r price_table, echo=FALSE}
library(kableExtra)

summary(df$Price) %>%
  as.array() %>%
  as.data.frame() %>%
  rename(Metric = Var1) %>%
  rename(Value = Freq) %>%
  kable() %>%
  kable_styling()
```

There are extreme housing prices in the data set.  The data is also not normal.  In order to correct this the price will be on a log scale and the outliers will be removed.

```{r log_transform}
# Log transformation
df <- df %>%
  mutate(Log_Price = log(Price))

# Remove the outliers
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

valid_prices <- unique(remove_outliers(df$Log_Price))
min_price <- min(valid_prices, na.rm=TRUE)
max_price <- max(valid_prices, na.rm=TRUE)

df <- df %>%
  filter(Log_Price > min_price) %>%
  filter(Log_Price < max_price)
```

Now that the outliers have been removed we have `r prettyNum(nrow(df), big.mark = ",", scientific = FALSE)` observations in our data set.  Here is what the log of the prices looks like:

##### Figure 3. Log Price Distribution
```{r log_price_hist, echo=FALSE, warning=FALSE}
ggplot(df, aes(Log_Price)) +
  geom_histogram(bins=30, fill = c("red"), color = c("red")) +
  scale_fill_fivethirtyeight() +
  theme_fivethirtyeight(base_size = 12, base_family = "sans") 
```

### Study Data

As previously stated there are `r prettyNum(nrow(df), big.mark = ",", scientific = FALSE)` observations in this data set.  Each case represents the most recent sale price for a residential property occuring in 2017, ranging in price from `r paste0("$", prettyNum(min(df$Price), big.mark = ",", scientific = FALSE))` to  `r paste0("$", prettyNum(max(df$Price), big.mark = ",", scientific = FALSE))`.  On a log scale these prices are ranging in price from `r paste0("$", prettyNum(min(df$Log_Price), big.mark = ",", scientific = FALSE))` to  `r paste0("$", prettyNum(max(df$Log_Price), big.mark = ",", scientific = FALSE))`.

## 3. Exploratory Data Analysis

As mentioned in the introduction, the median sale price for a home in the Rochester Metro is $146,000 [@nahb_hoi].  The median sale price in the study dataset is `r paste0("$", prettyNum(median(df$Price), big.mark = ",", scientific = FALSE))` which is pretty comparable.

This study posits the size and era a home was built are reliable predictors of the price (on a log scale).  The following scatterplot illustrates the relationship observed in the data.

##### Figure 4: Log Sales Price and Size by Era
```{r fig.height=10, echo=FALSE}
ggplot(df, aes(Log_Price, Size, color = Era)) + 
  geom_point() + 
  scale_color_fivethirtyeight() +
  theme_fivethirtyeight(base_size = 12, base_family = "sans") +
  facet_grid(Era ~ .)
```

It is clear to see that price and size are related.  The relationship between price and era is not as apparent.  The following visualization and summary table examines this relationship:

##### Figure 5: Log Sales Price by Era
```{r era_by_price_boxplot, echo=FALSE}
ggplot(df, aes(x=Era, y=Log_Price, color = Era)) + 
  geom_boxplot() + 
  theme_fivethirtyeight(base_size = 12, base_family = "sans") +
  theme(legend.position = "None")
```

##### Table 2: Log Sales Price by Era Summary
```{r price_by_era_table, echo=FALSE}
df %>%
  group_by(Era) %>%
  summarise(min = min(Log_Price), mean = mean(Log_Price), median = median(Log_Price), max = max(Log_Price)) %>%
  kable() %>%
  kable_styling()
```

The older houses tend to fetch less in the market than the newer houses.  There doesn't seem to be much difference between the pre and post-war housing.

### Checking for Multicollinearity

The two explanatory variables might be related.  It is reasonable to believe that home increased in size over time.  The plot below will check to see if the size and era are related.

##### Figure 6: Size by Era
```{r size_by_era_boxplot, echo=FALSE}
ggplot(df, aes(x=Era, y=Size, color = Era)) + 
  geom_boxplot() + 
  theme_fivethirtyeight(base_size = 12, base_family = "sans") +
  theme(legend.position = "None")
```

##### Table 3: Size by Era Summary
```{r size_by_era_table, echo=FALSE}
df %>%
  group_by(Era) %>%
  summarise(min = min(Size), mean = mean(Size), median = median(Size), max = max(Size)) %>%
  kable() %>%
  kable_styling()
```

The newer houses are generally larger than the older houses.  This trend does not seem as pronounced as I expected it would.

## 4. Inference

In order to evaluate how well the model generalizes the data will be split into training/test sets using an 80/20 split.

```{r train_test_split, message=FALSE}
library(caret)
set.seed(12345)
in_training <- createDataPartition(df$Log_Price, p = 0.8, list = FALSE)
training <- df[ in_training, ]
testing  <- df[-in_training, ]
```

There are `r prettyNum(nrow(training), big.mark = ",", scientific = FALSE)` observations in the training data set.  With the data split, the OLS line can be fit to the data.

```{r ols}
fit <- lm(Log_Price ~ Size + Era , training)
```

This is the summary of the model:

```{r summarize_model, comment=NA}
summary(fit)
```

This model's coefficients are a little hard to interpret given that they are on a log scale.  However one observes that all of the coefficients are statistically significant.  The square footage of living space (size) is positively correlated with the home price.  Also the later the home is built is positively correlated with the price.  60% of the variability in the log of the sales price is explained by these two variables.

```{r, echo=FALSE, eval=FALSE}
exp(coef(fit))
```

### Model Examination

##### Figure X: Residuals Scatterplot
```{r residual_plot, echo=FALSE}
ggplot(fit, aes(x = c(1:length(resid(fit))), y = resid(fit))) + 
  geom_point(color = "red") + 
  theme_fivethirtyeight(base_size = 12, base_family = "sans")
```

ADD TEXT ABOUT THE PLOT

```{r qqplot_residuals}
ggplot(fit, aes(sample=resid(fit)))+
  stat_qq(color="red") +
  geom_qq_line() + 
  theme_fivethirtyeight(base_size = 12, base_family = "sans") +
  theme(legend.position = "None")

```

The Q to Q plot above indicates that the residuals are more linear near the middle of the distribution but there are some major problems in the tails.

### Generalization

In order to assess how well this model generalizes, predictions for the testing data set will be made and compared to the actual.

```{r test_predictions}
testing$Log_Price_hat <- predict(fit, testing)
testing$Price_hat <- exp(testing$Log_Price_hat)
```

Let's visualize the predictions vs the actual sales prices.  In the following plot the sales price is on the x-axis and the prediction is on the y-axis.  Values are per one thousand.  The black dashed line represents where the prediction and the actual values are the same.

##### Figure X: Test Set Actual vs Model Prediction (in thousands)
```{r prediction_plot, echo=FALSE, message=FALSE, warning=FALSE}
over_under_equal <- function(price, price_hat){
  if(price == price_hat){"Equal"}
  else if(price > price_hat){"Under"}
  else {"Over"}
}
testing <- testing %>%
  rowwise() %>%
  mutate(Accuracy = over_under_equal(Price, Price_hat)) %>%
  ungroup

ggplot(testing, aes(Price / 1000, Price_hat / 1000, color = Accuracy)) + 
  geom_point() + 
  scale_color_fivethirtyeight() +
  xlim(0, max(testing$Price) / 1000) +
  ylim(0, max(testing$Price) / 1000) +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = 2) +
  theme_fivethirtyeight(base_size = 12, base_family = "sans") +
  guides(fill = FALSE, color = FALSE)
```

There are `r nrow(testing[testing$Accuracy=="Over", ])` cases where the model over estimates and `r nrow(testing[testing$Accuracy=="Under", ])` where it under estimates the home price. Let's examine to what degree it over/under estimates.

```{r test_error}
testing <- testing %>%
  mutate(Error = Price_hat - Price)
```

##### Figure X: Prediction Error Distribution
```{r prediction_error_distribution, echo=FALSE}
testing %>%
  ggplot(aes(Error / 1000)) +
  geom_histogram(bins = 50, fill="Red") + 
  scale_fill_fivethirtyeight() + 
  theme_fivethirtyeight(base_size = 12, base_family = "sans") +
  guides(fill = FALSE, color = FALSE)
```

##### Table 4?: Prediction Error Summary
```{r error_table, echo=FALSE}
summary(testing$Error) %>%
  as.array() %>%
  as.data.frame() %>%
  rename(Metric = Var1) %>%
  rename(Value = Freq) %>%
  kable() %>%
  kable_styling()
```

There is a wide variety in the errors.  It appears that the model tends to be really good or really bad.

```{r echo=F, eval=F}
If your data fails some conditions and you can’t use a theoretical method, then you should use simulation. If you can use both methods, then you should use both methods. It is your responsibility to figure out the appropriate methodology.

Check conditions

Theoretical inference (if possible) - hypothesis test and confidence interval

Simulation based inference - hypothesis test and confidence interval

Brief description of methodology that reflects your conceptual understanding
```

## 5. Conclusion

This study set out to see if the size of a home and the era in which it was built could be used to get an estimate of the sale price.  While these variables do seem to be relavant in predicting the home value, it appears that it is more complicated than just that.  Future research could include extending the model to capture a location variable, or employ a different modeling approach as  it appears that the relationship may be non-linear.

## 6. References

### 2018-12-07 DATA606/Silva-Project.Rmd
Additional Work

@@
 The Rochester, New York metro is acclaimed as a place with affordable housing.
-Recent data show 81% of the housing in the Rochester Metro is affordable [@nahb_hoi].
+Recent data shows 81% of home in the Rochester Metro are affordable [@nahb_hoi].
 The median sales price for a home in Rochester is $146,000.
@@
 There are also suburban homes that are larger and newer.
-Can the square footage and age of a home reliabily estimate the sale price?
+Can the square footage and age of a home estimate the sale price reliably?
 
@@
 
-The NYS Office of Real Property Services (NYS RPS) collects sales price data from adminstrative records.
+The NYS Office of Real Property Services (NYS RPS) collects sales price data from administrative records.
 The data for this study were scrapped using a [web scrapper](https://github.com/mikeasilva/CUNY-SPS/blob/21abb87c5a28204e75ba2944c574889e07289345/data/Build%20Monroe%20Real%20Property%20Data.py) in early November of 2018 [@mcrpp].
@@
 
-This study is purely observational in nature.  It does not assert that these variables have a causal relationship nor should it be interpreted that way.
-
-Scope of inference - generalizability: Identify the population of interest, and whether the findings from this analysis can be generalized to that population, or, if not, a subsection of that population.
-Explain why or why not.
-Also discuss any potential sources of bias that might prevent generalizability.
+This study is purely observational in nature.  It does not assert these variables have a causal relationship.
 
@@
 There are `r prettyNum(nrow(df), big.mark = ",", scientific = FALSE)` observations in the data set.
-This indicates the merging process did not introduce any dupplication.
+This indicates the merging process did not introduce any duplication.
 Now that the housing attributes are added in, some odd observations need to be filtered out.
@@
 This is the final count of cases for the study.
-The last data preperation step is to change the year built quantitative variable into a categorical variable and only keep the variables of interest.
+The next step is to create the categorical variable from the construction year.
 
-```{r finalize_df}
+```{r era}
 get_era <- function(year){
@@
   ungroup() %>%
-  mutate(Era = as.factor(Era)) %>%
+  mutate(Era = as.factor(Era))
+```
+
+The last data preparation step is renaming and selecting the variables of interest.
+
+```{r finalize_df}
+df <- df %>%
   rename(Size = SQFT_LIV) %>%
@@
 The American Community survey asks respondents about their home value.
-If we assume that the sale price is equivalent to the home value, how well does the study data represent the housing stock?
-The following figure takes the study data and compares with with Monroe County's housing stock [@acs].
+Assuming the sale price is comparable to the home value, how well does the study data represent the housing stock?
+The following figure takes the study data and compares with with Monroe County’s housing stock [@acs].
 
@@
 The study data under represents the two highest and two lowest categories.
-It also overrepresents the $100,000 to $149,999 price range.
-This could indicate that different segments of the housing stock "moved" faster/more than other segements.
+It also over represents the $100,000 to $149,999 price range.
+This could indicate different segments of the housing stock “moved” faster/more than other segments.
 
 This data above raise some concern.
-The data appears to be skewed.
+The data appears skewed.
 The following density plot illustrates the skew.
@@
 
-There are extreme housing prices in the data set.  The data is also not normal.  In order to correct this the price will be on a log scale and the outliers will be removed.
+There are extreme housing prices in the data set.  The data is also not normal.  To correct this, the price will be on a log scale and outliers will be removed.
 
@@
 
-This study posits the size and era a home was built are reliable predictors of the price (on a log scale).  The following scatterplot illustrates the relationship observed in the data.
+This study posits the size and era a home was built are reliable predictors of the price (on a log scale).  The following scatter plot illustrates the relationship observed in the data.
 
@@
 
-It is clear to see that price and size are related.  The relationship between price and era is not as apparent.  The following visualization and summary table examines this relationship:
+It is clear to see that price and size are related.  The relationship between price and era is not as clear  The following visualization and summary table examines this relationship:
 
@@
 
-The two explanatory variables might be related.  It is reasonable to believe that home increased in size over time.  The plot below will check to see if the size and era are related.
+The two explanatory variables might be related.  It is reasonable to believe that home increased in size over time.  The plot below will display how the size and era are related.
 
@@
 This model's coefficients are a little hard to interpret given that they are on a log scale.
-However one observes that all of the coefficients are statistically significant.
+Yet one observes that all the coefficients are statistically significant.
 The square footage of living space (size) is positively correlated with the home price.
@@
 
-### Model Examination
+### Residual Examination
 
-##### Figure X: Residuals Scatterplot
+##### Figure 7: Residuals Scatterplot
 ```{r residual_plot, echo=FALSE}
@@
 
-ADD TEXT ABOUT THE PLOT
+The above figure is the residuals ordered by their index.  There is no clear pattern in this figure.
 
-```{r qqplot_residuals}
+##### Figure 7: Residuals Q to Q
+```{r qqplot_residuals,echo=FALSE}
 ggplot(fit, aes(sample=resid(fit)))+
@@
 
-The Q to Q plot above indicates that the residuals are more linear near the middle of the distribution but there are some major problems in the tails.
+The Q to Q plot above indicates that the residuals are more linear near the middle of the distribution.  But there are some major problems in the tails.
 
@@
 ```{r test_predictions}
-testing$Log_Price_hat <- predict(fit, testing)
+predictions <- predict(fit, testing, interval = "confidence")
+testing$Log_Price_hat <- predictions[, 1]
 testing$Price_hat <- exp(testing$Log_Price_hat)
+testing$Log_Price_hat_lower <- predictions[, 2]
+testing$Price_hat_lower <- exp(testing$Log_Price_hat_lower)
+testing$Log_Price_hat_upper <- predictions[, 3]
+testing$Price_hat_upper <- exp(testing$Log_Price_hat_upper)
 ```
@@
 The black dashed line represents where the prediction and the actual values are the same.
+The blue are under estimates, red are over estimates, and the green points have a price that falls within the confidence interval.
 
-##### Figure X: Test Set Actual vs Model Prediction (in thousands)
+##### Figure 8: Test Set Actual vs Model Prediction (in thousands)
 ```{r prediction_plot, echo=FALSE, message=FALSE, warning=FALSE}
-over_under_equal <- function(price, price_hat){
-  if(price == price_hat){"Equal"}
-  else if(price > price_hat){"Under"}
-  else {"Over"}
+over_under_equal <- function(price, lower, upper){
+  if(price > upper){"Over Estimate"}
+  else if(price < lower){"Under Estimate"}
+  else {"Within CI"}
 }
@@
   rowwise() %>%
-  mutate(Accuracy = over_under_equal(Price, Price_hat)) %>%
+  mutate(Accuracy = over_under_equal(Price, Price_hat_lower, Price_hat_upper)) %>%
   ungroup
 
-ggplot(testing, aes(Price / 1000, Price_hat / 1000, color = Accuracy)) + 
+ggplot(testing, aes(Price_hat / 1000, Price / 1000, color = Accuracy)) + 
   geom_point() + 
@@
   geom_abline(intercept = 0, slope = 1, color = "black", linetype = 2) +
-  theme_fivethirtyeight(base_size = 12, base_family = "sans") +
-  guides(fill = FALSE, color = FALSE)
+  theme_fivethirtyeight(base_size = 12, base_family = "sans") 
 ```
 
-There are `r nrow(testing[testing$Accuracy=="Over", ])` cases where the model over estimates and `r nrow(testing[testing$Accuracy=="Under", ])` where it under estimates the home price.
+There are `r nrow(testing[testing$Accuracy=="Over Estimate", ])` cases where the model over estimates and `r nrow(testing[testing$Accuracy=="Under Estimate", ])` where it under estimates the home price.
+`r nrow(testing[testing$Accuracy=="Within CI", ])` are in within the confidence interval.
 Let's examine to what degree it over/under estimates.
@@
 
-##### Figure X: Prediction Error Distribution
+##### Figure 9: Prediction Error Distribution
 ```{r prediction_error_distribution, echo=FALSE}
@@
 
-##### Table 4?: Prediction Error Summary
+##### Table 4: Prediction Error Summary
 ```{r error_table, echo=FALSE}
@@
 
-There is a wide variety in the errors.  It appears that the model tends to be really good or really bad.
+There is a wide variety in the errors.  Let's look at the distribution of the error in absolute terms as a share of the total price.  It would be nice to see errors in the 5% range
 
-```{r echo=F, eval=F}
-If your data fails some conditions and you can’t use a theoretical method, then you should use simulation.
-If you can use both methods, then you should use both methods.
-It is your responsibility to figure out the appropriate methodology.
+```{r error_share}
+testing <- testing %>%
+  mutate(Error_Share = (abs(Error) / Price) * 100)
+```
 
-Check conditions
+##### Figure 10. Cumulative Distribution of Error Share (5% cutoff in red)
+```{r error_share_plot, echo=FALSE}
+testing %>%
+  ggplot(aes(Error_Share)) + 
+  stat_ecdf() + 
+  geom_vline(xintercept = 5, color = "Red", linetype = "dashed") +
+  scale_fill_fivethirtyeight() + 
+  theme_fivethirtyeight(base_size = 12, base_family = "sans") +
+  guides(fill = FALSE, color = FALSE)
+```
 
-Theoretical inference (if possible) - hypothesis test and confidence interval
-
-Simulation based inference - hypothesis test and confidence interval
-
-Brief description of methodology that reflects your conceptual understanding
-```
+Only `r round((nrow(testing[testing$Error_Share < 5, ]) / nrow(testing)) * 100, 0)`% of the testing set observations have an error which is +/-5% of the estimate.
 
@@
 This study set out to see if the size of a home and the era in which it was built could be used to get an estimate of the sale price.
-While these variables do seem to be relavant in predicting the home value, it appears that it is more complicated than just that.
-Future research could include extending the model to capture a location variable, or employ a different modeling approach as  it appears that the relationship may be non-linear.
-
+While these variables seem to be relevant, it appears that it is more complicated.
+Future research could expand the model.
+A location variable would be a next step.
+A different modeling approach may be employed as the relationship appears to be non-linear.
 ## 6. References

### 2018-12-07 DATA606/Silva-Project.Rmd
Refinement of Final Project

@@
 Finding affordable housing is a real challenge throughout the country.
-The Rochester, New York metro is acclaimed as a place with affordable housing.
-Recent data shows 81% of home in the Rochester Metro are affordable [@nahb_hoi].
-The median sales price for a home in Rochester is $146,000.
-This is about half as much as the national average ($268,000).
-With the high family income, housing proves to be affordable.
+The Rochester, New York metro is acclaimed fit its affordable housing.
+Recent data shows 81% of homes in the Rochester Metro are affordable [@nahb_hoi].
+The median sales price in Rochester is $146,000, or half of the national average ($268,000).
+With the relatively high family income, Rochester housing proves to be affordable.
 
@@
 
-The NYS Office of Real Property Services (NYS RPS) collects sales price data from administrative records.
-The data for this study were scrapped using a [web scrapper](https://github.com/mikeasilva/CUNY-SPS/blob/21abb87c5a28204e75ba2944c574889e07289345/data/Build%20Monroe%20Real%20Property%20Data.py) in early November of 2018 [@mcrpp].
-At that time the data was synced with NYS RPS on 22 October 2018.
+The NYS Office of Real Property Services (NYS RPS) collects sales price data.
+This administrative record data was scrapped using a [web scrapper](https://github.com/mikeasilva/CUNY-SPS/blob/21abb87c5a28204e75ba2944c574889e07289345/data/Build%20Monroe%20Real%20Property%20Data.py) in early November of 2018 [@mcrpp].
+At that time, the data was as of 22 October 2018.
 
-The NYS GIS clearinghouse compiled data on housing attributes [@nysgis].
-The 2017 data is used in this study.
+The NYS GIS clearinghouse compiles data on housing attributes [@nysgis].
+This data originates from local assessors.
+The 2017 data (latest available) was used.
 The square footage of living space is the measurement of size in this study.
@@
 Data on the age of the home are transformed into eras (a categorical variable).
-These eras come from an [article](https://www.forbes.com/sites/trulia/2013/05/02/american-homes-by-decade/#16c559821045) from the Trulia's Chief Economist.
+These eras come from an [article](https://www.forbes.com/sites/trulia/2013/05/02/american-homes-by-decade/#16c559821045) authored by Trulia's Chief Economist.
 They are:
@@
 
-The scrapped data is in a SQLite database stored in GitHub.  The first step is to download and extract this database.
+The scrapped data is in a SQLite database stored on GitHub.  The first step is to download and extract this database locally.
 
@@
 This is the final count of cases for the study.
-The next step is to create the categorical variable from the construction year.
+The next step is to translate the year of construction into an era.
 
@@
 
-The study data under represents the two highest and two lowest categories.
-It also over represents the $100,000 to $149,999 price range.
-This could indicate different segments of the housing stock “moved” faster/more than other segments.
+The study data is generally fairly representative.
+It under represent the two highest and two lowest categories.
+It also over represent home in the $100,000 to $149,999 price range.
+This could be due to different segments of the housing stock “moving” faster/more than other segments.
 
-This data above raise some concern.
-The data appears skewed.
+This data distribution depicted above raises some concern.
+The data appear skewed.
 The following density plot illustrates the skew.
-The x-axis is stopped at half a million even though the largest sale price in the data set is `r round(max(df$Price) / 1000000, 1)` million dollars.
+The x-axis is stopped at half a million for readability, even though the largest sale price in the data set is `r round(max(df$Price) / 1000000, 1)` million dollars.
 
@@
 
-There are extreme housing prices in the data set.  The data is also not normal.  To correct this, the price will be on a log scale and outliers will be removed.
+This confirms there are extreme housing prices in the data set.  The data is also not normal.  To correct this, the price will be put on a log scale and outliers will be removed.
 
@@
 Each case represents the most recent sale price for a residential property occuring in 2017, ranging in price from `r paste0("$", prettyNum(min(df$Price), big.mark = ",", scientific = FALSE))` to  `r paste0("$", prettyNum(max(df$Price), big.mark = ",", scientific = FALSE))`.
-On a log scale these prices are ranging in price from `r paste0("$", prettyNum(min(df$Log_Price), big.mark = ",", scientific = FALSE))` to  `r paste0("$", prettyNum(max(df$Log_Price), big.mark = ",", scientific = FALSE))`.
+On a log scale these prices range from `r paste0("$", prettyNum(min(df$Log_Price), big.mark = ",", scientific = FALSE))` to  `r paste0("$", prettyNum(max(df$Log_Price), big.mark = ",", scientific = FALSE))`.
 
@@
 As mentioned in the introduction, the median sale price for a home in the Rochester Metro is $146,000 [@nahb_hoi].
-The median sale price in the study dataset is `r paste0("$", prettyNum(median(df$Price), big.mark = ",", scientific = FALSE))` which is pretty comparable.
+The median sale price in the study dataset is `r paste0("$", prettyNum(median(df$Price), big.mark = ",", scientific = FALSE))` which is comparable.
 
@@
 
-It is clear to see that price and size are related.  The relationship between price and era is not as clear  The following visualization and summary table examines this relationship:
+It is clear that price and size are related.  The relationship between price and era is not as clear  The following visualization and summary table examines this relationship further:
 
@@
 
-The older houses tend to fetch less in the market than the newer houses.  There doesn't seem to be much difference between the pre and post-war housing.
+The older houses tend to fetch less in the market than the newer houses.  There also does not seem to be much difference between the pre and post-war housing.
 
@@
 
-The two explanatory variables might be related.  It is reasonable to believe that home increased in size over time.  The plot below will display how the size and era are related.
+The two explanatory variables might be related.  It is reasonable to believe that homes increased in size over time.  The plot below will display how the size and era are related.
 
@@
 Also the later the home is built is positively correlated with the price.
-60% of the variability in the log of the sales price is explained by these two variables.
+Only 60% of the variability in the log of the sales price is explained by these two variables.
 
@@
 The black dashed line represents where the prediction and the actual values are the same.
-The blue are under estimates, red are over estimates, and the green points have a price that falls within the confidence interval.
+The blue are under estimates, red are over estimates, and the green points have a price that falls within the prediction's confidence interval.
 
@@
 
-There are `r nrow(testing[testing$Accuracy=="Over Estimate", ])` cases where the model over estimates and `r nrow(testing[testing$Accuracy=="Under Estimate", ])` where it under estimates the home price.
+There are `r nrow(testing[testing$Accuracy=="Over Estimate", ])` cases where the model over estimates the home price, and `r nrow(testing[testing$Accuracy=="Under Estimate", ])` where it under estimates the home price.
 `r nrow(testing[testing$Accuracy=="Within CI", ])` are in within the confidence interval.
 Let's examine to what degree it over/under estimates.
+The model does not seem to accurately depict reality.
 
@@
 
-There is a wide variety in the errors.  Let's look at the distribution of the error in absolute terms as a share of the total price.  It would be nice to see errors in the 5% range
+There is a wide variety in the errors.  Let's look at the distribution of the error in absolute terms expressed as a share of the total price.  It would be nice to see errors in the 5% range
 
@@
 Only `r round((nrow(testing[testing$Error_Share < 5, ]) / nrow(testing)) * 100, 0)`% of the testing set observations have an error which is +/-5% of the estimate.
+Once again the model seems to be a poor predictor.
 
@@
 This study set out to see if the size of a home and the era in which it was built could be used to get an estimate of the sale price.
-While these variables seem to be relevant, it appears that it is more complicated.
+While these variables seem to be relevant, predictions made only using the variables are largely inaccurate.
+It appears that the relationship is more complicated.
 Future research could expand the model.
-A location variable would be a next step.
-A different modeling approach may be employed as the relationship appears to be non-linear.
+A location variable would be a next step along with the size of the plot.
+A different modeling approach should be employed as the relationship appears to be non-linear.
+
 ## 6. References

### 2019-01-30 DATA605/MSilva_Assign1.Rmd
Initial commit

---
title: "DATA 605 Assignment 1"
output:
  pdf_document: default
  html_document: default
---

## Problem Set 1

You can think of vectors representing many dimensions of related information. For
instance, Netflix might store all the ratings a user gives to movies in a vector. This is clearly a vector of very large dimensions (in the millions) and very sparse as the user might have rated only a few movies. Similarly, Amazon might store the items purchased by a user in a vector, with each slot or dimension representing a unique product and the value of the slot, the number of such items the user bought. One task that is frequently done in these settings is to find similarities between users. And, we can use dot-product between vectors to do just that. As you know, the dot-product is proportional to the length of two vectors and to the angle between them. In fact, the dot-product between two vectors, normalized by their lengths is called as the cosine distance and is frequently used in recommendation engines.

(1) Calculate the dot product u.v where u = [0.5; 0.5] and v = [3; −4]

$u \cdot v= u_1 \times v_1 + u_2 \times v_2 = 0.5 \times 3 + 0.5 \times 4 = 1.5 + 2 = 3.5$

```{r, warning=FALSE}
u <- c(0.5, 0.5)
v <- c(3, 4)

dot_product_u_times_v <- u %*% v
dot_product_u_times_v
```

(2) What are the lengths of u and v? Please note that the mathematical notion of the length of a vector is not the same as a computer science definition.

$||u|| = \sqrt{(u_1)^2 + (u_2)^2} = \sqrt{0.5^2 + 0.5^2} = \sqrt{0.25 + 0.25} = \sqrt{0.5} = 0.7071068$

```{r}
magnitude_of_u <- sqrt(u[1] ^ 2 + u[2] ^ 2)
magnitude_of_u
```

$||v|| = \sqrt{(v_1)^2 + (v_2)^2} = \sqrt{3^2 + 4^2} = \sqrt{9 + 16} = \sqrt{25} = 5$
```{r}
magnitude_of_v <- sqrt(v[1] ^ 2 + v[2] ^ 2)
magnitude_of_v
```

(3) What is the linear combination: 3u − 2v?

$3u - 2v = [1.5, 1.5] - [6, 8] = [-4.5, -6.5]$

```{r}
3 * u - 2 * v
```

(4) What is the angle between u and v

$\text{cos } \theta = \frac{u \cdot v}{||u||\times||v||} = \frac{3.5}{0.7071068 \times 5} = \frac{3.5}{3.535534} = 0.9899495$

```{r}
cosine <- dot_product_u_times_v / (magnitude_of_u * magnitude_of_v)
cosine
```

## Problem Set 2

Set up a system of equations with 3 variables and 3 constraints and solve for x. Please write a function in R that will take two variables (matrix A & constraint vector b) and solve using elimination. Your function should produce the right answer for the system of equations for any 3-variable, 3-equation system. You don’t have to worry about degenerate cases and can safely assume that the function will only be tested with a system of equations that has a solution. Please note that you do have to worry about zero pivots, though. Please note that you should not use the built-in function solve to solve this system or use matrix inverses. The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution. Alternatively, you can augment the matrix A with vector b and jointly apply the Gauss Jordan elimination procedure.

Please test it with the system below and it should produce a solution x = [−1.55, −0.32, 0.95]

 \[ \begin{bmatrix}
    1 & 1 & 3 \\
    2 & -1 & 5 \\
    -1 & -2 & 4
  \end{bmatrix}
  \begin{bmatrix}
    x_1 \\
    x_2 \\
    x_3 
  \end{bmatrix} = 
  \begin{bmatrix}
    1 \\
    2 \\
    6 
  \end{bmatrix}
\]

```{r, echo=FALSE, eval=FALSE}
my_solve <- function(A, b){
  # Find the largest coefficient of the first column (x)
  i <- which.max(A[, 1])
  # Swap the first row with the largest row
  if (i != 1){
    temp <- A[i, ]
    bi <- b[i]
    A[i, ] <- A[1, ]
    A[1, ] <- temp
    b[i] <- b[1]
    b[1] <- bi
  }
  # Divide the first row by A[1,1] so we have a one in the first pivot position
  b[1] <- b[1] / A[1, 1]
  A[1, ] <- A[1, ] / A[1, 1]
  # Multiply the first row with A[2, 1] and subtract it from the second row to eliminate A[2, 1]
  b[2] <- b[2] - (b[1] * A[2, 1])
  A[2, ] <- A[2, ] - (A[1, ] * A[2, 1])
  # Eliminate A[3,1] using the same method
  b[3] <- b[3] - (b[1] * A[3, 1])
  A[3, ] <- A[3, ] - (A[1, ] * A[3, 1])
  # Check to see if A[2,2] is not zero
  if (A[2, 2] == 0){
    # Swap the second and the third row
    temp <- A[3, ]
    b3 <- b[3]
    A[3, ] <- A[2, ]
    A[2, ] <- temp
    b[3] <- b[2]
    b[2] <- b3
  }
  # Divide the second row by A[2, 2] so we have a one in the second pivot position
  b[2] <- b[2] / A[2, 2]
  A[2, ] <- A[2, ] / A[2, 2]
  # Eliminate A[3,2] using the second row
  b[3] <- b[3] - (b[2] * A[3, 2])
  A[3, ] <- A[3, ] - (A[2, ] * A[3, 2])
  # Get a one in the third pivot position
  b[3] <- b[3] / A[3, 3]
  A[3, ] <- A[3, ] / A[3, 3]
  # I have an upper triangular matrix.  Now for the back substitution.
  x3 <- b[3]
  x2 <- b[2] - (A[2, 3] * x3)
  x1 <- b[1] - (A[1, 3] * x3) - (A[1, 2] * x2)
  # Return the results
  return(c(x1, x2, x3))
}

A <- matrix(c(1, 2, -1, 1, -1, -2, 3, 5, 4), 3, 3)
b <- c(1, 2, 6)
my_solve(A, b)
```

```{r}
my_solve <- function(A, b){
  n <- nrow(A)
  A <- cbind(A,  b)

  for (pivot_position in c(1:n)){
    row_with_largest_coefficient <- pivot_position - 1 + which.max(A[c(pivot_position:n), pivot_position])
    # Ensure the pivot is the largest value of the remaining rows
    # This protects against having a zero in the pivot position
    if (pivot_position != row_with_largest_coefficient){
      temp <- A[pivot_position, ]
      A[pivot_position, ] <- A[row_with_largest_coefficient, ]
      A[row_with_largest_coefficient, ] <- temp
    }
    # Get a 1 in the pivot position by dividing the row by the element in the pivot position
    A[pivot_position, ] <- A[pivot_position, ] / A[pivot_position, pivot_position]
    # Eliminate all other rows
    for (i in c(pivot_position:n)){
      if (i != pivot_position){
        # Subtract the scaled pivot position row from the row we want to eliminate
        A[i, ] <- A[i, ] - (A[pivot_position, ] * A[i, pivot_position])
      }
    }
  }
  # Now solve using back substituiton
  for (i in c( (n - 1):1)){
    for (j in c(n:2)){
      if (j > i){
        A[i, "b"] <- A[i, "b"] - (A[i, j] * A[j, "b"])
      }
    }
  }
  return(A[, "b"])
}

A <- matrix(c(1, 2, -1, 1, -1, -2, 3, 5, 4), 3, 3)
b <- c(1, 2, 6)

my_solve(A, b)
```

#### Further Verification

I want to run a couple more examples through the function to ensure it is working properly.  The first is taken from page 29 of the text book.  It should return [-3, 5, 2]
```{r}
A <- matrix(c(-7, 5, 1, -6, 5, 0, -12, 7, 4), 3, 3)
b <- c(-33, 24, 5)

my_solve(A, b)
```

### 2019-02-06 DATA605/MSilva_Assign2.Rmd
Work on Assignment 2

---
title: "DATA 605 Assignment 2"
author: "by Michael Silva"
output: html_document
---

## Problem Set 1

**(1) Show that $A^TA \neq AA^T$ in general.**

\[ A=\begin{bmatrix}
  a & b \\
  c & d
  \end{bmatrix}
\]

\[ A^T=\begin{bmatrix}
  a & c \\
  b & d
  \end{bmatrix}
\]

\[ A^TA=
  \begin{bmatrix}
  a & c \\
  b & d
  \end{bmatrix} \times
  \begin{bmatrix}
  a & b \\
  c & d
  \end{bmatrix} = \begin{bmatrix}
  a \times a + c \times c & c \times a + c \times c \\
  b \times b + d \times d & b \times d + d \times d
  \end{bmatrix} = \begin{bmatrix}
  a^2 + c^2 & ac + c^2 \\
  b^2 + d^2 & bd + d^2
  \end{bmatrix} 
\]

\[ AA^T=\begin{bmatrix}
  a & b \\
  c & d
  \end{bmatrix} \times
  \begin{bmatrix}
  a & c \\
  b & d
  \end{bmatrix} = \begin{bmatrix}
  a \times a + b \times b & a \times c + b \times d \\
  c \times a + d \times d & c \times c + d \times d
  \end{bmatrix} = \begin{bmatrix}
  a^2 + b^2 & ac + bd \\
  ac + d^2 & c^2 + d^2
  \end{bmatrix} 
\]

\[
\begin{bmatrix}
  a^2 + c^2 & ac + c^2 \\
  b^2 + d^2 & bd + d^2
  \end{bmatrix} 
  \neq
  \begin{bmatrix}
  a^2 + b^2 & ac + bd \\
  ac + d^2 & c^2 + d^2
  \end{bmatrix} 
\]

**(2) For a special type of square matrix A, we get $A^TA = AA^T$. Under what conditions could this be true?**

Any matrix where all elements are the same value.  For example the zero matrix:

\[ A=\begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix}
\]

\[ A^T=\begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix}
\]

\[ A^TA=
  \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} \times
  \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} = \begin{bmatrix}
  0 \times 0 + 0 \times 0 & 0 \times 0 + 0 \times 0 \\
  0 \times 0 + 0 \times 0 & 0 \times 0 + 0 \times 0
  \end{bmatrix} = \begin{bmatrix}
  0 + 0 & 0 + 0 \\
  0 + 0 & 0 + 0
  \end{bmatrix} 
  = \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} 
\]

\[ AA^T=
  \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} \times
  \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} = \begin{bmatrix}
  0 \times 0 + 0 \times 0 & 0 \times 0 + 0 \times 0 \\
  0 \times 0 + 0 \times 0 & 0 \times 0 + 0 \times 0
  \end{bmatrix} = \begin{bmatrix}
  0 + 0 & 0 + 0 \\
  0 + 0 & 0 + 0
  \end{bmatrix} 
  = \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} 
\]

\[
\begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} 
  =
  \begin{bmatrix}
  0 & 0 \\
  0 & 0
  \end{bmatrix} 
\]

## Problem Set 2

**Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars.**

**Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. Please submit your response in an R Markdown document using our class naming convention.**

**You don’t have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code. If you doing the entire assignment in R, then please submit only one markdown document for both the problems. **

```{r}
factorize <- function(A){
  n <- nrow(A)
  L <- diag(n)

  for (j in c(1:n)){
    for(i in c(2:n)){

    }
  }
  return(list(L=L, U=A))
}
```

I will test this on the example worked through on [this video](https://www.youtube.com/watch?v=UlWcofkUDDU&feature=youtu.be)

$ A = LU \text$ or

\[ \begin{bmatrix}
  2 & 4 & -4 \\
  1 & -4 & 3\\
  -6 & -9 & 5
  \end{bmatrix} = 
  \begin{bmatrix}
  1 & 0 & 0 \\
  \frac{1}{2} & 1 & 0\\
  -3 & -\frac{1}{2} & 1
  \end{bmatrix} \times
  \begin{bmatrix}
  2 & 4 & -4 \\
  0 & -6 & 5\\
  0 & 0 & -\frac{9}{2}
  \end{bmatrix}
\]

```
A <- matrix(c(2, 1, -6, 4, -4, -9, -4, 3, 5), 3, 3)

factors <- factorize(A)
factors$L

factors$U
```

### 2019-02-19 DATA605/MSilva_Assign4.Rmd
Initial commit

---
title: "DATA 605 Assignment 4"
author: "by Michael Silva"
output: html_document
---

### 2019-02-19 DATA605/MSilva_Assign4.Rmd
Work on assignment 4

@@
 ---
+
+## Problem Set 1
+
+**Given a 3 × 2 matrix A**
+
+\[ A=\begin{bmatrix}
+  1 & 2 & 3 \\
+  -1 & 0 & 4
+  \end{bmatrix}
+\]
+
+**write code in R to compute $X = AA^T$ and $Y = A^TA$.**
+
+```{r}
+A <- matrix(c(1, -1, 2, 0, 3, 4), 2)
+X <- A %*% t(A)
+Y <- t(A) %*% A
+```
+
+**Then, compute the eigenvalues and eigenvectors of X and Y using the built-in commans in R.**
+
+```{r}
+X_eigen <- eigen(X)
+Y_eigen <- eigen(Y)
+```
+
+**Then, compute the left-singular, singular values, and right-singular vectors of A using the svd command.**
+
+```{r}
+A_svd <- svd(A)
+```
+
+**Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and Y.**
+
+```{r}
+A_svd$u
+X_eigen$vectors
+```
+
+One might observe a difference in signs.  According to [this post](https://stat.ethz.ch/pipermail/r-help/2003-July/036504.html) on the R help forum by Roger Peng:
+
+>That is true, but using different compilers on the same machine and the 
+>same version of R may give different signs for the eigenvectors.  The 
+>moral is, don't rely on the signs of eigenvectors!  (This *is* on the help 
+>page.)
+
+```{r}
+A_svd$v
+Y_eigen$vectors[, 1:2] # The third value is basically zero
+```
+
+**In addition, the two non-zero eigenvalues (the 3rd value will be very close to zero, if not zero) of both X and Y are the same and are squares of the non-zero singular values of A.**
+
+```{r}
+X_eigen$values
+Y_eigen$values[1:2]
+A_svd$d ^ 2
+```
+
+## Problem Set 2
+
+**Using the procedure outlined in section 1 of the weekly handout, write a function to compute the inverse of a well-conditioned full-rank square matrix using co-factors.
+In order to compute the co-factors, you may use built-in commands to compute the determinant.**
+
+```{r}
+myinverse <- function(A){
+  # Initialize the cofactors matrix
+  matrix_of_cofactors <- matrix(NA, nrow = nrow(A), ncol = ncol(A))
+  for (i in c(1:nrow(A))){
+    for (j in c(1:ncol(A))){
+      # Get the determinant of the matrix except the current column and row
+      matrix_of_cofactors[i, j] <- det(A[-i, -j]) * (-1) ^ (i + j)
+    }
+  }
+  # Transpose all elements of the previous matrix of cofactors
+  adjugate <- t(matrix_of_cofactors)
+  # Divide the ajugate by the determinant of the original matrix
+  inverse <- adjugate / det(A)
+  return(inverse)
+}
+
+A <- matrix(c(3, 2, 0, 0, 0, 1, 2, -2, 1), 3)
+B <- myinverse(A)
+
+A %*% B
+```

### 2019-02-25 DATA605/MSilva_Assign5.Rmd
Initial commit

---
title: "DATA 605 Assignment 5"
author: "by Michael Silva"
output: html_document
---

Choose independently two numbers B and C at random from the interval [0, 1] with uniform density.

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
set.seed(12345)
n <- 1000000
B <- runif(n, min = 0, max = 1)
C <- runif(n, min = 0, max = 1)
data <- data.frame(B = B, C = C)
```

Prove that B and C are proper probability distributions.

```{r, echo=FALSE}
hist(data$B, main = "Histogram of B", xlab = "")
```

```{r, echo=FALSE}
hist(data$C, main = "Histogram of C", xlab = "")
```

Note that the point (B,C) is then chosen at random in the unit square.

Find the probability that

## (a) B + C < 1/2

```{r, message=FALSE}
data <- data %>%
  mutate(a = ifelse(B + C < 1/2, 1, 0))
sum(data$a) / n
```

This is represented by the area in blue

```{r echo=FALSE, cache=TRUE}
ggplot(data, aes(x = B, y = C, color = as.factor(a))) +
  geom_point() +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  theme(legend.position = "none", aspect.ratio=1)
```

## (b) BC < 1/2

```{r}
data <- data %>%
  mutate(b = ifelse(B * C < 1/2, 1, 0))
sum(data$b) / n
```

```{r echo=FALSE, cache=TRUE}
ggplot(data, aes(x = B, y = C, color = as.factor(b))) +
  geom_point() +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  theme(legend.position = "none", aspect.ratio=1)
```

## (c) |B - C| < 1/2

```{r}
data <- data %>%
  mutate(c = ifelse(abs(B - C) < 1/2, 1, 0))
sum(data$c) / n
```

```{r echo=FALSE, cache=TRUE}
ggplot(data, aes(x = B, y = C, color = as.factor(c))) +
  geom_point() +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  theme(legend.position = "none", aspect.ratio=1)
```

## (d) max(B, C) < 1/2

```{r}
data <- data %>%
  mutate(d = ifelse(pmax(B, C) < 1/2, 1, 0))
sum(data$d) / n
```

```{r echo=FALSE, cache=TRUE}
ggplot(data, aes(x = B, y = C, color = as.factor(d))) +
  geom_point() +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  theme(legend.position = "none", aspect.ratio=1)
```

## (e) min(B, C) < 1/2

```{r}
data <- data %>%
  mutate(e = ifelse(pmin(B, C) < 1/2, 1, 0))
sum(data$e) / n
```

```{r echo=FALSE, cache=TRUE}
ggplot(data, aes(x = B, y = C, color = as.factor(e))) +
  geom_point() +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  theme(legend.position = "none", aspect.ratio=1)
```

### 2019-03-07 DATA605/MSilva_Assign6.Rmd
More work

---
title: "DATA 605 Assignment 6"
author: "by Michael Silva"
output:
  html_document: default
  pdf_document: default
---

## Question 1

A box contains 54 red marbles, 9 white marbles, and 75 blue marbles. If a marble is randomly selected from the box, what is the probability that it is red or blue? Express your answer as a fraction or a decimal number rounded to four decimal places.

**0.9348**

```{r}
round((54 + 75) / (54 + 9 + 75), 4)
```

## Question 2

You are going to play mini golf. A ball machine that contains 19 green golf balls, 20 red golf balls, 24 blue golf balls, and 17 yellow golf balls, randomly gives you your ball. What is the probability that you end up with a red golf ball? Express your answer as a simplified fraction or a decimal rounded to four decimal places.

**0.25**

```{r}
round(20 / (19 + 20 + 24 + 17), 4)
```

## Question 3

A pizza delivery company classifies its customers by gender and location of residence. The research department has gathered data from a random sample of 1399 customers. The data is summarized in the table below.

#### Gender and Residence of Customers

|                           | Males  | Females |
|---------------------------|--------|---------|
| Apartment                 | 81     | 282     |
| Dorm                      | 116    | 79      |
| With Parent(s)            | 215    | 252     |
| Sorority/Fraternity House | 130    | 97      |
| Other                     | 129    | 72      |

What is the probability that a customer is not male or does not live with parents? Write your answer as a fraction or a decimal number rounded to four decimal places.

**0.8463**

```{r}
round((1399 - 215) / 1399, 4)
```

## Question 4

Determine if the following events are independent. Going to the gym. Losing weight. Answer: A) Dependent B) Independent

**Dependent**

## Question 5

A veggie wrap at City Subs is composed of 3 different vegetables and 3 different condiments wrapped up in a tortilla. If there are 8 vegetables, 7 condiments, and 3 types of tortilla available, how many different veggie wraps can be made?

**5,880**

```{r}
vegetable_combinations <- (factorial(8) / (factorial(3) * factorial(8 - 3))) 
condiment_combinations <- (factorial(7) / (factorial(3) * factorial(7 - 3)))
tortilla_combinations <- (factorial(3) / (factorial(1) * factorial(3 - 1)))

vegetable_combinations * condiment_combinations * tortilla_combinations

vegetables <- c('v1', 'v2', 'v3', 'v4', 'v5', 'v6', 'v7', 'v8')
condiments <- c('c1', 'c2', 'c3', 'c4', 'c5', 'c6', 'c7')
tortilla <- c('t1', 't2', 't3')

ncol(combn(vegetables, 3)) * ncol(combn(condiments, 3)) * ncol(combn(tortilla, 1))
```

## Question 6

Determine if the following events are independent. Jeff runs out of gas on the way to work. Liz watches the evening news. Answer: A) Dependent B) Independent

**Independent**

## Question 7

The newly elected president needs to decide the remaining 8 spots available in the cabinet he/she is appointing. If there are 14 eligible candidates for these positions (where rank matters), how many different ways can the members of the cabinet be appointed?

**121,080,960**

$P(n, r) = \frac{n!}{(n - r)!}$

so

$P(14, 8) = \frac{14!}{(14 - 8)!}= \frac{14!}{6!} = 121,080,960$

```{r}
factorial(14) / factorial(14 - 8)
```

## Question 8

A bag contains 9 red, 4 orange, and 9 green jellybeans. What is the probability of reaching into the bag and randomly withdrawing 4 jellybeans such that the number of red ones is 0, the number of orange ones is 1, and the number of green ones is 3? Write your answer as a fraction or a decimal number rounded to four decimal places.

**0.0124**

```{r, message=FALSE}
library(dplyr)
library(tidyr)

jellybeans <- c(rep("red", 9), rep("orange", 4), rep("green", 9))
combinations <- data.frame(matrix(combn(jellybeans, 4), ncol = 4))

total_combinations <- nrow(combinations)

combinations <- combinations %>%
  mutate(id = row_number()) %>%
  gather(col, color, -id) %>%
  mutate(count = 1) %>%
  select(-col) %>%
  group_by(id, color) %>%
  summarise(count = sum(count)) %>%
  spread(color, count, fill=0) %>%
  filter(orange == 1, green == 3)

round(nrow(combinations) / total_combinations, 4)
```

## Question 9

Evaluate the following expression.

$\frac{11!}{7!}$

**7,920**

$\frac{11!}{7!} = \frac{1 * 2 * 3* 4 * 5 * 6 * 7 * 8 * 9 * 10 * 11}{1 * 2 * 3* 4 * 5 * 6 * 7} = 8 * 9 * 10 * 11 = 7920$

```{r}
11 * 10 * 9 * 8

factorial(11) / factorial(7)
```

## Question 10

Describe the complement of the given event. 67% of subscribers to a fitness magazine are over the age of 34.

**33% of subscribers to a fitness magazine are not over the age of 34.**

## Question 11

If you throw exactly three heads in four tosses of a coin you win $97. If not, you pay me $30.

Step 1. Find the expected value of the proposition. Round your answer to two decimal places.

**$1.75**

The number of wining combinations is 4

\[ \begin{bmatrix}
  T & H & H & H \\
  H & T & H & H \\
  H & H & T & H \\
  H & H & H & T
  \end{bmatrix}
\]

The total number of combinations is $2^4 = 16$.

The probability of wining is $\frac{4}{16} = 0.25$ and the probability of loosing is the complement 0.75.

```{r}
(97 * 0.25) + (-30 * 0.75)
```

Step 2. If you played this game 559 times how much would you expect to win or lose? (Losses must be entered as negative.)

**$978.25**

```{r}
1.75 * 559
```

## Question 12

Flip a coin 9 times. If you get 4 tails or less, I will pay you $23. Otherwise you pay me $26.

Step 1. Find the expected value of the proposition. Round your answer to two decimal places.

**-$1.60**

```{r, warning=FALSE}
combinations <- data.frame(event1 = c("heads", "tails"))
for (i in 2:9){
  heads <- tails <- combinations
  heads$v <- 'heads'
  tails$v <- 'tails'
  n <- names(combinations)
  combinations <- rbind(heads, tails)
  names(combinations) <- c(n, paste0("event", i))
}

total_combinations <- nrow(combinations)

combinations <- combinations %>%
  mutate(id = row_number()) %>%
  gather(col, side, -id) %>%
  mutate(count = 1) %>%
  filter(side == "tails") %>%
  select(-col) %>%
  group_by(id, side) %>%
  summarise(count = sum(count)) %>%
  spread(side, count, fill=0) %>%
  filter(tails < 5)

p_win <- nrow(combinations) / total_combinations
p_loss <- 1 - p_win

expected <- (23 * p_win ) + (-26 * p_loss)
round(expected, 2)
```

Step 2. If you played this game 994 times how much would you expect to win or lose? (Losses must be entered as negative.)

**-$1,586.13**

```{r}
round(expected * 994, 2)
```

## Question 13

The sensitivity and specificity of the polygraph has been a subject of study and debate for years. A 2001 study of the use of polygraph for screening purposes suggested that the probability of detecting a liar was .59 (sensitivity) and that the probability of detecting a "truth teller" was .90 (specificity). We estimate that about 20% of individuals selected for the screening polygraph will lie.

a. What is the probability that an individual is actually a liar given that the polygraph detected him/her as such? (Show me the table or the formulaic solution or both.)

b. What is the probability that an individual is actually a truth-teller given that the polygraph detected him/her as such? (Show me the table or the formulaic solution or both.)

c. What is the probability that a randomly selected individual is either a liar or was identified as a liar by the polygraph? Be sure to write the probability statement.

### 2019-03-08 DATA605/MSilva_Assign6.Rmd
More work on HW 6

@@
 
+**0.118**
+
+Suppose you have 1,000 people.  Based on the given information, one would expect 200 of them to be liars and 800 to be truth tellers.  We know the formula for sensitivity is:
+
+$Sensitivity = \frac{TP}{TP + FN}$
+
+The denominator in this case would be the 200 liars.
+We know when you divide the TP by 200 you get .59, therefore we can figure that 118 of the 200 liars would be acurately predicted by the polygraph.
+The remaining 82 would be false negatives.
+
+Specificity is defined as:
+
+$Specificity = \frac{TN}{TN + FP}$
+
+Again we know that the denominator would be equal to the 800 truth tellers in our example and the specificity is 0.9.
+Therefore 720 individuals were correctly identified as truth tellers and 80 were false positives.
+
+Expresing this data in terms of probabilities we would have the following table:
+
+|              | Liar  | Truth Teller |
+|--------------|-------|--------------|
+| Liar         | 0.118 | 0.08         |
+| Truth Teller | 0.082 | 0.72         |
+
 b. What is the probability that an individual is actually a truth-teller given that the polygraph detected him/her as such? (Show me the table or the formulaic solution or both.)
 
+**0.89787556**
+
+Using the figures from the first exercise:
+
+```{r}
+TN <- 720
+FN <- 82
+TN / (TN + FN)
+```
+
 c. What is the probability that a randomly selected individual is either a liar or was identified as a liar by the polygraph? Be sure to write the probability statement.
+
+**0.28**
+
+This is either the 200 that are liars or the 80 that are detected incorectly as liars (FP) over the 1000 individuals.  That works out to 28%.

### 2019-03-15 DATA605/MSilva_Assign7.Rmd
Initial commit

---
title: "DATA 605 Week 7 Assignment"
author: "Mike Silva"
output: html_document
---

## Question 1

Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y .

## Question 2

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.)

### Part A

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a geometric.**

The probability the machine will fail within ten years is 1.  The probability that the machine will fail within k years is found using:

$$P(x=k) = (1-p)^{k-1} \times p$$
where p is the probability of the event happening that year.  We are given that p is 1/10.  The probability it will fail in the first eight years is displayed below

```{r echo=FALSE}
library(ggplot2)
n_failures_before_success <- 8
probability_of_success <- 1/10

p_ns <- c()
for (n in 1:n_failures_before_success){
  p_n <- probability_of_success * (1 - probability_of_success)^(n-1)
  p_ns <- c(p_ns, p_n)
}

ggplot(data.frame(year = 1:n_failures_before_success, p = p_ns), aes(year,p)) + 
  geom_bar(stat = "identity", fill = "steelblue", width = 1) +
  geom_text(aes(label = round(p,3)), vjust = 1.6, color = "white", size = 3.5) +
  theme_minimal() +
  ggtitle("Probability of Failure in the First 8 Years")
```

The probability of failure after the eight years will be one minus the sum of all the probabilities above (the area in blue).  Otherwise stated:

$$P(\text{failure after 8 years}) = 1 - \sum_{k=1}^{8} \frac{1}{10}\times\Big(\frac{9}{10}\Big)^{k-1} = 0.4304672$$

```{r}
n_failures_before_success <- 8
probability_of_success <- 1/10

p <- 0
for (n in 1:n_failures_before_success){
  p <- p + probability_of_success * (1 - probability_of_success) ^ (n - 1)
}

p_geometric <- 1 - p

message(paste("Probability that the machine will fail after 8 years:", p_geometric))
```

The [expected value](https://math.stackexchange.com/questions/2368304/how-do-you-calculate-the-expected-value-of-geometric-distribution-without-diffre) of a geometric distribution is

$$\text{Expected Value} = \frac{1}{P({success})} = \frac{1}{\frac{1}{10}} = 10$$

```{r}
expected_value <- 1 / probability_of_success

message(paste("Expected value:", expected_value))
```

The [standard deviation](https://www.youtube.com/watch?v=kDFiNt8nxSo) is:

$\text{Standard Deviation} = \sqrt{\frac{1-P(success)}{P(success)^2}} = \sqrt{\frac{1-0.1}{0.1^2}} = \sqrt{\frac{0.9}{0.01}} = \sqrt{90} \approx 9.486833$

```{r}
variance <- (1 - probability_of_success) / (probability_of_success ^ 2)

standard_deviation <- sqrt(variance)

message(paste("Standard deviation:", standard_deviation))
```

### Part B

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as an exponential.**

The probability the machine fails after 8 years is

$$P(\text{failure after 8 years}) = e^\frac{-8}{10} = e^{-0.8}=0.449329$$

```{r}
p_exponential <- exp(1) ^ (-n_failures_before_success /expected_value)

message(paste("Probability that the machine will fail after 8 years:", p_exponential))
```

The expected value is found using the same formula as above as demonstrated in [this video](https://www.youtube.com/watch?v=GJ2klfD0Q3g). Therefore

$$\text{Expected Value} = \frac{1}{P({success})} = \frac{1}{\frac{1}{10}} = 10$$

```{r}
expected_value <- 1 / probability_of_success

message(paste("Expected value:", expected_value))
```

The [standard deviation](https://www.youtube.com/watch?v=GJ2klfD0Q3g) is:

$\text{Standard Deviation} = \frac{1}{P(success)^2} = \frac{1}{0.1^2} = \frac{1}{0.01} = 100$

```{r}
standard_deviation <- 1 / (probability_of_success ^ 2)

message(paste("Standard deviation:", standard_deviation))
```

### Part C

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a binomial.**

## Part D

**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a Poisson**

## Summary

|        | Probability of Failure After 8 Years | Expected Value | Standard Deviation |
| Part A | 0.4304672                            | 10             | 9.48683298050514   |
| Part B | 0.449329                             | 10             | 100                |
| Part C |
| Part D |

### 2019-03-16 DATA605/MSilva_Assign7.Rmd
More work on Assignment 7

@@
 
-Let X1, X2, .
-.
-.
-, Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k.
-Let Y denote the minimum of the Xi’s.
-Find the distribution of Y .
+Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y.
+
+```{r}
+set.seed(42)
+n <- 100000
+k <- 50
+draws <- 10
+Y <- c()
+for (i in 1:n){
+  Xn <- sample(1:k, draws, TRUE)
+  Y <- c(Y, min(Xn))
+}
+```
+
+```{r, echo=FALSE}
+library(ggplot2)
+
+ggplot(data.frame(table(Y)), aes(Y, Freq)) + 
+  geom_bar(stat = "identity", fill = "steelblue", width = 1) +
+  theme_minimal() +
+  ylab("Count") + 
+  ggtitle("Distribution of Y's")
+```
+
+It's a long tail distribution like a power law distribution.
 
@@
 ```{r echo=FALSE}
-library(ggplot2)
 n_failures_before_success <- 8
@@
 ```{r}
-expected_value <- 1 / probability_of_success
+expected_value_geom <- 1 / probability_of_success
 
-message(paste("Expected value:", expected_value))
+message(paste("Expected value:", expected_value_geom))
 ```
@@
 
-$\text{Standard Deviation} = \sqrt{\frac{1-P(success)}{P(success)^2}} = \sqrt{\frac{1-0.1}{0.1^2}} = \sqrt{\frac{0.9}{0.01}} = \sqrt{90} \approx 9.486833$
+$$\text{Standard Deviation} = \sqrt{\frac{1-P(success)}{P(success)^2}} = \sqrt{\frac{1-0.1}{0.1^2}} = \sqrt{\frac{0.9}{0.01}} = \sqrt{90} \approx 9.486833$$
 
@@
 
-standard_deviation <- sqrt(variance)
+standard_deviation_geom <- sqrt(variance)
 
-message(paste("Standard deviation:", standard_deviation))
+message(paste("Standard deviation:", standard_deviation_geom))
 ```
@@
 ```{r}
-p_exponential <- exp(1) ^ (-n_failures_before_success /expected_value)
+p_exponential <- exp(1) ^ (-n_failures_before_success /expected_value_geom)
 
@@
 ```{r}
-expected_value <- 1 / probability_of_success
+expected_value_exp <- 1 / probability_of_success
 
-message(paste("Expected value:", expected_value))
+message(paste("Expected value:", expected_value_exp))
 ```
@@
 
-$\text{Standard Deviation} = \frac{1}{P(success)^2} = \frac{1}{0.1^2} = \frac{1}{0.01} = 100$
+$$\text{Standard Deviation} = \sqrt{\frac{1}{P(success)^2}} = \sqrt{\frac{1}{0.1^2}} = \sqrt{\frac{1}{0.01}} = \sqrt{100} = 10$$
 
 ```{r}
-standard_deviation <- 1 / (probability_of_success ^ 2)
+standard_deviation_exp <- sqrt(1 / (probability_of_success ^ 2))
 
-message(paste("Standard deviation:", standard_deviation))
+message(paste("Standard deviation:", standard_deviation_exp))
 ```
@@
 
-**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a binomial.**
+**What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)**
+
+The binomial probability formula is
+
+$$P(x) = \frac{N!}{x!(N-x)!}\pi^x(1-\pi)^{N-x}$$
+
+The probability of the maching failing after eight years is 1 minus the probability it does not fail in the first 8 years.
+Otherwise put 0 successes(x) in the first 8 years (N).
+The probability of failure is 1/10 therefore:
+
+$$P(0) = \frac{8!}{0!(8-0)!}\frac{1}{10}^0(1-\frac{1}{10})^{8-0} = \frac{40320}{40320}\times 1 \times \frac{9}{10}^8 = 1 \times 1 \times 0.9^8 = 0.4304672$$
+
+```{r}
+p_binomial <- pbinom(0, 8, 0.1)
+message(paste("Probability that the machine will fail after 8 years:", p_binomial))
+```
+
+The [expected value](http://onlinestatbook.com/2/probability/binomial.html) is the number of trials times the probability of success per trial.
+
+```{r}
+expected_value_binom <- n_failures_before_success * probability_of_success
+
+message(paste("Expected value:", expected_value_binom))
+```
+
+The [standard deviation](http://onlinestatbook.com/2/probability/binomial.html) is found using:
+
+$$\sigma = \sqrt{N\pi(1-\pi)}$$
+
+```{r}
+standard_deviation_binom <- sqrt(n_failures_before_success * probability_of_success*(1 - probability_of_success))
+
+message(paste("Standard deviation:", standard_deviation_binom))
+```
 
@@
 
-## Summary
+The Poisson formula is:
 
-|        | Probability of Failure After 8 Years | Expected Value | Standard Deviation |
-| Part A | 0.4304672                            | 10             | 9.48683298050514   |
-| Part B | 0.449329                             | 10             | 100                |
-| Part C |
-| Part D |
+$$P(x) = \frac{\lambda^xe^{-\lambda}}{x!}= \frac{10^8e^{-10}}{8!} =\frac{4539.993}{40320} = 0.112599$$
+
+```{r}
+x <- 8
+mu <- 10
+p_poisson <- ((mu^x) * (exp(1)^-mu)) / factorial(x)
+message(paste("Probability that the machine will fail after 8 years:", p_poisson))
+```
+
+For the Poisson distribution the mean and standard deviation are the same.
+
+```{r}
+expected_value_poisson <- mu
+
+message(paste("Expected value:", expected_value_poisson))
+```
+
+The variance is equal to mu so
+
+```{r}
+standard_deviation_poisson <- sqrt(mu)
+```
+
+### Summary
+
+The following table summarizes all the information found above
+
+```{r echo=FALSE, message=FALSE}
+library(kableExtra)
+library(dplyr)
+
+data.frame(Model = c("Geometric", "Exponential", "Binomial", "Poisson"),
+           Probability = c(p_geometric, p_exponential, p_binomial, p_poisson),
+           ex = c(expected_value_geom, expected_value_exp, expected_value_binom, expected_value_poisson),
+           sd = c(standard_deviation_geom, standard_deviation_exp, standard_deviation_binom, standard_deviation_poisson)) %>%
+  rename("Expected Value" = ex,
+         "Standard Deviation" = sd) %>%
+  kable() %>%
+  kable_styling()
+```

### 2019-03-18 DATA605/MSilva_Assign8.Rmd
Initial commit of Week 8 assignment

---
title: "DATA 605 Week 8 Assignment"
author: "Mike Silva"
output: html_document
---

## Question 11 on Pg. 303

A company buys 100 lightbulbs, each of which has an exponential lifetime of
1000 hours. What is the expected time for the first of these bulbs to burn
out?

## Question 14 on Pg. 303

Assume that X1 and X2 are independent random variables, each having an
exponential density with parameter λ. Show that Z = X1 − X2 has density

$$fz(z)=(1/2)\lambda e ^{-\lambda \mid z \mid}$$

## Question 1 Pg. 320-321

 Let X be a continuous random variable with mean µ = 10 and variance
$\sigma^2 = 100/3$. Using Chebyshev’s Inequality, find an upper bound for the
following probabilities.

### (a) P(|X − 10| ≥ 2).

### (b) P(|X − 10| ≥ 5).

### (c) P(|X − 10| ≥ 9).

### (d) P(|X − 10| ≥ 20).

### 2019-03-21 DATA605/MSilva_Assign8.Rmd
More work on Assignment 8

@@
 
-A company buys 100 lightbulbs, each of which has an exponential lifetime of
-1000 hours. What is the expected time for the first of these bulbs to burn
-out?
+**A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?**
+
+The exponential formula is:
+
+$$N(t) = N_0 e ^ {-\lambda t} $$
+
+lambda would be 1 failure every 1000 hours.  Otherwise stated:
+
+$$\lambda = \frac{1}{1000}$$
+
+Now we need to plug in the variables and solve for t
+
+$$100 e ^ {-\frac{1}{1000}t} = 99 \rightarrow e ^ {-\frac{1}{1000}t} = \frac{99}{100} \rightarrow -\frac{1}{1000}t = ln\bigg(\frac{99}{100}\bigg) \rightarrow t = ln\bigg(\frac{99}{100}\bigg)\times-1000 \approx 10.05034$$
+
+**Answer: 10.05034 hours**
 
@@
 
-Assume that X1 and X2 are independent random variables, each having an
-exponential density with parameter λ. Show that Z = X1 − X2 has density
+**Assume that X1 and X2 are independent random variables, each having an exponential density with parameter λ. Show that Z = X1 − X2 has density**
 
@@
 
- Let X be a continuous random variable with mean µ = 10 and variance
-$\sigma^2 = 100/3$. Using Chebyshev’s Inequality, find an upper bound for the
-following probabilities.
+**Let X be a continuous random variable with mean µ = 10 and variance $\sigma^2$ = 100/3. Using Chebyshev’s Inequality, find an upper bound for the following probabilities.**
 

### 2019-03-23 DATA605/MSilva_Assign8.Rmd
More work
Finished Assignment 8

@@
 
-**Assume that X1 and X2 are independent random variables, each having an exponential density with parameter λ. Show that Z = X1 − X2 has density**
+**Assume that X1 and X2 are independent random variables, each having an exponential density with parameter λ. Show that Z = X1 − X2 has density $fz(z)=\frac{1}{2}\lambda e ^{-\lambda \mid z \mid}$**
 
-$$fz(z)=(1/2)\lambda e ^{-\lambda \mid z \mid}$$
+*Note: In order to keep this straigth in my head I solved Z = X - Y instead of Z = X1 - X2.  This notation will be used in this solution.*
+
+The convolution formula of Z = X + Y is
+
+$$fx(z) = \int_{-\infty}^{\infty}fx(x)fy(z-x)dx$$
+
+But to do Z = X - Y we need to do Z = X + (-Y).  Pluging it into the convolution formula:
+
+$$fz(z) = \int_{-\infty}^{\infty}fx(x)f{-y}(z-x)dx$$
+Since $f{-y}(z-x) = fy(x-z)$ we can rewrite the convolution as
+
+$$fz(z) = \int_{-\infty}^{\infty}fx(x)fy(x-z)dx$$
+
+Looking at the case where z < 0
+
+$$ \int_{0}^{\infty}\lambda e ^{-\lambda x} \lambda e ^ {-\lambda(x-z)}dx = \lambda e ^ {\lambda z}\int_{0}^{\infty}\lambda e ^{-2\lambda x} dx = \lambda e ^{\lambda z}(-\frac{1}{2}e ^ {-\lambda x}|_0^{\infty}) = \frac{1}{2}\lambda e^{\lambda z}$$
+
+Now looking at the case where z is greater than or equal to zero.
+Since the two variables are independent Z = X - Y is the same as -Z = -X + Y.
+So the distribution is symetric arround zero thus fz(z) = fz(-z).
+Sp when z >= 0:
+
+$$fz(-z) = \frac{1}{2}\lambda e^{-\lambda z}$$
+
+This implies that $fz(z)=\frac{1}{2}\lambda e ^{-\lambda \mid z \mid}$
 
@@
 
+```{r}
+mu <- 10
+sigma <- sqrt(100/3)
+```
+
 ### (a) P(|X − 10| ≥ 2).
+
+```{r}
+k <- 2 / sigma
+1 / k^2
+```
 
@@
 
+```{r}
+k <- 5 / sigma
+1 / k^2
+```
+
 ### (c) P(|X − 10| ≥ 9).
 
+```{r}
+k <- 9 / sigma
+1 / k^2
+```
+
 ### (d) P(|X − 10| ≥ 20).
+
+```{r}
+k <- 20 / sigma
+1 / k^2
+```

### 2019-03-26 DATA605/MSilva_Assign9.Rmd
Initial Commit of Assignment 9

---
title: "DATA 605 Week 9 Assignment"
author: "Mike Silva"
output: html_document
---

## Problem 1

The price of one share of stock in the Pilsdorff Beer Company (see Exercise 8.2.12) is given by Yn on the nth day of the year. Finn observes that the differences Xn = Yn+1 − Yn appear to be independent random variables with a common distribution having mean µ = 0 and variance σ2 = 1/4. If Y1 = 100, estimate the probability that Y365 is

(a) ≥ 100.

(b) ≥ 110.

(c) ≥ 120.

## Problem 2

Calculate the expected value and variance of the binomial distribution using the moment generating function.

## Problem 3

Calculate the expected value and variance of the exponential distribution using the moment generating function.

### 2019-03-30 DATA605/MSilva_Assign9.Rmd
Work on Assignment 9
More work on Assignment 9

@@
 
+Since we are starting at 100 after day 1 the problem can be restated
+
+$$P(X_1 + ... + X_{364} \geq 0)$$
+
+$$X_1 + ... + X_{364} = N(364 \times 0, \sqrt{364} * \sqrt{\frac{1}{4}}) = N(0, 9.539392)$$
+
+$$P(X_1 + X_2 + ... + X_{364} \geq 0) = Z_0 = \frac{0 - 0}{9.539392} = 0$$
+
+Looking up a Z-Score of 0 gives a *probability of 0.5*
+
+```{r}
+pnorm(0/(sqrt(364)*.5), lower.tail = FALSE)
+```
+
 (b) ≥ 110.
 
+$$P(X_1 + X_2 + ... + X_{364} \geq 10) = Z_{10} = \frac{10 - 0}{9.539392} = 1.048285$$
+
+*The probability is 0.8527463*
+
+```{r}
+pnorm(10/(sqrt(364)*.5), lower.tail = FALSE)
+```
+
 (c) ≥ 120.
+
+$$P(X_1 + X_2 + ... + X_{364} \geq 20) = Z_{20} = \frac{20 - 0}{9.539392} = 2.09657$$
+
+*The probability is 0.9819842*
+
+```{r}
+pnorm(20/(sqrt(364)*.5), lower.tail = FALSE)
+```
 
@@
 
+The binomial function is $b(x, n, p) = \frac{n!}{x!(n-x)!}p^xq^{n-x}$ with $q=1-p$.
+
+The moment generating function is:
+
+$$Mx(t)=\sum_{x=0}^{n} e^{xt} \frac{n!}{x!(n-x)!}p^xq^{n-x} = \sum_{x=0}^{n} (pe^t)^x \frac{n!}{x!(n-x)!}q^{n-x} = (q+pe^t)^n$$
+
+Taking the derivative of the function with respect to t
+
+$$\frac{dM(t)}{dt} = pe^t n(q+pe^t)^{n-1}$$
+
+Evaluating at t=0
+
+$$E(x) = np(q+p)^{n-1} = np$$
+
+To find the variance we need to find E(x2)
+
+$$E(x^2) = np(q + p)^{n−2}(q + np) = np(q + np)$$
+
+We can now find the variance
+
+$$V = E(x^2) - (E(x))^2 = np(q + np) - (np)^2 = np(q+np) - n^2p^2 = npq$$
+
 ## Problem 3
@@
 Calculate the expected value and variance of the exponential distribution using the moment generating function.
+
+First we need to get the moment generating function:
+
+$$Mx(t) = \int_{-\infty}^\infty e^{tx}fx(x)dx = \int_0^\infty e^{tx}\lambda e^{-\lambda x}dx = \lambda \int_0^\infty e^{(t - \lambda)x}dx = \frac{\lambda}{t- \lambda}$$
+
+Now that the moment function is found we can find the first and second moments
+
+$$E(x) = \frac{\lambda}{(t- \lambda)^2}|_{t=0} = \frac{1}{\lambda}$$
+
+Here's the variance:
+
+$$V = E(x^2) = \frac{2\lambda}{(t- \lambda)^3}|_{t=0} = \frac{2}{\lambda^2}$$

### 2019-04-02 DATA605/MSilva_Assign10.Rmd
Completion of Assignment 10

---
title: "DATA 605 Week 10 Assignment"
author: "Mike Silva"
output: html_document
---

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.  A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if

(a) he bets 1 dollar each time (timid strategy).

**About 2%**

```{r timid}
states <- c("0", "1", "2", "3", "4", "5", "6", "7", "8")
transition_matrix <- matrix(c(1, 0, 0, 0, 0, 0, 0, 0, 0, 
                              0.6, 0, 0.4, 0, 0, 0, 0, 0, 0,
                              0, 0.6, 0, 0.4, 0, 0, 0, 0, 0,
                              0, 0, 0.6, 0, 0.4, 0, 0, 0, 0,
                              0, 0, 0, 0.6, 0, 0.4, 0, 0, 0,
                              0, 0, 0, 0, 0.6, 0, 0.4, 0, 0,
                              0, 0, 0, 0, 0, 0.6, 0, 0.4, 0,
                              0, 0, 0, 0, 0, 0, 0.6, 0, 0.4,
                              0, 0, 0, 0, 0, 0, 0, 0, 1), 
                            nrow = 9, 
                            byrow = TRUE,
                            dimnames = list(states, states))

starting_state <- c(0, 1, 0, 0, 0, 0, 0, 0, 0)

n <- 1000

for (i in 1:n){
  starting_state <- starting_state %*% transition_matrix
}
starting_state
```

```{r, echo = FALSE, eval=FALSE}
library(markovchain)
timid <- new("markovchain", states = states, byrow = TRUE, transitionMatrix = transition_matrix, name = "Timid")

c(0, 1, 0, 0, 0, 0, 0, 0, 0) * timid^1000
```

(b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).

**About 6.4%**

```{r bold}
transition_matrix <- matrix(c(1, 0, 0, 0, 0, 0, 0, 0, 0, 
                              0.6, 0, 0.4, 0, 0, 0, 0, 0, 0,
                              0.6, 0, 0, 0, 0.4, 0, 0, 0, 0,
                              0.6, 0, 0, 0, 0, 0, 0.4, 0, 0,
                              0.6, 0, 0, 0, 0, 0, 0, 0, 0.4,
                              0.6, 0, 0, 0, 0, 0, 0, 0, 0.4,
                              0.6, 0, 0, 0, 0, 0, 0, 0, 0.4,
                              0.6, 0, 0, 0, 0, 0, 0, 0, 0.4,
                              0, 0, 0, 0, 0, 0, 0, 0, 1), 
                            nrow = 9, 
                            byrow = TRUE,
                            dimnames = list(states, states))

starting_state <- c(0, 1, 0, 0, 0, 0, 0, 0, 0)

for (i in 1:n){
  starting_state <- starting_state %*% transition_matrix
}
starting_state
```

(c) Which strategy gives Smith the better chance of getting out of jail?

**The bold strategy**

### 2019-04-11 DATA605/MSilva_Exercise11.Rmd
Added Week 11 Exercise

---
title: "DATA 605 Week 11"
author: "Mike Silva"
date: "10 April 2019"
output: 
  rmdformats::readthedown:
    highlight: kate
---

## Introduction

As part of the Fundamentals of Computational Mathematics course work at the CUNY School of Professional Studies, we needed to complete a simple linear regression of data of our choice.  I have selected to look at the relationship between the total arcade revenue and the number of PhD computer science graduates.

## Data Acquisition

The data was scraped from [Tyler Vigen's site](http://www.tylervigen.com/view_correlation?id=97).  The total revenue from arcades is originally from the U.S. Census Bureau.  The computer science PhD's originally came from the National Science foundation.

```{r message=FALSE}
library(rvest)
library(stringr)
library(dplyr)
library(tidyr)
library(tibble)

df <- read_html("http://www.tylervigen.com/view_correlation?id=97") %>%
  html_nodes(xpath = "//table[@class='alldata']") %>%
  html_table(fill = TRUE) %>%
  as.data.frame() %>%
  select(-X12) %>% # Drop empty column
  slice(1:3) %>% # Keep the first 3 rows
  mutate(X1 = ifelse(X1 == "", "Year", str_extract(X1, "(.*?)\\s?\\(.*?\\)"))) %>% # Cleanup first column
  mutate(X1 = str_remove(X1, " \\(.*\\)")) %>% # More first column cleanup
  mutate(X1 = str_replace_all(X1, " ", ".")) %>% # Last bit of first column cleanup
  rownames_to_column %>% # Transpose dataframe
  gather(var, value, -rowname) %>% 
  spread(rowname, value) %>%
  select(-var) # Drop unneeded column

# Pull name from first row
names(df) <- df[1,]

df <- df %>%
  slice(2:11) %>% # Drop first row
  mutate(Total.revenue.generated.by.arcades = gsub(",", "", Total.revenue.generated.by.arcades)) %>% # Remove the commas
  mutate(Computer.science.doctorates.awarded = gsub(",", "", Computer.science.doctorates.awarded)) %>% # Remove the commas
  mutate_if(is.character, as.numeric) # Convert to numeric
```

The end result is the following data

```{r echo=FALSE, message=FALSE}
library(kableExtra)
df %>%
  rename("Computer Science Doctorates Awarded" = Computer.science.doctorates.awarded) %>%
  rename("Total Revenue Generated by Arcades" = Total.revenue.generated.by.arcades) %>%
  kable() %>%
  kable_styling()
```

## Model Building

Now that the data is cleaned up we will fit a line to the data

```{r}
model <- lm(Total.revenue.generated.by.arcades ~ Computer.science.doctorates.awarded, data = df)
```

## Model Evaluation

```{r echo = FALSE}
library(ggplot2)
ggplot(df, aes(Computer.science.doctorates.awarded, Total.revenue.generated.by.arcades)) +
  geom_point() +
  geom_abline(slope = model$coefficients[2], intercept = model$coefficients[1], color="red") +
  ylab("Total Revenue Generated by Arcades (millions $)") +
  xlab("Computer Science Doctorates Awarded")
```

```{r comment=NA}
summary(model)
```

### Residual Analysis

```{r}
plot(model)
```

## Conclusion

This model explains almost all of the variablitiy.  The R-squared is about 0.97.  The residual plots do not raise any concern.  The Q-Q is fairly linear.  The coefficients are statistically significant.  The only problem with this model is that the relationship is a [spurious correlations](http://www.tylervigen.com/view_correlation?id=97).

### 2019-04-13 DATA605/MSilva_Assign11.Rmd
Initial commit of week 11 work

---
title: "DATA 605 Week 11 Assignment"
author: "Mike Silva"
output: html_document
---

Using the "cars" dataset in R, build a linear model for stopping distance as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

## Visualization

```{r echo=FALSE}
library(ggplot2)
library(ggthemes)

ggplot(cars, aes(speed, dist)) +
  geom_point() +
  ylab("Stopping Distance (Feet)") +
  xlab("Speed (MPH)") + 
  theme_economist_white()
```

There appears to be a positive linear relationship between the speed of a car and the stopping distance.

## Model Creation & Evaluation

Let's fit a line to the data

```{r}
model <- lm(dist ~ speed, cars)
```

And let's visualize it

```{r echo=FALSE}
lm_eqn <- function(m){
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = unname(coef(m)[1]),
              b = unname(coef(m)[2]),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq))
}

ggplot(cars, aes(speed, dist)) +
  geom_point() +
  geom_abline(slope = model$coefficients[2], intercept = model$coefficients[1], color="red") +
  ylab("Stopping Distance (Feet)") +
  xlab("Speed (MPH)") + 
  geom_label(x = 10, y = 115, label.size = NA, label = lm_eqn(model), parse = TRUE) +
  theme_economist_white()
```

```{r comment=NA}
summary(model)
```

Every additional MPH in speed requires about 4 feet of stopping distance.  The standard error on this variable is about 9.5 times smaller than the estimate which is good.  The P-value of this coefficient is very small indicating that it is statistically significant.  This model explains about 65% of the variation.

## Residual Analysis

```{r echo=FALSE}
ggplot(cars, aes(resid(model))) +
  geom_density() +
  ylab("Density") +
  xlab("Residuals") + 
  geom_vline(xintercept = 0, linetype="dotted", color = "red") +
  theme_economist_white()
```

From the summary above and the chart of the residuals above we can note the residuals range from -29 to 43 with a median at -2.27 and the first and third quartiles are -9.5 and 9.2 respectively.  From that it would seem they are largely normaly distributed with a bit of an outlier on the the positive side.

```{r echo=FALSE}
ggplot(cars, aes(fitted(model),resid(model))) +
  geom_point() +
  ylab("Residuals") +
  xlab("Fitted Values") + 
  theme_economist_white()
```

There are three observations that look like outliers on the positive side of the distribution.  Outside of this the variation is fairly uniformly distributed.  Now for the Q-Q plot.

```{r echo=FALSE}
gg_qqplot <- function(fitted.lm) {
   # extract residuals from lm object
   res = residuals(fitted.lm)

   # calculate slope and interncept for qqline
   slope = (quantile(res, .75) - quantile(res, .25)) / (qnorm(.75) - qnorm(.25))
   intercept = quantile(res,.25) - slope*qnorm(.25)
   qq_line = data.frame(intercept = intercept, slope = slope)

   # generate ggplot for qqplot
   qq_plot <- ggplot(data = fitted.lm) +
      stat_qq(aes(sample = res)) +
      labs(x = "Theoretical Quantile", y = "Standardized Residual") +
      geom_abline(data = qq_line ,aes(intercept = intercept ,slope = slope), color = "red") +
    theme_economist_white()   
   qq_plot
}

gg_qqplot(model)
```

Again we see that most of the residuals lay on the line except for three observations which seem to be outliers.

### 2019-04-20 DATA605/MSilva_Assign12.Rmd
More work on Assignment 12

---
title: "DATA 605 Week 12 Assignment"
author: "Mike Silva"
output: html_document
---

## Part 1 - Original Data

### Data Acquisition/Wrangling

```{r}
df <- read.csv("who.csv")
```

Here's the first 5 observations of the data

```{r echo=FALSE, message=FALSE}
library(dplyr)
library(kableExtra)
df %>%
  head(5) %>%
  kable() %>%
  kable_styling()
```

### Visualization

```{r echo=FALSE}
library(ggplot2)
library(ggthemes)

ggplot(df, aes(TotExp, LifeExp)) +
  geom_point() +
  ylab("Avg Life Expectancy (years)") +
  xlab("Avg Personal and Government Expenditures (US dollars)") + 
  theme_economist_white()
```

### Model Building

```{r}
model <- lm(LifeExp~TotExp, df)
```

### Model Evaluation

The F statistic is `r summary(model)$fstatistic["value"]`.  It can give a yes-no response to "can I reject $H_0$".  In this case F statistic indicates that there's a relationship that is statistically significant and $R^2$ will indicate the strength of the relationship.

The $R^2$ is `r summary(model)$r.squared`.  It explains how much of the variation in life expectancy is explained by the total expenditures.  Only 25% of the variation is explained which is not a very good fit.

The standard error for total expenditures is `r summary(model)$coefficients[1,"Std. Error"]`.  It means the total expenditure coefficient is `r summary(model)$coefficients[1,"Estimate"]` +/- `r summary(model)$coefficients[1,"Std. Error"] * 1.96` at the 95% confidence level.

The P-value for total expenditures is `r summary(model)$coefficients[1,"Pr(>|t|)"]` which indicates the beta is statistically significant.

## Part 2 - Transformed Data

### Data Acquisition/Wrangling

```{r message=FALSE}
library(dplyr)
df <- df %>%
  mutate(TransformedTotExp = TotExp^.06,
         TransformedLifeExp = LifeExp^4.6)
```

### Visualization

```{r echo=FALSE}
ggplot(df, aes(TransformedTotExp, TransformedLifeExp)) +
  geom_point() +
  ylab("Avg Life Expectancy (years) to the 4.6 power") +
  xlab("Avg Personal and Government Expenditures (US dollars) to the 0.06 power") + 
  theme_economist_white()
```

### Model Building

```{r}
model <- lm(TransformedLifeExp~TransformedTotExp, df)
```

### Model Evaluation

The F statistic (`r summary(model)$fstatistic["value"]`) is larger than before and indicates that there's a relationship that is statistically significant. The $R^2$ (`r summary(model)$r.squared`) means about 73% of the variation is now explained which is a much better fit.

The standard error for total expenditures is `r summary(model)$coefficients[1,"Std. Error"]`.  It means the total expenditure coefficient is `r summary(model)$coefficients[1,"Estimate"]` +/- `r summary(model)$coefficients[1,"Std. Error"] * 1.96` at the 95% confidence level.

The P-value for the total expenditures is `r summary(model)$coefficients[1,"Pr(>|t|)"]` now which indicates the estimate is statistically significant.

This model is definately a better fit although the I personally find it harder to explain to less statistically litterate users due to the transformation.

## Part 3

```{r, results='asis'}
forecast_life_expectancy <- function(TransformedTotExp){
  d <- data.frame(TransformedTotExp = c(TransformedTotExp))
  prediction <- predict(model, d)
  cat(paste("Life Expectancy:", round(unname(log(prediction, 4.6)),1), "years"))
}
forecast_life_expectancy(1.5)
```

```{r, results='asis'}
forecast_life_expectancy(2.5)
```

## Part 4

### Model Building

```{r}
model <- lm(LifeExp ~ PropMD + TotExp + I(PropMD * TotExp), df)
```

### Model Evaluation

The F statistic (`r summary(model)$fstatistic["value"]`) is smaller than the model in Part 2.  It still indicates that there's a relationship that is statistically significant. The $R^2$ (`r summary(model)$r.squared`) means about 36% of the variation is explained by this model which is worse than the prvious model.

The standard error for Prop MD is `r summary(model)$coefficients[1,"Std. Error"]`.  It means the total expenditure coefficient is `r summary(model)$coefficients[1,"Estimate"]` +/- `r summary(model)$coefficients[1,"Std. Error"] * 1.96` at the 95% confidence level.

The standard error for total expenditures is `r summary(model)$coefficients[2,"Std. Error"]`.  It means the total expenditure coefficient is `r summary(model)$coefficients[2,"Estimate"]` +/- `r summary(model)$coefficients[2,"Std. Error"] * 1.96` at the 95% confidence level.

The standard error for the interaction term is `r summary(model)$coefficients[3,"Std. Error"]`.  It means the total expenditure coefficient is `r summary(model)$coefficients[3,"Estimate"]` +/- `r summary(model)$coefficients[3,"Std. Error"] * 1.96` at the 95% confidence level.

The P-value for the first term is `r summary(model)$coefficients[1,"Pr(>|t|)"]`  which indicates the estimate is statistically significant.  The second term is `r summary(model)$coefficients[2,"Pr(>|t|)"]`  which indicates the estimate is statistically significant as is the third term (`r summary(model)$coefficients[3,"Pr(>|t|)"]` ).

This model is not as good as the second model.

## Part 5

```{r, results='asis'}
forecast_life_expectancy <- function(PropMD, TotExp){
  d <- data.frame(PropMD = c(PropMD), TotExp = c(TotExp))
  prediction <- predict(model, d)
  cat(paste("Life Expectancy:", round(unname(prediction),1), "years"))
}
forecast_life_expectancy(0.03, 14)
```

This forecast does not seem realistic because the average live expectancy is not that high.

### 2019-04-21 DATA605/MSilva_Exercise12.Rmd
Added Week 12 Exercise

---
title: "DATA 605 Week 12"
author: "Mike Silva"
date: "16 April 2019"
output: 
  rmdformats::readthedown:
    highlight: kate
---

## Introduction

As part of the Fundamentals of Computational Mathematics course work at the CUNY School of Professional Studies, we need to build a multiple regression model.  The model needs at least one quadratic term, one dichotomous term, and one dichotomous vs. quantitative interaction term.  For this exercise I will be using the clasic Iris data set.

## Data Acquisition

I will be predicting the petal length.  I know that there is a strong relationship between the sepal length and the petal length (`r cor(iris$Sepal.Length, iris$Petal.Length)` correlation).  I will be using this in the model as a quadratic term.

Because of my knowledge of this dataset I know that the setosa species of iris is distinct from the versicolor and virginica species.  I will be using that as the dichotomous term.

I want to differentiate the versicolor and viginica group and will use the petal width as the quantitative interaction term.

```{r, message=FALSE}
library(dplyr)
df <- iris %>%
  mutate(Sepal.Length.Squared = Sepal.Length ^ 2) %>% # Quadratic term
  mutate(Not.Setosa = ifelse(Species == "setosa", 0, 1)) %>% # Dichotomous term
  mutate(Petal.Width.Interaction = Not.Setosa * Petal.Width) %>% # Dichotomous vs. quantitative interaction
  select(Petal.Length, Sepal.Length.Squared, Not.Setosa, Petal.Width.Interaction) %>%
  arrange(Petal.Length)
```

The end result is the following data

```{r echo=FALSE, message=FALSE}
library(kableExtra)
df %>%
  head(10) %>%
  kable() %>%
  kable_styling()
```

## Visualization

It's a bit harder to visualize this model because of the number of dimensions but here's the clearest viz I was able to work up:

```{r echo = FALSE}
library(ggplot2)
df %>%
  mutate(Not.Setosa = ifelse(Not.Setosa == 1, "Yes", "No")) %>%
  mutate(Not.Setosa = as.factor(Not.Setosa)) %>%
  ggplot(aes(Sepal.Length.Squared, Petal.Length, shape=Not.Setosa, color=Petal.Width.Interaction)) +
  geom_point() +
  ylab("Petal Length") +
  xlab("Sepal Length Squared") +
  scale_color_distiller(palette = "Spectral")
```

## Model Creation

Now to create the multivariate model

```{r}
model <- lm(Petal.Length ~ ., df)
```

## Model Evaluation

```{r comment=NA}
summary(model)
```

This model preformed very well.  The adjusted R-squared is `r summary(model)$adj.r.squared` which is phenominal!  All of the terms are statistically significant.

### Residual Analysis

Let's take a look at the residuals

```{r, comment=NA}
summary(resid(model))
```

The mean and median are close to zero which is good.  All of the other data hint that the residuals are fairly normally distributed.  Let's take a look at the plots:

```{r, echo=FALSE}
library(ggfortify)
autoplot(model)
```

There are a couple observations that are outliers but generally everything looks great!

## Conclusion

This model explains almost all of the variablitiy.  The R-squared is about 0.97.  The residual plots do not raise any concern.  The Q-Q is fairly linear.  The coefficients are statistically significant.

### 2019-04-23 DATA605/MSilva_Assign13.Rmd
Initial commit of Assignment 13

---
title: "DATA 605 Week 13"
author: "Mike Silva"
date: "April 23, 2019"
output: 
  rmdformats::readthedown:
    highlight: kate
---

## Question 1
**Use integration by substitution to solve the integral below.**

$$ \int{4e^{-7x}dx}$$

The integral of a constant times a function is the constant times the integral of the function so

$$ \int{4e^{-7x}dx} \rightarrow 4 \int{e^{-7x}dx}$$

Let u = -7x

$$ du = -7dx \rightarrow \frac{du}{-7}=dx\rightarrow dx = -\frac{1}{7}du$$

Substituded in

$$ 4 \int{e^{-7x}dx} \rightarrow 4\int{-\frac{e^u}{7}du} \rightarrow -\frac{4}{7}\int{e^u du}$$

Since  $\int{e^u du} = e^u$ we get

$$ -\frac{4}{7}\int{e^u du} \rightarrow -\frac{4}{7}e^u$$

Substitute for u we get

$$ -\frac{4}{7}e^{-7x} + C$$

## Question 2

**Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of $\frac{dN}{dt}=-\frac{3150}{t^4}-220$ bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function N(t) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.**

## Question 3

**Find the total area of the red rectangles in the figure below, where the equation of the line is f(x) = 2x-9.**

## Question 4

**Find the area of the region bounded by the graphs of the given equations **

$$y=x^2-2x-2, y=x+2$$

## Question 5

**A beauty supply store expects to sell 110 flat irons during the next year. It costs $3.75 to store one flat iron for one year. There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.**

## Question 6

**Use integration by parts to solve the integral below.**

$$\int{ln(9x) \times x^6 dx}$$

## Question 7

**Determine whether f(x) is a probability density function on the interval $[1, e^6]$ . If not, determine the value of the definite integral.**

$$f(x)=\frac{1}{6x}$$

### 2019-04-23 DATA605/MSilva_Assign13.Rmd
Completed Question 3

@@
 
+$$\int{-\frac{3150}{t^4}-220 dt}\rightarrow-220t-3150\frac{1}{-3t^3}\rightarrow-220t+\frac{1050}{t^3} + C$$
+
+To solve for the constant we solve for $C$ given that $N(t) = 6530$
+
+$$6530 = -220(1)+\frac{1050}{(1)^3} + C \rightarrow 6530 = -220 + 1050 + C \rightarrow 6530 = 830 + C \rightarrow 5700 = C$$
+
+The function is $N(t)=\frac{1050}{t^3}-220t+5700$
+
 ## Question 3
 
-**Find the total area of the red rectangles in the figure below, where the equation of the line is f(x) = 2x-9.**
+**Find the total area of the red rectangles in the figure below, where the equation of the line is $f(x) = 2x-9$.**
+
+![](MSilva_Assign13_Q3.png)
+
+The area under the line is a triangle from 4.5 to 8.5.  The height is found by pluging 8.5 into the function.
+
+$$f(8.5)=2(8.5)-9=17-9=8$$
+
+The area of the triangle is 1/2 base x height so
+
+$$\frac{1}{2}(8.5 - 4.5) \times 8 = \frac{1}{2}(4) \times 8 = 2 \times 8 = 16$$
+
+So the area is 16.  Verifying with R
+
+```{r, comment=NA}
+fx <- function(x){
+  2*x-9
+}
+integrate(fx, 4.5, 8.5)
+```
 

### 2019-04-24 DATA605/MSilva_Exercise13.Rmd
Initial commit of Week 13 Exercise

---
title: "DATA 605 Week 13"
author: "Mike Silva"
date: "April 23, 2019"
output: 
  rmdformats::readthedown:
    highlight: kate
---

## Fence Optimization

A rancher has 1000 feet of fencing in which to construct adjacent, equally sized rectangular pens. What dimensions should these pens have to maximize the enclosed area?

![](MSilva_Exercise13.png)

I will refer to the horizontal sections as X and the vertical sections as Y.  To solve this problem we need to maximize the area (Area = xy) subject to the materials constraint (1000 = 2x + 3y).

We have two equations with two unknowns.  I will solve for y in the second one:

$$1000 = 4x+3y \rightarrow 3y = -4x + 1000 \rightarrow y = -\frac{4}{3}x + \frac{1000}{3}$$

Plugging in the Y into the first equation gives:

$$ A(x) = x \times (-\frac{4}{3}x + \frac{1000}{3}) $$

$$ A(x) = -\frac{4}{3}x^2 + \frac{1000}{3}x $$

In order to maximize we take the derivative of A(x), set it equal to 0 and solve for x:

$$ A'(x) = -\frac{8}{3}x + \frac{1000}{3} $$

$$ 0 = -\frac{8}{3}x + \frac{1000}{3} \rightarrow \frac{8}{3}x = \frac{1000}{3}\rightarrow 3(\frac{8}{3}x) = 3(\frac{1000}{3})\rightarrow 8x = 1000 \rightarrow x = 125$$

Now to solve for Y.  Starting with the contraint equation

$$y = -\frac{4}{3}x + \frac{1000}{3} \rightarrow y = -\frac{4}{3}(125) + \frac{1000}{3} \approx 166.6667$$

So the dimensions of the pen to get the maximum area are *125 x 167 feet*.

### 2019-04-24 DATA605/MSilva_Assign13.Rmd
More work

@@
 
-$$y=x^2-2x-2, y=x+2$$
+$$f(x)=x^2-2x-2, g(x)=x+2$$
+
+First we will visualize the two lines.  The first is blue and the second is red.
+
+```{r, echo=FALSE, warning=FALSE}
+library(ggplot2)
+
+fx <- function(x){x^2 - 2*x - 2}
+gx <- function(x){x + 2}
+
+ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
+  stat_function(fun = gx, color="red") + 
+  stat_function(fun = fx, color="blue") +
+  geom_hline(yintercept = 0, color="black") +
+  geom_vline(xintercept = 0, color="black") +
+  xlim(-8, 8) +
+  ylim(-8, 8) +
+  theme(aspect.ratio=1)
+```
+
+The two points where the lines intersect are
+
+$$x^2-2x-2 = x+2 \rightarrow x^2-3x -4 = 0\rightarrow(x+1)(x-4) = 0\rightarrow x = -1 \text{ and } 4$$
+
+$$\int_{-1}^4{(x+2)-(x^2-2x-2)} \rightarrow (x^2+2x)-(\frac{x^3}{3}+x^2-2x)|_{-1}^4$$
+
+```{r, comment=NA}
+fx <- function(x){
+  (x+2)-(x^2 - 2*x - 2)
+}
+integrate(fx, -1, 4)
+```
 

### 2019-04-27 DATA605/MSilva_Assign13.Rmd
Completed Assignment 13

@@
 date: "April 23, 2019"
-output: 
-  rmdformats::readthedown:
-    highlight: kate
+output: html_document
 ---
@@
 ## Question 1
+
 **Use integration by substitution to solve the integral below.**
@@
 
-$$\int_{-1}^4{(x+2)-(x^2-2x-2)} \rightarrow (x^2+2x)-(\frac{x^3}{3}+x^2-2x)|_{-1}^4$$
+  $$\int_{-1}^4{(x+2)-(x^2-2x-2)} $$
 
@@
 
+I am assuming sales are uniform throughout the year (no seasonality).
+I am also assuming that there is no lag between ordering and receiving the items.
+I am also assumuing that the orders will be for the same lot size.
+I am also assuming that I am starting with zero irons in inventory on day one.
+
+Total costs are the sum of the carrying/holding costs (H(x)) and the reordering costs(R(x)).
+
+$$H(x) =  3.75 \times \frac{x}{2} \rightarrow H(x) = 1.875x$$
+and
+
+$$R(x) = 8\times\frac{110}{x}\rightarrow R(x)=\frac{880}{x}$$
+
+So total costs as a function is:
+
+$$C(x) = 1.875x + \frac{880}{x}$$
+
+We will take the first derivative and set it equal to zero and solve for x to find the number of orders to minimize the costs.
+
+$$C'(x) =1.875 - \frac{880}{x^2}$$
+
+$$0 =1.875 - \frac{880}{x^2}$$
+
+$$-1.875 = - \frac{880}{x^2}$$
+
+$$-1.875 x^2= - 880$$
+
+$$x^2 = \frac{-880}{-1.875}$$
+
+$$x = \sqrt{\frac{880}{1.875}} = 21.6641 $$
+**The costs are minimized at 22 orders with 5 irons per order**
+
+The blue line in the following graph is the cost function and the red point is where it is minimized.
+
+```{r, echo=FALSE}
+library(ggplot2)
+get_costs <- function(x){
+  return(1.875 * x + (880 / x))
+}
+
+df <- data.frame(orders = 1:110, costs = get_costs(1:110))
+m <- df[df$costs == min(df$costs),]
+
+ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
+  stat_function(fun = get_costs, color="blue") + 
+  xlim(0, 110) +
+  ylim(0, 1000) + 
+  geom_point(x=m$orders, y=m$costs, color="red") +
+  geom_hline(yintercept = 0, color="black") +
+  geom_vline(xintercept = 0, color="black") +
+  xlab("Orders") +
+  ylab("Total Costs") +
+  theme(aspect.ratio=1)
+```
+
 ## Question 6
@@
 
-$$\int{ln(9x) \times x^6 dx}$$
+$$\int{ln(9x) x^6 dx}$$
+
+The integration by part formula is
+
+$$\int{f(x)g'(x)}=f(x)g(x)-\int{f'(x)g(x)}$$
+
+Let $f(x) = ln(9x)$ and let $g'(x) = x^6$.  $f'(x)=\frac{1}{x}$ and $g(x) = \int{x^6}=\frac{x^7}{7}$.
+
+Substituting
+
+$$ln(9x)x^6- \int{\frac{1}{x}\frac{x^7}{7} }$$
+
+Evaluating the sub-integral
+
+$$\int{\frac{1}{x}\frac{x^7}{7}} = \int{\frac{x^6}{7}} = \frac{x^7}{49}$$
+
+$$ln(9x)x^6- \frac{x^7}{49}$$
 
@@
 $$f(x)=\frac{1}{6x}$$
+
+If it is a PDF then
+
+$$F(x) = \int_1^{e^6}{\frac{1}{6x}dx}=1$$
+
+$$\int_1^{e^6}{\frac{1}{6x}dx} = \frac{1}{6}ln(x) |_1^{e^6} = \frac{1}{6}ln(e^6) - \frac{1}{6}ln(1)=\frac{1}{6}\times6 - \frac{1}{6}\times 0 = 1 - 0 = 1$$
+
+**Yes f(x) is a probability density function on the interval $[1, e^6]$ **

### 2019-05-06 DATA605/MSilva_Assign15.Rmd
Initial commit of Assignment 15

---
title: "DATA 605 Week 15"
author: "Mike Silva"
date: "May 6, 2019"
output: html_document
---

## Question 1

Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.

( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

## Question 2

Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form ( x, y, z ). Separate multiple points with a comma.

$$f(x,y)=24x-6xy^2-8y^3$$

## Question 3

A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81  21x + 17y units of the "house" brand and 40 + 11x  23y units of the "name" brand.

**Step 1.** Find the revenue function R ( x, y ).

**Step 2.** What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

## Question 4

A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by $C(x,y)=\frac{1}{6}x^2+\frac{1}{6}y^2+7x+25y+700$, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

## Question 5

Evaluate the double integral on the given region.

$\int{\int{(e^{8x+3y})dA}}$; $R: 2 \leq x \leq 4$ and $2 \leq y \leq 4$

Write your answer in exact form without decimals.

### 2019-05-08 DATA605/MSilva_Assign15.Rmd
More work on Assigment 15

@@
 ( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )
+
+Given the set of (X, Y) points we need to find the best fit line $\hat{Y}= a + bX$ such that the sum of square errors in Y ($\sum{(Y-\hat{Y})^2}$) is minimized.
+
+$$SSE=\sum_{i=1}^{n}{(Y_i-\hat{Y})^2} = \sum_{i=1}^{n}{(Y_i-a - bX_i)^2}$$
+
+SSE will be minimized when $\frac{d{SSE}}{da} = 0$ and $\frac{d{SSE}}{db} = 0$.
+
+$$\frac{d{SSE}}{da} = \sum_{i=1}^{n}{-2(Y_i-a - bX_i)}=2\bigg(na+b\sum_{i=1}^{n}{X_i}-\sum_{i=1}^{n}{Y_i}\bigg)=0$$
+
+Dividing by 2 and solving for a
+
+$$a = \bar{Y}-b\bar{X}$$
+
+$$\frac{d{SSE}}{db} = \sum_{i=1}^{n}{-2X_i(Y_i-a - bX_i)}=\sum_{i=1}^{n}{-2(X_iY_i-aX_i-bX_i^2)}=0$$
+
+Substituting in a gives
+
+$$\sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y}-bX_i\bar{X}-bX_i^2)}=0 \rightarrow \sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y})}-b\sum_{i=1}^{n}{(X_i^2-X_i\bar{X})}=0$$
+
+Solving for b gives
+
+$$b=\frac{\sum_{i=1}^{n}{(X_iY_i-X_i\bar{Y})}}{\sum_{i=1}^{n}{(X_i^2-X_i\bar{X})}}$$
+
+```{r}
+X <- c(5.6, 6.3, 7, 7.7, 8.4)
+Y <- c(8.8, 12.4, 14.8, 18.2, 20.8)
+
+X_bar <- mean(X)
+Y_bar <- mean(Y)
+
+b <- sum(X * Y - X * Y_bar) / sum(X ^ 2 - X * X_bar)
+
+a <- Y_bar - b * X_bar
+```
+
+The least squares line is $\hat{Y}$ = `r a` + `r b` $X$
+
+Double checking my work
+```{r}
+lm(Y ~ X)
+```
 

### 2019-05-14 DATA605/MSilva_Assign15.Rmd
More work on assignment 15

@@
 Double checking my work
+
 ```{r}
@@
 
+First we find the first and second partial derivatives
+
+| First Derivative    | Second Derivative     |
+|---------------------|-----------------------|
+|$f_x = 24 - 6y^2$    | $f_{xx} = 0$          |
+|$f_y = -12xy -24y^2$ | $f_{yy} = -12x - 48y$ |
+
+and $f_{xy} = 12y$.  Now we find the critical points
+
+$$f_x(x,y) = 0 \rightarrow 24 - 6y^2=0 \rightarrow -6y^2 = -24 \rightarrow y^2 = 4 \rightarrow y = \pm 2 $$
+
+$$f_y(x,y) = 0 \rightarrow  -12xy - 24y^2 = 0 \rightarrow-12xy = 24y^2 \rightarrow  x = -2y$$
+
+When y = 2 x = -4 and when y = -2 x = 4.  In order to find the z we plug these values into the function.
+
+```{r}
+fxy <- function(x, y){
+  24*x-6*x*y^2-8*y^3
+}
+
+fxy(4, -2)
+fxy(-4, 2)
+```
+
+In order to figure out what type of point it is, we need to plug it into the second derivative test.
+
+```{r}
+second_derivative_test <- function(x,y){
+  #fxx * fyy - fxy^2
+  (0 * (-12*x - 48*y))-((12*y)^2)
+}
+
+second_derivative_test(4, -2)
+second_derivative_test(-4, 2)
+```
+
+| Point (x, y, z) | Second Derivative Test | Conclusion   |
+|-----------------|------------------------|--------------|
+|(4, -2, 64)      | Negative               | Saddle Point |
+|(-4, 2, -64)     | Negative               | Saddle Point |
+
 ## Question 3
@@
 A grocery store sells two brands of a product, the "house" brand and a "name" brand.
-The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81  21x + 17y units of the "house" brand and 40 + 11x  23y units of the "name" brand.
+The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell 81 - 21x + 17y units of the "house" brand and 40 + 11x -23y units of the "name" brand.
 
-**Step 1.** Find the revenue function R ( x, y ).
+**Step 1.** Find the revenue function R (x, y).
+
+$R(x,y) = x(81 - 21x + 17y) + y(40 + 11x - 23y)$
 
 **Step 2.** What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?
+
+```{r}
+revenue <- function(x,y){
+  x * (81 - 21 * x + 17 * y) + y * (40 + 11 * x -23 * y)
+}
+
+revenue(2.30, 4.10)
+```
+
+R(x,y) = $116.62
 
@@
 
+Let's express the cost equation as a function of x.  Given $x + y = 96$, $y = 96 - x$.
+
+$$C(x)=\frac{1}{6}x^2+\frac{1}{6}(96-x)^2+7x+25(96-x)+700$$
+
+$$C(x)=\frac{x^2}{6}+\frac{1}{6}(96-x)^2-18x+3100$$
+
+$$C(x)=\frac{x^2}{3}-50x+4636$$
+
+Now minimize it
+
+$$C'(x) =  \frac{2x}{3}-50 \rightarrow 0 = \frac{2x}{3}-50 \rightarrow\frac{2x}{3} = 50 \rightarrow 2x = 150 \rightarrow x = 75$$
+
+$$y = 96 - (75) \rightarrow y = 21$$
+
+| Location    | Units |
+|-------------|-------|
+| Los Angeles | 75    |
+| Denver      | 21    |
+
 ## Question 5

### 2019-05-15 DATA605/MSilva_Final.Rmd
Initial commit

---
title: "DATA 605 Final"
author: "Mike Silva"
output: html_document
---

## Problem 1

Using R, generate a random variable X that has 10,000 random uniform numbers from 1 to N, where N can be any number of your choosing greater than or equal to 6. Then generate a random variable Y that has 10,000 random normal numbers with a mean of $\mu=\sigma=\frac{N+1}{2}$.

```{r}
set.seed(12345)
N <- 10
X <- runif(10000, min=1, max=N)
mu <- sigma <- (N + 1)/2
Y <- rnorm(10000, mean=mu, sd=sigma)
```

*Probability.* Calculate as a minimum the below probabilities a through c. Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable. Interpret the meaning of all probabilities.

a. P(X>x | X>y)

b. P(X>x, Y>y)

c. P(X<x | X>y)

Investigate whether P(X>x and Y>y)=P(X>x)P(Y>y) by building a table and evaluating the marginal and joint probabilities.

Check to see if independence holds by using Fisher’s Exact Test and the Chi Square Test. What is the difference between the two? Which is most appropriate?

## Problem 2

You are to register for Kaggle.com (free) and compete in the House Prices: Advanced Regression Techniques competition. https://www.kaggle.com/c/house-prices-advanced-regression-techniques. I want you to do the following.

*Descriptive and Inferential Statistics.* Provide univariate descriptive statistics and appropriate plots for the training data set. Provide a scatterplot matrix for at least two of the independent variables and the dependent variable. Derive a correlation matrix for any three quantitative variables in the dataset. Test the hypotheses that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval. Discuss the meaning of your analysis. Would you be worried about familywise error? Why or why not?

*Linear Algebra and Correlation.* Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.) Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix. Conduct LU decomposition on the matrix.

*Calculus-Based Probability & Statistics.* Many times, it makes sense to fit a closed form distribution to data. Select a variable in the Kaggle.com training dataset that is skewed to the right, shift it so that the minimum value is absolutely above zero if necessary. Then load the MASS package and run
fitdistr to fit an exponential probability density function. (See
https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/fitdistr.html). Find the optimal value of $\lambda$ for this distribution, and then take 1000 samples from this exponential distribution using this value (e.g., rexp(1000, $\lambda$)). Plot a histogram and compare it with a histogram of your original variable. Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF). Also generate a 95% confidence interval from the empirical data, assuming normality. Finally, provide the empirical 5th percentile and 95th percentile of the data. Discuss.

*Modeling.* Build some type of multiple regression model and submit your model to the competition board. Provide your complete model summary and results with analysis. Report your Kaggle.com user name and score.

### 2019-05-16 DATA605/MSilva_Final.Rmd
More Work on DATA605 Final

@@
 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(kableExtra)
+```
+
 ## Problem 1
@@
 
-```{r}
+```{r, comment=NA}
 set.seed(12345)
 N <- 10
-X <- runif(10000, min=1, max=N)
+n <- 10000
 mu <- sigma <- (N + 1)/2
-Y <- rnorm(10000, mean=mu, sd=sigma)
+df <- data.frame(X = runif(n, min=1, max=N), 
+                 Y = rnorm(n, mean=mu, sd=sigma))
+summary(df$X)
+summary(df$Y)
 ```
@@
 
-a. P(X>x | X>y)
+```{r}
+x <- median(df$X)
+y <- as.numeric(quantile(df$Y)["25%"])
+```
 
-b. P(X>x, Y>y)
+small letter "x" is `r x` and small letter "y" is `r y`.
 
-c. P(X<x | X>y)
+a. $P(X>x | X>y)$
+
+```{r}
+P_a_and_b <- df %>%
+  filter(X > x,
+         X > y) %>%
+  nrow() / n
+
+P_b <- df %>%
+  filter(X > y) %>%
+  nrow() / n
+
+problem_1a <- P_a_and_b / P_b
+```
+
+**The probablity of a random number uniformly ranging from 1 to `r N` being greater than `r x` given that it is greater than `r y` is `r problem_1a`**
+
+b. $P(X>x, Y>y)$
+
+```{r}
+problem_1b <- df %>%
+  filter(X > x,
+         Y > y) %>%
+  nrow() / n
+```
+
+**The probablity of a random number uniformly ranging from 1 to `r N` being greater than `r x` and a random normally distributed number with a mean and standard deviation of `r mu` being greater than `r y` is `r problem_1b`**
+
+c. $P(X<x | X>y)$
+
+```{r}
+P_a_and_b <- df %>%
+  filter(X < x,
+         X > y) %>%
+  nrow() / n
+
+P_b <- df %>%
+  filter(X > y) %>%
+  nrow() / n
+
+problem_1c <- P_a_and_b / P_b
+```
+
+**The probablity of a random number uniformly ranging from 1 to `r N` being less than `r x` given that it is greater than `r y` is `r problem_1c`**
 
@@
 
-Check to see if independence holds by using Fisher’s Exact Test and the Chi Square Test. What is the difference between the two? Which is most appropriate?
+```{r}
+# Create Joint Probabilities
+temp <- df %>%
+  mutate(A = ifelse(X > x, " X greater than x", " X not greater than x"),
+         B = ifelse(Y > y, " Y greater than y", " Y not greater than y")) %>%
+  group_by(A, B) %>%
+  summarise(count = n()) %>%
+  mutate(probability = count / n)
+
+# Create Marginal Probabilities
+temp <- temp %>%
+  ungroup() %>%
+  group_by(A) %>%
+  summarise(count = sum(count),
+            probability = sum(probability)) %>%
+  mutate(B = "Total") %>%
+  bind_rows(temp)
+
+temp <- temp %>%
+  ungroup() %>%
+  group_by(B) %>%
+  summarise(count = sum(count),
+            probability = sum(probability)) %>%
+  mutate(A = "Total") %>%
+  bind_rows(temp)
+
+# Create Table
+temp %>%
+  select(-count) %>%
+  spread(A, probability) %>%
+  rename(" " = B) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**P(X>x and Y>y) is `r temp[temp$A == " X greater than x" & temp$B == " Y greater than y",]$probability`.
+P(X>x)P(Y>y) is `r temp[temp$A == " X greater than x" & temp$B == "Total",]$probability` $\times$ `r temp[temp$A == "Total" & temp$B == " Y greater than y",]$probability` which is `r temp[temp$A == " X greater than x" & temp$B == "Total",]$probability * temp[temp$A == "Total" & temp$B == " Y greater than y",]$probability`.
+They are not the same.**
+
+Check to see if independence holds by using Fisher's Exact Test and the Chi Square Test. What is the difference between the two? Which is most appropriate?
+
+```{r, comment=NA}
+count_data <- temp %>%
+  filter(A != "Total",
+         B != "Total") %>%
+  select(-probability) %>%
+  spread(A, count) %>%
+  as.data.frame()
+
+row.names(count_data) <- count_data$B
+
+count_data <- count_data %>%
+  select(-B) %>%
+  as.matrix()
+
+fisher.test(count_data)
+```
+
+```{r, comment=NA}
+chisq.test(count_data)
+```
+
+**Fisher's Exact Test is for is used when you have small cell sizes (less than 5).  The Chi Square Test is used when the cell sizes are large.  It would be appropriate in this case.**
 

### 2019-05-18 DATA605/MSilva_Final.Rmd
Work on DATA605 Final
More work on 605 final

@@
 ---
-title: "DATA 605 Final"
+title: "CUNY SPS DATA 605 Final"
 author: "Mike Silva"
-output: html_document
+date: "`r Sys.Date()`"
+output:
+  prettydoc::html_pretty:
+    theme: architect
+    highlight: github
 ---
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
+#output: html_document
+#output:
+#  rmdformats::html_clean:
+#    highlight: kate
+library(knitr)
+#library(rmdformats)
 library(dplyr)
@@
 
-```{r, comment=NA}
+```{r, echo=TRUE, eval=TRUE, comment=NA}
 set.seed(12345)
@@
 
-*Probability.* Calculate as a minimum the below probabilities a through c.
+### Probability
+
+Calculate as a minimum the below probabilities a through c.
 Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable.
@@
 
-```{r}
+```{r echo=TRUE, eval=TRUE}
 x <- median(df$X)
@@
 
-```{r}
+```{r echo=TRUE, eval=TRUE}
 P_a_and_b <- df %>%
@@
 
-```{r}
+```{r echo=TRUE, eval=TRUE}
 problem_1b <- df %>%
@@
 
-```{r}
+```{r echo=TRUE, eval=TRUE}
 # Create Joint Probabilities
@@
 
-```{r, comment=NA}
+```{r, echo=TRUE, eval=TRUE, comment=NA}
 count_data <- temp %>%
@@
 
-```{r, comment=NA}
+```{r, echo=TRUE, eval=TRUE, comment=NA}
 chisq.test(count_data)
@@
 
-*Descriptive and Inferential Statistics.* Provide univariate descriptive statistics and appropriate plots for the training data set.
+```{r, echo=TRUE, eval=TRUE}
+kaggle <- read.csv("Kaggle_House_Prices/train.csv") %>%
+  mutate(MSSubClass = as.factor(MSSubClass))
+```
+
+### Descriptive and Inferential Statistics
+
+Provide univariate descriptive statistics and appropriate plots for the training data set.
+
+```{r echo=FALSE, eval=TRUE, results='asis', warning=FALSE}
+for (variable in  names(kaggle)){
+  cat(paste("<h4>", variable, "</h4>"))
+
+  d <- kaggle[,names(kaggle) == variable]
+
+  if(is.factor(d)){
+    p <- ggplot(kaggle, aes_string(variable)) + 
+      geom_bar() + 
+      theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+    d_table <- kaggle %>%
+      group_by(.dots = variable) %>%
+      summarise(count = n()) %>%
+      kable() %>%
+      kable_styling()
+  } else {
+    p <- ggplot(kaggle, aes_string(variable)) + 
+      geom_histogram(binwidth = 10)
+    q <- quantile(d, na.rm = TRUE)
+    d_table <- data.frame(Min = min(d, na.rm = TRUE),
+                          First = q[2],
+                          Median = median(d, na.rm = TRUE), 
+                          Mean = mean(d, na.rm = TRUE),
+                          Third = q[4],
+                          Max = max(d, na.rm = TRUE)) %>%
+      rename("First Q" = First,
+             "Third Q" = Third) %>%
+      gather(Measure, Value) %>%
+      kable() %>%
+      kable_styling()
+  }
+  print(p)
+  print(d_table)
+}
+```
+
 Provide a scatterplot matrix for at least two of the independent variables and the dependent variable.
+
+```{r, echo=TRUE}
+kaggle %>%
+  select(TotalBsmtSF, GrLivArea, YearBuilt, SalePrice) %>%
+  pairs()
+```
+
 Derive a correlation matrix for any three quantitative variables in the dataset.
-Test the hypotheses that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval.
+
+```{r, echo=TRUE}
+kaggle %>%
+  select(TotalBsmtSF, GrLivArea, YearBuilt) %>%
+  cor() %>%
+  kable() %>%
+  kable_styling()
+```
+
+Test the hypothesis that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval.
 Discuss the meaning of your analysis.
@@
 
-*Linear Algebra and Correlation.* Invert your correlation matrix from above.
+```{r}
+d_table <- data.frame(Min = min(d, na.rm = TRUE),
+                          "First Qu." = q[2],
+                          Median = median(d, na.rm = TRUE), 
+                          Mean = mean(d, na.rm = TRUE),
+                          "Third Qu." = q[4],
+                          Max = max(d, na.rm = TRUE)) %>%
+  gather(Measure, Value)
+
+```
+
+### Linear Algebra and Correlation
+
+Invert your correlation matrix from above.
 (This is known as the precision matrix and contains variance inflation factors on the diagonal.) Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix.
@@
 
-*Calculus-Based Probability & Statistics.* Many times, it makes sense to fit a closed form distribution to data.
+### Calculus-Based Probability & Statistics
+
+Many times, it makes sense to fit a closed form distribution to data.
 Select a variable in the Kaggle.com training dataset that is skewed to the right, shift it so that the minimum value is absolutely above zero if necessary.
-Then load the MASS package and run
-fitdistr to fit an exponential probability density function. (See
-https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/fitdistr.html).
+Then load the MASS package and run fitdistr to fit an exponential probability density function.
+(See https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/fitdistr.html).
 Find the optimal value of $\lambda$ for this distribution, and then take 1000 samples from this exponential distribution using this value (e.g., rexp(1000, $\lambda$)).
@@
 
-*Modeling.* Build some type of multiple regression model and submit your model to the competition board.
-Provide your complete model summary and results with analysis.
-Report your Kaggle.com user name and score.
+### Modeling
+
+Build some type of multiple regression model and submit your model to the competition board. Provide your complete model summary and results with analysis. Report your Kaggle.com user name and score.
+
+(http://jse.amstat.org/v19n3/decock.pdf)

### 2019-05-21 DATA605/MSilva_Final.Rmd
More work on the final

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-#output: html_document
-#output:
-#  rmdformats::html_clean:
-#    highlight: kate
+library(MASS)
+library(Matrix)
 library(knitr)
-#library(rmdformats)
+library(matlib)
 library(dplyr)
@@
 kaggle <- read.csv("Kaggle_House_Prices/train.csv") %>%
-  mutate(MSSubClass = as.factor(MSSubClass))
+  filter(GrLivArea < 4000) %>% # Removing outliers per http://jse.amstat.org/v19n3/decock.pdf
+  # Filling in missing values with zeros where it makes sense
+  mutate(BedroomAbvGr = replace_na(BedroomAbvGr, 0),
+         BsmtFullBath = replace_na(BsmtFullBath, 0),
+         BsmtHalfBath = replace_na(BsmtHalfBath, 0),
+         BsmtUnfSF = replace_na(BsmtUnfSF, 0),
+         EnclosedPorch = replace_na(EnclosedPorch, 0),
+         Fireplaces = replace_na(Fireplaces, 0),
+         GarageArea = replace_na(GarageArea, 0),
+         GarageCars = replace_na(GarageCars, 0),
+         HalfBath = replace_na(HalfBath, 0),
+         KitchenAbvGr = replace_na(KitchenAbvGr, 0),
+         LotFrontage = replace_na(LotFrontage, 0),
+         OpenPorchSF = replace_na(OpenPorchSF, 0),
+         PoolArea = replace_na(PoolArea, 0),
+         ScreenPorch = replace_na(ScreenPorch, 0),
+         TotRmsAbvGrd = replace_na(TotRmsAbvGrd, 0),
+         WoodDeckSF = replace_na(WoodDeckSF, 0))
+```
+
+There are some numerical features that are really categories that will need to be cleaned up
+
+```{r}
+clean_MSSubClass <- function(x){
+  return(paste0("SC", as.character(x)))
+}
+
+clean_MoSold <- function(x){
+  if(x > 9){return(as.character(x))}
+  else{
+    return(paste0("0", as.character(x)))
+  }
+}
+
+kaggle <- kaggle %>%
+  rowwise() %>%
+  mutate(MSSubClass = clean_MSSubClass(MSSubClass)) %>%
+  mutate(MoSold = clean_MoSold(MoSold)) %>%
+  ungroup() %>%
+  mutate(MSSubClass = as.factor(MSSubClass),
+         MoSold = as.factor(MoSold)) %>%
+  as.data.frame()
 ```
@@
     p <- ggplot(kaggle, aes_string(variable)) + 
-      geom_histogram(binwidth = 10)
+      geom_density()
     q <- quantile(d, na.rm = TRUE)
@@
                           Max = max(d, na.rm = TRUE)) %>%
-      rename("First Q" = First,
-             "Third Q" = Third) %>%
+      rename("First Quartile" = First,
+             "Third Quartile" = Third) %>%
       gather(Measure, Value) %>%
@@
 ```{r, echo=TRUE}
-kaggle %>%
+correlation_matrix <- kaggle %>%
   select(TotalBsmtSF, GrLivArea, YearBuilt) %>%
   cor() %>%
+  as.matrix()
+
+correlation_matrix %>%
   kable() %>%
@@
 Test the hypothesis that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval.
-Discuss the meaning of your analysis.
-Would you be worried about familywise error?
-Why or why not?
 
-```{r}
-d_table <- data.frame(Min = min(d, na.rm = TRUE),
-                          "First Qu." = q[2],
-                          Median = median(d, na.rm = TRUE), 
-                          Mean = mean(d, na.rm = TRUE),
-                          "Third Qu." = q[4],
-                          Max = max(d, na.rm = TRUE)) %>%
-  gather(Measure, Value)
+```{r, results="asis", tidy=TRUE}
+zero_vars <- c()
 
+variables <- kaggle %>%
+  select(-SalePrice) %>%
+  names()
+
+cat("<table><thead><tr><th>Variable</th><th>Correlation</th><th>Equal to Zero</th></tr></thread><tbody>")
+
+for(variable in variables){
+  d <- kaggle[,names(kaggle) == variable]
+  if(is.numeric(d)){
+    results <- cor.test(kaggle$SalePrice, d, conf.level = 0.8)
+    if(0 > results$conf.int[1] & results$conf.int[2] > 0){
+      hypothesis_test_results <- "Yes"
+      zero_vars <- c(zero_vars, variable)
+    } else {
+      hypothesis_test_results <- "No"
+    }
+    cat(paste("<tr><td>",variable,"</td><td>", cor(kaggle$SalePrice, d), "</td><td>", hypothesis_test_results, "</td>"))
+  }
+}
+cat("</tbody></table>")
 ```
+
+Discuss the meaning of your analysis. Would you be worried about familywise error? Why or why not?
+
+**There are `r length(zero_vars)` variables that have no correlation with the sale price.
+I would be worried be woried about family-wise error because I just ran `r length(variable)` hypothesis tests.
+The family-wise error rate would be: ${FWER} = 1 - (1 - .2)^`r length(zero_vars)` = 1 - `r .8 ^ length(zero_vars)` = `r 1 - (.8^ length(zero_vars))`$.
+This means there is a high probability of committing a Type I error.**
 
@@
 
-Invert your correlation matrix from above.
-(This is known as the precision matrix and contains variance inflation factors on the diagonal.) Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix.
+Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.)
+
+```{r}
+precision_matrix <- inv(correlation_matrix)
+
+precision_matrix %>%
+  kable() %>%
+  kable_styling()
+```
+
+Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix.
+
+**This will generate the identity matrix.**
+
+```{r}
+correlation_matrix %*% precision_matrix %>%
+  round() %>%
+  kable() %>%
+  kable_styling()
+```
+
+```{r}
+precision_matrix %*% correlation_matrix %>%
+  round() %>%
+  kable() %>%
+  kable_styling()
+```
+
 Conduct LU decomposition on the matrix.
+
+```{r}
+A <- Matrix::expand(lu(correlation_matrix))
+
+A$L %>%
+  as.matrix() %>%
+  kable() %>%
+  kable_styling()
+
+A$U %>%
+  as.matrix() %>%
+  kable() %>%
+  kable_styling()
+```
 

### 2019-05-22 DATA605/MSilva_Final.Rmd
More work on the final

@@
 title: "CUNY SPS DATA 605 Final"
-author: "Mike Silva"
+author: "by Mike Silva"
 date: "`r Sys.Date()`"
@@
 
+**NOTE:** The instructions/questions are in **bold**. My responses and comments are in *italics*.
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 library(Matrix)
-library(knitr)
 library(matlib)
@@
 
-Using R, generate a random variable X that has 10,000 random uniform numbers from 1 to N, where N can be any number of your choosing greater than or equal to 6.
-Then generate a random variable Y that has 10,000 random normal numbers with a mean of $\mu=\sigma=\frac{N+1}{2}$.
+**Using R, generate a random variable X that has 10,000 random uniform numbers from 1 to N, where N can be any number of your choosing greater than or equal to 6.
+Then generate a random variable Y that has 10,000 random normal numbers with a mean of $\mu=\sigma=\frac{N+1}{2}$.**
 
@@
 
-Calculate as a minimum the below probabilities a through c.
+**Calculate as a minimum the below probabilities a through c.
 Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable.
-Interpret the meaning of all probabilities.
+Interpret the meaning of all probabilities.**
 
@@
 
-small letter "x" is `r x` and small letter "y" is `r y`.
+*small letter "x" is `r x` and small letter "y" is `r y`.*
 
@@
 
-**The probablity of a random number uniformly ranging from 1 to `r N` being greater than `r x` given that it is greater than `r y` is `r problem_1a`**
+*The probablity of a random number uniformly ranging from 1 to `r N` being greater than `r x` given that it is greater than `r y` is `r problem_1a`*
 
@@
 
-**The probablity of a random number uniformly ranging from 1 to `r N` being greater than `r x` and a random normally distributed number with a mean and standard deviation of `r mu` being greater than `r y` is `r problem_1b`**
+*The probablity of a random number uniformly ranging from 1 to `r N` being greater than `r x` and a random normally distributed number with a mean and standard deviation of `r mu` being greater than `r y` is `r problem_1b`*
 
@@
 
-**The probablity of a random number uniformly ranging from 1 to `r N` being less than `r x` given that it is greater than `r y` is `r problem_1c`**
+*The probablity of a random number uniformly ranging from 1 to `r N` being less than `r x` given that it is greater than `r y` is `r problem_1c`*
 
-Investigate whether P(X>x and Y>y)=P(X>x)P(Y>y) by building a table and evaluating the marginal and joint probabilities.
+**Investigate whether P(X>x and Y>y)=P(X>x)P(Y>y) by building a table and evaluating the marginal and joint probabilities.**
 
@@
 
-**P(X>x and Y>y) is `r temp[temp$A == " X greater than x" & temp$B == " Y greater than y",]$probability`.
+*P(X>x and Y>y) is `r temp[temp$A == " X greater than x" & temp$B == " Y greater than y",]$probability`.
 P(X>x)P(Y>y) is `r temp[temp$A == " X greater than x" & temp$B == "Total",]$probability` $\times$ `r temp[temp$A == "Total" & temp$B == " Y greater than y",]$probability` which is `r temp[temp$A == " X greater than x" & temp$B == "Total",]$probability * temp[temp$A == "Total" & temp$B == " Y greater than y",]$probability`.
-They are not the same.**
+They are not the same.*
 
-Check to see if independence holds by using Fisher's Exact Test and the Chi Square Test. What is the difference between the two? Which is most appropriate?
+**Check to see if independence holds by using Fisher's Exact Test and the Chi Square Test. What is the difference between the two? Which is most appropriate?**
 
@@
 
-**Fisher's Exact Test is for is used when you have small cell sizes (less than 5).  The Chi Square Test is used when the cell sizes are large.  It would be appropriate in this case.**
+*Fisher's Exact Test is for is used when you have small cell sizes (less than 5).  The Chi Square Test is used when the cell sizes are large.  It would be appropriate in this case.*
 
@@
 
-You are to register for Kaggle.com (free) and compete in the House Prices: Advanced Regression Techniques competition.
+**You are to register for Kaggle.com (free) and compete in the House Prices: Advanced Regression Techniques competition.
 https://www.kaggle.com/c/house-prices-advanced-regression-techniques.
-I want you to do the following.
+I want you to do the following.**
 
@@
 kaggle <- read.csv("Kaggle_House_Prices/train.csv") %>%
-  filter(GrLivArea < 4000) %>% # Removing outliers per http://jse.amstat.org/v19n3/decock.pdf
+  # Removing outliers per http://jse.amstat.org/v19n3/decock.pdf
+  filter(GrLivArea < 4000) %>% 
   # Filling in missing values with zeros where it makes sense
@@
 
-There are some numerical features that are really categories that will need to be cleaned up
+*There are some numerical features that are really categories that will need to be cleaned up*
 
@@
 
+clean_text_number <- function(x){
+  if(x < 10){return(paste0("0", x))}
+  return(as.character(x))
+}
+
 kaggle <- kaggle %>%
+  # Clean up variables
   rowwise() %>%
@@
   mutate(MoSold = clean_MoSold(MoSold)) %>%
+  mutate(OverallQual = clean_text_number(OverallQual)) %>%
+  mutate(OverallCond = clean_text_number(OverallCond)) %>%
   ungroup() %>%
   mutate(MSSubClass = as.factor(MSSubClass),
-         MoSold = as.factor(MoSold)) %>%
+         MoSold = as.factor(MoSold),
+         OverallQual = as.factor(OverallQual),
+         OverallCond = as.factor(OverallCond)) %>%
   as.data.frame()
@@
 
-Provide univariate descriptive statistics and appropriate plots for the training data set.
+**Provide univariate descriptive statistics and appropriate plots for the training data set.**
 
@@
 
-Provide a scatterplot matrix for at least two of the independent variables and the dependent variable.
+**Provide a scatterplot matrix for at least two of the independent variables and the dependent variable.**
 
@@
 
-Derive a correlation matrix for any three quantitative variables in the dataset.
+**Derive a correlation matrix for any three quantitative variables in the dataset.**
 
@@
 
-Test the hypothesis that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval.
+**Test the hypothesis that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval.**
 
@@
 
-Discuss the meaning of your analysis. Would you be worried about familywise error? Why or why not?
+**Discuss the meaning of your analysis. Would you be worried about familywise error? Why or why not?**
 
-**There are `r length(zero_vars)` variables that have no correlation with the sale price.
+*There are `r length(zero_vars)` variables that have no correlation with the sale price.
 I would be worried be woried about family-wise error because I just ran `r length(variable)` hypothesis tests.
 The family-wise error rate would be: ${FWER} = 1 - (1 - .2)^`r length(zero_vars)` = 1 - `r .8 ^ length(zero_vars)` = `r 1 - (.8^ length(zero_vars))`$.
-This means there is a high probability of committing a Type I error.**
+This means there is a really high probability of committing a Type I error.*
 
@@
 
-Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.)
+**Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.)**
 
@@
 
-Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix.
+**Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix.**
 
-**This will generate the identity matrix.**
+*This will generate the identity matrix.*
 
@@
 
+*This will too.*
+
 ```{r}
@@
 
-Conduct LU decomposition on the matrix.
+**Conduct LU decomposition on the matrix.**
 
 ```{r}
-A <- Matrix::expand(lu(correlation_matrix))
+lu_decomposition <- Matrix::expand(lu(correlation_matrix))
+```
 
-A$L %>%
-  as.matrix() %>%
+*The LU decomposition should yield the correlation matrix after multiplying the two components.*
+
+```{r}
+A <- lu_decomposition$L %*% lu_decomposition$U %>%
+  as.matrix() 
+colnames(A) <- colnames(correlation_matrix)
+rownames(A) <- rownames(correlation_matrix)
+
+A %>%
   kable() %>%
@@
 
-A$U %>%
-  as.matrix() %>%
+correlation_matrix %>%
   kable() %>%
@@
 
-Many times, it makes sense to fit a closed form distribution to data.
+**Many times, it makes sense to fit a closed form distribution to data.
 Select a variable in the Kaggle.com training dataset that is skewed to the right, shift it so that the minimum value is absolutely above zero if necessary.
@@
 (See https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/fitdistr.html).
-Find the optimal value of $\lambda$ for this distribution, and then take 1000 samples from this exponential distribution using this value (e.g., rexp(1000, $\lambda$)).
-Plot a histogram and compare it with a histogram of your original variable.
-Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF).
-Also generate a 95% confidence interval from the empirical data, assuming normality.
-Finally, provide the empirical 5th percentile and 95th percentile of the data.
-Discuss.
+Find the optimal value of $\lambda$ for this distribution, and then take 1000 samples from this exponential distribution using this value (e.g., rexp(1000, $\lambda$)).**
+
+*I will be using the Square Footage of the Unfinished Basement for this part of the exam.
+I know it is right skewed because the mean (`r mean(kaggle$BsmtUnfSF)`) is larger than the median (`r median(kaggle$BsmtUnfSF)`) as shown in the figure below:*
+
+```{r, echo=FALSE}
+ggplot(kaggle, aes(BsmtUnfSF)) + geom_density() + 
+  geom_vline(xintercept = mean(kaggle$BsmtUnfSF), color = "red")+ 
+  geom_vline(xintercept = median(kaggle$BsmtUnfSF), color = "blue") +
+  annotate("text", x = 700, y = 0.0005, label = "Mean", color="red") + 
+  annotate("text", x = 300, y = 0.0005, label = "Median", color="blue")
+```
+
+*The minimum (`r min(kaggle$BsmtUnfSF)`) is not smaller than zero so we don't need to shift the data.
+Remember that the maximum value is `r max(kaggle$BsmtUnfSF)`.
+I have already loaded the MASS package.*
+
+```{r}
+exponential_pdf <- fitdistr(kaggle$BsmtUnfSF, densfun = "exponential")
+```
+
+*The optimal value for $\lambda$ is `r exponential_pdf$estimate`.*
+
+```{r}
+samples <- rexp(10000, rate = exponential_pdf$estimate)
+```
+
+**Plot a histogram and compare it with a histogram of your original variable.**
+
+```{r}
+histogram_data <- data.frame(Value = samples, Data = c("Sample"))
+histogram_data <- data.frame(Value = kaggle$BsmtUnfSF, Data = c("Original")) %>%
+  rbind(histogram_data) %>%
+  mutate(Data = as.factor(Data))
+
+ggplot(histogram_data, aes(Value, fill = Data)) + geom_histogram(bins = 50, alpha = 0.5)
+```
+
+*The original data is not quite exponential, but it's not a bad approximation.  It has a longer tail than the data.*
+
+**Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF).**
+
+```{r}
+
+```
+
+**Also generate a 95% confidence interval from the empirical data, assuming normality.**
+
+```{r, comment=NA}
+mu <- mean(kaggle$BsmtUnfSF)
+s <- sd(kaggle$BsmtUnfSF)
+n <- nrow(kaggle)
+error <- qnorm(0.975) * s / sqrt(n)
+
+ci <- c(mu - error, mu + error)
+names(ci) <- c("5%", "95%")
+ci
+```
+
+*The 95% confidence interval (assuming normality) is `r round(ci[1])` to `r round(ci[2])`.*
+
+**Finally, provide the empirical 5th percentile and 95th percentile of the data. Discuss.**
+
+```{r, comment=NA}
+quantile(kaggle$BsmtUnfSF, c(0.05, 0.95))
+```
+
+*The actual 95% confidence interval is `r quantile(kaggle$BsmtUnfSF, c(0.05))` to `r quantile(kaggle$BsmtUnfSF, c(0.95))`.*
 
@@
 
-Build some type of multiple regression model and submit your model to the competition board. Provide your complete model summary and results with analysis. Report your Kaggle.com user name and score.
+**Build some type of multiple regression model and submit your model to the competition board.
+Provide your complete model summary and results with analysis.
+Report your Kaggle.com user name and score.**
 
-(http://jse.amstat.org/v19n3/decock.pdf)
+*Since there are so many variables I will be using a random forest to detect which variables are the most important and will be using them to build the model.*
+
+```{r}
+library(randomForest)
+library(caret)
+
+rf_data <- kaggle %>%
+  # Dropping variables with a lot of missing values
+  select(-Alley, -MasVnrType, -MasVnrArea, -BsmtQual, -BsmtCond, -BsmtExposure, -BsmtFinType1, -BsmtFinType2, -Electrical, -FireplaceQu, -GarageType, -GarageYrBlt, -GarageFinish, -GarageQual, -GarageCond, -PoolQC, -Fence, -MiscFeature)
+
+ctrl <- trainControl(method = "cv", number = 5)
+rf_fit <- train(SalePrice ~ ., data = rf_data, method = "rf", trControl = ctrl, importance = TRUE, verbose = TRUE)
+```

### 2019-05-23 DATA605/MSilva_Final.Rmd
More work on the final

@@
 ```{r}
-exponential_pdf <- fitdistr(kaggle$BsmtUnfSF, densfun = "exponential")
+lambda <- fitdistr(kaggle$BsmtUnfSF, densfun = "exponential")$estimate
 ```
 
-*The optimal value for $\lambda$ is `r exponential_pdf$estimate`.*
-
+*The optimal value for $\lambda$ is `r lambda`.*
 ```{r}
-samples <- rexp(10000, rate = exponential_pdf$estimate)
+samples <- rexp(10000, rate = lambda)
 ```
@@
 
-*The original data is not quite exponential, but it's not a bad approximation.  It has a longer tail than the data.*
+*The original data is not quite exponential, but it's not a bad approximation.  The exponential distribution data has a longer tail than the basement square footage data.*
 
@@
 
-```{r}
+*The PDF would be $f(x;\lambda) = \lambda e^{-\lambda x}$ where $x \geq 0$ and otherwise zero.*
 
-```
+*The CDF would be $F(x;\lambda) = 1 - e^{-\lambda x}$.  lambda is `r lambda`. To find the 5th percentile we need to solve for x in:*
+
+$$0.05 = 1 - e^{-`r lambda` x}$$
+
+$$-0.95 = - e^{-`r lambda` x}$$
+
+$$-ln(0.95) = `r lambda` x$$
+
+$$x = \frac{-ln(0.95)}{`r lambda`} = `r -log(0.95)/lambda`$$
+*For the 95th percentile we need to solve for x in:*
+
+$$0.95 = 1 - e^{-`r lambda` x}$$
+
+$$-0.05 = - e^{-`r lambda` x}$$
+
+$$-ln(0.05) = `r lambda` x$$
+
+$$x = \frac{-ln(0.05)}{`r lambda`} = `r -log(0.05)/lambda`$$
+
+*So the 5th and 95th percentiles are approximately `r round(-log(0.95)/lambda,0)` and `r round(-log(0.05)/lambda,0)`, respectively.*
 
@@
 
-```{r}
+```{r, cache=TRUE, message=FALSE, warning=FALSE}
 library(randomForest)
 library(caret)
+
+set.seed(12345)
 
@@
 
-ctrl <- trainControl(method = "cv", number = 5)
+ctrl <- trainControl(method = "cv", number = 10)
 rf_fit <- train(SalePrice ~ ., data = rf_data, method = "rf", trControl = ctrl, importance = TRUE, verbose = TRUE)
 ```
+
+*Now we can examine the variable importance*
+
+```{r}
+imp_var <- importance(rf_fit$finalModel) 
+variables <- rownames(imp_var)
+
+imp_var <- imp_var %>%
+  as.data.frame()
+
+names(imp_var) <- c("IncMSE", "IncNodePurity")
+imp_var$Variable <- variables
+
+imp_var %>%
+  select(Variable, IncMSE, IncNodePurity) %>%
+  arrange(-IncMSE) %>%
+  kable() %>%
+  kable_styling()
+```

### 2019-05-24 DATA605/MSilva_Final.Rmd
More work on the final

@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE}
+library(randomForest)
+library(caret)
 library(MASS)
@@
 
-*small letter "x" is `r x` and small letter "y" is `r y`.*
+*Small letter "x" is `r x` and small letter "y" is `r y`.*
 
@@
 
-*There are some numerical features that are really categories that will need to be cleaned up*
-
-```{r}
-clean_MSSubClass <- function(x){
-  return(paste0("SC", as.character(x)))
-}
-
-clean_MoSold <- function(x){
-  if(x > 9){return(as.character(x))}
-  else{
-    return(paste0("0", as.character(x)))
-  }
-}
-
-clean_text_number <- function(x){
-  if(x < 10){return(paste0("0", x))}
-  return(as.character(x))
-}
-
-kaggle <- kaggle %>%
-  # Clean up variables
-  rowwise() %>%
-  mutate(MSSubClass = clean_MSSubClass(MSSubClass)) %>%
-  mutate(MoSold = clean_MoSold(MoSold)) %>%
-  mutate(OverallQual = clean_text_number(OverallQual)) %>%
-  mutate(OverallCond = clean_text_number(OverallCond)) %>%
-  ungroup() %>%
-  mutate(MSSubClass = as.factor(MSSubClass),
-         MoSold = as.factor(MoSold),
-         OverallQual = as.factor(OverallQual),
-         OverallCond = as.factor(OverallCond)) %>%
-  as.data.frame()
-```
-
 ### Descriptive and Inferential Statistics
@@
 
-```{r echo=FALSE, eval=TRUE, results='asis', warning=FALSE}
+```{r stats_and_plots, echo=FALSE, eval=TRUE, results='asis', warning=FALSE}
 for (variable in  names(kaggle)){
@@
 
-```{r, results="asis", tidy=TRUE}
+```{r, results="asis"}
 zero_vars <- c()
@@
 
-cat("<table><thead><tr><th>Variable</th><th>Correlation</th><th>Equal to Zero</th></tr></thread><tbody>")
+cat("<table><thead><tr><th>Variable</th><th>Correlation Equal to Zero</th></tr></thread><tbody>")
 
@@
     }
-    cat(paste("<tr><td>",variable,"</td><td>", cor(kaggle$SalePrice, d), "</td><td>", hypothesis_test_results, "</td>"))
+    cat(paste("<tr><td>",variable,"</td><td>", hypothesis_test_results, "</td>"))
   }
@@
 
-*The LU decomposition should yield the correlation matrix after multiplying the two components.*
+*The LU decomposition should yield the correlation matrix after multiplying the two components.  In other words this:*
 
@@
   kable_styling()
+```
 
+*Should match this:*
+
+```{r}
 correlation_matrix %>%
@@
 ```
+
+*Which it does.*
 
@@
 *The optimal value for $\lambda$ is `r lambda`.*
+
 ```{r}
@@
 
-```{r}
+```{r, echo=FALSE}
 histogram_data <- data.frame(Value = samples, Data = c("Sample"))
@@
 
-*The 95% confidence interval (assuming normality) is `r round(ci[1])` to `r round(ci[2])`.*
+*The 95% confidence interval (assuming normality) is `r round(ci[1])` to `r round(ci[2])`.
+Although what I think was meant was assume the data is normally distributed.
+Calculate the 5th and 95th percentile.
+That is found using $x = \mu + Z \sigma$.
+The Z is -1.645 for the 5th percentile, and 1.645 for the 95th.
+So the percentiles would be `r round(mu + (-1.645 * s), 0)` and `r round(mu + (1.645 * s), 0)`.*
 
@@
 
-*The actual 95% confidence interval is `r quantile(kaggle$BsmtUnfSF, c(0.05))` to `r quantile(kaggle$BsmtUnfSF, c(0.95))`.*
+*The actual 5th percentile is `r quantile(kaggle$BsmtUnfSF, c(0.05))` and the 95th is `r quantile(kaggle$BsmtUnfSF, c(0.95))`.  So the findings are summarized in the following table:*
+
+| Method | 5% | 95% |
+| ------ |:--:|:---:|
+| Exponential CDF | `r round(-log(0.95)/lambda,0)` | `r round(-log(0.05)/lambda,0)` |
+| Normal 95% CI | `r round(ci[1])`      | `r round(ci[2])` |
+| Normal Percentiles | `r round(mu + (-1.645 * s), 0)`  | `r round(mu + (1.645 * s), 0)` | 
+| Emperical Percentiles | `r round(quantile(kaggle$BsmtUnfSF, c(0.05)),0)`       | `r round(quantile(kaggle$BsmtUnfSF, c(0.95)),0)` |
+
+*If we model the data as exponentially distributed the 5th percentile is `r round(-log(0.95)/lambda,0)`.
+If we model it as normally distributed the 5th is at `r round(mu + (-1.645 * s), 0)` which in the context of square footage does not make any sense.
+The actual 5th percentile is `r round(quantile(kaggle$BsmtUnfSF, c(0.05)),0)`.
+The difference is explained to the assumed shape/distibution of the underlying data.*
+
+*Looking at the 95th percentile we have `r round(-log(0.5)/lambda,0)` if the data are exponentially distributed, `r round(mu + (1.645 * s), 0)` if it is normally distributed and `r round(quantile(kaggle$BsmtUnfSF, c(0.95)),0)` in reality.
+Again the difference is due to the assumed shape.*
+
+*I have left out the 95% CI from the discussion because the confidence interval is a way of estimating the mean of the population.
+We would know if we took 100 estimates that the actual mean falls within the confidence intervals 95% of the time.
+Within this context it is meaningless.*
 
@@
 
-*Since there are so many variables I will be using a random forest to detect which variables are the most important and will be using them to build the model.*
+*Since there are so many variables I will be using a random forest to detect which variables are the most important and will be using them to build the model.
+First I will drop out the variables that are missing a lot of data or otherwise meaningless.*
 
-```{r, cache=TRUE, message=FALSE, warning=FALSE}
-library(randomForest)
-library(caret)
-
+```{r rf, cache=TRUE, message=FALSE, warning=FALSE}
 set.seed(12345)
@@
   # Dropping variables with a lot of missing values
-  select(-Alley, -MasVnrType, -MasVnrArea, -BsmtQual, -BsmtCond, -BsmtExposure, -BsmtFinType1, -BsmtFinType2, -Electrical, -FireplaceQu, -GarageType, -GarageYrBlt, -GarageFinish, -GarageQual, -GarageCond, -PoolQC, -Fence, -MiscFeature)
+  select(-Alley, -MasVnrType, -MasVnrArea, -BsmtQual, -BsmtCond, -BsmtExposure, -BsmtFinType1, -BsmtFinType2, -Electrical, -FireplaceQu, -GarageType, -GarageYrBlt, -GarageFinish, -GarageQual, -GarageCond, -PoolQC, -Fence, -MiscFeature) %>%
+  # Dropping the ID because it is meaningless
+  select(-Id)
 
-ctrl <- trainControl(method = "cv", number = 10)
-rf_fit <- train(SalePrice ~ ., data = rf_data, method = "rf", trControl = ctrl, importance = TRUE, verbose = TRUE)
+ctrl <- trainControl(method = "cv", number = 5)
+rf_fit <- train(SalePrice ~ ., 
+                data = rf_data, 
+                method = "rf", 
+                ntree = 1000, 
+                trControl = ctrl, 
+                tuneGrid = data.frame(mtry = 10), 
+                importance = TRUE, 
+                verbose = TRUE)
 ```
 
-*Now we can examine the variable importance*
+*Now we can examine the variables to in order of importance:*
 
-```{r}
+```{r rf_importance}
 imp_var <- importance(rf_fit$finalModel) 
@@
 ```
+
+*The takeaways from the preceeding table is that size seems to consistently rise to the top as a variable to include in the model.
+I don't want to have multiple variables that are measuring the same thing (i.e.
+Total SF of Living Area, Total SF of Basement, etc.).
+I will use this table to select the variables that I think should be included in the linear regression model.*
+
+```{r model, comment=NA}
+fit <- lm(SalePrice ~ GrLivArea + GarageArea + LotArea + YearRemodAdd + Fireplaces + YearBuilt + OverallQual, kaggle)
+summary(fit)
+```
+
+*These seven variables explain 80% of the variability in the sale price.  I will use this model to make the Kaggle submission.*
+
+```{r predict, comment=NA}
+test <- read.csv("Kaggle_House_Prices/test.csv")
+y_hat <- predict(fit, test)
+submission <- data.frame(Id = test$Id, SalePrice = y_hat)
+summary(submission)
+```
+
+*It appears there is an NA in the submission.  In the interest of getting my final turned in on time I will be using the median to plug that hole.*
+
+```{r}
+fill <- median(submission$SalePrice, na.rm = TRUE)
+submission <- submission %>%
+  mutate(SalePrice = replace_na(SalePrice, fill))
+write.csv(submission, "submission.csv", quote = FALSE, row.names =  FALSE)
+```
+
+*My Kaggle score is* **0.47768** *and my username is* **mikesilva***.  I am ranked 4520 on the leaderboard.  I will try to improve this score when I have time.*
+
+*This analysis is available on  [GitHub](https://github.com/mikeasilva/CUNY-SPS/blob/master/DATA605/MSilva_Final.Rmd).*

### 2019-05-24 DATA605/MSilva_Final.Rmd
More work on the final

@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
 library(randomForest)
@@
   # Removing outliers per http://jse.amstat.org/v19n3/decock.pdf
-  filter(GrLivArea < 4000) %>% 
+  filter(GrLivArea < 4000)
+
+fill_holes <- function(df){
+  df %>%
     # Filling in missing values with zeros where it makes sense
@@
           WoodDeckSF = replace_na(WoodDeckSF, 0)) 
+}
+
+kaggle <- fill_holes(kaggle)
 ```
@@
 ```{r predict, comment=NA}
-test <- read.csv("Kaggle_House_Prices/test.csv")
+test <- read.csv("Kaggle_House_Prices/test.csv") %>%
+  fill_holes()
 y_hat <- predict(fit, test)
@@
 
-*It appears there is an NA in the submission.  In the interest of getting my final turned in on time I will be using the median to plug that hole.*
+*It appears that some of the predictions are incorrect since the minimum is negative.*
 
 ```{r}
-fill <- median(submission$SalePrice, na.rm = TRUE)
-submission <- submission %>%
-  mutate(SalePrice = replace_na(SalePrice, fill))
 write.csv(submission, "submission.csv", quote = FALSE, row.names =  FALSE)
@@
 
-*My Kaggle score is* **0.47768** *and my username is* **mikesilva***.  I am ranked 4520 on the leaderboard.  I will try to improve this score when I have time.*
+*My Kaggle score is* **0.47769** *and my username is* **mikesilva***.  I am ranked 4520 on the leaderboard.  I will try to improve this score when I have time.*
 
-*This analysis is available on  [GitHub](https://github.com/mikeasilva/CUNY-SPS/blob/master/DATA605/MSilva_Final.Rmd).*
+*This analysis is available on  [https://github.com/mikeasilva/CUNY-SPS](https://github.com/mikeasilva/CUNY-SPS/blob/master/DATA605/MSilva_Final.Rmd).*

### 2019-05-24 DATA605/MSilva_Final.Rmd
More work on final

@@
 ```{r, results="asis"}
-zero_vars <- c()
+zero_vars <- 0
+test <- 0
 
@@
   if(is.numeric(d)){
+    test <- test + 1
     results <- cor.test(kaggle$SalePrice, d, conf.level = 0.8)
@@
       hypothesis_test_results <- "Yes"
-      zero_vars <- c(zero_vars, variable)
+      zero_vars <- zero_vars + 1
     } else {
@@
 
-*There are `r length(zero_vars)` variables that have no correlation with the sale price.
-I would be worried be woried about family-wise error because I just ran `r length(variable)` hypothesis tests.
-The family-wise error rate would be: ${FWER} = 1 - (1 - .2)^`r length(zero_vars)` = 1 - `r .8 ^ length(zero_vars)` = `r 1 - (.8^ length(zero_vars))`$.
+*There are `r zero_vars` variables that have no correlation with the sale price.
+I would be worried be woried about family-wise error because I just ran `r test` hypothesis tests.
+The family-wise error rate would be: ${FWER} = 1 - (1 - .2)^`r zero_vars` = 1 - `r .8 ^ zero_vars` = `r 1 - (.8^ zero_vars)`$.
 This means there is a really high probability of committing a Type I error.*
@@
 ```{r}
-samples <- rexp(10000, rate = lambda)
+samples <- rexp(1000, rate = lambda)
 ```
@@
 
-*This analysis is available on  [https://github.com/mikeasilva/CUNY-SPS](https://github.com/mikeasilva/CUNY-SPS/blob/master/DATA605/MSilva_Final.Rmd).*
+*This analysis is available on  [https://github.com/mikeasilva/CUNY-SPS](https://github.com/mikeasilva/CUNY-SPS/blob/master/DATA605/MSilva_Final.Rmd).
+A walkthough of this analysis is available on [YouTube](https://youtu.be/W5iE-ItRvrQ).*

### 2019-08-30 DATA608/hw1.rmd
Work on DATA608 hw1

---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```
```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

*First let's start with some aggregate counts of the data:*

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(kableExtra)

temp <- inc %>%
  select(Industry) %>%
  unique() %>%
  tally() %>%
  rename(Industries = n)

temp <- inc %>%
  select(City) %>%
  unique() %>%
  tally() %>%
  rename(Cities = n) %>%
  bind_cols(temp)

inc %>%
  select(State) %>%
  unique() %>%
  tally() %>%
  rename(States = n) %>%
  bind_cols(temp) %>%
  kable() %>%
  kable_styling()
```

*Hmmm.  There are 52 states in the data set.  I'm guessing there is DC but I wonder if there is an error in the data.  I will take a closer look at this:*

```{r, comment=NA}
unique(inc$State)
```

*Ahh!  There's Puerto Rico.  My hunch about DC was correct.  No need to clean that up.  Great!*

*I am wondering how some of the variables differ the industry.  I will look at the size of the work force and the revenue.  I will also construct an output per worker to see which industry is the most productive in economic terms:*

```{r}
inc %>%
  group_by(Industry) %>%
  mutate(`Output per Worker` = Revenue / Employees) %>%
  summarise(`Avg Employees` = round(mean(Employees, na.rm = TRUE), 0),
            `Avg Revenue` = mean(Revenue, na.rm = TRUE),
            `Avg Output per Worker` = round(mean(`Output per Worker`, na.rm = TRUE), 0)) %>%
  arrange(desc(`Avg Output per Worker`)) %>%
  kable() %>%
  kable_styling()
```

*This is not supprising that the top industries are the most "productive."  They are very capital intensive.*

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r, fig.height=10}
library(ggplot2)
library(ggthemes)

inc %>%
  group_by(State) %>%
  tally() %>%
  ggplot(aes(x = n, y=reorder(State, n))) +
  geom_point() +
  ggtitle("Number of Fastest Growing Companies by State") +
  ylab(element_blank()) + 
  xlab("Companies") + 
  theme_economist()
```

## Question 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r, warning=FALSE}
inc %>%
  filter(State == "NY") %>%
  select(Industry, Employees) %>%
  na.omit() %>%
  group_by(Industry) %>%
  mutate(Median = median(Employees)) %>%
  ggplot(aes(x = reorder(Industry, Median), y = Employees)) +
  ylim(0, 1000) +
  xlab(element_blank()) + 
  ggtitle("Workforce Size by Industry", subtitle = "(employment over 1,000 not shown)") +
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size = 0.9) +
  coord_flip() +
  theme_economist()
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

*Funny!  I guess I think like this investor.*

```{r, warning=FALSE}
inc %>%
  select(Industry, Revenue, Employees) %>%
  na.omit() %>%
  mutate(`Revenue Per Employee` = Revenue / Employees / 1000000) %>%
  group_by(Industry) %>%
  mutate(Median = median(`Revenue Per Employee`, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(Industry, Median), y = `Revenue Per Employee`)) +
  ylim(0, 1) +
  xlab(element_blank()) + 
  ylab("Millions of Dollars") + 
  ggtitle("Revenue per Employee by Industry", subtitle = "(revenue over $1MM not shown)") +
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size = 0.9) +
  coord_flip() +
  theme_economist()
```

### 2019-08-30 DATA621/Week 1 Textbook Exercises.Rmd
Initial commit of Week 1 Textbook Exercises

---
title: "DATA 612 Week #1 Textbook Exercises"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## LMR 1.1

The dataset *teengam* concers a study of teenage gambline in Britain.  Make a numerical and graphical summary of the data, commenting on any features that you find interesting.  Limit the output you present to a quantity that a busy reader would find sufficeint to get a basic understanding of the data.

## LMR 1.3

The dataset *prostate* is from a study on 97 men with prostate conacer who were due to receive a radical prostatectomy.  Make a numerical and graphical summary of the data as in the first question.

## LMR 1.4

The dataset *sat* comes from a study entitled "Getting What You Pay For: The Debate Overty Equity in Public School Expenditures."  Make a numerical and graphical summary of the data as in the first question.

## LMR 1.5

The dataset *divusa* contains data on divorces in the United States from 1920 to 1996.  Make a numerical and graphical summary of the data as in the first question.

### 2019-08-31 DATA621/Week 1 Textbook Exercises.Rmd
Work on Week 1 Textbook Exercises

@@
 
-```{r knitr_init, echo=FALSE, cache=FALSE}
+```{r knitr_init, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
 library(knitr)
 library(rmdformats)
-
+library(tidyverse)
+library(kableExtra)
 ## Global options
@@
 
-The dataset *teengam* concers a study of teenage gambline in Britain.
+The dataset *teengamb* concers a study of teenage gambline in Britain.
 Make a numerical and graphical summary of the data, commenting on any features that you find interesting.
-Limit the output you present to a quantity that a busy reader would find sufficeint to get a basic understanding of the data.
+Limit the output you present to a quantity that a busy reader would find sufficient to get a basic understanding of the data.
+
+**Estimates show teens in Britain gamble between 0 and 150 pounds annually.  Males gamble more than females.  They gamble on average nearly 8 times as much.**
+
+```{r}
+# A paper describing the original study can be found at http://www.utstat.utoronto.ca/reid/sta2201s/2012/teengamb.pdf.
+First I will read in the data.
+I will recode the sex variable making it easier to interpret.
+The income variable is per week and the gambling is per year.
+I will adjust the income to put it on the same basis.
+
+teengamb <- read.delim("data/teengamb.txt") %>%
+  mutate(sex = as.factor(ifelse(sex == 0, "Male", "Female")),
+         income = income * 52)
+
+ggplot(teengamb, aes(x = sex, y = gamble, color=sex)) + 
+  geom_boxplot() +
+  ylab("Pounds") +
+  xlab(element_blank()) +
+  theme(legend.position = "none") +
+  ggtitle("Estimated Annual Gambling by Sex")
+```
+
+```{r}
+temp <- teengamb %>%
+  group_by(sex) %>%
+  summarise(mpg = list(enframe(quantile(gamble, probs=c(0.25, 0.5, 0.75))))) %>% 
+  unnest() %>%
+  spread(name, value)
+
+teengamb %>%
+  group_by(sex) %>%
+  summarise(`Min.` = min(gamble),
+            Mean = mean(gamble),
+            `Max.` = max(gamble)) %>%
+  merge(temp) %>%
+  select(sex, `Min.`, `25%`, `50%`, Mean, `75%`, `Max.`) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**Teens earned between 31 and 780 pounds annually.  Those who earned more were more likely to gammble more.**
+
+```{r}
+ggplot(teengamb, aes(x = income, y = gamble)) + 
+  geom_point() + 
+  geom_smooth(method = "lm") +
+  xlab("Annual Income (Pounds)") +
+  ylab("Annual Gambling (Pounds)") +
+  ggtitle("Income vs. Gambling")
+```
 
@@
 The dataset *prostate* is from a study on 97 men with prostate conacer who were due to receive a radical prostatectomy.  Make a numerical and graphical summary of the data as in the first question.
+
+```{r}
+#lcavol log(cancer volume)
+#lweight log(prostate weight)
+#age age
+#lbph log(benign prostatic hyperplasia amount)
+#svi seminal vesicle invasion
+#lcp log(capsular penetration)
+#gleason Gleason score
+#pgg45 percentage Gleason scores 4 or 5
+#lpsa log(prostate specific antigen)
+
+prostate <- read.delim("data/prostate.txt") 
+pairs(prostate)
+```
 

### 2019-09-04 DATA621/Week 1 Textbook Exercises.Rmd
More work on Textbook Exercises 1

@@
 library(kableExtra)
+library(gridExtra)
+library(ggthemes)
 ## Global options
@@
 
-ggplot(teengamb, aes(x = sex, y = gamble, color=sex)) + 
+plot1 <- ggplot(teengamb, aes(x = sex, color = sex, fill = sex)) +
+  geom_bar(stat = "count") +
+  ylab("Teens") + 
+  xlab(element_blank()) +
+  theme(legend.position = "none") +
+  ggtitle("Respondents by Sex")
+
+plot2 <- ggplot(teengamb, aes(x = sex, y = gamble, color=sex)) + 
   geom_boxplot() +
@@
   ggtitle("Estimated Annual Gambling by Sex")
+
+grid.arrange(plot1, plot2, ncol = 2)
 ```
@@
             Mean = mean(gamble),
-            `Max.` = max(gamble)) %>%
+            `Max.` = max(gamble), 
+            Count = n()) %>%
   merge(temp) %>%
-  select(sex, `Min.`, `25%`, `50%`, Mean, `75%`, `Max.`) %>%
+  select(sex, `Min.`, `25%`, `50%`, Mean, `75%`, `Max.`, Count) %>%
   kable() %>%
@@
 
-**Teens earned between 31 and 780 pounds annually.  Those who earned more were more likely to gammble more.**
+**Teens earned between 31 and 780 pounds annually.  Those who earned more were more likely to gamble more.**
 
@@
 prostate <- read.delim("data/prostate.txt") 
-pairs(prostate)
+```
+
+**These men are generally older.  Their age ranges from 41 to 79 years.  The median age is 65.**
+
+```{r}
+ggplot(prostate, aes(x=age)) + 
+  geom_histogram(bins=30) +
+  ylab("Men") +
+  xlab(element_blank()) + 
+  ggtitle("Age of Men")
+
+temp <- prostate %>%
+  summarise(mpg = list(enframe(quantile(age, probs=c(0.25, 0.5, 0.75))))) %>% 
+  unnest() %>%
+  spread(name, value)
+
+prostate %>%
+  summarise(`Min.` = min(age),
+            Mean = mean(age),
+            `Max.` = max(age)) %>%
+  merge(temp) %>%
+  select(`Min.`, `25%`, `50%`, Mean, `75%`, `Max.`) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**There is a direct relationship between the prostate cancer volume and the capsular penetration.**
+
+```{r}
+ggplot(prostate, aes(lcp, lcavol)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  xlab("Capsular Penetration (log)") +
+  ylab("Cancer Volume (log)") +
+  ggtitle("Cancer Volume and Capsular Penetration")
+```
+
+```{r}
+temp <- prostate %>%
+  summarise(mpg = list(enframe(quantile(lcp, probs=c(0.25, 0.5, 0.75))))) %>% 
+  unnest() %>%
+  spread(name, value)
+
+prostate %>%
+  summarise(`Min.` = min(lcp),
+            Mean = mean(lcp),
+            `Max.` = max(lcp)) %>%
+  merge(temp) %>%
+  select(`Min.`, `25%`, `50%`, Mean, `75%`, `Max.`) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**The volume of the cancer is also positively correlated with the PSA.**
+
+```{r}
+ggplot(prostate, aes(lpsa, lcavol)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  xlab("PSA (log)") +
+  ylab("Cancer Volume (log)") +
+  ggtitle("Cancer Volume and PSA (Prostate Specific Antigen)")
+```
+
+```{r}
+temp <- prostate %>%
+  summarise(mpg = list(enframe(quantile(lpsa, probs=c(0.25, 0.5, 0.75))))) %>% 
+  unnest() %>%
+  spread(name, value)
+
+prostate %>%
+  summarise(`Min.` = min(lpsa),
+            Mean = mean(lpsa),
+            `Max.` = max(lpsa)) %>%
+  merge(temp) %>%
+  select(`Min.`, `25%`, `50%`, Mean, `75%`, `Max.`) %>%
+  kable() %>%
+  kable_styling()
 ```
@@
 
-The dataset *sat* comes from a study entitled "Getting What You Pay For: The Debate Overty Equity in Public School Expenditures."  Make a numerical and graphical summary of the data as in the first question.
+The dataset *sat* comes from a study entitled "Getting What You Pay For: The Debate Over Equity in Public School Expenditures."  Make a numerical and graphical summary of the data as in the first question.
+
+**SAT scores range from 844 to 1107 with an average arround 966.  These data have considerable variability.**
+
+```{r}
+#  These variables include average annual expenditures per student, the student/teacher ratio, the average annual salary for teachers in each state, and the percentage of all eligible students taking the SAT.
+Finally, the three remaining variables in the dataset measure educational "output" (e.g., performance) using the SAT.
+Verbal and math scores appear separately, and an average total score is given for each state.
+
+sat <- read.delim("data/sat.txt") 
+```
+
+**At first blush it appears that high schools that spend less per pupil do better on the SAT than those who spend more.  Strange!?!**
+
+```{r}
+ggplot(sat, aes(expend, total)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  xlab("1994-95 per Student Spending (thousands of dollars)") +
+  ylab("Avg T SAT Score, 1994-95") +
+  ggtitle("Spend Less for Better SAT Scores")
+```
+
+**But when you examine the popularity of the SAT test a different pattern emerges.
+In certain states the ACT is the college enterance test of choice.
+Thus the SAT scores in these state are only for the students who want to apply to a school back East.**
+
+```{r}
+ggplot(sat, aes(takers, total)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  xlab("Percent of Students Taking the SAT") +
+  ylab("Avg. SAT Score, 1994-95") +
+  ggtitle("SAT Self-Selection Bias")
+```
 
@@
 The dataset *divusa* contains data on divorces in the United States from 1920 to 1996.  Make a numerical and graphical summary of the data as in the first question.
+
+```{r}
+# year the year from 1920-1996
+# divorce divorce per 1000 women aged 15 or more
+# unemployed unemployment rate
+# femlab percent female participation in labor force aged 16+
+# marriage marriages per 1000 unmarried women aged 16+
+# birth births per 1000 women aged 15-44
+# military military personnel per 1000 population
+
+divusa <- read.delim("data/divusa.txt")
+```
+
+**The US divorce rate has generally been increasing over time.  There was a spike in the 1940s and it has been trending down since the 1980.**
+
+```{r}
+ggplot(divusa, aes(year, divorce)) +
+  geom_line() +
+  ylab("Per 1,000 Women") +
+  xlab(element_blank()) +
+  ggtitle("U.S. Divorce Rate Over Time")
+```
+
+```{r}
+divusa <- divusa %>%
+  mutate(Decade = floor(year/10)*10)
+
+temp <- divusa %>%
+  group_by(Decade) %>%
+  summarise(mpg = list(enframe(quantile(divorce, probs=c(0.25, 0.5, 0.75))))) %>% 
+  unnest() %>%
+  spread(name, value)
+
+divusa %>%
+  group_by(Decade) %>%
+  summarise(`Min.` = min(divorce),
+            Mean = mean(divorce),
+            `Max.` = max(divorce), 
+            Count = n()) %>%
+  merge(temp) %>%
+  select(Decade, `Min.`, `25%`, `50%`, Mean, `75%`, `Max.`, Count) %>%
+  kable() %>%
+  kable_styling()
+```

### 2019-09-04 DATA621/Week 1 Textbook Exercises.Rmd
Wrapped up work on week 1 textbook exercises

@@
 sat <- read.delim("data/sat.txt")
+
+temp <- sat %>%
+  summarise(mpg = list(enframe(quantile(total, probs=c(0.25, 0.5, 0.75))))) %>% 
+  unnest() %>%
+  spread(name, value)
+
+sat %>%
+  summarise(`Min.` = min(total),
+            Mean = mean(total),
+            `Max.` = max(total),
+            Count = n()) %>%
+  merge(temp) %>%
+  select(`Min.`, `25%`, `50%`, Mean, `75%`, `Max.`, Count) %>%
+  kable() %>%
+  kable_styling()
 ```
@@
 ```
+
+**There are a couple of things that have a similar spike.  For example military personnel per 1000...**
+
+```{r}
+ggplot(divusa, aes(x = year)) +
+  geom_line(aes(y = divorce, color = "Divorce")) +
+  geom_line(aes(y = military, color = "Military")) + 
+  scale_y_continuous(sec.axis = sec_axis(~.*1, name = "Per 1,000 People"))+ 
+  scale_colour_manual(values = c("blue", "red")) +
+  labs(y = "Per 1,000 Women", x = element_blank(), color = element_blank()) +
+  theme(legend.position = "bottom") +
+  ggtitle("Divorce and Military Personnel Rate")
+```
+
+**...the marriage rate...**
+
+```{r}
+ggplot(divusa, aes(x = year)) +
+  geom_line(aes(y = divorce, color = "Divorce Rate")) +
+  geom_line(aes(y = marriage, color = "Marriage")) +
+  scale_colour_manual(values = c("blue", "red")) +
+  labs(y = "Per 1,000 Women", x = element_blank(), color = element_blank()) +
+  theme(legend.position = "bottom") +
+  ggtitle("Divorce and Marriage Rate")
+```
+
+**...and most strikingly, the female labor force participation rate.**
+
+```{r}
+ggplot(divusa, aes(x = year)) +
+  geom_line(aes(y = divorce, color = "Divorce Rate")) +
+  geom_line(aes(y = femlab, color = "Female Labor Force Participation")) +
+  scale_y_continuous(sec.axis = sec_axis(~.*1, name = "Percent of Working Age Females"))+ 
+  scale_colour_manual(values = c("blue", "red")) +
+  labs(y = "Per 1,000 Women", x = element_blank(), color = element_blank()) +
+  theme(legend.position = "bottom") +
+  ggtitle("Divorce Rate and Female Labor Force Participation")
+```

### 2019-09-04 DATA621/Week 2 Textbook Exercises.Rmd
Initial commit

--- 2019-08-31 DATA621/Week 1 Textbook Exercises.Rmd
@@
 ---
-title: "DATA 612 Week #1 Textbook Exercises"
+title: "DATA 621 Week #2 Textbook Exercises"
 author: "Mike Silva"
@@
 
-## LMR 1.1
+## MARR 2.1
 
-The dataset *teengamb* concers a study of teenage gambline in Britain.
-Make a numerical and graphical summary of the data, commenting on any features that you find interesting.
-Limit the output you present to a quantity that a busy reader would find sufficient to get a basic understanding of the data.
+## MARR 2.2
 
-**Estimates show teens in Britain gamble between 0 and 150 pounds annually.  Males gamble more than females.  They gamble on average nearly 8 times as much.**
+## LMR 2.4
 
-```{r}
-# A paper describing the original study can be found at http://www.utstat.utoronto.ca/reid/sta2201s/2012/teengamb.pdf.
-First I will read in the data.
-I will recode the sex variable making it easier to interpret.
-The income variable is per week and the gambling is per year.
-I will adjust the income to put it on the same basis.
-
-teengamb <- read.delim("data/teengamb.txt") %>%
-  mutate(sex = as.factor(ifelse(sex == 0, "Male", "Female")),
-         income = income * 52)
-
-ggplot(teengamb, aes(x = sex, y = gamble, color=sex)) + 
-  geom_boxplot() +
-  ylab("Pounds") +
-  xlab(element_blank()) +
-  theme(legend.position = "none") +
-  ggtitle("Estimated Annual Gambling by Sex")
-```
-
-```{r}
-temp <- teengamb %>%
-  group_by(sex) %>%
-  summarise(mpg = list(enframe(quantile(gamble, probs=c(0.25, 0.5, 0.75))))) %>% 
-  unnest() %>%
-  spread(name, value)
-
-teengamb %>%
-  group_by(sex) %>%
-  summarise(`Min.` = min(gamble),
-            Mean = mean(gamble),
-            `Max.` = max(gamble)) %>%
-  merge(temp) %>%
-  select(sex, `Min.`, `25%`, `50%`, Mean, `75%`, `Max.`) %>%
-  kable() %>%
-  kable_styling()
-```
-
-**Teens earned between 31 and 780 pounds annually.  Those who earned more were more likely to gammble more.**
-
-```{r}
-ggplot(teengamb, aes(x = income, y = gamble)) + 
-  geom_point() + 
-  geom_smooth(method = "lm") +
-  xlab("Annual Income (Pounds)") +
-  ylab("Annual Gambling (Pounds)") +
-  ggtitle("Income vs. Gambling")
-```
-
-## LMR 1.3
-
-The dataset *prostate* is from a study on 97 men with prostate conacer who were due to receive a radical prostatectomy.  Make a numerical and graphical summary of the data as in the first question.
-
-```{r}
-#lcavol log(cancer volume)
-#lweight log(prostate weight)
-#age age
-#lbph log(benign prostatic hyperplasia amount)
-#svi seminal vesicle invasion
-#lcp log(capsular penetration)
-#gleason Gleason score
-#pgg45 percentage Gleason scores 4 or 5
-#lpsa log(prostate specific antigen)
-
-prostate <- read.delim("data/prostate.txt") 
-pairs(prostate)
-```
-
-## LMR 1.4
-
-The dataset *sat* comes from a study entitled "Getting What You Pay For: The Debate Overty Equity in Public School Expenditures."  Make a numerical and graphical summary of the data as in the first question.
-
-## LMR 1.5
-
-The dataset *divusa* contains data on divorces in the United States from 1920 to 1996.  Make a numerical and graphical summary of the data as in the first question.
+## LMR 2.5

### 2019-09-05 DATA621/Week 2 Textbook Exercises.Rmd
More work on week 2 exercises

@@
 opts_knit$set(width=75)
+
+# A Prefix nulling hook.
+
+# Make sure to keep the default for normal processing.
+default_output_hook <- knitr::knit_hooks$get("output")
+
+# Output hooks handle normal R console output.
+knitr::knit_hooks$set( output = function(x, options) {
+
+  comment <- knitr::opts_current$get("comment")
+  if( is.na(comment) ) comment <- ""
+  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
+                     x, perl = TRUE)
+  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
+  if( can_null && do_null ) {
+    # By default R print output aligns at the right brace.
+    align_index <- regexpr( "\\]", x )[1] - 1
+    # Two cases: start or newline
+    re <- paste0( "^.{", align_index, "}\\]")
+    rep <- comment
+    x <- gsub( re, rep,  x )
+    re <- paste0( "\\\n.{", align_index, "}\\]")
+    rep <- paste0( "\n", comment )
+    x <- gsub( re, rep,  x )
+  }
+
+  default_output_hook( x, options )
+
+})
+
+knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
 ```
@@
 
+The dataset *teengamb* concerns a study of teenage gambling in Britain.
+Fit a regression model with the expenditure on gambling as the response and the sex, status, income and verbal score as predictors.
+Present the output.
+
+```{r, comment=NA}
+teengamb <- read.delim("data/teengamb.txt") %>%
+  mutate(sex = as.factor(ifelse(sex == 0, "Male", "Female")))
+model <- lm(gamble ~ sex + status + income + verbal, data = teengamb)
+summary(model)
+```
+
+A. What percentage of variation in the response is explained by these predictors?
+
+```{r, null_prefix=TRUE}
+summary(model)$r.squared
+```
+
+B. Which observation has the largest (positive) residual?  Give the case number
+
+```{r, null_prefix=TRUE}
+largest_positive_residual <- max(residuals(model))
+
+seq(1, length(residuals(model)))[residuals(model) == largest_positive_residual]
+```
+
+C. Compute the mean and median of the residuals.
+
+```{r, null_prefix=TRUE}
+noquote(paste("Mean:", round(mean(residuals(model)),3)))
+
+noquote(paste("Median:", round(median(residuals(model)),3)))
+```
+
+D. Compute the correlation of the residuals with the fitted values.
+
+```{r, null_prefix=TRUE}
+fitted_vals <- predict(model, teengamb)
+round(cor(fitted_vals, residuals(model)),3)
+```
+
+E. Compute the correlation of the residual with the income
+
+```{r, null_prefix=TRUE}
+round(cor(teengamb$income, residuals(model)),3)
+```
+
+F. For all other predictors held constant, what would be the difference in predicted expenditure on gambling for a male compared to a female?
+
+```{r, null_prefix=TRUE}
+unname(model$coefficients["sexMale"])
+```
+
 ## MARR 2.2
+
+The dataset *uswages* is drawn as a sample from the Current Population Survey in 1988.
+Fit a model with weekly wages as the response and years of education and experience as predictors.
+Report and give a simple interpretation to the regression coefficient for the years of education.
+Now fit the same model but with the logged weekly wages.
+Give an interpretation to the regression coefficeint for the years of education.
+Which interpretation is more natural?
 

### 2019-09-05 DATA608/hw1.rmd
More work on HW1

@@
             `Avg Revenue` = mean(Revenue, na.rm = TRUE),
-            `Avg Output per Worker` = round(mean(`Output per Worker`, na.rm = TRUE), 0)) %>%
+            `Avg Output per Worker` = round(mean(`Output per Worker`, na.rm = TRUE), 0),
+            Count = n()) %>%
   arrange(desc(`Avg Output per Worker`)) %>%
@@
   xlab("Companies") + 
-  theme_economist()
+  theme_clean()
 ```
@@
   ggtitle("Workforce Size by Industry", subtitle = "(employment over 1,000 not shown)") +
-  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size = 0.9) +
+  geom_boxplot(outlier.colour="gray", outlier.shape=8, outlier.size = 0.9) +
   coord_flip() +
-  theme_economist()
+  theme_clean()
 ```
@@
   ggtitle("Revenue per Employee by Industry", subtitle = "(revenue over $1MM not shown)") +
-  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size = 0.9) +
+  geom_boxplot(outlier.colour="gray", outlier.shape=8, outlier.size = 0.9) +
   coord_flip() +
-  theme_economist()
+  theme_clean()
 ```

### 2019-09-06 DATA621/Week 2 Textbook Exercises.Rmd
More work

--- 2019-09-04 DATA621/Week 2 Textbook Exercises.Rmd
@@
 ## MARR 2.1
+
+The web site www.playbill.com provides weekly reports on the box office ticket sales for plays on Broadway in New York.
+We shall consider the data for the week Octover 11-17, 2004 (reffered to below as the current week).
+The data are in the form of the gross box office results for the current week and the gross box office results for the previous wek (i.e., October 3-10, 2004).
+
+Fit the following model to the data: $Y=\beta_0 + \beta_1 x + e$ where $Y$ is the gross box office results for the current week (in $) and $x$ is the gross box office results for the previous week (in $).
+Compleete the following tasks:
+
+```{r, comment=NA}
+playbill <- read.csv("data/playbill.csv")
+# teengamb <- read.delim("data/teengamb.txt") %>%
+#   mutate(sex = as.factor(ifelse(sex == 0, "Male", "Female")))
+fit <- lm(CurrentWeek ~ LastWeek, data = playbill)
+summary(fit)
+```
+
+A. Find a 95% confidence interval for the slope of the regression model, $\beta_1$.  Is 1 a plausible value for $\beta_1$?  Give a reason to support your answer.
+
+```{r, null_prefix=TRUE}
+confint(fit, "LastWeek")
+```
+
+**Yes.  It is plausible because it falls within the range of the 95% confidence interval.**
+
+B. Test the null hypothesis $H_0 : \beta_0 = 10000$ against a two-sided alternative.  Interpret your result.
 

### 2019-09-06 DATA621/Week 2 Textbook Exercises.Rmd
More Work

@@
 
-```{r knitr_init, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
+```{r knitr_init, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
 library(knitr)
@@
 library(tidyverse)
-library(kableExtra)
-library(gridExtra)
-library(ggthemes)
 ## Global options
-options(max.print="75")
-opts_chunk$set(echo=FALSE,
-	             cache=TRUE,
-               prompt=FALSE,
-               tidy=TRUE,
+options(max.print="100")
+opts_chunk$set(tidy=TRUE,
                comment=NA,
@@
                warning=FALSE)
-opts_knit$set(width=75)
-
+opts_knit$set(width=200)
 # A Prefix nulling hook.
@@
 The web site www.playbill.com provides weekly reports on the box office ticket sales for plays on Broadway in New York.
-We shall consider the data for the week Octover 11-17, 2004 (reffered to below as the current week).
+We shall consider the data for the week October 11-17, 2004 (reffered to below as the current week).
 The data are in the form of the gross box office results for the current week and the gross box office results for the previous wek (i.e., October 3-10, 2004).
@@
 Fit the following model to the data: $Y=\beta_0 + \beta_1 x + e$ where $Y$ is the gross box office results for the current week (in $) and $x$ is the gross box office results for the previous week (in $).
-Compleete the following tasks:
 
@@
 playbill <- read.csv("data/playbill.csv")
-# teengamb <- read.delim("data/teengamb.txt") %>%
-#   mutate(sex = as.factor(ifelse(sex == 0, "Male", "Female")))
 fit <- lm(CurrentWeek ~ LastWeek, data = playbill)
-summary(fit)
 ```
+
+Complete the following tasks:
 
@@
 
-```{r, null_prefix=TRUE}
+```{r, comment=NA}
 confint(fit, "LastWeek")
@@
 
+```{r}
+confint(fit, "(Intercept)")
+```
+
+**The 95% confidence interval ranges from -14,244 to 27,854.  Since the null hypothesis falls outside of this range we would reject the null hypothesis.**
+
+C.
+Use the fitted regression model to estimate the gross box office results for the current week (in $) for a production with $400,000 in gross box office the previous week.
+Find a 95% prediction interval for the gross box office results for the current week (in $) for a production with $400,000 in gross box office the previous week.
+
+```{r, null_prefix=TRUE}
+predict(fit, data.frame("LastWeek" = 400000), interval="predict")
+```
+
+Is $450,000 a feasible value for the gross box office results in the current week, for a production with $400,000 in gros box office the previous week?  Give a reason to support your answer.
+
+**No, because it falls outside of the upper bounds of the prediction interval.**
+
+D. Some promoters of Brodway plays use the rediction rult that next weeks gross box office results will be equal to this weeks's gross box office results.  Comment on the appropriateness of this rule.
+
+**Let's examine this model graphically.  We will plot the data and add a red line that represents the model:**
+
+```{r}
+ggplot(playbill, aes(x=LastWeek/1000, y=CurrentWeek/1000)) +
+  geom_point() +
+  labs(title = "Gross Box Office Results",
+       subtitle = "(in thousands of dollars)",
+       x = "Previous Week",
+       y = "Current Week") +
+  geom_abline(slope=1, intercept = 0, color = "red")
+```
+
+**Most of the data lines up with the line.  It is not a bad model.  This also relates to question A.  1 falls in the 95% confidence interval.**
+
 ## MARR 2.2
+
+A story by James R.
+Hagerty entitled *With Buyers Sidelined, Home Prices Slide* published in the Thursday October 25, 2007 edition of the *Wall Street Journal* contained data on so-called fundamental housing indicators in major real estate markets across the US.
+The author argues that...*prices are generally falling and overdue loan payments are pilling up*.
+Thus, we shall consider data presented in the article on
+
+$Y$ = Percentage change in average price from July 2006 to July 2007 (based on the S&P/Case-Shiller national housing index); and
+
+$x$ = Percentage of mortgage loans 30 days or more overdue in latest quarter (based on data from Equifax and Moody's).
+
+Fit the following model to the data: $Y = \beta_0 + \beta_1 x + e$.
+
+```{r}
+indicators <- read.delim("data/indicators.txt")
+fit <- lm(PriceChange ~ LoanPaymentsOverdue, data = indicators)
+```
+
+Complete the following tasks:
+
+A.
+Find a 95% confidence interval for the slope of the regression model, $\beta_1$.
+On the basis of this confidence interval decide whether there is evidence of a significant negative linear association.
+
+```{r}
+confint(fit, "LoanPaymentsOverdue")
+```
+
+**The data suggest that there may be a negative linear association.  However this relationship is not very strong.**
+
+B. Use the fitted regression model to estimate $E(Y|X=4)$.  Find a 95% confidence interval for $E(Y|X=4)$.  Is 0% a feasible value for $E(Y|X=4)$?  Give a reason to support your answer.
+
+```{r, null_prefix=TRUE}
+predict(fit, data.frame("LoanPaymentsOverdue" = 4), interval="predict")
+```
+
+**The expected value is -3.79 (roughly).  The 95% CI is -14.48 to 6.9.  0 is a feasible value because it falls within the confidence interval.**
 

### 2019-09-06 DATA621/Week 06 Textbook Exercises.Rmd
Renamed files

--- 2019-08-31 DATA621/Week 1 Textbook Exercises.Rmd
@@
 ---
-title: "DATA 612 Week #1 Textbook Exercises"
+title: "DATA 621 Week #6 Textbook Exercises"
 author: "Mike Silva"
@@
 
-## LMR 1.1
+## MARR 6.3
 
-The dataset *teengamb* concers a study of teenage gambline in Britain.
-Make a numerical and graphical summary of the data, commenting on any features that you find interesting.
-Limit the output you present to a quantity that a busy reader would find sufficient to get a basic understanding of the data.
+## MARR 6.5
 
-**Estimates show teens in Britain gamble between 0 and 150 pounds annually.  Males gamble more than females.  They gamble on average nearly 8 times as much.**
+## LMR 7.3
 
-```{r}
-# A paper describing the original study can be found at http://www.utstat.utoronto.ca/reid/sta2201s/2012/teengamb.pdf.
-First I will read in the data.
-I will recode the sex variable making it easier to interpret.
-The income variable is per week and the gambling is per year.
-I will adjust the income to put it on the same basis.
-
-teengamb <- read.delim("data/teengamb.txt") %>%
-  mutate(sex = as.factor(ifelse(sex == 0, "Male", "Female")),
-         income = income * 52)
-
-ggplot(teengamb, aes(x = sex, y = gamble, color=sex)) + 
-  geom_boxplot() +
-  ylab("Pounds") +
-  xlab(element_blank()) +
-  theme(legend.position = "none") +
-  ggtitle("Estimated Annual Gambling by Sex")
-```
-
-```{r}
-temp <- teengamb %>%
-  group_by(sex) %>%
-  summarise(mpg = list(enframe(quantile(gamble, probs=c(0.25, 0.5, 0.75))))) %>% 
-  unnest() %>%
-  spread(name, value)
-
-teengamb %>%
-  group_by(sex) %>%
-  summarise(`Min.` = min(gamble),
-            Mean = mean(gamble),
-            `Max.` = max(gamble)) %>%
-  merge(temp) %>%
-  select(sex, `Min.`, `25%`, `50%`, Mean, `75%`, `Max.`) %>%
-  kable() %>%
-  kable_styling()
-```
-
-**Teens earned between 31 and 780 pounds annually.  Those who earned more were more likely to gammble more.**
-
-```{r}
-ggplot(teengamb, aes(x = income, y = gamble)) + 
-  geom_point() + 
-  geom_smooth(method = "lm") +
-  xlab("Annual Income (Pounds)") +
-  ylab("Annual Gambling (Pounds)") +
-  ggtitle("Income vs. Gambling")
-```
-
-## LMR 1.3
-
-The dataset *prostate* is from a study on 97 men with prostate conacer who were due to receive a radical prostatectomy.  Make a numerical and graphical summary of the data as in the first question.
-
-```{r}
-#lcavol log(cancer volume)
-#lweight log(prostate weight)
-#age age
-#lbph log(benign prostatic hyperplasia amount)
-#svi seminal vesicle invasion
-#lcp log(capsular penetration)
-#gleason Gleason score
-#pgg45 percentage Gleason scores 4 or 5
-#lpsa log(prostate specific antigen)
-
-prostate <- read.delim("data/prostate.txt") 
-pairs(prostate)
-```
-
-## LMR 1.4
-
-The dataset *sat* comes from a study entitled "Getting What You Pay For: The Debate Overty Equity in Public School Expenditures."  Make a numerical and graphical summary of the data as in the first question.
-
-## LMR 1.5
-
-The dataset *divusa* contains data on divorces in the United States from 1920 to 1996.  Make a numerical and graphical summary of the data as in the first question.
+## LMR 7.8

### 2019-09-06 DATA608/hw1.rmd
More work on hw1

@@
 output:
-  html_document: default
+  html_document:
+    code_folding: "hide"
   html_notebook: default
 ---
+```{r}
+library(knitr)
+# A Prefix nulling hook.
 
+# Make sure to keep the default for normal processing.
+default_output_hook <- knitr::knit_hooks$get("output")
+
+# Output hooks handle normal R console output.
+knitr::knit_hooks$set( output = function(x, options) {
+
+  comment <- knitr::opts_current$get("comment")
+  if( is.na(comment) ) comment <- ""
+  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
+                     x, perl = TRUE)
+  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
+  if( can_null && do_null ) {
+    # By default R print output aligns at the right brace.
+    align_index <- regexpr( "\\]", x )[1] - 1
+    # Two cases: start or newline
+    re <- paste0( "^.{", align_index, "}\\]")
+    rep <- comment
+    x <- gsub( re, rep,  x )
+    re <- paste0( "\\\n.{", align_index, "}\\]")
+    rep <- paste0( "\n", comment )
+    x <- gsub( re, rep,  x )
+  }
+
+  default_output_hook( x, options )
+
+})
+
+knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
+```
 **Principles of Data Visualization and Introduction to ggplot2**
@@
 
-```{r}
+```{r comment=NA}
 
@@
 ```
-```{r}
+```{r comment=NA}
 
@@
 
-*First let's start with some aggregate counts of the data:*
+**First let's start with some aggregate counts of the data:**
 
@@
 
-*Hmmm.  There are 52 states in the data set.  I'm guessing there is DC but I wonder if there is an error in the data.  I will take a closer look at this:*
+**Hmmm.  There are 52 states in the data set.  I'm guessing there is DC but I wonder if there is an error in the data.  I will take a closer look at this:**
 
-```{r, comment=NA}
+```{r, comment=NA, null_prefix=TRUE}
 unique(inc$State)
@@
 
-*Ahh!  There's Puerto Rico.  My hunch about DC was correct.  No need to clean that up.  Great!*
+**Ahh!  There's Puerto Rico.  My hunch about DC was correct.  No need to clean that up.  Great!**
 
-*I am wondering how some of the variables differ the industry.
+**I am wondering how some of the variables differ the industry.
 I will look at the size of the work force and the revenue.
-I will also construct an output per worker to see which industry is the most productive in economic terms:*
+I will also construct an output per worker to see which industry is the most productive in economic terms:**
 
@@
 
-*This is not supprising that the top industries are the most "productive."  They are very capital intensive.*
+**This is not supprising that the top industries are the most "productive."  They are very capital intensive.**
 
@@
 library(ggplot2)
-library(ggthemes)
 
@@
   ggplot(aes(x = n, y=reorder(State, n))) +
-  geom_point() +
-  ggtitle("Number of Fastest Growing Companies by State") +
-  ylab(element_blank()) + 
-  xlab("Companies") + 
-  theme_clean()
+  geom_point(color = "dodger blue") +
+  labs(title = "Number of Fastest Growing Companies by State",
+       y = element_blank(),
+       x = element_blank()) +
+  theme(panel.background = element_blank(),
+        panel.grid.major = element_line(color = "gray95"),
+        panel.grid.major.x = element_line(color = "gray95", linetype="dashed"))
 ```
@@
   ylim(0, 1000) +
-  xlab(element_blank()) + 
-  ggtitle("Workforce Size by Industry", subtitle = "(employment over 1,000 not shown)") +
-  geom_boxplot(outlier.colour="gray", outlier.shape=8, outlier.size = 0.9) +
-  coord_flip() +
-  theme_clean()
+  labs(title = "Workforce Size by Industry",
+       subtitle = "(employment over 1,000 not shown)",
+       y = element_blank(),
+       x = element_blank()) +
+  theme(panel.background = element_blank(),
+        panel.grid.major = element_line(color = "gray95"),
+        panel.grid.major.x = element_line(color = "gray95", linetype="dashed")) +
+  geom_boxplot(outlier.colour="gray", outlier.shape = 8, outlier.size = 0.9) +
+  coord_flip()
 ```
@@
 
-*Funny!  I guess I think like this investor.*
+**Funny!  I guess I think like this investor.**
 
@@
   ylim(0, 1) +
-  xlab(element_blank()) + 
-  ylab("Millions of Dollars") + 
-  ggtitle("Revenue per Employee by Industry", subtitle = "(revenue over $1MM not shown)") +
-  geom_boxplot(outlier.colour="gray", outlier.shape=8, outlier.size = 0.9) +
-  coord_flip() +
-  theme_clean()
+  labs(title = "Revenue per Employee by Industry", 
+       subtitle = "(revenue over $1MM not shown)",
+       y = "Millions of Dollars",
+       x = element_blank()) +
+  theme(panel.background = element_blank(),
+        panel.grid.major = element_line(color = "gray95"),
+        panel.grid.major.x = element_line(color = "gray95", linetype="dashed")) +
+  geom_boxplot(outlier.colour="gray", outlier.shape = 8, outlier.size = 0.9) +
+  coord_flip()
 ```

### 2019-09-07 DATA621/Week 02 Textbook Exercises.Rmd
More work on Week 2 Textbook Exercises

--- 2019-09-06 DATA621/Week 2 Textbook Exercises.Rmd
@@
     highlight: kate
+    code_folding: "hide"
 ---
@@
        y = "Current Week") +
-  geom_abline(slope=1, intercept = 0, color = "red")
+  geom_abline(slope=1, intercept = 0, color = "red") + 
+  theme(panel.background = element_blank(),
+        panel.grid.major = element_line(color = "gray90"))
 ```
@@
 
+The dataset *prostate* comes from a study on 97 men with prostate cancer who were due to receive a radical prostatectomy.
+Fit a model with *lpsa* as the response and *lcavol* as the predictor.
+Record the residual standard error and the $R^2$.
+Now add *lweight*, *svi*, *lbph*, *age*, *lcp*, *pgg45* and *gleason* to the model on one at a time.
+For each model record the residual standard error and the $R^2$.
+Plot the trends in these two statistics.
+
+```{r}
+prostate <- read.delim("data/prostate.txt")
+
+get_row <- function(fit, n_vars){
+  r2 <- summary(fit)$r.squared
+  rse <- sigma(fit)
+  data.frame("Predictors" = n_vars, "RSE" = rse, "R2" = r2)
+}
+
+df <- get_row(lm(lpsa ~ lcavol, prostate), 1) %>% 
+  bind_rows(get_row(lm(lpsa ~ lcavol + lweight, prostate), 2)) %>%
+  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi, prostate), 3)) %>%
+  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph, prostate), 4)) %>%
+  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph, prostate), 5)) %>%
+  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph + age, prostate), 6)) %>%
+  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp, prostate), 7)) %>%
+  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp + pgg45, prostate), 8)) %>%
+  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp + pgg45 + gleason, prostate), 9)) %>%
+  gather(Measure, Value, -Predictors) %>%
+  mutate(Measure = as.factor(Measure))
+
+ggplot(df, aes(x=Predictors, y = Value, color = Measure)) +
+  geom_line() +
+  scale_x_continuous(breaks = round(seq(min(df$Predictors), max(df$Predictors), by = 1), 0)) +
+  scale_y_continuous(breaks = round(seq(0, 1, by = 0.05), 2)) +
+  annotate("text", x = 5, y = max(df[df$Measure == "R2",]$Value) + 0.01, label ="R2") +
+  annotate("text", x = 5, y = min(df[df$Measure == "RSE",]$Value) + 0.02, label ="RSE") +
+  labs(title="R2 and RSE by Number of Variables", 
+       x = element_blank(), 
+       y = element_blank()) +
+  theme(legend.position = "None",
+        panel.background = element_blank(),
+        panel.grid.major = element_line(color = "gray90"))
+```
+
 ## LMR 2.5
+
+Using the *prostate* data, plot *lpsa* against *lcavol*.
+Fit the regressions of *lpsa* on *lcavol* and *lcavol* on *lpsa*.
+Display both regression lines on the plot.
+At what point do the two lines intersect?
+**They intersect where the cancer volume roughly 1.3 and the log of PSA is 2.5**
+
+```{r}
+fit <- lm(lpsa ~ lcavol, prostate)
+df <- data.frame(lcavol = prostate$lcavol, 
+                 lpsa = predict(fit, prostate), 
+                 Model = "lpsa ~ lcavol")
+
+fit <- lm(lcavol ~ lpsa, prostate)
+df <- df %>% 
+  bind_rows(data.frame(lcavol = predict(fit, prostate), 
+                       lpsa = prostate$lpsa, 
+                       Model = "lcavol ~ lpsa")) %>%
+  mutate(Model = as.factor(Model))
+
+ggplot(prostate, aes(lcavol, lpsa)) + 
+  geom_point() + 
+  geom_line(aes(lcavol, lpsa, color = Model), data = df) + 
+  labs(x = "Cancer Volume", 
+       y = "PSA (log)") +
+  theme(legend.position = "bottom",
+        panel.background = element_blank(),
+        panel.grid.major = element_line(color = "gray90")) +
+  scale_x_continuous(breaks = round(seq(min(prostate$lcavol), max(prostate$lcavol), by = 1), 0)) +
+  scale_y_continuous(breaks = round(seq(min(prostate$lpsa), max(prostate$lpsa), by = 1), 0)) 
+```

### 2019-09-07 DATA621/Week 01 Textbook Exercises.Rmd
Minor change

--- 2019-09-04 DATA621/Week 1 Textbook Exercises.Rmd
@@
 ---
-title: "DATA 612 Week #1 Textbook Exercises"
+title: "DATA 621 Week #1 Textbook Exercises"
 author: "Mike Silva"

### 2019-09-23 DATA621/HW1.Rmd
Initial commit of DATA 621 HW1 files

---
title: "DATA 621 Homework #1"
author: "Michael Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(corrplot)
```

## Introduction

```{r read_data, echo=FALSE}
# Read in the training data
training <- read.csv("./data/moneyball-training-data.csv") %>%
  select(-INDEX) # Dropping meaningless index
# Read in the evaluation data
evaluation <- read.csv("./data/moneyball-evaluation-data.csv")
```

We have been given a dataset with `r nrow(training)` records summarizing a major league baseball team's season. The records span 1871 to 2006 inclusive.  All statistics have been adjusted to match the performance of a 162 game season.  The objective is to build a linear regression model to predict the number of wins for a team.

### Working Theory

We are working on the premise that there are "good" teams and there are "bad" teams.  The good teams win more than the bad teams.  We are assuming that some of the predictors will be higher for the good teams than for the bad teams.  Consequently we can use these variables to predict how many times a team will win in a season.

### Notes About the Data

There are some difficulties with this dataset.  First it covers such a wide time period.  We know there are different "eras" of baseball.  This data will span multiple eras.  Has the fundamental relationships between wining and these predictors change over time?  We think it has.  If so this will be a challenge.

## Data Exploration

### First Look at the Data

We will first look at the data to get a sense of what we have.

```{r small_multiples_density, warning=FALSE}
training %>%
  gather(variable, value, TARGET_WINS:TEAM_FIELDING_DP) %>%
  ggplot(., aes(value)) + 
  geom_density(fill = "indianred4", color="indianred4") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

```{r}
quick_summary <- function(df){
  df %>%
    summary() %>%
    kable() %>%
    kable_styling()
}

quick_summary(training)
```

Some initial observations:

* The response variable (`TARGET_WINS`) looks to be normally distributed.  This supports the working theory that there are good teams and bad teams.  There are also a lot of average teams.
* There are also quite a few variables with missing values.  We may need to deal with these in order to have the largest data set possible for modeling.
* A couple variables are bimodal (`TEAM_BATTING_HR`, `TEAM_BATTING_SO` `TEAM_PITCHING_HR`).  This may be a challenge as some of them are missing values and that may be a challenge in filling in missing values.
* Some variables are right skewed (`TEAM_BASERUN_CS`, `TEAM_BASERUN_SB`, etc.).  This might support the good team theory.  It may also introduce non-normally distributed residuals in the model.  We shall see.

### Correlations

Let's take a look at the correlations.  The following is the correlations from the complete cases only:

```{r correlation plot}
training %>% 
  cor(., use = "complete.obs") %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", diag = FALSE)
```

```{r}
temp <- training %>% 
  cor(., use = "complete.obs") #%>%

temp[lower.tri(temp, diag=TRUE)] <- ""
temp <- temp %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  gather(Variable, Correlation, -rowname) %>%
  filter(Variable != rowname) %>%
  filter(Correlation != "") %>%
  mutate(Correlation = as.numeric(Correlation)) %>%
  rename(` Variable` = rowname) %>%
  arrange(desc(abs(Correlation))) 
```

#### Correlations with Response Variable

Let's take a look at how the predictors are correlated with the response variable:

```{r}
temp %>%
  filter(` Variable` == "TARGET_WINS") %>%
  kable() %>%
  kable_styling()
```

It looks like the hits, walks, home runs, and errors have the strongest correlations with wins.  None of these correlations are particularly strong.  This suggests there is a lot of 'noise' in these relationships.

It is interesting to note allowing hits is positively correlated with wins. How strange!  Let's take a closer look at this correlation:

```{r}
ggplot(training, aes(TEAM_PITCHING_H, TARGET_WINS)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Hits Allowed") +
  ylab("Wins")
```

It is also noteworthy that pitching strikeouts is negatively correlated with winning.  That does not make any sense.

#### Strong Correlations (Absolute Value > 0.5)

Let's see if the are any predictors are correlated with each other.  We will only look for "strong" correlations:

```{r}
temp %>%
  filter(abs(Correlation) > .5) %>%
  kable() %>%
  kable_styling()
```

There are `r temp %>% filter(Correlation > .99) %>% nrow(.)` variables that have a correlation that is almost 1!  We will need to be careful to prevent adding autocorrelation errors to our model.

### Missing Values

During our first look at the data it was noted that there were variables that are missing data.  Here's a look at what variables are missing data and how big of a problem is it:

```{r}
training %>% 
  gather(variable, value) %>%
  filter(is.na(value)) %>%
  group_by(variable) %>%
  tally() %>%
  mutate(percent = n / nrow(training) * 100) %>%
  mutate(percent = paste0(round(percent, ifelse(percent < 10, 1, 0)), "%")) %>%
  arrange(desc(n)) %>%
  rename(`Variable Missing Data` = variable,
         `Number of Records` = n,
         `Share of Total` = percent) %>%
  kable() %>%
  kable_styling()
```

The hit by pitcher varriable is missing over 90% of it's data.  Based on the weak correlation with the response variable we will exclude it from consideration in our model.

Caught stealling a base (`TEAM_BASERUN_CS`) is next on the list.  It may be possible to predict it using `TEAM_BASERUN_SB` since they are strongly correlated, but there are `r training %>% filter(is.na(TEAM_BASERUN_SB) & is.na(TEAM_BASERUN_SB)) %>% nrow()` times they both are missing data.

## Data Preparation

## Model Building

## Model Selection

## Appendix

### 2019-09-25 DATA621/HW1.Rmd
More work on HW1

@@
 title: "DATA 621 Homework #1"
-author: "Michael Silva"
+subtitle: "From Work by Critical Thinking Group 3"
+author: "Mike Silva"
 date: "`r Sys.Date()`"
@@
     toc_depth: 3
-    code_folding: "hide"
+    code_folding: "show"
 ---
@@
 library(corrplot)
+library(caret)
+library(DMwR)
 ```
@@
 This data will span multiple eras.
-Has the fundamental relationships between wining and these predictors change over time?
+Has the fundamental relationships between winning and these predictors change over time?
 We think it has.
@@
 
+```{r warning=FALSE}
+training %>%
+  gather(variable, value, -TARGET_WINS) %>%
+  ggplot(., aes(value, TARGET_WINS)) + 
+  geom_point(fill = "indianred4", color="indianred4") + 
+  geom_smooth(method = "lm", se = FALSE, color = "black") + 
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = "Wins")
+```
+
 ```{r}
@@
 
-It is interesting to note allowing hits is positively correlated with wins. How strange!  Let's take a closer look at this correlation:
-
-```{r}
-ggplot(training, aes(TEAM_PITCHING_H, TARGET_WINS)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  xlab("Hits Allowed") +
-  ylab("Wins")
-```
-
-It is also noteworthy that pitching strikeouts is negatively correlated with winning.  That does not make any sense.
+It is interesting to note allowing hits is positively correlated with wins.
+How strange!
+It is also noteworthy that pitching strikeouts is negatively correlated with winning.
+That does not make any sense.
+When one examines the scatter plots above it becomes apparent that these correlations are being effected by some outliers.
 
@@
 
-Let's see if the are any predictors are correlated with each other.  We will only look for "strong" correlations:
+Are any predictors are correlated with each other?  We will only look for "strong" correlations:
 
@@
 
-### Missing Values
+### Strange Data Values
 
-During our first look at the data it was noted that there were variables that are missing data.  Here's a look at what variables are missing data and how big of a problem is it:
+#### Missing Values
+
+During our first look at the data it was noted that there were variables that are missing data.  Here's a look at what variables are missing data and how big of a problem it is:
 
@@
 
-The hit by pitcher varriable is missing over 90% of it's data.  Based on the weak correlation with the response variable we will exclude it from consideration in our model.
+The hit by pitcher varriable is missing over 90% of it's data.  We will exclude it from consideration in our model.
 
@@
 
+The strike outs are going to be a little tricky because of their bimodal distribution.  All `r training %>% filter(is.na(TEAM_BATTING_SO) & is.na(TEAM_PITCHING_SO)) %>% nrow()`
+
+#### Zero Values
+
+There are also variables that have verly low values.  Let's see how big of a problem this is:
+
+```{r}
+training %>% 
+  gather(variable, value) %>%
+  filter(value == 0) %>%
+  group_by(variable) %>%
+  tally() %>%
+  mutate(percent = n / nrow(training) * 100) %>%
+  mutate(percent = paste0(round(percent, ifelse(percent < 10, 1, 0)), "%")) %>%
+  arrange(desc(n)) %>%
+  rename(`Variable With Zeros` = variable,
+         `Number of Records` = n,
+         `Share of Total` = percent) %>%
+  kable() %>%
+  kable_styling()
+```
+
+This isn't nearly as large of a problem as the missing values.
+
+### Deeper Dive into the Variables
+
+#### TARGET_WINS
+
+```{r}
+training %>% 
+  ggplot(aes(TARGET_WINS)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TARGET_WINS)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TARGET_WINS)), col = "green", lty = 2) +
+  labs(x = element_blank(), 
+       y = "Count",
+       title = "Distribution of Wins",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+The range of data looks good.  There is a single zero value, which is the Washington Nationals 1872 season.
+
+#### TEAM_BATTING_3B
+
+This field represents triples hit by the team. Triples aren't very common because the ball is still in the field of play (unlike a homerun) but the batter still has enough time to get 3 bases.
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_BATTING_3B)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_BATTING_3B)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_BATTING_3B)), col = "green", lty = 2) +
+  labs(x = element_blank(), 
+       y = "Count",
+       title = "Distribution of Triples",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+Looking at the distribution, the value of zero doesn't look too unusual. Even if it were, the value is not likely to have a large impact.
+
+#### TEAM_BATTING_HR
+
+Although homeruns are more common in modern baseball (thank you PDEs!), there are some low values in the data. So zero doesn't seem too unusual here either.
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_BATTING_HR)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_BATTING_HR)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_BATTING_HR)), col = "green", lty = 2) +
+  labs(x = element_blank(), 
+       y = "Count",
+       title = "Distribution of Homeruns",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+#### TEAM_BATTING_BB
+
+This variable represents when the batter is "walked" by the pitcher (also known as Base on Balls):
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_BATTING_BB)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_BATTING_BB)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_BATTING_BB)), col = "green", lty = 2) +
+  labs(x = element_blank(), 
+       y = "Count",
+       title = "Distribution of Walks (Base on Balls)",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+Four balls will walk a batter in modern baseball, however that wasn't always the case.
+A century or more ago (within the date range of this data set) walks took as many as 9 balls to happen[^1].
+Knowing this, and looking at the left-tail of the values above, it is not unreasonable that there might be a season with no walks.
+Like triples above, leaving the one zero data point in is unlikely to adversely impact any regression, since there are valid values nearby.
+
+#### TEAM_BATTING_SO
+
+Here we saw some NA values, `r training %>% filter(is.na(TEAM_BATTING_SO)) %>% nrow(.)` of them to be specific.
+Plus we have `r training %>% filter(TEAM_BATTING_SO == 0) %>% nrow(.)` zero values as well.
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_BATTING_SO)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_BATTING_SO,na.rm = T)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_BATTING_SO, na.rm = T)), col = "green", lty = 2) +
+  labs(x = element_blank(), 
+       y = "Count",
+       title = "Distribution of Strikeouts (Batter)",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+First, the zero values seem nigh-impossible. An entire season (162 games) without a single batter being struck out seems highly suspect, let alone 20 of them in the dataset.
+
+We will replace these values with imputed values, but the distribution looks to be bimodal, so using a mean or median (which is squarely between those peaks) may cause some issues with the model.
+So, instead, we will impute values using regression.
+
+We will impute values for this variable by looking at it's nearest neighbors (based on other variables) and taking a weighted average of their values.
+
+#### TEAM_BASERUN_SB
+
+With this variable, we have a good number of NA values, and 2 zeroes:
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_BASERUN_SB)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_BASERUN_SB, na.rm = T)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_BASERUN_SB, na.rm = T)), col = "green", lty = 2) +
+  labs(x = element_blank(),
+       y = "Count",
+       title = "Distribution of Stolen Bases",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+The zeroes may be legitimate here so we will leave them alone. For the NAs, we will use the same KNN imputation we used above for strikeouts to impute values.
+
+#### TEAM_BASERUN_CS
+
+This variable is NA for nearly a third of records and only 2 zero values (which could be legitimate values):
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_BASERUN_CS)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_BASERUN_CS, na.rm = T)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_BASERUN_CS, na.rm = T)),col = "green", lty = 2) +
+  labs(x = element_blank(),
+       y = "Count",
+       title = "Distribution of Caught Stealing Bases",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+Despite the high number of missing values (and a potential for increased error), we will use the KNN imputed values.
+
+#### TEAM_BATTING_HBP
+
+With this variable, we see nearly all entries are missing:
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_BATTING_HBP)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_BATTING_HBP, na.rm = T)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_BATTING_HBP, na.rm = T)), col = "green", lty = 2) +
+  labs(x = element_blank(),
+       y = "Count",
+       title = "Distribution of Being Hit By a Pitch",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+We will drop this variable as we cannot responsibly impute values with such thin data.
+
+#### TEAM_PITCHING_HR
+
+This variable has no NA values, but there are a few zero values. However, the zero values seem to be legitimate given the distribution:
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_PITCHING_HR)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_PITCHING_HR, na.rm = T)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_PITCHING_HR, na.rm = T)), col = "green", lty = 2) +
+  labs(x = element_blank(),
+       y = "Count",
+       title = "Distribution of Homeruns Pitched",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+#### TEAM_PITCHING_BB
+
+Here we have no NA values and a single zero:
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_PITCHING_BB)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_PITCHING_BB, na.rm = T)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_PITCHING_BB, na.rm = T)), col = "green", lty = 2) +
+  labs(x = element_blank(),
+       y = "Count",
+       title = "Distribution of Walks Pitched",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+As we did with walks above, we can assume that is is possible to have no walks (and therefore pitch no walks). So, we will leave the zero alone.
+
+However, there are some **really** high values in the data, which strains reality a little.
+We could take anything defined as an outlier ($1.5 \cdot \text{IQR}$) and set it to NA so those records will be excluded from any model we build with this variable.
+But, when you do the math it seems extreme, but plausible.
+For example, the most number of games in a season in MLB is 162 (currently).
+With a max value or 3,645 walks pitched you get 22.5 walks per game on average.
+Divided equally amongst 9 innings, it comes out to 2.5 walks per inning.
+
+I'd be surprised that any pitcher wouldn't be removed after an inning or two of 2-3 walks, but neither can we rule it out as a possibility.
+
+#### TEAM_PITCHING_SO
+
+This variable represents strikeouts pitched. We see that there are 102 NA values and a *lot* of extremely high values:
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_PITCHING_SO)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_PITCHING_SO, na.rm = T)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_PITCHING_SO, na.rm = T)), col = "green", lty = 2) +
+  labs(x = element_blank(),
+       y = "Count",
+       title = "Distribution of Strikeouts Pitched",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+The extreme values need to be handled.
+First, a typical game will be 9 innings in length, and in each inning you can only pitch 3 strikeouts (because then your part of the inning is over).
+Those 27 potential strikeouts multiplied by 162 games means an upper limit near 4,374 a season.
+
+Games can go beyond 9 innings, but even if every game in a season was as long as the longest ever MLB game (26 innings) you can only have 12,636 strikeouts.
+So, the max value of `r max(training$TEAM_PITCHING_SO,na.rm=T)` is invalid.
+
+We'll make a high-yet-reasonable assumption of a mean 11 innings per game.
+We will call anything more than 5,346 strikeouts an invalid data point by setting them to NA so they will be imputed prior to modeling.
+
+#### TEAM_FIELDING_DP
+
+The values in this variable seem reasonable, however we do have some NA values.
+
+```{r}
+training %>% 
+  ggplot(aes(TEAM_FIELDING_DP)) + 
+  geom_histogram(bins = 50) +
+  geom_vline(aes(xintercept = mean(TEAM_FIELDING_DP, na.rm = T)), col = "red", lty = 2) +
+  geom_vline(aes(xintercept = median(TEAM_FIELDING_DP, na.rm = T)), col = "green", lty = 2) +
+  labs(x = element_blank(),
+       y = "Counts",
+       title = "Distribution of Double Plays",
+       caption = "* Red line is the mean value and green is the median")
+```
+
+Again, we will use the KNN imputation from earlier to fill in NAs with imputed values.
+
 ## Data Preparation
+
+### Fixing Missing/Zero Values
+
+First we will remove the invalid data and prep it for imputation.  We will drop the hit by pitcher variable from the dataset.
+
+```{r}
+remove_bad_values <- function(df){
+  # Change 0's to NA so they too can be imputed
+  df <- df %>%
+  mutate(TEAM_BATTING_SO = ifelse(TEAM_BATTING_SO == 0, NA, TEAM_BATTING_SO))
+  # Remove the high pitching strikeout values
+  df[which(df$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
+  # Drop the hit by pitcher variable
+  df %>%
+    select(-TEAM_BATTING_HBP)
+}
+
+training <- remove_bad_values(training)
+evaluation <- remove_bad_values(evaluation)
+```
+
+```{r}
+set.seed(42)
+knn <- training %>% knnImputation()
+
+impute_values <- function(df, knn){
+  impute_me <- is.na(df$TEAM_BATTING_SO)
+  df[impute_me,"TEAM_BATTING_SO"] <- knn[impute_me,"TEAM_BATTING_SO"] 
+  impute_me <- is.na(df$TEAM_BASERUN_SB)
+  df[impute_me,"TEAM_BASERUN_SB"] <- knn[impute_me,"TEAM_BASERUN_SB"] 
+  impute_me <- is.na(df$TEAM_BASERUN_CS)
+  df[impute_me,"TEAM_BASERUN_CS"] <- knn[impute_me,"TEAM_BASERUN_CS"] 
+  impute_me <- is.na(df$TEAM_PITCHING_SO)
+  df[impute_me,"TEAM_PITCHING_SO"] <- knn[impute_me,"TEAM_PITCHING_SO"]
+  impute_me <- is.na(df$TEAM_FIELDING_DP)
+  df[impute_me,"TEAM_FIELDING_DP"] <- knn[impute_me,"TEAM_FIELDING_DP"]
+  return(df)
+}
+
+training <- impute_values(training, knn)
+```
+
+### Feature Engineering
+
+The batting singles is not included but we can back it out of the hits.  We will do this.
+
+```{r}
+add_features <- function(df){
+  df %>%
+    mutate(TEAM_BATTING_1B = TEAM_BATTING_H - TEAM_BATTING_2B - TEAM_BATTING_3B - TEAM_BATTING_HR)
+}
+
+training <- add_features(training)
+evaluation <- add_features(evaluation)
+```
+
+### Results
+
+Here's what the data look like after imputation and correction:
+
+```{r warning=FALSE}
+training %>%
+  gather(variable, value) %>%
+  ggplot(., aes(value)) + 
+  geom_density(fill = "indianred4", color="indianred4") + 
+  facet_wrap(~variable, scales ="free", ncol = 4) +
+  labs(x = element_blank(), y = element_blank())
+```
+
+```{r}
+quick_summary <- function(df){
+  df %>%
+    summary() %>%
+    kable() %>%
+    kable_styling()
+}
+
+quick_summary(training)
+```
 
@@
 
+We will divide the data into training and test sets using a 70/30 split.  We will build our models on the training set and evaluate it on the test set.
+
+```{r}
+set.seed(42)
+train_index <- createDataPartition(training$TARGET_WINS, p = .7, list = FALSE, times = 1)
+
+moneyball_train <- training[train_index,]
+moneyball_test <- training[-train_index,]
+```
+
+### Kitchen Sink Model
+
+We will begin with a "kitchen sink" model.
+
+```{r}
+kitchen_sink <- lm(TARGET_WINS ~ ., moneyball_train)
+summary(kitchen_sink)
+```
+
+It does a fairly good job predicting, but there are a lot of variables that are not statistically significant.
+
+```{r}
+plot(kitchen_sink)
+```
+
+### Simple Model
+
+Let's try to create a simplier model. We will pick variables that had high correlations and exclude the pitching variables which would introduce autocorrelation issues.
+
+```{r}
+simple_fit <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_BB + TEAM_FIELDING_E, data = moneyball_train)
+summary(simple_fit)
+```
+
+This model isn't as fitted to the data as well but the variables are statistically significant.
+
+```{r}
+plot(simple_fit)
+```
+
+### Higher Order Stepwise Regression
+
+For the third model we will use a stepwise regression method using a backwards elimination process.  We also introduce some higher order polynomial variables.
+
+```{r}
+full_formula <- "TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP + TEAM_BATTING_1B + I(TEAM_BATTING_2B^2) + I(TEAM_BATTING_3B^2) + I(TEAM_BATTING_HR^2) + I(TEAM_BATTING_BB^2) + I(TEAM_BATTING_SO^2) + I(TEAM_BASERUN_SB^2) + I(TEAM_BASERUN_CS^2) + I(TEAM_PITCHING_H^2) + I(TEAM_PITCHING_HR^2) + I(TEAM_PITCHING_BB^2) + I(TEAM_PITCHING_SO^2) + I(TEAM_FIELDING_E^2) + I(TEAM_FIELDING_DP^2) + I(TEAM_BATTING_1B^2) + I(TEAM_BATTING_2B^3) + I(TEAM_BATTING_3B^3) + I(TEAM_BATTING_HR^3) + I(TEAM_BATTING_BB^3) + I(TEAM_BATTING_SO^3) + I(TEAM_BASERUN_SB^3) + I(TEAM_BASERUN_CS^3) + I(TEAM_PITCHING_H^3) + I(TEAM_PITCHING_HR^3) + I(TEAM_PITCHING_BB^3) + I(TEAM_PITCHING_SO^3) + I(TEAM_FIELDING_E^3) + I(TEAM_FIELDING_DP^3) + I(TEAM_BATTING_1B^3) + I(TEAM_BATTING_2B^4) + I(TEAM_BATTING_3B^4) + I(TEAM_BATTING_HR^4) + I(TEAM_BATTING_BB^4) + I(TEAM_BATTING_SO^4) + I(TEAM_BASERUN_SB^4) + I(TEAM_BASERUN_CS^4) + I(TEAM_PITCHING_H^4) + I(TEAM_PITCHING_HR^4) + I(TEAM_PITCHING_BB^4) + I(TEAM_PITCHING_SO^4) + I(TEAM_FIELDING_E^4) + I(TEAM_FIELDING_DP^4) + I(TEAM_BATTING_1B^4)"
+
+full_model <- lm(full_formula, moneyball_train)
+step_back <- MASS::stepAIC(full_model, direction="backward", trace = F)
+poly_call <- summary(step_back)$call
+step_back <- lm(poly_call[2], moneyball_train)
+summary(step_back)
+```
+
+This model has the highest adjusted R-squared but has some variables that are statistically insignificant.
+
+```{r}
+plot(step_back)
+```
+
 ## Model Selection
 
-## Appendix
+In order to select which model is the "best" we will test it against a test set.
+We will examine the difference between the predicted and actual values.
+Since the wins are in whole numbers and the `predict` function will generate floating point numbers, I will be rounding the results.
+
+### Kitchen Sink Results
+
+```{r}
+moneyball_test$kitchen_sink <- round(predict(kitchen_sink, moneyball_test), 0)
+moneyball_test <- moneyball_test %>%
+  mutate(kitchen_sink_error = TARGET_WINS - kitchen_sink)
+
+ggplot(moneyball_test, aes(kitchen_sink_error)) +
+  geom_histogram(bins = 50)
+
+summary(moneyball_test$kitchen_sink_error)
+```
+
+### Simple Model
+
+```{r}
+moneyball_test$simple <- round(predict(simple_fit, moneyball_test), 0)
+moneyball_test <- moneyball_test %>%
+  mutate(simple_error = TARGET_WINS - simple)
+
+ggplot(moneyball_test, aes(simple_error)) +
+  geom_histogram(bins = 50)
+
+summary(moneyball_test$simple_error)
+```
+
+### Stepwise Results
+
+```{r}
+moneyball_test$step_back <- round(predict(step_back, moneyball_test), 0)
+moneyball_test <- moneyball_test %>%
+  mutate(step_back_error = TARGET_WINS - step_back)
+
+moneyball_test %>%
+  filter(step_back_error > -100) %>%
+  ggplot(., aes(step_back_error)) +
+  geom_histogram(bins = 50) +
+  labs(caption = "Outlier removed")
+
+summary(moneyball_test$step_back_error)
+```
+
+This model did not preform very well on the out of sample data.
+
+## Final Model Selection
+
+In deciding which model to use to make my final predictions I will look at the amount of error in the predictions:
+
+```{r}
+moneyball_test %>%
+  select(kitchen_sink_error, simple_error, step_back_error) %>%
+  rename(`Kitchen Sink` = kitchen_sink_error,
+         `Simple` = simple_error,
+         `Stepwise` = step_back_error) %>%
+  summary() %>%
+  kable() %>%
+  kable_styling()
+```
+
+The stepwise regression model did not preform well with out of sample data.
+The high adjusted R-squared is probably evidence of overfitting.
+The simple and "kitchen sink" model have similar preformance.
+I will use the simple model because I prefer to keep it simple stupid.
+The residuals of the simple model are mostly normally distributed.
+The adjusted R squared indicates about a quarter of the variation is explained by the model.
+I will complete this homework by generating predictions off of the evaluation set using the simple model.
+
+```{r}
+evaluation$TARGET_WINS <- round(predict(simple_fit, evaluation), 0)
+evaluation %>%
+  select(INDEX, TARGET_WINS) %>%
+  write.csv(., "./data/moneyball-predictions.csv", row.names = F)
+```
+
+[^1]: https://en.wikipedia.org/wiki/Base_on_balls

### 2019-09-25 DATA621/HW1.Rmd
Updated HW1.Rmd

@@
 
-First we will remove the invalid data and prep it for imputation.  We will drop the hit by pitcher variable from the dataset.
+First we will remove the invalid data and prep it for imputation.  We will drop the hit by pitcher variable from the dataset.  We will then impute the data using KNN.
 
 ```{r}
-remove_bad_values <- function(df){
-  # Change 0's to NA so they too can be imputed
-  df <- df %>%
-  mutate(TEAM_BATTING_SO = ifelse(TEAM_BATTING_SO == 0, NA, TEAM_BATTING_SO))
-  # Remove the high pitching strikeout values
-  df[which(df$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
-  # Drop the hit by pitcher variable
-  df %>%
+training <- training %>%
+  mutate(TEAM_BATTING_SO = ifelse(TEAM_BATTING_SO == 0, NA, TEAM_BATTING_SO)) %>%
+  mutate(TEAM_PITCHING_SO = ifelse(TEAM_PITCHING_SO > 5346, NA, TEAM_PITCHING_SO)) %>%
   select(-TEAM_BATTING_HBP)
-}
 
-training <- remove_bad_values(training)
-evaluation <- remove_bad_values(evaluation)
-```
-
-```{r}
 set.seed(42)
 knn <- training %>% knnImputation()
-
-impute_values <- function(df, knn){
-  impute_me <- is.na(df$TEAM_BATTING_SO)
-  df[impute_me,"TEAM_BATTING_SO"] <- knn[impute_me,"TEAM_BATTING_SO"] 
-  impute_me <- is.na(df$TEAM_BASERUN_SB)
-  df[impute_me,"TEAM_BASERUN_SB"] <- knn[impute_me,"TEAM_BASERUN_SB"] 
-  impute_me <- is.na(df$TEAM_BASERUN_CS)
-  df[impute_me,"TEAM_BASERUN_CS"] <- knn[impute_me,"TEAM_BASERUN_CS"] 
-  impute_me <- is.na(df$TEAM_PITCHING_SO)
-  df[impute_me,"TEAM_PITCHING_SO"] <- knn[impute_me,"TEAM_PITCHING_SO"]
-  impute_me <- is.na(df$TEAM_FIELDING_DP)
-  df[impute_me,"TEAM_FIELDING_DP"] <- knn[impute_me,"TEAM_FIELDING_DP"]
-  return(df)
-}
-
-training <- impute_values(training, knn)
+impute_me <- is.na(training$TEAM_BATTING_SO)
+training[impute_me,"TEAM_BATTING_SO"] <- knn[impute_me,"TEAM_BATTING_SO"] 
+impute_me <- is.na(training$TEAM_BASERUN_SB)
+training[impute_me,"TEAM_BASERUN_SB"] <- knn[impute_me,"TEAM_BASERUN_SB"] 
+impute_me <- is.na(training$TEAM_BASERUN_CS)
+training[impute_me,"TEAM_BASERUN_CS"] <- knn[impute_me,"TEAM_BASERUN_CS"] 
+impute_me <- is.na(training$TEAM_PITCHING_SO)
+training[impute_me,"TEAM_PITCHING_SO"] <- knn[impute_me,"TEAM_PITCHING_SO"]
+impute_me <- is.na(training$TEAM_FIELDING_DP)
+training[impute_me,"TEAM_FIELDING_DP"] <- knn[impute_me,"TEAM_FIELDING_DP"]
 ```

### 2019-09-29 DATA621/Week 03 Textbook Exercises.Rmd
Updated Week 3 Exercises

---
title: "DATA 621 Week #3 Textbook Exercises"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r knitr_init, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(kableExtra)
library(gridExtra)
library(ggthemes)
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)

# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})

knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

## MARR 2.3

The manager of the purchasing department of a large company would like to develop a regression model to predict the average amount of times it takes to process a given number of invoices.  Over a 30-day period, data are collected on the number of invoices processed and the total time taken (in hours). The data are avialable in the file inovices.txt. The following modle was fit to the data $Y = \beta_0 + \beta_1 x + e$ where $Y$ is the processing time and $x$ is the number of invoices.  Utilizing the output of the fit of theis model provided below, complete the following tasks.

```{r comment=NA}
invoices <- read.delim('./data/invoices.txt')
fit <- lm(Time ~ Invoices, invoices)
summary(fit)
```

a) Find a 95% confidence interval for the start-up time, i.e., $\beta_0$.

```{r}
confint(fit, "(Intercept)")
```

b) Suppose that a best practice benchmark for the average processing time for an additonal invoice is 0.01 hours (or 0.6 minutes).  Test the null hypothesis $H_0:\beta_1 = 0.01$ against a two-sided alternative.  Interpret your results.

```{r, null_prefix=TRUE}
# T statistic
abs(coef(summary(fit))[2, 1] - 0.01 / coef(summary(fit))[2, 2])
```

since the t is greater than  2.048 we reject the null hypothsis.  This suggests that the the observed invoice processing time is not equal to the best practice time.

c) Find a point estimate and a 95% prediction interval for the itme taken to process 130 invoices.

```{r, null_prefix=TRUE}
predict(fit, data.frame(Invoices = 130), interval="predict") 
```

## LMR 3.6

Thirty-nine MBA students were asked about hapiness and how it relates to their income and social life.  the data is found in `happy`.  Fit a regression model with `happy` as the response and the other four variables as predictors.

```{r, comment=NA}
happy <- read.delim('./data/happy.txt')
fit <- lm(happy ~ ., data = happy)
summary(fit)
```

a) What predictors were statistically significant at  the 1% level? **money**

b) Use the `table` function to produce a numerical summary of the response.  What assumption used to preform the t-tests seems questionable in light of this summary?  **Normally distributed residuals**

```{r, eval=FALSE}
table(happy$happy)
```

```{r}
happy %>%
  group_by(happy) %>%
  tally() %>%
  kable() %>%
  kable_styling()
```

c) Use the permutation process described in section 3.3 to test the significance of the `money` predictor.

```{r, null_prefix=TRUE}
nreps <- 4000
set.seed(123)
t_stats <- numeric(nreps)
p_values <- numeric(nreps)
for(i in 1:nreps){
  fit <- lm(sample(happy) ~ ., data = happy)
  t_stats[i] <- summary(fit)[["coefficients"]][2, "t value"]
  p_values[i] <- summary(fit)[["coefficients"]][2, "Pr(>|t|)"]

}

mean(p_values)
```

d) Plot a histogram of the permutation t-statistics.  Make sure you use the probability rather than the frequency version of the histogram.

```{r}
hist(t_stats, freq = FALSE)
```

e)  Overlay an appropriate t-densisty over the histogram.  Hint: Use `grid <- seq(-3, 3, length = 300)` to create a grid of values, then use the `dt` function to compute the t-density on this grid and the `lines` function to superimpose the result.

```{r}
grid <- seq(-3, 3, length = 300)
hist(t_stats, freq = FALSE)
lines(grid, dt(grid, 34))
```

f) Use the bootstrap procedure from Section 3.6 to compute 90% and 95% confidence intervals for $\beta_\text{money}$.  Does zero fall within these confidence intervals?  Are the results consistent with previous tests?

```{r}
set.seed(123)
nb <- 4000
bootstrap_coefficients <- matrix(NA, nb, length(coef(fit)))
for (i in 1:nb){
  bfit <- lm(sample(happy) ~ ., data = happy)
  bootstrap_coefficients[i,] <- coef(bfit)
}

bootstrap_coefficients <- as.data.frame(bootstrap_coefficients)
names(bootstrap_coefficients) <- names(coef(bfit))
```

```{r}
# 95% CI
apply(bootstrap_coefficients, 2, function(x) quantile(x, c(0.025, 0.975)))
```

```{r}
# 90% CI
apply(bootstrap_coefficients, 2, function(x) quantile(x, c(0.05, 0.95)))
```

**Zero falls within the CI's.  This is consistent with the other results.**

## LMR 4.5

## LMR 5.2

## LMR 14.2

### 2019-10-04 DATA608/app.R
Initial commit

@@ -1,0 +1,86 @@
+library(shiny)
+library(dplyr)
+library(ggplot2)
+library(plotly)
+#library(usmap)
+#library(gridExtra)
+#library(socviz)
+#library(ggthemes)
+
+xwalk <- read.csv("data/xwalk.csv", colClasses=c("state_fips"="character"))
+
+df <- read.csv("data/cleaned-cdc-mortality-1999-2010-2.csv") %>%
+    merge(xwalk)
+
+# 2010 Ranking options
+options <- filter(df, Year == 2010) %>%
+    mutate(ICD.Chapter = as.character(ICD.Chapter))
+options <- unique(sort(options$ICD.Chapter))
+
+
+ui <- navbarPage("Cause of Death",
+                 tabPanel("2010 Ranking",
+                          selectInput("cause_of_death", "Cause of Death:", options, width = "100%"),
+                          mainPanel(
+                              plotlyOutput("rank_plot", height = "600px")
+                            )
+                          )
+                 )
+
+server <- function(input, output) {
+    
+    # change_over_time <- reactive({
+    #     temp <- df %>%
+    #         filter(ICD.Chapter == cause_of_death)
+    #     
+    #     us <- temp %>%
+    #         group_by(Year) %>%
+    #         summarise(Deaths = sum(Deaths),
+    #                   Population = sum(Population)) %>%
+    #         mutate(ICD.Chapter = input$cause_of_death,
+    #                State = "US",
+    #                Crude.Rate = round(Deaths / Population * 100000, 1)) %>%
+    #         select(ICD.Chapter, State, Year, Deaths, Population, Crude.Rate)
+    #     
+    #     return(bind_rows(temp, us))
+    # })
+    
+    output$rank_plot <- renderPlotly({
+        temp <- df %>%
+            filter(Year == 2010 & ICD.Chapter == input$cause_of_death) %>%
+            arrange(Crude.Rate) %>%
+            rename(`Mortality Rate` = Crude.Rate,
+                   `State Name` = state_name) %>%
+            mutate(Rank = row_number())
+        n_group <- nrow(temp) / 5
+        p <- temp %>%
+            mutate(color = as.factor(ceiling(Rank / n_group))) %>%
+            ggplot(aes(y = `Mortality Rate`, x = Rank, text = `State Name`,  fill = color)) +
+            scale_fill_brewer(palette = "Set1") +
+            geom_bar(stat = "identity")+
+            #geom_text(aes(label=Crude.Rate), color="white", size=2, nudge_x = 2)+
[…]

### 2019-10-04 DATA608/app.R
Updated data608 shiny app

@@ -22,18 +22,18 @@ ui <- navbarPage("Cause of Death",
                  tabPanel("2010 Ranking",
                           selectInput("cause_of_death", "Cause of Death:", options, width = "100%"),
                           mainPanel(
-                              plotlyOutput("rank_plot", height = "600px")
-                            )
+                              plotlyOutput("rank_plot", height = "700px")
+                            , width = 12)
                           )
                  )
 
 server <- function(input, output) {
     
     # change_over_time <- reactive({
-    #     temp <- df %>%
+    #     plot_df <- df %>%
     #         filter(ICD.Chapter == cause_of_death)
     #     
-    #     us <- temp %>%
+    #     us <- plot_df %>%
     #         group_by(Year) %>%
     #         summarise(Deaths = sum(Deaths),
     #                   Population = sum(Population)) %>%
@@ -42,34 +42,37 @@ server <- function(input, output) {
     #                Crude.Rate = round(Deaths / Population * 100000, 1)) %>%
     #         select(ICD.Chapter, State, Year, Deaths, Population, Crude.Rate)
     #     
-    #     return(bind_rows(temp, us))
+    #     return(bind_rows(plot_df, us))
     # })
     
     output$rank_plot <- renderPlotly({
-        temp <- df %>%
+        # Wrangle the data
+        plot_df <- df %>%
             filter(Year == 2010 & ICD.Chapter == input$cause_of_death) %>%
             arrange(Crude.Rate) %>%
             rename(`Mortality Rate` = Crude.Rate,
                    `State Name` = state_name) %>%
             mutate(Rank = row_number())
-        n_group <- nrow(temp) / 5
-        p <- temp %>%
-            mutate(color = as.factor(ceiling(Rank / n_group))) %>%
-            ggplot(aes(y = `Mortality Rate`, x = Rank, text = `State Name`,  fill = color)) +
+        n_group <- nrow(plot_df) / 5
+        plot_df <- plot_df %>%
+            mutate(color = as.factor(ceiling(Rank / n_group))) 
+        # This is to help position the text
+        text_y <- max(plot_df$`Mortality Rate`) / - 100
+        # Create the ggplot
+        p <- ggplot(plot_df, aes(y = `Mortality Rate`, x = Rank, text = paste("State:", `State Name`),  fill = color)) +
             scale_fill_brewer(palette = "Set1") +
             geom_bar(stat = "identity")+
-            #geom_text(aes(label=Crude.Rate), color="white", size=2, nudge_x = 2)+
[…]

### 2019-10-04 DATA621/Week 06 Textbook Exercises.Rmd
Updated Week 06 Textbook Exercises

@@
 library(gridExtra)
+library(faraway)
 library(ggthemes)
@@
 options(max.print="75")
-opts_chunk$set(echo=FALSE,
-	             cache=TRUE,
-               prompt=FALSE,
-               tidy=TRUE,
+opts_chunk$set(echo=TRUE,
                comment=NA,
@@
 opts_knit$set(width=75)
+# A Prefix nulling hook.
+
+# Make sure to keep the default for normal processing.
+default_output_hook <- knitr::knit_hooks$get("output")
+
+# Output hooks handle normal R console output.
+knitr::knit_hooks$set( output = function(x, options) {
+
+  comment <- knitr::opts_current$get("comment")
+  if( is.na(comment) ) comment <- ""
+  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
+                     x, perl = TRUE)
+  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
+  if( can_null && do_null ) {
+    # By default R print output aligns at the right brace.
+    align_index <- regexpr( "\\]", x )[1] - 1
+    # Two cases: start or newline
+    re <- paste0( "^.{", align_index, "}\\]")
+    rep <- comment
+    x <- gsub( re, rep,  x )
+    re <- paste0( "\\\n.{", align_index, "}\\]")
+    rep <- paste0( "\n", comment )
+    x <- gsub( re, rep,  x )
+  }
+
+  default_output_hook( x, options )
+
+})
+
+knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
 ```
@@
 
+Using the `divusa` data:
+
+a) Fit a regression model with `divorce` as the response and `unemployed`, `femlab`, `marriage`, `birth` and `military` as predictors. Compute the condition numbers and interpret their meanings.
+
+```{r, null_prefix=TRUE}
+divusa <- read.delim("data/divusa.txt")
+fit <- lm(divorce ~ unemployed + femlab + marriage + birth + military, data = divusa)
+summary(fit)
+x <- model.matrix(fit)[,-1]
+e <- eigen(t(x) %*% x)
+sqrt(e$val[1]/e$val)
+```
+
+b) For the same model, compute the VIFs.  Is there evidence that collinearity causes some predictors not to be significant? Explain.
+
+```{r, null_prefix=TRUE}
+vif(x)
+```
+
+c) Does the removal of insignificant predictors from the model reduce the collinearity? Investigate.
+
+```{r, null_prefix=TRUE}
+fit <- lm(divorce ~ femlab + marriage + birth, data = divusa)
+summary(fit)
+x <- model.matrix(fit)[,-1]
+e <- eigen(t(x) %*% x)
+sqrt(e$val[1]/e$val)
+vif(x)
+```
+
 ## LMR 7.8

### 2019-10-05 DATA608/app.R
Updated app.R

@@ -2,6 +2,7 @@ library(shiny)
 library(dplyr)
 library(ggplot2)
 library(plotly)
+library(scales)
 #library(usmap)
 #library(gridExtra)
 #library(socviz)
@@ -19,7 +20,7 @@ options <- unique(sort(options$ICD.Chapter))
 
 
 ui <- navbarPage("Cause of Death",
-                 tabPanel("2010 Ranking",
+                 tabPanel("States by Rank in 2010",
                           selectInput("cause_of_death", "Cause of Death:", options, width = "100%"),
                           mainPanel(
                               plotlyOutput("rank_plot", height = "700px")
@@ -28,22 +29,6 @@ ui <- navbarPage("Cause of Death",
                  )
 
 server <- function(input, output) {
-    
-    # change_over_time <- reactive({
-    #     plot_df <- df %>%
-    #         filter(ICD.Chapter == cause_of_death)
-    #     
-    #     us <- plot_df %>%
-    #         group_by(Year) %>%
-    #         summarise(Deaths = sum(Deaths),
-    #                   Population = sum(Population)) %>%
-    #         mutate(ICD.Chapter = input$cause_of_death,
-    #                State = "US",
-    #                Crude.Rate = round(Deaths / Population * 100000, 1)) %>%
-    #         select(ICD.Chapter, State, Year, Deaths, Population, Crude.Rate)
-    #     
-    #     return(bind_rows(plot_df, us))
-    # })
     
     output$rank_plot <- renderPlotly({
         # Wrangle the data
@@ -58,12 +43,15 @@ server <- function(input, output) {
             mutate(color = as.factor(ceiling(Rank / n_group))) 
         # This is to help position the text
         text_y <- max(plot_df$`Mortality Rate`) / - 100
+        max_rank <- max(plot_df$Rank) - 2
         # Create the ggplot
-        p <- ggplot(plot_df, aes(y = `Mortality Rate`, x = Rank, text = paste("State:", `State Name`),  fill = color)) +
+        p <- ggplot(plot_df, aes(x = Rank, y = `Mortality Rate`, text = paste("State:", `State Name`),  fill = color)) +
             scale_fill_brewer(palette = "Set1") +
             geom_bar(stat = "identity")+
+            scale_x_continuous(limits=c(2, max_rank), oob = rescale_none) +
             theme_minimal() +
             theme(legend.position = "none", 
+                  panel.grid.major.y = element_blank(),
                   axis.title.y = element_blank(),
[…]

### 2019-10-05 DATA621/Week 03 Textbook Exercises.Rmd
Worked on DATA 621 HW 3

@@
 options(max.print="75")
-opts_chunk$set(echo=FALSE,
+opts_chunk$set(echo=TRUE,
 	             cache=FALSE,
@@
 
+For the `fat` data used in this chapter, a smaller model using only `age`, and `weight`, `height` and `abdom` was proposed on the grounds that these predictors are either known by the individual or easily measured.
+
+a. Compare this model to the full thirteen predictor model used earlier in the chapter.  Is it justifiable to use the smaller model?
+
+```{r}
+fat <- read.delim('./data/fat.txt') 
+full <- lm(brozek ~ age + weight + height + neck + chest + abdom + hip + thigh + knee + ankle + biceps + forearm + wrist, data = fat)
+small <- lm(brozek ~ age + weight + height + abdom, data = fat)
+summary(small)
+```
+
+b.
+Compute a 95% prediction interval for median predictor values and compare to the results to the interval for the full model.
+Do the intervals differ by a practically important amount?
+**No it doesn't differ.
+See below:**
+
+### Small Model
+
+```{r}
+x0 <- data.frame(t(apply(model.matrix(full), 2, median)))
+predict(small, x0, interval = "prediction")
+```
+
+### Full Model
+
+```{r}
+predict(full, x0, interval = "prediction")
+```
+
+c. For the smaller model, examine all the observations from case number 25 to 50.  Which two observations are particularly anomalous? **Case 25 and 33.**
+
+```{r}
+data.frame(actual = fat[25:50, "brozek"], yhat = round(predict(small, fat)[25:50], 1)) %>% 
+  rownames_to_column() %>%
+  mutate(error = abs((yhat - actual)/yhat)) %>% 
+  arrange(desc(error)) %>%
+  top_n(5) %>%
+  kable() %>% 
+  kable_styling()
+```
+
+d.
+Recompute the 95% prediction interval for the median predictors after these two anomoulous cases have been excluded from the data.
+Did this make much difference to the outcome?
+**It certainly did.
+See below:**
+
+```{r}
+fat <- fat[!row.names(fat) %in% c(25,33),]
+
+full <- lm(brozek ~ age + weight + height + neck + chest + abdom + hip + thigh + knee + ankle + biceps + forearm + wrist, data = fat)
+small <- lm(brozek ~ age + weight + height + abdom, data = fat)
+
+x0 <- data.frame(t(apply(model.matrix(full), 2, median)))
+```
+
+### Small Model
+
+```{r}
+predict(small, x0, interval = "prediction")
+```
+
+### Full Model
+
+```{r}
+predict(full, x0, interval = "prediction")
+```
+
 ## LMR 5.2
 
+Use the `odor` dataset with odor as the response variable and `temp` as the predictor.
+Consider all possible models that also include all, some or none of the other two predictors.
+Report the coefficient for temperature, its standard error, t-statistic and p-value in each case.
+Discuss what stays the same, what changes and why.
+Which model is the best?
+
+```{r}
+odor <- read.delim("data/odor.txt")
+models <- c("odor ~ temp", "ordor ~ temp + gas", "ordor ~ temp + pack", "ordor ~ temp + gas + pack")
+coefficients <- matrix(NA, 4, 6)
+
+fit <- lm(odor ~ temp, data = odor)
+coefficients[1,] <- c("odor ~ temp", summary(fit)$coefficients[2,], summary(fit)$adj.r.squared)
+fit <- lm(odor ~ temp + gas, data = odor)
+coefficients[2,] <- c("odor ~ temp + gas", summary(fit)$coefficients[2,], summary(fit)$adj.r.squared)
+fit <- lm(odor ~ temp + pack, data = odor)
+coefficients[3,] <- c("odor ~ temp + pack", summary(fit)$coefficients[2,], summary(fit)$adj.r.squared)
+fit <- lm(odor ~ temp + gas + pack, data = odor)
+coefficients[4,] <- c("odor ~ temp + gas + pack", summary(fit)$coefficients[2,], summary(fit)$adj.r.squared)
+coefficients <- as.data.frame(coefficients)
+names(coefficients) <- c("Formula", names(summary(fit)$coefficients[2,]), "Adjusted R-squared")
+coefficients %>%
+  kable() %>%
+  kable_styling()
+```
+
 ## LMR 14.2

### 2019-10-05 DATA608/app.R
Updated app.R

@@ -13,6 +13,10 @@ xwalk <- read.csv("data/xwalk.csv", colClasses=c("state_fips"="character"))
 df <- read.csv("data/cleaned-cdc-mortality-1999-2010-2.csv") %>%
     merge(xwalk)
 
+# us <- read.csv("data/us-cdc-mortality-1999-2017.csv") %>%
+#     filter(Year <= max(df$Year)) %>%
+#     mutate(Crude.Rate = as.numeric(as.character(Crude.Rate)))
+
 # 2010 Ranking options
 options <- filter(df, Year == 2010) %>%
     mutate(ICD.Chapter = as.character(ICD.Chapter))
@@ -22,11 +26,14 @@ options <- unique(sort(options$ICD.Chapter))
 ui <- navbarPage("Cause of Death",
                  tabPanel("States by Rank in 2010",
                           selectInput("cause_of_death", "Cause of Death:", options, width = "100%"),
-                          mainPanel(
-                              plotlyOutput("rank_plot", height = "700px")
-                            , width = 12)
-                          )
+                          mainPanel(plotlyOutput("rank_plot", height = "700px"), width = 12)
+                          ),
+                 tabPanel("Improvement vs National Average",
+                          selectInput("cause_of_death_2", "Cause of Death:", options, width = "100%"),
+                          #uiOutput("state_options"),
+                          mainPanel(plotlyOutput("improvement_plot"), width = 12)
                  )
+)
 
 server <- function(input, output) {
     
@@ -58,6 +65,65 @@ server <- function(input, output) {
             labs(y = "Deaths per 100,000", x = "") +
             geom_text(data = plot_df, aes(x = Rank, y = text_y, label = State), color = "black", size = 2) + 
             coord_flip()
+        # Load it into plotly
+        ggplotly(p) %>% 
+            config(displayModeBar = F) %>%
+            layout(margin = list(t = 0))
+    })
+    
+    output$state_options <- reactive({
+        options <- df %>%
+            filter(ICD.Chapter == input$cause_of_death_2) %>%
+            select(state_name)
+        options <- unique(sort(options$state_name))
+        
+        selectInput("highlight", "Highlight:", options, width = "100%")
+    })
+    
+    output$improvement_plot <- renderPlotly({
+        # Wrangle the data
+        plot_df <- df %>%
+            merge(xwalk) %>%
+            filter(ICD.Chapter == input$cause_of_death_2)
+        
+        n_years <- plot_df %>%
[…]

### 2019-10-05 DATA608/app.R
Updated app.R

@@ -13,9 +13,7 @@ xwalk <- read.csv("data/xwalk.csv", colClasses=c("state_fips"="character"))
 df <- read.csv("data/cleaned-cdc-mortality-1999-2010-2.csv") %>%
     merge(xwalk)
 
-# us <- read.csv("data/us-cdc-mortality-1999-2017.csv") %>%
-#     filter(Year <= max(df$Year)) %>%
-#     mutate(Crude.Rate = as.numeric(as.character(Crude.Rate)))
+state_options <- sort(df$state_name)
 
 # 2010 Ranking options
 options <- filter(df, Year == 2010) %>%
@@ -30,7 +28,7 @@ ui <- navbarPage("Cause of Death",
                           ),
                  tabPanel("Improvement vs National Average",
                           selectInput("cause_of_death_2", "Cause of Death:", options, width = "100%"),
-                          #uiOutput("state_options"),
+                          selectInput("highlight", "Highlight:", state_options, width = "100%"),
                           mainPanel(plotlyOutput("improvement_plot"), width = 12)
                  )
 )
@@ -71,15 +69,6 @@ server <- function(input, output) {
             layout(margin = list(t = 0))
     })
     
-    output$state_options <- reactive({
-        options <- df %>%
-            filter(ICD.Chapter == input$cause_of_death_2) %>%
-            select(state_name)
-        options <- unique(sort(options$state_name))
-        
-        selectInput("highlight", "Highlight:", options, width = "100%")
-    })
-    
     output$improvement_plot <- renderPlotly({
         # Wrangle the data
         plot_df <- df %>%
@@ -110,14 +99,14 @@ server <- function(input, output) {
             select(State, base_rate) %>%
             merge(plot_df) %>%
             mutate(Index = round(Crude.Rate / base_rate * 100, 0),
-                   color = ifelse(State == "US", "red", "gray")) %>%
+                   color = ifelse(State == "US", "2", "1")) %>%
+            mutate(color = ifelse(state_name == input$highlight, "3", color)) %>%
             rename(abbr = State,
                    State = state_name,
                    `Mortality Rate` = Crude.Rate)
         plot_df$color <- as.factor(plot_df$color)
         p <- ggplot(plot_df) +
-            geom_line(aes(x = Year, y = Index, group = State, color = color)) +
-            geom_point(aes(x = Year, y = Index, group = State, color = color, text = paste("Mortality Rate:", `Mortality Rate`))) +
[…]

### 2019-10-07 DATA608/app.R
Update DATA608 app.R

@@ -3,16 +3,9 @@ library(dplyr)
 library(ggplot2)
 library(plotly)
 library(scales)
-#library(usmap)
-#library(gridExtra)
-#library(socviz)
-#library(ggthemes)
 
 xwalk <- read.csv("data/xwalk.csv", colClasses=c("state_fips"="character"))
-
-df <- read.csv("data/cleaned-cdc-mortality-1999-2010-2.csv") %>%
-    merge(xwalk)
-
+df <- read.csv("data/cleaned-cdc-mortality-1999-2010-2.csv") %>% merge(xwalk)
 state_options <- sort(df$state_name)
 
 # 2010 Ranking options
@@ -20,12 +13,13 @@ options <- filter(df, Year == 2010) %>%
     mutate(ICD.Chapter = as.character(ICD.Chapter))
 options <- unique(sort(options$ICD.Chapter))
 
-
 ui <- navbarPage("Cause of Death",
+                 
                  tabPanel("States by Rank in 2010",
                           selectInput("cause_of_death", "Cause of Death:", options, width = "100%"),
                           mainPanel(plotlyOutput("rank_plot", height = "700px"), width = 12)
                           ),
+                 
                  tabPanel("Improvement vs National Average",
                           selectInput("cause_of_death_2", "Cause of Death:", options, width = "100%"),
                           selectInput("highlight", "Highlight:", state_options, width = "100%"),
@@ -40,17 +34,19 @@ server <- function(input, output) {
         plot_df <- df %>%
             filter(Year == 2010 & ICD.Chapter == input$cause_of_death) %>%
             arrange(Crude.Rate) %>%
-            rename(`Mortality Rate` = Crude.Rate,
-                   `State Name` = state_name) %>%
             mutate(Rank = row_number())
+        
         n_group <- nrow(plot_df) / 5
+        
         plot_df <- plot_df %>%
             mutate(color = as.factor(ceiling(Rank / n_group))) 
+        
         # This is to help position the text
-        text_y <- max(plot_df$`Mortality Rate`) / - 100
+        text_y <- max(plot_df$Crude.Rate) / - 100
         max_rank <- max(plot_df$Rank) - 2
+        
         # Create the ggplot
-        p <- ggplot(plot_df, aes(x = Rank, y = `Mortality Rate`, text = paste("State:", `State Name`),  fill = color)) +
[…]

### 2019-10-10 DATA621/HW2.Rmd
Initial commit of DATA621 HW 2 files

---
title: "DATA 621 Homework #2"
subtitle: "From Work by Critical Thinking Group 3"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(caret)
library(pROC)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv("./data/classification-output-data.csv")
actual <- "class"
predicted <- "scored.class"
probability <- "scored.probability"
# Load our functions
source("hw2.R")
```

## Caret's Output

```{r}
confusionMatrix(factor(df[[predicted]]), factor(df[[actual]]))
```

## Our Function's Output

**Confusion Matrix**

```{r}
confusion_matrix(df, actual, predicted) %>%
  kable() %>%
  kable_styling()
```

**Accuracy**: `r accuracy(df, actual, predicted)`

**Classification Error Rate**: `r classification_error_rate(df, actual, predicted)`

**Precision**: `r precision(df, actual, predicted)`

**Sensitivity (Recall)**: `r sensitivity(df, actual, predicted)`

**Specificity**: `r specificity(df, actual, predicted)`

**F1 Score**: `r f1_score(df, actual, predicted)`

### F1 Score Will Always be Between 0 and 1
Before we move on, let’s consider a question that was asked: What are the bounds on the F1 score? Show that the F1 score will always be between 0 and 1. (Hint: If 0<a<1 and 0<b<1 then ab<a).

**Proving the range of the F1 Score must be between 0 and 1**

**Let's expand the F1 equation using the definitions of precision and sensitivity.**

$$\textrm{F1 Score}= \frac{2 \times\mathrm{Precision}\times\mathrm{Sensitivity}}{\mathrm{Precision}+\mathrm{Sensitivity}}$$ 
**The numerator is equal to**
$$=\frac{2 \times TP \times TP}{(TP + FP)(TP+FN)}$$ 
**The denominator is equal to**
$$= \frac{1}{\frac{TP}{TP+FP}+\frac{TP}{TP+FN}}$$  
**Combine the Numerator and the denominator:**
$$F1=\frac{2 \times TP \times TP}{(TP + FP)(TP+FN)} \times \frac{1}{\frac{TP}{TP+FP}+\frac{TP}{TP+FN}}$$  
**Multiply the denominator by (TP + FP)(TP + FN):**

$$denominator = \frac{(TP+FP)(TP+FN)}{TP(TP+FN)+TP(TP+FP)}$$  
$$F1=\frac{2(TP)^2}{(TP+FP)(TP+FN)}\times \frac{(TP+FP)(TP+FN)}{TP(TP+FN)+TP(TP+FP)}$$  
**Cancel the denominator of the first term with the numerator of the second term:**

$$F1 =\frac{2(TP)^2}{TP(TP+FN)+TP(TP+FP)}$$
**Simplify the Denominator:**
$$F1 =\frac{2(TP)^2}{TP^2+FNTP+TP^2+TPFP}$$  
$$F1 =\frac{2(TP)^2}{TP(TP+FN+TP+FP)}$$  
$$F1 =\frac{2(TP)^2}{TP(TP+FN+TP+FP)}$$  
$$F1 =\frac{2(TP)^2}{TP(2TP +FN +FP)}$$ 
**Cancel the TP terms and get final answer**
$$F1=\frac{2TP}{2TP+FN+FP}$$
<br>  
In the equation above, $TP,FN,FP \in \mathbb{N}$ where $\mathbb{N} = \{0,1,2,3...\}$

$2TP \le2TP+FN+FP$ this equation shows that the numerator is at most equal to the denominator (i.e $F1\le1$).

There would be a Maximum value of $1$ when $FN=FP=0$ and $TP > 0$

The fraction $=\frac{2TP}{2TP+FN+FP}$ has a minimum value of zero when the numerator is minimized (i.e. $TP=0$).

We have shown that $0\le F1\le1$
<br>

## pROC Output

```{r}
par(pty = "s")
roc(df[[actual]], df[[probability]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

## Our ROC Curve

```{r}
our_roc <- roc_curve(df, actual, probability)

our_roc$plot
```

```{r}
our_roc$auc_vector
```

## Source Code

```{r source, code=readLines("hw2.R")}
```

### 2019-10-10 DATA621/hw2.R

@@ -1,0 +1,130 @@
+library(dplyr)
+library(ggplot2)
+
+confusion_matrix <- function(df, actual, predicted){
+  cm <- table(df[[predicted]], df[[actual]])
+  if (nrow(cm) == ncol(cm)){
+    return(cm)
+  }
+  # Make sure all missing values are present in the confusion matrix
+  fixed_cm <- matrix(0, ncol(cm), ncol(cm))
+  colnames(fixed_cm) <- colnames(cm)
+  rownames(fixed_cm) <- colnames(cm)
+  # Horribly inefficient at large scale but since we're dealing with a
+  # confusion matrix...
+  for (r in rownames(cm)){
+    for (c in colnames(cm)){
+      fixed_cm[r, c] <- cm[r, c]
+    }
+  }
+  return(fixed_cm)
+}
+
+confusion_matrix_outcomes <- function(cm){
+  # This will obviously need some work to handle more than a 2x2 matrix
+  TP <- cm[1]
+  FN <- cm[2]
+  FP <- cm[3]
+  TN <- cm[4]
+  list("TN" = TN, "FP" = FP, "FN" = FN, "TP" = TP)
+}
+
+# Write a function that takes the data set as a dataframe, with actual and 
+# predicted classifications identified, and returns the accuracy of the 
+# predictions.
+
+accuracy <- function(df, actual, predicted){
+  cm <- confusion_matrix_outcomes(confusion_matrix(df, actual, predicted))
+  (cm$TP + cm$TN) / (cm$TP + cm$FP + cm$TN + cm$FN)
+}
+
+# Write a function that takes the data set as a dataframe, with actual and 
+# predicted classifications identified, and returns the classification error 
+# rate of the predictions.
+
+classification_error_rate <- function(df, actual, predicted){
+  cm <- confusion_matrix_outcomes(confusion_matrix(df, actual, predicted))
+  (cm$FP + cm$FN) / (cm$TP + cm$FP + cm$TN + cm$FN)
+}
+
+# Write a function that takes the data set as a dataframe, with actual and 
+# predicted classifications identified, and returns the precision of the 
+# predictions.
+
+precision <- function(df, actual, predicted){
+  cm <- confusion_matrix_outcomes(confusion_matrix(df, actual, predicted))
+  cm$TP / (cm$TP + cm$FP)
+}
+
+# Write a function that takes the data set as a dataframe, with actual and 
+# predicted classifications identified, and returns the sensitivity of the 
[…]

### 2019-10-31 DATA621/HW3.Rmd
Initial commit of DATA 621 HW3

---
title: "DATA 621 Homework #3"
subtitle: "From Work by Critical Thinking Group 3"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(corrplot)
library(caret)
library(Amelia)
library(pROC)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv('data/crime-training-data_modified.csv')
evaluation <- read.csv("data/crime-evaluation-data_modified.csv")
```

## Data Exploration

### Are There Missing Values?
First we look at the data to see if any variables have missing data:

```{r}
missmap(df, main = "Missing vs Observed Values")
```

It looks like we have a complete data set.  No need to impute values.

### Splitting the Data
Next, we look to split our data between a training set (`train`) and a test data set (`test`). We'll use a 70-30 split between train and test, respectively.
```{r}
set.seed(42)
train_index <- createDataPartition(df$target, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

### Exploratory Data Analysis
Now, let's look at our training data. By looking at a correlation matrix, we can see which variables may be too correlated to be included together in a model as predictor variables. This will help us later during the model selection process.

```{r}
train %>% 
  cor(.) %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", diag = FALSE)
```

Next we will look at each potential predictor and how it is distributed across the target variable.

The following plots show how predictors are distributed between a positive target variable (areas with crime rates higher than the median, i.e. blue) and a negative target variable (areas with crime rates below the median, i.e. red). What we are looking for is variables that show way to split data into two groups.

```{r}
for (var in names(train)){
  if(var != "target"){
    plot_df <- train
    plot_df$x <- plot_df[,var]
    p <- ggplot(plot_df, aes(x, color = factor(target))) +
      geom_density() +
      theme_light() +
      ggtitle(var) +
      scale_color_brewer(palette = "Set1") +
      theme(legend.position = "none",
            axis.title.y = element_blank(),
            axis.title.x = element_blank())
    print(p)
  }
}
```

Looking at the plots above, the `nox` variable seems to be the best variable to divide the data into the two groups.

## Basic Model Building

We start by applying occam's razor and create a baseline model that only has one predictor. Any model we build beyond that will have to outperform this simplest model.

```{r}
baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
summary(baseline)
test$baseline <- ifelse(predict.glm(baseline, test,"response") >= 0.5,1,0)
cm <- confusionMatrix(factor(test$baseline), factor(test$target),"1")
results <- tibble(model = "baseline",predictors = 1,F1 = cm$byClass[7],
                  deviance=baseline$deviance, 
                  r2 = 1 - baseline$deviance/baseline$null.deviance,
                  aic=baseline$aic)
cm
```

Our baseline model is ok with an F1 score of `r cm$byClass[7]`.

Next we try adding every other variable, to build a full model. From here we can work backwards and eliminate non-significant predictors:

```{r}
fullmodel <- glm(target ~ ., family = binomial(link = "logit"), train)
summary(fullmodel)
test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
results <- rbind(results,tibble(model = "fullmodel",
                                predictors = 12,F1 = cm$byClass[7],
                                deviance=fullmodel$deviance, 
                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
                                aic=fullmodel$aic))
cm
```

The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, many variables do not seem to be significant.

After some backward elimination of non-significant predictor variables, we arrive at the following model:

```{r}
model1 <- glm(target ~ . -tax -rm -chas - age -zn -indus, 
              family = binomial(link = "logit"), 
              train)
summary(model1)
test$model1 <- ifelse(predict.glm(model1, test,"response") < 0.5, 0, 1)
cm <- confusionMatrix(factor(test$model1), factor(test$target),"1")
results <- rbind(results,tibble(model = "model1",
                                predictors = 6,F1 = cm$byClass[7],
                                deviance=model1$deviance, 
                                r2 = 1-model1$deviance/model1$null.deviance,
                                aic=model1$aic))
cm
```

## Alternative Models

### Feature Engineering
In this portion below we explored using the difference between the class distributions as a feature itself. So instead of the raw feature itself the we created density features that are essientially the differences in probability of the positve and negative classes at that certain point for the feature.

```{r}
library(fuzzyjoin)
density_diff <- function(data,eval_data,var,target,new_column){
  data[[paste0(var,"temp")]] <- data[[var]]
  eval_data[[paste0(var,"temp")]] <- eval_data[[var]]

  #standardize variable between 0 and 1
  data[[var]] <- (data[[var]] - min(data[[var]]))/(max(data[[var]]) - min(data[[var]]))
  eval_data[[var]] <- (eval_data[[var]] - min(eval_data[[var]]))/(max(eval_data[[var]]) - min(eval_data[[var]]))

  ## Calculate density estimates
  g1 <- ggplot(data, aes(x=data[[var]], group=target, colour=target)) +
    geom_density(data = data) + xlim(min(data[[var]]), max(data[[var]]))
  gg1 <- ggplot_build(g1)
  # construct the dataset
  x <- gg1$data[[1]]$x[gg1$data[[1]]$group == 1]
  y1 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 1]
  y2 <- gg1$data[[1]]$y[gg1$data[[1]]$group == 2]
  df2 <- data.frame(x = x, ymin = pmin(y1, y2), ymax = pmax(y1, y2), side=(y1<y2), ydiff = y2-y1)
  ##creating the second graph object
  g3 <- ggplot(df2) +
    geom_line(aes(x = x, y = ydiff, colour = side)) +
    geom_area(aes(x = x, y = ydiff, fill = side, alpha = 0.4)) +
    guides(alpha = FALSE, fill = FALSE)

  data$join <- data[[var]]
  df2$join <- df2$x
  temp <- difference_left_join(data,df2, by="join", max_dist =.001)
  means <- aggregate(ydiff ~ temp$join.x, temp, mean)

  colnames(means) <- c("std_var","density_diff")

  new_data <- merge(means,data, by.x ="std_var", by.y =var)
  new_data$std_var <- NULL
  new_data$join <- NULL
  new_data[new_column] <- new_data$density_diff
  new_data$density_diff <- NULL
  ###################same thing with eval data ############################
  eval_data$join <- eval_data[[var]]
  df2$join <- df2$x
  temp <- difference_left_join(eval_data,df2, by="join", max_dist =.001)
  means <- aggregate(ydiff ~ temp$join.x, temp, mean)

  ##new stuff
  colnames(means) <- c("std_var","density_diff")
  #data["key"] <- (var - min(var))/(max(var) - min(var))

  eval_data2 <- merge(means,eval_data, by.x ="std_var", by.y =var)
  eval_data2$std_var <- NULL
  eval_data2$join <- NULL
  eval_data2[new_column] <- eval_data2$density_diff
  eval_data2$density_diff <- NULL

  eval_data2[[var]] <- eval_data2[[paste0(var,"temp")]]
  eval_data2[[paste0(var,"temp")]] <- NULL

  new_data[[var]] <- new_data[[paste0(var,"temp")]]
  new_data[[paste0(var,"temp")]] <- NULL

  mylist <- list(g1,g3,new_data,eval_data2)
  names(mylist)<- c("dist","dist_diff","new_data_training","new_data_eval")
  return(mylist)
}
```

### Creating Some New Variables

```{r}
str(train)
str(test)
## try it out 
train <- density_diff(train,test,"nox",train$target,"nox_density")$new_data_training
train <- density_diff(train,test,"age",train$target,"age_density")$new_data_training 
train <- density_diff(train,test,"indus",train$target,"indus_density")$new_data_training
train <- density_diff(train,test,"dis",train$target,"dis_density")$new_data_training
train <- density_diff(train,test,"rad",train$target,"rad_density")$new_data_training
train <- density_diff(train,test,"tax",train$target,"tax_density")$new_data_training
train <- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_training
train <- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_training
train <- density_diff(train,test,"medv",train$target,"medv_density")$new_data_training

test<- density_diff(train,test,"nox",train$target,"nox_density")$new_data_eval
test<- density_diff(train,test,"age",train$target,"age_density")$new_data_eval 
test<- density_diff(train,test,"indus",train$target,"indus_density")$new_data_eval
test<- density_diff(train,test,"dis",train$target,"dis_density")$new_data_eval
test<- density_diff(train,test,"rad",train$target,"rad_density")$new_data_eval
test<- density_diff(train,test,"tax",train$target,"tax_density")$new_data_eval
test<- density_diff(train,test,"ptratio",train$target,"ptratio_density")$new_data_eval
test<- density_diff(train,test,"lstat",train$target,"lstat_density")$new_data_eval
test<- density_diff(train,test,"medv",train$target,"medv_density")$new_data_eval

str(train)
str(test)
```

### Example of the new meta feature
The first graph below is the difference between the distributions and the second graph is the new derived predictive feature
```{r}
density_diff(train,test,"age",train$target,"age_density")$dist
density_diff(train,test,"age",train$target,"age_density")$dist_diff
```

With our new density variables, we can run another model:

```{r}
density_models <- glm(target ~ nox_density +indus_density+age_density+dis_density+rad_density+tax_density+ptratio_density+lstat_density+medv_density, family = binomial(link = "logit"), train)
summary(density_models)
test$density_models <- ifelse(predict(density_models, test) < 0.5, 0, 1)
test$density_models_yhat <- predict(density_models, test)
cm <- confusionMatrix(factor(test$density_models), factor(test$target))
results <- rbind(results,tibble(model = "density models",
                                predictors = 9,
                                F1 = cm$byClass[7],
                                deviance = density_models$deviance,
                                r2 = 1 - density_models$deviance / density_models$null.deviance,
                                aic = density_models$aic))
cm
```

## Model Selection

Since the assignment mentions that the purpose is prediction, we will prefer F1 score as our measure of model success.

```{r}
kable(results)
```

Looking the three models together, `model1` looks like the best one. Although the full model scored an F1 that was ever-so-slightly higher on our test data set, `model1` has half the predictors and it's scores are almost exactly the same as `fullmodel`.

### pROC Output

ROC curves can give us another look at which model might be better suited for prediction.

#### Baseline
```{r}
par(pty = "s")
roc(test[["target"]], test[["baseline"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Full Model
```{r}
par(pty = "s")
roc(test[["target"]], test[["fullmodel"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Model1
```{r}
par(pty = "s")
roc(test[["target"]], test[["model1"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Density Model 
```{r}
par(pty = "s")
roc(test[["target"]], test[["density_models_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

### 2019-10-31 DATA621/HW3.Rmd
Finalize HW3

@@
 
+## Introduction
+
+We have been given a dataset with `r nrow(df)` records summarizing a attributes of various neighborhoods of a major city.
+Each record in the training dataset has a response variable that is 1 when the neighborhood's crime rate is above the median and 0 when it is not.
+In all there are `r ncol(df)-1` predictors.
+We will construct a logistic regression model to predict the classification of the evaluation data set.
+
 ## Data Exploration
@@
 ### Splitting the Data
+
 Next, we look to split our data between a training set (`train`) and a test data set (`test`). We'll use a 70-30 split between train and test, respectively.
@@
 ### Exploratory Data Analysis
-Now, let's look at our training data.
+
+Now, let's look at our training data. The following table summarizes the predictor variables of the training set:
+
+```{r}
+train %>%
+  select(-target) %>%
+  summary() %>%
+  kable() %>%
+  kable_styling()
+```
+
 By looking at a correlation matrix, we can see which variables may be too correlated to be included together in a model as predictor variables.
@@
 ```
+
+## Final Selection and Predictions
+
+Having looked at the F1 statistics and the ROC curves we will use the model1 to form our final predictions.
+
+```{r}
+evaluation$target <- ifelse(predict.glm(model1, evaluation, "response") < 0.5, 0, 1)
+evaluation %>%
+  write.csv(., "./data/hw3-predictions.csv", row.names = F)
+```

### 2019-11-01 DATA608/module 5/app.py
Initial commit of DATA 608 Module 5

@@ -1,0 +1,74 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Oct 31 08:52:11 2019
+
+@author: Michael Silva
+"""
+
+import os
+import pandas as pd
+from flask import Flask, jsonify, request
+
+APP_ROOT = os.path.dirname(os.path.abspath(__file__))
+
+
+def get_tree_list():
+    soql_url = (
+        "https://data.cityofnewyork.us/resource/nwxe-4ae8.json?"
+        + "$select=spc_common,count(tree_id)"
+        + "&$group=spc_common"
+    ).replace(" ", "%20")
+    temp = pd.read_json(soql_url).dropna()
+    return temp.spc_common.tolist()
+
+
+app = Flask(__name__)
+
+
+@app.route("/")
+def index():
+    html = open(os.path.join(APP_ROOT, "js_in_webpage.html")).read()
+    return html
+
+
+@app.route("/api/v1/predict/<tree>")
+def predict(tree):
+    # Predict the Borough based on the probability the user is seeing a tree
+    soql_url = (
+        "https://data.cityofnewyork.us/resource/nwxe-4ae8.json?"
+        + "$select=boroname,count(tree_id)"
+        + "&$where=spc_common='"
+        + tree
+        + "'"
+        + "&$group=boroname"
+    ).replace(" ", "%20")
+    # Get the data, sort it by tree count, and clean up the columns
+    df = pd.read_json(soql_url).sort_values(by='count_tree_id', ascending=False).rename(columns={"count_tree_id": "n", "boroname": "Borough"})
+    # Get the total number of trees
+    total = df["n"].sum()
+    # Compute the probability
+    df['Probability'] = round(df['n'] / total, 2)
+    # Get the first row (highest probability)
+    df = df.drop(columns = ['n']).iloc[0]
+    json = df.to_dict()
+    # Fix The Bronx
+    if json['Borough'] == "Bronx":
+        json['Borough'] = "The Bronx"
+    
+    return jsonify(json)
+
+
+@app.route("/api/v1/trees")
+def trees():
+    soql_url = (
+        "https://data.cityofnewyork.us/resource/nwxe-4ae8.json?"
+        + "$select=spc_common,count(tree_id)"
+        + "&$group=spc_common"
+    ).replace(" ", "%20")
+    temp = pd.read_json(soql_url).dropna()
[…]

### 2019-11-16 DATA621/HW4.Rmd
DATA 621 HW #4

---
title: "DATA 621 Homework #4"
subtitle: "From Work by Critical Thinking Group 3"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 4
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
require(gridExtra)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
library(scales)
library(purrr)
library(RColorBrewer)
library(ROCR)
library(corrplot)
# Thank you Stack Overflow!
# A Prefix nulling hook.
options(scipen = 999)
# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv("./data/insurance_training_data.csv")
evaluation <- read.csv("./data/insurance-evaluation-data.csv")

strip_dollars <- function(x){
  x <- as.character(x)
  x <- gsub(",", "", x)
  x <- gsub("\\$", "", x)
  as.numeric(x)
}
```

## Introduction

We have been given a dataset with `r nrow(df)` records representing customers of an auto insurance company.  Each record has two response variables.  The first objective is to train a logistic regression classifier to predict if a person was in a car crash.  The response variable here is `TARGET_FLAG` respresenting whether a person had an accident (1) or did not have one (0).

```{r, echo=FALSE}
df %>% 
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
  ggplot(aes(x=TARGET_FLAG,fill=TARGET_FLAG)) +
  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
  theme_light() +
  theme(legend.position = "none") +
  labs(x="TARGET_FLAG", y="# Observations")
```

The second objective will be to train a regression model to predict the cost of a crash, if one occurred.  The second reponse variable, `TARGET_VALUE`, is the amount it will cost if the person crashes their car. The value is zero if the person did not crash their car.

```{r, echo=FALSE}
df %>% filter(TARGET_FLAG == 1) %>%
  ggplot(aes(x=TARGET_AMT)) + 
  geom_density() +
  geom_vline(aes(xintercept = mean(TARGET_AMT)), lty=2, col="red") +
  geom_label(aes(x=mean(TARGET_AMT),y=1,label="mu"),parse=T) +
  geom_vline(aes(xintercept = median(TARGET_AMT)), lty=2, col="darkgreen") +
  geom_label(aes(x=median(TARGET_AMT),y=.5,label="median")) +
  scale_x_log10(labels=comma) + theme_light() +
  labs(title="TARGET_AMT Density Plot", caption="x-axis is log 10 scale",
       y="Density", x="LOG(TARGET_AMT)")
```

Looking at the distribution of the `TARGET_AMT` variable, we can see that the variable is considerably right-skewed. Thus, a LOG transform might be best here.

## Data Preparation & Exploration

We will first look at the summary statistics for the data

```{r}
df %>%
  summary() %>%
  kable() %>%
  kable_styling()
```

There are some missing values that we will need to deal with. There are also some values that seem invalid (i.e. -3 CAR_AGE).

### Fix Data Types

There are some variables that are currently factors that are dollar values that need to be transformed into numeric variables.  We will do this to both the `df` and `evaluation` data frames.  There are also some invalid data that will be changed to NAs.

```{r}
strip_dollars <- function(x){
  x <- as.character(x)
  x <- gsub(",", "", x)
  x <- gsub("\\$", "", x)
  as.numeric(x)
}

fix_data_types <- function(messy_df){
  messy_df %>%
    rowwise() %>%
    mutate(INCOME = strip_dollars(INCOME),
           HOME_VAL = strip_dollars(HOME_VAL),
           BLUEBOOK = strip_dollars(BLUEBOOK),
           OLDCLAIM = strip_dollars(OLDCLAIM)) %>%
    ungroup()
}

na_bad_values <- function(messy_df){
  messy_df %>%
    rowwise() %>%
    mutate(CAR_AGE = ifelse(CAR_AGE < 0, NA, CAR_AGE))%>%
    ungroup()
}

df$TARGET_FLAG <- factor(df$TARGET_FLAG)

df <- df %>%
  fix_data_types() %>%
  na_bad_values()
evaluation <- evaluation %>%
  fix_data_types() %>%
  na_bad_values()
```

Now that we have corrected the invalid variables, we can look at a sumamry of the data:

```{r}
df %>%
  summary() %>%
  kable() %>%
  kable_styling()
```

### Fix Missing Values

There are `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable is missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, `r df %>% filter(is.na(AGE)) %>% nrow() ` observations where the variable `AGE` is missing, `r df %>% filter(is.na(INCOME)) %>% nrow() ` observations where the variable `INCOME` is missing, and `r df %>% filter(is.na(HOME_VAL)) %>% nrow() ` observations where the variable `HOME_VAL` is missing.  There are `r nrow(df) - nrow(na.omit(df))`, or `r round(((nrow(df) - nrow(na.omit(df))) / nrow(df)) * 100)`% of the observations missing varables.  We will fill in the missing data with the median value.

```{r}
fix_missing <- function(df) {
  df %>% 
    mutate_at(vars(c("CAR_AGE", "YOJ", "AGE", "INCOME", "HOME_VAL")), ~ifelse(is.na(.), median(., na.rm = TRUE), .))
}

df <- fix_missing(df)
evaluation <- fix_missing(evaluation)
```

### Feature Creation

We will create a log transformed income and home value feature.  We will also create an average claim amount that will hopefully track better with the `TARGET_AMT` variable.  We will also flag the `TARGET_AMT` observations that are outliers.  These would be cases where the vehicle is totaled and are qualitatively different than the minor accidents.

```{r}
# Function to add features

outlier <- min(boxplot(df[df$TARGET_FLAG==1,]$TARGET_AMT, plot=FALSE)$out)

create_features <- function(d){
  d %>%
    mutate(LOG_INCOME = log(INCOME + 1),
           LOG_HOME_VAL = log(HOME_VAL + 1),
           AVG_CLAIM = ifelse(CLM_FREQ > 0, OLDCLAIM / CLM_FREQ, 0),
           PRIOR_ACCIDENT = factor(ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)),
           COLLEGE_EDUCATED = factor(ifelse(EDUCATION %in% c("Bachelors", "Masters", "PhD"), 1, 0)),
           URBAN_DRIVER = factor(ifelse(URBANICITY == "Highly Urban/ Urban", 1, 0)),
           YOUNG_MALE = factor(ifelse(SEX == "M" & AGE < 25, 1, 0)),
           YOUNG = factor(ifelse(AGE < 25, 1, 0)),
           STUDENT = factor(ifelse(JOB == "Student", 1, 0)),
           MVR_PTS_GT_0 = factor(ifelse(MVR_PTS > 0, 1, 0)),
           RED_SPORTS_CAR = factor(ifelse(CAR_TYPE == "Sports Car" & RED_CAR == "yes", 1, 0)),
           HAS_KIDS = factor(ifelse(HOMEKIDS == 0, 0, 1)),
           KID_DRIVERS = factor(ifelse(KIDSDRIV == 0, 0, 1)),
           TARGET_AMT_OUTLIER = ifelse(TARGET_AMT < outlier, 0, 1)) %>%
    select(-URBANICITY)
}

df <- create_features(df)
evaluation <- create_features(evaluation)
```

### Creating Training/Test Data Sets

#### For Classifier Model

Now that we have a complete data set we will split the data into a training (`train`) and test set (`test`) for the classifier. We'll use a 70-30 split between train and test, respectively.

```{r}
set.seed(42)
train_index <- createDataPartition(df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

There are `r nrow(train[train$TARGET_FLAG == 1,])` out of `r nrow(train)` records in the training data set that have been in an accident.  We want to correct the imbalance in the dataset by over sampling this group so our classifier will do a better job identifying this minority group.

```{r}
set.seed(42)
minority <- nrow(train[train$TARGET_FLAG == 1,])
majority <- nrow(train[train$TARGET_FLAG == 0,])
diff <- majority - minority
minority_index <- train[train$TARGET_FLAG == 1,]$INDEX
over_sample_train <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
  merge(train, .) %>%
  bind_rows(train)
```

The over sampled data frame has `r nrow(over_sample_train)` records, and as you can see in the following figure is now balanced:

```{r, echo=FALSE}
over_sample_train %>% 
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
  ggplot(aes(x=TARGET_FLAG, fill=TARGET_FLAG)) +
  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
  theme_light() +
  theme(legend.position = "none") +
  labs(x="TARGET_FLAG", y="# Observations")

over_sample_train$TARGET_FLAG <- factor(over_sample_train$TARGET_FLAG)
```

#### For Linear Regression Model

Since the linear regression model's goal is to predict the cost of the claim we will subset the `r nrow(df)` records for those involved in an accident.  We will split these `r df %>% filter(TARGET_FLAG == 1) %>% nrow()` records into training and test sets using the same 70-30 split.

```{r}
set.seed(42)
accidents <- df %>%
  filter(TARGET_FLAG == 1)
amt_train_index <- createDataPartition(accidents$TARGET_AMT, p = .7, list = FALSE, times = 1)
amt_train <- accidents[amt_train_index,]
amt_test <- accidents[-amt_train_index,]
```

There are `r nrow(amt_train)` out of `r nrow(accidents)` records in the training data set.

### Exploratory Data Analysis

In exploring the data we are looking for two things.  We are looking for variables that will help use divide the data into those who have been in an accident and those who have not.  We are also looking for variables that are linearly correlated with the claim amount so it may be used as a predictor for the linear regression model.  We will be looking at both training set for the two model types.

First we will examine the numeric variables found in the oversampled classification data set:

```{r}
plot_vars <- c("TARGET_FLAG", names(keep(over_sample_train, is.numeric)))

over_sample_train[plot_vars] %>%
  select(-INDEX, -TARGET_AMT) %>%
  gather(variable, value, -TARGET_FLAG) %>%
  ggplot(., aes(TARGET_FLAG, value, color=TARGET_FLAG)) + 
  geom_boxplot() +
  scale_color_brewer(palette="Set1") +
  theme_light() +
  theme(legend.position = "none") +
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

There are a few variables that seem to have a difference between the two groups.  For example `CLM_FREQ` might be useful in differentiating.  There isn't much difference between a lot of groups which is somewhat understandable.

Next we will look at the catagorical variables in the oversampled classifcation data set.  These charts look to see if a variable can be uses to distinguish those that have had an accident (blue) from those who haven't (red).

```{r}
plot_vars <- names(keep(over_sample_train, is.factor))

temp <- over_sample_train[plot_vars] %>%
  gather(variable, value, -TARGET_FLAG) %>%
  group_by(TARGET_FLAG, variable, value) %>%
  tally()

temp %>%
  group_by(variable, value) %>%
  summarise(total = sum(n)) %>%
  merge(temp,.) %>%
  mutate(percent = n / total) %>%
  ggplot(., aes(value, percent, fill=TARGET_FLAG)) + 
  geom_col() +
  scale_fill_brewer(palette="Set1") +
  theme_light() +
  theme(legend.position = "none") +
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

The `URBAN_DRIVER` is much more likely to be in an accident than the rural driver.  The `YOUNG` and those with a `PRIOR_ACCIDENT` are also more likely to be in an accident.  There doesn't seem to be much evidence for a difference between the sexes.

Now let's focus on the relationship between these variables and the size of the claim.  As a reminder this data only includes those who have been in an accident.  For starters we will look at the distribution of the claims for those in an accident.

```{r}
ggplot(amt_train, aes(x=TARGET_AMT)) + 
  geom_density() +
  theme_light() +
  geom_vline(aes(xintercept = mean(TARGET_AMT)), lty=2, col="red") +
  geom_label(aes(x=25000, y=0.00015, label=paste("mean =", round(mean(TARGET_AMT),0)))) +
  geom_vline(aes(xintercept = median(TARGET_AMT)), lty=2, col="darkgreen") +
  geom_label(aes(x=25000, y=0.00010, label=paste("median = ", round(median(TARGET_AMT), 0)))) +
  labs(title="TARGET_AMT Density Plot", y="Density", x="TARGET_AMT")
```

As was previously noted this distribution has a long tail.  The mean payout is $`r round(mean(amt_train$TARGET_AMT), 0)` and the median is $`r round(median(amt_train$TARGET_AMT), 0)`.  The median and mean are higher, of course for those observations we classified as outliers.  The outlier cutoff point is $`r round(outlier, 0)`.

```{r}
amt_train %>%
  mutate(TARGET_AMT_OUTLIER = ifelse(TARGET_AMT_OUTLIER == 1, "Yes", "No")) %>%
  group_by(TARGET_AMT_OUTLIER) %>%
  summarise(Mean = mean(TARGET_AMT),
            Median = median(TARGET_AMT)) %>%
  kable() %>%
  kable_styling()
```

Now to look at the data for the linear regression model.  We are looking for good predictors of the claim amount.  We will look at the numeric variables scatter and correlation plots:

```{r}
amt_train %>%
  keep(is.numeric) %>%
  gather(variable, value, -TARGET_AMT) %>%
  ggplot(., aes(value, TARGET_AMT)) + 
  geom_point() +
  scale_color_brewer(palette="Set1") +
  theme_light() +
  theme(legend.position = "none") +
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

```{r}
M <- amt_train %>%
  select(-INDEX) %>%
  keep(is.numeric) %>%
  cor(.)
corrplot(M, type = "upper")
```

```{r}
amt_train %>%
  keep(is.numeric)%>%
  select(-INDEX, -TARGET_AMT) %>%
  cor(., amt_train$TARGET_AMT) %>%
  kable() %>%
  kable_styling()
```

These two points highlight that there isn't a really strong correlation between most of the predictors and the claim amount.  The only one that is strong is the outlier flag that we created.  Let's finish our exploration by looking at the categoricial variables.  We are removing the `TARGET_AMT_OUTLIERS` from the plot to increase their readability.

```{r}
plot_vars <- c("TARGET_AMT", "TARGET_AMT_OUTLIER", names(keep(amt_train, is.factor)))

amt_train[plot_vars] %>% 
  filter(TARGET_AMT_OUTLIER == 0) %>%
  gather(variable, value, -TARGET_AMT) %>%
  ggplot(., aes(value, TARGET_AMT)) + 
  geom_boxplot() +
  theme_light() +
  theme(legend.position = "none") +
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

These boxplots show that there isn't really much difference in the claim amounts between these groups.

#### Summary of Findings

Through this analysis we have found that there is a lot of noise associated with this signal.  When it comes the the classification task, most variables offer a similar picture.  This is probably due to the random nature of vehicle accidents.  There are a few variables that may be helpful in building our models.

The claim prediciton task will definately be more of a challenge.  Little is correlated with the claim amounts.  The model fit will not be as strong as the classification models.

## Model Creation & Evaluation

Now that we have created datasets and explored them, it is time to build some predictive models and evaluate them.

For the classification models we will look at the prediction accuracy on the test set.  Especially with the accuracy at discovering those that are in an accident.  Accurate prediction of this minority class indicates a good model.

```{r}
evaluate_model <- function(model, test_df, target = "TARGET_FLAG", threshold = 0.5){
  test_df$yhat <- ifelse(predict.glm(model, test_df, "response") >= threshold, 1, 0)
  cm <- confusionMatrix(factor(test_df$yhat), factor(test_df[[target]]), "1")
  deviance <- model$deviance
  r2 <- 1 - model$deviance / model$null.deviance

  cat("F1 =", cm$byClass[7],"\nR2 =", r2, "\n\n")
  print(cm)

  eval <- data.frame(actual = test_df$TARGET_FLAG, 
                     predicted = test_df$yhat, 
                     prob = predict(model, test_df))

  pred <- prediction(eval$prob, eval$actual)
  auc <- performance(pred, measure = "auc")@y.values[[1]]
  perf <- performance(pred, "tpr", "fpr")
  plot(perf,main="ROC Curve", sub = paste0("AUC: ", round(auc, 3)))

  return(cm)
}
```

### Classification Model

#### Baseline Model

We will create a simple model to serve as the baseline.  This model posits that drivers history is a good representation of their future.  So drivers that have been in an accident (`OLDCLAIM` > 0 or `AVG_CLAIM` > 0) are more likely to be in an accident.  Conversely those who haven't aren't likely to be in one in the future.

```{r, null_prefix = TRUE, comment=NA}
baseline_model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), over_sample_train)
summary(baseline_model)
results <- evaluate_model(baseline_model, test)
```

This simple model has statistically significant predictors.  Applying this model to the test data set indicates this simple model has a `r  round(unname(results$overall['Accuracy'])*100,1)`% accuracy rate.  It correcly identified `r round(unname(results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(results$byClass['Specificity'])*100,1)`% of those without.  It is better than flipping a coin but not by much.  All future models must outpreform this baseline.

#### Risk Taker Model

For this model we are going assume that risk takers are more likely to be in an accident.  To identify those who tend to be risk takers, we are going with the urban legend that those who drive red sports cars are risk takers.  Also young men (16-24) also tend to be risk takers.

```{r, null_prefix = TRUE, comment=NA}
risk_taker_model <- glm(TARGET_FLAG ~ RED_SPORTS_CAR + YOUNG_MALE, family = binomial(link = "logit"), over_sample_train)
summary(risk_taker_model)
risk_taker_results <- evaluate_model(risk_taker_model, test)
```

This model only has one statistically significant predictor, however it has a `r  round(unname(risk_taker_results$overall['Accuracy'])*100,1)`% accuracy rate.  This is because it identified `r  round(unname(risk_taker_results$byClass['Specificity'])*100,1)`% of those who didn't have an accident (and they are more numerous).  When looking at the model's performance on identifying the minority group of interest we see it correcly identified `r round(unname(risk_taker_results$byClass['Sensitivity'])*100,1)`% of the people with accidents.  So while the accuracy rate out preformed the baseline model, this one will not be the model we will use.

#### Traditional Model

Traditional wisdom holds that there are a few tried and true predictors of someone's risk of an automobile accident. Namely: age (under 25 vs. 25+), marital status, driving record, distance to work, and the driver's sex (which may or may not be proper to use, but seems to have held throughout the years).

```{r}
traditional_model <- glm(TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX + REVOKED + MVR_PTS + TRAVTIME + CAR_USE, family = binomial(link = "logit"), over_sample_train)
summary(traditional_model)
model_results <- evaluate_model(traditional_model, test)
```

This model has statistically significant predictors.  Applying this model to the test data set indicates this model has a `r  round(unname(model_results$overall['Accuracy'])*100,1)`% accuracy rate.  It correcly identified `r round(unname(model_results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(model_results$byClass['Specificity'])*100,1)`% of those without.  This model out preforms the baseline model.

#### Traditional Model with Cross-Validation

Here we try to use cross-validation techniques to improve the traditional model. We go back to the raw training data, because it is typically not-advised to use data that is oversampled prior to cross-validation.

```{r}
# Get complete cases
cases <- train %>%
  select(YOUNG,MSTATUS,PRIOR_ACCIDENT,SEX,REVOKED,MVR_PTS, TRAVTIME, CAR_USE) %>%
  complete.cases()
temp <- train[cases,]

# Use 5-fold cross-validation
train_control <- trainControl(method = "cv", number = 5, sampling = "up")
traditional_cv <- train(form = TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX + REVOKED + MVR_PTS + TRAVTIME + CAR_USE, method = "glm", family = "binomial", data = temp, trControl = train_control)

traditional_cv
```

```{r}
# Evaluating the model
eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(traditional_cv,newdata=test,type="raw"), prob = predict(traditional_cv,newdata=test,type="prob"))

confusionMatrix(eval$predicted, eval$actual, positive = "1")

pred <- prediction(eval$prob.1, eval$actual)
auc <- performance(pred, measure = "auc")@y.values[[1]]
perf <- performance(pred,"tpr","fpr")
plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
```

#### Alternate Traditional Model

This model is an alternate to the traditional model.

```{r, null_prefix = TRUE, comment=NA}
model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME + SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER, family = binomial(link = "logit"), over_sample_train)
summary(model)
model_results <- evaluate_model(model, test)
```

This model's predictors are all statistically significant.  Applying this model to the test data set indicates this model has a `r  round(unname(model_results$overall['Accuracy'])*100,1)`% accuracy rate.  It correcly identified `r round(unname(model_results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(model_results$byClass['Specificity'])*100,1)`% of those without.  This model also out preforms the baseline model.  Now that we have a few viable classification models, we can turn our attention to the linear claim model.

### Linear Claim Model

#### Baseline Model

Given the unusual distribution one posible solution would be to predict the claim is the median value.  This is a gross simplification but will serve as the baseline for the linear models to compete against. The way we are looking at how well these models preform is calculating the prediction error as a percentage of the actual and then averaging it across the data set.

```{r}
the_median <- median(amt_train$TARGET_AMT)

evaluate_lm <- function(yhat, actual){
  data.frame(yhat = yhat, actual = actual) %>%
    mutate(error = yhat - actual) %>%
    mutate(error_percent = error / actual) %>%
    summarise(`Average Error` = mean(error), `Average Percent` = mean(error_percent)) %>%
    kable() %>%
    kable_styling()
}

evaluate_lm(c(the_median), amt_test$TARGET_AMT)
```

By simply using the median to predict the values of the claims we would underestimate the actual value on average by $1,600 dollars.  On a percentage basis the error would be 50% of the actual.

#### Simple Linear Model

This simple linerar model is that the amount of the claim is based off of the value of the vehicle.  More expensive vehicles should be more costly to repair than less expensive vehicles.  It is overly simplistic.

```{r}
baseline_lm <- lm(TARGET_AMT ~ BLUEBOOK, amt_train)
summary(baseline_lm)
```

While this predictor is statistically significant and positive, the model's overall explanitory power is weak.  Let's see how it preformed on the test set:

```{r}
evaluate_lm(predict(baseline_lm, amt_test), amt_test$TARGET_AMT)
```

Well the average error has decreased which is good but the average error share of the actual has doubled.  This is a garbage model.

#### Outlier Model

We previously flagged the outliers in the data set.  IF we can build another classification model, how well will it preform?

```{r}
outlier_lm <- lm(TARGET_AMT ~ TARGET_AMT_OUTLIER, amt_train)
summary(outlier_lm)
```

This model is unfair as it is predicting the outcomes based on a predictor derived from the outcome.  Unsuprisingly this model does a really good job.  It has an adjusted $R^2$ of `r summary(outlier_lm)$adj.r.squared`.

The reason we explored this model is to see if we has a reason to build another classifier.  If we build a second classifier that can predict this variable would it be worth it.  Let's see how the model preforms on the test set:

```{r}
evaluate_lm(predict(outlier_lm, amt_test), amt_test$TARGET_AMT)
```

The predictions on average understate the actual value by $140.  The error as a percent of the actual is close to what we get with just using the median.  The improvement really comes from using a higher value to estimate the claim for the outliers.

The challenge will be to create a classifier.  There are `r nrow(amt_train[amt_train$TARGET_AMT_OUTLIER == 1,])` out of `r nrow(amt_train)` that are outliers.  Let's create another balanced data set and see if we can build another classifier.

```{r}
set.seed(42)
minority <- nrow(amt_train[amt_train$TARGET_AMT_OUTLIER == 1,])
majority <- nrow(amt_train[amt_train$TARGET_AMT_OUTLIER == 0,])
diff <- majority - minority
minority_index <- amt_train[amt_train$TARGET_AMT_OUTLIER == 1,]$INDEX
over_sample_train_2 <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
  merge(amt_train, .) %>%
  bind_rows(amt_train)
```

Let's explore the correlations to see what is correlated:

```{r eval=FALSE, echo=FALSE}
over_sample_train_2 %>%
  keep(is.numeric)%>%
  select(-TARGET_AMT_OUTLIER, -INDEX, -TARGET_AMT) %>%
  cor(., over_sample_train_2$TARGET_AMT_OUTLIER) %>%
  kable() %>%
  kable_styling()
```

```{r}
M <- over_sample_train_2 %>%
  select(-INDEX, -TARGET_AMT) %>%
  keep(is.numeric) %>%
  cor(.)
corrplot(M, type = "upper")
```

It looks like the `BLUEBOOK` and `PARENT1` are the only variables that are correlated with this outcome.  So let's build the classifier.

```{r, null_prefix = TRUE, comment=NA}
outlier_model <- glm(TARGET_AMT_OUTLIER ~ BLUEBOOK + PARENT1, family = binomial(link = "logit"), over_sample_train_2)
summary(outlier_model)
outlier_model_results <- evaluate_model(outlier_model, test, "TARGET_AMT_OUTLIER")
```

Now to test this out.  Based on our linear regression test set we will predict if the observation is likely to be an outlier or not.  We will then estimate the `TARGET_AMT` using the linear model.  We will then evaluate the results of the "model".

```{r}
temp <- amt_test
temp$TARGET_AMT_OUTLIER <- ifelse(predict.glm(outlier_model, amt_test, "response") >= 0.5, 1, 0)
temp$TARGET_AMT_yhat <- predict(outlier_lm, amt_test)
evaluate_lm(temp$TARGET_AMT_yhat, temp$TARGET_AMT)
```

So through this somewhat convoluted approach we have done slightly better than using the median.

## Predictions

The final predictions will use the alternative traditional model to predict the `TARGET_FLAG` variable, then pass it through the second classifier which predicts the `TARGET_AMT_OUTLIER` flag.  We then use the linear model to make the final amount prediction.

```{r}
predictions <- function(df, classifier_1, linear_model, classifier_2){
  df$TARGET_FLAG <- ifelse(predict.glm(classifier_1, df, "response") >= 0.5, 1, 0)
  # We assume that everyone with a TARGET_FLAG = 0 has a TARGET_AMT as zero.  
  # We then refine it with the outlier model.
  df$TARGET_AMT_OUTLIER <- ifelse(predict.glm(classifier_2, df, "response") >= 0.5, 1, 0)
  df$yhat <- predict(linear_model, df)
  df <- df %>%
    mutate(TARGET_AMT = ifelse(TARGET_AMT_OUTLIER == 1, yhat, 0)) %>%
    mutate(TARGET_AMT = ifelse(TARGET_FLAG == 0, 0, TARGET_AMT)) %>%
    select(-yhat, -TARGET_AMT_OUTLIER)
  return(df)
}

evaluation <- predictions(evaluation, model, outlier_lm, outlier_model)
```

### 2019-12-06 DATA621/HW5.Rmd
Initial commit of DATA 621 HW 5 files

---
title: "DATA 621 Homework #5"
subtitle: "From Work by Critical Thinking Group 3"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 4
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(caret)
library(RColorBrewer)
library(corrplot)
library(gridExtra)
options(scipen = 999)
```

```{r data, echo=FALSE}
df <- read.csv("./data/wine-training-data.csv")
hold_out_data <- read.csv("./data/wine-evaluation-data.csv")
df <- df[,names(df)[2:length(names(df))]] # Drop the INDEX
predictors <- names(df)[2:length(names(df))]
```

## Introduction

We have been given a dataset with information on `r nrow(df)` commercially available wines.  The variables in the dataset are mostly related to the chemical properties of the wine being sold.  A large wine manufacturer is studying the data in order to predict the number of wine cases ordered based upon the wine characteristics. If the wine manufacturer can predict the number of cases, then that manufacturer will be able to adjust their wine offering to maximize sales.  The objective is explore, analyze and build a count regression model to predict the number of cases of wine that will be sold given certain properties of the wine.

## Data Exploration & Preparation

As previously stated we have `r nrow(df)` observations for developing the model.  The evaluation dataset (which we will set aside for now) consists of `r nrow(hold_out_data)` observations.  We will begin by taking a look at the modeling data:

```{r}
summary(df) %>%
  kable() %>%
  kable_styling()
```

Two things jump out about this dataset.  The first thing is that there are variables with negative values.  These negative values don't make sense in the context of chemical properties of wine.  After all, how can you have negative chlorides?

Second, there are quite a few records with missing values.  We will need to either impute values or drop the records with incomplete data.  These two issues will need to be corrected in the data.

### Correcting Invalid Values

We will begin by correcting the invalid values.  We will assume the negative values were a data entry issue and a negative value was entered when a positive value was desired.  If these negative values were changed into positive numbers, they would be within the range of plausible values based on the other data.

```{r}
fix_negative_values <- function(df){
  df %>%
    mutate(FixedAcidity = abs(FixedAcidity),
           VolatileAcidity = abs(VolatileAcidity),
           CitricAcid = abs(CitricAcid),
           ResidualSugar = abs(ResidualSugar),
           Chlorides = abs(Chlorides),
           FreeSulfurDioxide = abs(FreeSulfurDioxide),
           TotalSulfurDioxide = abs(TotalSulfurDioxide),
           Sulphates = abs(Sulphates),
           Alcohol = abs(Alcohol))
}
df <- fix_negative_values(df)
hold_out_data <- fix_negative_values(hold_out_data)
```

### Missing Values

The instructions indicate that sometimes, the fact that a variable is missing is actually predictive of the target.  Let's see what we can learn about `TARGET` from the missing values.

```{r}
temp <- df %>%
  select(TARGET, STARS, Alcohol, ResidualSugar, Chlorides, FreeSulfurDioxide, TotalSulfurDioxide, pH, Sulphates) %>%
  gather("variable", "value", -TARGET) %>%
  mutate(na = ifelse(is.na(value), "Yes", "No")) %>%
  group_by(TARGET, na, variable) %>%
  tally() %>%
  ungroup()

 temp <- temp %>%
    group_by(variable, na) %>%
    summarise(total = sum(n)) %>%
    merge(temp) %>%
    mutate(share = n / total)

 ggplot(temp, aes(TARGET, share, fill = na)) +
   geom_bar(stat = "identity", position = "dodge") +
   scale_fill_brewer(palette = "Set1") +
   facet_wrap(~variable, ncol = 4) + 
   ylab("Percent of Group")
```

We learn that the observations that are missing `STARS` are much more likely to have a zero `TARGET`.  So let's look at how many stars wine with a zero `TARGET` have:

```{r}
df %>%
  filter(TARGET == 0) %>%
  select(STARS) %>%
  na.omit() %>%
  group_by(STARS) %>%
  tally() %>%
  kable() %>%
  kable_styling()
```

Based on this it looks like we should assign 1 for all the missing `STARS`.

There is little to no information encoded in the NAs for `Alcohol`,  `ResidualSugar`, `Chlorides`, `FreeSulfurDioxide`, `TotalSulfurDioxide`, `pH`, and `Sulphates`.  We will impute using the median value.

```{r}
fix_missing_values <- function(df){
  df %>%
    mutate(
      ResidualSugar = ifelse(is.na(ResidualSugar), median(ResidualSugar, na.rm = T), ResidualSugar),
      Chlorides = ifelse(is.na(Chlorides), median(Chlorides, na.rm = T), Chlorides),
      FreeSulfurDioxide = ifelse(is.na(FreeSulfurDioxide), median(FreeSulfurDioxide, na.rm = T), FreeSulfurDioxide),
      TotalSulfurDioxide = ifelse(is.na(TotalSulfurDioxide), median(TotalSulfurDioxide, na.rm = T), TotalSulfurDioxide),
      pH = ifelse(is.na(pH), median(pH, na.rm = T), pH),
      Sulphates = ifelse(is.na(Sulphates), median(Sulphates, na.rm = T), Sulphates),
      Alcohol = ifelse(is.na(Alcohol), median(Alcohol, na.rm = T), Alcohol),
      STARS_imputed = ifelse(is.na(STARS), 1, 0),
      STARS = ifelse(is.na(STARS), 1, STARS)
    )
}
df <- fix_missing_values(df)
hold_out_data <- fix_missing_values(hold_out_data)
```

### Training Test Split

Next we will look at splitting the data into a training and testing set at a standard 70-30 split between train and test:

```{r}
set.seed(42)
train_index <- createDataPartition(df$TARGET, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

### Univariate Analysis

#### Response Variable

```{r echo=FALSE, message=FALSE, warning=FALSE}
train %>%
  group_by(TARGET) %>%
  tally() %>%
  ggplot(., aes(x = factor(TARGET), y = n, fill = factor(TARGET))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Set1") + 
  theme(legend.position = "none") +
  labs(x="Number of Wine Cases Purchased (TARGET)",y = "Count")
```

In order to model this as a Poisson process the mean should be equal to the variance

```{r}
train %>%
  summarise(mean = mean(TARGET), variance = var(TARGET)) %>%
  kable() %>%
  kable_styling()
```

They are not the same.  The implications of this requirement not being met is the the $\beta$s are correct but the standard errors will be wrong.

#### Predictors

```{r}
dens <- lapply(predictors, FUN=function(var) {
  ggplot(df, aes_string(x = var)) + 
    geom_density(fill = "gray") +
    geom_vline(aes(xintercept = mean(train[, var])), color = "blue", size=1) +
    geom_vline(aes(xintercept = median(train[, var])), color = "red", size=1) +
    geom_vline(aes(xintercept = quantile(train[, var], 0.25)), linetype = "dashed", size = 0.5) + 
    geom_vline(aes(xintercept = quantile(train[, var], 0.75)), linetype = "dashed", size = 0.5)
  })
do.call(grid.arrange, args = c(dens, list(ncol = 3)))
```

Some of the predictors (the chemical composition variables) are right skewed.

### Bivariate Analysis

Now let's examine the correlations between the variables:

```{r}
corrplot.mixed(cor(train))
```

Hmm.  This suggests that the chemical makeup of the wine has little bearing on the number of cases sold.  The label appeal, acidity index value and star rating have the strongest correlation.  This also suggests that it may be worth the effort to do a better job filling in the missing values in the `STARS` variable.

## Model Building & Evaluation

### Evaluation Approach

In order to assess the models we will return the the original intent of the exercise: If the wine manufacturer can predict the number of cases, then that manufacturer will be able to adjust their wine offering to maximize sales.  We will evaluate the models based on the number of cases that are likely to be sold.  We will assume that when the model overestimates the cases sold, these cases will never be sold and represent a loss.  When the model underestimates there is also a loss but it is unseen.  It is potential sales that are not realized.

We will summarize the evaluation by first constructing a confusion matrix like summary looking at what our model predicts vs what actually occured.  We will also summarize the total number of cases that would be sold, the net cases sold (after adjusting for the glut due to overestimating), and a final adjusted net which factors in "losses" from under estimation.

```{r}

evaluate_model <- function(model, test_df, yhat = FALSE){
  temp <- data.frame(yhat=c(0:8), TARGET = c(0:8), n=c(0))

  if(yhat){
    test_df$yhat <- yhat
  } else {
    test_df$yhat <- round(predict.glm(model, newdata=test_df, type="response"), 0)
  }

  test_df <- test_df %>%
    group_by(yhat, TARGET) %>%
    tally() %>%
    mutate(accuracy = ifelse(yhat > TARGET, "Over", ifelse(yhat < TARGET, "Under", "Accurate"))) %>%
    mutate(cases_sold = ifelse(yhat > TARGET, TARGET, yhat) * n,
           glut = ifelse(yhat > TARGET, yhat - TARGET, 0) * n,
           missed_opportunity = ifelse(yhat < TARGET, TARGET - yhat, 0) * n) %>%
    mutate(net_cases_sold = cases_sold - glut,
           adj_net_cases_sold = cases_sold - glut - missed_opportunity)

  results <- test_df %>%
    group_by(accuracy) %>%
    summarise(n = sum(n)) %>%
    spread(accuracy, n)

  accurate <- results$Accurate
  over <- results$Over
  under <- results$Under

  cases_sold <- sum(test_df$cases_sold)
  net_cases_sold <- sum(test_df$net_cases_sold)
  adj_net_cases_sold <- sum(test_df$adj_net_cases_sold)
  missed_opportunity <- sum(test_df$missed_opportunity)
  glut <- sum(test_df$glut)

  confusion_matrix <- test_df %>%
    bind_rows(temp) %>%
    group_by(yhat, TARGET) %>%
    summarise(n = sum(n)) %>%
    spread(TARGET, n, fill = 0)

  return(list("confusion_matrix" = confusion_matrix, "results" = results, "df" = test_df, "accurate" = accurate, "over" = over, "under" = under, "cases_sold" = cases_sold, "net_cases_sold" = net_cases_sold, "adj_net_cases_sold" = adj_net_cases_sold, "glut" = glut, "missed_opportunity" = missed_opportunity))
}
```

### Chemical Property Model

We will begin by first constructing a count regression model based on the chemical properties of the wine.  So we will be excluding the label, star rating and the acid index as it is a weighted average of the acidity variables.  Based on our bivariate analysis we would not expect this model to preform well.  Since the mean is not equal to the variance we will be using the quasi-Poisson model.

```{r}
train1 <- train %>%
  select(-LabelAppeal, -AcidIndex, -STARS, -STARS_imputed)
model1 <- glm(TARGET ~ ., family = quasipoisson, train1)
summary(model1)
model1_results <- evaluate_model(model1, test)
```

The following is a confusion matrix like examination of the model with the predicted cases going down the rows and the actual cases going across the columns.

```{r}
model1_results$confusion_matrix %>% 
  kable() %>% 
  kable_styling()
```

The model is only accurate `r model1_results$accurate` times.  It overestimates `r model1_results$over` times and underestimates `r model1_results$under` times.  If decisions were based off of this model, one would expect to sell `r model1_results$cases_sold`.  If we subtract out the extra cases due to overestimation we would have `r model1_results$net_cases_sold` net cases sold. If we adjust this for the cases that could have been sold if the model didn’t underestimate we would have `r model1_results$adj_net_cases_sold` cases sold.

### Second Chemical Property Model

We will try one more chemical property model but only select the variables that were statistically significant in the preious model (`FixedAcidity`, `VolatileAcidity` and `Alcohol`).

```{r}
model2 <- glm(TARGET ~ FixedAcidity + VolatileAcidity + Alcohol, family = quasipoisson, train)
summary(model2)
model2_results <- evaluate_model(model2, test)
```

We would expect to have `r model2_results$cases_sold` cases sold using this model (with a net of `r model2_results$net_cases_sold` and adjusted net of `r model2_results$adj_net_cases_sold`). Looking at the confusion matrix we see this model is accurate only `r model2_results$accurate` out of `r nrow(test)` times.

```{r}
model2_results$confusion_matrix %>% 
  kable() %>% 
  kable_styling()
```

The previous model is slightly better but is still a horible model.  Building a model based off of the chemical properties does not seem like a good solution.  Let's turn our attention to the other variables with stronger correlations.

### Preceived Quality Model

This model relies on the consumer preception of quality based on the label, and the star ratings of the experts.

```{r}
model3 <- glm(TARGET ~ LabelAppeal + STARS, family = quasipoisson, train)
summary(model3)
model3_results <- evaluate_model(model3, test)
```

Here is a confusion matrix for this model:

```{r}
model3_results$confusion_matrix %>% 
  kable() %>% 
  kable_styling()
```

### Perceived Quality Plus Model

We will begin with the preceeding model but add in the Acid Index and the flag if the STARS was imputed.

```{r}
model4 <- glm(TARGET ~ LabelAppeal + STARS + AcidIndex + STARS_imputed, family = quasipoisson, train)
summary(model4)
model4_results <- evaluate_model(model4, test)
```

Let's see how the predictions line up with the reality:

```{r}
model4_results$confusion_matrix %>% 
  kable() %>% 
  kable_styling()
```

We would expect to have `r model3_results$cases_sold` net cases sold using this model (with a net of `r model3_results$net_cases_sold` and adjusted net of `r model3_results$adj_net_cases_sold`). Looking at the confusion matrix we see this model is accurate only `r model3_results$accurate` out of `r nrow(test)` times.

There are no zero predictions in this model which is not accurate.  We will try to address this in our last model

### Adjusted Perceived Quality Plus Model

We will start with the predictions from preceived quality plus model.  We will then manually override the the predictions that have the stars filled in with a predicted zero.

```{r}
temp <- test
temp$yhat <- round(predict.glm(model4, newdata=temp, type="response"), 0)
temp <- temp %>%
  mutate(yhat = ifelse(STARS_imputed == 1, 0, yhat))
model5_results <- evaluate_model(model4, test, temp$yhat)
```

```{r}
model5_results$confusion_matrix %>% 
  kable() %>% 
  kable_styling()
```

We see that applying the adjustment based on if the stars are imputed increased the correct number of cases where it should be predicted as a zero.  It also introduced some underestimation.

We would expect to have `r model4_results$cases_sold` net cases sold using this model (with a net of `r model4_results$net_cases_sold` and adjusted net of `r model4_results$adj_net_cases_sold`). Looking at the confusion matrix we see this model is accurate only `r model4_results$accurate` out of `r nrow(test)` times.

## Model Selection

The following is a summary of the five models

<table>
<thead>
  <tr>
    <th>Model</th>
    <th>Number of Cases Sold</th>
    <th>Number of Cases Overestimated to Sell</th>
    <th>Missed Opportunities</th>
    <th>Times Accurate</th>
  </tr>
</thead>
<tbody>
  <tr>
    <td>Chemical Property Model 1</td>
    <td>`r model1_results$cases_sold`</td>
    <td>`r model1_results$glut`</td>
    <td>`r model1_results$missed_opportunity`</td>
    <td>`r model1_results$accurate`</td>
  </tr>
  <tr>
    <td>Chemical Property Model 2</td>
    <td>`r model2_results$cases_sold`</td>
    <td>`r model2_results$glut`</td>
    <td>`r model2_results$missed_opportunity`</td>
    <td>`r model2_results$accurate`</td>
  </tr>
  <tr>
    <td>Perceived Quality Model</td>
    <td>`r model3_results$cases_sold`</td>
    <td>`r model3_results$glut`</td>
    <td>`r model3_results$missed_opportunity`</td>
    <td>`r model3_results$accurate`</td>
  </tr>
  <tr>
    <td>Perceived Quality Plus Model</td>
    <td>`r model4_results$cases_sold`</td>
    <td>`r model4_results$glut`</td>
    <td>`r model4_results$missed_opportunity`</td>
    <td>`r model4_results$accurate`</td>
  </tr>
  <tr>
    <td>Adjusted Perceived Quality Plus Model</td>
    <td>`r model5_results$cases_sold`</td>
    <td>`r model5_results$glut`</td>
    <td>`r model5_results$missed_opportunity`</td>
    <td>`r model5_results$accurate`</td>
  </tr>
</tbody>
</table>

The `Perceived Quality Plus Model` has the most cases sold out of the models.  However the `Adjusted Perceived Quality Plus Model` has the highest accuracy.  One of the reasons why the `Perceived Quality Plus Model` predicted more cases sold is because we forced more zero predictions in the `Adjusted Perceived Quality Plus Model`.

If we subtract out the waste from overestimation the `Adjusted Perceived Quality Plus Model` is the best model.  This has real business consequences as having extra cases of wine on hand that don't sell becomes a real cost to a business.  For these two reasons it will be the model of choice in producing estimates for the hold out datase.

```{r}
hold_out_data$TARGET <- round(predict.glm(model4, newdata=hold_out_data, type="response"), 0)
hold_out_data %>%
  mutate(TARGET = ifelse(STARS_imputed == 1, 0, TARGET)) %>%
  write.csv(., "./data/hw5-predictions.csv", row.names = F)
```

### 2019-12-06 DATA608/final project/app.py
Added assets

@@ -1,7 +1,7 @@
 # -*- coding: utf-8 -*-
 """
+app.py
 Created on Wed Nov 15 08:52:11 2019
-
 @author: Michael Silva
 """
 
@@ -17,9 +17,18 @@ def index():
     html = open(os.path.join(APP_ROOT, "webpage.html")).read()
     return html
 
-@app.route("/arrow-menu-down-gray.svg")
-def arrow_menu_down_gray():
-    return send_file(os.path.join(APP_ROOT, "arrow-menu-down-gray.svg"), mimetype='image/svg+xml')
+@app.route("/assets/<file_name>")
+def assets(file_name):
+    file_extension = file_name.split(".")[-1]
+    mime_type = {
+        "css": "text/css",
+        "ico": "image/x-icon",
+        "js": "text/javascript",
+        "png": "image/png",
+        "svg": "image/svg+xml"
+    }
+    file_path = os.path.join(APP_ROOT, "assets/", file_name)
+    return send_file(file_path, mimetype = mime_type[file_extension])
 
 @app.route("/api/v1/all-data")
 def all_data():
@@ -44,7 +53,7 @@ def all_data():
 
 @app.route("/api/v1/options")
 def options():
-    seen = set();
+    seen = set()
     options = {}
     html = ""
     data = csv.DictReader(open(os.path.join(APP_ROOT, "study_data.csv")))

### 2019-12-06 DATA608/final project/assets/nys-shrinking-middle-class.js

@@ -1,0 +1,27 @@
+$(document).ready(function () {
+    $.get("api/v1/options", function (options) {
+        $("#dropdown").empty().html(options);
+        //$("#stacked-viz-dropdown").html(options);
+    });
+
+    // UI EVENTS
+
+    // Prevent form submission
+    $("#form").submit(function (event) {
+        event.preventDefault();
+    });
+
+    // Navigation menu click event handling
+    $('.menu_item').click(function () {
+        var wrapper_name = $(this).attr('data-section-id');
+        $('main').attr('class', '').addClass(wrapper_name);
+        $('.menu_item').removeClass('current');
+        $(this).addClass('current');
+    });
+
+    // Button at the bottom of page click event handling
+    $('.back_or_next').click(function () {
+        var name = $(this).attr('data-menu-item-name');
+        $('a[data-section-id=' + name + ']').click();
+    });
+});

### 2019-12-07 DATA608/final project/assets/nys-shrinking-middle-class.js
Added visualization

@@ -1,10 +1,169 @@
-$(document).ready(function () {
-    $.get("api/v1/options", function (options) {
-        $("#dropdown").empty().html(options);
-        //$("#stacked-viz-dropdown").html(options);
+var all_data;
+var geoid;
+var key = 0;
+var viz_data;
+var viz_plotly_data;
+
+var viz_plotly_layout = {
+    xaxis: { range: [0, 60] },
+    showlegend: false,
+    title: "Share of Households by Income Level (%)"
+};
+
+var viz_plotly_options = {
+    responsive: true,
+    scrollZoom: false,
+    showLink: false
+};
+
+function x() {
+    viz_data = all_data[geoid];
+    return [Math.round(viz_data[key]["low_share"] * 100, 0), Math.round(viz_data[key]["middle_share"] * 100, 0), Math.round(viz_data[key]["upper_share"] * 100, 0)];
+}
+
+function update_viz() {
+    Plotly.animate("viz", {
+        data: [{ x: x(), }],
+        traces: [0],
+        layout: {}
+    }, {
+        transition: {
+            duration: 500,
+            easing: "cubic-in-out"
+        },
+        frame: {
+            duration: 500
+        }
+    });
+}
+
+function change_dropdown() {
+    if ($("#dropdown option:selected").val() == "36123") {
+        $("#dropdown option:selected").attr("selected", false);
+        $("#dropdown option:first").attr("selected", "selected");
+    } else {
+        $("#dropdown option:selected").attr("selected", false).next().attr("selected", "selected");
+    }
+}
+
+function plot_stacked_viz_data() {
+    var stacked_viz_layout = {
+        barmode: "overlay",
+        xaxis: {type: 'category'},
+        title: "Are Households Moving Up or Down?"
+    };
+
+    stacked_viz_geoid = $("#stacked-viz-dropdown").val();
+    var high = {
+        x: [],
+        y: [],
+        textposition: 'auto',
+        name: 'High',
+        hoverinfo: 'none',
+        marker: {color: '#d63031'},
+        type: 'bar'
+    };
+
+    var middle = {
+        x: [],
+        y: [],
+        textposition: 'auto',
+        name: 'Middle',
+        hoverinfo: 'none',
+        marker: {color: '#0984e3'},
+        type: 'bar'
+    };
+
+    var low = {
+        x: [],
+        y: [],
+        textposition: 'auto',
+        name: 'Low',
+        hoverinfo: 'none',
[…]

### 2019-12-07 DATA608/final project/assets/nys-shrinking-middle-class.js
Lots of cool changes

@@ -1,6 +1,7 @@
 var all_data;
 var geoid;
 var key = 0;
+var viz_mode = "animated";
 var viz_data;
 var viz_plotly_data;
 
@@ -22,19 +23,23 @@ function x() {
 }
 
 function update_viz() {
-    Plotly.animate("viz", {
-        data: [{ x: x(), }],
-        traces: [0],
-        layout: {}
-    }, {
-        transition: {
-            duration: 500,
-            easing: "cubic-in-out"
-        },
-        frame: {
-            duration: 500
-        }
-    });
+    if(viz_mode == "animated"){
+        Plotly.animate("viz", {
+            data: [{ x: x(), }],
+            traces: [0],
+            layout: {}
+        }, {
+            transition: {
+                duration: 500,
+                easing: "cubic-in-out"
+            },
+            frame: {
+                duration: 500
+            }
+        });
+    } else {
+        plot_stacked_bar_chart();
+    }
 }
 
 function change_dropdown() {
@@ -46,10 +51,10 @@ function change_dropdown() {
     }
 }
 
-function plot_stacked_viz_data() {
+function plot_stacked_bar_chart() {
     var stacked_viz_layout = {
         barmode: "overlay",
-        xaxis: {type: 'category'},
+        xaxis: {type: "category"},
         title: "Are Households Moving Up or Down?"
     };
 
@@ -57,31 +62,31 @@ function plot_stacked_viz_data() {
     var high = {
         x: [],
         y: [],
-        textposition: 'auto',
-        name: 'High',
-        hoverinfo: 'none',
-        marker: {color: '#d63031'},
-        type: 'bar'
+        textposition: "auto",
+        name: "High",
+        hoverinfo: "none",
+        marker: {color: "#d63031"},
+        type: "bar"
     };
 
     var middle = {
         x: [],
         y: [],
-        textposition: 'auto',
-        name: 'Middle',
-        hoverinfo: 'none',
-        marker: {color: '#0984e3'},
-        type: 'bar'
+        textposition: "auto",
+        name: "Middle",
+        hoverinfo: "none",
+        marker: {color: "#0984e3"},
+        type: "bar"
     };
 
     var low = {
         x: [],
         y: [],
-        textposition: 'auto',
-        name: 'Low',
-        hoverinfo: 'none',
-        marker: {color: '#fdcb6e'},
-        type: 'bar'
+        textposition: "auto",
+        name: "Low",
+        hoverinfo: "none",
+        marker: {color: "#fdcb6e"},
[…]

### 2019-12-07 DATA608/final project/assets/nys-shrinking-middle-class.js
Total awesomeness!

@@ -4,6 +4,7 @@ var key = 0;
 var viz_mode = "animated";
 var viz_data;
 var viz_plotly_data;
+var viz_timeout = 1000;
 
 var viz_plotly_layout = {
     xaxis: { range: [0, 60] },
@@ -58,7 +59,7 @@ function plot_stacked_bar_chart() {
         title: "Are Households Moving Up or Down?"
     };
 
-    stacked_viz_geoid = $("#stacked-viz-dropdown").val();
+    geoid = $("#dropdown").val();
     var high = {
         x: [],
         y: [],
@@ -93,7 +94,7 @@ function plot_stacked_bar_chart() {
     middle_vals = [];
     high_vals = [];
 
-    $.each(all_data[stacked_viz_geoid], function (index, d) {
+    $.each(all_data[geoid], function (index, d) {
         l = Math.round(d["low_share"] * 100, 0);
         m = Math.round(d["middle_share"] * 100, 0);
         h = Math.round(d["upper_share"] * 100, 0);
@@ -130,6 +131,7 @@ function change_viz() {
     var chart_showing = $("main").hasClass("part_1");
     if(chart_showing){
         if(viz_mode == "animated"){
+            viz_timeout = 1000;
             if (auto_advance) {
                 key++;
                 if (key == 5) {
@@ -140,11 +142,14 @@ function change_viz() {
                 update_viz();
             } 
         } else {
-            change_dropdown();
+            viz_timeout = 3000;
+            if(auto_advance){
+                change_dropdown();
+            }
             update_viz();
         }
     }
-    setTimeout(change_viz, 1000);
+    setTimeout(change_viz, viz_timeout);
 }
 
 
@@ -160,7 +165,7 @@ function init(){
         all_data = response;
         plot_animated_bar_chart();
         // Start the auto advancer
-        setTimeout(change_viz, 1000);
+        setTimeout(change_viz, viz_timeout);
     });
 }
 

### 2019-12-08 DATA608/final project/app.py
Added inital tiles

@@ -41,6 +41,7 @@ def all_data():
             year = str(int(year)-4) + "-" + year[-2:]
         json_data.append({
             "label": year,
+            "name": row['name'].replace(" County, New York", ""),
             "low_share": row["low_class"],
             "middle_share": row["middle_class"],
             "upper_share": row["upper_class"],

### 2019-12-08 DATA608/final project/assets/nys-shrinking-middle-class.js

@@ -5,6 +5,8 @@ var viz_mode = "animated";
 var viz_data;
 var viz_plotly_data;
 var viz_timeout = 1000;
+var year_filter;
+var group_filter;
 
 var viz_plotly_layout = {
     xaxis: { range: [0, 60] },
@@ -152,9 +154,33 @@ function change_viz() {
     setTimeout(change_viz, viz_timeout);
 }
 
+function build_ranking_tiles() {
+    console.log(all_data);
+    $.each(all_data, function (geoid, data) {
+        div = $("<div />").attr("id", geoid).addClass("tile");
+        label = $("<label />").text(data[0]['name']);
+        div.append(label);
+        $.each(data, function(index, d){
+            c = "data_for_" + d["label"].replace("-", "_");
+            l = Math.round(d["low_share"] * 100, 0) + "%";
+            m = Math.round(d["middle_share"] * 100, 0) + "%";
+            h = Math.round(d["upper_share"] * 100, 0) + "%";
+            p = $("<p />").addClass(c);
+            l_span = $("<span />").addClass("low_value percent").text(l);
+            m_span = $("<span />").addClass("mid_value percent").text(m);
+            h_span = $("<span />").addClass("high_value percent").text(h);
+            p.append(l_span).append(m_span).append(h_span);
+            div.append(p);
+        });
+        $('#ranking_tiles_container').append(div);
+    });
+}
+
 
 function init(){
     viz = $("#viz_container").addClass("viz");
+    group_filter = $("#group_filter").val();
+    year_filter = $("#year_filter").val();
     geoid = $("#dropdown").val();
 
     $.get("api/v1/options", function (options) {
@@ -164,12 +190,16 @@ function init(){
     $.getJSON("api/v1/all-data", function (response) {
         all_data = response;
         plot_animated_bar_chart();
+        // Build the ranking tiles
+        build_ranking_tiles();
         // Start the auto advancer
         setTimeout(change_viz, viz_timeout);
     });
 }
 
-
+function shuffle_tiles(){
+    $("#ranking_tiles_container").removeClass().addClass(year_filter).addClass(group_filter);
+}
 
 $(document).ready(function () {
     var geoid = $("#dropdown").val();
@@ -177,6 +207,16 @@ $(document).ready(function () {
     $("#slider").change(function () {
         key = $(this).val();
         update_viz();
+    });
+
+    $("#year_filter").change(function(){
+        year_filter = $(this).val();
[…]

### 2019-12-08 DATA608/final project/app.py
UI enchancements

@@ -8,6 +8,7 @@ Created on Wed Nov 15 08:52:11 2019
 import os
 import csv
 from flask import Flask, jsonify, request, send_file
+import pandas as pd
 
 APP_ROOT = os.path.dirname(os.path.abspath(__file__))
 app = Flask(__name__)
@@ -51,6 +52,29 @@ def all_data():
         })
         json[row['geoid']] = json_data
     return jsonify(json)
+
+@app.route("/api/v1/rankings")
+def rankings():
+    df = pd.read_csv(os.path.join(APP_ROOT, "study_data.csv"))
+    df["low_rank"] = df.groupby("year")["low_class"].rank("dense", ascending=False)
+    df["middle_rank"] = df.groupby("year")["middle_class"].rank("dense", ascending=False)
+    df["high_rank"] = df.groupby("year")["upper_class"].rank("dense", ascending=False)
+
+    rankings = {}
+    for index, row in df.iterrows():
+        geoid = row['geoid']
+        year = str(row['year'])
+        low_rank = int(row['low_rank'])
+        middle_rank = int(row['middle_rank'])
+        high_rank = int(row['high_rank'])
+        if int(year) > 2000:
+            year = str(int(year) - 4) + "_" + year[-2:]
+        year = "show_" + year
+        r = rankings.get(geoid, dict())
+        r[year] = {"low": low_rank, "middle": middle_rank, "high": high_rank}
+        rankings[geoid] = r
+
+    return jsonify(rankings)
 
 @app.route("/api/v1/options")
 def options():

### 2019-12-08 DATA608/final project/assets/nys-shrinking-middle-class.js

@@ -7,6 +7,7 @@ var viz_plotly_data;
 var viz_timeout = 1000;
 var year_filter;
 var group_filter;
+var rankings;
 
 var viz_plotly_layout = {
     xaxis: { range: [0, 60] },
@@ -155,7 +156,6 @@ function change_viz() {
 }
 
 function build_ranking_tiles() {
-    console.log(all_data);
     $.each(all_data, function (geoid, data) {
         div = $("<div />").attr("id", geoid).addClass("tile");
         label = $("<label />").text(data[0]['name']);
@@ -173,6 +173,11 @@ function build_ranking_tiles() {
             div.append(p);
         });
         $('#ranking_tiles_container').append(div);
+    });
+
+    $.getJSON("api/v1/rankings", function (response) {
+        rankings = response;
+        shuffle_tiles();
     });
 }
 
@@ -198,6 +203,10 @@ function init(){
 }
 
 function shuffle_tiles(){
+    $.each(rankings, function (geoid, d) {
+        rank = "rank_" + d[year_filter][group_filter];
+        $('#'+geoid).removeClass().addClass("tile").addClass(rank);
+    });
     $("#ranking_tiles_container").removeClass().addClass(year_filter).addClass(group_filter);
 }
 

### 2019-12-19 DATA621/Manuscript.Rmd
Added data 621 final project

---
title: "A Logistic Regression Approach to CoIL Challenge 2000"
author: "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
bibliography: "Manuscript.bib"
output: 
  word_document:
    reference_docx: word-template.docx
---

```{r echo=FALSE, message=FALSE}
temp <- read.delim("data/ticdata2000.txt", header = FALSE, nrows = 1)
df_cols <- length(names(temp))
source("CoIL.R")
library(pander)
knitr::opts_chunk$set(comment = NA, echo = FALSE)
knitr::knit_hooks$set(inline = function(x) { if(!is.numeric(x)){ x }else{ prettyNum(round(x,2), big.mark=",") } })
```

# Abstract

<!-- In the abstract, the authors should explain the major objective of the study in an objective section, explain how the study was done in a Methods section, describe the findings in a Results section, and report whether the major goal was met in a Conclusion. In general, the reader should ask, If I could not read the entire manuscript, would the abstract adequately summarize it? -->

This paper describes a logistic regression based solution to the CoIL Challenge 2000. The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics. Models were trained on over sampled data.  The model out preformed other's attemps at solving this classification problem.

*Key words: CoIL Challenge, Logistic Regression*

# Introduction

<!-- Readers need to understand all the reasons for performing the study, the assumptions underlying the methodology, and the nuances of the performance of the study. The purposes of the Introduction are, first, to provide the rationale for the study and, second, to explain the study's goals. The Introduction should include a problem statement that conveys the important issues and provides the context for the study. -->

Businesses use data science to extract insights from data.  One pratical application is identifying households to include in a marketing campaign.  In this paper we set out to identify potential customers for an insurance product using real world data from the Computational Intelligence and Learning (CoIL) Challenge. Specifically we are predicting if a customer is likely candidate for a caravan (mobile home/camper) insurance policy.  This is particularly challenging because the data is imballanced (only `r sum(as.numeric(as.character(df$CARAVAN)))` of the `r nrow(df)` records for model training/testing are policy holders).

# Literature Review

<!--Discuss how other researchers have addressed similar problems, what their achievements are, and what the advantage and drawbacks of each reviewed approach are. Explain how your investigation is similar or different to the state-of-the-art. Please cite the relevant papers where appropriate.-->

Fourty-three other research teams have attempted to identify potential insurance policy customers [@Putten].  They used a variety of approaches including: Boosted Decision Tree [@McKone], Classification and Regression Tree (CART) [@Simmonds], Classification Trees with Bagging [@White], C4.5 [@Rickets; @Seewald], Evolutionary Algorithm [@Koudijs], Fuzzy Classifier [@Abonyi; @Kaymak], Genetic Algorithms and Hill-climbers [@Carter], Inductive Learning by Logic Minimization (ILLM) [@Gamberger; @Smuc], Instance Based Reasoning (iBARET) [@Miksovsky], K-Means [@Vesanto], KXEN [@Bera], LOGIT [@Doornik], Mask Perceptron with Boosting [@Leckie], Midos Algorithm [@Krogel], N-Tuple Classifier [@Jorgensen], Naïve Bayes [@Elkan; @Kontkanen], Neural Networks[@Brierley; @Crocoll; @Kim; @Shtovba_Mashnitskiy], Phase Intervals and Genetic Algorithms [@Shtovba], Scoring System [@Lewandowski], Support Vector Machines[@Keerthi], and XCS [@Greenyer].

The maximum number of potential policy owners that could be found is 238.  Previous researchers identified 95 policy owners on average.  The best preforming model [@Elkan] during the initial challenge identified 121 policy owners.  It was a Naïve Bayes, suggesting that probabilities of some of the variables will be useful in identifing potential customers.

A meta analysis of the initial researchers found that simpler algorithms tended to outpreform more complicated ones [@Putten].  With the benefit of these findings, we set out to create a simple logistic regression model that preforms as well or better than the original CoIL Challenge cohort.  In the end, our model outpreformed the orignal researcher's model in correctly predicting the customers that would purchase the insurance policy.

```{r}
# Similar the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121), n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
ggplot(data = prediction_scores, aes(x = score, y = n)) +
  geom_bar(stat = "identity", fill = "#377eb8") +
  xlab(element_blank()) +
  ylab(element_blank()) + 
  xlim(37, 250) +
  annotate("text", x = 65, y = 4, label = "Previous Researchers") +
  geom_vline(xintercept = 238, color = "#e41a1c", linetype="dashed") +
  annotate("text", x = 220, y = 4, label = "Maximum") +
  geom_vline(xintercept = final_model$correct, color = "#4daf4a") +
  annotate("text", x = final_model$correct - 20, y = 4, label = "Our Model") +
  ggtitle("Number of Customers Correctly Predicted")
```

# Methodology

<!-- Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection. -->

The CoIL Challenge dataset is composed of `r df_cols` variables accross `r nrow(df)` observations.  An evaluation dataset is provided with `r nrow(eval)` observations. Five of the predictors are categorical and the remainder are numeric.  Most of the predictors have little to no correlation with the variable of interest (CARAVAN).

## Experimentation and Results

<!-- Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).

Needs to systematically and clearly articulate the study findings. If the results are unclear, the reviewer must decide whether the analysis of the data was poorly executed or whether the Results section is poorly organized. -->

We split the data into training/test sets using a 70/30 split.  We corrected the imballance by oversampling the minority class (caravan policy holders).  Given the large number of possible predictors we used a random forest to aid in variable selection.  A logistic regression model was trained on the oversampled training set using the top five variables selected by the random forest (MOSTYPE, PPERSAUT, MOSHOOFD, PBRAND, APERSAUT).

The MOSTYPE variable has 40 customer types.  Not all customer types were statiscally significant predictors.  We identified those in the oversampled dataset with a probability greater than 0.5 to purchase the insurance product as the LIKELY_CUSTOMER variable.  A figure in the appendix shows all 40 customer types and the probability that they purched the insurance policy in the oversampled dataset.

We fit our second model to the LIKELY_CUSTOMER and PPERSAUT variables. Although it only has two variables, the model preformed well.  This seemed in-line with expectations that simple models preform best.

The third model was fit the same as the second model other than the driven growers variable, which is a variable derived from MOSHOOFD variable, specifically it is when the MOSHOOFD variable is equal to 2, which means the specific customer's main type is "Driven Grower" hence the variable name. We chose this because when running decision trees we noticed this specific factor of MOSHOOFD variable often stood out in the decision trees.

In evaluating the models we examined we focused on the specificity.  The goal of the CoIL challenge was to accurately predict those would would purchase the insurance policy, so focusing on the model's specificity was the best evaluation metric.  In order to get a better sense of how well the model generalizes, we repeatedly retrained and evaluated the model using diffrent samplings of the training dataset.  The following figure summarizes the distribution of the specificty the models produced on the test dataset.  The dashed line is how our model preformed:

```{r}
temp <- rbind(data.frame("Model" = c("1"), "Specificity" = model1_robust_results$specificity),
              data.frame("Model" = c("2"), "Specificity" = model2_robust_results$specificity))
temp <- 
  rbind(temp, data.frame("Model" = c("3"), "Specificity" = model3_robust_results$specificity))%>%
  mutate(Model = as.factor(Model))

ggplot(temp, aes(x = Specificity, fill = Model)) +
  geom_density(alpha = 0.5, color = NA) +
  ylab("Density") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  geom_vline(xintercept = mean(model1_results$specificity), linetype="dashed", color = "#e41a1c") +
  geom_vline(xintercept = mean(model2_results$specificity), linetype="dashed", color = "#377eb8") +
  geom_vline(xintercept = mean(model3_results$specificity), linetype="dashed", color = "#4daf4a") 
```

## Discussion

<!-- Should state whether their hypotheses were verified or proven untrue or, if no hypotheses were given, whether their research questions were answered. The authors should also comment on their results in light of previous studies and explain what differences (if any) exist between their findings and those reported by others and attempt to provide an explanation for the discrepancies.-->

Our model outpreforms the original cohort of CoIL Challenge researchers.  We found that the being a member of one of the following customer types to be a significant predictor of purchasing the insurance produce (listed in order of the probability of purchasing with the MOSTYPE in parenthesis):

* Affluent young families (12), 
* Middle class families (8), 
* Career and childcare (6), 
* Double income no kids (7), 
* High Income, expensive child (1), 
* Very Important Provincials (2), 
* Couples with teens 'Married with children' (36), 
* High status seniors (3), 
* Mixed small town dwellers (37), 
* Stable family (10), 
* Ethnically diverse (20), 
* Traditional families (38), 
* Family starters (11)

People at the top of the list (most likely to buy) are generally those who are well to do.  This is further reinforced in the model with the inclusion of PPERSAUT or the contribution to car policies.  Those with higher contribution levels are more likely to purchase caravan insurance.

"Driven Growers" are made up of the Career and childcare, Double income no kids, and Middle class families customer types.  Once again these customer types generally have more disposible income and are likely customers.

These findings are similar to what the winner of the CoIL challenge found [@Elkan].  They found that a high PPERSAUT (meaning a level 6) or having two car policies to be the strongest single predictors of having caravan insurance.  They found the other most statistically significant predictors are:

* "purchasing power class" is high (5 or higher, especially 7 or higher)
* a private third party insurance policy
* a boat policy
* a social security insurance policy
* a single fire policy with higher contribution (level 4)

Elkan explain that "Intuitively, these predictors identify customers who have a car and are wealthier than average, and who in general carry more insurance coverage than average. It is not surprising that these are the people who are most likely to have caravan insurance."  Our findings are inline with this, although we arrived at this conclusion by examining other variables.

# Conclusions

<!-- Conclude your findings, limitations, and suggest areas for future work. -->

This paper set out to define a model that was effective at identifying potential customers for an insurance policy.  We found that a simple logistic regression model with three variables (two which we derived) outpreformed previous research.  In general terms the three variables correlate with the wealth of the household.  Thus customers that are wealthier than average are most likely candidates to purchase insurance.

Areas for future work include building an ensemble model using the specification of the third model.  Our repeated modeling has a distribution where the mean number of correct predictions is `r round(mean(model3_robust_results$correct), 0)` which is higher than the `r model3_results$correct`.  It appears that incorporating downsampling into the ensemble might have limited impact on the performance as the mean number of correct predictions is `r round(mean(model3_down_robust_results$correct), 0)`.

\newpage

# Appendix

## Probability of Purchasing Product (in blue) by Customer Type

```{r, fig.width=10, fig.height=11}
temp <- up_train %>%
  group_by(MOSTYPE) %>%
  tally() %>%
  rename(total = n)

temp <- up_train %>%
  filter(CARAVAN == 1) %>%
  group_by(MOSTYPE) %>%
  tally() %>%
  merge(temp, ., all.x = TRUE) %>%
  mutate(share_1 = ifelse(is.na(n), 0, n / total)) %>%
  mutate(order = row_number(share_1)) %>%
  select(-n) %>%
  merge(temp)

temp <- up_train %>%
  group_by(CARAVAN, MOSTYPE) %>%
  tally() %>%
  merge(temp) %>%
  mutate(share = n / total)

ggplot(temp, aes(x = reorder(MOSTYPE, order), y = share, fill = CARAVAN))+
  geom_bar(stat="identity") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip() +
  xlab(element_blank()) +
  ylab(element_blank()) +
  geom_hline(yintercept = 0.5, color = "white", linetype = "dashed") +
  theme_minimal() +
  theme(legend.position = "none") 
```

\newpage

## Correlation Coefficients for Variables of Interest
```{r}
vars <- c("CARAVAN", "MOSTYPE", "PPERSAUT", "PBRAND", "MOSHOOFD", "APERSAUT", "LIKELY_CUSTOMERS", "DRIVEN_GROWERS")
temp <- up_train %>%
  mutate(CARAVAN = as.numeric(as.character(CARAVAN)),
         LIKELY_CUSTOMERS = as.numeric(as.character(LIKELY_CUSTOMERS)))
temp <- temp[vars]

numeric_vars <- unlist(lapply(temp, is.numeric))

pander(cor(temp[numeric_vars], temp[numeric_vars]))
```

## Data Dictionary for Variables of Interest

```{r, results='asis'}
data_dictionary <- data.frame(
  Name = c("CARAVAN", "MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
           "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV",
           "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD",
           "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER", "MBERMIDD",
           "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2", "MSKC", "MSKD",
           "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0", "MZFONDS", "MZPART",
           "MINKM30", "MINK3045", "MINK4575", "MINK7512", "MINK123M", "MINKGEM",
           "MKOOPKLA", "PWAPART", "PWABEDR", "PWAAND", "PPERSAUT", "PBESAUT",
           "PMOTSCO", "PVRAAUT", "PAANHANG", "PTRACTOR", "PWERKT", "PBROM",
           "PLEVEN", "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
           "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART", "AWABEDR",
           "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO", "AVRAAUT", "AAANHANG",
           "ATRACTOR", "AWERKT", "ABROM", "ALEVEN", "APERSONG", "AGEZONG",
           "AWAOREG", "ABRAND", "AZEILPL", "APLEZIER", "AFIETS", "AINBOED",
           "ABYSTAND", "LIKELY_CUSTOMERS", "DRIVEN_GROWERS"),
  Description = c("Number of mobile home policy",
                  "Customer Subtype", "Number of houses", "Avg size household",
                  "Avg age", "Customer main type", "Roman catholic",
                  "Protestant", "Other religion", "No religion", "Married",
                  "Living together", "Other relation", "Singles",
                  "Household without children", "Household with children",
                  "High level education", "Medium level education",
                  "Lower level education", "High status", "Entrepreneur",
                  "Farmer", "Middle management", "Skilled labourers",
                  "Unskilled labourers", "Social class A", "Social class B1",
                  "Social class B2", "Social class C", "Social class D",
                  "Rented house", "Home owners", "1 car", "2 cars", "No car",
                  "National Health Service", "Private health insurance",
                  "Income < 30.000", "Income 30-45.000", "Income 45-75.000",
                  "Income 75-122.000", "Income >123.000", "Average income",
                  "Purchasing power class",
                  "Contribution private third party insurance",
                  "Contribution third party insurance (firms)",
                  "Contribution third party insurane (agriculture)",
                  "Contribution car policies",
                  "Contribution delivery van policies",
                  "Contribution motorcycle/scooter policies",
                  "Contribution lorry policies",
                  "Contribution trailer policies",
                  "Contribution tractor policies",
                  "Contribution agricultural machines policies",
                  "Contribution moped policies",
                  "Contribution life insurances",
                  "Contribution private accident insurance policies",
                  "Contribution family accidents insurance policies",
                  "Contribution disability insurance policies",
                  "Contribution fire policies",
                  "Contribution surfboard policies",
                  "Contribution boat policies",
                  "Contribution bicycle policies",
                  "Contribution property insurance policies",
                  "Contribution social security insurance policies",
                  "Number of private third party insurance",
                  "Number of third party insurance (firms)",
                  "Number of third party insurane (agriculture)",
                  "Number of car policies",
                  "Number of delivery van policies",
                  "Number of motorcycle/scooter policies",
                  "Number of lorry policies",
                  "Number of trailer policies", "Number of tractor policies",
                  "Number of agricultural machines policies",
                  "Number of moped policies", "Number of life insurances",
                  "Number of private accident insurance policies",
                  "Number of family accidents insurance policies",
                  "Number of disability insurance policies",
                  "Number of fire policies", "Number of surfboard policies",
                  "Number of boat policies", "Number of bicycle policies",
                  "Number of property insurance policies",
                  "Number of social security insurance policies",
                  "MOSTYPE = 12, 8, 6, 7, 1, 2, 36, 3, 37, 10, 20, 38, or 11",
                  "MOSHOOFD = 2"))
temp <- data_dictionary[data_dictionary$Name %in% vars,]
row.names(temp) <- NULL
pander(temp)
```

\newpage

## Model Summary

```{r}
summary(model3)
```

\newpage

## Confusion Matrix and Statistics for our Model

```{r}
final_model$cm
```

## R statistical programming code.

```{r}
conn <- file("CoIL.r", open = "r")
lines <- readLines(conn)
n_lines <- length(lines)
for (i in 1:n_lines) {
  cat(lines[i], "\n")
}
close(conn)
```

\newpage

## References

### 2019-12-19 DATA621/CoIL.R

@@ -1,0 +1,329 @@
+# CoIL Challenge Source Code
+library(tidyverse)
+library(caret)
+
+## Download the data sets from UCI if they are not present
+url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/tic-mld/"
+files <- c("ticdata2000.txt", "ticeval2000.txt", "tictgts2000.txt")
+for (file_name in files) {
+  file_path <- paste0("data/", file_name)
+  file_url <- paste0(url, file_name)
+  if (!file.exists(file_path)) {
+    message(paste("Downloading", file_name))
+    download.file(file_url, file_path)
+  }
+}
+
+## Read in and clean the data
+prepare_data <- function(df){
+  names(df) <- c(
+    "MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
+    "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV", "MFALLEEN",
+    "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD", "MOPLLAAG", "MBERHOOG",
+    "MBERZELF", "MBERBOER", "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA",
+    "MSKB1", "MSKB2", "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2",
+    "MAUT0", "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
+    "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART", "PWABEDR", 
+    "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO", "PVRAAUT",  "PAANHANG",
+    "PTRACTOR", "PWERKT", "PBROM", "PLEVEN", "PPERSONG", "PGEZONG",
+    "PWAOREG", "PBRAND", "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED",
+    "PBYSTAND", "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT",
+    "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM",
+    "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
[…]

### 2020-01-27 DATA624/HW1.Rmd
Initial commit of HW1

---
title: "DATA 624 Homework 1"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(fpp2)
```

## 2.1

Use the help function to explore what the series `gold`, `woolyrnq` and `gas` represent.

a. Use `autoplot()` to plot each of these in separate plots.

b. What is the frequency of each series? Hint: apply the `frequency()` function.

c. Use `which.max()` to spot the outlier in the gold series. Which observation was it?

## 2.2

Download the file `tute1.csv` from the book website, open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

a. You can read the data into R with the following script:

```{r}
download.file("http://otexts.com/fpp2/extrafiles/tute1.csv", "tute1.csv")
tute1 <- read.csv("tute1.csv", header=TRUE)
View(tute1)
```

b. Convert the data to time series

```{r}
mytimeseries <- ts(tute1[,-1], start=1981, frequency=4)
```

(The [,-1] removes the first column which contains the quarters as we don’t need them now.)

c. Construct time series plots of each of the three series

```{r}
autoplot(mytimeseries, facets=TRUE)
```

Check what happens when you don’t include facets=TRUE.

```{r}
autoplot(mytimeseries)
```

## 2.3

Download some monthly Australian retail data from the book website. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

a. You can read the data into R with the following script:

```{r}
download.file("https://otexts.com/fpp2/extrafiles/retail.xlsx", "retail.xlsx")
retaildata <- readxl::read_excel("retail.xlsx", skip=1)
```

The second argument (skip=1) is required because the Excel sheet has two header rows.

b. Select one of the time series as follows (but replace the column name with your own chosen column):

```{r}
myts <- ts(retaildata[,"A3349873A"], frequency=12, start=c(1982,4))
```

c. Explore your chosen retail time series using the following functions:

`autoplot()`, `ggseasonplot()`, `ggsubseriesplot()`, `gglagplot()`, `ggAcf()`

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

## 2.6

Use the following graphics functions: `autoplot()`, `ggseasonplot()`, `ggsubseriesplot()`, `gglagplot()`, `ggAcf()` and explore features from the following time series: `hsales`, `usdeaths`, `bricksq`, `sunspotarea`, `gasoline`.

* Can you spot any seasonality, cyclicity and trend?

* What do you learn about the series?

### 2020-01-31 DATA624/HW1.Rmd
More work on HW1

@@
     toc_depth: 3
-    code_folding: "show"
+    code_folding: "hide"
 ---
 
-```{r setup, echo=FALSE, cache=FALSE}
+```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
 knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
@@
 library(fpp2)
+library(gridExtra)
 ```
 
-## 2.1
+## Question 2.1
 
@@
 
+**`gold` is the daily morning gold prices in US dollars from 1 Jan 1985 to 31 March 1989.
+`woolyrnq` is the quarterly production of woolen yarn in Australia in tons from March 1965 to September 1994.
+`gas` is monthly gas production in Australia from 1956-1995.**
+
 a. Use `autoplot()` to plot each of these in separate plots.
+
+```{r}
+autoplot(gold)
+autoplot(woolyrnq)
+autoplot(gas)
+```
 
@@
 
+```{r}
+frequency(gold)
+frequency(woolyrnq)
+frequency(gas)
+```
+
 c. Use `which.max()` to spot the outlier in the gold series. Which observation was it?
 
-## 2.2
+```{r}
+goldoutlier <- which.max(gold)
+```
+
+**It is the `r goldoutlier` observation.  The price of gold was `r gold[goldoutlier]`**
+
+## Queston 2.2
 
@@
 ```{r}
+if(!file.exists("tute1.csv")){
   download.file("http://otexts.com/fpp2/extrafiles/tute1.csv", "tute1.csv")
+}
 tute1 <- read.csv("tute1.csv", header=TRUE)
@@
 
-## 2.3
+## Question 2.3
 
@@
 ```{r}
+if(!file.exists("retail.xlsx")){
   download.file("https://otexts.com/fpp2/extrafiles/retail.xlsx", "retail.xlsx")
+}
 retaildata <- readxl::read_excel("retail.xlsx", skip=1)
@@
 
+```{r}
+autoplot(myts)
+ggseasonplot(myts)
+ggsubseriesplot(myts)
+gglagplot(myts)
+ggAcf(myts)
+```
+
 Can you spot any seasonality, cyclicity and trend? What do you learn about the series?
 
-## 2.6
+**There is a clear seasonality increase in retail sales from October to the end of the year.
+This is the Christmas shopping season.
+There is also a trend of increasing retail sales over time.
+The trend has been rising untill the 2000's where it flatened out for roughly a decade.
+Since 2010 it looks like the trend increases again.**
+
+## Question 2.6
 
@@
 
+### hsales
+
 * Can you spot any seasonality, cyclicity and trend?
 
+```{r}
+autoplot(hsales)
+ggseasonplot(hsales)
+ggsubseriesplot(hsales)
+gglagplot(hsales)
+ggAcf(hsales)
+```
+
 * What do you learn about the series?
+
+### usdeaths
+
+* Can you spot any seasonality, cyclicity and trend?
+
+```{r}
+autoplot(usdeaths)
+ggseasonplot(usdeaths)
+ggsubseriesplot(usdeaths)
+gglagplot(usdeaths)
+ggAcf(usdeaths)
+```
+
+* What do you learn about the series?
+
+### bricksq
+
+* Can you spot any seasonality, cyclicity and trend?
+
+```{r}
+autoplot(bricksq)
+ggseasonplot(bricksq)
+ggsubseriesplot(bricksq)
+gglagplot(bricksq)
+ggAcf(bricksq)
+```
+
+* What do you learn about the series?
+
+### sunspotarea
+
+* Can you spot any seasonality, cyclicity and trend?
+
+```{r}
+autoplot(sunspotarea)
+#ggseasonplot(sunspotarea)
+#ggsubseriesplot(sunspotarea)
+gglagplot(sunspotarea)
+ggAcf(sunspotarea)
+```
+
+* What do you learn about the series?
+
+### gasoline
+
+* Can you spot any seasonality, cyclicity and trend?
+
+```{r}
+autoplot(gasoline)
+ggseasonplot(gasoline)
+#ggsubseriesplot(gasoline)
+gglagplot(gasoline)
+ggAcf(gasoline)
+```
+
+* What do you learn about the series?

### 2020-02-04 DATA612/data612.py
Initial commit

@@ -1,0 +1,117 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sun Feb  1 08:25:31 2020
+
+@author: Michael Silva
+"""
+import numpy as np
+import pandas as pd
+
+
+def train_test_split(user_item_df, train_proportion=0.8, random_seed=42):
+    """Splits a data frame into two data frames.
+    Args:
+        user_item_df (DataFrame): the pandas dataframe of that is a user item matrix.
+        train_proportion (float): the proportion of the non N/A data in the training set (Optional - 80% default)
+        random_seed (int): the random number seed (Optional - 42 default).
+    Returns:
+        train_df (DataFrame): The training set dataframe.
+        test_df (DataFrame): The testing set dataframe.
+    """
+    train_df = user_item_df.copy()
+    np.random.seed(random_seed)
+    # Count how many non N/A values are in the data frame
+    has_data = train_df.count().sum(axis=0)
+    # Determine how many values we need in the test set (there is no N/A's in the test set)
+    n_test = has_data - int(round(has_data * train_proportion, 0))
+    # Create an empty test data frame
+    test_df = pd.DataFrame(np.nan, index=train_df.index, columns=train_df.columns)
+    # Fill it with randomly selected values
+    while test_df.count().sum(axis=0) < n_test:
+        # Randomly select a row & column
+        row_id = np.random.choice(list(train_df.index), 1)
+        col_id = np.random.choice(list(train_df.columns), 1)
+        # Get the data at that location
+        val = train_df.iloc[row_id][col_id].values[0][0]
+        # Check to see if it is not N/A
+        if not np.isnan(val):
+            # Remove it from the training set
+            train_df.at[row_id, col_id] = None
+            # Save it to the test set
+            test_df.at[row_id, col_id] = val
+    return (train_df, test_df)
+
+
+def one_or_na(x):
+    """Returns NA if it's an NA or 1.
+    Args:
+        x (mixed): value from pandas dataframe
+    Returns:
+        one_or_na (mixed): Either 1 or NA
+    """
+    if np.isnan(x):
+        return x
+    else:
+        return 1
+
+
+def get_RMSE(user_item_df, predictor):
+    """Calculates the RMSE for the predictor for the dataframe
+    Args:
+        user_item_df (DataFrame): the pandas dataframe of that is a user item matrix.
+        predictor (float): the predicted value.
+    Returns:
[…]

### 2020-02-05 DATA624/HW1.Rmd
More work on HW1

@@
 ```
+
+**`gold` is an annual time series, `woolyrnq` is a quarterly time series and `gas` is a monthly time series**
 
@@
 
+**One-family home sales in the US tend to be highest in March.  The ACF plot suggests that there is some annual cycle, but it is noisy.**
+
 * What do you learn about the series?
+
+**If I am a realtor I'm not going to be busy in the winter months.  Early spring (march through may) will be my busy time of the year.**
 
@@
 
+**There appears to be a monthly pattern to the data.**
+
 * What do you learn about the series?
+
+**Accidental deaths in the US tends to be highest in July.**
 
@@
 
+**There isn't much variation from Q2 through Q4.  The trend was genrally increasing until about the 1980's.**
+
 * What do you learn about the series?
+
+**Q1 is a slow quarter for Australian clay brick producers.  There has been quite a bit more irregularity since 1975.**
 
@@
 
+**There appears to be a cycle in the data.  It looks like it's about a 10 to 11 year cycle.**
+
 * What do you learn about the series?
+
+**If the pattern holds 2020 should be a year with low sunspot area.**
 

### 2020-02-08 DATA624/HW1.Rmd
Updated HW1

@@
 
+**It stacks the visualization instead of having small multiples.**
+
 ## Question 2.3
@@
 
-**There appears to be a monthly pattern to the data.**
+**There appears to be a seasonal pattern to the data.**
 
@@
 ggseasonplot(gasoline)
-#ggsubseriesplot(gasoline)
+gasoline %>%
+  as.vector()%>%
+  ts(., frequency=52) %>%
+  ggsubseriesplot()
 gglagplot(gasoline)
@@
 
+**There is a trend and some seasonality to the data.  I thought you would see some cyclical behavior coinsiding with U.S. regression dates but that is not present.**
+
 * What do you learn about the series?
+
+**The trend of the supply of gasoline has been generally increasing.  It increases slightly during the summer months.**

### 2020-02-08 DATA624/HW2.Rmd
Initial commit of Homework 2

---
title: "DATA 624 Homework 2"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(fpp2)
```

## Question 3.1

For the following series, find an appropriate Box-Cox transformation in order to stabilise the variance.

* `usnetelec`
* `usgdp`
* `mcopper`
* `enplanements`

## Question 3.2

Why is a Box-Cox transformation unhelpful for the `cangas` data?

## Question 3.3

What Box-Cox transformation would you select for your `retail` data?

## Question 3.8

For your `retail` time series:

Split the data into two parts using
```{r}
myts.train <- window(myts, end=c(2010,12))
myts.test <- window(myts, start=2011)
```

Check that your data have been split appropriately by producing the following plot.

```{r}
autoplot(myts) +
  autolayer(myts.train, series="Training") +
  autolayer(myts.test, series="Test")
```

Calculate forecasts using snaive applied to myts.train.

```{r}
fc <- snaive(myts.train)
```

Compare the accuracy of your forecasts against the actual values stored in myts.test.

```{r}
accuracy(fc,myts.test)
```

Check the residuals.

```{r}
checkresiduals(fc)
```

Do the residuals appear to be uncorrelated and normally distributed?

How sensitive are the accuracy measures to the training/test split?

### 2020-02-11 DATA612/data612.py
Added docstrings and reordered functions

@@ -8,6 +8,76 @@ import numpy as np
 import pandas as pd
 
 
+def get_baseline_predictions(raw_avg, user_bias, item_bias):
+    """Creates baseline predictions using the user and item biases
+    Args:
+        raw_avg (float): the mean of the user item matrix.
+        user_bias (pandas.Series): A vector with the user biases.
+        item_bias (pandas.Series): A vector with the item bises.
+    Returns:
+        baseline_predictions (DataFrame): The predictions.
+    """
+    # Create a data farme by repeating the raw avg
+    baseline = pd.DataFrame(
+        raw_avg, index=np.arange(len(user_bias)), columns=item_bias.index
+    )
+    # Create a data frame by repeating the user bias series
+    user = pd.concat([pd.DataFrame(user_bias)] * len(item_bias), axis=1)
+    user.columns = item_bias.index
+    # Create a data frame by repeating the item bias series
+    item = pd.concat([pd.DataFrame(item_bias)] * len(user_bias), axis=1)
+    item.columns = user_bias.index
+    item = item.transpose()
+    # Bring the three components together to make the baseline predictions
+    baseline_predictions = baseline + user + item
+    # Round the values to the nearest integer between -10 and 10
+    baseline_predictions = (
+        baseline_predictions.round(0).astype("Int64").applymap(valid_val)
+    )
+    return baseline_predictions
+
+def get_biases(user_item_df, predictor):
+    """Calculates the user and item biases for the dataframe
+    Args:
+        user_item_df (DataFrame): the pandas dataframe of that is a user item matrix.
+        predictor (float): the predicted value.
+    Returns:
+        user_bias (pandas.Series): The biases for all of the rows.  
+        item_bias (pandas.Series): The biases for all of the columns.
+    """
+    user_mean = user_item_df.mean(axis=1)
+    item_mean = user_item_df.mean(axis=0)
+    user_bias = user_mean - predictor
+    item_bias = item_mean - predictor
+    return (user_bias, item_bias)
+
+def get_RMSE(user_item_df, predictor):
+    """Calculates the RMSE for the predictor for the dataframe
+    Args:
+        user_item_df (DataFrame): the pandas dataframe of that is a user item matrix.
+        predictor (float): the predicted value.
+    Returns:
+        RMSE_df (DataFrame): a data frame with the RMSE values.
+    """
[…]

### 2020-02-12 DATA612/data612.py
Finalized Project 1 Files

@@ -17,9 +17,9 @@ def get_baseline_predictions(raw_avg, user_bias, item_bias):
     Returns:
         baseline_predictions (DataFrame): The predictions.
     """
-    # Create a data farme by repeating the raw avg
+    # Create a data frame by repeating the raw avg
     baseline = pd.DataFrame(
-        raw_avg, index=np.arange(len(user_bias)), columns=item_bias.index
+        raw_avg, index=user_bias.index, columns=item_bias.index
     )
     # Create a data frame by repeating the user bias series
     user = pd.concat([pd.DataFrame(user_bias)] * len(item_bias), axis=1)
@@ -30,10 +30,6 @@ def get_baseline_predictions(raw_avg, user_bias, item_bias):
     item = item.transpose()
     # Bring the three components together to make the baseline predictions
     baseline_predictions = baseline + user + item
-    # Round the values to the nearest integer between -10 and 10
-    baseline_predictions = (
-        baseline_predictions.round(0).astype("Int64").applymap(valid_val)
-    )
     return baseline_predictions
 
 def get_biases(user_item_df, predictor):
@@ -65,6 +61,18 @@ def get_RMSE(user_item_df, predictor):
     mean_squared_errors = squared_errors.stack().mean()
     RMSE = mean_squared_errors ** (1 / 2)
     return RMSE
+
+def get_valid_jester_predictions(baseline_predictions):
+    """Round the values to the nearest integer between -10 and 10
+    Args:
+        baseline_predictions (DataFrame): the pandas dataframe of predictions.
+    Returns:
+        baseline_predictions (DataFrame): a data frame with valid predictions.
+    """
+    baseline_predictions = (
+        baseline_predictions.round(0).astype("Int64").applymap(valid_jester_val)
+    )
+    return baseline_predictions
 
 def one_or_na(x):
     """Returns NA if it's an NA or 1.
@@ -111,8 +119,8 @@ def train_test_split(user_item_df, train_proportion=0.8, random_seed=42):
             test_df.at[row_id, col_id] = val
     return (train_df, test_df)
 
-def valid_val(x):
-    """Validates the predicted rating
+def valid_jester_val(x):
+    """Validates the jester predicted rating
     Args:
         x (float): the predicted rating to validate.
     Returns:

### 2020-02-14 DATA624/Prophet Example.R
Initial commit

@@ -1,0 +1,30 @@
+## Prophet Example
+library(tidyverse)
+library(prophet)
+
+## Download NIFRS data
+if(!file.exists("NIFRS.csv")){
+  download.file("https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA624/NIFRS.csv", "NIFRS.csv")
+}
+
+## Read in the data
+df <- read.csv("NIFRS.csv")
+
+## Split out training set
+training <- df %>%
+  filter(YEAR < 2018) %>%
+  rename(ds = IN_DATE) %>%
+  group_by(ds) %>%
+  tally() %>%
+  rename(y = n)
+
+## Build a model & produce forecasts
+model <- prophet(training)
+future_df <- make_future_dataframe(model, periods = 365)
+forecast <- predict(model, future_df)
+
+## Plot the model forecasts
+plot(model, forecast)
+
+## Plot the decomposition
+prophet_plot_components(model, forecast)

### 2020-02-14 DATA624/HW2.Rmd
More work on HW2

@@
 library(fpp2)
+library(gridExtra)
+library(readxl)
 ```
@@
 
-* `usnetelec`
-* `usgdp`
-* `mcopper`
-* `enplanements`
+**I will answer these questions with side by side visualizations of the data.  The appropriate Box-Cox transformation will be on the right hand side with the lambda in the title.**
+
+```{r}
+side_by_side <- function(x, y){
+  lambda <- BoxCox.lambda(x)
+  plot1 <- autoplot(x) +
+    ggtitle("Original") +
+    ylab(y) + 
+    theme(axis.title.x = element_blank())
+  plot2 <- autoplot(BoxCox(x, lambda)) +
+    ggtitle(paste0("Box-Cox Transformed (lambda=", round(lambda, 4),")")) +
+    theme(axis.title.x = element_blank(),
+        axis.title.y = element_blank()) 
+  grid.arrange(plot1, plot2, ncol = 2)
+}
+```
+
+### usnetelec
+
+```{r}
+side_by_side(usnetelec, "US Net Electricity Generation")
+```
+
+### usgdp
+
+```{r}
+side_by_side(usgdp, "US GDP")
+```
+
+### mcopper
+
+```{r}
+side_by_side(mcopper, "Monthly Copper Prices")
+```
+
+### enplanements
+
+```{r}
+side_by_side(enplanements, "Monthly US Domestic Enplanements")
+```
 
@@
 
+```{r}
+side_by_side(cangas, "Monthly Canadian Gas Production")
+```
+
+**The Box-Cox transformation does not help with this timeseries because the variation is initially small, then gets large, then gets small again.
+Box-Cox was not designed to handle this case.
+It was designed for cases where the variance increases or decreases over time.**
+
 ## Question 3.3
@@
 What Box-Cox transformation would you select for your `retail` data?
+
+```{r}
+retaildata <- read_excel("retail.xlsx", skip = 1)
+myts <- ts(retaildata[, "A3349873A"], frequency = 12, start = c(1982, 4))
+side_by_side(myts, "Retail Sales")
+```
+
+**The variation was increasing over time in the original data.
+It has become significantly more uniform once it is transformed with a lambda of 0.13.
+Because the variance was increasing over time this was an effective transformation.**
 
@@
 ```{r}
-myts.train <- window(myts, end=c(2010,12))
-myts.test <- window(myts, start=2011)
+myts.train <- window(myts, end = c(2010, 12))
+myts.test <- window(myts, start = 2011)
 ```
@@
 autoplot(myts) +
-  autolayer(myts.train, series="Training") +
-  autolayer(myts.test, series="Test")
+  autolayer(myts.train, series = "Training") +
+  autolayer(myts.test, series = "Test")
 ```
@@
 ```{r}
-accuracy(fc,myts.test)
+accuracy(fc, myts.test)
 ```
@@
 
+**They do appear to be normally distributed however with a sligh positve skew.
+The residuals do no appear to be uncorrelated.
+The Ljung-Box test has a p value that is less than 0.05.
+This suggests there is more information that can be discovered and that the seasonal naive model is not the best model.**
+
 How sensitive are the accuracy measures to the training/test split?
+
+**The accuracy measures are quite sensitive to the training/test split. The values are significantly different between the two.  This would suggest that the model does not generalize well.**

### 2020-02-18 DATA624/HW3.Rmd
Initial commit

---
title: "DATA 624 Homework 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(fpp2)
library(forcats)
```

## Question 6.2

The `plastics` data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.

* Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?

```{r}
autoplot(plastics)
```

***The `plastics` data has has an increasing trend and a cycle of 1 year***

* Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

```{r}
decompose_plastics <- decompose(plastics, type="multiplicative")

decompose_plastics %>% 
  autoplot() +
  ggtitle("Multiplicative Decomposition of Plastic Product A Sales")
```

```{r, eval=FALSE, echo=FALSE}
trend_cycle_component <- ma(ma(plastics, order = 12, centre = TRUE), order = 2)
detrended_plastics <- plastics / trend_cycle_component
seasonal_component <- tapply(detrended_plastics, cycle(detrended_plastics), mean)

decompose(plastics)
```

* Do the results support the graphical interpretation from part a?

**Yes**

* Compute and plot the seasonally adjusted data.

```{r}
seasonally_adjusted_plastics <- plastics - decompose_plastics$seasonal
autoplot(seasonally_adjusted_plastics)
```

* Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

* Does it make any difference if the outlier is near the end rather than in the middle of the time series?

## Question 6.3

Recall your retail time series data (from Exercise 3 in Section 2.10). Decompose the series using X11. Does it reveal any outliers, or unusual features that you had not noticed previously?

### 2020-02-20 DATA624/HW3.Rmd
More work on HW3

@@
 library(tidyverse)
-library(kableExtra)
 library(fpp2)
 library(forcats)
+library(readxl)
+library(seasonal)
 ```
@@
 
-***The `plastics` data has has an increasing trend and a cycle of 1 year***
+***The `plastics` data has has an increasing trend and a seasonal component where sales are higher in the summer and lower in the winter.***
 
@@
 detrended_plastics <- plastics / trend_cycle_component
-seasonal_component <- tapply(detrended_plastics, cycle(detrended_plastics), mean)
+seasonal_component <- tapply(detrended_plastics, cycle(detrended_plastics), mean, na.rm=TRUE)
+seasonal_component
 
-decompose(plastics)
+decompose(plastics, type = "multiplicative")
 ```
@@
 
-**Yes**
+**Yes.  The trend is increasing and there's a seasonal component.**
 
@@
 ```{r}
-seasonally_adjusted_plastics <- plastics - decompose_plastics$seasonal
-autoplot(seasonally_adjusted_plastics)
+seasonally_adjusted_plastics <- plastics / decompose_plastics$seasonal
+autoplot(plastics, series = "original data") +
+  autolayer(seasonally_adjusted_plastics, series = "seasonally adjusted") +
+  ylab("Sales (thousands)") +
+  ggtitle("Plastic Product A Sales") +
+  scale_color_brewer(palette = "Set1")
 ```
@@
 
+```{r}
+plastics_with_outlier <- plastics
+plastics_with_outlier[20] <- plastics_with_outlier[20] + 500
+decompose_plastics_with_outlier <- decompose(plastics_with_outlier, type="multiplicative")
+seasonally_adjusted_plastics_with_outlier <- plastics_with_outlier / decompose_plastics_with_outlier$seasonal
+
+autoplot(plastics, series = 'original data') +
+  autolayer(seasonally_adjusted_plastics, series = 'without outlier') +
+  autolayer(seasonally_adjusted_plastics_with_outlier, series = 'with outlier') +
+  ylab("Sales (thousands)") +
+  ggtitle("Seasonally Adjusted Plastic Product A Sales") +
+  scale_color_brewer(palette = "Set1")
+```
+
+**The outlier causes the series to be slighly larger, and the troughs to be overstated.**
+
 * Does it make any difference if the outlier is near the end rather than in the middle of the time series?
+
+```{r}
+plastics_with_outlier_near_the_end <- plastics
+plastics_with_outlier_near_the_end[50] <- plastics_with_outlier_near_the_end[50] + 500
+decompose_plastics_with_outlier_near_the_end <- decompose(plastics_with_outlier_near_the_end, type="multiplicative")
+seasonally_adjusted_plastics_with_outlier_near_the_end <- plastics_with_outlier_near_the_end / decompose_plastics_with_outlier_near_the_end$seasonal
+
+plastics_with_outlier_in_the_middle <- plastics
+plastics_with_outlier_in_the_middle[30] <- plastics_with_outlier_in_the_middle[30] + 500
+decompose_plastics_with_outlier_in_the_middle <- decompose(plastics_with_outlier_in_the_middle, type="multiplicative")
+seasonally_adjusted_plastics_with_outlier_in_the_middle <- plastics_with_outlier_in_the_middle / decompose_plastics_with_outlier_in_the_middle$seasonal
+
+autoplot(plastics, series = 'original data') +
+  autolayer(seasonally_adjusted_plastics_with_outlier_near_the_end, series = 'with outlier near the end') +
+  autolayer(seasonally_adjusted_plastics_with_outlier_in_the_middle, series = 'with outlier in the middle') +
+  ylab("Sales (thousands)") +
+  ggtitle("Seasonally Adjusted Plastic Product A Sales") +
+  scale_color_brewer(palette = "Set1")
+```
+
+**The outlier has less of an impact if it's in the middle of the series as opposed to near the end.
+It's interesting that when the outlier is near the end the seasonal adjustment doesn't adjust out the troughs of the orignal series.
+When it's in the middle it seems to just introduce some noise.**
 
@@
 Recall your retail time series data (from Exercise 3 in Section 2.10). Decompose the series using X11. Does it reveal any outliers, or unusual features that you had not noticed previously?
+
+```{r}
+retaildata <- read_excel("retail.xlsx", skip = 1)
+retail <- ts(retaildata[, "A3349337W"], frequency = 12, start = c(1982, 4))
+x11_retail <- seas(retail, x11="")
+autoplot(x11_retail) +
+  ggtitle("X11 Decomposition of Retail Sales Data")
+```
+
+**There are some spikes in the remainder early on (circa 1983) and around 2000.  I also find it interesting to see the seasonal component is decreasing over time in the X11 decomposition.**

### 2020-02-21 DATA624/HW3.Rmd
More work on HW 3

@@
 
-**The outlier causes the series to be slighly larger, and the troughs to be overstated.**
+**The outlier causes the series to be slighly higher than the seasonally adjusted series without an outlier.
+There are also some troughs in the series with an outlier that don't exist in the series without an outlier.
+So the addition of the outlier causes the model to find troughs in the data it wouldn't otherwise, and slighly overstate the level of the seasonal component.**
 
@@
 ```{r}
+# Near the End
 plastics_with_outlier_near_the_end <- plastics
@@
 
+# In the Middle
 plastics_with_outlier_in_the_middle <- plastics
@@
 
+# Plot both series
 autoplot(plastics, series = 'original data') +
@@
 
-**There are some spikes in the remainder early on (circa 1983) and around 2000.  I also find it interesting to see the seasonal component is decreasing over time in the X11 decomposition.**
+**There are some spikes in the remainder early on (circa 1983) and around 2000.  That indicates the presense of some outliers.**

### 2020-02-24 DATA624/HW4.Rmd
Initial commit

---
title: "DATA 624 Homework 4"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(fpp2)
library(forecast)
library(readxl)
library(seasonal)
```

## Question 3.1

The UC Irvine Machine Learning Repository contains a data set related to glass identification.  The data  consists of 214 glass samples labeled as one of sever class categories.  There are nine predictiors, including the refractive index and percentages of eight elements: Na, Mg, Al, Si, K, Ca, Ba, and Fe.  The data can be accessed via:

```{r}
library(mlbench)
data(Glass)
str(Glass)
```

(a) Using visualizations, explore the predictor variables to understand their distributions as well as the relationships between predictors

(b) Do there appear to be any outliers in the data?  Are any predictors skewed?

(c) Are there any relevant transformations of one or more predictors that might improve the classification model?

## Question 3.2

The soybean data can also be found at the UC Irvine Machine Learning Repository.  Data were collected to predict disease in 683 soybeans.  The 35 predictors are mostly categorical and include information on the environemental conditions (e.g. temperature, precipitation) and plant conditions (e.g., left spots, mold growth).  The outcome labels consist of 19 distinct classes.  The data can be loaded via:

```{r}
library(mlbench)
data(Soybean)
## See ?Soybean for details
```

(a) Investigate the frequency distributions for the categhorical predictors.  Are any of the distributions degenerate in the ways discussed earlier in this chapter?

(b) Roughly 18% of the data are missing.  Are there particular predictors that are more likely to be missing?  Is the pattern of missing data related to the classes?

(c) Develop a strategy for handling missing data, either by eliminating predictors or imputation.

### 2020-02-25 DATA612/data612.py
More work on Project 2

@@ -4,6 +4,7 @@ Created on Sun Feb  1 08:25:31 2020
 
 @author: Michael Silva
 """
+import re
 import numpy as np
 import pandas as pd
 
@@ -74,6 +75,20 @@ def get_valid_jester_predictions(baseline_predictions):
     )
     return baseline_predictions
 
+def is_plus_or_minus_five(x):
+    """Returns a value between -5 and 5.
+    Args:
+        x (int): value that may be outside of the -5 to 5 range
+    Returns:
+        valid_x (int): x within the -5 to 5 range
+    """
+    if x > 5:
+        return 5
+    elif x < -5:
+        return -5
+    else:
+        return x
+
 def one_or_na(x):
     """Returns NA if it's an NA or 1.
     Args:
@@ -86,6 +101,43 @@ def one_or_na(x):
     else:
         return 1
 
+def read_joke(n):
+    """Reads in the text of the jester joke
+    Args:
+        n (int): the joke number (1 to 100)
+    Returns:
+        joke_text (str): the content of the joke in one line
+    """
+    joke_text = ""
+    joke_begin = False
+    with open('jokes/init'+str(n)+'.html') as f:
+        for line in f.readlines():
+            if "end of joke" in line:
+                joke_begin = False
+            if joke_begin:
+                joke_text += line.strip() + " "
+            elif "begin of joke" in line:
+                joke_begin = True
+    # Strip all html tags
+    html = re.compile('<.*?>')
+    joke_text = re.sub(html, '', joke_text).strip()
+    # Remove all the extra spaces
+    joke_text = ' '.join(joke_text.split())
+    return joke_text
+
+def rescale_jester_ratings(df):
+    """Convert data to a -5 to 5 integer scale
+    Args:
+        df (DataFrame): the pandas dataframe of jester ratings.
+    Returns:
+        rescaled_df (DataFrame): a data frame with valid rescaled ratings.
+    """
+    df = df / 2
+    rescaled_df = (
+        df.round(0).astype("Int64").applymap(is_plus_or_minus_five)
+    )
+    return rescaled_df
+    
 def train_test_split(user_item_df, train_proportion=0.8, random_seed=42):
     """Splits a data frame into two data frames.
     Args:

### 2020-02-25 DATA612/data612.py
More work on Project 2

@@ -9,6 +9,42 @@ import numpy as np
 import pandas as pd
 
 
+def bag_of_words(sentences, case_sensitive=False, na_fill=0):
+    """Transforms a list of sentences into a bag of words matrix
+    Args:
+        sentences (list): A list of sentences
+        case_sensitive (bool): Should the bag of words be case sensitive (Optional - False default)
+        na_fill (mixed): What should fill the NA's? (Optional - 0 default)
+    Returns:
+        bag_of_words_df (DataFrame): Dataframe with word counts by sentence.
+    """
+    # Get the count of words in each sentence
+    bag = dict()
+    n = 0
+    for sentence in sentences:
+        n += 1
+        for word in sentence.split():
+            if not case_sensitive:
+                # Make bag of words case insensitive
+                word = word.lower()
+            # Use a tuple as the key holding the word and the sentence index number
+            key = (word, n)
+            # Count the word
+            bag[key] = bag.get(key, 0) + 1
+    # Convert the count into a long data frame
+    bag_of_words_data = list()
+    for k in bag.keys():
+        row = {"word": k[0], "index": k[1], "count": bag[k]}
+        bag_of_words_data.append(row)
+    bag_of_words_df = pd.DataFrame(bag_of_words_data)
+    # Convert from long to wide dataframe
+    bag_of_words_df = bag_of_words_df.pivot_table(
+        index="index", columns="word", values="count", fill_value=na_fill
+    )
+
+    return bag_of_words_df
+
+
 def get_baseline_predictions(raw_avg, user_bias, item_bias):
     """Creates baseline predictions using the user and item biases
     Args:
@@ -19,9 +55,7 @@ def get_baseline_predictions(raw_avg, user_bias, item_bias):
         baseline_predictions (DataFrame): The predictions.
     """
     # Create a data frame by repeating the raw avg
-    baseline = pd.DataFrame(
-        raw_avg, index=user_bias.index, columns=item_bias.index
-    )
+    baseline = pd.DataFrame(raw_avg, index=user_bias.index, columns=item_bias.index)
     # Create a data frame by repeating the user bias series
     user = pd.concat([pd.DataFrame(user_bias)] * len(item_bias), axis=1)
     user.columns = item_bias.index
@@ -32,6 +66,7 @@ def get_baseline_predictions(raw_avg, user_bias, item_bias):
[…]

### 2020-02-26 DATA612/data612.py
More work on Project 2

@@ -4,17 +4,22 @@ Created on Sun Feb  1 08:25:31 2020
 
 @author: Michael Silva
 """
-import re
+import contractions
+import nltk
 import numpy as np
 import pandas as pd
+import re
+import sklearn.preprocessing
+import sklearn.metrics.pairwise
+import string
 
 
-def bag_of_words(sentences, case_sensitive=False, na_fill=0):
+def bag_of_words(sentences, na_fill=0, normalize=False):
     """Transforms a list of sentences into a bag of words matrix
     Args:
         sentences (list): A list of sentences
-        case_sensitive (bool): Should the bag of words be case sensitive (Optional - False default)
         na_fill (mixed): What should fill the NA's? (Optional - 0 default)
+        normalize: If the bag of words should be normalized (Optional - False default)
     Returns:
         bag_of_words_df (DataFrame): Dataframe with word counts by sentence.
     """
@@ -23,10 +28,8 @@ def bag_of_words(sentences, case_sensitive=False, na_fill=0):
     n = 0
     for sentence in sentences:
         n += 1
+        sentence = sentence.lower()
         for word in sentence.split():
-            if not case_sensitive:
-                # Make bag of words case insensitive
-                word = word.lower()
             # Use a tuple as the key holding the word and the sentence index number
             key = (word, n)
             # Count the word
@@ -41,8 +44,34 @@ def bag_of_words(sentences, case_sensitive=False, na_fill=0):
     bag_of_words_df = bag_of_words_df.pivot_table(
         index="index", columns="word", values="count", fill_value=na_fill
     )
+    if normalize:
+        normalized_bag_of_words = sklearn.preprocessing.normalize(bag_of_words_df)
+        bag_of_words_df = pd.DataFrame(
+            normalized_bag_of_words,
+            columns=bag_of_words_df.columns,
+            index=bag_of_words_df.index,
+        )
 
     return bag_of_words_df
+
+
+def cosine_similarity(bag_of_words, return_difference=False):
+    """Measures the cosine similarity or difference of a term frequency matrix
+    Args:
+        bag_of_words (DataFrame): The term frequency matrix
+        return_difference (bool): Do you want the cosine difference returned? (Optional False default)
+    Returns:
+        cosine_df (DataFrame): The cosine similarity or difference in data frame form
+    """
+    bag_of_words_cosine_metric = sklearn.metrics.pairwise.cosine_similarity(
+        bag_of_words
+    )
[…]

### 2020-02-26 DATA624/HW4.Rmd
Minor change

@@
 library(tidyverse)
-library(fpp2)
-library(forecast)
-library(readxl)
-library(seasonal)
 ```
@@
 
-(a) Investigate the frequency distributions for the categhorical predictors.  Are any of the distributions degenerate in the ways discussed earlier in this chapter?
+(a) Investigate the frequency distributions for the categorical predictors.  Are any of the distributions degenerate in the ways discussed earlier in this chapter?
 

### 2020-02-29 DATA612/data612.py
More work on Project 2

@@ -11,6 +11,7 @@ import pandas as pd
 import re
 import sklearn.preprocessing
 import sklearn.metrics.pairwise
+import sklearn.feature_extraction.text
 import string
 
 
@@ -55,21 +56,21 @@ def bag_of_words(sentences, na_fill=0, normalize=False):
     return bag_of_words_df
 
 
-def cosine_similarity(bag_of_words, return_difference=False):
-    """Measures the cosine similarity or difference of a term frequency matrix
+def cosine_similarity(df, return_difference=False):
+    """Measures the cosine similarity or difference of a data frame
     Args:
-        bag_of_words (DataFrame): The term frequency matrix
+        df (DataFrame): The data frame of values to compute the cosine similarity/difference 
         return_difference (bool): Do you want the cosine difference returned? (Optional False default)
     Returns:
         cosine_df (DataFrame): The cosine similarity or difference in data frame form
     """
-    bag_of_words_cosine_metric = sklearn.metrics.pairwise.cosine_similarity(
-        bag_of_words
+    cosine_metric = sklearn.metrics.pairwise.cosine_similarity(
+        df
     )
     if return_difference:
-        bag_of_words_cosine_metric = 1 - bag_of_words_cosine_metric
+        cosine_metric = 1 - cosine_metric
     cosine_df = pd.DataFrame(
-        bag_of_words_cosine_metric, index=bag_of_words.index, columns=bag_of_words.index
+        cosine_metric, index=df.index, columns=df.index
     )
     return cosine_df
 
@@ -111,6 +112,26 @@ def get_biases(user_item_df, predictor):
     user_bias = user_mean - predictor
     item_bias = item_mean - predictor
     return (user_bias, item_bias)
+
+
+def get_cosine_based_joke_recommendations(cosine_similarity_df, joke_id, n_recommendations = 10):
+    """Translate a cosine similarity data frame into an ordered list of recommendations.
+    Args:
+        cosine_similarity_df (DataFrame):
+        joke_id (int): the id of the joke (column number)
+        n_recommendations (int): The number of recommendations (Optional - 10 default)
+    Returns:
+        recommendations (list): An ordered list of recommendations
+    """
+    # Transform the row of cosine values into a three column data frame (level_0 = joke_id, level_1 = cosine_similarity_df.columns, 0 = cosine value)
+    temp = cosine_similarity_df[:joke_id].stack().reset_index()
[…]

### 2020-02-29 DATA624/HW4.Rmd
More work on HW4

@@
 library(tidyverse)
+library(corrplot)
+library(kableExtra)
 ```
@@
 The UC Irvine Machine Learning Repository contains a data set related to glass identification.
-The data  consists of 214 glass samples labeled as one of sever class categories.
-There are nine predictiors, including the refractive index and percentages of eight elements: Na, Mg, Al, Si, K, Ca, Ba, and Fe.
+The data  consists of 214 glass samples labeled as one of several class categories.
+There are nine predictors, including the refractive index and percentages of eight elements: Na, Mg, Al, Si, K, Ca, Ba, and Fe.
 The data can be accessed via:
@@
 (a) Using visualizations, explore the predictor variables to understand their distributions as well as the relationships between predictors
+
+**First we will take a look at the distribution of the predictors:**
+
+```{r}
+Glass %>%
+  pivot_longer(-Type, names_to = "Element", values_to = "Value", values_drop_na = TRUE) %>%
+  mutate(Element = as.factor(Element)) %>%
+  ggplot(aes(Value, color = Element, fill = Element)) +
+  geom_histogram(binwidth = 1) +
+  facet_wrap(~ Element, ncol=3) +
+  scale_fill_brewer(palette = "Set1") +
+  scale_color_brewer(palette = "Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  ggtitle("Distribution of Predictor Variables")
+```
+
+**Next let's look at the correlation plot.**
+
+```{r}
+col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
+
+Glass %>%
+  select(-Type) %>%
+  cor() %>%
+  round(., 2) %>%
+  corrplot(., method="color", col=col(200), type="upper", order="hclust", addCoef.col = "black", tl.col="black", tl.srt=45, diag=FALSE )
+```
+
+```{r}
+Glass %>%
+  pivot_longer(-Type, names_to = "Element", values_to = "Value", values_drop_na = TRUE) %>%
+  mutate(Element = as.factor(Element)) %>%
+  group_by(Element) %>%
+  summarise(Min = min(Value),
+            Median = median(Value),
+            Mean = mean(Value),
+            Max = max(Value)) %>%
+  kable() %>%
+  kable_styling()
+#1st Qu. and 3rd Qu.
+```
+
+**RI is the refractive index.
+All other predictors are elements.
+Elements unit of measurement, according to [UCI documentation](http://archive.ics.uci.edu/ml/datasets/Glass+Identification) is the "weight percent in corresponding oxide."  There is no documentation on the refractive index provided by UCI.**
+
+**It is not suprising that Silicon has the highest weight percentage in the glass samples.**
 

### 2020-02-29 DATA624/HW4.Rmd
More work on HW4

--- 2020-02-26 DATA624/HW4.Rmd
@@
 
+**Before looking at the relationships I think it would be helpful to know a bit more about the predictors.
+T element concentrations predictors' unit of measurement, according to [UCI documentation](http://archive.ics.uci.edu/ml/datasets/Glass+Identification) is the "weight percent in corresponding oxide."  There is no documentation on the refractive index provided by UCI.
+But according to Wikipedia, the refractive index measure how much light slows down by passing through a material.
+An increasing refractive index corresponds to decreasing speed of light in the material.**
+
+**Now that we have an inkling of what the data represent, we will take a look at the distribution of the predictors:**
+
+```{r}
+long_glass <- Glass %>%
+  pivot_longer(-Type, names_to = "Predictor", values_to = "Value", values_drop_na = TRUE) %>%
+  mutate(Predictor = as.factor(Predictor))
+
+long_glass %>%
+  ggplot(aes(Value, color = Predictor, fill = Predictor)) +
+  geom_histogram(bins = 20) +
+  facet_wrap(~ Predictor, ncol = 3, scales = "free") +
+  scale_fill_brewer(palette = "Set1") +
+  scale_color_brewer(palette = "Set1") +
+  theme_light() +
+  theme(legend.position = "none") +
+  ggtitle("Distribution of Predictor Variables")
+```
+
+**Glass is primarly made of silica (Si), soda (Na) and lime (Ca).  Seeing these predictors at higher concentrations is not suprising.**
+
+**Now we will examine how the predictors are related to eachother.  We will do that with a correlation plot.**
+
+```{r}
+#ColorBrewer's 5 class spectral color palette
+col <- colorRampPalette(c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"))
+
+Glass %>%
+  select(-Type) %>%
+  cor() %>%
+  round(., 2) %>%
+  corrplot(., method="color", col=col(200), type="upper", order="hclust", addCoef.col = "black", tl.col="black", tl.srt=45, diag=FALSE )
+```
+
+**Most of the predictors are negatively correlated, which makes sense.
+They are measuring chemical concentrations on a percentage basis.
+As one element increases we would expect a decrease in the others.**
+
+**Most of the correlations are not very strong.
+The exception to this is the correlation between calcium oxide and the refraction index is strongly positively correlated.
+I am going to take some liberties and summarize the data in a tabular form, because this "visualization" speaks to me:**
+
+```{r, echo=FALSE}
+long_glass %>%
+  group_by(Predictor) %>%
+  summarise(Min = min(Value),
+            `1st Qu.` = quantile(Value, .25),
+            Median = median(Value),
+            Mean = mean(Value),
+            `3rd Qu.` = quantile(Value, .75),
+            Max = max(Value)) %>%
+  kable() %>%
+  kable_styling()
+```
+
 (b) Do there appear to be any outliers in the data?  Are any predictors skewed?
 
+**I want to see how the predictors are distributed by the type of glass.  I will use a scatter plot to do this but will be excluding scilica because of the difference in scale.**
+
+```{r}
+long_glass %>%
+  ggplot(aes(x = Type, y = Value, color = Predictor)) +
+  geom_jitter() +
+  ylim(0, 20) + 
+  scale_color_brewer(palette = "Set1") +
+  theme_light()
+```
+
+**It looks like glass type 1, 2 and 3 are very similar in chemical composition.
+There are a couple of observations that appear to be outliers.
+For example there are a couple of potasium (K) observations in the type 5 glass that are unusually high.
+There is a barium (Ba) observation in type 2 glass that apears to be an outlier along with some calcium (Ca) observations in type 2 glass.**
+
+**Magnesium is bimodal and left skewed. Iron, potasium and barium are right skewed.  The other predictors are somewhat normal.**
+
 (c) Are there any relevant transformations of one or more predictors that might improve the classification model?
+
+**Something like a Box-Cox transformation might improve the classification model's preformance.**
 

### 2020-02-29 DATA624/HW4.Rmd
More work on HW4

@@
 library(corrplot)
+library(Amelia)
 library(kableExtra)
+library(caret)
 ```
@@
 
+**I am assuming the degenerate distibuted variaviables discussed earlier in the chapter referes to section 3.5 on removing predictors.  Here's some frequency tables:**
+
+```{r, results='asis'}
+for (predictor in names(select(Soybean, -Class))){
+  temp <- Soybean %>%
+    group_by(.dots=predictor) %>%
+    tally() %>%
+    arrange(desc(n)) 
+  temp %>%
+    summarise(total = sum(n)) %>%
+    merge(temp) %>%
+    mutate(share = n / total) %>%
+    select(-total) %>%
+    kable() %>%
+    kable_styling() %>%
+    print()
+}
+```
+
+**There's a lot of missing variables.
+The authors recommended removing variables with near zero variance.
+I know that the `caret` package has a function for that.
+Here's the output from that function:**
+
+```{r}
+nearZeroVar(Soybean, saveMetrics = T) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**There are three variables (`r names(Soybean)[caret::nearZeroVar(Soybean)]`) that have a near zero variance, and should probably be removed.**
+
 (b) Roughly 18% of the data are missing.  Are there particular predictors that are more likely to be missing?  Is the pattern of missing data related to the classes?
 
+```{r}
+Soybean %>%
+  arrange(Class) %>%
+  missmap(main = "Missing vs Observed")
+```
+
+**There are blocks of observations that are missing.  Since the data are arranged by the classes this suggests that the patterns of missing data are related to the classes.**
+
 (c) Develop a strategy for handling missing data, either by eliminating predictors or imputation.
+
+**In my training it has been encouraged to never get rid of data, so my bias would be towards filling them with imputation.**
+
+**I would first drop all missing variables and run it through a random forest to determine which variables are important.
+If the near zero variance variables are not high in the list, I would drop them since they have little informational value.**
+
+**If one of the important variables has a lot of missing data I would use k-means to fill in the holes with this categorical data.
+If the variables were quantitative I would probably use the mean or median to fill in the holes.**

### 2020-03-01 DATA612/data612.py
More work on 612 Project 2

@@ -9,9 +9,9 @@ import nltk
 import numpy as np
 import pandas as pd
 import re
-import sklearn.preprocessing
-import sklearn.metrics.pairwise
-import sklearn.feature_extraction.text
+from sklearn.preprocessing import normalize as sklearn_normalize
+from sklearn.metrics.pairwise import cosine_similarity as sklearn_cosine_similarity
+from sklearn.feature_extraction.text import TfidfVectorizer as sklearn_TfidfVectorizer
 import string
 
 
@@ -46,7 +46,7 @@ def bag_of_words(sentences, na_fill=0, normalize=False):
         index="index", columns="word", values="count", fill_value=na_fill
     )
     if normalize:
-        normalized_bag_of_words = sklearn.preprocessing.normalize(bag_of_words_df)
+        normalized_bag_of_words = sklearn_normalize(bag_of_words_df)
         bag_of_words_df = pd.DataFrame(
             normalized_bag_of_words,
             columns=bag_of_words_df.columns,
@@ -64,14 +64,10 @@ def cosine_similarity(df, return_difference=False):
     Returns:
         cosine_df (DataFrame): The cosine similarity or difference in data frame form
     """
-    cosine_metric = sklearn.metrics.pairwise.cosine_similarity(
-        df
-    )
+    cosine_metric = sklearn_cosine_similarity(df)
     if return_difference:
         cosine_metric = 1 - cosine_metric
-    cosine_df = pd.DataFrame(
-        cosine_metric, index=df.index, columns=df.index
-    )
+    cosine_df = pd.DataFrame(cosine_metric, index=df.index, columns=df.index)
     return cosine_df
 
 
@@ -114,7 +110,29 @@ def get_biases(user_item_df, predictor):
     return (user_bias, item_bias)
 
 
-def get_cosine_based_joke_recommendations(cosine_similarity_df, joke_id, n_recommendations = 10):
+def get_cosine_based_neighbors(cosine_similarity_df, col_id, n_neighbors = 10):
+    """Returns the n_neighbors closest neighbors using cosine similarity.
+    Args:
+        cosine_similarity_df (DataFrame): the data frame of cosine similarity
+        col_id (int): the column id to find neighbors for
+        n_neighbors (int): the number of neighbors (Optional - 10 default)
+    Returns:
+        neighbors (list): ordered list of closest neighbor ids
+    """
[…]

### 2020-03-01 DATA624/HW4.Rmd
More work on HW4

@@
 
-**Before looking at the relationships I think it would be helpful to know a bit more about the predictors.
-T element concentrations predictors' unit of measurement, according to [UCI documentation](http://archive.ics.uci.edu/ml/datasets/Glass+Identification) is the "weight percent in corresponding oxide."  There is no documentation on the refractive index provided by UCI.
-But according to Wikipedia, the refractive index measure how much light slows down by passing through a material.
-An increasing refractive index corresponds to decreasing speed of light in the material.**
-
-**Now that we have an inkling of what the data represent, we will take a look at the distribution of the predictors:**
+**First we will take a look at the distribution of the predictors:**
 
@@
 
-**Now we will examine how the predictors are related to eachother.  We will do that with a correlation plot.**
+**Now we will examine how the predictors are related to each other.  We will do that with a correlation plot.**
 
@@
 
-**I am assuming the degenerate distibuted variaviables discussed earlier in the chapter referes to section 3.5 on removing predictors.  Here's some frequency tables:**
+**I am assuming the degenerate distibuted variaviables discussed earlier in the chapter refers to section 3.5 on removing predictors.  Here's some frequency tables:**
 
@@
 
-**If one of the important variables has a lot of missing data I would use k-means to fill in the holes with this categorical data.
-If the variables were quantitative I would probably use the mean or median to fill in the holes.**
+**If one of the important variables has a lot of missing data I would use knn to fill in the holes with this categorical data.
+If the variables were quantitative I would probably use the mean or median or k-means to fill in those holes.**

### 2020-03-02 DATA624/HW4.Rmd
Wrapping up HW 4

@@
 
-**In my training it has been encouraged to never get rid of data, so my bias would be towards filling them with imputation.**
+**I will be eliminating the three near zero variance predictiors.
+For all other predictors I will be imputing values.
+I don't have any domain expertise that would inform the imputations, so I will be using decision trees to (hopefully) produce good imputations.
+It has been my experience that decision trees preform really well.
+The `dlookr` package integrates well with the tidyverse.**
 
-**I would first drop all missing variables and run it through a random forest to determine which variables are important.
-If the near zero variance variables are not high in the list, I would drop them since they have little informational value.**
+```{r}
+library(dlookr)
 
-**If one of the important variables has a lot of missing data I would use knn to fill in the holes with this categorical data.
-If the variables were quantitative I would probably use the mean or median or k-means to fill in those holes.**
+Soybean_complete <- Soybean %>%
+  # Impute missing values using rpart
+  mutate(
+    date = imputate_na(Soybean, date, Class, method = "rpart", no_attrs = TRUE),
+    plant.stand = imputate_na(Soybean, plant.stand, Class, method = "rpart", no_attrs = TRUE),
+    precip = imputate_na(Soybean, precip, Class, method = "rpart", no_attrs = TRUE),
+    temp = imputate_na(Soybean, temp, Class, method = "rpart", no_attrs = TRUE),
+    hail = imputate_na(Soybean, hail, Class, method = "rpart", no_attrs = TRUE),
+    crop.hist = imputate_na(Soybean, crop.hist, Class, method = "rpart", no_attrs = TRUE),
+    area.dam = imputate_na(Soybean, area.dam, Class, method = "rpart", no_attrs = TRUE),
+    sever = imputate_na(Soybean, sever, Class, method = "rpart", no_attrs = TRUE),
+    seed.tmt = imputate_na(Soybean, seed.tmt, Class, method = "rpart", no_attrs = TRUE),
+    germ = imputate_na(Soybean, germ, Class, method = "rpart", no_attrs = TRUE),
+    plant.growth = imputate_na(Soybean, plant.growth, Class, method = "rpart", no_attrs = TRUE),
+    leaf.halo = imputate_na(Soybean, leaf.halo, Class, method = "rpart", no_attrs = TRUE),
+    leaf.marg = imputate_na(Soybean, leaf.marg, Class, method = "rpart", no_attrs = TRUE),
+    leaf.size = imputate_na(Soybean, leaf.size, Class, method = "rpart", no_attrs = TRUE),
+    leaf.shread = imputate_na(Soybean, leaf.shread, Class, method = "rpart", no_attrs = TRUE),
+    leaf.malf = imputate_na(Soybean, leaf.malf, Class, method = "rpart", no_attrs = TRUE),
+    stem = imputate_na(Soybean, stem, Class, method = "rpart", no_attrs = TRUE),
+    lodging = imputate_na(Soybean, lodging, Class, method = "rpart", no_attrs = TRUE),
+    stem.cankers = imputate_na(Soybean, stem.cankers, Class, method = "rpart", no_attrs = TRUE),
+    canker.lesion = imputate_na(Soybean, canker.lesion, Class, method = "rpart", no_attrs = TRUE),
+    fruiting.bodies = imputate_na(Soybean, fruiting.bodies, Class, method = "rpart", no_attrs = TRUE),
+    ext.decay = imputate_na(Soybean, ext.decay, Class, method = "rpart", no_attrs = TRUE),
+    int.discolor = imputate_na(Soybean, int.discolor, Class, method = "rpart", no_attrs = TRUE),
+    fruit.pods = imputate_na(Soybean, fruit.pods, Class, method = "rpart", no_attrs = TRUE),
+    seed = imputate_na(Soybean, seed, Class, method = "rpart", no_attrs = TRUE),
+    mold.growth = imputate_na(Soybean, mold.growth, Class, method = "rpart", no_attrs = TRUE),
+    seed.discolor = imputate_na(Soybean, seed.discolor, Class, method = "rpart", no_attrs = TRUE),
+    seed.size = imputate_na(Soybean, seed.size, Class, method = "rpart", no_attrs = TRUE),
+    shriveling = imputate_na(Soybean, shriveling, Class, method = "rpart", no_attrs = TRUE),
+    fruit.spots = imputate_na(Soybean, fruit.spots, Class, method = "rpart", no_attrs = TRUE),
+    roots = imputate_na(Soybean, roots, Class, method = "rpart", no_attrs = TRUE)) %>%
+  # Drop the near zero variance predictors
+  select(-leaf.mild, -mycelium, -sclerotia) 
+```
+
+**To prove that it worked I present the following**
+
+```{r}
+Soybean_complete %>%
+  arrange(Class) %>%
+  missmap(main = "Missing vs Observed")
+```

### 2020-03-02 DATA612/data612.py
Updated for Project 2

@@ -152,8 +152,8 @@ def get_RMSE(user_item_df, predictor):
     Returns:
         RMSE_df (DataFrame): a data frame with the RMSE values.
     """
-    predictors = user_item_df.applymap(one_or_na) * predictor
-    errors = user_item_df - predictors
+    predictions = user_item_df.applymap(one_or_na) * predictor
+    errors = predictions - user_item_df
     squared_errors = errors ** 2
     mean_squared_errors = squared_errors.stack().mean()
     RMSE = mean_squared_errors ** (1 / 2)

### 2020-03-03 DATA624/HW5.Rmd
Initial commit

---
title: "DATA 624 Homework 5"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(fpp2)
```

## Question 7.1

Consider the `pigs` series — the number of pigs slaughtered in Victoria each month.

a. Use the ses() function in R to find the optimal values of $\alpha$ and $\iota_0$, and generate forecasts for the next four months.

b. Compute a 95% prediction interval for the first forecast using  $\hat{y} \pm$ 1.96 $s$ where $s$ is the standard deviation of the residuals.

c. Compare your interval with the interval produced by R.

## Question 7.5

Data set `books` contains the daily sales of paperback and hardcover books at the same store. The task is to forecast the next four days’ sales for paperback and hardcover books.

a. Plot the series and discuss the main features of the data.

b. Use the `ses()` function to forecast each series, and plot the forecasts.

c. Compute the RMSE values for the training data in each case.

## Question 7.6

We will continue with the daily sales of paperback and hardcover books in data set `books`.

a. Apply Holt’s linear method to the `paperback` and `hardback` series and compute four-day forecasts in each case.

b. Compare the RMSE measures of Holt’s method for the two series to those of simple exponential smoothing in the previous question. (Remember that Holt’s method is using one more parameter than SES.) Discuss the merits of the two forecasting methods for these data sets.

c. Compare the forecasts for the two series using both methods. Which do you think is best?

d. Calculate a 95% prediction interval for the first forecast for each series, using the RMSE values and assuming normal errors. Compare your intervals with those produced using `ses` and `holt`.

## Question 7.7

For this exercise use data set `eggs`, the price of a dozen eggs in the United States from 1900–1993. Experiment with the various options in the `holt()` function to see how much the forecasts change with damped trend, or with a Box-Cox transformation. Try to develop an intuition of what each argument is doing to the forecasts.

[Hint: use `h=100` when calling `holt()` so you can clearly see the differences between the various options when plotting the forecasts.]

Which model gives the best RMSE?

## Question 7.8

Recall your retail time series data (from Exercise 3 in Section 2.10).

a. Why is multiplicative seasonality necessary for this series?

b. Apply Holt-Winters’ multiplicative method to the data. Experiment with making the trend damped.

c. Compare the RMSE of the one-step forecasts from the two methods. Which do you prefer?

d. Check that the residuals from the best method look like white noise.

e. Now find the test set RMSE, while training the model to the end of 2010. Can you beat the seasonal naïve approach from Exercise 8 in Section 3.7?

## Question 7.9

For the same retail data, try an STL decomposition applied to the Box-Cox transformed series, followed by ETS on the seasonally adjusted data. How does that compare with your best previous forecasts on the test set?

### 2020-03-03 DATA624/Project 1-A ATM.Rmd
Initial commit

---
title: "DATA 624 Project 1 Part A - ATM"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
```

## Directions

In part A, I want you to forecast how much cash is taken out of 4 different ATM machines for May 2010.  The data is given in a single file.  The variable ‘Cash’ is provided in hundreds of dollars, other than that it is straight forward.   I am being somewhat ambiguous on purpose to make this have a little more business feeling.  Explain and demonstrate your process, techniques used and not used, and your actual forecast.  I am giving you data via an excel file, please provide your written report on your findings, visuals, discussion and your R code via an RPubs link along with the actual.rmd file  Also please submit the forecast which you will put in an Excel readable file.

### 2020-03-03 DATA624/Project 1-B Power.Rmd
Initial commit

---
title: "DATA 624 Project 1 Part B - Power"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
```

## Directions

Part B consists of a simple dataset of residential power usage for January 1998 until December 2013.  Your assignment is to model these data and a monthly forecast for 2014.  The data is given in a single file.  The variable ‘KWH’ is power consumption in Kilowatt hours, the rest is straight forward.    Add this to your existing files above.

### 2020-03-03 DATA624/Project 1-C Waterflow.Rmd
Initial commit

---
title: "DATA 624 Project 1 Part C - Waterflow"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
```

## Directions

Part C consists of two data sets.  These are simple 2 columns sets, however they have different time stamps.  Your optional assignment is to time-base sequence the data and aggregate based on hour (example of what this looks like, follows).  Note for multiple recordings within an hour, take the mean.  Then to determine if the data is stationary and can it be forecast.  If so, provide a week forward forecast and present results via Rpubs and .rmd and the forecast in an Excel readable file.

### 2020-03-05 DATA624/HW5.Rmd
More work on HW5

@@
 library(fpp2)
+options(scipen=999)
 ```
@@
 
+```{r}
+ses_pigs <- ses(pigs, h = 4)
+summary(ses_pigs)
+```
+
+**The `ses` function generates and $\alpha$ of `r ses_pigs$model$par[1]` and a $\iota_0$ of `r ses_pigs$model$par[2]`.**
+
 b. Compute a 95% prediction interval for the first forecast using  $\hat{y} \pm$ 1.96 $s$ where $s$ is the standard deviation of the residuals.
 
+```{r}
+# Get the first forecast
+y_hat <- ses_pigs$mean[1]
+# Get the standard deviation of the residuals
+s <- sd(ses_pigs$residuals)
+lower_ci <- y_hat - 1.96 * s
+upper_ci <- y_hat + 1.96 * s
+```
+
+**The 95% CI is from `r round(lower_ci, 0)` to `r round(upper_ci, 0)`,**
+
 c. Compare your interval with the interval produced by R.
+
+**The 95% CI produced by R is `r round(78611.97, 0)` to `r round(119020.8, 0 )`.R's CI is slightly wider than the CI we computed.**
 
@@
 
+```{r}
+autoplot(books) + 
+  ggtitle("Daily Sale of Books")
+```
+
+**The series has an upwards trend.
+There is only 30 days of data so I can't really speak to an seasonality or weekly effects with much confidence.
+The paperback sales lagging hardback, or hardback and paperback being counter cyclical, but I would stress it's really too soon to make these claims.**
+
 b. Use the `ses()` function to forecast each series, and plot the forecasts.
 
+```{r}
+ses_paperback <- ses(books[, 1], h = 4)
+autoplot(ses_paperback) +
+  ylab("Paperback Book Sales")
+```
+
+```{r}
+ses_hardback <- ses(books[, 2], h = 4)
+autoplot(ses_hardback) +
+  ylab("Hardback Book Sales")
+```
+
 c. Compute the RMSE values for the training data in each case.
+
+```{r}
+accuracy(ses_paperback)
+```
+
+```{r}
+accuracy(ses_hardback)
+```
+
+**The RMSE for the paperback model is 33.6 and 31.9 for the hardback model.**
 

### 2020-03-06 DATA624/HW5.Rmd
More work on HW5

@@
 
-```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
+```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
 knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
@@
 library(fpp2)
+library(kableExtra)
+library(gridExtra)
+library(readxl)
 options(scipen=999)
@@
 
-**The RMSE for the paperback model is 33.6 and 31.9 for the hardback model.**
+**The RMSE for the paperback model is `r round(accuracy(ses_paperback)[2], 2)` and `r round(accuracy(ses_hardback)[2], 2)` for the hardback model.**
 
@@
 
+```{r}
+holt_paperback <- holt(books[, 1], h = 4)
+summary(holt_paperback)
+```
+
+```{r}
+holt_hardback <- holt(books[, 2], h = 4)
+autoplot(holt_hardback) +
+  ylab("Hardback Book Sales")
+summary(holt_hardback)
+```
+
 b.
@@
 
+**Holt's method does a better job than the SES models, because the RMSE is smaller for both the paperback and hardback series.
+This is understandable because Holt's includes a trend component.
+SES assumes there is no trend.
+This does not really fit these timeseries.**
+
+```{r, echo=FALSE}
+data.frame(Type = c('Paperback', 'Hardback'),
+           SES = c(accuracy(ses_paperback)[2], accuracy(ses_hardback)[2]),
+           Holt = c(accuracy(holt_paperback)[2], accuracy(holt_hardback)[2])
+           ) %>%
+  rename(`Holt's Method` = Holt) %>%
+  kable() %>%
+  kable_styling()
+```
+
 c. Compare the forecasts for the two series using both methods. Which do you think is best?
 
+```{r}
+p1 <- autoplot(ses_paperback) + ylab("Paperback Book Sales") + ggtitle("SES Forecast")
+p2 <- autoplot(holt_paperback) + ylab("Paperback Book Sales") + ggtitle("Holt's Forecast")
+grid.arrange(p1, p2, ncol = 2)
+```
+
+```{r}
+p1 <- autoplot(ses_hardback) + ylab("Hardback Book Sales") + ggtitle("SES Forecast")
+p2 <- autoplot(holt_hardback) + ylab("Hardback Book Sales") + ggtitle("Holt's Forecast")
+grid.arrange(p1, p2, ncol = 2)
+```
+
+**I think Holt's method is a better forecast.  I like that factors in the trend.**
+
 d. Calculate a 95% prediction interval for the first forecast for each series, using the RMSE values and assuming normal errors. Compare your intervals with those produced using `ses` and `holt`.
+
+```{r}
+cat("Paperback:", holt_paperback$mean[1] - 1.96 * accuracy(holt_paperback)[2], "to", holt_paperback$mean[1] + 1.96 * accuracy(holt_paperback)[2])
+```
+```{r}
+cat("Hardback:", holt_hardback$mean[1] - 1.96 * accuracy(holt_hardback)[2], "to", holt_hardback$mean[1] + 1.96 * accuracy(holt_hardback)[2])
+```
+
+**The 95% CI for the paperback series generated by Holt's method is 143.9130  to 275.0205.
+This is wider than the CI computed above.
+The CI for hardbacks from Holt's method is 192.9222 to 307.4256, which again is wider than the CI above.**
 
@@
 
+```{r}
+h <- 100
+
+holt_eggs <- holt(eggs, h=h)
+box_cox_eggs <- holt(eggs, h=h, lambda=TRUE)
+damped_eggs <- holt(eggs, h=h, damped=TRUE)
+damped_box_cox_eggs <- holt(eggs, h=h, damped=TRUE, lambda=TRUE)
+exponential_eggs <- holt(eggs, h=h, exponential=TRUE)
+
+autoplot(holt_eggs) + ggtitle("Holt's Method") 
+autoplot(damped_eggs) + ggtitle("Damped")
+autoplot(box_cox_eggs) + ggtitle("Box-Cox")
+autoplot(damped_box_cox_eggs) + ggtitle("Damped & Box-Cox")
+autoplot(exponential_eggs) + ggtitle("Exponential")
+```
+
+```{r}
+get_rmse <- function(model){
+  accuracy(model)[2]
+}
+
+Model <- c("Holt's Linear", 
+           "Box-Cox Transformed",
+           "Damped",
+           "Damped and Box-Cox",
+           "Exponential")
+
+RMSE <- c(get_rmse(holt_eggs), 
+          get_rmse(box_cox_eggs),
+          get_rmse(damped_eggs),
+          get_rmse(damped_box_cox_eggs),
+          get_rmse(exponential_eggs))
+
+eggs_rmse_df <- data.frame(Model, RMSE)
+
+eggs_rmse_df %>%
+  kable() %>%
+  kable_styling()
+```
+
+**The model with the lowest (best) RMSE is the `r eggs_rmse_df[eggs_rmse_df$RMSE == min(eggs_rmse_df$RMSE),]$Model` model.**
+
 ## Question 7.8
@@
 b. Apply Holt-Winters’ multiplicative method to the data. Experiment with making the trend damped.
+
+```{r}
+retaildata <- read_excel("retail.xlsx", skip = 1)
+retail_ts <- ts(retaildata[, "A3349873A"], frequency = 12, start = c(1982, 4))
+
+hw_multiplicative <- hw(retail_ts, seasonal = "multiplicative")
+autoplot(hw_multiplicative) + 
+  ggtitle("Multiplicative") + 
+  ylab("Retail Sales")
+
+hw_multiplicative_damped <- hw(retail_ts, seasonal = "multiplicative", damped = TRUE)
+autoplot(hw_multiplicative_damped) + 
+  ggtitle("Multiplicative & Damped") + 
+  ylab("Retail Sales")
+```
 

### 2020-03-06 DATA624/HW5.Rmd
More work on HW5

@@
 
-b. Apply Holt-Winters’ multiplicative method to the data. Experiment with making the trend damped.
+**As we have seen in the previous sections, the variability in the series is increasing over time.
+This calls for a method that accounts for multiplicative seasonality.
+Here's the plot in case you have forgotten what the time series looks like:**
 
@@
 retail_ts <- ts(retaildata[, "A3349873A"], frequency = 12, start = c(1982, 4))
+autoplot(retail_ts) + ylab("Retail Sales")
+```
 
+b. Apply Holt-Winters’ multiplicative method to the data. Experiment with making the trend damped.
+
+```{r}
 hw_multiplicative <- hw(retail_ts, seasonal = "multiplicative")
@@
   ylab("Retail Sales")
+summary(hw_multiplicative)
 
@@
   ylab("Retail Sales")
+summary(hw_multiplicative_damped)
 ```
@@
 
+```{r}
+cat("RMSE of Multiplicative = ", accuracy(hw_multiplicative)[2])
+```
+
+```{r}
+cat("RMSE of Multiplicative & Damped = ", accuracy(hw_multiplicative_damped)[2])
+```
+
+**The non-damped model is preforming better.**
+
 d. Check that the residuals from the best method look like white noise.
 
+```{r}
+checkresiduals(hw_multiplicative)
+```
+
+**The residuals looks like white noise to me.  They are normally distributed with a mean of zero, with no pattern over time.**
+
 e. Now find the test set RMSE, while training the model to the end of 2010. Can you beat the seasonal naïve approach from Exercise 8 in Section 3.7?
+
+```{r}
+retail_train <- window(retail_ts, end = c(2010, 12))
+retail_test <- window(retail_ts, start = 2011)
+
+get_retail_test_rmse <- function(model){
+  accuracy(model, retail_test)[4]
+}
+
+Model <- c("Seasonal Naïve (Baseline)", 
+           "SES",
+           "Holt's Method",
+           "Damped Holt's Method",
+           "Holt-Winters Additive",
+           "Holt-Winters Multiplicative",
+           "Damped Holt-Winters Additive",
+           "Damped Holt-Winters Multiplicative")
+
+RMSE <- c(get_retail_test_rmse(snaive(retail_train)), 
+          get_retail_test_rmse(ses(retail_train)),
+          get_retail_test_rmse(holt(retail_train)),
+          get_retail_test_rmse(holt(retail_train, damped = TRUE)),
+          get_retail_test_rmse(hw(retail_train, seasonal = "additive")),
+          get_retail_test_rmse(hw(retail_train, seasonal = "multiplicative")),
+          get_retail_test_rmse(hw(retail_train, seasonal = "additive", damped = TRUE)),
+          get_retail_test_rmse(hw(retail_train, seasonal = "multiplicative", damped = TRUE)))
+
+rmse_df <- data.frame(Model, RMSE)
+
+rmse_df %>%
+  kable() %>%
+  kable_styling()
+```
+
+**This was and interesting exercise.
+Some of the more sophisticated approaches preformed less well on the test set than the seasonal naïve model.
+In the end the `r rmse_df[rmse_df$RMSE == min(rmse_df$RMSE),]$Model` preformed the best on the test set.**
 

### 2020-03-07 DATA624/HW5.Rmd
Additional work on HW5

@@
 **This was and interesting exercise.
+I decided to run it through a buch of models, even some that didn't make sense given this data set.
 Some of the more sophisticated approaches preformed less well on the test set than the seasonal naïve model.
@@
 How does that compare with your best previous forecasts on the test set?
+
+```{r}
+# Window doesn't return a ts object so we need to redefine
+retail_train <- ts(as.vector(retail_ts), start=c(1982,4), end=c(2010,12), frequency = 12)
+lambda <- BoxCox.lambda(retail_train)
+bc_retail_train <- BoxCox(retail_train, lambda = lambda)
+stl_retail_train <- stl(bc_retail_train, s.window="periodic") # Guessing on s.window setting based on documentation
+stl_time_series <- data.frame(stl_retail_train$time.series)
+sa_retail_train <- bc_retail_train - stl_time_series$seasonal
+
+ets_retail_train <- forecast(sa_retail_train)
+bc_retail_test <- BoxCox(retail_test, lambda = lambda)
+accuracy(ets_retail_train, bc_retail_test)
+```

### 2020-03-07 DATA624/Project 1-A ATM.Rmd
More work on Project 1-A

@@
     toc_depth: 3
-    code_folding: "show"
+    code_folding: "hide"
 ---
@@
 library(kableExtra)
+library(readxl)
+library(fpp2)
 ```
 
-## Directions
+## Introduction
 
-In part A, I want you to forecast how much cash is taken out of 4 different ATM machines for May 2010.
-The data is given in a single file.
-The variable ‘Cash’ is provided in hundreds of dollars, other than that it is straight forward.
-I am being somewhat ambiguous on purpose to make this have a little more business feeling.
-Explain and demonstrate your process, techniques used and not used, and your actual forecast.
-I am giving you data via an excel file, please provide your written report on your findings, visuals, discussion and your R code via an RPubs link along with the actual.rmd file  Also please submit the forecast which you will put in an Excel readable file.
+I am to forecast how much cash will be taken out of 4 different ATM machines for May 2010.  Data was provided for this project.  The cash is in hundreds of dollars.
+
+## Data Exploration
+
+I will begin by plotting the cash withdrawn by each ATM.
+
+```{r}
+atm <- read_excel("ATM624Data.xlsx") %>%
+  mutate(DATE = as.Date(DATE, origin = "1899-12-30"),
+         ATM = as.factor(ATM))
+
+atm %>%
+  ggplot(aes(DATE, Cash, color=ATM)) +
+  geom_line(lwd=1) +
+  ggtitle("Cash Withdrawal by Date and ATM") +
+  scale_color_brewer(palette = "Set1") + 
+  theme(legend.title = element_blank(),
+        legend.position = "bottom",
+        axis.title = element_blank())
+```
+
+ATM #4 is utilized at a higher rate than every other ATM in the data series.
+There is an unusual spike in ATM #4.
+According to a [2010 article in Time magazine](https://business.time.com/2010/10/07/more-thevies-are-making-total-atm-withdrawls/) an average sized ATM can hold $200,000.
+So the outlier is an implausible figure.
+There are some missing ATM labels in the data.
+Let's remove ATM #4 and the missing label from the dataset and replot the visualization:
+
+```{r}
+atm %>%
+  filter(ATM != "ATM4") %>%
+  filter(!is.na(ATM)) %>%
+  ggplot(aes(DATE, Cash, color=ATM)) +
+  geom_line(lwd=1) +
+  ggtitle("Cash Withdrawal by Date and ATM") +
+  scale_color_brewer(palette = "Set1") + 
+  theme(legend.title = element_blank(),
+        legend.position = "bottom",
+        axis.title = element_blank())
+```
+
+It looks like ATM #3 was not utilized until recently.
+ATM #1 and #2 seem very similar to each other.
+It doesn't seem like there is much of a trend in the data.
+Let's explore the distributions further with this visualization:
+
+```{r}
+atm %>%
+  na.omit() %>%
+  ggplot(aes(ATM, Cash, color = ATM)) +
+  geom_boxplot() +
+  ggtitle("Cash Withdrawal by ATM") +
+  scale_color_brewer(palette = "Set1") +
+  facet_wrap(.~ATM, nrow = 2, scales = "free") +
+  theme(legend.position = "none",
+        axis.title.y = element_blank(),
+        axis.title.x = element_blank(),
+        axis.text.x = element_blank(),
+        axis.ticks.x = element_blank())
+
+atm %>%
+  filter(!is.na(ATM)) %>%
+  group_by(ATM) %>%
+  summarise(Minimum = min(Cash, na.rm = T),
+            `1st Qu.` = quantile(Cash, .25, na.rm = T),
+            Mean = mean(Cash, na.rm = T),
+            Median = median(Cash, na.rm = T),
+            `3rd Qu.` = quantile(Cash, .75, na.rm = T),
+            Maximum = max(Cash, na.rm = T),
+            `NA's` = sum(is.na(Cash))) %>%
+  kable() %>%
+  kable_styling()
+```
+
+### Dealing with Missing Data & Outliers
+
+ATM #1 is missing 3 observations and ATM #2 is missing 2.  I will replace the NA's with the median values.  We will also replace all the outliers in ATM #4 with the $200,000.
+
+```{r}
+atm <- atm %>% 
+  rowwise() %>%
+  mutate(Cash = if_else(is.na(Cash) & ATM == "ATM1", 91, Cash)) %>%
+  mutate(Cash = if_else(is.na(Cash) & ATM == "ATM2", 67, Cash)) %>%
+  mutate(Cash = if_else(Cash > 2000, 2000, Cash))
+```

### 2020-03-07 DATA624/Project 1-A ATM.Rmd
Minor change

@@
 library(fpp2)
+library(prophet)
 ```
@@
 ```{r}
-atm <- atm %>% 
+adjusted_atm <- atm %>% 
   rowwise() %>%
@@
 ```
+
+## Model Building
+
+I will construct a series of models using different approaches.  I will evaluate each model's preformance on a test set.

### 2020-03-07 DATA624/HW5.Rmd
Finalized HW5

--- 2020-03-06 DATA624/HW5.Rmd
@@
 How does that compare with your best previous forecasts on the test set?
+
+```{r}
+# Window doesn't return a ts object so we need to redefine
+retail_train <- ts(as.vector(retail_ts), start=c(1982,4), end=c(2010,12), frequency = 12)
+# Get the optimal lambda
+lambda <- BoxCox.lambda(retail_train)
+# Preform a Box-Cox transformation
+bc_retail_train <- BoxCox(retail_train, lambda = lambda)
+# Preform the STL decomposition
+stl_retail_train <- mstl(bc_retail_train)
+## Create a seasonally adjusted series
+sa_stl_retail_train <- stl_retail_train[,1] - stl_retail_train[,3]
+# Reverse the Box-Cox transformation
+sa_retail_train <- InvBoxCox(sa_stl_retail_train, lambda = lambda)
+# Check the output
+autoplot(retail_train) + autolayer(sa_retail_train)
+```
+```{r}
+ets_retail_train <- forecast(sa_retail_train)
+
+rmse_df %>%
+  mutate(Model = as.character(Model)) %>%
+  rbind(c("ETS on STL Seasonally-Adjusted data", get_retail_test_rmse(ets_retail_train))) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**The ETS model on the transformed data preformed worse than the Seasonal Naïve model.  Again a simpler approach can be a better one.**

### 2020-03-07 DATA624/Project 1-A ATM.Rmd
More work on Project 1

--- 2020-03-07 DATA624/Project 1-A ATM.Rmd
@@
 library(tidyverse)
+library(lubridate)
 library(kableExtra)
@@
 library(fpp2)
+library(prophet)
+library(forcats)
 ```
@@
   mutate(DATE = as.Date(DATE, origin = "1899-12-30"),
-         ATM = as.factor(ATM))
+         ATM = as.factor(ATM)) %>%
+  data.frame()
 
@@
   mutate(Cash = if_else(is.na(Cash) & ATM == "ATM2", 67, Cash)) %>%
-  mutate(Cash = if_else(Cash > 2000, 2000, Cash))
+  mutate(Cash = if_else(Cash > 2000, 2000, Cash)) %>%
+  ungroup() %>%
+  data.frame()
 ```
+
+## Model Building
+
+I will construct a series of models using different approaches.
+I will start from the most simple approach and move on to more sophisticated ones.
+I beleive the "best" model will come from the middle of the pack.
+Let's see how it plays out.
+
+Since the task is to predict the cash withdrawls for May 2010 (31 days) we will set up test sets that include the last 31 days.
+Each model will be evaluated on how accurate they are in forecasting the test set's values.
+I will use the RMSE as the metric to assess the accuracy.
+
+```{r}
+train_test_split <- function(atm_df, atm_number, n_in_test = 31, return_ts=TRUE){
+  ### Splits the data into training and test sets
+  # Args:
+  #     atm_df (data.frame): The ATM data
+  #     atm_number (int): Which ATM do you want the time series for?
+  #     n_in_test (int): The number of observations in the test set taken from the end (optional-31 default)
+  #     return_ts (logical): Should the function return ts objects (optional-TRUE default)
+  # Returns:
+  #     train_test_ts (list): Two ts objects (train and test).
+
+  # Subset the data frame
+  df <- atm_df %>%
+    filter(ATM == paste0("ATM", atm_number)) %>%
+    select(-ATM)
+
+  # Split the data frame into training and test sets
+  test <- tail(df, n_in_test)
+  train <- head(df, -n_in_test)
+
+  # Tranform into ts objects
+  if(return_ts){
+    train_start <- (min(train$DATE))
+    train <- ts(train$Cash, start = decimal_date(as.Date(train_start)), frequency = 365)
+    test_start <- (min(test$DATE))
+    test <- ts(test$Cash, start = decimal_date(as.Date(test_start)), frequency = 365)
+  }
+
+  # Return the data to the user
+  train_test <- list(train = train, test = test)
+  return(train_test)
+}
+
+atm_1_ts <- train_test_split(adjusted_atm, 1, 31)
+atm_2_ts <- train_test_split(adjusted_atm, 2, 31)
+atm_3_ts <- train_test_split(adjusted_atm, 3, 31)
+atm_4_ts <- train_test_split(adjusted_atm, 4, 31)
+
+atm_1_df <- train_test_split(adjusted_atm, 1, 31, return_ts = FALSE)
+atm_2_df <- train_test_split(adjusted_atm, 2, 31, return_ts = FALSE)
+atm_3_df <- train_test_split(adjusted_atm, 3, 31, return_ts = FALSE)
+atm_4_df <- train_test_split(adjusted_atm, 4, 31, return_ts = FALSE)
+```
+
+### ATM 1 Models
+
+I will start by looking at the training data and the decomposition.
+I cannot use the `decompose()` or `stl()` functions to decompose it because I don't have enough data.
+Facebook Prophet can decompose it using its Basian process.
+
+```{r}
+atm_prophet <- function(df, forcast_window){
+  m <- df %>%
+    mutate(y = Cash, ds = as.character(DATE)) %>%
+    prophet()
+  future <- make_future_dataframe(m, periods = forcast_window)
+  atm_forecast <- predict(m, future)
+  results = list(
+    model = m,
+    future = future,
+    forecast = atm_forecast
+  )
+  return(results)
+}
+
+atm_1_prophet <- atm_prophet(atm_1_df$train, 31)
+prophet_plot_components(atm_1_prophet$model, atm_1_prophet$forecast)
+```
+
+That is interesting! There appears to be a weekly effect in the data.  Wednesdays and especially Thursdays are low traffic days.  The trend apears to be decreasing but at a very slow rate.

### 2020-03-13 DATA624/Project 1-A ATM.Rmd
More work on Project 1

@@
 ```{r}
-train_test_split <- function(atm_df, atm_number, n_in_test = 31, return_ts=TRUE){
+
+forecast_window <- 31
+
+df_to_ts <- function(df){
+  ### Converts ATM data frames into ts objects
+  # Args:
+  #     df (data.frame): The ATM data frame 
+  # Returns:
+  #     ts (Time-Series): An ATM time series object
+  ts(df$Cash, start = decimal_date(as.Date(min(df$DATE))), frequency = 365)
+}
+
+train_test_split <- function(atm_df, atm_number, n_in_test, return_ts=TRUE){
   ### Splits the data into training and test sets
@@
   # Returns:
-  #     train_test_ts (list): Two ts objects (train and test).
+  #     train_eval_test_ts (list): Data split into sets (train, eval and test).
 
@@
 
-  # Split the data frame into training and test sets
+  # Split the data frame into training, eval and test sets
   test <- tail(df, n_in_test)
   train <- head(df, -n_in_test)
+  eval_ts <- tail(train, n_in_test)
+  short_train <- head(df, -(2 * n_in_test))
 
@@
   if(return_ts){
-    train_start <- (min(train$DATE))
-    train <- ts(train$Cash, start = decimal_date(as.Date(train_start)), frequency = 365)
-    test_start <- (min(test$DATE))
-    test <- ts(test$Cash, start = decimal_date(as.Date(test_start)), frequency = 365)
+    train <- df_to_ts(train)
+    test <- df_to_ts(test)
+    eval_ts <- df_to_ts(eval_ts)
+    short_train <- df_to_ts(short_train)
   }
@@
   # Return the data to the user
-  train_test <- list(train = train, test = test)
-  return(train_test)
+  train_eval_test_ts <- list(train = train, test = test, evaluation = eval_ts, short_train = short_train)
+
+  return(train_eval_test_ts)
 }
 
-atm_1_ts <- train_test_split(adjusted_atm, 1, 31)
-atm_2_ts <- train_test_split(adjusted_atm, 2, 31)
-atm_3_ts <- train_test_split(adjusted_atm, 3, 31)
-atm_4_ts <- train_test_split(adjusted_atm, 4, 31)
-
-atm_1_df <- train_test_split(adjusted_atm, 1, 31, return_ts = FALSE)
-atm_2_df <- train_test_split(adjusted_atm, 2, 31, return_ts = FALSE)
-atm_3_df <- train_test_split(adjusted_atm, 3, 31, return_ts = FALSE)
-atm_4_df <- train_test_split(adjusted_atm, 4, 31, return_ts = FALSE)
+# Get time series objects
+atm_1_ts <- train_test_split(adjusted_atm, 1, forecast_window)
+atm_2_ts <- train_test_split(adjusted_atm, 2, forecast_window)
+atm_3_ts <- train_test_split(adjusted_atm, 3, forecast_window)
+atm_4_ts <- train_test_split(adjusted_atm, 4, forecast_window)
+# Get data frame objects
+atm_1_df <- train_test_split(adjusted_atm, 1, forecast_window, return_ts = FALSE)
+atm_2_df <- train_test_split(adjusted_atm, 2, forecast_window, return_ts = FALSE)
+atm_3_df <- train_test_split(adjusted_atm, 3, forecast_window, return_ts = FALSE)
+atm_4_df <- train_test_split(adjusted_atm, 4, forecast_window, return_ts = FALSE)
 ```
@@
 I cannot use the `decompose()` or `stl()` functions to decompose it because I don't have enough data.
-Facebook Prophet can decompose it using its Basian process.
+Facebook Prophet can decompose it using its Bayesian estimation process.
 
@@
 That is interesting! There appears to be a weekly effect in the data.  Wednesdays and especially Thursdays are low traffic days.  The trend apears to be decreasing but at a very slow rate.
+
+```{r}
+
+atm_1_df$train %>%
+  mutate(`Day of Week` = recode(as.factor(wday(DATE)), "1" = "Sunday", "2" = "Monday", "3" = "Tuesday", "4"="Wednesday", "5"="Thursday", "6"="Friday","7"="Saturday")) %>%
+  ggplot(aes(x=DATE, y=Cash, color=`Day of Week`)) +
+  geom_point() +
+  #geom_line() +
+  facet_wrap(.~`Day of Week`) +
+  theme(legend.position = "None", axis.title = element_blank())
+```
+
+Saturday looks fairly consistent.  Other days (Tuesday and Thursday) show recent changes in their trend.
+
+```{r}
+atm_1_holt <- holt(atm_1_ts$train, h=forecast_window)
+autoplot(atm_1_holt) + autolayer(atm_1_ts$test)
+```
+
+```{r}
+plot(atm_1_prophet$model, atm_1_prophet$forecast)
+```
+
+```{r}
+atm_1_mean <- mean(atm_1_df$train)
+atm_1_last_observation <- tail(atm_1_df$train, 1)$Cash
+atm_1_mean <- mean(atm_1_df$train$Cash)
+atm_1_results <- 
+  atm_1_prophet$forecast %>%
+  mutate(DATE = as.Date(ds), prophet_yhat = yhat) %>%
+  select(DATE, prophet_yhat) %>%
+  mutate(naive_yhat = atm_1_last_observation, 
+         mean_yhat = atm_1_mean) %>%
+  merge(atm_1_df$evaluation) %>%
+  rename(actual = Cash)
+atm_1_results$holt <- tail(atm_1_holt$fitted, forecast_window)
+```

### 2020-03-14 DATA624/HW6.Rmd
Initial commit of HW6

---
title: "DATA 624 Homework 6"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(fpp2)
library(kableExtra)
library(gridExtra)
library(readxl)
options(scipen=999)
```

## Question 8.1

Figure 8.31 shows the ACFs for 36 random numbers, 360 random numbers and 1,000 random numbers.

a. Explain the differences among these figures. Do they all indicate that the data are white noise?

<img src ="https://otexts.com/fpp2/fpp_files/figure-html/wnacfplus-1.png" alt="Left: ACF for a white noise series of 36 numbers. Middle: ACF for a white noise series of 360 numbers. Right: ACF for a white noise series of 1,000 numbers.">

Figure 8.31: Left: ACF for a white noise series of 36 numbers. Middle: ACF for a white noise series of 360 numbers. Right: ACF for a white noise series of 1,000 numbers.

b. Why are the critical values at different distances from the mean of zero? Why are the autocorrelations different in each figure when they each refer to white noise?

## Question 8.2

A classic example of a non-stationary series is the daily closing IBM stock price series (data set `ibmclose`). Use R to plot the daily closing prices for IBM stock and the ACF and PACF. Explain how each plot shows that the series is non-stationary and should be differenced.

## Question 8.3

For the following series, find an appropriate Box-Cox transformation and order of differencing in order to obtain stationary data.

a. `usnetelec`

b. `usgdp`

c. `mcopper`

d. `enplanements`

e. `visitors`

## Question 8.5

For your retail data (from Exercise 3 in Section 2.10), find the appropriate order of differencing (after transformation if necessary) to obtain stationary data.

## Question 8.6

Use R to simulate and plot some data from simple ARIMA models.

a. Use the following R code to generate data from an AR(1) model with $ϕ_1 = 0.6$ and $σ_2 = 1$.  The process starts with $y_1=0$.

```{r}
y <- ts(numeric(100))
e <- rnorm(100)
for(i in 2:100)
  y[i] <- 0.6*y[i-1] + e[i]
```

b. Produce a time plot for the series. How does the plot change as you change $ϕ_1$?

c. Write your own code to generate data from an MA(1) model with $θ_1=0.6$ and $σ_2=1$.

d. Produce a time plot for the series. How does the plot change as you change $θ_1$?

e. Generate data from an ARMA(1,1) model with $ϕ_1=0.6$, $θ_1=0.6$ and $σ_2=1$.

f. Generate data from an AR(2) model with $ϕ_1=−0.8$, $ϕ_2=0.3$ and $σ_2=1$. (Note that these parameters will give a non-stationary series.)

g. Graph the latter two series and compare them.

## Question 8.7

Consider `wmurders`, the number of women murdered each year (per 100,000 standard population) in the United States.

a. By studying appropriate graphs of the series in R, find an appropriate ARIMA(p,d,q) model for these data.

b. Should you include a constant in the model? Explain.

c. Write this model in terms of the backshift operator.

d. Fit the model using R and examine the residuals. Is the model satisfactory?

e. Forecast three times ahead. Check your forecasts by hand to make sure that you know how they have been calculated.

f. Create a plot of the series with forecasts and prediction intervals for the next three periods shown.

g. Does auto.arima() give the same model you have chosen? If not, which model do you think is better?

### 2020-03-14 DATA624/HW6.Rmd
More work on HW6

@@
 library(readxl)
+library(urca)
 options(scipen=999)
@@
 
-a. Explain the differences among these figures. Do they all indicate that the data are white noise?
-
 <img src ="https://otexts.com/fpp2/fpp_files/figure-html/wnacfplus-1.png" alt="Left: ACF for a white noise series of 36 numbers.
@@
 
-Figure 8.31: Left: ACF for a white noise series of 36 numbers. Middle: ACF for a white noise series of 360 numbers. Right: ACF for a white noise series of 1,000 numbers.
+a. Explain the differences among these figures. Do they all indicate that the data are white noise?
+
+**The ACF bands become narrower as the number of numbers increase.
+There is no discernable pattern in the bars of the ACF charts and the bars are within the bands, which indicates the data are white noise.**
 
 b. Why are the critical values at different distances from the mean of zero? Why are the autocorrelations different in each figure when they each refer to white noise?
+
+**It's related to the law of large numbers.
+As the number of observations increase, the number of large outliers from the mean decreases.
+We are more certain that a large observation is really an outlier with more data.**
 
@@
 
+```{r}
+ggtsplot <- function(ts, title){
+  # A ggplot2 version of tsdisplay(df)
+  # Args:
+  #    ts (Time-Series): The time series we want to plot
+  #    title (str): The title of the graph
+  grid.arrange( 
+    autoplot(ts) +
+      ggtitle(title) +
+      theme(axis.title = element_blank()),
+    grid.arrange(
+      ggAcf(ts) + 
+        ggtitle(element_blank()), 
+      ggPacf(ts) + 
+        ggtitle(element_blank()), ncol = 2), nrow = 2)
+}
+
+ggtsplot(ibmclose, "IBM Closing Stock Price")
+```
+
+**Time series with trends or with seasonality are not stationary.
+It is clear from the top plot that there is a trend in IBM's stock price.
+The ACF will drop to zero quickly for a stationary time series.
+This ACF does not have that property.
+The gradual decrease is a tell that it is not stationary.
+The first PACF is almost one which is another sign that the data is not stationary.**
+
 ## Question 8.3
@@
 
+```{r}
+ggtsplot(diff(usnetelec, 1), "Differenced US Net Electricity Generation (lag=1)")
+```
+
+```{r}
+#ur.kpss(diff(usnetelec, 1)) 
+#Box.test(diff(usnetelec, 1), lag = 1)
+usnetelec %>% diff() %>% ur.kpss() %>% summary()
+```
+
 b. `usgdp`
+
+```{r}
+ggtsdisplay(diff(usgdp), "US GDP")
+```
 

### 2020-03-14 DATA624/Project 1-A ATM.Rmd
More work on Project 1

@@
 library(forcats)
+options(scipen=999)
 ```
@@
   scale_color_brewer(palette = "Set1") +
+  facet_wrap(.~ATM, ncol = 2) +
   theme(legend.title = element_blank(),
@@
   geom_line(lwd=1) +
+  facet_wrap(.~ATM, ncol = 2) +
   ggtitle("Cash Withdrawal by Date and ATM") +
@@
   #     ts (Time-Series): An ATM time series object
-  ts(df$Cash, start = decimal_date(as.Date(min(df$DATE))), frequency = 365)
+  ts(df$Cash, start = decimal_date(as.Date(min(df$DATE))), frequency = 7)
+  #msts(df$Cash, seasonal.periods=c(7, 365.25))
 }
@@
   # Split the data frame into training, eval and test sets
+  full <- df
   test <- tail(df, n_in_test)
@@
   if(return_ts){
+    full <- df_to_ts(full)
     train <- df_to_ts(train)
@@
   # Return the data to the user
-  train_eval_test_ts <- list(train = train, test = test, evaluation = eval_ts, short_train = short_train)
+  train_eval_test_ts <- list(train = train, test = test, evaluation = eval_ts, short_train = short_train, full = full)
 
@@
 ```{r}
+autoplot(hw(atm_1_ts$train, h=forecast_window))
+```
+
+```{r}
 atm_1_holt <- holt(atm_1_ts$train, h=forecast_window)
@@
 ```
+```{r}
+day_of_week_averages <- atm_1_df$train %>%
+  mutate(day_of_week = wday(DATE)) %>%
+  group_by(day_of_week) %>%
+  summarise(yhat = median(Cash))
+temp <- atm_1_df$test %>%
+  mutate(day_of_week = wday(DATE)) %>%
+  merge(day_of_week_averages) %>%
+  arrange(DATE) %>%
+  select(DATE, yhat)
+atm_1_results$day_of_week_model <- temp$yhat
+```
+
+### Random Forest
+
+```{r}
+library(randomForest)
+set.seed(42)
+rf_data <- function(df){
+  df %>%
+    mutate(day_of_week = as.factor(wday(DATE)),
+           DATE = as.numeric(DATE))
+}
+
+rf_fit <- randomForest(Cash~., rf_data(atm_1_df$train))
+predict(rf_fit, rf_data(atm_1_df$test))
+
+atm_1_results$rf <- predict(rf_fit, rf_data(atm_1_df$test))
+```
+
+```{r}
+lm_fit <- lm(Cash~day_of_week, rf_data(atm_1_df$train))
+atm_1_results$lm <- predict(lm_fit, rf_data(atm_1_df$test))
+```

### 2020-03-17 DATA624/HW6.Rmd
More work on HW 6

@@
 ```{r}
-ggtsplot(diff(usnetelec, 1), "Differenced US Net Electricity Generation (lag=1)")
+autoplot(usnetelec, main = "US Net Electricity Generation") +
+  theme(axis.title = element_blank())
 ```
 
+**The trend looks nearly linear.  It will probably only need to be differenced to be stationary.  I will use the `ndiffs` function to determine the order of differencing.***
+
 ```{r}
-#ur.kpss(diff(usnetelec, 1)) 
-#Box.test(diff(usnetelec, 1), lag = 1)
+ggtsplot(diff(usnetelec, ndiffs(usnetelec)), paste0("Differenced US Net Electricity Generation (lag=", ndiffs(usnetelec), ")"))
+```
+
+**It needed to be differenced by `r ndiffs(usnetelec)`.  This is further confirmed using the KPSS Unit Root test below:**
+
+```{r}
 usnetelec %>% diff() %>% ur.kpss() %>% summary()
@@
 ```{r}
-ggtsdisplay(diff(usgdp), "US GDP")
+autoplot(usgdp, main = "Quarterly US Gross Domestic Product") +
+  theme(axis.title = element_blank())
+```
+
+**Since this series is dollars we probably need to log transform it.  It also looks like it needs to be differenced.**
+
+```{r}
+usgdp_lambda <- BoxCox.lambda(usgdp)
+bc_usgdp <- BoxCox(usgdp, lambda = usgdp_lambda)
+ggtsplot(diff(bc_usgdp, ndiffs(bc_usgdp)), paste0("Differenced Quarterly US GDP (lag=", ndiffs(bc_usgdp), ", lambda=",round(usgdp_lambda,2),")"))
+```
+
+```{r}
+bc_usgdp %>% diff() %>% ur.kpss() %>% summary()
 ```
@@
 
+```{r}
+autoplot(mcopper, main = "Monthly Grade A Copper Prices") +
+  theme(axis.title = element_blank())
+```
+
+```{r}
+mcopper_lambda <- BoxCox.lambda(mcopper)
+bc_mcopper <- BoxCox(mcopper, lambda = mcopper_lambda)
+ggtsplot(diff(bc_mcopper, ndiffs(bc_mcopper)), paste0("Differenced Monthly Grade A Copper Prices (lag=", ndiffs(bc_mcopper), ", lambda=",round(mcopper_lambda,2),")"))
+```
+
 d. `enplanements`
 
+```{r}
+autoplot(enplanements, main = "Monthly US Domestic Enplanements") +
+  theme(axis.title = element_blank())
+```
+
+```{r}
+enplanements_lambda <- BoxCox.lambda(enplanements)
+bc_enplanements <- BoxCox(enplanements, lambda = enplanements_lambda)
+ggtsplot(diff(bc_enplanements, ndiffs(bc_enplanements)), paste0("Differenced Monthly US Domestic Enplanements (lag=", ndiffs(bc_enplanements), ", lambda=",round(enplanements_lambda,2),")"))
+```
+
 e. `visitors`
+
+```{r}
+autoplot(visitors, main = "Monthly Australian Overseas Visitors") +
+  theme(axis.title = element_blank())
+```
+
+```{r}
+visitors_lambda <- BoxCox.lambda(visitors)
+bc_visitors <- BoxCox(visitors, lambda = visitors_lambda)
+ggtsplot(diff(bc_visitors, ndiffs(bc_visitors)), paste0("Differenced Monthly Australian Overseas Visitors (lag=", ndiffs(bc_visitors), ", lambda=",round(visitors_lambda,2),")"))
+```
 

### 2020-03-18 DATA624/HW6.Rmd
More work on HW 6

@@
 library(readxl)
-library(urca)
+library(tseries)
 options(scipen=999)
@@
 
-**It needed to be differenced by `r ndiffs(usnetelec)`.  This is further confirmed using the KPSS Unit Root test below:**
+**It needed to be differenced by `r ndiffs(usnetelec)`, or a first order difference.  Now to check if the data is stationary.**
 
 ```{r}
-usnetelec %>% diff() %>% ur.kpss() %>% summary()
+is_stationary <- function(data){
+  results <- kpss.test(data)
+  if (results$p.value > 0.05){
+    "Data IS Stationary"
+  } else {
+    "Data IS NOT Stationary"
+  }
+}
+
+cat(is_stationary(diff(usnetelec, ndiffs(usnetelec))))
 ```
+
+**The p-value of 0.36 indicates the data are stationary.**
 
@@
 
+**The Box Cox transformation with a lambda of `r round(usgdp_lambda,2)` and a differencing of `r ndiffs(bc_usgdp)`.**
+
 ```{r}
-bc_usgdp %>% diff() %>% ur.kpss() %>% summary()
+cat(is_stationary(diff(bc_usgdp, ndiffs(bc_usgdp))))
 ```
@@
 
+**This has the spike in copper prices that happened late in the the 2000's.**
+
 ```{r}
@@
 ggtsplot(diff(bc_mcopper, ndiffs(bc_mcopper)), paste0("Differenced Monthly Grade A Copper Prices (lag=", ndiffs(bc_mcopper), ", lambda=",round(mcopper_lambda,2),")"))
+```
+
+**The Box Cox transformation with a lambda of `r round(mcopper_lambda,2)` and a differencing of `r ndiffs(bc_mcopper)`.**
+
+```{r}
+cat(is_stationary(diff(bc_mcopper, ndiffs(bc_mcopper))))
 ```
@@
 
+**The variance is not constant so this series needs a Box-Cox transformation and differencing to be stationary.**
+
 ```{r}
@@
 ggtsplot(diff(bc_enplanements, ndiffs(bc_enplanements)), paste0("Differenced Monthly US Domestic Enplanements (lag=", ndiffs(bc_enplanements), ", lambda=",round(enplanements_lambda,2),")"))
+```
+
+**The Box Cox transformation with a lambda of `r round(enplanements_lambda,2)` and a differencing of `r ndiffs(bc_enplanements)` (1st order).**
+
+```{r}
+cat(is_stationary(diff(bc_enplanements, ndiffs(bc_enplanements))))
 ```
@@
 
+**This series also exhibits an increasing variance and trend so Box-Cox and differencing will be needed.**
+
 ```{r}
@@
 
+**The Box Cox transformation had a lambda of `r round(visitors_lambda,2)` and a differencing of `r ndiffs(bc_visitors)` (1st order).**
+
+```{r}
+cat(is_stationary(diff(bc_visitors, ndiffs(bc_visitors))))
+```
+
 ## Question 8.5
@@
 For your retail data (from Exercise 3 in Section 2.10), find the appropriate order of differencing (after transformation if necessary) to obtain stationary data.
+
+```{r}
+retaildata <- read_excel("retail.xlsx", skip = 1)
+retail <- ts(retaildata[, "A3349873A"], frequency = 12, start = c(1982, 4))
+retail_lambda <- BoxCox.lambda(retail)
+bc_retail <- BoxCox(retail, lambda = retail_lambda)
+ggtsplot(diff(bc_retail, ndiffs(bc_retail)), paste0("Differenced Monthly Retail Sales (lag=", ndiffs(bc_retail), ", lambda=",round(retail_lambda,2),")"))
+```
+
+**The data is differenced with a first order difference and adjusted with a Box Cox transformation with a lambda of `r round(retail_lambda,2)`.  The result:**
+
+```{r}
+cat(is_stationary(diff(bc_retail, ndiffs(bc_retail))))
+```
 

### 2020-03-19 DATA624/HW6.Rmd
More work on HW6

@@
 
+**In order to see the effects of a change in $ϕ_1$ we will need to modify the code a little bit:**
+
+```{r}
+sim_AR1 <- function(phi){
+  set.seed(42)
+  y <- ts(numeric(100))
+  e <- rnorm(100)
+  for(i in 2:100){
+    y[i] <- phi*y[i-1] + e[i]
+  }
+  return(y)
+}
+
+p <- autoplot(sim_AR1(0.6))
+for(phi in seq(0.1, 0.9, 0.1)){
+  p <- p + autolayer(sim_AR1(phi), series = paste(phi))
+}
+p +
+  labs(title="The effects of changing Phi", color = "Phi") +
+  theme(axis.title = element_blank(), legend.position = "bottom") +
+  scale_color_brewer(palette = "Set1")
+```
+
+**Now with a fixed random number seed we can see the effect.  As $ϕ_1$ increases, the distance from zero increases.  It increases the autocorrelation with the preceeding value.**
+
 c. Write your own code to generate data from an MA(1) model with $θ_1=0.6$ and $σ_2=1$.
+
+```{r}
+sim_MA1 <- function(theta){
+  set.seed(42)
+  y <- ts(numeric(100))
+  e <- rnorm(100)
+  e[1] <- 0
+  for(i in 2:100){
+    y[i] <- theta*e[i-1] + e[i]
+  }
+  return(y)
+}
+```
 
@@
 
+```{r}
+p <- autoplot(sim_MA1(0.6))
+for(theta in seq(0.1, 0.9, 0.1)){
+  p <- p + autolayer(sim_MA1(theta), series = paste(theta))
+}
+
+p +
+  labs(title="The effects of changing Theta", color = "Theta") +
+  theme(axis.title = element_blank(), legend.position = "bottom") +
+  scale_color_brewer(palette = "Set1")
+```
+
+**This is a bit harder to see but as $ϕ_1$ increases, the distance from zero again increases.  The $θ_1$ deals with the error so it is not as pronounced as a change in $ϕ_1$ though.**
+
 e. Generate data from an ARMA(1,1) model with $ϕ_1=0.6$, $θ_1=0.6$ and $σ_2=1$.
+
+```{r}
+sim_ARMA <- function(phi, theta){
+  set.seed(42)
+  y <- ts(numeric(100))
+  e <- rnorm(100)
+  e[1] <- 0
+  for(i in 2:100)
+    y[i] <- phi*y[i-1] + theta*e[i-1] + e[i]
+  return(y)
+}
+```
 
@@
 
+```{r}
+sim_AR2 <- function(phi_1, phi_2){
+  set.seed(42)
+  y <- ts(numeric(100))
+  e <- rnorm(100)
+  for(i in 3:100)
+    y[i] <- phi_1*y[i-1] + phi_2*y[i-2] + e[i]
+  return(y)
+}
+```
+
 g. Graph the latter two series and compare them.
+
+```{r}
+autoplot(sim_ARMA(0.6, 0.6), series = "ARMA(1,1)") +
+  autolayer(sim_AR2(-0.8, 0.3), series = "AR(2)") +
+  theme(axis.title = element_blank(), legend.position = "bottom", legend.title = element_blank()) +
+  scale_color_brewer(palette = "Set1")
+```
 
@@
 
+```{r}
+autoplot(wmurders) +
+  ggtitle("Women Murdered in the U.S.") +
+  theme(axis.title = element_blank())
+```
+
+**The trends (both up and down) suggest that the series should be differenced before moving forward.
+`ndiffs` estimates that `r ndiffs(wmurders)` differened are required to make this time series stationary.**
+
+```{r}
+ggtsplot(diff(wmurders, ndiffs(wmurders)), "Differenced Women Murdered in the U.S.")
+```
+
+**The first lag of the PACF is the only spike which indicates the value of p in the ARIMA model should be 1.  The high first value in the ACF indicates that the p is also one.**
+
 b. Should you include a constant in the model? Explain.
+
+**No.  Because the differencing term (d) is 2.  We should not include it in the model.**
 
@@
 g. Does auto.arima() give the same model you have chosen? If not, which model do you think is better?
+
+```{r}
+auto.arima(wmurders)
+```

### 2020-03-19 DATA624/HW6.Rmd
Updated HW6.Rmd

@@
 
-**The first lag of the PACF is the only spike which indicates the value of p in the ARIMA model should be 1.  The high first value in the ACF indicates that the p is also one.**
+**The first lag of the PACF is the only spike which indicates the value of p in the ARIMA model should be 1.  The high first value in the ACF indicates that the d is also one.**
 
@@
 
-**No.  Because the differencing term (d) is 2.  We should not include it in the model.**
+**No.
+Based on the [blog post](https://robjhyndman.com/hyndsight/arimaconstants/) by the text book author, since the differencing term (d) is 2 if c = 0 the eventual forecast function (EFF) will follow a straight line with intercept and slope determined by the last few observations.
+If c != 0 then the EFF  will follow a quadratic trend.
+Opting for the simplier model We should not include it in the model.**
 
@@
 
+**$(1-\phi_1B) (1-B)^2y_t = c + (1 + \theta_1B)e_t$**
+
 d. Fit the model using R and examine the residuals. Is the model satisfactory?
+
+```{r}
+fit <- arima(wmurders, order = c(1, 2, 1))
+checkresiduals(fit)
+```
+
+**The ACF show all autocorrelations are withing acceptable limits, indicating residuals are white noise.  The p-value of the Ljung-Box test also indicate the residuals are white noise.**
 
@@
 
+```{r}
+forecast(fit, h=3) %>%
+  kable() %>%
+  kable_styling()
+```
+
 f. Create a plot of the series with forecasts and prediction intervals for the next three periods shown.
+
+```{r}
+autoplot(forecast(fit, h=3))
+```
 
@@
 ```
+
+**Yes auto.arima() gives the same model that I have chosen.**

### 2020-03-20 DATA624/Project 1-B Power.Rmd
Work on Project 1-B

@@
 library(tidyverse)
+library(readxl)
+library(lubridate)
 library(kableExtra)
+library(fpp2)
+library(forcats)
+library(gridExtra)
+options(scipen=999)
 ```
 
-## Directions
+## Introduction
 
-Part B consists of a simple dataset of residential power usage for January 1998 until December 2013.
-Your assignment is to model these data and a monthly forecast for 2014.
-The data is given in a single file.
-The variable ‘KWH’ is power consumption in Kilowatt hours, the rest is straight forward.
-Add this to your existing files above.
+I am to forecast how much power will be consumed by residential customers in kilowatt hours for the next 12 months.  Data was provided for this project spaning January 1998 until December 2013.
+
+## Data Exploration
+
+I will begin by reading in the data and ploting the series.
+
+```{r}
+KWH <- read_excel("ResidentialCustomerForecastLoad-624.xlsx") %>%
+  select(KWH) %>%
+  ts(start = decimal_date(date("1998-01-01")), frequency = 12)
+
+autoplot(KWH, main = "Residential Power Usage (KWH)") +
+  theme(axis.title = element_blank())
+```
+
+```{r}
+summary(KWH)
+```
+
+We have outliers and missing data.  We will clean them up with `tsclean` from the forcasts package, and replot the data with the ACF and PACF:
+
+```{r}
+KWH <- tsclean(KWH)
+
+ggtsplot <- function(ts, title){
+  # A ggplot2 version of tsdisplay(df)
+  # Args:
+  #    ts (Time-Series): The time series we want to plot
+  #    title (str): The title of the graph
+  grid.arrange( 
+    autoplot(ts) +
+      ggtitle(title) +
+      theme(axis.title = element_blank()),
+    grid.arrange(
+      ggAcf(ts) + 
+        ggtitle(element_blank()), 
+      ggPacf(ts) + 
+        ggtitle(element_blank()), ncol = 2), nrow = 2)
+}
+
+ggtsplot(KWH, "Cleaned Residential Power Usage (KWH)")
+```
+
+This looks much better.  There is a seasonal component that looks to be additive.  I will need to use a model that captures seasonality.
+
+## Model Approach
+
+Since the data is highly seasonal, I we will need to use an ARIMA or Holt-Winters model.  Let's fit an models using `auto.arima()` and `hw()` and visualize their projections.
+
+```{r}
+h <- 12
+lambda <- BoxCox.lambda(KWH)
+autoplot(KWH) +
+  autolayer(hw(KWH, h=h), series="Holt-Winters") +
+  autolayer(hw(KWH, h=h, lambda = lambda), series="Holt-Winters (Box-Cox Adjusted)") +
+  autolayer(forecast(auto.arima(KWH), h=h), series="ARIMA") +
+  facet_grid(.~series) +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "none", axis.title = element_blank())
+```
+
+### Residual Analysis
+
+All models do a good job picking up the seasonality.  Let's look at the residuals to see which model did the best job.
+
+#### Holt-Winters Model
+
+```{r}
+hw_fit <- hw(KWH, h=h)
+checkresiduals(hw_fit)
+```
+
+#### Adjusted Holt-Winters Model
+
+```{r}
+adj_hw_fit <- hw(KWH, h=h, lambda = lambda)
+checkresiduals(adj_hw_fit)
+```
+
+#### ARIMA Model
+
+```{r}
+arima_fit <- auto.arima(KWH)
+#autoplot(forecast(arima_fit, h=h))
+checkresiduals(arima_fit)
+```

### 2020-03-21 DATA624/Project 1-B Power.Rmd
More work on Project 1-B

--- 2020-03-03 DATA624/Project 1-B Power.Rmd
@@
 ```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
-knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = T, comment = NA, message = F, warning = F)
 library(tidyverse)
+library(readxl)
+library(lubridate)
 library(kableExtra)
+library(fpp2)
+library(forcats)
+library(gridExtra)
+options(scipen = 999)
 ```
 
-## Directions
+## Introduction
 
-Part B consists of a simple dataset of residential power usage for January 1998 until December 2013.
-Your assignment is to model these data and a monthly forecast for 2014.
-The data is given in a single file.
-The variable ‘KWH’ is power consumption in Kilowatt hours, the rest is straight forward.
-Add this to your existing files above.
+I am to forecast how much power will be consumed by residential customers in kilowatt hours for the next 12 months.  Data was provided for this project spaning January 1998 until December 2013.
+
+## Data Exploration
+
+I will begin by reading in the data and ploting the series.
+
+```{r}
+kwh <- read_excel("ResidentialCustomerForecastLoad-624.xlsx") %>%
+  select(KWH) %>%
+  ts(start = decimal_date(date("1998-01-01")), frequency = 12)
+
+autoplot(kwh, main = "Residential Power Usage (KWH)") +
+  theme(axis.title = element_blank())
+```
+
+```{r}
+summary(kwh)
+```
+
+We have outliers and missing data.  We will clean them up with `tsclean` from the forcasts package, and replot the data with the ACF and PACF:
+
+```{r}
+kwh <- tsclean(kwh)
+
+ggtsplot <- function(ts, title) {
+  # A ggplot2 version of tsdisplay(df)
+  # Args:
+  #    ts (Time-Series): The time series we want to plot
+  #    title (str): The title of the graph
+  grid.arrange(
+    autoplot(ts) +
+      ggtitle(title) +
+      theme(axis.title = element_blank()),
+    grid.arrange(
+      ggAcf(ts) + ggtitle(element_blank()),
+      ggPacf(ts) + ggtitle(element_blank()), ncol = 2)
+    , nrow = 2)
+}
+
+ggtsplot(kwh, "Cleaned Residential Power Usage (KWH)")
+```
+
+This looks much better.  There is a seasonal component that looks to be additive.  I will need to use a model that captures seasonality.
+
+## Candidate Models
+
+Since the data is highly seasonal, I we will need to use an ARIMA or Holt-Winters model.
+Let's fit an models using `auto.arima()` and `hw()` and visualize their projections.
+I will use a couple of different parameters with the Holt-Winters model.
+
+```{r}
+h <- 12
+lambda <- BoxCox.lambda(kwh)
+autoplot(kwh) +
+  autolayer(hw(kwh, h = h), series = "Holt-Winters") +
+  autolayer(hw(kwh, h = h, lambda = lambda), series = "Holt-Winters (Box-Cox Adjusted)") +
+  autolayer(hw(kwh, h = h, lambda = lambda, damped = TRUE), series = "Holt-Winters (Damped & Box-Cox Adjusted)") +
+  autolayer(forecast(auto.arima(kwh), h = h), series = "ARIMA") +
+  facet_wrap(. ~ series, ncol = 2) +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "none", axis.title = element_blank())
+```
+
+## Residual Analysis
+
+All models do a good job picking up the seasonality.  Let's look at the residuals to see if any of the models should not be used.
+
+### Holt-Winters Model
+
+```{r}
+hw_fit <- hw(kwh, h = h)
+checkresiduals(hw_fit)
+```
+
+### Adjusted Holt-Winters Model
+
+```{r}
+adj_hw_fit <- hw(kwh, h = h, lambda = lambda)
+checkresiduals(adj_hw_fit)
+```
+
+### Damped & Adjusted Holt-Winters Model
+
+```{r}
+damp_adj_hw_fit <- hw(kwh, h = h, lambda = lambda, damped = TRUE)
+checkresiduals(damp_adj_hw_fit)
+```
+
+### ARIMA Model
+
+```{r}
+arima_fit <- auto.arima(kwh)
+checkresiduals(arima_fit)
+```
+
+## Cross Validation
+
+In order to understand how well a model is likely to preform at predicting out of sample data, I will use the `tsCV` function and evaluate the models.
+My goal is to minimize the RMSE.
+First I will get the errors from the cross validation process, then I will compute the RMSE.
+
+```{r}
+get_rmse <- function(e) {
+  sqrt(mean(e^2, na.rm=TRUE))
+}
+
+hw_error <- tsCV(kwh, hw, h = h)
+adj_hw_error <- tsCV(kwh, hw, h = h, lambda = lambda)
+damped_adj_hw_error <- tsCV(kwh, hw, h = h, lambda = lambda, damped = TRUE)
+arima_errors <- tsCV(kwh, forecast(auto.arima(kwh), h=h), h = h)
+
+data.frame(Model = c("ARIMA", "Holt-Winters", "Adjusted Holt-Winters", "Damped & Adjusted Holt-Winters"),
+           RMSE = c(get_rmse(arima_errors[,h]), get_rmse(hw_error[,h]), get_rmse(adj_hw_error[,h]), get_rmse(damped_adj_hw_error[,h]))) %>%
+  arrange(RMSE) %>%
+  kable() %>%
+  kable_styling()
+```

### 2020-03-21 DATA624/Project 1-B Power.Rmd
More work on Project 1-B

@@
 ```{r}
-kwh <- read_excel("ResidentialCustomerForecastLoad-624.xlsx") %>%
+raw_data <- read_excel("ResidentialCustomerForecastLoad-624.xlsx")
+kwh <- raw_data %>%
   select(KWH) %>%
@@
 
+There are some spikes in the ACF that indicate this is not the best model.
+
 ### Adjusted Holt-Winters Model
@@
 ```
+
+This is a bit better than the previous model but there are still ACF spikes outside of the bands.
 
@@
 
+The preformance of this model is about the same as above.
+
 ### ARIMA Model
@@
 ```
+
+This model seems to be the best.
+There is only one ACF spike outside of the threshold and it only by a small amount.
+My working theory is this is the model I should use, but I will check this using cross validation.
 
@@
 
+arima_forecast <- function(x, h){forecast(Arima(x, order = c(3, 0, 1), seasonal = c(2, 1, 0), include.drift = TRUE), h = h)}
+
 hw_error <- tsCV(kwh, hw, h = h)
@@
 damped_adj_hw_error <- tsCV(kwh, hw, h = h, lambda = lambda, damped = TRUE)
-arima_errors <- tsCV(kwh, forecast(auto.arima(kwh), h=h), h = h)
+arima_errors <- tsCV(kwh, arima_forecast, h = h)
 
@@
 ```
+
+The ARIMA model had the best cross validation results.
+
+## Model Selection
+
+Given that the ARIMA model minimized the RMSE in the cross validation, I will use it to forecast residential power consumption.
+
+## Forecast
+
+Here's my forcast for the next `r h` months of power usage:
+
+```{r}
+my_forecast <- forecast(arima_fit, h = h)
+autoplot(my_forecast)
+  ggtitle("Forecasted Residential Power Consumption (KWH)") +
+  theme(axis.title = element_blank())
+
+my_forecast %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+```{r}
+data.frame(my_forecast$mean) %>%
+  rename(KWH = my_forecast.mean)
+```

### 2020-03-21 DATA624/Project-1-B-Power.Rmd
More work on Project 1-B

--- 2020-03-21 DATA624/Project 1-B Power.Rmd
@@
 library(gridExtra)
+library(openxlsx)
 options(scipen = 999)
@@
 ```{r}
+if (!file.exists("ResidentialCustomerForecastLoad-624.xlsx")) {
+  download.file("https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA624/ResidentialCustomerForecastLoad-624.xlsx", "ResidentialCustomerForecastLoad-624.xlsx")
+}
 raw_data <- read_excel("ResidentialCustomerForecastLoad-624.xlsx")
@@
 get_rmse <- function(e) {
-  sqrt(mean(e^2, na.rm=TRUE))
+  sqrt(mean(e^2, na.rm = TRUE))
 }
 
-arima_forecast <- function(x, h){forecast(Arima(x, order = c(3, 0, 1), seasonal = c(2, 1, 0), include.drift = TRUE), h = h)}
+arima_forecast <- function(x, h) {
+  forecast(Arima(x, order = c(3, 0, 1), seasonal = c(2, 1, 0), include.drift = TRUE), h = h)
+}
 
@@
 data.frame(Model = c("ARIMA", "Holt-Winters", "Adjusted Holt-Winters", "Damped & Adjusted Holt-Winters"),
-           RMSE = c(get_rmse(arima_errors[,h]), get_rmse(hw_error[,h]), get_rmse(adj_hw_error[,h]), get_rmse(damped_adj_hw_error[,h]))) %>%
+           RMSE = c(get_rmse(arima_errors[, h]), get_rmse(hw_error[, h]), get_rmse(adj_hw_error[, h]), get_rmse(damped_adj_hw_error[, h]))) %>%
   arrange(RMSE) %>%
@@
 my_forecast <- forecast(arima_fit, h = h)
-autoplot(my_forecast)
+autoplot(my_forecast) +
   ggtitle("Forecasted Residential Power Consumption (KWH)") +
@@
 ```{r}
-data.frame(my_forecast$mean) %>%
-  rename(KWH = my_forecast.mean)
+max_case_sequence <- max(raw_data$CaseSequence)
+projections <- data.frame(my_forecast$mean) %>%
+  rename(KWH = my_forecast.mean) %>%
+  mutate(CaseSequence = row_number()) %>%
+  rowwise() %>%
+  mutate(`YYYY-MMM` = format(ymd(paste0("2014", ifelse(CaseSequence < 10, paste0("0", CaseSequence), CaseSequence), "01")), "%Y-%b")) %>%
+  mutate(CaseSequence = CaseSequence + max_case_sequence) %>%
+  ungroup() %>%
+  select(CaseSequence, `YYYY-MMM`, KWH) %>%
+  write.xlsx("2014ResidentialCustomerForecastLoad.xlsx")
 ```

### 2020-03-22 DATA624/Project 1-A ATM.Rmd
A lot of good work on Project 1-A

--- 2020-03-07 DATA624/Project 1-A ATM.Rmd
@@
 library(fpp2)
-library(prophet)
 library(forcats)
+library(gridExtra)
+options(scipen=999)
 ```
@@
   ggplot(aes(DATE, Cash, color=ATM)) +
-  geom_line(lwd=1) +
+  geom_line() +
   ggtitle("Cash Withdrawal by Date and ATM") +
   scale_color_brewer(palette = "Set1") +
+  facet_wrap(.~ATM, ncol = 2) +
   theme(legend.title = element_blank(),
@@
   ggplot(aes(DATE, Cash, color=ATM)) +
-  geom_line(lwd=1) +
+  geom_line() +
+  facet_wrap(.~ATM, ncol = 2) +
   ggtitle("Cash Withdrawal by Date and ATM") +
@@
 
-### Dealing with Missing Data & Outliers
+### Is there any Daily Effects?
 
-ATM #1 is missing 3 observations and ATM #2 is missing 2.  I will replace the NA's with the median values.  We will also replace all the outliers in ATM #4 with the $200,000.
+There is a pattern in the time series but fluctuations are very tight.
+I wonder if it is explained by the day of the week.
+For example, Friday and Saturday might have heavier usage and a Tuesday night might be calm.
+Let's examine this hypothesis:
 
 ```{r}
-adjusted_atm <- atm %>% 
-  rowwise() %>%
-  mutate(Cash = if_else(is.na(Cash) & ATM == "ATM1", 91, Cash)) %>%
-  mutate(Cash = if_else(is.na(Cash) & ATM == "ATM2", 67, Cash)) %>%
-  mutate(Cash = if_else(Cash > 2000, 2000, Cash)) %>%
-  ungroup() %>%
-  data.frame()
+atm %>%
+  na.omit() %>%
+  mutate(`Day of Week` = recode(as.factor(wday(DATE)), "1" = "Sunday", "2" = "Monday", "3" = "Tuesday", "4"="Wednesday", "5"="Thursday", "6"="Friday","7"="Saturday"))  %>%
+  ggplot(aes(`Day of Week`, Cash, color = `Day of Week`)) +
+  geom_boxplot() +
+  facet_grid(ATM ~ `Day of Week`, scales = "free") +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "none", axis.title = element_blank(), axis.text.x = element_blank(), axis.ticks = element_blank())
 ```
 
-## Model Building
+It looks like there is some varriation based on the day of the week.
+Thursday looks to be less busy than the other days of the week.
+There is so much variation between the 4 ATMs that each will be modeled seperately.
 
-I will construct a series of models using different approaches.
-I will start from the most simple approach and move on to more sophisticated ones.
-I beleive the "best" model will come from the middle of the pack.
-Let's see how it plays out.
+## A Quick Note on Process
 
-Since the task is to predict the cash withdrawls for May 2010 (31 days) we will set up test sets that include the last 31 days.
-Each model will be evaluated on how accurate they are in forecasting the test set's values.
-I will use the RMSE as the metric to assess the accuracy.
+I will clean up the data to produce the forecasts.
+I will then explore some candidate modeling techniques.
+Once I have a handful of candidates, I will prefrom a cross-validation on the models and get the RMSE.
+The model that minimizes the RMSE during cross-validation will be selected as the model of choice.
+
+## ATM #1
+
+### Data Cleanup
+
+I will begin ATM #1.  There are 3 missing observations that will need to be cleaned up.
 
 ```{r}
-train_test_split <- function(atm_df, atm_number, n_in_test = 31, return_ts=TRUE){
-  ### Splits the data into training and test sets
+atm1 <- atm %>%
+  filter(ATM == "ATM1")
+
+atm1$clean_cash <- c(tsclean(ts(atm1$Cash, start = decimal_date(as.Date(min(atm1$DATE))), frequency = 365)))
+```
+
+Now that we have a complete data set I will create time series objects.
+Since there is a weekly effect I will be using a frequency of 7.
+This does mess up the dates in the plot however so please pay no reguard to them.
+
+```{r}
+ggtsplot <- function(ts, title) {
+  # A ggplot2 version of tsdisplay(df)
   # Args:
-  #     atm_df (data.frame): The ATM data
-  #     atm_number (int): Which ATM do you want the time series for?
-  #     n_in_test (int): The number of observations in the test set taken from the end (optional-31 default)
-  #     return_ts (logical): Should the function return ts objects (optional-TRUE default)
-  # Returns:
-  #     train_test_ts (list): Two ts objects (train and test).
-
-  # Subset the data frame
-  df <- atm_df %>%
-    filter(ATM == paste0("ATM", atm_number)) %>%
-    select(-ATM)
-
-  # Split the data frame into training and test sets
-  test <- tail(df, n_in_test)
-  train <- head(df, -n_in_test)
-
-  # Tranform into ts objects
-  if(return_ts){
-    train_start <- (min(train$DATE))
-    train <- ts(train$Cash, start = decimal_date(as.Date(train_start)), frequency = 365)
-    test_start <- (min(test$DATE))
-    test <- ts(test$Cash, start = decimal_date(as.Date(test_start)), frequency = 365)
+  #    ts (Time-Series): The time series we want to plot
+  #    title (str): The title of the graph
+  grid.arrange(
+    autoplot(ts) +
+      ggtitle(title) +
+      theme(axis.title = element_blank()),
+    grid.arrange(
+      ggAcf(ts) + ggtitle(element_blank()),
+      ggPacf(ts) + ggtitle(element_blank()), ncol = 2)
+    , nrow = 2)
 }
 
-  # Return the data to the user
-  train_test <- list(train = train, test = test)
-  return(train_test)
+atm1_ts <- ts(atm1$clean_cash, start = decimal_date(as.Date(min(atm1$DATE))), frequency = 7)
+ggtsplot(atm1_ts, "ATM #1")
+```
+
+One readily observes the repeating peaks on every 7th lag.  There also seems to be an interesting negative correlation between the 1st and both the 3rd and the 5th lag.
+
+### Model Creation
+
+#### STL Decomposition Models
+
+I will try a couple of seasonal decomposition models.  I will set the seasonal window to 7 so it picks up the day of the week variation.
+
+```{r}
+atm1_ts %>%
+  stl(s.window = 7, robust = TRUE) %>%
+  autoplot()
+```
+
+This seems promising to me.  I will do STL decomposition forecasts using both the ETS and ARIMA models and check their residual plots
+
+#### STL + ETS
+
+```{r}
+h <- 31
+atm1_stl_ets_fit <- atm1_ts %>%
+  stlf(h = h, s.window = 7, robust = TRUE, method="ets")
+checkresiduals(atm1_stl_ets_fit)
+```
+
+This model seems to hold merit and should be taken under consideration.
+
+#### STL + ARIMA
+
+```{r}
+atm1_stl_arima_fit <- atm1_ts %>%
+  stlf(h = h, s.window = 7, robust = TRUE, method="arima") 
+checkresiduals(atm1_stl_arima_fit)
+```
+
+This model also preformed well.  It is definately a candidate for the cross-validation stage.
+
+#### Holt-Winters
+
+```{r}
+atm1_hw_fit <- hw(atm1_ts, h = h)
+checkresiduals(atm1_hw_fit)
+```
+
+This method did a fairly good job.  I will move it on to the next phase.
+
+### Holt-Winters with Box Cox Adjustment
+
+```{r}
+atm1_lambda <- BoxCox.lambda(atm1_ts)
+atm1_adj_hw_fit <- hw(atm1_ts, h = h, lambda = atm1_lambda)
+checkresiduals(atm1_adj_hw_fit)
+```
+
+This seems to be a strong candidate.  We will see how if fares in the cross-validation.
+
+#### ARIMA
+
+```{r}
+atm1_arima_fit <- auto.arima(atm1_ts)
+checkresiduals(atm1_arima_fit)
+```
+
+This model looks like it is preforming well.  Let's see how all of them stack up.
+
+### Cross Validation
+
+In order to understand how well a model is likely to preform at predicting out of sample data, I will use the `tsCV` function and evaluate the models.
+As prevously noted my goal is to minimize the RMSE.
+First I will get the errors from the cross validation process, then I will compute the RMSE.
+
+```{r}
+get_rmse <- function(e) {
+  sqrt(mean(e^2, na.rm = TRUE))
 }
 
-atm_1_ts <- train_test_split(adjusted_atm, 1, 31)
-atm_2_ts <- train_test_split(adjusted_atm, 2, 31)
-atm_3_ts <- train_test_split(adjusted_atm, 3, 31)
-atm_4_ts <- train_test_split(adjusted_atm, 4, 31)
+atm1_arima_forecast <- function(x, h) {
+  forecast(Arima(x, order = c(0, 0, 1), seasonal = c(0, 1, 2)), h = h)
+}
 
-atm_1_df <- train_test_split(adjusted_atm, 1, 31, return_ts = FALSE)
-atm_2_df <- train_test_split(adjusted_atm, 2, 31, return_ts = FALSE)
-atm_3_df <- train_test_split(adjusted_atm, 3, 31, return_ts = FALSE)
-atm_4_df <- train_test_split(adjusted_atm, 4, 31, return_ts = FALSE)
+stl_ets_errors <- tsCV(atm1_ts, stlf, h = h, s.window = 7, robust = TRUE, method="ets")
+stl_arima_errors <- tsCV(atm1_ts, stlf, h = h, s.window = 7, robust = TRUE, method="arima")
+hw_errors <- tsCV(atm1_ts, hw, h = h)
+adj_hw_errors <- tsCV(atm1_ts, hw, h = h, lambda = atm1_lambda)
+arima_errors <- tsCV(atm1_ts, atm1_arima_forecast, h = h)
+
+data.frame(Model = c("STL ETS", "STL ARIMA", "ARIMA", "Holt-Winters", "Adjusted Holt-Winters"),
+           RMSE = c(get_rmse(stl_ets_errors[, h]), get_rmse(stl_arima_errors[, h]), get_rmse(arima_errors[, h]), get_rmse(hw_errors[, h]), get_rmse(adj_hw_errors[, h]))) %>%
+  arrange(RMSE) %>%
+  kable() %>%
+  kable_styling()
 ```
 
-### ATM 1 Models
+It looks like the ARIMA model did the best job predicting on the out of sample data set.
 
-I will start by looking at the training data and the decomposition.
-I cannot use the `decompose()` or `stl()` functions to decompose it because I don't have enough data.
-Facebook Prophet can decompose it using its Basian process.
+### Model Selection
+
+I will select the ARIMA model to produce the forecast for ATM #1.  I have decided on this because it is the model that preformed the best in the cross validation.
+
+## ATM #2
+
+Now I will repeat the above process for ATM #2.  As this is a repeat I will not include as much explanatory text.
+
+### Data Cleanup
+
+There are 2 missing observations.  This will be cleaned up using the `tsclean` function again.
 
 ```{r}
-atm_prophet <- function(df, forcast_window){
-  m <- df %>%
-    mutate(y = Cash, ds = as.character(DATE)) %>%
-    prophet()
-  future <- make_future_dataframe(m, periods = forcast_window)
-  atm_forecast <- predict(m, future)
-  results = list(
-    model = m,
-    future = future,
-    forecast = atm_forecast
-  )
-  return(results)
+atm2 <- atm %>%
+  filter(ATM == "ATM2")
+
+atm2$clean_cash <- c(tsclean(ts(atm2$Cash, start = decimal_date(as.Date(min(atm2$DATE))), frequency = 365)))
+
+atm2_ts <- ts(atm2$clean_cash, start = decimal_date(as.Date(min(atm2$DATE))), frequency = 7)
+ggtsplot(atm2_ts, "ATM #2")
+```
+
+Once again the ACF plot has the regular spikes on evry multiple of 7.
+
+### Model Creation
+
+#### STL Decomposition Models
+
+Again I will try a couple of seasonal decomposition models.  I will set the seasonal window to 7 so it picks up the day of the week variation.
+
+```{r}
+atm2_ts %>%
+  stl(s.window = 7, robust = TRUE) %>%
+  autoplot()
+```
+
+This seems very similar to ATM #1.
+
+#### STL + ETS
+
+```{r}
+atm2_stl_ets_fit <- atm2_ts %>%
+  stlf(h = h, s.window = 7, robust = TRUE, method="ets")
+checkresiduals(atm2_stl_ets_fit)
+```
+
+This model preformed much better than the it did for the ATM #1 time series.  It will be interesting to see how it does in cross-validation
+
+#### STL + ARIMA
+
+```{r}
+atm2_stl_arima_fit <- atm2_ts %>%
+  stlf(h = h, s.window = 7, robust = TRUE, method="arima") 
+checkresiduals(atm2_stl_arima_fit)
+```
+
+Interesting.  The residuals have a bit more spread but the left tail is shorter than the STL+ETS model.
+
+#### Holt-Winters
+
+```{r}
+atm2_hw_fit <- hw(atm2_ts, h = h)
+checkresiduals(atm2_hw_fit)
+```
+
+This method did not preform as well as the others.  I will, however keep it in so I can compare it to ATM #1's statistics.
+
+### Holt-Winters with Box Cox Adjustment
+
+```{r}
+atm2_lambda <- BoxCox.lambda(atm2_ts)
+atm2_adj_hw_fit <- hw(atm2_ts, h = h, lambda = atm2_lambda)
+checkresiduals(atm2_adj_hw_fit)
+```
+
+#### ARIMA
+
+```{r}
+atm2_arima_fit <- auto.arima(atm2_ts)
+checkresiduals(atm2_arima_fit)
+```
+
+This model looks like it is preforming well.  Let's see how all of them stack up.
+
+### Cross Validation
+
+In order to understand how well a model is likely to preform at predicting out of sample data, I will use the `tsCV` function and evaluate the models.
+As prevously noted my goal is to minimize the RMSE.
+First I will get the errors from the cross validation process, then I will compute the RMSE.
+
+```{r}
+atm2_arima_forecast <- function(x, h) {
+  forecast(Arima(x, order = c(2, 0, 2), seasonal = c(0, 1, 1)), h = h)
 }
 
-atm_1_prophet <- atm_prophet(atm_1_df$train, 31)
-prophet_plot_components(atm_1_prophet$model, atm_1_prophet$forecast)
+stl_ets_errors <- tsCV(atm2_ts, stlf, h = h, s.window = 7, robust = TRUE, method="ets")
+stl_arima_errors <- tsCV(atm2_ts, stlf, h = h, s.window = 7, robust = TRUE, method="arima")
+hw_errors <- tsCV(atm2_ts, hw, h = h)
+adj_hw_errors <- tsCV(atm2_ts, hw, h = h, lambda = atm2_lambda)
+arima_errors <- tsCV(atm2_ts, atm2_arima_forecast, h = h)
+
+data.frame(Model = c("STL ETS", "STL ARIMA", "ARIMA", "Holt-Winters", "Adjusted Holt-Winters"),
+           RMSE = c(get_rmse(stl_ets_errors[, h]), get_rmse(stl_arima_errors[, h]), get_rmse(arima_errors[, h]), get_rmse(hw_errors[, h]), get_rmse(adj_hw_errors[, h]))) %>%
+  arrange(RMSE) %>%
+  kable() %>%
+  kable_styling()
 ```
 
-That is interesting! There appears to be a weekly effect in the data.  Wednesdays and especially Thursdays are low traffic days.  The trend apears to be decreasing but at a very slow rate.
+Interesting.
+It looks like the ARIMA model was the top preformer again.
+I find it interesting that the RMSE is higher and more spread out for ATM #2.
+ATM #1's RMSEs ranged from roughyl 30 to 39.
+The RMSEs for ATM #2 ranged from 34 to 55.
+There is more variability in this data that isn't captured by the model.
+
+### Model Selection
+
+As stated in the previous section I will use the ARIMA model to produce the forecast for ATM #2.
+
+## ATM #3

### 2020-03-23 DATA624/Project 1-A ATM.Rmd
More work on Project 1

@@
 ## ATM #3
+
+This model is quite different from the two proceeding cases.  You can see it in the plot:
+
+```{r}
+atm %>%
+  filter(ATM == "ATM3") %>%
+  mutate(nonzero = if_else(Cash == 0, "No", "Yes")) %>%
+  ggplot(aes(DATE, Cash, color=nonzero)) +
+  geom_point() +
+  ggtitle("ATM #3") +
+  scale_color_brewer(palette = "Set1") +
+  theme(axis.title = element_blank(), legend.position = "none")
+```
+
+Most of the values are zeros (points in red above), except for 3 points (shown in blue).  The three non-zero points are the most current.
+
+## ATM #4
+
+This ATM is different too, but not as radically different as ATM #3.  I will be using the same approach used for ATM #1 and #2 with this ATM.
+
+### Data Cleanup
+
+There is one major outlier in the data set.  We will clean it up by simply by using the `tsclean` function once again.
+
+```{r}
+atm4 <- atm %>%
+  filter(ATM == "ATM4")
+
+atm4$clean_cash <- c(tsclean(ts(atm4$Cash, start = decimal_date(as.Date(min(atm4$DATE))), frequency = 365)))
+
+atm4_ts <- ts(atm4$clean_cash, start = decimal_date(as.Date(min(atm4$DATE))), frequency = 7)
+ggtsplot(atm4_ts, "ATM #4")
+```
+
+Once again there is the familiar patter of peaks on the multiples of seven.
+
+### Model Creation
+
+#### STL Decomposition Models
+
+Again I will try a couple of seasonal decomposition models.  I will set the seasonal window to 7 so it picks up the day of the week variation.
+
+```{r}
+atm2_ts %>%
+  stl(s.window = 7, robust = TRUE) %>%
+  autoplot()
+```
+
+This seems very similar to ATM #1.
+
+#### STL + ETS
+
+```{r}
+atm4_stl_ets_fit <- atm4_ts %>%
+  stlf(h = h, s.window = 7, robust = TRUE, method="ets")
+checkresiduals(atm4_stl_ets_fit)
+```
+
+This model did a fair job.  There are a couple of ACF spikes that are outside the bands.  Let's see if the ARIMA model does better:
+
+#### STL + ARIMA
+
+```{r}
+atm4_stl_arima_fit <- atm4_ts %>%
+  stlf(h = h, s.window = 7, robust = TRUE, method="arima") 
+checkresiduals(atm4_stl_arima_fit)
+```
+
+This is an improvement over the STL+ETS model.  There are still a couple of spikes on the ACF that falls outside the threshold.
+
+#### Holt-Winters
+
+```{r}
+atm4_hw_fit <- hw(atm4_ts, h = h)
+checkresiduals(atm4_hw_fit)
+```
+
+This method seems to have preformed better than any of the STL models.
+
+### Holt-Winters with Box Cox Adjustment
+
+```{r}
+atm4_lambda <- BoxCox.lambda(atm4_ts)
+atm4_adj_hw_fit <- hw(atm4_ts, h = h, lambda = atm4_lambda)
+checkresiduals(atm4_adj_hw_fit)
+```
+
+This model preformed very well.  It doesn't look like any of the ACF spikes are outside the bands.  This is a contender for sure.
+
+#### ARIMA
+
+```{r}
+atm4_arima_fit <- auto.arima(atm4_ts)
+checkresiduals(atm4_arima_fit)
+```
+
+This model did fairly well, but it seems like the Holt-Winters with Box-Cox adjustment did better.  It's time to cross-validate and see how all of them preform.
+
+### Cross Validation
+
+Once again, I will use the `tsCV` function and evaluate the models.  My goal to minimize the RMSE remains for this dataset.
+
+```{r}
+atm4_arima_forecast <- function(x, h) {
+  forecast(Arima(x, order = c(0, 0, 3), seasonal = c(1, 0, 0)), h = h)
+}
+
+stl_ets_errors <- tsCV(atm4_ts, stlf, h = h, s.window = 7, robust = TRUE, method="ets")
+stl_arima_errors <- tsCV(atm4_ts, stlf, h = h, s.window = 7, robust = TRUE, method="arima")
+hw_errors <- tsCV(atm4_ts, hw, h = h)
+adj_hw_errors <- tsCV(atm4_ts, hw, h = h, lambda = atm4_lambda)
+arima_errors <- tsCV(atm4_ts, atm4_arima_forecast, h = h)
+
+data.frame(Model = c("STL ETS", "STL ARIMA", "ARIMA", "Holt-Winters", "Adjusted Holt-Winters"),
+           RMSE = c(get_rmse(stl_ets_errors[, h]), get_rmse(stl_arima_errors[, h]), get_rmse(arima_errors[, h]), get_rmse(hw_errors[, h]), get_rmse(adj_hw_errors[, h]))) %>%
+  arrange(RMSE) %>%
+  kable() %>%
+  kable_styling()
+```
+
+That's an unexpected result.  The Adjusted Holt-Winters (the one I thought would preform the best), did not preform as well.  The ARIMA model, once again, rose to the top.
+
+### Model Selection
+
+Once again I will use the forecasts produced by the ARIMA model as the projections for ATM #4.

### 2020-03-23 DATA624/Project 1-A ATM.Rmd
More work on Project 1

@@
 library(gridExtra)
+library(openxlsx)
 options(scipen=999)
@@
 ```{r}
+if(!file.exists("ATM624Data.xlsx")) {
+  download.file("https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA624/ATM624Data.xlsx", "ATM624Data.xlsx")
+}
+
 atm <- read_excel("ATM624Data.xlsx") %>%
@@
 
-Most of the values are zeros (points in red above), except for 3 points (shown in blue).  The three non-zero points are the most current.
+Most of the values are zeros (points in red above), except for 3 points (shown in blue).  The three non-zero points are the most current.  This presents a serious challenge.
+
+### Model Creation/Selection
+
+There is one fundamental question with this dataset.
+Are the three points an outlier, or is it the begining of the new normal?
+If they are outliers one would expect the cash value to return to zero.
+If the three points are an indication of change, then the historical data have little relevance.
+
+I will be assuming the new data are the begining of the new normal.
+The challenge is we only have three data points.
+In the absense of more data I will calculate the average of these three points and use it for the forecast with the recommendation to revise it frequently.
+As the average is only based off of three data points it should not be considered stable.
+
+```{r}
+atm3 <- atm %>%
+  filter(ATM == "ATM3", Cash > 0)
+
+atm3_mean <- mean(atm3$Cash)
+```
 
@@
 Once again I will use the forecasts produced by the ARIMA model as the projections for ATM #4.
+
+## Summary
+
+I set out to create predictions for 4 different ATMs.
+After testing multiple approaches using cross-validation, I selected ARIMA models for ATM #1, #2 and #4, as it was the modeling technique with the lowest RMSE.
+For ATM #3 we used the mean of all non-zero data (3 observations).
+This model needs to be updated once more data becomes available.
+I will finish this project by exporting my forcasts in the same file format as the original data.
+
+```{r, echo=FALSE}
+dates <- seq(ymd('2010-05-01'), ymd('2010-05-31'), by = "1 day")
+
+atm1_forecast <- forecast(atm1_arima_fit, h=h)
+atm2_forecast <- forecast(atm2_arima_fit, h=h)
+atm3_forecast <- rep(atm3_mean, h)
+atm4_forecast <- forecast(atm4_arima_fit, h=h)
+
+forecasts_df <- data.frame("DATE" = dates, "ATM" = c("ATM1"), "Cash" = c(atm1_forecast$mean))
+forecasts_df <- data.frame("DATE" = dates, "ATM" = c("ATM2"), "Cash" = c(atm2_forecast$mean)) %>%
+  rbind(forecasts_df, .)
+forecasts_df <- data.frame("DATE" = dates, "ATM" = c("ATM3"), "Cash" = atm3_forecast) %>%
+  rbind(forecasts_df, .)
+forecasts_df <- data.frame("DATE" = dates, "ATM" = c("ATM4"), "Cash" = c(atm4_forecast$mean)) %>%
+  rbind(forecasts_df, .)
+
+write.xlsx(forecasts_df, "ATM_Forecasts.xlsx")
+```

### 2020-03-23 DATA624/Project 1-B Power.Rmd
More work on Project 1

@@
 
-## Forecast
+## Summary
 
+I set out to forecast residential power consumption.
+Four modeling techniques were tested for accuracy using cross-validation.
+In the end an ARIMA model was selected bercause it minimized the RMSE.
 Here's my forcast for the next `r h` months of power usage:
@@
 
-```{r}
-data.frame(my_forecast$mean) %>%
-  rename(KWH = my_forecast.mean)
+```{r, echo=FALSE}
+max_case_sequence <- max(raw_data$CaseSequence)
+data.frame(KWH = round(my_forecast$mean)) %>%
+  mutate(CaseSequence = row_number()) %>%
+  rowwise() %>%
+  mutate(`YYYY-MMM` = format(ymd(paste0("2014", ifelse(CaseSequence < 10, paste0("0", CaseSequence), CaseSequence), "01")), "%Y-%b")) %>%
+  mutate(CaseSequence = CaseSequence + max_case_sequence) %>%
+  ungroup() %>%
+  select(CaseSequence, `YYYY-MMM`, KWH) %>%
+  rbind(raw_data, .) %>%
+  write.xlsx("Residential_Customer_Load_Forecast.xlsx")
 ```

### 2020-03-24 DATA624/Project 1-A ATM.Rmd
Linted Project 1-A

@@
 ```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
-knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
 library(tidyverse)
@@
 library(openxlsx)
-options(scipen=999)
+options(scipen = 999)
 ```
@@
 ```{r}
-if(!file.exists("ATM624Data.xlsx")) {
+if (!file.exists("ATM624Data.xlsx")) {
   download.file("https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA624/ATM624Data.xlsx", "ATM624Data.xlsx")
@@
 atm %>%
-  ggplot(aes(DATE, Cash, color=ATM)) +
+  ggplot(aes(DATE, Cash, color = ATM)) +
   geom_line() +
@@
   filter(!is.na(ATM)) %>%
-  ggplot(aes(DATE, Cash, color=ATM)) +
+  ggplot(aes(DATE, Cash, color = ATM)) +
   geom_line() +
@@
   na.omit() %>%
-  mutate(`Day of Week` = recode(as.factor(wday(DATE)), "1" = "Sunday", "2" = "Monday", "3" = "Tuesday", "4"="Wednesday", "5"="Thursday", "6"="Friday","7"="Saturday"))  %>%
+  mutate(`Day of Week` = recode(as.factor(wday(DATE)), "1" = "Sunday", "2" = "Monday", "3" = "Tuesday", "4" = "Wednesday", "5" = "Thursday", "6" = "Friday", "7" = "Saturday"))  %>%
   ggplot(aes(`Day of Week`, Cash, color = `Day of Week`)) +
@@
 atm1_stl_ets_fit <- atm1_ts %>%
-  stlf(h = h, s.window = 7, robust = TRUE, method="ets")
+  stlf(h = h, s.window = 7, robust = TRUE, method = "ets")
 checkresiduals(atm1_stl_ets_fit)
@@
 atm1_stl_arima_fit <- atm1_ts %>%
-  stlf(h = h, s.window = 7, robust = TRUE, method="arima") 
+  stlf(h = h, s.window = 7, robust = TRUE, method = "arima")
 checkresiduals(atm1_stl_arima_fit)
@@
 
-stl_ets_errors <- tsCV(atm1_ts, stlf, h = h, s.window = 7, robust = TRUE, method="ets")
-stl_arima_errors <- tsCV(atm1_ts, stlf, h = h, s.window = 7, robust = TRUE, method="arima")
+stl_ets_errors <- tsCV(atm1_ts, stlf, h = h, s.window = 7, robust = TRUE, method = "ets")
+stl_arima_errors <- tsCV(atm1_ts, stlf, h = h, s.window = 7, robust = TRUE, method = "arima")
 hw_errors <- tsCV(atm1_ts, hw, h = h)
@@
 atm2_stl_ets_fit <- atm2_ts %>%
-  stlf(h = h, s.window = 7, robust = TRUE, method="ets")
+  stlf(h = h, s.window = 7, robust = TRUE, method = "ets")
 checkresiduals(atm2_stl_ets_fit)
@@
 atm2_stl_arima_fit <- atm2_ts %>%
-  stlf(h = h, s.window = 7, robust = TRUE, method="arima") 
+  stlf(h = h, s.window = 7, robust = TRUE, method = "arima")
 checkresiduals(atm2_stl_arima_fit)
@@
 
-stl_ets_errors <- tsCV(atm2_ts, stlf, h = h, s.window = 7, robust = TRUE, method="ets")
-stl_arima_errors <- tsCV(atm2_ts, stlf, h = h, s.window = 7, robust = TRUE, method="arima")
+stl_ets_errors <- tsCV(atm2_ts, stlf, h = h, s.window = 7, robust = TRUE, method = "ets")
+stl_arima_errors <- tsCV(atm2_ts, stlf, h = h, s.window = 7, robust = TRUE, method = "arima")
 hw_errors <- tsCV(atm2_ts, hw, h = h)
@@
   mutate(nonzero = if_else(Cash == 0, "No", "Yes")) %>%
-  ggplot(aes(DATE, Cash, color=nonzero)) +
+  ggplot(aes(DATE, Cash, color = nonzero)) +
   geom_point() +
@@
 atm4_stl_ets_fit <- atm4_ts %>%
-  stlf(h = h, s.window = 7, robust = TRUE, method="ets")
+  stlf(h = h, s.window = 7, robust = TRUE, method = "ets")
 checkresiduals(atm4_stl_ets_fit)
@@
 atm4_stl_arima_fit <- atm4_ts %>%
-  stlf(h = h, s.window = 7, robust = TRUE, method="arima") 
+  stlf(h = h, s.window = 7, robust = TRUE, method = "arima")
 checkresiduals(atm4_stl_arima_fit)
@@
 
-stl_ets_errors <- tsCV(atm4_ts, stlf, h = h, s.window = 7, robust = TRUE, method="ets")
-stl_arima_errors <- tsCV(atm4_ts, stlf, h = h, s.window = 7, robust = TRUE, method="arima")
+stl_ets_errors <- tsCV(atm4_ts, stlf, h = h, s.window = 7, robust = TRUE, method = "ets")
+stl_arima_errors <- tsCV(atm4_ts, stlf, h = h, s.window = 7, robust = TRUE, method = "arima")
 hw_errors <- tsCV(atm4_ts, hw, h = h)
@@
 ```{r, echo=FALSE}
-dates <- seq(ymd('2010-05-01'), ymd('2010-05-31'), by = "1 day")
+dates <- seq(ymd("2010-05-01"), ymd("2010-05-31"), by = "1 day")
 
-atm1_forecast <- forecast(atm1_arima_fit, h=h)
-atm2_forecast <- forecast(atm2_arima_fit, h=h)
+atm1_forecast <- forecast(atm1_arima_fit, h = h)
+atm2_forecast <- forecast(atm2_arima_fit, h = h)
 atm3_forecast <- rep(atm3_mean, h)
-atm4_forecast <- forecast(atm4_arima_fit, h=h)
+atm4_forecast <- forecast(atm4_arima_fit, h = h)
 

### 2020-03-27 DATA624/Project 1-C Waterflow.Rmd
More work on Project 1-C

@@
 ```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
-knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
 library(tidyverse)
+library(lubridate)
 library(kableExtra)
+library(readxl)
+library(fpp2)
+library(forcats)
+library(gridExtra)
+library(openxlsx)
+library(tseries)
+library(scales)
+knitr::knit_hooks$set(inline = function(x) { if(!is.numeric(x)){ x }else{ prettyNum(round(x,2), big.mark=",") } })
+options(scipen = 999)
 ```
 
-## Directions
+## Introduction
 
-Part C consists of two data sets.
-These are simple 2 columns sets, however they have different time stamps.
-Your optional assignment is to time-base sequence the data and aggregate based on hour (example of what this looks like, follows).
-Note for multiple recordings within an hour, take the mean.
-Then to determine if the data is stationary and can it be forecast.
-If so, provide a week forward forecast and present results via Rpubs and .rmd and the forecast in an Excel readable file.
+For this project I have data on water flows from two pipes.
+The data are captured with timestamps.
+I am to forecast a week forward.
+In the instructions for this project I am given the following hint: For multiple recordings within an hour, take the mean.
+I will be on the lookout for data under the hour.
+
+## Data Exploration
+
+I will begin by plotting the datasets:
+
+```{r}
+pipe_1 <- read_excel("Waterflow_Pipe1.xlsx") %>%
+  mutate(`Date Time` = as.POSIXct(`Date Time` * (60 * 60 * 24), origin = "1899-12-30", tz = "GMT"))
+
+pipe_2 <- read_excel("Waterflow_Pipe2.xlsx") %>%
+  mutate(`Date Time` = as.POSIXct(`Date Time` * (60 * 60 * 24), origin = "1899-12-30", tz = "GMT"))
+
+pipes <- pipe_1 %>%
+  mutate(Pipe = "Pipe 1") %>%
+  rbind(mutate(pipe_2, Pipe = "Pipe 2"))
+
+ggplot(pipes, aes(`Date Time`, WaterFlow, color=Pipe)) +
+  geom_line() +
+  scale_color_brewer(palette = "Set1") +
+  facet_wrap(. ~ Pipe) +
+  theme(legend.position = "none", axis.title = element_blank())
+```
+
+The two pipes are quite different in terms of the data they offer.
+Pipe 1 has a median waterflow of about `r round(median(pipe_1$WaterFlow))`, while pipe 2 has a median of `r round(median(pipe_2$WaterFlow))`.
+Pipe 1 also has what looks like over a week of data, while pipe 2 has over a month.
+Both datasets have 1000 observations.
+This suggests that pipe one has more fine grain observations (i.e.
+multiple observations in a hour or minute) while pipe 2's data captures a longer period of time.
+That would explain the difference in median and the length of time covered by the respective data sets.
+This also is hinted at in the instructions
+
+### Closer Look at Pipe 1
+
+I will look at the head and tail of pipe 1's data to get a sense of what's included:
+
+```{r}
+head(pipe_1) %>% kable() %>% kable_styling()
+```
+
+There is multiple observations per minute which confirms my suspicions.  I will need to create the averages per minute as directed.
+
+### Closer Look at Pipe 2
+
+I will also take a look at pipe 2's data to see if there is anything I need to be aware of:
+
+```{r}
+head(pipe_2) %>% kable() %>% kable_styling()
+```
+
+Hmmm.  Some of the timestamps are just a bit off from the hour mark.  I will need to clean that up.
+
+## Data Cleanup
+
+### Pipe 1
+
+I will clean up the pipe 1 data by creating the averages per minute.
+
+```{r}
+pipe_1 <- pipe_1 %>%
+  mutate(Date = as.Date(`Date Time`),
+         Time = paste0(format(`Date Time`, "%H"),":00:00")) %>%
+  group_by(Date, Time) %>%
+  summarise(WaterFlow = mean(WaterFlow)) %>%
+  ungroup() %>%
+  mutate(`Date Time` = paste(as.character(Date), Time)) %>%
+  select(`Date Time`, WaterFlow)
+```
+
+After doing this I have `r nrow(pipe_1)` observations.  Not to convert to a time series object and plot it:
+
+```{r}
+pipe_1_ts <- ts(pipe_1$WaterFlow, start = c(2015,10,23), frequency = 24) %>%
+  tsclean
+
+is_stationary <- function(ts){
+  results <- kpss.test(ts)
+  if (results$p.value > 0.05){
+    "data IS stationary"
+  } else {
+    "data is NOT stationary"
+  }
+}
+
+ggtsplot <- function(ts, title) {
+  # A ggplot2 version of tsdisplay()
+  # Args:
+  #    ts (Time-Series): The time series we want to plot
+  #    title (str): The title of the graph
+  grid.arrange(
+    autoplot(ts) +
+      scale_y_continuous(labels = comma) +
+      ggtitle(paste0(title, " (", is_stationary(ts),")")) +
+      theme(axis.title = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank()),
+    grid.arrange(
+      ggAcf(ts) + ggtitle(element_blank()),
+      ggPacf(ts) + ggtitle(element_blank()), ncol = 2)
+    , nrow = 2)
+}
+
+ggtsplot(pipe_1_ts, "Pipe #1 Waterflow")
+```
+
+This time series falls within the ACF and the data is stationary.  With stationary data I am done with this data set and can focus on pipe #2.
+
+### Pipe 2
+
+I will clean up pipe 2's `Date Time` field by rounding it to the nearest hour.
+
+```{r}
+pipe_2 <- pipe_2 %>%
+  mutate(Date = as.Date(`Date Time`), 
+         Time = format(round(`Date Time`, units="hours"), format="%H:%M"),
+         `Date Time` = paste(as.character(Date), Time)) %>%
+  select(`Date Time`, WaterFlow)
+```
+
+Now to create the time series object and plot it
+
+```{r}
+pipe_2_ts <- ts(pipe_2$WaterFlow, start = c(2015, 10, 23, 1), frequency = 24) %>%
+  tsclean()
+ggtsplot(pipe_2_ts, "Pipe #2 Water Flow")
+```
+
+The data is stationary so we can begin forecasting.
+
+## Model Creation
+
+### A Note on Process
+
+I will develop a variety of models using different methods and will validate their performance using cross validation. The model that minimizes the error wil
+
+### Pipe 1 Candidates
+
+#### STL + ETS
+
+```{r}
+h <- 24 * 7
+pipe_1_stl_ets_fit <- pipe_1_ts %>%
+  stlf(h = h, s.window = 10, robust = TRUE, method = "ets")
+checkresiduals(pipe_1_stl_ets_fit)
+```
+
+#### STL + ARIMA
+
+```{r}
+pipe_1_stl_arima_fit <- pipe_1_ts %>%
+  stlf(h = h, s.window = 10, robust = TRUE, method = "arima")
+checkresiduals(pipe_1_stl_arima_fit)
+```
+
+#### Holt-Winters
+
+```{r}
+pipe_1_hw_fit <- hw(pipe_1_ts, h = h)
+checkresiduals(pipe_1_hw_fit)
+```
+
+#### Holt-Winters with Box Cox Adjustment
+
+```{r}
+pipe_1_lambda <- BoxCox.lambda(pipe_1_ts)
+pipe_1_adj_hw_fit <- hw(pipe_1_ts, h = h, lambda = pipe_1_lambda)
+checkresiduals(pipe_1_adj_hw_fit)
+```
+
+#### ARIMA
+
+```{r}
+pipe_1_arima_fit <- auto.arima(pipe_1_ts)
+checkresiduals(pipe_1_arima_fit)
+```
+
+#### Summary
+
+```{r}
+autoplot(pipe_1_ts) +
+  autolayer(pipe_1_stl_ets_fit, series = "STL ETS") +
+  autolayer(pipe_1_stl_arima_fit, series = "STL ARIMA") +
+  autolayer(pipe_1_hw_fit, series = "Holt-Winters") +
+  autolayer(pipe_1_adj_hw_fit, series = "Holt-Winters with Adj.") +
+  autolayer(forecast(pipe_1_arima_fit, h=h), series = "ARIMA") +
+  facet_wrap(.~series) +
+  scale_color_brewer(palette = "Set1") +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title = element_blank(), legend.position = "none")
+```
+
+### Pipe 2 Candidates
+
+#### STL + ETS
+
+```{r}
+pipe_2_stl_ets_fit <- pipe_2_ts %>%
+  stlf(h = h, s.window = 6, robust = TRUE, method = "ets")
+checkresiduals(pipe_2_stl_ets_fit)
+```
+
+#### STL + ARIMA
+
+```{r}
+pipe_2_stl_arima_fit <- pipe_2_ts %>%
+  stlf(h = h, s.window = 6, robust = TRUE, method = "arima")
+checkresiduals(pipe_2_stl_arima_fit)
+```
+
+#### Holt-Winters
+
+```{r}
+pipe_2_hw_fit <- hw(pipe_2_ts, h = h)
+checkresiduals(pipe_2_hw_fit)
+```
+
+#### Holt-Winters with Box Cox Adjustment
+
+```{r}
+pipe_2_lambda <- BoxCox.lambda(pipe_2_ts)
+pipe_2_adj_hw_fit <- hw(pipe_2_ts, h = h, lambda = pipe_2_lambda)
+checkresiduals(pipe_2_adj_hw_fit)
+```
+
+#### ARIMA
+
+```{r}
+pipe_2_arima_fit <- auto.arima(pipe_2_ts)
+checkresiduals(pipe_2_arima_fit)
+```
+
+#### Summary
+
+```{r}
+autoplot(pipe_2_ts) +
+  autolayer(pipe_2_stl_ets_fit, series = "STL ETS") +
+  autolayer(pipe_2_stl_arima_fit, series = "STL ARIMA") +
+  autolayer(pipe_2_hw_fit, series = "Holt-Winters") +
+  autolayer(pipe_2_adj_hw_fit, series = "Holt-Winters with Adj.") +
+  autolayer(forecast(pipe_2_arima_fit, h=h), series = "ARIMA") +
+  facet_wrap(. ~ series) +
+  scale_color_brewer(palette = "Set1") +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title = element_blank(), legend.position = "none")
+```
+
+## Cross Validation
+
+### Pipe 1 Models
+
+```{r pipe_1_cv}
+get_rmse <- function(e) {
+  sqrt(mean(e^2, na.rm = TRUE))
+}
+
+pipe_1_arima_forecast <- function(x, h) {
+  forecast(Arima(x), h = h)
+}
+
+stl_ets_errors <- tsCV(pipe_1_ts, stlf, h = h, s.window = 10, robust = TRUE, method = "ets")
+stl_arima_errors <- tsCV(pipe_1_ts, stlf, h = h, s.window = 10, robust = TRUE, method = "arima")
+hw_errors <- tsCV(pipe_1_ts, hw, h = h)
+adj_hw_errors <- tsCV(pipe_1_ts, hw, h = h, lambda = pipe_1_lambda)
+arima_errors <- tsCV(pipe_1_ts, pipe_1_arima_forecast, h = h)
+
+data.frame(Model = c("STL ETS", "STL ARIMA", "ARIMA", "Holt-Winters", "Adjusted Holt-Winters"),
+           RMSE = c(get_rmse(stl_ets_errors[, h]), get_rmse(stl_arima_errors[, h]), get_rmse(arima_errors[, h]), get_rmse(hw_errors[, h]), get_rmse(adj_hw_errors[, h]))) %>%
+  arrange(RMSE) %>%
+  kable() %>%
+  kable_styling()
+```
+
+### Pipe 2 Models
+
+```{r pipe_2_cv}
+pipe_2_arima_forecast <- function(x, h) {
+  forecast(Arima(x, seasonal = c(0,0,1)), h = h)
+}
+
+stl_ets_errors <- tsCV(pipe_2_ts, stlf, h = h, s.window = 6, robust = TRUE, method = "ets")
+stl_arima_errors <- tsCV(pipe_2_ts, stlf, h = h, s.window = 6, robust = TRUE, method = "arima")
+hw_errors <- tsCV(pipe_2_ts, hw, h = h)
+adj_hw_errors <- tsCV(pipe_2_ts, hw, h = h, lambda = pipe_2_lambda)
+arima_errors <- tsCV(pipe_2_ts, pipe_2_arima_forecast, h = h)
+
+data.frame(Model = c("STL ETS", "STL ARIMA", "ARIMA", "Holt-Winters", "Adjusted Holt-Winters"),
+           RMSE = c(get_rmse(stl_ets_errors[, h]), get_rmse(stl_arima_errors[, h]), get_rmse(arima_errors[, h]), get_rmse(hw_errors[, h]), get_rmse(adj_hw_errors[, h]))) %>%
+  arrange(RMSE) %>%
+  kable() %>%
+  kable_styling()
+```
+
+## Model Selection
+
+## Forecasts

### 2020-03-28 DATA624/Project 1-A ATM.Rmd
Finalize DATA 624 Midterm Files

@@
 library(fpp2)
-library(forcats)
+library(forecast)
 library(gridExtra)
 library(openxlsx)
+library(scales)
+knitr::knit_hooks$set(inline = function(x) {
+  if (!is.numeric(x)) {
+    x
+  } else {
+    prettyNum(round(x, 2), big.mark = ",")
+  }
+})
 options(scipen = 999)
@@
   scale_color_brewer(palette = "Set1") +
+  scale_y_continuous(labels = comma) +
   facet_wrap(.~ATM, ncol = 2) +
@@
       ggtitle(title) +
+      scale_y_continuous(labels = comma) +
       theme(axis.title = element_blank()),

### 2020-03-28 DATA624/Project 1-B Power.Rmd
Finalize DATA 624 Midterm Files

@@
 library(openxlsx)
+library(scales)
+knitr::knit_hooks$set(inline = function(x) {
+  if (!is.numeric(x)) {
+    x
+  } else {
+    prettyNum(round(x, 2), big.mark = ",")
+  }
+})
 options(scipen = 999)
@@
 autoplot(kwh, main = "Residential Power Usage (KWH)") +
+  scale_y_continuous(labels = comma) +
   theme(axis.title = element_blank())
@@
 ```{r}
-summary(kwh)
+summary(kwh) %>% kable() %>% kable_styling()
 ```
@@
     autoplot(ts) +
+      scale_y_continuous(labels = comma) +
       ggtitle(title) +
@@
   scale_color_brewer(palette = "Set1") +
+  scale_y_continuous(labels = comma) +
   theme(legend.position = "none", axis.title = element_blank())

### 2020-03-28 DATA624/Project 1-C Waterflow.Rmd
Finalize DATA 624 Midterm Files

--- 2020-03-03 DATA624/Project 1-C Waterflow.Rmd
@@
 ```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
-knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
 library(tidyverse)
+library(lubridate)
 library(kableExtra)
+library(readxl)
+library(fpp2)
+library(forcats)
+library(gridExtra)
+library(openxlsx)
+library(tseries)
+library(scales)
+knitr::knit_hooks$set(inline = function(x) {
+  if (!is.numeric(x)) {
+    x
+  } else {
+    prettyNum(round(x, 2), big.mark = ",")
+  }
+})
+options(scipen = 999)
 ```
 
-## Directions
+## Introduction
 
-Part C consists of two data sets.
-These are simple 2 columns sets, however they have different time stamps.
-Your optional assignment is to time-base sequence the data and aggregate based on hour (example of what this looks like, follows).
-Note for multiple recordings within an hour, take the mean.
-Then to determine if the data is stationary and can it be forecast.
-If so, provide a week forward forecast and present results via Rpubs and .rmd and the forecast in an Excel readable file.
+For this project I have data on water flows from two pipes.
+The data are captured with timestamps.
+I am to forecast a week forward.
+In the instructions for this project I am given the following hint: For multiple recordings within an hour, take the mean.
+I will be on the lookout for data under the hour.
+
+## Data Exploration
+
+I will begin by plotting the datasets:
+
+```{r}
+pipe_1 <- read_excel("Waterflow_Pipe1.xlsx") %>%
+  mutate(`Date Time` = as.POSIXct(`Date Time` * (60 * 60 * 24), origin = "1899-12-30", tz = "GMT"))
+
+pipe_2 <- read_excel("Waterflow_Pipe2.xlsx") %>%
+  mutate(`Date Time` = as.POSIXct(`Date Time` * (60 * 60 * 24), origin = "1899-12-30", tz = "GMT"))
+
+pipes <- pipe_1 %>%
+  mutate(Pipe = "Pipe 1") %>%
+  rbind(mutate(pipe_2, Pipe = "Pipe 2"))
+
+ggplot(pipes, aes(`Date Time`, WaterFlow, color = Pipe)) +
+  geom_line() +
+  scale_color_brewer(palette = "Set1") +
+  facet_wrap(. ~ Pipe) +
+  theme(legend.position = "none", axis.title = element_blank())
+```
+
+The two pipes are quite different in terms of the data they offer.
+Pipe 1 has a median waterflow of about `r round(median(pipe_1$WaterFlow))`, while pipe 2 has a median of `r round(median(pipe_2$WaterFlow))`.
+Pipe 1 also has what looks like over a week of data, while pipe 2 has over a month.
+Both datasets have 1000 observations.
+This suggests that pipe one has more fine grain observations (i.e.
+multiple observations in a hour or minute) while pipe 2's data captures a longer period of time.
+That would explain the difference in median and the length of time covered by the respective data sets.
+This also is hinted at in the instructions.
+
+### Closer Look at Pipe 1
+
+I will look at the head and tail of pipe 1's data to get a sense of what's included:
+
+```{r}
+head(pipe_1) %>% kable() %>% kable_styling()
+```
+
+There is multiple observations per minute which confirms my suspicions.  I will need to create the averages per minute as directed.
+
+### Closer Look at Pipe 2
+
+I will also take a look at pipe 2's data to see if there is anything I need to be aware of:
+
+```{r}
+head(pipe_2) %>% kable() %>% kable_styling()
+```
+
+Hmmm.  Some of the timestamps are just a bit off from the hour mark.  I will need to clean that up.
+
+## Data Cleanup
+
+### Pipe 1
+
+I will clean up the pipe 1 data by creating the averages per minute.
+
+```{r}
+pipe_1 <- pipe_1 %>%
+  mutate(Date = as.Date(`Date Time`),
+         Time = paste0(format(`Date Time`, "%H"), ":00:00")) %>%
+  group_by(Date, Time) %>%
+  summarise(WaterFlow = mean(WaterFlow)) %>%
+  ungroup() %>%
+  mutate(`Date Time` = as.POSIXct(paste(as.character(Date), Time)), format = "%Y-%m-%d %H:%M:%OS") %>%
+  select(`Date Time`, WaterFlow)
+```
+
+After doing this I have `r nrow(pipe_1)` observations.
+
+#### Sidebar: Is there any pattern in the usage by hour?
+
+I want to pause right here to see if there is any pattern in the the water flow by hour.
+
+```{r}
+pipe_1 %>%
+  mutate(Time = format(`Date Time`, "%H")) %>%
+  ggplot(aes(Time, WaterFlow)) +
+  geom_boxplot() +
+  ggtitle("Pipe #1 Water Flow by Hour") +
+  theme(axis.title = element_blank())
+```
+
+There is not a discernable pattern in the water flow by hour.  With that exploration done I will convert the data to a time series object and plot it:
+
+```{r}
+pipe_1_ts <- ts(pipe_1$WaterFlow, start = c(2015, 10, 23), frequency = 24)
+
+is_stationary <- function(ts) {
+  results <- kpss.test(ts)
+  if (results$p.value > 0.05) {
+    "data IS stationary"
+  } else {
+    "data is NOT stationary"
+  }
+}
+
+ggtsplot <- function(ts, title) {
+  # A ggplot2 version of tsdisplay()
+  # Args:
+  #    ts (Time-Series): The time series we want to plot
+  #    title (str): The title of the graph
+  grid.arrange(
+    autoplot(ts) +
+      scale_y_continuous(labels = comma) +
+      ggtitle(paste0(title, " (", is_stationary(ts), ")")) +
+      theme(axis.title = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank()),
+    grid.arrange(
+      ggAcf(ts) + ggtitle(element_blank()),
+      ggPacf(ts) + ggtitle(element_blank()), ncol = 2)
+    , nrow = 2)
+}
+
+ggtsplot(pipe_1_ts, "Pipe #1 Waterflow")
+```
+
+This time series falls within the ACF and the data is stationary.  With stationary data I am done with this data set and can focus on pipe #2.
+
+### Pipe 2
+
+I will clean up pipe 2's `Date Time` field by rounding it to the nearest hour.
+
+```{r}
+pipe_2 <- pipe_2 %>%
+  mutate(Date = as.Date(`Date Time`),
+         Time = format(round(`Date Time`, units = "hours"), format = "%H:%M"),
+         `Date Time` = as.POSIXct(paste(as.character(Date), Time)), format = "%Y-%m-%d %H:%M:%OS") %>%
+  select(`Date Time`, WaterFlow)
+```
+
+#### Sidebar: Is there any pattern in the usage by hour?
+
+I again want to see if there is a pattern in this pipe's usage by hour:
+
+```{r}
+pipe_2 %>%
+  mutate(Time = format(`Date Time`, "%H")) %>%
+  ggplot(aes(Time, WaterFlow)) +
+  geom_boxplot() +
+  ggtitle("Pipe #2 Water Flow by Hour") +
+  theme(axis.title = element_blank())
+```
+
+Again there is not a discernable pattern in the water flow by hour.  Now I will convert the data to a time series object and plot it:
+
+```{r}
+pipe_2_ts <- ts(pipe_2$WaterFlow, start = c(2015, 10, 23, 1), frequency = 24)
+ggtsplot(pipe_2_ts, "Pipe #2 Water Flow")
+```
+
+The data is stationary so we can begin forecasting.
+
+## Model Creation
+
+### A Note on Process
+
+I will develop a variety of models using different methods and will validate their performance using cross validation. The model that minimizes the error will be the model of choice.
+
+### Pipe 1 Candidates
+
+#### STL + ETS
+
+I will begin with a STL decomposition based model.
+
+```{r}
+h <- 24 * 7
+pipe_1_stl_ets_fit <- pipe_1_ts %>%
+  stlf(h = h, s.window = 24, robust = TRUE, method = "ets")
+checkresiduals(pipe_1_stl_ets_fit)
+```
+
+This model appears to have preformed fairly well.  The residuals are normal and there's only one minor ACF spike.  I'll see how this preforms in the cross validation stage.
+
+#### STL + ARIMA
+
+Next I will create an ARIMA model on the STL decomposed time series.
+
+```{r}
+pipe_1_stl_arima_fit <- pipe_1_ts %>%
+  stlf(h = h, s.window = 24, robust = TRUE, method = "arima")
+checkresiduals(pipe_1_stl_arima_fit)
+```
+
+That's strange.  It is a ARIMA(0,0,0) model.  That is a white-noise model.  Let's see if this is an artifact of the STL decomposition or if we get a similar result training the ARIMA model.
+
+#### ARIMA
+
+Last of all I will try an ARIMA model using the `auto.arima` function to define the (p,d,q) parameters.
+
+```{r}
+pipe_1_arima_fit <- auto.arima(pipe_1_ts)
+checkresiduals(pipe_1_arima_fit)
+```
+
+Again the best model was an ARIMA(0,0,0) model!  This suggests there is no reliable way to forecast the waterflow.
+
+### Pipe 2 Candidates
+
+#### STL + ARIMA
+
+I will first check with a STL decomposition and ARIMA to see if it results in a white noise model.
+
+```{r}
+pipe_2_stl_arima_fit <- pipe_2_ts %>%
+  stlf(h = h, s.window = 24, robust = TRUE, method = "arima")
+checkresiduals(pipe_2_stl_arima_fit)
+```
+
+It did.  Let's see what kind of model `auto.arima` comes up with.
+
+#### ARIMA
+
+```{r}
+pipe_2_arima_fit <- auto.arima(pipe_2_ts)
+checkresiduals(pipe_2_arima_fit)
+```
+
+We have another white noise model.  The water flow from pipe 2 cannot be forecasted reliably.
+
+## Summary
+
+White noise models were recommended when modeling the water flow of both pipes using the `auto.arima` function.  This suggests that there is not a reliable way to model the waterflow.

### 2020-04-04 DATA624/HW7.Rmd
Initial commit

---
title: "DATA 624 Homework 6"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(gridExtra)
options(scipen=999)
```

## Question 6.2

Developing a model to predict permeability (See Sect. 1.4) could save significant resources for a pharmaceutical company, while at the same time more rapidly identifying molecules that have a sufficient permeability to become a drug:

(A) Start `R` and use these commands to load the data:
```{r}
library(AppliedPredictiveModelings)
data(permeability)
```

## Question 6.3

### 2020-04-04 DATA624/HW7.Rmd
More work on HW7

@@
 ```{r}
-library(AppliedPredictiveModelings)
+library(AppliedPredictiveModeling)
 data(permeability)
@@
 
+The matrix `fingerprints` contains the 1,107 binary molecular predictors for the 165 compounds, while `permeability` contains permeability response.
+
+(B) The fingerprint predictors indicate the presense or absense of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure.
+Filter out the predictors that have low frequencieis using the `nearZeroVar` function from the `caret` package.
+How many are left for modeling?
+
+```{r}
+library(caret)
+length(nearZeroVar(fingerprints))
+```
+
+(C) Split the data into training and test sets, pre-process the data, and tune a PLS model.  How many latent variables are optimal and what is the corresponding resampled estimate of $R^2$?
+
+(D) Predict the response of the test set.  What is the test set estimate of $R^2$?
+
+(E) Try building other models discussed in this chapter.  Do any have better predictive performance?
+
+(F)  Would you recommend any of your models to replace the permeability laboratory experiment?
+
 ## Question 6.3
+
+A chemical manufacturing process for a pharmaceutical produce was discussed in Sect.
+1.4.
+In this problem, the objective is to understand the relationship between biological measurement of the raw materials (predictors), measurements of the manufacutring process (predictors), and the response of product yield.
+Biological predictors cannot be changed but can be used to assess the quality of the raw materials before processing.
+On the other hand, manufacturing process predictors can be changed in the manufacturing process.
+Improving product yield by 1% will boot revenue by approximately one hundred thousand dollars per batch:
+
+(A) Start `R` and use these commands to laod the data:
+
+```{r}
+library(AppliedPredictiveModeling)
+data(chemicalManufacturing)
+```
+
+the matrix `processPredictors` contains the 57 predictors (12 describing the input biological material and 45 describing the process predictors) for the 176 manufacturing runs, `yield` contains the percent yueld for each run.
+
+(B) A small percentage of cells in the predictor set contain missing values.  Use an imputation function to fill in these missing values (e.g., see Sect. 3.8).
+
+(C) Split the data into a training and a test set, pre-process the data, and tune a model of your choice from this chapter.  What is the optimal value of the performance metric?
+
+(D) Predict the response for the test set.  What is the value of the performance metric and how does this compare with the resampled performance metric on the training set?
+
+(E) Which predictors are most important in the model you have trained?  Do either the biological or process predictors dominate the list?
+
+(F) Explore the relationships between each of the top predictors and the response.  How could this information be helpful in improving yield in future runds of the manufacturing process?

### 2020-04-05 DATA624/HW7.Rmd
More work on HW7

@@
 ```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
 library(tidyverse)
+library(caret)
 library(kableExtra)
-library(gridExtra)
-options(scipen=999)
+library(glmnet)
+library(pls)
+knitr::knit_hooks$set(inline = function(x) {
+  if (!is.numeric(x)) {
+    x
+  } else {
+    prettyNum(round(x, 2), big.mark = ",")
+  }
+})
+options(scipen = 999)
+# Make this reproducible
+set.seed(42)
 ```
@@
 
-(A) Start `R` and use these commands to load the data:
+### Part A
+
+Start `R` and use these commands to load the data:
+
 ```{r}
@@
 
-(B) The fingerprint predictors indicate the presense or absense of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure.
+### Part B
+
+the `fingerprints` predictors indicate the presense or absense of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure.
 Filter out the predictors that have low frequencieis using the `nearZeroVar` function from the `caret` package.
@@
 
+**`r length(nearZeroVar(fingerprints))`**
+
+### Part C
+
+Split the data into training and test sets, pre-process the data, and tune a PLS model.  How many latent variables are optimal and what is the corresponding resampled estimate of $R^2$?
+
 ```{r}
-library(caret)
-length(nearZeroVar(fingerprints))
+df <- as.data.frame(fingerprints[, nearZeroVar(fingerprints)]) %>%
+  mutate(y = permeability)
+in_train <- createDataPartition(df$y, times = 1, p = .8, list = FALSE)
+train_df <- df[in_train, ]
+test_df <- df[-in_train, ]
+
+pls_model <- train(
+  y ~ ., data = train_df, method = "pls",
+  center = TRUE,
+  trControl = trainControl("cv", number = 10),
+  tuneLength = 25
+)
+
+# Plot model RMSE vs different values of components
+plot(pls_model, main=paste("Training Set RMSE Minimized at", pls_model$bestTune$ncomp, "Components"))
 ```
 
-(C) Split the data into training and test sets, pre-process the data, and tune a PLS model.  How many latent variables are optimal and what is the corresponding resampled estimate of $R^2$?
+```{r}
+pls_model$results %>%
+  filter(ncomp == pls_model$bestTune$ncomp) %>%
+  select(ncomp, RMSE, Rsquared) %>%
+  kable() %>%
+  kable_styling()
+```
 
-(D) Predict the response of the test set.  What is the test set estimate of $R^2$?
+**The optimal number of components in the PLS model is `r pls_model$bestTune$ncomp`.
+This captures `r round(pls_model$results$Rsquared[pls_model$bestTune$ncomp],2)*100`% of the variation in the permeability.**
 
-(E) Try building other models discussed in this chapter.  Do any have better predictive performance?
+### Part D
 
-(F)  Would you recommend any of your models to replace the permeability laboratory experiment?
+Predict the response of the test set.  What is the test set estimate of $R^2$?
+
+```{r}
+# Make predictions
+pls_predictions <- predict(pls_model, test_df)
+# Model performance metrics
+results <- data.frame(Model = "PLS", RMSE = caret::RMSE(pls_predictions, test_df$y), Rsquared = caret::R2(pls_predictions, test_df$y)) 
+results %>%
+  kable() %>%
+  kable_styling()
+```
+
+**The $R^2$ is `r round(caret::R2(pls_predictions, test_df$y), 2)`.**
+
+### Part E
+
+Try building other models discussed in this chapter.  Do any have better predictive performance?
+
+#### PCR Model
+
+```{r}
+pcr_model <- train(
+  y ~ ., data = train_df, method = "pcr",
+  center = TRUE,
+  trControl = trainControl("cv", number = 10),
+  tuneLength = 25
+)
+
+plot(pcr_model, main = paste("Training Set RMSE Minimized at", pcr_model$bestTune, "Components"))
+
+# Make predictions
+pcr_predictions <- predict(pcr_model, test_df)
+# Model performance metrics
+pcr_results <- data.frame(Model = "PCR", RMSE = caret::RMSE(pcr_predictions, test_df$y), Rsquared = caret::R2(pcr_predictions, test_df$y)) 
+pcr_results %>%
+  kable() %>%
+  kable_styling()
+```
+
+#### Ridge Regression
+
+```{r}
+x <- model.matrix(y ~ ., data=train_df)
+x_test <- model.matrix(y ~ ., data=test_df)
+rr_cv <- cv.glmnet(x, train_df$y, alpha = 0)
+rr_model <- glmnet(x, train_df$y, alpha = 0, lambda = rr_cv$lambda.min)
+rr_predictions <- as.vector(predict(rr_model, x_test))
+rr_results <- data.frame(Model = "Ridge Regression", RMSE = caret::RMSE(rr_predictions, test_df$y), Rsquared = caret::R2(rr_predictions, test_df$y)) 
+rr_results %>%
+  kable() %>%
+  kable_styling()
+```
+
+#### Lasso Regression
+
+```{r}
+lr_cv <- cv.glmnet(x, train_df$y, alpha = 1)
+lr_model <- glmnet(x, train_df$y, alpha = 1, lambda = lr_cv$lambda.min)
+lr_predictions <- as.vector(predict(lr_model, x_test)) 
+lr_results <- data.frame(Model = "Lasso Regression", RMSE = caret::RMSE(lr_predictions, test_df$y), Rsquared = caret::R2(lr_predictions, test_df$y))  
+lr_results %>%
+  kable() %>%
+  kable_styling()
+```
+
+#### Elastic Net Regession
+
+```{r}
+en_model <- train(
+  y ~ ., data = train_df, method = "glmnet",
+  trControl = trainControl("cv", number = 10),
+  tuneLength = 10
+)
+# Best tuning parameters
+en_model$bestTune
+```
+
+```{r}
+
+en_predictions <- en_model %>% predict(x_test)
+# Model performance metrics
+en_results <- data.frame(Model = "Elastic Net Regression", RMSE = caret::RMSE(en_predictions, test_df$y), Rsquared = caret::R2(en_predictions, test_df$y))
+en_results %>%
+  kable() %>%
+  kable_styling()
+```
+
+#### Summary
+
+```{r}
+pls_model$results %>%
+  filter(ncomp == pls_model$bestTune$ncomp) %>%
+  mutate("Model" = "PLS") %>%
+  select(Model, RMSE, Rsquared) %>%
+  as.data.frame() %>%
+  bind_rows(pcr_results) %>%
+  bind_rows(rr_results) %>%
+  bind_rows(lr_results) %>%
+  bind_rows(en_results) %>%
+  arrange(desc(Rsquared)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+### Part F
+
+Would you recommend any of your models to replace the permeability laboratory experiment?
 

### 2020-04-05 DATA624/HW7.Rmd
More work on HW 7

@@
     highlight: kate
-    toc_depth: 3
+    toc_depth: 4
     code_folding: "show"
@@
 ```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
+knitr::opts_chunk$set(echo = TRUE, comment = NA, 
+                      message = FALSE, warning = FALSE)
 library(tidyverse)
@@
 library(pls)
+library(corrplot)
+library(tidyr)
+library(RColorBrewer )
 knitr::knit_hooks$set(inline = function(x) {
@@
 options(scipen = 999)
-# Make this reproducible
-set.seed(42)
 ```
@@
   mutate(y = permeability)
-in_train <- createDataPartition(df$y, times = 1, p = .8, list = FALSE)
+
+# Make this reproducible
+set.seed(42)
+
+in_train <- createDataPartition(df$y, times = 1, p = 0.8, list = FALSE)
 train_df <- df[in_train, ]
@@
 # Plot model RMSE vs different values of components
-plot(pls_model, main=paste("Training Set RMSE Minimized at", pls_model$bestTune$ncomp, "Components"))
+title <- paste("Training Set RMSE Minimized at",
+               pls_model$bestTune$ncomp,
+               "Components")
+plot(pls_model, main = title)
 ```
@@
 # Model performance metrics
-results <- data.frame(Model = "PLS", RMSE = caret::RMSE(pls_predictions, test_df$y), Rsquared = caret::R2(pls_predictions, test_df$y)) 
+results <- data.frame(Model = "PLS",
+                      RMSE = caret::RMSE(pls_predictions, test_df$y),
+                      Rsquared = caret::R2(pls_predictions, test_df$y))
 results %>%
@@
 
-plot(pcr_model, main = paste("Training Set RMSE Minimized at", pcr_model$bestTune, "Components"))
+title <- paste("Training Set RMSE Minimized at",
+               pcr_model$bestTune,
+               "Components")
+plot(pcr_model, main = title)
 
@@
 # Model performance metrics
-pcr_results <- data.frame(Model = "PCR", RMSE = caret::RMSE(pcr_predictions, test_df$y), Rsquared = caret::R2(pcr_predictions, test_df$y)) 
+pcr_results <- data.frame(Model = "PCR",
+                          RMSE = caret::RMSE(pcr_predictions, test_df$y),
+                          Rsquared = caret::R2(pcr_predictions, test_df$y))
 pcr_results %>%
@@
 ```{r}
-x <- model.matrix(y ~ ., data=train_df)
-x_test <- model.matrix(y ~ ., data=test_df)
+x <- model.matrix(y ~ ., data = train_df)
+x_test <- model.matrix(y ~ ., data = test_df)
 rr_cv <- cv.glmnet(x, train_df$y, alpha = 0)
@@
 rr_predictions <- as.vector(predict(rr_model, x_test))
-rr_results <- data.frame(Model = "Ridge Regression", RMSE = caret::RMSE(rr_predictions, test_df$y), Rsquared = caret::R2(rr_predictions, test_df$y)) 
+rr_results <- data.frame(Model = "Ridge Regression",
+                         RMSE = caret::RMSE(rr_predictions, test_df$y),
+                         Rsquared = caret::R2(rr_predictions, test_df$y))
 rr_results %>%
@@
 lr_predictions <- as.vector(predict(lr_model, x_test))
-lr_results <- data.frame(Model = "Lasso Regression", RMSE = caret::RMSE(lr_predictions, test_df$y), Rsquared = caret::R2(lr_predictions, test_df$y))  
+lr_results <- data.frame(Model = "Lasso Regression",
+                         RMSE = caret::RMSE(lr_predictions, test_df$y),
+                         Rsquared = caret::R2(lr_predictions, test_df$y))
 lr_results %>%
@@
 # Model performance metrics
-en_results <- data.frame(Model = "Elastic Net Regression", RMSE = caret::RMSE(en_predictions, test_df$y), Rsquared = caret::R2(en_predictions, test_df$y))
+en_results <- data.frame(Model = "Elastic Net Regression",
+                         RMSE = caret::RMSE(en_predictions, test_df$y),
+                         Rsquared = caret::R2(en_predictions, test_df$y))
 en_results %>%
@@
 
+**None of the models did a better job (had a higher $R^2$) than the PLS model.**
+
 ### Part F
@@
 Would you recommend any of your models to replace the permeability laboratory experiment?
+
+**No, because the best model's (PLS) $R^2$ was really low.  The model did not have much explanatory power.**
 
@@
 
-(A) Start `R` and use these commands to laod the data:
+### Part A
+
+Start `R` and use these commands to load the data:
 
@@
 library(AppliedPredictiveModeling)
-data(chemicalManufacturing)
+data(ChemicalManufacturingProcess)
 ```
@@
 
-(B) A small percentage of cells in the predictor set contain missing values.  Use an imputation function to fill in these missing values (e.g., see Sect. 3.8).
+### Part B
 
-(C) Split the data into a training and a test set, pre-process the data, and tune a model of your choice from this chapter.  What is the optimal value of the performance metric?
+A small percentage of cells in the predictor set contain missing values.  Use an imputation function to fill in these missing values (e.g., see Sect. 3.8).
 
-(D) Predict the response for the test set.  What is the value of the performance metric and how does this compare with the resampled performance metric on the training set?
+**I will use KNN to impute values.  The `caret` package makes it simple:**
 
-(E) Which predictors are most important in the model you have trained?  Do either the biological or process predictors dominate the list?
+```{r}
+# Make this reproducible
+set.seed(42)
 
-(F) Explore the relationships between each of the top predictors and the response.  How could this information be helpful in improving yield in future runds of the manufacturing process?
+knn_model <- preProcess(ChemicalManufacturingProcess, "knnImpute")
+df <- predict(knn_model, ChemicalManufacturingProcess)
+```
+
+### Part C
+
+Split the data into a training and a test set, pre-process the data, and tune a model of your choice from this chapter.  What is the optimal value of the performance metric?
+
+**I will pre-process the data removing the zero variance variables prior to splitting the data.**
+
+```{r}
+df <- df %>%
+  select_at(vars(-one_of(nearZeroVar(., names = TRUE))))
+```
+
+**Now the data will be split into training and test set using an 80:20 split.**
+
+```{r}
+in_train <- createDataPartition(df$Yield, times = 1, p = 0.8, list = FALSE)
+train_df <- df[in_train, ]
+test_df <- df[-in_train, ]
+```
+
+**Given the preformance of the PLS in the previous exercise, I will use it again.  I'm going to further pre-process by centering and scaling the predictors.**
+
+```{r}
+pls_model <- train(
+  Yield ~ ., data = train_df, method = "pls",
+  center = TRUE,
+  scale = TRUE,
+  trControl = trainControl("cv", number = 10),
+  tuneLength = 25
+)
+
+# Plot model RMSE vs different values of components
+title <- paste("Training Set RMSE Minimized at",
+               pls_model$bestTune$ncomp,
+               "Components")
+plot(pls_model, main = title)
+```
+
+**For the performance metric I will again look at the RMSE and the $R^2$.
+The optimal values for RMSE would be zero, and the optimal $R^2$ would be 1.
+Here's the performance metrics for the optimal model on the training set.**
+
+```{r}
+pls_model$results %>%
+  filter(ncomp == pls_model$bestTune$ncomp) %>%
+  select(ncomp, RMSE, Rsquared) %>%
+  kable() %>%
+  kable_styling()
+```
+
+### Part D
+
+Predict the response for the test set.  What is the value of the performance metric and how does this compare with the resampled performance metric on the training set?
+
+```{r}
+# Make predictions
+pls_predictions <- predict(pls_model, test_df)
+# Model performance metrics
+results <- data.frame(RMSE = caret::RMSE(pls_predictions, test_df$Yield),
+           Rsquared = caret::R2(pls_predictions, test_df$Yield))
+results %>%
+  kable() %>%
+  kable_styling()
+```
+
+**We have an RMSE of about `r round(results$RMSE, 2)` and a $R^2$ of roughly `r round(results$Rsquared, 2)` on the test set.**
+
+### Part E
+
+Which predictors are most important in the model you have trained?  Do either the biological or process predictors dominate the list?
+
+```{r}
+varImp(pls_model) %>%
+  plot(, top = 20, main = "Important Variables (Cliped at 20)")
+```
+
+```{r}
+pls_importance <- varImp(pls_model)$importance %>% 
+  as.data.frame() %>%
+  rownames_to_column("Variable") %>%
+  arrange(desc(Overall)) %>%
+  mutate(importance = row_number())
+
+pls_importance %>%
+  mutate(Variable = gsub("[0-9]+", "", Variable)) %>%
+  group_by(Variable) %>%
+  tally() %>%
+  arrange(desc(n)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**The process predictors dominate the list.**
+
+### Part F
+
+Explore the relationships between each of the top predictors and the response.  How could this information be helpful in improving yield in future rounds of the manufacturing process?
+
+```{r}
+df %>%
+  select_at(vars(Yield, pls_importance$Variable)) %>%
+  cor() %>%
+  corrplot(type = "lower", order = "hclust", tl.col = "black", tl.srt = 45)
+  #corrplot(., type = "upper", order = "hclust",
+  #       col = brewer.pal(n = 8, name = "RdBu"))
+```
+  merge(pls_importance) %>%
+  ggplot(aes(value, Yield, color = Variable)) +
+  theme(legend.position = "none") +
+  geom_point() %>%
+  facet_wrap(~ Variable)
+```

### 2020-04-05 DATA624/HW7.Rmd
More work on HW 7

@@
 library(tidyr)
-library(RColorBrewer )
+library(RColorBrewer)
 knitr::knit_hooks$set(inline = function(x) {
@@
 options(scipen = 999)
+
+# Make sure to keep the default for normal processing.
+default_output_hook <- knitr::knit_hooks$get("output")
+
+# Output hooks handle normal R console output.
+knitr::knit_hooks$set(output = function(x, options) {
+
+  comment <- knitr::opts_current$get("comment")
+  if (is.na(comment)) comment <- ""
+  can_null <- grepl(paste0(comment, "\\s*\\[\\d?\\]"),
+                     x, perl = TRUE)
+  do_null <- isTRUE(knitr::opts_current$get("null_prefix"))
+  if (can_null && do_null) {
+    # By default R print output aligns at the right brace.
+    align_index <- regexpr("\\]", x)[1] - 1
+    # Two cases: start or newline
+    re <- paste0("^.{", align_index, "}\\]")
+    rep <- comment
+    x <- gsub(re, rep,  x)
+    re <- paste0("\\\n.{", align_index, "}\\]")
+    rep <- paste0("\n", comment)
+    x <- gsub(re, rep,  x)
+  }
+
+  default_output_hook(x, options)
+
+})
+
+knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
+
 ```
@@
 the `fingerprints` predictors indicate the presense or absense of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure.
-Filter out the predictors that have low frequencieis using the `nearZeroVar` function from the `caret` package.
+Filter out the predictors that have low frequencies using the `nearZeroVar` function from the `caret` package.
 How many are left for modeling?
 
-**`r length(nearZeroVar(fingerprints))`**
+```{r, null_prefix=TRUE}
+fingerprints %>%
+  nearZeroVar() %>%
+  length()
+```
+
+**There are `r length(nearZeroVar(fingerprints))` variables left.**
 
@@
 Split the data into training and test sets, pre-process the data, and tune a PLS model.  How many latent variables are optimal and what is the corresponding resampled estimate of $R^2$?
+
+**I will split it using a 80:20 split.  I will center the data.**
 
@@
 
+**No improvement here.**
+
 #### Ridge Regression
@@
 
+**PLS preformed better.**
+
 #### Lasso Regression
@@
 
+**PLS has more explanatory power.**
+
 #### Elastic Net Regession
@@
 ```{r}
-
 en_predictions <- en_model %>% predict(x_test)
@@
 ```
+
+**There isn't any improvment over the PLS.**
 
@@
 
-```{r}
-varImp(pls_model) %>%
-  plot(, top = 20, main = "Important Variables (Cliped at 20)")
-```
+**I will select the variables with a `varImp` sore greater than or equal to 60 to be the "important" ones.**
 
@@
   rownames_to_column("Variable") %>%
+  filter(Overall >= 50) %>%
   arrange(desc(Overall)) %>%
@@
 
+varImp(pls_model) %>%
+  plot(., top = max(pls_importance$importance), main = "Important Variables")
+```
+
+```{r}
 pls_importance %>%
@@
 
-**The process predictors dominate the list.**
+**There are only `r max(pls_importance$importance)` variables and the process predictors dominate the list.**
 
@@
 ```{r}
-df %>%
-  select_at(vars(Yield, pls_importance$Variable)) %>%
+important_vars <- df %>%
+  select_at(vars(Yield, pls_importance$Variable))
+
+important_vars_p <- cor.mtest(important_vars)$p
+
+important_vars %>%
   cor() %>%
-  corrplot(type = "lower", order = "hclust", tl.col = "black", tl.srt = 45)
-  #corrplot(., type = "upper", order = "hclust",
-  #       col = brewer.pal(n = 8, name = "RdBu"))
+  corrplot(method = "color", type = "lower", order = "hclust",
+           tl.cex = 0.8, tl.col = "black", tl.srt = 45,
+           addCoef.col = "black", number.cex = 0.7,
+           p.mat = important_vars_p,  sig.level = 0.05, diag = FALSE)
 ```
-  merge(pls_importance) %>%
-  ggplot(aes(value, Yield, color = Variable)) +
-  theme(legend.position = "none") +
-  geom_point() %>%
-  facet_wrap(~ Variable)
-```
+
+**All variables are positively correlated with the Yield.
+Manufacuring process 32 is the most important variable.
+It is positively correlated the biological variables.
+It is also negatively correlated with manufacturing process 13.**
+
+**Further exploration of these relationships could improve the yield.  This would be valuable information for these chemical manufacturers.**

### 2020-04-08 DATA624/HW8.Rmd
Fixed typo

---
title: "DATA 624 Homework 6"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 4
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(caret)
library(kableExtra)
library(glmnet)
library(pls)
library(corrplot)
library(tidyr)
library(RColorBrewer)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set(output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if (is.na(comment)) comment <- ""
  can_null <- grepl(paste0(comment, "\\s*\\[\\d?\\]"),
                     x, perl = TRUE)
  do_null <- isTRUE(knitr::opts_current$get("null_prefix"))
  if (can_null && do_null) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr("\\]", x)[1] - 1
    # Two cases: start or newline
    re <- paste0("^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub(re, rep,  x)
    re <- paste0("\\\n.{", align_index, "}\\]")
    rep <- paste0("\n", comment)
    x <- gsub(re, rep,  x)
  }

  default_output_hook(x, options)

})

knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))

```

## Question  7.2

Friedman (1991) introduced several benchmark data sets created by simulation.  On of these simulations used the following nonlinear equations to create data:

$y = 10 sin(\pi x_1x_2) + 20(x_3- 0.5)^2 + 10x_4 + 5x_5 + N(0, \sigma^2)$

where the $x$ values are randome variables uniformly distributed between $[0,1]$ (there are also 5 other non-informative variables also created in the simulation).  The package `mlbench` contains a function called `mlbench.friedman1` that simulates these data:

```{r}
library(mlbench)
set.seed(200)
trainingData <- mlbench.friedman1(200, sd = 1)
## We convert the 'x' data from a matrix to a data frame
## One reason is that this will five the columns names.
trainingData$x <- data.frame(trainingData$x)
## Look at the data using
featurePlot(trainingData$x, trainingData$y)
## or other methods.

## This creates a list with a vector 'y' and a matrix
## of predictors 'x'.  Also simulate a large test set to
## estimate the true error rate with good precision:
testData <- mlbench.friedman1(5000, sd = 1)
testData$x <- data.frame(testData$x)
```

Tune several models on these data.  For example:

```{r}
library(caret)
knnModel <- train(x = trainingData$x, 
                  y = trainingData$y,
                  method = "knn",
                  preProcess = c("center", "scale"),
                  tuneLength = 10)
knnModel
```

```{r}
knnPred <- predict(knnModel, newdata = testData$x)
## The function 'postResample' can be used to get the test set
## performance values
postResample(pred = knnPred, obs = testData$y)
```

Which models appear to give the best performance?  Does MARS select the infomrative predictors (those named X1-X5)?

## Question 7.5

Exercise 6.3 descrives data for a chemical manufacturing process.  Use the same data imputation, data splitting, and pre-processing steps as before and traing several nonlinear regression models.

```{r}
library(AppliedPredictiveModeling)
data(permeability)
df <- as.data.frame(fingerprints[, nearZeroVar(fingerprints)]) %>%
  mutate(y = permeability)

# Make this reproducible
set.seed(42)

in_train <- createDataPartition(df$y, times = 1, p = 0.8, list = FALSE)
train_df <- df[in_train, ]
test_df <- df[-in_train, ]

pls_model <- train(
  y ~ ., data = train_df, method = "pls",
  center = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)
```

### Part A

Which nonlinear regression model give the optimal resampling and test set performance?

### Part B

Which predictors are most important in the optimal nonlinear regression model?  Do either the biological or process variables dominate the list?  How do the top ten important predictors compare to the top ten predictors from the optimal linear model?

### Part C

Explore the relationships between the top predictors and the response for the predictors that are unique to the optimal nonlinear regression model. Do these plots reveal intuition about the biological or process predictors and theri relationship with yield?

### 2020-04-09 DATA624/HW8.Rmd
More work on HW 8
More work on HW8

@@
 ---
-title: "DATA 624 Homework 6"
+title: "DATA 624 Homework 8"
 date: "`r Sys.Date()`"
@@
 
-Which models appear to give the best performance?  Does MARS select the infomrative predictors (those named X1-X5)?
+Which models appear to give the best performance?  Does MARS select the informative predictors (those named X1-X5)?
 
+### MARS Model
+
+```{r mars_model}
+MARS_grid <- expand.grid(.degree = 1:2, .nprune = 2:15)
+MARS_model <- train(x = trainingData$x, 
+                  y = trainingData$y,
+                  method = "earth",
+                  tuneGrid = MARS_grid,
+                  preProcess = c("center", "scale"),
+                  tuneLength = 10)
+MARS_model
+```
+
+**The optimal MARS model minimized the RMSE when the nprune = `r MARS_model$bestTune$nprune`  and the degree = `r MARS_model$bestTune$degree`.**
+
+```{r}
+MARS_predictions <- predict(MARS_model, newdata = testData$x)
+postResample(pred = MARS_predictions, obs = testData$y)
+```
+
+**The RMSE of the MARS model is a lot lower than the KNN model.  Let's see what variables are important.**
+
+```{r}
+varImp(MARS_model)
+```
+
+**The MARS model picks up the X1-X5 varaibles.**
+
+### SVM Model
+
+```{r svm_model}
+SVM_model <- train(x = trainingData$x,
+                   y = trainingData$y,
+                   method = "svmRadial",
+                   preProcess = c("center", "scale"),
+                   tuneLength = 10,
+                   trControl = trainControl(method = "cv"))
+SVM_model
+```
+
+```{r}
+SVM_predictions <- predict(SVM_model, newdata = testData$x)
+postResample(pred = SVM_predictions, obs = testData$y)
+```
+
+### Summary
+
+```{r}
+results <- data.frame(t(postResample(pred = knnPred, obs = testData$y))) %>% 
+  mutate("Model" = "KNN")
+
+results <- data.frame(t(postResample(pred = MARS_predictions, obs = testData$y))) %>%
+  mutate("Model"= "MARS") %>%
+  bind_rows(results)
+
+results <- data.frame(t(postResample(pred = SVM_predictions, obs = testData$y))) %>%
+  mutate("Model"= "SVM") %>%
+  bind_rows(results)
+
+results %>%
+  select(Model, RMSE, Rsquared, MAE) %>%
+  arrange(RMSE) %>%
+  kable() %>%
+  kable_styling()
+```
 ## Question 7.5
 
-Exercise 6.3 descrives data for a chemical manufacturing process.  Use the same data imputation, data splitting, and pre-processing steps as before and traing several nonlinear regression models.
+Exercise 6.3 describes data for a chemical manufacturing process.  Use the same data imputation, data splitting, and pre-processing steps as before and traing several nonlinear regression models.
 
@@
 
-# Make this reproducible
 set.seed(42)

### 2020-04-09 DATA624/HW8.Rmd
More work on HW8

@@
 
-where the $x$ values are randome variables uniformly distributed between $[0,1]$ (there are also 5 other non-informative variables also created in the simulation).
+where the $x$ values are random variables uniformly distributed between $[0,1]$ (there are also 5 other non-informative variables also created in the simulation).
 The package `mlbench` contains a function called `mlbench.friedman1` that simulates these data:
@@
 
-```{r}
+### KNN Model
+
+```{r knn_model}
 library(caret)
@@
 
+**The optimal SVM mdoel has a $\sigma$ of `r SVM_model$bestTune$sigma` and an C of `r SVM_model$bestTune$C`.**
+
 ```{r}
@@
 ```
+
+```{r}
+varImp(SVM_model)
+```
+
+**The SVM picked up the important varaibles.**
+
+### Neural Network Model
+
+```{r}
+nnet_grid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)
+nnet_maxnwts <- 5 * (ncol(trainingData$x) + 1) + 5 + 1
+nnet_model <- train(x = trainingData$x,
+                    y = trainingData$y,
+                    method = "avNNet",
+                    preProcess = c("center", "scale"),
+                    tuneGrid = nnet_grid,
+                    trControl = trainControl(method = "cv"),
+                    linout = TRUE,
+                    trace = FALSE,
+                    MaxNWts = nnet_maxnwts,
+                    maxit = 500)
+nnet_model
+```
+
+**The best neural network has a size = `r nnet_model$bestTune$size` and a decay of `r nnet_model$bestTune$decay`.**
+
+```{r}
+nnet_predictions <- predict(nnet_model, newdata = testData$x)
+postResample(pred = nnet_predictions, obs = testData$y)
+```
+
+```{r}
+varImp(nnet_model)
+```
+
+**The top 5 variables are the ones we want to see listed.**
 
@@
 
+results <- data.frame(t(postResample(pred = nnet_predictions, obs = testData$y))) %>%
+  mutate("Model"= "Neural Network") %>%
+  bind_rows(results)
+
 results %>%
@@
 ```
+
+**The MARS model preformed the best and identified the right variables as the important ones.  The $R^2$ on it is extremely high.  I'm impressed by it's performance on the test set!**
+
 ## Question 7.5

### 2020-04-09 DATA624/HW8.Rmd
More work on HW 8

@@
 
-```{r}
+```{r neural_network_model}
 nnet_grid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)
@@
 )
+
+pls_predictions <- predict(pls_model, test_df)
+
+results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$y))) %>%
+  mutate("Model"= "PLS")
 ```
@@
 Which nonlinear regression model give the optimal resampling and test set performance?
+
+#### KNN Model
+
+```{r}
+knn_model <- train(y ~ ., data = train_df, method = "knn",
+                   preProcess = c("center"), 
+                   trControl = trainControl("cv", number = 10),
+                   tuneLength = 25)
+knn_model
+
+knn_predictions <- predict(knn_model, test_df)
+
+results <- data.frame(t(postResample(pred = knn_predictions, obs = test_df$y))) %>%
+  mutate("Model"= "KNN") %>% rbind(results)
+```
+
+#### MARS Model
+
+```{r}
+MARS_grid <- expand.grid(.degree = 1:2, .nprune = 2:15)
+MARS_model <- train(y ~ ., data = train_df, method = "earth",
+                    tuneGrid = MARS_grid,
+                    preProcess = c("center"),
+                    tuneLength = 25)
+MARS_model
+
+MARS_predictions <- predict(MARS_model, test_df)
+
+results <- data.frame(t(postResample(pred = MARS_predictions, obs = test_df$y))) %>%
+  mutate("Model"= "MARS") %>% rbind(results)
+```
+
+#### SVM Model
+
+```{r}
+SVM_model <- train(y ~ ., data = train_df, method = "svmRadial",
+                   preProcess = c("center"),
+                   tuneLength = 25,
+                   trControl = trainControl(method = "cv"))
+SVM_model
+
+SVM_predictions <- predict(SVM_model, test_df)
+
+results <- data.frame(t(postResample(pred = SVM_predictions, obs = test_df$y))) %>%
+  mutate("Model"= "SVM") %>% rbind(results)
+```
+
+#### Neural Network Model
+
+```{r}
+nnet_grid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)
+nnet_maxnwts <- 5 * ncol(train_df) + 5 + 1
+nnet_model <- train(y ~ ., data = train_df, method = "avNNet",
+                    preProcess = c("center"),
+                    tuneGrid = nnet_grid,
+                    trControl = trainControl(method = "cv"),
+                    linout = TRUE,
+                    trace = FALSE,
+                    MaxNWts = nnet_maxnwts,
+                    maxit = 500)
+nnet_model
+
+nnet_predictions <- predict(nnet_model, test_df)
+
+results <- data.frame(t(postResample(pred = nnet_predictions, obs = test_df$y))) %>%
+  mutate("Model"= "Neural Network") %>% rbind(results)
+```
+
+#### Summary
+
+```{r}
+results %>%
+  select(Model, RMSE, Rsquared, MAE) %>%
+  arrange(RMSE) %>%
+  kable() %>%
+  kable_styling()
+```
 

### 2020-04-10 DATA624/HW8.Rmd
More work on HW8

@@
 
+pls_model
+```
+
+```{r}
 pls_predictions <- predict(pls_model, test_df)
@@
 
-#### Neural Network Model
-
-```{r}
-nnet_grid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)
-nnet_maxnwts <- 5 * ncol(train_df) + 5 + 1
-nnet_model <- train(y ~ ., data = train_df, method = "avNNet",
-                    preProcess = c("center"),
-                    tuneGrid = nnet_grid,
-                    trControl = trainControl(method = "cv"),
-                    linout = TRUE,
-                    trace = FALSE,
-                    MaxNWts = nnet_maxnwts,
-                    maxit = 500)
-nnet_model
-
-nnet_predictions <- predict(nnet_model, test_df)
-
-results <- data.frame(t(postResample(pred = nnet_predictions, obs = test_df$y))) %>%
-  mutate("Model"= "Neural Network") %>% rbind(results)
-```
-
 #### Summary

### 2020-04-14 DATA624/HW8.Rmd
More work on HW 8

@@
 
+**I will run the data through the models in the chapter trying to keep each model as close to the original linear model in terms of the parameters.**
+
 ```{r}
 library(AppliedPredictiveModeling)
-data(permeability)
-df <- as.data.frame(fingerprints[, nearZeroVar(fingerprints)]) %>%
-  mutate(y = permeability)
+data(ChemicalManufacturingProcess)
 
+# Make this reproducible
 set.seed(42)
 
-in_train <- createDataPartition(df$y, times = 1, p = 0.8, list = FALSE)
+knn_model <- preProcess(ChemicalManufacturingProcess, "knnImpute")
+df <- predict(knn_model, ChemicalManufacturingProcess)
+
+df <- df %>%
+  select_at(vars(-one_of(nearZeroVar(., names = TRUE))))
+
+in_train <- createDataPartition(df$Yield, times = 1, p = 0.8, list = FALSE)
 train_df <- df[in_train, ]
@@
 pls_model <- train(
-  y ~ ., data = train_df, method = "pls",
+  Yield ~ ., data = train_df, method = "pls",
   center = TRUE,
+  scale = TRUE,
   trControl = trainControl("cv", number = 10),
@@
 pls_model
-```
 
-```{r}
 pls_predictions <- predict(pls_model, test_df)
 
-results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$y))) %>%
+results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$Yield))) %>%
   mutate("Model"= "PLS")
@@
 ```{r}
-knn_model <- train(y ~ ., data = train_df, method = "knn",
-                   preProcess = c("center"), 
+knn_model <- train(
+  Yield ~ ., data = train_df, method = "knn",
+  center = TRUE,
+  scale = TRUE,
   trControl = trainControl("cv", number = 10),
-                   tuneLength = 25)
+  tuneLength = 25
+)
 knn_model
@@
 
-results <- data.frame(t(postResample(pred = knn_predictions, obs = test_df$y))) %>%
+results <- data.frame(t(postResample(pred = knn_predictions, obs = test_df$Yield))) %>%
   mutate("Model"= "KNN") %>% rbind(results)
@@
 MARS_grid <- expand.grid(.degree = 1:2, .nprune = 2:15)
-MARS_model <- train(y ~ ., data = train_df, method = "earth",
+
+MARS_model <- train(
+  Yield ~ ., data = train_df, method = "earth",
   tuneGrid = MARS_grid,
-                    preProcess = c("center"),
-                    tuneLength = 25)
+  # If the following lines are uncommented, it throws an error
+  #center = TRUE,
+  #scale = TRUE,
+  trControl = trainControl("cv", number = 10),
+  tuneLength = 25
+)
 MARS_model
@@
 
-results <- data.frame(t(postResample(pred = MARS_predictions, obs = test_df$y))) %>%
+results <- data.frame(t(postResample(pred = MARS_predictions, obs = test_df$Yield))) %>%
   mutate("Model"= "MARS") %>% rbind(results)
@@
 ```{r}
-SVM_model <- train(y ~ ., data = train_df, method = "svmRadial",
-                   preProcess = c("center"),
-                   tuneLength = 25,
-                   trControl = trainControl(method = "cv"))
+SVM_model <- train(
+  Yield ~ ., data = train_df, method = "svmRadial",
+  center = TRUE,
+  scale = TRUE,
+  trControl = trainControl(method = "cv"),
+  tuneLength = 25
+)
 SVM_model
@@
 
-results <- data.frame(t(postResample(pred = SVM_predictions, obs = test_df$y))) %>%
+results <- data.frame(t(postResample(pred = SVM_predictions, obs = test_df$Yield))) %>%
   mutate("Model"= "SVM") %>% rbind(results)
+```
+
+#### Neural Network Model
+
+```{r}
+nnet_grid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)
+nnet_maxnwts <- 5 * ncol(train_df) + 5 + 1
+nnet_model <- train(
+  Yield ~ ., data = train_df, method = "avNNet",
+  center = TRUE,
+  scale = TRUE,
+  tuneGrid = nnet_grid,
+  trControl = trainControl(method = "cv"),
+  linout = TRUE,
+  trace = FALSE,
+  MaxNWts = nnet_maxnwts,
+  maxit = 500
+)
+
+nnet_model
+
+nnet_predictions <- predict(nnet_model, test_df)
+
+results <- data.frame(t(postResample(pred = nnet_predictions, obs = test_df$Yield))) %>%
+  mutate("Model"= "Neural Network") %>% rbind(results)
+
 ```
@@
 
+**The SVM Model was the best non-linear model.**
+
 ### Part B
@@
 
+```{r}
+varImp(SVM_model, 10)
+```
+
+```{r}
+varImp(pls_model, 10)
+```
+
+**The SVM model was very similar to the PLS model.
+In either case the manufacuturing process variables dominate the list.
+The SVM model found `BiologicalMaterial12` to be important and didn't find `BiologicalMaterial02` to be important.**
+
 ### Part C
@@
 Explore the relationships between the top predictors and the response for the predictors that are unique to the optimal nonlinear regression model.
-Do these plots reveal intuition about the biological or process predictors and theri relationship with yield?
+Do these plots reveal intuition about the biological or process predictors and their relationship with yield?
+
+```{r}
+ggplot(train_df, aes(BiologicalMaterial12, Yield)) +
+  geom_point()
+```
+
+**This indicates a positive relationship between the `BiologicalMaterial02` and the yeild, although it seems that there is a sweet spot, or a point of diminishing marginal returns on the material as the highest yeilds are not the furthest to the right on the graph.**

### 2020-04-15 DATA624/HW9.Rmd
Initial commit

--- 2020-04-10 DATA624/HW8.Rmd
@@
 ---
-title: "DATA 624 Homework 8"
+title: "DATA 624 Homework 9"
 date: "`r Sys.Date()`"
@@
 library(tidyverse)
-library(caret)
 library(kableExtra)
-library(glmnet)
-library(pls)
-library(corrplot)
 library(tidyr)
@@
 knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
-
 ```
 
-## Question  7.2
+## Question 8.1
 
-Friedman (1991) introduced several benchmark data sets created by simulation.  On of these simulations used the following nonlinear equations to create data:
+## Question 8.2
 
-$y = 10 sin(\pi x_1x_2) + 20(x_3- 0.5)^2 + 10x_4 + 5x_5 + N(0, \sigma^2)$
+## Question 8.3
 
-where the $x$ values are random variables uniformly distributed between $[0,1]$ (there are also 5 other non-informative variables also created in the simulation).
-The package `mlbench` contains a function called `mlbench.friedman1` that simulates these data:
-
-```{r}
-library(mlbench)
-set.seed(200)
-trainingData <- mlbench.friedman1(200, sd = 1)
-## We convert the 'x' data from a matrix to a data frame
-## One reason is that this will five the columns names.
-trainingData$x <- data.frame(trainingData$x)
-## Look at the data using
-featurePlot(trainingData$x, trainingData$y)
-## or other methods.
-
-## This creates a list with a vector 'y' and a matrix
-## of predictors 'x'.  Also simulate a large test set to
-## estimate the true error rate with good precision:
-testData <- mlbench.friedman1(5000, sd = 1)
-testData$x <- data.frame(testData$x)
-```
-
-Tune several models on these data.  For example:
-
-### KNN Model
-
-```{r knn_model}
-library(caret)
-knnModel <- train(x = trainingData$x, 
-                  y = trainingData$y,
-                  method = "knn",
-                  preProcess = c("center", "scale"),
-                  tuneLength = 10)
-knnModel
-```
-
-```{r}
-knnPred <- predict(knnModel, newdata = testData$x)
-## The function 'postResample' can be used to get the test set
-## performance values
-postResample(pred = knnPred, obs = testData$y)
-```
-
-Which models appear to give the best performance?  Does MARS select the informative predictors (those named X1-X5)?
-
-### MARS Model
-
-```{r mars_model}
-MARS_grid <- expand.grid(.degree = 1:2, .nprune = 2:15)
-MARS_model <- train(x = trainingData$x, 
-                  y = trainingData$y,
-                  method = "earth",
-                  tuneGrid = MARS_grid,
-                  preProcess = c("center", "scale"),
-                  tuneLength = 10)
-MARS_model
-```
-
-**The optimal MARS model minimized the RMSE when the nprune = `r MARS_model$bestTune$nprune`  and the degree = `r MARS_model$bestTune$degree`.**
-
-```{r}
-MARS_predictions <- predict(MARS_model, newdata = testData$x)
-postResample(pred = MARS_predictions, obs = testData$y)
-```
-
-**The RMSE of the MARS model is a lot lower than the KNN model.  Let's see what variables are important.**
-
-```{r}
-varImp(MARS_model)
-```
-
-**The MARS model picks up the X1-X5 varaibles.**
-
-### SVM Model
-
-```{r svm_model}
-SVM_model <- train(x = trainingData$x,
-                   y = trainingData$y,
-                   method = "svmRadial",
-                   preProcess = c("center", "scale"),
-                   tuneLength = 10,
-                   trControl = trainControl(method = "cv"))
-SVM_model
-```
-
-**The optimal SVM mdoel has a $\sigma$ of `r SVM_model$bestTune$sigma` and an C of `r SVM_model$bestTune$C`.**
-
-```{r}
-SVM_predictions <- predict(SVM_model, newdata = testData$x)
-postResample(pred = SVM_predictions, obs = testData$y)
-```
-
-```{r}
-varImp(SVM_model)
-```
-
-**The SVM picked up the important varaibles.**
-
-### Neural Network Model
-
-```{r neural_network_model}
-nnet_grid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)
-nnet_maxnwts <- 5 * (ncol(trainingData$x) + 1) + 5 + 1
-nnet_model <- train(x = trainingData$x,
-                    y = trainingData$y,
-                    method = "avNNet",
-                    preProcess = c("center", "scale"),
-                    tuneGrid = nnet_grid,
-                    trControl = trainControl(method = "cv"),
-                    linout = TRUE,
-                    trace = FALSE,
-                    MaxNWts = nnet_maxnwts,
-                    maxit = 500)
-nnet_model
-```
-
-**The best neural network has a size = `r nnet_model$bestTune$size` and a decay of `r nnet_model$bestTune$decay`.**
-
-```{r}
-nnet_predictions <- predict(nnet_model, newdata = testData$x)
-postResample(pred = nnet_predictions, obs = testData$y)
-```
-
-```{r}
-varImp(nnet_model)
-```
-
-**The top 5 variables are the ones we want to see listed.**
-
-### Summary
-
-```{r}
-results <- data.frame(t(postResample(pred = knnPred, obs = testData$y))) %>% 
-  mutate("Model" = "KNN")
-
-results <- data.frame(t(postResample(pred = MARS_predictions, obs = testData$y))) %>%
-  mutate("Model"= "MARS") %>%
-  bind_rows(results)
-
-results <- data.frame(t(postResample(pred = SVM_predictions, obs = testData$y))) %>%
-  mutate("Model"= "SVM") %>%
-  bind_rows(results)
-
-results <- data.frame(t(postResample(pred = nnet_predictions, obs = testData$y))) %>%
-  mutate("Model"= "Neural Network") %>%
-  bind_rows(results)
-
-results %>%
-  select(Model, RMSE, Rsquared, MAE) %>%
-  arrange(RMSE) %>%
-  kable() %>%
-  kable_styling()
-```
-
-**The MARS model preformed the best and identified the right variables as the important ones.  The $R^2$ on it is extremely high.  I'm impressed by it's performance on the test set!**
-
-## Question 7.5
-
-Exercise 6.3 describes data for a chemical manufacturing process.  Use the same data imputation, data splitting, and pre-processing steps as before and traing several nonlinear regression models.
-
-```{r}
-library(AppliedPredictiveModeling)
-data(permeability)
-df <- as.data.frame(fingerprints[, nearZeroVar(fingerprints)]) %>%
-  mutate(y = permeability)
-
-set.seed(42)
-
-in_train <- createDataPartition(df$y, times = 1, p = 0.8, list = FALSE)
-train_df <- df[in_train, ]
-test_df <- df[-in_train, ]
-
-pls_model <- train(
-  y ~ ., data = train_df, method = "pls",
-  center = TRUE,
-  trControl = trainControl("cv", number = 10),
-  tuneLength = 25
-)
-
-pls_model
-```
-
-```{r}
-pls_predictions <- predict(pls_model, test_df)
-
-results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$y))) %>%
-  mutate("Model"= "PLS")
-```
-
-### Part A
-
-Which nonlinear regression model give the optimal resampling and test set performance?
-
-#### KNN Model
-
-```{r}
-knn_model <- train(y ~ ., data = train_df, method = "knn",
-                   preProcess = c("center"), 
-                   trControl = trainControl("cv", number = 10),
-                   tuneLength = 25)
-knn_model
-
-knn_predictions <- predict(knn_model, test_df)
-
-results <- data.frame(t(postResample(pred = knn_predictions, obs = test_df$y))) %>%
-  mutate("Model"= "KNN") %>% rbind(results)
-```
-
-#### MARS Model
-
-```{r}
-MARS_grid <- expand.grid(.degree = 1:2, .nprune = 2:15)
-MARS_model <- train(y ~ ., data = train_df, method = "earth",
-                    tuneGrid = MARS_grid,
-                    preProcess = c("center"),
-                    tuneLength = 25)
-MARS_model
-
-MARS_predictions <- predict(MARS_model, test_df)
-
-results <- data.frame(t(postResample(pred = MARS_predictions, obs = test_df$y))) %>%
-  mutate("Model"= "MARS") %>% rbind(results)
-```
-
-#### SVM Model
-
-```{r}
-SVM_model <- train(y ~ ., data = train_df, method = "svmRadial",
-                   preProcess = c("center"),
-                   tuneLength = 25,
-                   trControl = trainControl(method = "cv"))
-SVM_model
-
-SVM_predictions <- predict(SVM_model, test_df)
-
-results <- data.frame(t(postResample(pred = SVM_predictions, obs = test_df$y))) %>%
-  mutate("Model"= "SVM") %>% rbind(results)
-```
-
-#### Summary
-
-```{r}
-results %>%
-  select(Model, RMSE, Rsquared, MAE) %>%
-  arrange(RMSE) %>%
-  kable() %>%
-  kable_styling()
-```
-
-### Part B
-
-Which predictors are most important in the optimal nonlinear regression model?
-Do either the biological or process variables dominate the list?
-How do the top ten important predictors compare to the top ten predictors from the optimal linear model?
-
-### Part C
-
-Explore the relationships between the top predictors and the response for the predictors that are unique to the optimal nonlinear regression model.
-Do these plots reveal intuition about the biological or process predictors and theri relationship with yield?
+## Question 8.7

### 2020-04-15 DATA624/HW9.Rmd
More work on Homework 9

@@
 
+Recreate the simulated data from Exercise 7.2
+
+```{r}
+library(mlbench)
+set.seed(200)
+simulated <- mlbench.friedman1(200, sd = 1)
+simulated <- cbind(simulated$x, simulated$y)
+simulated <- as.data.frame(simulated)
+colnames(simulated)[ncol(simulated)] <- "y"
+```
+
+### Part A
+
+Fit a random fores model to all the predictors, then estimate the variable importance scores:
+
+```{r}
+library(randomForest)
+library(caret)
+rf_model_1 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
+rf_imp_1 <- varImp(rf_model_1, scale = FALSE)
+rf_imp_1
+```
+
+Did the random forest model significantly use the uninformative predictors (V6-V10)?
+
+### Part B
+
+Now add an additional predictor that is highly correlated with one of the informative predictors.  For example:
+
+```{r}
+simulated$duplicate1 <- simulated$V1 + rnorm(200) * .1
+cor(simulated$duplicate1, simulated$V1)
+```
+
+Fit another random forest model to these data.  Did the importance score for V1 change?  What happens when you add another predictor that is also highly correlated with V1?
+
+```{r}
+rf_model_2 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
+rf_imp_2 <- varImp(rf_model_2, scale = FALSE)
+rm_imp_2
+```
+
+### Part C
+
+Use the `cforest` function in the `party` package to fit a random forest model using conditional inference trees.
+The `party` package function `varimp` can calculate predictor importance.
+The `conditional` argument of that function toggles between the traditional importance measure and the modified version described in Stobl et al.
+(2007).
+Do these importance show the same pattern as the traditional random forest model?
+
+```{r}
+library(party)
+rf_model_3 <- cforest(y ~ ., data = simulated)
+```
+
+### Part D
+
+Repeat this process with different tree models, such as boosted trees and Cubist.  Does the same pattern occur?
+
 ## Question 8.2
+
+Use a simulation to show tree bias with different granularities.
 
@@
 
+In stocastic gradient boosting the bagging fraction and learning rate will govern the construction of the trees as they are guided by the gradient.
+Although the optimal values of these parameters should be obtained through the tuning process, it is helpful to understand how the magnitudes of these parameters affect magnitudes of variable importance.
+Figure 8.24 provides the variable importance plots for boosting using two extreme values for the baggin fraction (0.1.
+and 0.9) and the learning rate (0.1 and 0.9) for the solubility data.
+The left-hand plot has both parameters set to 0.1, and the right-hand plot has both set to 0.9:
+
+### Part A
+
+Why does the model on the right focus its importance on just the first few of predictors, whereas the model on the left spreads importance across more predictors?
+
+### Part B
+
+Which model do you think would be more predictive of other samples?
+
+### Part C
+
+How would increasing interaction depth affect the slope of predictor importance for either model in Fig. 8,24?
+
 ## Question 8.7
+
+Refer to Exercises 6.3 and 7.5 which describe a chemical manufacturing process.  Use the same data imputation, data splitting, and pre-processing steps as before and train several tree-based models:
+
+### Part A
+
+Which tree-based regression model gives the optimal resampling and test set performance?
+
+### Part B
+
+Which predictors are most important in the optimal tree-based regression model?
+Do either the biological or process variables dominate the list?
+How do the top 10 important predictors compare to the top 10 predictors for the optimal linear and nonlinear models?
+
+### Part C
+
+Plot the optimal single tree with the distribution of yield in the terminal nodes.
+Does this view of the data provide additional knowledge about the biological or process predictors and their relationship with yield?

### 2020-04-15 DATA624/HW9.Rmd
More work on HW9

@@
 library(RColorBrewer)
+library(gbm)
+library(randomForest)
+library(caret)
+library(party)
+library(mlbench)
 knitr::knit_hooks$set(inline = function(x) {
@@
 ```{r}
-library(mlbench)
 set.seed(200)
@@
 ```{r}
-library(randomForest)
-library(caret)
 rf_model_1 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
 rf_imp_1 <- varImp(rf_model_1, scale = FALSE)
-rf_imp_1
+rf_imp_1  %>% 
+  as.data.frame() %>% 
+  rownames_to_column() %>% 
+  rename(Variable = rowname) %>%
+  arrange(desc(Overall)) %>%
+  kable() %>%
+  kable_styling()
 ```
@@
 Did the random forest model significantly use the uninformative predictors (V6-V10)?
+
+**No.  The scores are very close to zero indicating they are uninformative.**
 
@@
 
-```{r}
+```{r, null_prefix = TRUE}
 simulated$duplicate1 <- simulated$V1 + rnorm(200) * .1
@@
 rf_imp_2 <- varImp(rf_model_2, scale = FALSE)
-rm_imp_2
+rf_imp_2 %>% 
+  as.data.frame() %>% 
+  rownames_to_column() %>% 
+  rename(Variable = rowname) %>%
+  arrange(desc(Overall)) %>%
+  kable() %>%
+  kable_styling()
 ```
+
+**The importance of `v1` has decreased significantly since adding the additional correlated predictor.
+When adding the correlated predictor, both were included in the decision trees as significant predictors.
+This caused the overall significance to decrease.**
 
@@
 ```{r}
-library(party)
 rf_model_3 <- cforest(y ~ ., data = simulated)
+rf_imp_3 <- varImp(rf_model_3, conditional = TRUE)
+rf_imp_3 %>% 
+  as.data.frame() %>% 
+  rownames_to_column() %>% 
+  rename(Variable = rowname) %>%
+  arrange(desc(Overall)) %>%
+  kable() %>%
+  kable_styling()
 ```
@@
 Repeat this process with different tree models, such as boosted trees and Cubist.  Does the same pattern occur?
+
+#### Boosted Tree
+
+**Training it without the duplicate predictor:**
+
+```{r}
+gbm_grid <- expand.grid(.interaction.depth = seq(1, 7, by = 2),
+                        .n.trees = seq(100, 1000, by = 100),
+                        .shrinkage = c(0.01, 0.1),
+                        .n.minobsinnode = 10)
+gbm_model_1 <- train(y ~ ., data = select(simulated, -duplicate1), method="gbm", tuneGrid = gbm_grid, verbose = FALSE)
+
+gbm_imp_1 <- varImp(gbm_model_1)
+
+gbm_imp_1$importance %>% 
+  as.data.frame() %>% 
+  rownames_to_column() %>% 
+  rename(Variable = rowname) %>%
+  arrange(desc(Overall)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**The boosted tree did not pick up the uninformative predictors.  Now let's see what happens when I train it WITH the duplicate predictor:**
+
+```{r}
+gbm_model_2 <- train(y ~ ., data = simulated, method="gbm", tuneGrid = gbm_grid, verbose = FALSE)
+
+gbm_imp_2 <- varImp(gbm_model_2)
+
+gbm_imp_2$importance %>% 
+  as.data.frame() %>% 
+  rownames_to_column() %>% 
+  rename(Variable = rowname) %>%
+  arrange(desc(Overall)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**This model exhibits the same pattern.  The duplicate becomes one of the important variables and the importance of `v1` decreases.**
+
+#### Cubist
+
+**Again training it without the duplicate predictor:**
+
+```{r}
+cubist_model_1 <- train(y ~ ., data = select(simulated, -duplicate1), method="cubist")
+
+cubist_imp_1 <- varImp(cubist_model_1)
+
+cubist_imp_1$importance %>% 
+  as.data.frame() %>% 
+  rownames_to_column() %>% 
+  rename(Variable = rowname) %>%
+  arrange(desc(Overall)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**Once again, the cubist model does not pick up the un-important variables.  Let's train it WITH the duplicate predictor and see what happens:**
+
+```{r}
+cubist_model_2 <- train(y ~ ., data = simulated, method="cubist")
+
+cubist_imp_2 <- varImp(cubist_model_2)
+
+cubist_imp_2$importance %>% 
+  as.data.frame() %>% 
+  rownames_to_column() %>% 
+  rename(Variable = rowname) %>%
+  arrange(desc(Overall)) %>%
+  kable() %>%
+  kable_styling()
+```
+
+**The cubist model also exhibits a similar behavior as the random forest model, but is not as pronounced.**
 

### 2020-04-17 DATA624/HW9.Rmd
More work on HW9

@@
 library(party)
+library(rpart)
 library(mlbench)
@@
 
+***I will create a simulated dataset generated by a non-linear function.
+I will then train the tree based models with varying levels of pruning.
+I will then look at the MSE on the training set and test set in relation to the complexity of the tree.**
+
+```{r}
+set.seed(42)
+
+n_sample <- 350
+
+nonlinear_function <- function(x){
+  sin(1.25 * x) + 2 * cos(.25*x)
+}
+
+x <- runif(n_sample, 1, 25)
+f_of_x <- nonlinear_function(x)
+noise <- rnorm(n_sample, 0, 2)
+y <- f_of_x + noise
+
+df <- data.frame(y=y, x=x)
+in_train <- createDataPartition(df$y, p = .8, list = FALSE, times = 1)
+train_df <- df[in_train,]
+test_df <- df[-in_train,]
+
+results <- data.frame(Granularity = c(NA), MSE = c(NA), data = c(NA)) %>% na.omit()
+
+get_mse <- function(model, data){
+  y_hat <- predict(model, data)
+  mse <- mean((y_hat - data$y)^2)
+  return(mse)
+}
+
+for(depth in seq(1:10)){
+  rtree_model <- rpart(y ~ x, data = train_df, control=rpart.control(maxdepth=depth))
+  results <- rbind(results, data.frame(Granularity = depth, MSE = get_mse(rtree_model, train_df), data = "Training"))
+  results <- rbind(results, data.frame(Granularity = depth, MSE = get_mse(rtree_model, test_df), data = "Test"))
+}
+
+ggplot(results, aes(Granularity, MSE, color = data, group = data)) +
+  geom_line() +
+  geom_point() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom", legend.title = element_blank())
+```
+
+**One can see that as the granularity of the tree model increases, the MSE on the training set decreases.
+However the MSE on the test set initially begins to decline then increases again as the model starts to overfit the training data.**
+
 ## Question 8.3
@@
 
-How would increasing interaction depth affect the slope of predictor importance for either model in Fig. 8,24?
+How would increasing interaction depth affect the slope of predictor importance for either model in Fig. 8.24?
 

### 2020-04-20 DATA624/HW9.Rmd
More work on HW 9

@@
 
-```{r}
+```{r 8_1_A}
 rf_model_1 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
@@
 
-```{r}
+```{r 8_1_B}
 rf_model_2 <- randomForest(y ~ ., data = simulated, importance = TRUE, ntree = 1000)
@@
 
-```{r}
+```{r 8_1_C}
 rf_model_3 <- cforest(y ~ ., data = simulated)
@@
 
-```{r}
+```{r 8_1_D_gbm}
 gbm_grid <- expand.grid(.interaction.depth = seq(1, 7, by = 2),
@@
 
-```{r}
+```{r 8_1_D_cubist}
 cubist_model_1 <- train(y ~ ., data = select(simulated, -duplicate1), method="cubist")
@@
 
-***I will create a simulated dataset generated by a non-linear function.
+**I will create a simulated dataset generated by a non-linear function.
 I will then train the tree based models with varying levels of pruning.
@@
 
-```{r}
+```{r 8_2}
 set.seed(42)
@@
 Although the optimal values of these parameters should be obtained through the tuning process, it is helpful to understand how the magnitudes of these parameters affect magnitudes of variable importance.
-Figure 8.24 provides the variable importance plots for boosting using two extreme values for the baggin fraction (0.1.
+Figure 8.24 provides the variable importance plots for boosting using two extreme values for the bagging fraction (0.1.
 and 0.9) and the learning rate (0.1 and 0.9) for the solubility data.
 The left-hand plot has both parameters set to 0.1, and the right-hand plot has both set to 0.9:
+
+![Fig 8.24](fig_8_24.png)
 
@@
 
+**Because the learning rate is set to 0.1, the importance get's spread out over more predictors.  The higher learning rate will focus the importance on a smaller set of variables.**
+
 ### Part B
@@
 
+**The one on the left.  It will generalize while the one on the right will overfit the training data.  Always go for an ensemble of weak predictors.**
+
 ### Part C
@@
 How would increasing interaction depth affect the slope of predictor importance for either model in Fig. 8.24?
+
+**The predictor importance would get spread across more predictors as the interaction depth would increase.  The slope would decrease.**
 
@@
 
+```{r}
+library(AppliedPredictiveModeling)
+data(ChemicalManufacturingProcess)
+
+# Make this reproducible
+set.seed(42)
+
+knn_model <- preProcess(ChemicalManufacturingProcess, "knnImpute")
+df <- predict(knn_model, ChemicalManufacturingProcess)
+
+df <- df %>%
+  select_at(vars(-one_of(nearZeroVar(., names = TRUE))))
+
+in_train <- createDataPartition(df$Yield, times = 1, p = 0.8, list = FALSE)
+train_df <- df[in_train, ]
+test_df <- df[-in_train, ]
+
+pls_model <- train(
+  Yield ~ ., data = train_df, method = "pls",
+  center = TRUE,
+  scale = TRUE,
+  trControl = trainControl("cv", number = 10),
+  tuneLength = 25
+)
+
+pls_predictions <- predict(pls_model, test_df)
+
+pls_in_sample <- pls_model$results[pls_model$results$ncomp == pls_model$bestTune$ncomp,]
+results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$Yield))) %>%
+  mutate("In Sample RMSE" = pls_in_sample$RMSE,
+         "In Sample Rsquared" = pls_in_sample$Rsquared,
+         "In Sample MAE" = pls_in_sample$MAE,
+         "Model"= "PLS")
+```
+
+**I'm going to fit the data against a series of tree based models.  I will use caret with vanilla settings to see how well the preform against a tuned PLS model.**
+
+#### Boosted Tree
+
+```{r}
+gbm_model <- train(Yield ~ ., data = train_df, method="gbm", verbose = FALSE)
+gbm_predictions <- predict(gbm_model, test_df)
+gbm_in_sample <- merge(gbm_model$results, gbm_model$bestTune)
+results <- data.frame(t(postResample(pred = gbm_predictions, obs = test_df$Yield))) %>%
+  mutate("In Sample RMSE" = gbm_in_sample$RMSE,
+         "In Sample Rsquared" = gbm_in_sample$Rsquared,
+         "In Sample MAE" = gbm_in_sample$MAE,
+         "Model"= "Boosted Tree") %>%
+  rbind(results)
+```
+
+#### Random Forest
+
+```{r}
+rf_model <- train(Yield ~ ., data = train_df, method = "rf")
+rf_predictions <- predict(rf_model, test_df)
+rf_in_sample <- merge(rf_model$results, rf_model$bestTune)
+results <- data.frame(t(postResample(pred = rf_predictions, obs = test_df$Yield))) %>%
+  mutate("In Sample RMSE" = rf_in_sample$RMSE,
+         "In Sample Rsquared" = rf_in_sample$Rsquared,
+         "In Sample MAE" = rf_in_sample$MAE,
+         "Model"= "Random Forest") %>%
+  rbind(results)
+```
+
+#### Cubist Model
+
+```{r}
+cubist_model <- train(Yield ~ ., data = train_df, method="cubist")
+
+cubist_predictions <- predict(cubist_model, test_df)
+cubist_in_sample <- merge(cubist_model$results, cubist_model$bestTune)
+results <- data.frame(t(postResample(pred = cubist_predictions, obs = test_df$Yield))) %>%
+  mutate("In Sample RMSE" = cubist_in_sample$RMSE,
+         "In Sample Rsquared" = cubist_in_sample$Rsquared,
+         "In Sample MAE" = cubist_in_sample$MAE,
+         "Model"= "Cubist") %>%
+  rbind(results)
+```
+
+#### Results
+
+```{r}
+results %>%
+  kable() %>%
+  kable_styling()
+```
+
+**All tree model's fit the in sample data better than the PLS model.
+However they may have overfit the data as the RMSE on the test set is higher for the random forest and boosted tree.
+The cubist model is the only on that out preformed the PLS on the test set.
+These findings are on vanilla models.
+Additional tuning could be used to improve the model's performance**
+
 ### Part B
@@
 
+**Here's the top ten predictors:**
+
+```{r}
+varImp(cubist_model, 10)
+```
+
+**The manufacturing process variables continue to dominate the list.  Some of the variables on on both lists (in different orders of importance) while there are differences between the two lists:**
+
+```{r}
+varImp(pls_model, 10)
+```
+
 ### Part C

### 2020-04-21 DATA624/HW9.Rmd
More work on HW9

@@
 library(caret)
+library(rattle)
 library(party)
@@
 )
-
 pls_predictions <- predict(pls_model, test_df)
-
 pls_in_sample <- pls_model$results[pls_model$results$ncomp == pls_model$bestTune$ncomp,]
@@
          "Model"= "PLS")
+pls_model
 ```
 
-**I'm going to fit the data against a series of tree based models.  I will use caret with vanilla settings to see how well the preform against a tuned PLS model.**
+**I'm going to fit the data against a series of tree based models.  I will use caret and try to match the same paramaters to guague th preformance against the similar PLS model.**
 
+#### Bagged Tree
+
+```{r}
+set.seed(42)
+bagControl = bagControl(fit = ctreeBag$fit, predict = ctreeBag$pred, aggregate = ctreeBag$aggregate)
+bag_model <- train(Yield ~ ., data = train_df, method="bag", bagControl = bagControl,
+                   center = TRUE,
+                   scale = TRUE,
+                   trControl = trainControl("cv", number = 10),
+                   tuneLength = 25)
+bag_predictions <- predict(bag_model, test_df)
+bag_in_sample <- merge(bag_model$results, bag_model$bestTune)
+results <- data.frame(t(postResample(pred = bag_predictions, obs = test_df$Yield))) %>%
+  mutate("In Sample RMSE" = bag_in_sample$RMSE,
+         "In Sample Rsquared" = bag_in_sample$Rsquared,
+         "In Sample MAE" = bag_in_sample$MAE,
+         "Model"= "Bagged Tree") %>%
+  rbind(results)
+bag_model
+```
 #### Boosted Tree
@@
 ```{r}
-gbm_model <- train(Yield ~ ., data = train_df, method="gbm", verbose = FALSE)
+set.seed(42)
+gbm_model <- train(Yield ~ ., data = train_df, method="gbm", verbose = FALSE,
+                   trControl = trainControl("cv", number = 10),
+                   tuneLength = 25)
 gbm_predictions <- predict(gbm_model, test_df)
@@
   rbind(results)
+gbm_model
 ```
@@
 ```{r}
-rf_model <- train(Yield ~ ., data = train_df, method = "rf")
+set.seed(42)
+rf_model <- train(Yield ~ ., data = train_df, method = "ranger", 
+                  scale = TRUE,
+                  trControl = trainControl("cv", number = 10),
+                  tuneLength = 25)
 rf_predictions <- predict(rf_model, test_df)
@@
   rbind(results)
+rf_model
 ```
 
-#### Cubist Model
+#### Conditional Inference Random Forest
 
 ```{r}
-cubist_model <- train(Yield ~ ., data = train_df, method="cubist")
-
-cubist_predictions <- predict(cubist_model, test_df)
-cubist_in_sample <- merge(cubist_model$results, cubist_model$bestTune)
-results <- data.frame(t(postResample(pred = cubist_predictions, obs = test_df$Yield))) %>%
-  mutate("In Sample RMSE" = cubist_in_sample$RMSE,
-         "In Sample Rsquared" = cubist_in_sample$Rsquared,
-         "In Sample MAE" = cubist_in_sample$MAE,
-         "Model"= "Cubist") %>%
+set.seed(42)
+crf_model <- train(Yield ~ ., data = train_df, method = "cforest",
+                   trControl = trainControl("cv", number = 10),
+                   tuneLength = 25)
+crf_predictions <- predict(crf_model, test_df)
+crf_in_sample <- merge(crf_model$results, crf_model$bestTune)
+results <- data.frame(t(postResample(pred = crf_predictions, obs = test_df$Yield))) %>%
+  mutate("In Sample RMSE" = crf_in_sample$RMSE,
+         "In Sample Rsquared" = crf_in_sample$Rsquared,
+         "In Sample MAE" = crf_in_sample$MAE,
+         "Model"= "Conditional Random Forest") %>%
   rbind(results)
+crf_model
 ```
@@
 results %>%
+  arrange(RMSE) %>%
   kable() %>%
@@
 However they may have overfit the data as the RMSE on the test set is higher for the random forest and boosted tree.
-The cubist model is the only on that out preformed the PLS on the test set.
-These findings are on vanilla models.
-Additional tuning could be used to improve the model's performance**
+The boosted tree model is the only on that out preformed the PLS on the test set.**
 
@@
 ```{r}
-varImp(cubist_model, 10)
+varImp(bag_model)
 ```
 
-**The manufacturing process variables continue to dominate the list.  Some of the variables on on both lists (in different orders of importance) while there are differences between the two lists:**
+**The manufacturing process variables continue to dominate the list.  The same variables are found in both lists (in different orders of importance)**
 
 ```{r}
-varImp(pls_model, 10)
+varImp(pls_model)
 ```
@@
 Does this view of the data provide additional knowledge about the biological or process predictors and their relationship with yield?
+
+```{r}
+set.seed(1)
+cart_model <- train(Yield ~ ., data = train_df, method = "rpart",
+                    trControl = trainControl("cv", number = 10),
+                    tuneLength = 25)
+fancyRpartPlot(cart_model$finalModel, sub="")
+```
+
+**This indicates that if you want to maximize yield, then the manufacturing process 32 must be greater than or equal to 0.19, process 13 < -0.85 and the Biological material 3 >= 0.49.
+This recipie will produce the greatest yield.
+Caution should be used however as this model is likely overfitting the training data.**

### 2020-04-21 DATA624/HW9.Rmd
Finalizing HW 9

@@
 
-```{r}
+```{r 8_7_A_pls}
 library(AppliedPredictiveModeling)
@@
 
-```{r}
+```{r 8_7_A_bag}
 set.seed(42)
@@
 
-```{r}
+```{r 8_7_A_gbm}
 set.seed(42)
@@
 
-```{r}
+```{r 8_7_A_rf}
 set.seed(42)
@@
 
-```{r}
+```{r 8_7_A_crf}
 set.seed(42)
@@
 
-```{r}
+```{r 8_7_A_results}
 results %>%
@@
 
-```{r}
+```{r 8_7_B_bag}
 varImp(bag_model)
@@
 
-```{r}
+```{r 8_7_B_pls}
 varImp(pls_model)
@@
 
-```{r}
+```{r 8_7_C}
 set.seed(1)

### 2020-04-29 DATA624/HW10.Rmd
Initial commit

---
title: "DATA 624 Homework 10"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 4
    code_folding: "show"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE)
library(dplyr)
library(kableExtra)
library(tidyr)
library(igraph)
library(arules)
library(arulesViz)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set(output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if (is.na(comment)) comment <- ""
  can_null <- grepl(paste0(comment, "\\s*\\[\\d?\\]"),
                     x, perl = TRUE)
  do_null <- isTRUE(knitr::opts_current$get("null_prefix"))
  if (can_null && do_null) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr("\\]", x)[1] - 1
    # Two cases: start or newline
    re <- paste0("^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub(re, rep,  x)
    re <- paste0("\\\n.{", align_index, "}\\]")
    rep <- paste0("\n", comment)
    x <- gsub(re, rep,  x)
  }

  default_output_hook(x, options)

})

knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
```

## Introduction

The assignment is to use R to mine the data for association rules. It should report support, confidence and lift and your top 10 rules by lift.

## Market Basket Analysis

```{r}
transactions <- read.transactions("GroceryDataSet.csv", sep=",")
```

```{r}
itemFrequencyPlot(transactions, topN=10, type="absolute", main="Top 10 Items")
```

In order to complete this market basket analysis, I will use the Apriori algorithm.

```{r}
apriori(transactions, parameter=list(supp=0.001, conf=0.5) , control=list(verbose=FALSE)) %>%
  DATAFRAME() %>%
  arrange(desc(lift)) %>%
  top_n(10) %>%
  kable() %>%
  kable_styling()
```

## Cluster Analysis

If I am given the choice, I will use a network graph to preform the cluster analysis.

```{r}
temp <- read.csv("GroceryDataSet.csv", header = FALSE) %>%
  mutate(shoper_id = row_number()) %>%
  pivot_longer(-shoper_id) %>%
  filter(value != "") %>%
  select(-name)

edges <- temp %>%
  rename(to = value) %>%
  merge(temp) %>%
  rename(from = value) %>%
  filter(to != from) %>%
  select(-shoper_id)

G <- graph_from_data_frame(edges)
```

```{r}
newman_girvan_clusters <- cluster_edge_betweenness(G)
dendPlot(newman_girvan_clusters, mode="hclust")
```

```{r}
plot(newman_girvan_clusters, G) 
```

### 2020-04-29 DATA624/Project 2.Rmd
Initial commit

---
title: "DATA 624 Project 2"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidyr)
library(kableExtra)
library(caret)
library(scales)
library(readxl)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)
```

## Introduction

We are to develop a model that predicts the pH level of a beverage.

## Data Exploration

```{r}
df <- read_excel("StudentData.xlsx")
test <- read_excel("StudentEvaluation.xlsx")
```

### Missing Values

```{r}

```

### 2020-04-30 DATA624/HW10.Rmd
Finished HW 10

@@
 library(dplyr)
+library(ggplot2)
 library(kableExtra)
@@
 library(arules)
-library(arulesViz)
-knitr::knit_hooks$set(inline = function(x) {
-  if (!is.numeric(x)) {
-    x
-  } else {
-    prettyNum(round(x, 2), big.mark = ",")
-  }
-})
-options(scipen = 999)
-
-# Make sure to keep the default for normal processing.
-default_output_hook <- knitr::knit_hooks$get("output")
-
-# Output hooks handle normal R console output.
-knitr::knit_hooks$set(output = function(x, options) {
-
-  comment <- knitr::opts_current$get("comment")
-  if (is.na(comment)) comment <- ""
-  can_null <- grepl(paste0(comment, "\\s*\\[\\d?\\]"),
-                     x, perl = TRUE)
-  do_null <- isTRUE(knitr::opts_current$get("null_prefix"))
-  if (can_null && do_null) {
-    # By default R print output aligns at the right brace.
-    align_index <- regexpr("\\]", x)[1] - 1
-    # Two cases: start or newline
-    re <- paste0("^.{", align_index, "}\\]")
-    rep <- comment
-    x <- gsub(re, rep,  x)
-    re <- paste0("\\\n.{", align_index, "}\\]")
-    rep <- paste0("\n", comment)
-    x <- gsub(re, rep,  x)
-  }
-
-  default_output_hook(x, options)
-
-})
-
-knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
+library(stats)
 ```
@@
 
+First I will read in the data and explore it a bit by looking at the top 10 items purchased
+
 ```{r}
 transactions <- read.transactions("GroceryDataSet.csv", sep=",")
-```
-
-```{r}
 itemFrequencyPlot(transactions, topN=10, type="absolute", main="Top 10 Items")
@@
 
-In order to complete this market basket analysis, I will use the Apriori algorithm.
+Whole milk is the most frequently purchased item.
+
+In order to complete this market basket analysis, I will use the Apriori algorithm. I will print out the top 10 rules with their support, confidence and lift.
 
@@
 
-If I am given the choice, I will use a network graph to preform the cluster analysis.
+I am going to look for item groupings.
+I will use a network graph to preform the cluster analysis.
+First I will need to create a network graph from the transaction data.
+The I will detect the communities in the graph using the Louvain algorthym.
 
@@
 
-edges <- temp %>%
-  rename(to = value) %>%
-  merge(temp) %>%
-  rename(from = value) %>%
-  filter(to != from) %>%
-  select(-shoper_id)
-
-G <- graph_from_data_frame(edges)
+louvain_communities <- temp %>%
+  rename(to = value, from = shoper_id) %>%
+  graph_from_data_frame(directed = FALSE) %>%
+  cluster_louvain() %>%
+  communities()
 ```
 
-```{r}
-newman_girvan_clusters <- cluster_edge_betweenness(G)
-dendPlot(newman_girvan_clusters, mode="hclust")
-```
+Now that all customers and items are assigned to one of `r length(louvain_communities)` clusters I will summarize them by naming them the items and giving the count of the shopers in the cluster.
 
 ```{r}
-plot(newman_girvan_clusters, G) 
+items <- as.character(unique(temp$value))
+
+cluster_df <- data.frame(name = c(NA), members = c(NA)) %>% na.omit()
+
+for (i in 1:length(louvain_communities)){
+  cluster_name <- paste0(i,": ")
+  cluster_members <- 0
+  for (member in louvain_communities[[i]]){
+    if (member %in% items){
+      cluster_name <- paste0(cluster_name, member, " + ")
+    } else {
+      cluster_members <- cluster_members + 1
+    }
+  }
+  cluster_name <- substr(cluster_name,1,nchar(cluster_name)-3)
+  cluster_df <- rbind(cluster_df, data.frame(name = cluster_name, members = cluster_members))
+}
+
+cluster_df %>%
+  arrange(desc(members)) %>%
+  kable() %>%
+  kable_styling()
 ```

### 2020-05-07 DATA624/Project 2.Rmd
More work on Project 2

@@
 library(kableExtra)
+library(corrplot)
 library(caret)
@@
 library(readxl)
+library(naniar)
+library(xgboost)
+library(Matrix)
 knitr::knit_hooks$set(inline = function(x) {
@@
 options(scipen = 999)
+seed <- 42
 ```
@@
 
-```{r}
-df <- read_excel("StudentData.xlsx")
-test <- read_excel("StudentEvaluation.xlsx")
+```{r read_data}
+df <- read_excel("StudentData.xlsx") %>%
+  data.frame()
+
+eval_df <- read_excel("StudentEvaluation.xlsx") %>%
+  data.frame()
+```
+
+```{r summary_df}
+summary(df)
 ```
@@
 
-```{r}
+```{r missing_vals_df}
+df %>%
+  select(-PH) %>%
+  vis_miss()
+```
 
+```{r df_drop_na_ph}
+df <- df %>% 
+  drop_na(PH)
 ```
+
+```{r missing_vals_eval_df}
+eval_df %>%
+  select(-PH) %>%
+  vis_miss()
+```
+
+```{r knn_impute}
+set.seed(seed)
+impute_model <- df %>%
+  select(-PH) %>%
+  preProcess(., method = "knnImpute")
+
+df_imputed <- predict(impute_model, df) %>%
+  na.omit()
+
+eval_df_imputed <- predict(impute_model, eval_df)
+```
+
+## Near-Zero Variance
+
+```{r near_zero_variance}
+nearZeroVar(df_imputed, saveMetrics = TRUE) %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  filter(nzv == TRUE)
+```
+
+## Correlations
+
+```{r corrplot}
+temp <- df_imputed %>%
+  select(-Brand.Code) %>%
+  na.omit()
+
+p_mat <- cor.mtest(temp, conf.level = 0.95)$p
+
+df_imputed %>%
+  select(-Brand.Code) %>%
+  na.omit() %>%
+  cor() %>%
+  corrplot(p.mat = p_mat, sig.level = 0.05, insig = "blank", order = "hclust", type = "upper")
+```
+
+## Train/Test Split
+
+```{r train_test_split}
+set.seed(seed)
+in_train <- createDataPartition(df_imputed$PH, p = 0.8, list = FALSE, times = 1)
+train_df_imputed <- df_imputed[in_train,]
+test_df_imputed <- df_imputed[-in_train,]
+train_df <- df[in_train,]
+test_df <- df[-in_train,]
+```
+
+## Random Forest
+
+### Ranger
+
+```{r ranger_rf}
+set.seed(seed)
+tc <- trainControl(method = "cv", number = 10)
+rf_fit <- train(PH ~ ., data = train_df_imputed, method = "ranger", importance = "permutation", trControl = tc)
+rf_fit
+
+varImp(rf_fit)
+
+rf_predictions <- predict(rf_fit, test_df_imputed)
+postResample(pred = rf_predictions, obs = test_df_imputed$PH)
+```
+
+### RF
+
+```{r rf, cache=TRUE}
+set.seed(seed)
+rf_fit2 <- train(PH ~ ., data = train_df_imputed, method = "rf", trControl = tc)
+rf_fit2
+
+varImp(rf_fit2)
+
+rf2_predictions <- predict(rf_fit2, test_df_imputed)
+postResample(pred = rf2_predictions, obs = test_df_imputed$PH)
+```
+
+### Brand Code Random Forests
+
+```{r brand_ranger_rf}
+for (brand_code in unique(train_df_imputed$Brand.Code)){
+  print(paste("Brand Code", brand_code))
+  temp_df <- train_df_imputed %>%
+    filter(Brand.Code == brand_code) %>%
+    select(-Brand.Code)
+  set.seed(seed)
+  temp_rf <- train(PH ~ ., data = temp_df, method = "ranger", importance = "permutation", trControl = tc)
+  print(temp_rf)
+  print(varImp(temp_rf))
+  temp_test <- test_df_imputed %>%
+    filter(Brand.Code == brand_code) %>%
+    select(-Brand.Code) 
+  temp_predictions <- predict(temp_rf, temp_test)
+  print(postResample(pred = temp_predictions, obs = temp_test$PH))
+}
+```
+
+## XGBOOST
+
+```{r xgboost}
+#converting datasets to matrices
+#options(na.action="na.pass")
+training2 <- train_df_imputed %>% drop_na(Brand.Code)
+testing2 <- test_df_imputed %>% drop_na(Brand.Code)
+
+trainingmx <- model.matrix(~.+0,data=training2[,names(training2) != c("PH")])
+testingmx <- model.matrix(~.+0,data=testing2[,names(testing2) != c("PH")])
+
+trainingdmx <- xgb.DMatrix(data = trainingmx, label=training2$PH)
+testingdmx <- xgb.DMatrix(data = testingmx, label=testing2$PH)
+
+#default parameters
+params <- list(booster = "gbtree", objective = "reg:linear", eta=0.3, gamma=0, max_depth=6, min_child_weight=1, subsample=1, colsample_bytree=1)
+
+#determine the best nround parameter (It controls the maximum number of iterations. For classification, it is similar to the number of trees to grow.)
+xgbcv <- xgb.cv( params = params, data = trainingdmx, nrounds = 300, nfold = 5, showsd = T, stratified = T, print_every_n = 10, early_stop_rounds = 20, maximize = F) # best at 250 iterations
+
+xgb_model1 <- xgb.train (params = params, data = trainingdmx, nrounds = 250, watchlist = list(val=testingdmx,train=trainingdmx), print_every_n = 10, early_stop_round = 10, maximize = F)
+
+xgbpred <- predict(xgb_model1,testingdmx)
+
+mat <- xgb.importance (feature_names = colnames(trainingmx),model = xgb_model1)
+
+xgb.plot.importance (importance_matrix = mat)
+```

### 2020-05-10 DATA624/Project 2 Group Work.Rmd
More work on Group 2

---
title: "DATA 624 Final Project"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "show"
---

# R Libraries

```{r, echo = T, results = 'hide'}
library(caret)
library(tidyverse)
library(corrplot)
library(kableExtra)
library(mlbench)
library(dplyr)
library(readxl)
library(VIM)
library(psych)
library(xgboost)
library(Matrix)
library(writexl)
seed <- 12345
```

# Exploratory Data Analysis and PreProcessing

## Data Loading

```{r}
train <- read_excel("StudentData.xlsx")
test <- read_excel("StudentEvaluation.xlsx")
```

## Data Statistics

```{r}
dim(train)
dim(test)
```

#### Complete observations

```{r}
nrow(train[complete.cases(train),])
```

#### All predictors are numeric except for "Brand Code, 1st Variable)

```{r}
summary(train)
summary(test)
```

## Handling missing values (Train and Test sets)

```{r, warning=FALSE}
# Response
train <- train %>% drop_na(PH)

#Predictors
train <- kNN(train, imp_var=FALSE)
```

## Correlation Analysis

```{r}
# Analyze Correlations with the response variable

names <- colnames(train[,-26])
pairs.panels(train[, c("PH", names[1:8])])
pairs.panels(train[, c("PH", names[9:17])])
pairs.panels(train[, c("PH", names[18:26])])
pairs.panels(train[, c("PH", names[27:32])])
```

Top correlated features to PH:
- Mnf Flow (-0.45)
- Bowl Setpoint (0.35)
- Filler Level (0.32)
- Usage Cont (-0.32)
- Pressure Setpoint (-0.31)
- Hyd Pressure3 (-0.24)
- Pressure Vacuum (0.22)
- Hyd Pressure2 (-0.20)

```{r}
train2 <- train %>% dplyr::select(-'Brand Code')

mydata.cor = cor(train2, method = c("spearman"))
corrplot(mydata.cor,cl.cex = 0.7,tl.cex = .7,diag = TRUE)
```

## Correlated Predictors

```{r}
corr <- cor(train[,-c(1,26)], use='complete.obs')
topcorr <- findCorrelation(corr) #top correlated predictors that could be removed to improve modeling
colnames(train[,topcorr])
corrplot(cor(train[,topcorr], use='complete.obs'))
```

## Near Zero Variance Predictors

```{r}
nzv <- nearZeroVar(train)
colnames(train[,nzv])
```

"Hyd Pressure1" should be removed from the dataset as it is constant across observations

## Data Distribution & Variability

```{r}
boxplot(train[, c(names[2:6])])
boxplot(train[, c(names[7:11])])
boxplot(train[, c(names[12:15])])
boxplot(train[, c(names[16:20])])
boxplot(train[, c(names[21:26])])
boxplot(train[, c(names[27:32])])

############ Look at some Decision Trees ##################
source("https://raw.githubusercontent.com/crarnouts/Data_605_Final/master/RandomForestNulls_testing.R")
 colnames(train)<- make.names(colnames(train), unique=TRUE)
 colnames(test)<- make.names(colnames(test), unique=TRUE)
train <- as.data.frame(train)
test <- as.data.frame(test)

test <- RF_with_Nulls(train,test,"PH",.5,5,10,.01,5,1)
```

## Data Split

```{r}
set.seed(seed)
train_index <- createDataPartition(train$PH, p = .7, list = FALSE, times = 1)
training <- train[train_index,]
testing <- train[-train_index,]

# Preparing test set for validation
Xtest <- testing[,-grep("PH", colnames(testing))]
```

# Modeling & Evaluation

## Linear Regression

```{r}
set.seed(seed)
lm <- lm(PH~.,data = training)
summary(lm)
```

## Bagged Tree

```{r, warning=FALSE}
set.seed(seed)
bagControl = bagControl(fit = ctreeBag$fit, predict = ctreeBag$pred, aggregate = ctreeBag$aggregate)
bag_model <- train(PH ~., 
                    data = training, method="bag", bagControl = bagControl,
                   center = TRUE,
                   scale = TRUE,
                   trControl = trainControl("cv", number = 5),
                   tuneLength = 25)
bag_model

bag_pred <- predict(bag_model, newdata = Xtest)
postResample(obs = testing$PH, pred=bag_pred)

varImp(bag_model)
```

## XGBoost (Extreme Gradient Boosting)

```{r}
# Converting datasets to matrices
training2 <- training %>% drop_na(`Brand.Code`)
testing2 <- testing %>% drop_na(`Brand.Code`)
trainingmx<-model.matrix(~.+0,data=training2[,names(training2) != c("PH")])
testingmx<-model.matrix(~.+0,data=testing2[,names(testing2) != c("PH")])
trainingdmx <- xgb.DMatrix(data = trainingmx, label=training2$PH) 
testingdmx <- xgb.DMatrix(data = testingmx, label=testing2$PH) 
# Default parameters
params <- list(booster = "gbtree", objective = "reg:linear", eta=0.3, gamma=0, max_depth=6, min_child_weight=1, subsample=1, colsample_bytree=1)
# Determine the best nround parameter (It controls the maximum number of iterations. For classification, it is similar to the number of trees to grow.)
xgbcv <- xgb.cv( params = params, data = trainingdmx, nrounds = 300, nfold = 5, showsd = T, stratified = T, print_every_n = 10, early_stop_rounds = 20, maximize = F) 
# Best at 260 iterations:
```

```{r}
set.seed(seed)
xgb_model1 <- xgb.train (params = params, data = trainingdmx, nrounds = 260, watchlist = list(val=testingdmx,train=trainingdmx), print_every_n = 10, early_stop_round = 10, maximize = F)
# Best result at iteration 211 for training and validation sets:
```

```{r}
mat <- xgb.importance (feature_names = colnames(trainingmx),model = xgb_model1)
xgb.plot.importance (importance_matrix = mat) 
```

## SVM

```{r}
ctrl = trainControl(method='cv', number = 10)
set.seed(seed)
svmRad <- train(PH ~.,
                data=training,
                method = "svmRadial",
                preProc = c("center", "scale"),
                tuneLength = 14,
                trControl = ctrl)
svmRad
svm_pred <- predict(svmRad, newdata = Xtest)
postResample(obs = testing$PH, pred=svm_pred)
varImp(svmRad)
```

## Cubist

```{r}
set.seed(seed)
cubist <- train(PH ~., 
                data = training,
                method='cubist')
cubist  #display model performance
varImp(cubist) #display variable importance
cubist_pred <- predict(cubist, newdata=Xtest) # generate preds
postResample(obs=testing$PH, pred=cubist_pred) # evaluate model over test set
```

## Random Forest

```{r}
ctrl = trainControl(method='cv', number = 10, allowParallel = TRUE)
set.seed(seed)
rforest <- train(PH ~., 
                 data = training,
                 method = "ranger", 
                 importance = "permutation",
                 tuneLength = 10,
                 trControl = ctrl
                 )
rforest #display model performance
varImp(rforest) #display variable importance
rf_pred <- predict(rforest, newdata = Xtest) # generate preds
postResample(obs = testing$PH, pred=rf_pred) # Evaluate model over test set
```

#### Training Random Forest over Individual Brand Codes

```{r brand_ranger_rf}
for (brand_code in unique(training$Brand.Code)){
  print(paste("Brand Code", brand_code))

  temp_df <- training %>%
    filter(Brand.Code == brand_code) %>%
    select(-Brand.Code)
  set.seed(seed)
  temp_rf <- train(PH ~ ., data = temp_df, method = "ranger", importance = "permutation", trControl = ctrl)
  print(temp_rf)
  print(varImp(temp_rf))
  temp_test <- testing %>%
    filter(Brand.Code == brand_code) %>%
    select(-Brand.Code)
  temp_predictions <- predict(temp_rf, temp_test)
  print(postResample(pred = temp_predictions, obs = temp_test$PH))
}
```

# Predicting New Data w/Random Forest

```{r, warning=FALSE}
pfile <- read_excel("StudentEvaluation.xlsx")
#Preparing the dataset we will ultimately predict PH on
test <- pfile[,-grep("PH", colnames(pfile))]
test <- kNN(test, imp_var=FALSE)
colnames(test)<- make.names(colnames(test), unique=TRUE)

ctrl = trainControl(method='cv', number = 10)
set.seed(seed)
rf_model <- train(PH ~., 
                 data = train,
                 method = "ranger", 
                 importance = "permutation",
                 tuneLength = 10,
                 trControl = ctrl
                 )

final_rf_pred <- predict(rf_model, newdata=as.data.frame(test))
pfile$PH <- final_rf_pred # applying predictions to unimputed dataset

write_xlsx(pfile, "Predictions_file.xlsx") # write to excel
```

### 2020-05-10 DATA624/Project 2 Personal Work.Rmd
More work on Group 2

--- 2020-05-07 DATA624/Project 2.Rmd
@@
 library(Matrix)
+library(VIM)
 knitr::knit_hooks$set(inline = function(x) {
@@
 options(scipen = 999)
-seed <- 42
+seed <- 12345
 ```
@@
 set.seed(seed)
-impute_model <- df %>%
-  select(-PH) %>%
-  preProcess(., method = "knnImpute")
 
-df_imputed <- predict(impute_model, df) %>%
-  na.omit()
-
-eval_df_imputed <- predict(impute_model, eval_df)
+df_imputed <- kNN(df, imp_var = FALSE)
+eval_df_imputed <- kNN(select(eval_df, -PH), imp_var = FALSE)
 ```
@@
 set.seed(seed)
-tc <- trainControl(method = "cv", number = 10)
+tc <- trainControl(method = "cv", number = 10, allowParallel = TRUE)
 rf_fit <- train(PH ~ ., data = train_df_imputed, method = "ranger", importance = "permutation", trControl = tc)
@@
 
-```{r rf, cache=TRUE}
+```{r rf}
 set.seed(seed)
@@
 ```
+
+## Stepwise Linear Regression
+
+```{r stepwise}
+# Loading latter to avoid the select clash with dplyr
+library(MASS)
+full_model <- lm(PH ~., data = train_df_imputed)
+# Stepwise regression model
+lm_model <- stepAIC(full_model, direction = "both", trace = FALSE)
+summary(lm_model)
+```
+
+```{r}
+set.seed(seed)
+tc <- trainControl(method = "cv", number = 10, allowParallel = TRUE)
+lm_fit <- train(PH ~ ., data = train_df_imputed, method = "glmStepAIC", trControl = tc)
+lm_fit
+
+varImp(lm_fit)
+
+lm_predictions <- predict(lm_fit, test_df_imputed)
+postResample(pred = lm_predictions, obs = test_df_imputed$PH)
+```

### 2020-09-15 DATA622/M02-glm.rmd
Initial commit

---
title: "Chapter 02: Our first model,Grand Old Logistic Regression"
output:
html_document: M02-glm.html
df_print: paged
pdf_document: default
---
```{r Startup}

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

We agree to maintain all the data in
/home/data/ML.Data directory in 12.42.205.8 and 9. 
If you wish to run these in your local then create a directory. 
I created c:\Users\rkannan\rk\data 
and copied these csv files.
We assume data will be in csv format.

#Load input data heart.csv and ecoli.csv
```{r MainCode}
unixpath<-"/home/data/ML.Data/heart.csv"
unixecolipath<-"/home/data/ML.Data/ecoli.csv"
path<-"c://Users/rkannan/rk/data/heart.csv"
ecolipath<-"c://Users/rkannan/rk/data/ecoli.csv"
heart<-read.csv(path,header=TRUE,sep=',',stringsAsFactors=FALSE)
ecoli<-read.csv(ecolipath,head=T,sep=',',stringsAsFactors=F)
```
Let us write a one-line utility function to read in a CSV as above
We assume, comma separated and includes a header record.
```{r}
loadData<-function(csvfile) { read.csv(csvfile,head=T,sep=',',stringsAsFactors=F) }
ecoli_1<-ecoli
ecoli_2<-loadData(ecolipath)
TBL=table(ecoli_1==ecoli_2)
TBL
```
## Review the data, quick and easy EDA
View a few records (head/tail), dimension, number of rows and records
and the variable names.
Everything is an object and mostly a list. We can index into a list to 
access the elements in the list. R index runs from 1 to the length.
```{r}
head(heart)
```
### unix head and tail commands
R commands *head* and *tail* are like the Unix head and tail shell commands
These commands output a few records from the beginning or at the end of the file.
```{r}
tail(heart)
```
### What is the shape of this dataset
How many  observations does it have (each row is an observation)
How many variables are available to us -- number of columns

```{r}
dim(heart)
```
*dim* prints the row X column -- equivalent to *shape in python*
```{r}
nrow(heart)
```
we can get just the number of rows in a data.frame
or just the number of columns
```{r}
ncol(heart)
```
### Manipulating column names 
Recall **heart<-read.csv(path,*header=TRUE*,sep=',',stringsAsFactors=FALSE)**
so R would load the data into a data.frame with column names
Here are some ways to manipulate header
```{r}
names(heart)
names(heart)[3]
length(names(heart))
```
let us change the name of the first column

```{r}
names(heart)[[1]]<-'age'
```
Always trust but verify that our indexing worked...
```{r}
names(heart)
```
# Are the covariates correlated?
Numerically review the pairwise correlation amongst independent variables
```{r}
(corheart<-cor(heart[,1:13]))
```
# Are there any constant feature vectors?
Let us find out if there are any constant variables.
Constant variables cannot possibly contribute to predicting the target.
For example, given this simple birds dataset, and that I saw a bird with two legs,
can you identify which bird that might be?
```{r}
birds<-data.frame(legs=rep(2,8),name=c('crow','pigeon','owl','eagle','sparrow',
'duck', 'woodpecker', 'humming bird'),stringsAsFactors=F)
birds
```
Let us write a utility function we can call on any list...
we can also do length(unique(x)) < 1
```{r}
isConstant<-function(x) length(names(table(x)))<2
apply(heart,2,isConstant)
```
Let us plot,correlations using pairs function
```{r}
library(plotly)
p <- pairs(heart)
```
## What is our class/label variable?
Is the problem at hand *binary classification* or *multi-class classification!*?
Let us assume, our customers want us to predict **target** given the rest.
```{r}
classLabels<-table(heart$target) # this is the actual class distribution both test/train partitions must be close to this.
print(classLabels)
names(classLabels)
```
# Class Imbalance
Let us try to understand the *distribution of the target variable.* Let us make sure
the distribution is balanced. 
### When is a class imbalanced?
If we find 1(minority) in 10 in the extreme case
or anything less than 1 (minority) in 6, we will consider our dataset to be imbalanced.
These are not hard rules but my intuition is as follows
if we are going to create training data with 70/30 split then a random 
sample that contains none or very few of the minority class becomes likely.
Our efforts to train a model is not going to be very fruitful. The model
may not know how to detect minority class or overwhelmed by majority class signature.

```{r}
length(names(classLabels))
ifelse(length(names(classLabels))==2,"binary classification", "multi-class classification")

```
CONSIDER WRITING SMALL ONE LINERS isBinaryClassification(classVector)
and one for, class imbalanced, isImbalancedData(classVector)
IMHO, only one[or two] instance of a certain class -- better to remove that class initially.

## First Classification run
Let us run a simple generalized linear model and try to
understand each variable and how important they are in 
determining the target variable.

we will **NOT** split the data as this is a trial run
and this run is to develop a rough draft script for
any classification exercise.
```{r}
glm_model<-glm(target~.,data=heart, family='binomial')
glm_model
summary_glm_model<-summary(glm_model)
coef_summary_glm_model<-coef(summary_glm_model)
dim(coef_summary_glm_model)
coef_summary_glm_model[[1]]
row.names(coef_summary_glm_model)
```
We will reject any variable with a p-value greater than 0.05.

# p-value
There is much debate regarding p-value. Nothing prevents us from saying
we will reject if p-value is greater than 0.07 after we run the model.

While I respect that viewpoint, I continue to use
p-value as I use 0.05 always. Sometimes, I do experiment with variables
that are indicated as not-significant (p-value > 0.05 for that variable).
For the overall model significance I consistently use 0.05. I do not change that
to accept unreliable model.
I interpret p-value to be the probability of rejecting H0 (the hypothesis) when it is
actually, valid/true.

The p value is the probability of observing a sample value as extreme as, or more extreme than, the value actually observed, given that the null hypothesis is true.
p value represents the risk of rejecting a true null hypothesis. It is the probability of a Type I error if the null hypothesis is rejected.

# Type I and Type II error
When a Type I error (a) is committed, a true null hypothesis is rejected; the innocent person is unjustly convicted. The value is called the level of significance and is the probability of rejecting the true null. With
a Type II error (b), one fails to reject a false null hypothesis; the result is an unjust acquittal, with the guilty person going free. In our system of justice, it is more important to reduce the probability of convicting
the innocent than that of acquitting the guilty. Similarly, hypothesis testing places a greater emphasis
on Type I errors than on Type II errors.

# Ethical Practice
Statistical tools cannot enforce ethical science or human behavior.
So when running  a ML experiment, write ahead of time what your 
acceptance criteria is prior to running the experiment. Include the result
in your reporting regardless of the results.

# File Drawer Problem
Include all results from all the experiments, regardless of the results whether it is in agreement with your expectation or NOT. Eliminate the so called File Drawer problem -- of not including unfavorable results --.
```{r}
coef_summary_glm_model
coef_summary_glm_model[,4]<0.05
row_names<-row.names(coef_summary_glm_model[coef_summary_glm_model[,4]<0.05,])
row_names
summary_glm_model$aic
summary_glm_model$null.deviance
summary_glm_model$deviance
ifelse(summary_glm_model$deviance<summary_glm_model$null.deviance,"model has improved","model has not helped")
```
### p-values
Now let us rerun the model with the signifcant variables, that is, with p-value <0.05
```{r}
formulastr<-paste('target~',paste(row.names(coef_summary_glm_model[coef_summary_glm_model[,4]<0.05,]),collapse='+'),sep='')
formulastr
paste(row.names(coef_summary_glm_model[coef_summary_glm_model[,4]<0.05,]),collapse='+')
model2<-glm(formulastr,data=heart,family='binomial')
summ_model2<-summary(model2)
coef_summ_model2<-coef(summ_model2)
coef_summ_model2
summ_model2$aic
summ_model2$null.deviance
summ_model2$deviance
summary_glm_model$deviance
ifelse(summary_glm_model$deviance<summary_glm_model$null.deviance,"model has improved","model has not helped")
```
So we removed predictors based on p-value and our deviance of the
revised model is higher by about 6%.
So let us compute the VIF factors for the model predictors and 
determine if we removed some variables that should not have been removed.
```{r}
if(!require(car))install.packages('car')
library(car)
vif_model<-vif(glm_model)
```
VIF close to 1 indicates the variables are not correlated
Above 4 is highly correlated.
```{r}
vif_model
vif_model[vif_model>4]
names(vif_model)
nl<-names(vif_model[vif_model<4])
(newformulastr<-paste('target~',paste(nl,collapse='+')))
```
We learned from the entire dataset. Let us get the target variable into a vector.
```{r}
Y<-heart[[14]]
head(Y)
```
Now let us predict using our model. The model has already seen the data.
We do not expect our model, to reproduce at 100%. But we do expect that our model performs better than random guess. If our model under-performs random guess, then that would be an indication of the model's inability to learn at all.
Here in such cases that model is deficient and is a case of under-fitting.
```{r}
newmodel<-glm(formulastr,data=heart,family='binomial')
summary(newmodel)
summnewmodel<-summary(newmodel)
(summnewmodel$aic)
(summnewmodel$deviance)
(summnewmodel$null.deviance)
(p_values<-coef(summnewmodel)[,4])
```
Now let us compute the performance of the classifier.As indicated above,
a random choice would have scored 0.5. If our model is any good it must perform
better than 0.5.
```{r}
table(p_values<0.05)
predYprob<-predict(newmodel,heart[,1:13],type='response')
predY<-ifelse(predYprob<0.5,0,1)
table(predY)
table(heart[[14]])
table(heart[[14]],predY)
library(caret)
require(caret)
cfm<-caret::confusionMatrix(table(heart[[14]],predY))
cfm
```
Accuracy is 84.49%.
But this is from data the model has already seen.

# Underfitting 
A Model with an accuracy of 70% or higher is performant, and therefore, is not underfitting. We can surmise the model is capable of learning. It is better than
random guess which would yield 50% accuracy.

Business is not interested in predicting what has already happened. Business need help
when there is uncertainty and not enough information is available. How will this model perform when a new patient or a new customer arrives. 
*How can we estimate model performance over never seen before data? That is the capability,
**generalize**, we are seeking. Is Life not the response to un-controlled, un-anticipated, unpredictable events?*

# Never seen before data -- HOLD-OUT
We need to devise a process to find out how our model performs when new data is
presented. To do that, we will split our dataset into two disjoint partitions: a training set **70%** and a test set **30%**.
We will hold out some of the data given to us as test set.
We will train the model on the training set. 
Measure its performance on the training set.
Make sure it is at or near 0.85 accuracy, affirming it still is capable of learning with **70%** of the data.
Then we predict the lable for the observations in the test set. 
Measure the performance on the test set.

# Over-fitting, learning from noise
We normally expect performance over test set to be less than training set. If the performance over test set decreases by 20% or more, our model is over-fitting. Two well known causes of over-fitting, (1) *model complexity* and (2) *over-training*. The model has captured both noise and signal and the noise is reflected in the predictions, resulting in sub-optimal **generalization**.

*I have quoted many numerical values as above, but they are my discipline definitely not a rule. They are my choice. You can set your own limits, please do so before running the model and do not change.*

# Data Partition Repeatability/Reproducibility
Let us split our data into "random" test/train disjoint partitions. Always set the seed.
We will train our model on the training set. We will predict using training set to
ensure the model is NOT under-fitting and then predict using the test set, some data
model is not trained on. The test set represents new data. We will evaluate the model's ability to generalize.
We can determine if our model is *over-fitting* influenced by noise in the training set.
```{r}
set.seed(43)
tstidx<-sample(1:nrow(heart),0.30*nrow(heart),replace=F)
trdata<-heart[-tstidx,]
tsdata<-heart[tstidx,]
```
Now let us run the model.
```{r}
glm.trmodel<-glm(formulastr,data=trdata,family='binomial')
summary(glm.trmodel)
predtr<-predict(glm.trmodel,trdata[,1:13],type='response')
head(predtr)

predtrclass<-ifelse(predtr<0.5,0,1)
table(trdata[[14]])
table(predtrclass)

length(predtrclass)==length(trdata[[14]])
(trcfm<-caret::confusionMatrix(table(trdata[[14]],predtrclass)))
```
Accuracy on training data is 0.8404 not bad..model is capable of learning.
Now let us predict the class for never seen before data.
That is our 'held out' test dataset. This is what matters to the business.
```{r}
predts<-predict(glm.trmodel,tsdata[,1:13],type='response')
predtsclass<-ifelse(predts<0.5,0,1)
table(predtsclass)
table(tsdata[[14]])
table(tsdata[[14]],predtsclass)
tscfm<-caret::confusionMatrix(table(tsdata[[14]],predtsclass))
tscfm
(precision <- tscfm$byClass['Pos Pred Value'])    #https://stackoverflow.com/questions/13548092
(recall <- tscfm$byClass['Sensitivity'])
(f_measure <- 2 * ((precision * recall) / (precision + recall))) #geometric mean instead of arithmatic mean
```
Accuracy on the test set (held out data or never seen before data) 
is 0.8111, and has fallen down from accuracy obtained during training phase 0.84. The drop  is about 4% drop in performance and therefore we conclude Model is not over-fitting.
WHAT IF, our performance had fallen quite a bit, say more than 25%, from training to testing phase. That would be a classic case of overfitting. That limit 25% is my discipline. It is not a rule. In consultation with your sponsor you set, these limits and follow it with discipline.

* There are many advanced techniques to mitigate overfitting: 
    +   LASSO, 
    +   RIDGE, and 
    +   Cross Validation.

Here, We will remove some variables and see if it has any bearing on our results.
```{r}
tst.model2<-glm(target~cp+ca+thal+oldpeak,data=trdata,family='binomial')
summary(tst.model2)
tr.tst.pred<-predict(tst.model2,trdata[,1:13],family='binomial')
tr.pred.class<-ifelse(tr.tst.pred<0.5,0,1)
tr.pred.table<-table(trdata[[14]],tr.pred.class)
tr.pred.table
(tr.pred.cfm<-confusionMatrix(tr.pred.table))
(precision <- tr.pred.cfm$byClass['Pos Pred Value'])    #https://stackoverflow.com/questions/13548092
(recall <- tr.pred.cfm$byClass['Sensitivity'])
(f_measure <- 2 * ((precision * recall) / (precision + recall)))
```
Our accuracy for the training phase is `r tr.pred.cfm$overall['Accuracy']`. 
Now let us repeat it over test to find out if there is over-fitting.
```{r}
tst.pred2<-predict(tst.model2,tsdata[,1:13],type='response')
tst.pred2.class<-ifelse(tst.pred2<0.5,0,1)
tst.pred2.table<-table(tsdata[[14]],tst.pred2.class)
tst.pred2.table
(tst.pred2.cfm<-confusionMatrix(tst.pred2.table))
(accuracy<-tst.pred2.cfm$overall['Accuracy'])
(precision <- tst.pred2.cfm$byClass['Pos Pred Value'])    #https://stackoverflow.com/questions/13548092
(recall <- tst.pred2.cfm$byClass['Sensitivity'])
(f_measure <- 2 * ((precision * recall) / (precision + recall)))
```
Accuracy is `r accuracy`. Accuracy on training data was `r tr.pred.cfm$overall['Accuracy']`.
There is no over-fitting. So we move on.
Now let us visualize our performance using ROC plots.
We need ROCR or pROC package and there are other packages.
I use pROC.
```{r echo=FALSE}
graphics.off() 
par("mar") 
par(mar=c(1,1,1,1))
if(!require(pROC))install.packages('pROC')
library(pROC)
par(pty="s") 
glmROC <- roc(tsdata[[14]]~ tst.pred2.class,plot=TRUE,print.auc=TRUE,col="green",lwd =4,legacy.axes=TRUE,main="ROC Curves")
```
Now let us compute AUC and plot Receiver Operating Curve (ROC) using ROCR 
package.
```{r}
getMetrics<-function(actual_class,predicted_response)
{
X=list()
if ( require(ROCR) ) {
auc_1=prediction(predicted_response,actual_class)
prf=performance(auc_1, measure="tpr",x.measure="fpr")
slot_fp=slot(auc_1,"fp")
slot_tp=slot(auc_1,"tp")

fpr=unlist(slot_fp)/unlist(slot(auc_1,"n.neg"))
tpr=unlist(slot_tp)/unlist(slot(auc_1,"n.pos"))

auc<-performance(auc_1,"auc")
AUC<-auc@y.values[[1]]
X=list(fpr=fpr,tpr=tpr,auc=AUC)
}
X
}
```
time to test our utility function...
```{r}
L<-getMetrics(tsdata[[14]],tst.pred2)
plot(L$fpr,L$tpr,main=" ROC Plot tpr vs fpr")
print(paste("AUC=",L$auc,sep=''))
```

We would prefer 90 or above...but AUC=0.89 and accuracy = 0.8 is acceptable.
the recall for this learner is 0.8718 and precision is 0.7727 and their geometric mean, aka F-Score is 0.8193.

## Logistic Regression is a goto algo!!
Definitely, GLM is doing better than a random guess!!! 
Logistic Regression is an excellent baseline algorithm,
My opinion, one must always must perform logistic regression in any ML exercise as a base line.
To show some respect, to grand old **Ockham and his razor**, do NOT use, ANY exotic hyper-parameters, at this stage,
Use just plain vanilla models, as much as possible using objective criteria.
Try to refine programmatically and numerically without requiring human intervention.
That is smart!

*Most importantly there is significant mathematics, statistics and probability behind these algorithms. In the word document, we take a deeper look.*
This module will serve as a template, for all our clinical modules in this course we will provide.

* We will include
   +   R script, 
   +   RMD Script, 
   +   HTML (from the RMD) and 
   +   an essay in word format.

Next Module we will take up Naive Bayes, another parametric model.

### 2020-09-15 DATA622/M02-glm.R

@@ -1,0 +1,174 @@
+unixpath<-"/home/data/ML.Data/heart.csv"
+unixecolipath<-"/home/data/ML.Data/ecoli.csv"
+path<-"c://Users/rkannan/rk/data/heart.csv"
+ecolipath<-"c://Users/rkannan/rk/data/ecoli.csv"
+heart<-read.csv(path,header=TRUE,sep=',',stringsAsFactors=FALSE)
+ecoli<-read.csv(ecolipath,head=T,sep=',',stringsAsFactors=F)
+loadData<-function(csvfile) { read.csv(csvfile,head=T,sep=',',stringsAsFactors=F) }
+ecoli_1<-ecoli
+ecoli_2<-loadData(ecolipath)
+TBL=table(ecoli_1==ecoli_2)
+TBL
+head(heart)
+tail(heart)
+dim(heart)
+nrow(heart)
+ncol(heart)
+names(heart)
+
+names(heart)[[1]]<-'age'
+
+(corheart<-cor(heart[,1:13]))
+
+isConstant<-function(x) length(names(table(x)))<2
+apply(heart,2,isConstant)
+
+isConstant<-function(x) length(names(table(x)))<2
+apply(heart,2,isConstant)
+
+classLabels<-table(heart$target)
+print(classLabels)
+names(classLabels)
+length(names(classLabels))
+ifelse(length(names(classLabels))==2,"binary classification", "multi-class classification")
+
+glm_model<-glm(target~.,data=heart, family='binomial')
+glm_model
+summary_glm_model<-summary(glm_model)
+coef_summary_glm_model<-coef(summary_glm_model)
+dim(coef_summary_glm_model)
+coef_summary_glm_model[[1]]
+row.names(coef_summary_glm_model)
+
+coef_summary_glm_model
+coef_summary_glm_model[,4]<0.05
+row_names<-row.names(coef_summary_glm_model[coef_summary_glm_model[,4]<0.05,])
+row_names
+summary_glm_model$aic
+summary_glm_model$null.deviance
+summary_glm_model$deviance
+ifelse(summary_glm_model$deviance<summary_glm_model$null.deviance,"model has improved","model has not helped")
+
+if(!require(car))install.packages('car')
+library(car)
+vif_model<-vif(glm_model)
+
+vif_model
+vif_model[vif_model>4]
[…]

### 2020-09-15 DATA622/M02-glm.rmd
Updated path variables

@@
 unixecolipath<-"/home/data/ML.Data/ecoli.csv"
-path<-"c://Users/rkannan/rk/data/heart.csv"
-ecolipath<-"c://Users/rkannan/rk/data/ecoli.csv"
+#path<-"c://Users/rkannan/rk/data/heart.csv"
+#ecolipath<-"c://Users/rkannan/rk/data/ecoli.csv"
+path<-"heart.csv"
+ecolipath<-"ecoli.csv"
 heart<-read.csv(path,header=TRUE,sep=',',stringsAsFactors=FALSE)

### 2020-09-19 DATA698/01 - Gather iNaturalist JSON Data.py
Initial commit

@@ -1,0 +1,102 @@
+# -*- coding: utf-8 -*-
+"""
+Gather iNaturalist JSON Data
+
+Created on Tue Sep 15 20:23:39 2020
+
+@author: Michael Silva
+
+This script gathers data on iNaturalist observations.  The data are organized
+by iNaturalist taxons.  It will request the maximum allowed number of data and
+save them as JSON files for further processing.
+"""
+
+from os import path
+import json
+import requests
+
+# iNaturalist Taxon ID and the Label (does not have to be unique)
+taxon_id_and_label = {
+    "58732": "Poison Ivy",
+    "58729": "Poison Ivy",
+    "50278": "Virginia Creeper",
+    "47726": "Box Elder",
+    "138680": "Thicket Bean",
+    "62671": "Kudzu",
+    "62666": "Kudzu",
+    "82529": "Grapevine",
+    "119936": "Grapevine",
+    "50625": "Hoptree",
+    "58738": "Fragrant Sumac",
+    "55911": "Bramble",
+    "54436": "Bramble",
+    "47544": "Bramble",
+    "125489": "Bramble",
+    "50298": "Strawberry",
+    "243824": "Strawberry",
+    "77835": "Japanese Honeysuckle",
+    "50310": "Jack in the Pulpit",
+    "166162": "Boston Ivy",
+    "129021": "Honewort",
+    "51435": "Trillium",
+    "50855": "Trillium",
+    "51741": "Goutweed",
+    "85485": "Hog Peanut",
+}
+# Number of observations per API call
+per_page = 100
+# iNaturalist currently only allows 10,000 results
+inaturalist_max = int(round(10000 / per_page, 0))
+
+
+for taxon_id, label in taxon_id_and_label.items():
+    print("Gathering " + taxon_id_and_label[taxon_id] + " Data...")
+    scrape_and_save = True
+    page = 1
+
+    while scrape_and_save:
+        # Determine if this has already been scrapped
+        page_str = str(page)
+        page_str = page_str.zfill(3)
+        file_path = (
+            "./iNaturalist/json/"
+            + taxon_id_and_label[taxon_id]
+            + "-"
+            + taxon_id
+            + "-"
+            + page_str
+            + ".json"
+        )
+        if not path.isfile(file_path):
+            # File does not exist so scrape and save
+            url = (
+                "https://api.inaturalist.org/v1/observations?verifiable=true&order_by=observations.id&order=desc&page="
+                + str(page)
[…]

### 2020-09-19 DATA698/02 - Gather iNaturalist Image URLs.py

@@ -1,0 +1,103 @@
+# -*- coding: utf-8 -*-
+"""
+Gather iNaturalist Image URLs
+
+Created on Fri Sep 18 16:24:40 2020
+
+@author: Michael Silva
+
+Having previously gathered iNaturalist observation records, and saved them to
+JSON files, this script processess the data and gets the URL for the image. All
+data are stored in a SQLite database.
+"""
+
+import time
+import json
+import requests
+from glob import glob
+import sqlite3
+
+# It may take more than one pass to gather the image URLs.  Set this to False
+# if this is not the first run on the script.
+
+create_tables = False
+cool_down = False
+
+conn = sqlite3.connect("images.db")
+c = conn.cursor()
+
+if create_tables:
+    c.execute("DROP TABLE IF EXISTS images;")
+    c.execute("DROP TABLE IF EXISTS json_file;")
+    c.execute(
+        "CREATE TABLE images (id INTEGER PRIMARY KEY AUTOINCREMENT, label TEXT, url text, img_url TEXT, lat REAL, lon REAL, observed_on TEXT, assigned INTEGER DEFAULT 0, downloaded INTEGER DEFAULT 0);"
+    )
+    c.execute("CREATE INDEX IF NOT EXISTS idx_assigned ON images (assigned);")
+    c.execute("CREATE INDEX IF NOT EXISTS idx_downloaded ON images (downloaded);")
+    c.execute("CREATE TABLE json_file (file_name TEXT);")
+    conn.commit()
+
+def gather(cool_down):
+    for file in glob("./iNaturalist/json/*.json"):
+        c.execute("SELECT COUNT(*) FROM json_file WHERE file_name = ?", (file,))
+        result = c.fetchone()
+        if result[0] == 0:
+            label = file.split("\\")[1].split("-")[0]
+            with open(file) as json_file:
+                data = json.load(json_file)
+                if "results" in data and not cool_down:
+                    for result in data["results"]:
+                        url = result["uri"].replace(
+                            "https://www.inaturalist.org/",
+                            "https://api.inaturalist.org/v1/",
+                        )
+                        print(label + " >> " + url)
+                        c.execute("SELECT COUNT(*) FROM images WHERE url = ?", (url,))
+                        r = c.fetchone()
+                        if r[0] == 0:
+                            response = requests.get(url)
+                            if response.status_code == 200:
+                                response_json = response.json()
+        
+                                img_url = response_json["results"][0]["photos"][0]["url"]
+                                img_url = img_url.replace("square", "large")
[…]

### 2020-09-20 DATA622/M02-glm.rmd
Minor bugfix

@@
 we can also do length(unique(x)) < 1
+```{r}
+isConstant<-function(x) length(names(table(x)))<2
+apply(birds,2,isConstant)
+```
 ```{r}
@@
 ```{r}
-newmodel<-glm(formulastr,data=heart,family='binomial')
+newmodel<-glm(newformulastr,data=heart,family='binomial')
 summary(newmodel)

### 2020-09-20 DATA698/01 - Gather iNaturalist JSON Data.py
Improvements

@@ -27,7 +27,7 @@ taxon_id_and_label = {
     "82529": "Grapevine",
     "119936": "Grapevine",
     "50625": "Hoptree",
-    "58738": "Fragrant Sumac",
+    "58738": "Sumac",
     "55911": "Bramble",
     "54436": "Bramble",
     "47544": "Bramble",

### 2020-09-20 DATA698/02 - Gather iNaturalist Image URLs.py

@@ -23,6 +23,9 @@ import sqlite3
 create_tables = False
 cool_down = False
 
+pause_in_seconds = 5
+cool_down_in_seconds = 120
+
 conn = sqlite3.connect("images.db")
 c = conn.cursor()
 
@@ -42,51 +45,57 @@ def gather(cool_down):
         c.execute("SELECT COUNT(*) FROM json_file WHERE file_name = ?", (file,))
         result = c.fetchone()
         if result[0] == 0:
-            label = file.split("\\")[1].split("-")[0]
             with open(file) as json_file:
+                label = file.split("\\")[1].split("-")[0]
                 data = json.load(json_file)
                 if "results" in data and not cool_down:
                     for result in data["results"]:
-                        url = result["uri"].replace(
-                            "https://www.inaturalist.org/",
-                            "https://api.inaturalist.org/v1/",
-                        )
-                        print(label + " >> " + url)
-                        c.execute("SELECT COUNT(*) FROM images WHERE url = ?", (url,))
-                        r = c.fetchone()
-                        if r[0] == 0:
-                            response = requests.get(url)
-                            if response.status_code == 200:
-                                response_json = response.json()
-        
-                                img_url = response_json["results"][0]["photos"][0]["url"]
-                                img_url = img_url.replace("square", "large")
-                                
-                                lat = response_json["results"][0]["geojson"]["coordinates"][
-                                    1
-                                ]
-                                lon = response_json["results"][0]["geojson"]["coordinates"][
-                                    0
-                                ]
-                                observed_on = response_json["results"][0]["observed_on"]
-        
-                                insert_data = (
-                                    label,
-                                    url,
-                                    img_url,
-                                    lat,
-                                    lon,
-                                    observed_on,
-                                )
-        
-                                c.execute(
-                                    "INSERT INTO images (label, url, img_url, lat, lon, observed_on) VALUES (?,?,?,?,?,?)",
-                                    insert_data,
-                                )
-                                conn.commit()
-                            else:
-                                cool_down = True
-                                return cool_down
+                        if "inaturalist" in result["uri"]:
+                            url = result["uri"].replace(
+                                "https://www.inaturalist.org/",
+                                "https://api.inaturalist.org/v1/",
+                            ).replace(
+                                "http://www.inaturalist.org/",
+                                "https://api.inaturalist.org/v1/",
+                            )
+                            c.execute("SELECT COUNT(*) FROM images WHERE url = ?", (url,))
+                            r = c.fetchone()
+                            if r[0] == 0:
[…]

### 2020-09-20 DATA698/03 - iNaturalist Image Download Controller.py
Initial commit

@@ -1,0 +1,73 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sat Sep 19 07:15:13 2020
+
+@author: Michael Silva
+"""
+
+import sqlite3
+from flask import Flask, jsonify, request, g
+import json
+
+DATABASE = "images.db"
+debug = False
+app = Flask(__name__)
+
+
+def make_dicts(cursor, row):
+    return dict((cursor.description[idx][0], value) for idx, value in enumerate(row))
+
+
+def get_db():
+    db = getattr(g, "_database", None)
+    if db is None:
+        db = g._database = sqlite3.connect(DATABASE)
+        db.row_factory = make_dicts
+    return db
+
+
+def db_save():
+    get_db().commit()
+
+
+@app.teardown_appcontext
+def close_connection(exception):
+    db = getattr(g, "_database", None)
+    if db is not None:
+        db.close()
+
+
+@app.route("/")
+def home():
+    return ""
+
+
+@app.route("/job")
+def job():
+    cur = get_db().cursor()
+    cur.execute(
+        "SELECT images.* FROM images, (SELECT MIN(id) AS id FROM images WHERE assigned = 0) AS f WHERE images.id = f.id;"
+    )
+    try:
+        task = cur.fetchall()[0]
+        cur.execute("UPDATE images SET assigned = 1 WHERE id = ?", (task["id"],))
+        db_save()
+    except:
+        # We have assigned all the images
+        task = {"id": "Done"}
+    return jsonify(task)
+
+
+@app.route("/image_downloaded")
+def image_downloaded():
+    cur = get_db().cursor()
+    image_id = request.args.get("id")
+    cur.execute(
+        "UPDATE images SET assigned = 1, downloaded = 1 WHERE id = ?", (image_id,)
+    )
+    db_save()
+    return json.dumps({"success": True}), 200, {"ContentType": "application/json"}
+
+
+if __name__ == "__main__":
+    app.run(debug=debug)

### 2020-09-20 DATA698/04 - Download iNaturalist Images.py

@@ -1,0 +1,48 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sat Sep 19 07:15:13 2020
+
+@author: Michael Silva
+"""
+
+import requests
+from os import path, mkdir
+
+base_url = "http://127.0.0.1:5000/"
+has_work = True
+
+
+def relay_image_downloaded(id):
+    url = base_url + "image_downloaded?id=" + str(id)
+    response = requests.get(url)
+    response = response.json()
+    if "success" not in response:
+        print("Something bad happened > " + url)
+
+
+while has_work:
+    # Get a new job
+    job = requests.get(base_url + "job")
+    job = job.json()
+
+    if job["id"] == "Done":
+        has_work = False
+    else:
+        # Download the image
+        img_url = job["img_url"]
+        img_id = job["id"]
+        img_id_str = str(img_id)
+        label = job["label"]
+        img_path = "./iNaturalist/images/" + label + "/"
+        # Create directories if needed
+        if not path.exists(img_path):
+            mkdir(img_path)
+        # Check if image path exists
+        file_extension = img_url.split(".")[-1].split("?")[0]
+        file_path = img_path + img_id_str.zfill(8) + "." + file_extension
+        if not path.exists(file_path):
+            print("Downloading " + img_url)
+            img = requests.get(img_url)
+            with open(file_path, "wb") as f:
+                f.write(img.content)
+        relay_image_downloaded(img_id)

### 2020-09-20 DATA698/03 - iNaturalist Image Download Controller.py
Added reset endpoint

@@ -42,6 +42,17 @@ def home():
     return ""
 
 
+@app.route("/image_downloaded")
+def image_downloaded():
+    cur = get_db().cursor()
+    image_id = request.args.get("id")
+    cur.execute(
+        "UPDATE images SET assigned = 1, downloaded = 1 WHERE id = ?", (image_id,)
+    )
+    db_save()
+    return json.dumps({"success": True}), 200, {"ContentType": "application/json"}
+
+
 @app.route("/job")
 def job():
     cur = get_db().cursor()
@@ -58,15 +69,12 @@ def job():
     return jsonify(task)
 
 
-@app.route("/image_downloaded")
-def image_downloaded():
+@app.route("/reset")
+def reset():
     cur = get_db().cursor()
-    image_id = request.args.get("id")
-    cur.execute(
-        "UPDATE images SET assigned = 1, downloaded = 1 WHERE id = ?", (image_id,)
-    )
+    cur.execute("UPDATE images SET assigned = 0 WHERE downloaded = 0")
     db_save()
-    return json.dumps({"success": True}), 200, {"ContentType": "application/json"}
+    return "Success"
 
 
 if __name__ == "__main__":

### 2020-09-20 DATA622/M02-glm.R
Removed M02-glm.R

@@ -1,174 +1,0 @@
-unixpath<-"/home/data/ML.Data/heart.csv"
-unixecolipath<-"/home/data/ML.Data/ecoli.csv"
-path<-"c://Users/rkannan/rk/data/heart.csv"
-ecolipath<-"c://Users/rkannan/rk/data/ecoli.csv"
-heart<-read.csv(path,header=TRUE,sep=',',stringsAsFactors=FALSE)
-ecoli<-read.csv(ecolipath,head=T,sep=',',stringsAsFactors=F)
-loadData<-function(csvfile) { read.csv(csvfile,head=T,sep=',',stringsAsFactors=F) }
-ecoli_1<-ecoli
-ecoli_2<-loadData(ecolipath)
-TBL=table(ecoli_1==ecoli_2)
-TBL
-head(heart)
-tail(heart)
-dim(heart)
-nrow(heart)
-ncol(heart)
-names(heart)
-
-names(heart)[[1]]<-'age'
-
-(corheart<-cor(heart[,1:13]))
-
-isConstant<-function(x) length(names(table(x)))<2
-apply(heart,2,isConstant)
-
-isConstant<-function(x) length(names(table(x)))<2
-apply(heart,2,isConstant)
-
-classLabels<-table(heart$target)
-print(classLabels)
-names(classLabels)
-length(names(classLabels))
-ifelse(length(names(classLabels))==2,"binary classification", "multi-class classification")
-
-glm_model<-glm(target~.,data=heart, family='binomial')
-glm_model
-summary_glm_model<-summary(glm_model)
-coef_summary_glm_model<-coef(summary_glm_model)
-dim(coef_summary_glm_model)
-coef_summary_glm_model[[1]]
-row.names(coef_summary_glm_model)
-
-coef_summary_glm_model
-coef_summary_glm_model[,4]<0.05
-row_names<-row.names(coef_summary_glm_model[coef_summary_glm_model[,4]<0.05,])
-row_names
-summary_glm_model$aic
-summary_glm_model$null.deviance
-summary_glm_model$deviance
-ifelse(summary_glm_model$deviance<summary_glm_model$null.deviance,"model has improved","model has not helped")
-
-if(!require(car))install.packages('car')
-library(car)
-vif_model<-vif(glm_model)
-
-vif_model
-vif_model[vif_model>4]
[…]

### 2020-09-20 DATA622/M03-nb-script.rmd
Initial commit of Module 3

---
title: "NaiveBayes"
output:  
html_document:
df_print: paged
pdf_document: default
---
Let us run a classification task using NaiveBayes.

```{r Startup}

```
##NaiveBayes
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Let us train NaiveBayes model using standard implementation 
and compare it with our own computation. TO make our computation simpler
we will include only categorical variables in the dataset.
```{r MainCode}
library(e1071)
library(caret)
file<-"heart.csv"
```

As practitioners we must always be wary of Overfitting and underfitting.

We consider when the performance of a classifier falls by 50% from 
training to test dataset to be due to overfitting.

We consider when the performance of a classifier is below 50% to be
an indication of a weak classifier unable to learn. Also known as
underfitting.

Let us create some utility functions.

# usage isOverfitting(tr.cfm,tst.cfm)
```{r}
isOverfitting<-function(trcfm,tstcfm)
{
  # if the accuracy from training is much more than accuracy testing phase
  # then we conclude Model is overfitting 
  # unable to generalize
  tracc=trcfm$overall['Accuracy']
  tstacc=tstcfm$overall['Accuracy']
  tracc > 2*tstacc 
}
```
# usage isUnderfitting(tr.cfm)
```{r}
isUnderfitting<-function(trcfm)
{
  # if the accuracy from training is much more than accuracy testing phase
  # then we conclude Model is overfitting 
  # unable to generalize
  tracc=trcfm$overall['Accuracy']
  0.5 >  tracc
}
```
# Load data and EDA
Let us load the data.
And subset it to include only categorical predictors.
```{r}
file<-'c:/Users/rkannan/rk/data/heart.csv'
heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)
head(heart)
catheart<-heart[,c(2,3,6,7,9,11,12,13,14)]
```
#Preliminary EDA

```{r}
head(catheart)
dim(catheart)
table(catheart$target)
sum(table(catheart$target))
```
# Repeatable and Reproducible Experiments
Set a seed for repeatable/reproducible random number sequence
and split the data into training/test sets.

```{r}
set.seed(43)
trdidx<-sample(1:nrow(catheart),0.7*nrow(catheart),replace=F)
trcatheart<-catheart[trdidx,]
tstcatheart<-catheart[-trdidx,]
```
# Check for class imbalance
Here we check visually to see if there are labels which occur in the single digit.
If a class exists in just once, if we wish to learn then we cannot test the algorithm for that class and if we include it in the test then we cannot train on it.
There are many statistical techniques to correct for class imbalance, left to another follow up workshop.
```{r}
table(catheart$target)
```
# Training and testing sets must be similar to the entire dataset
Are the training and test set distributed approximately similar to
given dataset. This is important so that statistics that applies 
to training and test sets, are applicable to the entire dataset.
## Otherwise, we cannot generalize.
```{r}
dim(trcatheart)
table(trcatheart$target)
dim(tstcatheart)
table(tstcatheart$target)
```
Train NB model, standard implementation using training set

```{r}
nbtr.model<-naiveBayes(target~.,data=trcatheart)
```
Performance over the training set

```{r}
nbtr.trpred<-predict(nbtr.model,trcatheart[,-c(9)],type='raw')
nbtr.trclass<-unlist(apply(round(nbtr.trpred),1,which.max))-1
nbtr.trtbl<-table(trcatheart[[9]], nbtr.trclass)
tr.cfm<-caret::confusionMatrix(nbtr.trtbl)
tr.cfm
```
Performance over held out data, the test set

```{r}
nbtr.tspred<-predict(nbtr.model,tstcatheart[,-c(9)],type='raw')
#roc.nbtr.tspred<-unlist(apply(nbtr.tspred,1,which.max)) # this returns the position
#roc.nbtr.tspred<-apply(nbtr.tspred,1,FUN=function(v)v[which.max(v)]) # this returns the value
#https://stackoverflow.com/questions/47883541/how-can-i-implement-roc-curve-analysis-for-naive-bayes-classification-algorithm
roc.nbtr.tspred<-nbtr.tspred[,2]
nbtr.tsclass<-unlist(apply(round(nbtr.tspred),1,which.max))-1
nbtr.tstbl<-table(tstcatheart[[9]], nbtr.tsclass)
tst.cfm<-caret::confusionMatrix(nbtr.tstbl)
tst.cfm
```
#Underfitting / Overfitting
```{r}
ifelse(isUnderfitting(tr.cfm),"model is deficient",paste("There is no underfitting", "model is an effective learner@ [",tr.cfm$overall['Accuracy'],"] accuracy"))
ifelse(isOverfitting(tr.cfm,tst.cfm),"model is overfitting -- too complex",
       paste(" There is no overfitting, model is an effective learner@ [",tr.cfm$overall['Accuracy'],"] training accuracy vs testing accuracy=[",tst.cfm$overall['Accuracy'],']'))
```
#Conditionally Independent Class Probabilities
Let us now compute NaiveBayes probabilities for each class
directly from data.
Given any predictor and a value
we compute the class conditional probabilities
p(c1|x) and p(c2|x) using 
p(c1|X)=p(x1|c1)p(x2|c1)p(x3|c1)..p(xn|c1)p(c1)
and
p(c2|X)=p(x1|c2)p(x2|c2)p(x3|c2)..p(xn|c2)p(c2)

Here p(c1)=frequency of c1 / total frequency
p(x1|c1)=frequency of x1 given c1/frequency of c1

we assign the class with the higher probability

```{r}
trcathearttarget_0<-trcatheart[trcatheart$target==0,]

trcathearttarget_1<-trcatheart[trcatheart$target==1,]

tstcathearttarget_0<-tstcatheart[tstcatheart$heart==0,]
tstcathearttarget_1<-tstcatheart[tstcatheart$heart==1,]
```
We compute class frequencies as follows, 
the number of rows in each class is the frequency of that class.
# class c1 proportion
```{r}
c0freq<-nrow(trcathearttarget_0)
c1freq<-nrow(trcathearttarget_1)
p0<-c0freq/(c0freq+c1freq)
p1<-c1freq/(c0freq+c1freq)
c(p0,p1)
```
We count, the number of observations in which the attribute has the given value
#We return the two proportions for a binary classifier.
```{r}

classProbabilities<-
  function(dsetc1,dsetc2,prop,val,c1prob,c2prob)
  {

    propdsetc1=dsetc1[dsetc1[prop]==val,]
    propdsetc2=dsetc2[dsetc2[prop]==val,]
    c(nrow(propdsetc1)/nrow(dsetc1),nrow(propdsetc2)/nrow(dsetc1))
  }
```
Let us put our utility functions to test, calculate conditional probability
```{r}
trcathearttarget_0[1,]
```
sex cp fbs restecg exang slope ca thal target
276   1  0   0       1     0     2  2    3      0

Let us now compute the probabilities for each of these attribute/value pairs.

```{r}
classProbabilities(trcathearttarget_0,trcathearttarget_1,'sex',1,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'cp',0,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'fbs',0,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'restecg',0,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'exang',1,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'slope',2,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'ca',2,p0,p1)
classProbabilities(trcathearttarget_0,trcathearttarget_1,'thal',3,p0,p1)
```

```{r}
xx<-c(
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'sex',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'cp',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'fbs',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'restecg',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'exang',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'slope',2,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'ca',2,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'thal',3,p0,p1),
  c(p0,p1))
nr<-9
c(p0,p1)
mx<-matrix(xx,nr,ncol=2,byrow=T)
#mx
cp<-unlist(apply(mx,2,cumprod))
#cp[nr,]
pred_class<-which.max(cp[nr,])-1
trcathearttarget_0[1,]$target
ifelse(trcathearttarget_0[1,]$target==0,
       ifelse(pred_class==0,'TP','FN'),
       ifelse(pred_class==1,'TN','FP'))
```

```{r}
tstcatheart[7,]
```
sex cp fbs restecg exang slope ca thal target
26   0  1   0       1     0     2  2    2      1
```{r}
tstcathearttarget_0<-tstcatheart[tstcatheart$heart==0,]
tstcathearttarget_1<-tstcatheart[tstcatheart$heart==1,]

xx<-c(
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'sex',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'cp',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'fbs',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'restecg',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'exang',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'slope',2,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'ca',2,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'thal',2,p0,p1),
  c(p0,p1))

mx<-matrix(xx,nrow=9,ncol=2,byrow=T)

unlist(apply(mx,2,cumprod))
cp<-unlist(apply(mx,2,cumprod))
which.max(cp[9,])-1
pred_class<-which.max(cp[nr,])-1
pred_class
tstcatheart[7,]$target
ifelse(tstcatheart[7,]$target==0,
       ifelse(pred_class==0,'TP','FN'),
       ifelse(pred_class==1,'TN','FP'))
```
#NOTE THAT we are using the distribution of training set 
#to evaluate the test set probabilities
# otherwise we could not claim to be generalizing
Let us do one more 23rd observation from the test set.seed
```{r}
tstcatheart[23,]

xx<-c(
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'sex',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'cp',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'fbs',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'restecg',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'exang',1,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'slope',2,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'ca',0,p0,p1),
  classProbabilities(trcathearttarget_0,trcathearttarget_1,'thal',2,p0,p1),
  c(p0,p1))

mx<-matrix(xx,nrow=9,ncol=2,byrow=T)

unlist(apply(mx,2,cumprod))
cp<-unlist(apply(mx,2,cumprod))
which.max(cp[9,])-1
pred_class<-which.max(cp[nr,])-1
pred_class
tstcatheart[23,]$target
ifelse(tstcatheart[23,]$target==0,
       ifelse(pred_class==0,'TP','FN'),
       ifelse(pred_class==1,'TN','FP'))

```
```{r}
evaluate_naiveBayes<-
function(nv,vv,targetclass,trdsetcl1,trdsetcl2,cl1prob,cl2prob)
{
nvl<-length(nv)
xx<-c()
for (idx in 1:nvl)
{
attr<-nv[idx]
value<-vv[[idx]]
#print(c(attr,value))
#xx<-c(xx,classProbabilities(trdsetcl1,trdsetcl2,nv[idx],vv[[idx]],cl1prob,cl2prob))
xx<-c(xx,classProbabilities(trdsetcl1,trdsetcl2,attr,value,cl1prob,cl2prob))
}
xx<-unlist(c(xx,  c(p0,p1)))
nvl<-nvl+1
mx<-matrix(xx,nrow=nvl,ncol=2,byrow=T)

unlist(apply(mx,2,cumprod))
cp<-unlist(apply(mx,2,cumprod))

pred_class<-which.max(cp[nvl,])-1
pred_class
targetclass
metric<-ifelse(targetclass==0,
       ifelse(pred_class==0,'TP','FN'),
       ifelse(pred_class==1,'TN','FP'))
c(targetclass,pred_class,metric)
}

```
Now we can iterate over the entire test set.
```{r}
tstcount<-nrow(tstcatheart)
tstcol<-ncol(tstcatheart)
featuremx<-tstcatheart[,-c(which(names(tstcatheart)=='target'))]
nv<-names(featuremx)
tstclass<-tstcatheart$target
tstpred<-c()
for(r in 1:tstcount)
{
  tstpred<-c(tstpred,
             evaluate_naiveBayes(nv,featuremx[r,],
                            tstclass[r],trcathearttarget_0,
                            trcathearttarget_1,p0,p1))
}
tstresults<-matrix(tstpred,nrow=tstcount,ncol=3,byrow=T,dimnames=
                     list(1:tstcount,c("GT",'Pred','metric')))
tstresults
table(tstresults[,3])
table(tstresults[,1],tstresults[,2])
```
# Final observation 
Note that we estimated the class probabilities using the training set
however, we compute the metric using the test set.
# Is that not what we want?
Statistics and probabilities is one big hat from which we can pull many rabbits,
and if you are NOT cautious you can end up with a 3 legged rabbit.
This classifier assumes that features are uncorrelated and yet we achieved significant
inferences.

If you recall, the e1071:naiveBayes achieved 
```{r}
nbtr.tstbl
```
in comparison to our naive implementation.
```{r}
table(tstresults[,1],tstresults[,2])
```
Let us plot the ROCR curve and determine AUC for the e1071 standard
implementation.
```{r}
getMetrics<-function(actual_class,predicted_response)
{
  test.set=data.frame(target=actual_class,prediction=predicted_response)
  TPrates = c()
TNrates = c()
thresholds = seq(0, 1, by = 0.05)
total_positive=nrow(test.set[test.set$target==0 ,])
total_negative=nrow(test.set[test.set$target==1,])
for (threshold in thresholds) {
  TPrateForThisThreshold = nrow(test.set[test.set$target == 0 & test.set$prediction <= threshold,])/total_positive
  TNrateForThisThreshold = nrow(test.set[test.set$target == 1 & test.set$prediction > threshold,])/total_negative
  TPrates = c(TPrates, TPrateForThisThreshold)
  TNrates = c(TNrates, TNrateForThisThreshold)
}
if(!require(pracma))install.packages('pracma')
library(pracma)
AUC <- trapz(TPrates,TNrates)
X=list(fpr=1-TNrates,tpr=TPrates,auc=AUC)
X
#X=list()
#if ( require(ROCR) ) {
#auc_1=prediction(predicted_response,actual_class)
#prf=performance(auc_1, measure="tpr",x.measure="fpr")
#slot_fp=slot(auc_1,"fp")
#slot_tp=slot(auc_1,"tp")

#fpr=unlist(slot_fp)/unlist(slot(auc_1,"n.neg"))
#tpr=unlist(slot_tp)/unlist(slot(auc_1,"n.pos"))

#auc<-performance(auc_1,"auc")
#AUC<-auc@y.values[[1]]
#X=list(fpr=fpr,tpr=tpr,auc=AUC)
#}
#X
}
```
time to test our utility function...
```{r}
L<-getMetrics(tstcatheart[[9]],roc.nbtr.tspred)
plot(L$fpr, L$tpr, type="l",main=" NB ROC Plot tpr vs fpr")
#plot(L$fpr,L$tpr,main=" NB ROC Plot tpr vs fpr")
print(paste(" NB AUC=",L$auc,sep=''))
```

### 2020-09-20 DATA622/M03-nb-script.rmd
Modified file path

@@
 ```{r}
-file<-'c:/Users/rkannan/rk/data/heart.csv'
+file<-'heart.csv'
 heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)

### 2020-09-21 DATA698/02 - Process iNaturalist JSON Data.py
More work on DATA 698 Project

@@ -1,0 +1,61 @@
+# -*- coding: utf-8 -*-
+"""
+Process iNaturalist JSON Data
+
+Created on Fri Sep 18 16:24:40 2020
+
+@author: Michael Silva
+
+Having previously gathered iNaturalist observation records, and saved them to
+JSON files, this script processess the data and builds out an index of API calls.
+All data will be stored in a SQLite database.
+"""
+
+import json
+from glob import glob
+import sqlite3
+
+create_tables = False
+
+conn = sqlite3.connect("images.db")
+c = conn.cursor()
+
+if create_tables:
+    c.execute("DROP TABLE IF EXISTS image;")
+    c.execute("DROP TABLE IF EXISTS json_file;")
+    c.execute(
+        "CREATE TABLE images (id INTEGER PRIMARY KEY AUTOINCREMENT, label TEXT, url text, img_url TEXT DEFAULT NULL, lat REAL DEFAULT NULL, lon REAL DEFAULT NULL, observed_on TEXT DEFAULT NULL, scrapped INTEGER default 0, assigned INTEGER DEFAULT 0, downloaded INTEGER DEFAULT 0, error INTEGER DEFAULT 0);"
+    )
+    c.execute("CREATE INDEX IF NOT EXISTS idx_assigned ON images (assigned);")
+    c.execute("CREATE INDEX IF NOT EXISTS idx_downloaded ON images (downloaded);")
+    c.execute("CREATE INDEX IF NOT EXISTS idx_scrapped ON images (scrapped);")
+    c.execute("CREATE INDEX IF NOT EXISTS idx_error ON images (error);")
+    c.execute("CREATE TABLE json_file (file_name TEXT);")
+    conn.commit()
+
+for file in glob("./iNaturalist/json/*.json"):
+        c.execute("SELECT COUNT(*) FROM json_file WHERE file_name = ?", (file,))
+        result = c.fetchone()
+        if result[0] == 0:
+            print(file)
+            with open(file) as json_file:
+                label = file.split("\\")[1].split("-")[0]
+                data = json.load(json_file)
+                if "results" in data:
+                    for result in data["results"]:
+                        if "inaturalist.org" in result["uri"]:
+                            url = result["uri"].replace(
+                                "https://www.inaturalist.org/",
+                                "https://api.inaturalist.org/v1/",
+                            ).replace(
+                                "http://www.inaturalist.org/",
+                                "https://api.inaturalist.org/v1/",
+                            )
[…]

### 2020-09-21 DATA698/04 - Download iNaturalist Images.py

@@ -41,7 +41,7 @@ while has_work:
         file_extension = img_url.split(".")[-1].split("?")[0]
         file_path = img_path + img_id_str.zfill(8) + "." + file_extension
         if not path.exists(file_path):
-            print("Downloading " + img_url)
+            print("Downloading " + label + " >> " + img_url)
             img = requests.get(img_url)
             with open(file_path, "wb") as f:
                 f.write(img.content)

### 2020-09-21 DATA698/03 - iNaturalist Image Download Controller.py
More work

@@ -42,6 +42,15 @@ def home():
     return ""
 
 
+@app.route("/done-scrapping")
+def done_scrapping():
+    cur = get_db().cursor()
+    image_id = request.args.get("id")
+    cur.execute("UPDATE images SET scrapped = 1 WHERE id = ?", (image_id,))
+    db_save()
+    return json.dumps({"success": True}), 200, {"ContentType": "application/json"}
+
+
 @app.route("/image_downloaded")
 def image_downloaded():
     cur = get_db().cursor()
@@ -73,8 +82,34 @@ def job():
 def reset():
     cur = get_db().cursor()
     cur.execute("UPDATE images SET assigned = 0 WHERE downloaded = 0")
+    cur.execute("UPDATE images SET scrapped = 0 WHERE scrapped = -1")
     db_save()
     return "Success"
+
+
+@app.route("/scrape-me")
+def scrape_me():
+    cur = get_db().cursor()
+    starts_with = request.args.get("starts_with")
+    if starts_with is None:
+        cur.execute(
+            "SELECT images.* FROM images, (SELECT MIN(id) AS id FROM images WHERE scrapped = 0 AND error = 0) AS f WHERE images.id = f.id;"
+        )
+    else:
+        cur.execute(
+            "SELECT images.* FROM images, (SELECT MIN(id) AS id FROM images WHERE scrapped = 0 AND error = 0 AND label LIKE '"
+            + starts_with
+            + "%') AS f WHERE images.id = f.id;"
+        )
+        
+    try:
+        task = cur.fetchall()[0]
+        cur.execute("UPDATE images SET scrapped = -1 WHERE id = ?", (task["id"],))
+        db_save()
+    except:
+        # We have assigned all the images
+        task = {"id": "Done"}
+    return jsonify(task)
 
 
 if __name__ == "__main__":

### 2020-09-21 DATA698/03 - iNaturalist Image Download Controller.py
More work on the project

@@ -8,6 +8,7 @@ Created on Sat Sep 19 07:15:13 2020
 import sqlite3
 from flask import Flask, jsonify, request, g
 import json
+from collections import OrderedDict 
 
 DATABASE = "images.db"
 debug = False
@@ -42,15 +43,6 @@ def home():
     return ""
 
 
-@app.route("/done-scrapping")
-def done_scrapping():
-    cur = get_db().cursor()
-    image_id = request.args.get("id")
-    cur.execute("UPDATE images SET scrapped = 1 WHERE id = ?", (image_id,))
-    db_save()
-    return json.dumps({"success": True}), 200, {"ContentType": "application/json"}
-
-
 @app.route("/image_downloaded")
 def image_downloaded():
     cur = get_db().cursor()
@@ -62,7 +54,35 @@ def image_downloaded():
     return json.dumps({"success": True}), 200, {"ContentType": "application/json"}
 
 
-@app.route("/job")
+@app.route("/image_scrapped")
+def image_scrapped():
+    cur = get_db().cursor()
+    image_id = request.args.get("id")
+    cur.execute("UPDATE images SET scrapped = 1 WHERE id = ?", (image_id,))
+    db_save()
+    return json.dumps({"success": True}), 200, {"ContentType": "application/json"}
+
+
+@app.route("/image_data", methods=["POST"])
+def data_intake():
+    cur = get_db().cursor()
+    image_id = request.args.get("id")
+    data = OrderedDict(request.get_json())
+    val = tuple(data.values())
+    sql = "UPDATE images SET "
+    for k in data.keys():
+        sql = sql + k + " = ?, "
+    sql = sql[: len(sql) - 2] + " WHERE id = " + image_id + ";"
+    
+    try:
+        cur.execute(sql, val)
+        db_save()
+        return json.dumps({"success": True}), 200, {"ContentType": "application/json"}
+    except:
+        return json.dumps({"failure": True}), 200, {"ContentType": "application/json"}
+
+
+@app.route("/download_job")
 def job():
     cur = get_db().cursor()
     cur.execute(
@@ -87,8 +107,8 @@ def reset():
     return "Success"
 
 
-@app.route("/scrape-me")
-def scrape_me():
+@app.route("/scrape_job")
+def scrape_job():
     cur = get_db().cursor()
     starts_with = request.args.get("starts_with")
     if starts_with is None:
@@ -101,7 +121,7 @@ def scrape_me():
             + starts_with
[…]

### 2020-09-21 DATA698/04 - Download iNaturalist Images.py

@@ -7,13 +7,27 @@ Created on Sat Sep 19 07:15:13 2020
 
 import requests
 from os import path, mkdir
+import json
+import time
 
 base_url = "http://127.0.0.1:5000/"
 has_work = True
+cooling_down = False
+job_count = 0
+starts_with = None
+
+
+work_modes = {"1": "scrape_job", "2": "download_job", "3": "reset"}
+work_mode = input("Enter work mode (1=scrape, 2=download image, 3=reset): ")
+endpoint = work_modes[work_mode]
+if work_mode == "1":
+    starts_with = input("What should the label start with (Press enter for nothing)? ")
+    if starts_with == "":
+        starts_with = None
 
 
 def relay_image_downloaded(id):
-    url = base_url + "image_downloaded?id=" + str(id)
+    url = base_url + "image_downloaded?id=" + id
     response = requests.get(url)
     response = response.json()
     if "success" not in response:
@@ -21,28 +35,80 @@ def relay_image_downloaded(id):
 
 
 while has_work:
-    # Get a new job
-    job = requests.get(base_url + "job")
-    job = job.json()
-
-    if job["id"] == "Done":
+    if cooling_down:
+        print(
+            "========> COMPLETED " + str(job_count) + " JOBS BEFORE COOL DOWN <========"
+        )
+        time.sleep(60)
+        cooling_down = False
+        job_count = 0
+    if endpoint == work_modes["3"]:
+        # Reset
+        response = requests.get(base_url + "reset")
+        print("Reset complete")
         has_work = False
     else:
-        # Download the image
-        img_url = job["img_url"]
-        img_id = job["id"]
-        img_id_str = str(img_id)
-        label = job["label"]
-        img_path = "./iNaturalist/images/" + label + "/"
-        # Create directories if needed
-        if not path.exists(img_path):
-            mkdir(img_path)
-        # Check if image path exists
-        file_extension = img_url.split(".")[-1].split("?")[0]
-        file_path = img_path + img_id_str.zfill(8) + "." + file_extension
-        if not path.exists(file_path):
-            print("Downloading " + label + " >> " + img_url)
-            img = requests.get(img_url)
-            with open(file_path, "wb") as f:
-                f.write(img.content)
-        relay_image_downloaded(img_id)
+        # Get a new job
+        job_count += 1
+        request_url = base_url + endpoint
+        if starts_with is not None:
+            request_url = request_url + "?starts_with=" + starts_with
+
[…]

### 2020-09-21 DATA698/02 - Gather iNaturalist Image URLs.py
Minor changes

@@ -1,112 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Gather iNaturalist Image URLs
-
-Created on Fri Sep 18 16:24:40 2020
-
-@author: Michael Silva
-
-Having previously gathered iNaturalist observation records, and saved them to
-JSON files, this script processess the data and gets the URL for the image. All
-data are stored in a SQLite database.
-"""
-
-import time
-import json
-import requests
-from glob import glob
-import sqlite3
-
-# It may take more than one pass to gather the image URLs.  Set this to False
-# if this is not the first run on the script.
-
-create_tables = False
-cool_down = False
-
-pause_in_seconds = 5
-cool_down_in_seconds = 120
-
-conn = sqlite3.connect("images.db")
-c = conn.cursor()
-
-if create_tables:
-    c.execute("DROP TABLE IF EXISTS images;")
-    c.execute("DROP TABLE IF EXISTS json_file;")
-    c.execute(
-        "CREATE TABLE images (id INTEGER PRIMARY KEY AUTOINCREMENT, label TEXT, url text, img_url TEXT, lat REAL, lon REAL, observed_on TEXT, assigned INTEGER DEFAULT 0, downloaded INTEGER DEFAULT 0);"
-    )
-    c.execute("CREATE INDEX IF NOT EXISTS idx_assigned ON images (assigned);")
-    c.execute("CREATE INDEX IF NOT EXISTS idx_downloaded ON images (downloaded);")
-    c.execute("CREATE TABLE json_file (file_name TEXT);")
-    conn.commit()
-
-def gather(cool_down):
-    for file in glob("./iNaturalist/json/*.json"):
-        c.execute("SELECT COUNT(*) FROM json_file WHERE file_name = ?", (file,))
-        result = c.fetchone()
-        if result[0] == 0:
-            with open(file) as json_file:
-                label = file.split("\\")[1].split("-")[0]
-                data = json.load(json_file)
-                if "results" in data and not cool_down:
-                    for result in data["results"]:
-                        if "inaturalist" in result["uri"]:
-                            url = result["uri"].replace(
-                                "https://www.inaturalist.org/",
-                                "https://api.inaturalist.org/v1/",
-                            ).replace(
-                                "http://www.inaturalist.org/",
-                                "https://api.inaturalist.org/v1/",
-                            )
-                            c.execute("SELECT COUNT(*) FROM images WHERE url = ?", (url,))
-                            r = c.fetchone()
-                            if r[0] == 0:
[…]

### 2020-09-21 DATA698/03 - iNaturalist Work Controller.py
Minor correction to SQL query

@@ -86,7 +86,7 @@ def data_intake():
 def job():
     cur = get_db().cursor()
     cur.execute(
-        "SELECT images.* FROM images, (SELECT MIN(id) AS id FROM images WHERE assigned = 0) AS f WHERE images.id = f.id;"
+        "SELECT images.* FROM images, (SELECT MIN(id) AS id FROM images WHERE assigned = 0 and img_url IS NOT NULL and error = 0) AS f WHERE images.id = f.id;"
     )
     try:
         task = cur.fetchall()[0]

### 2020-09-21 DATA698/04 - iNaturalist Worker.py
Minor change

@@ -60,7 +60,7 @@ while has_work:
         if job["id"] == "Done":
             has_work = False
 
-        if endpoint == work_modes["1"]:
+        if endpoint == work_modes["1"] and has_work:
             # Scrape observation data
             print(job["label"] + " >> Scrapping " + job["url"])
             response = requests.get(job["url"])
@@ -94,7 +94,7 @@ while has_work:
                 )
             else:
                 cooling_down = True
-        if endpoint == work_modes["2"]:
+        if endpoint == work_modes["2"] and has_work:
             # Download the image
             img_url = job["img_url"]
             img_id_str = str(job["id"])

### 2020-09-21 DATA698/04 - iNaturalist Worker.py
Minor change

@@ -10,22 +10,24 @@ from os import path, mkdir
 import json
 import time
 
+# Initial values
 base_url = "http://127.0.0.1:5000/"
 has_work = True
 cooling_down = False
 job_count = 0
 starts_with = None
 
-
-work_modes = {"1": "scrape_job", "2": "download_job", "3": "reset"}
-work_mode = input("Enter work mode (1=scrape, 2=download image, 3=reset): ")
+# How will this worker operate?  Which mode?
+work_modes = {"1": "scrape_job", "2": "download_job", "3": "scrape_job", "4":"reset"}
+work_mode = input("Enter work mode (1=scrape and download image, 2=download image ONLY, 3=scrape ONLY, 4=reset): ")
 endpoint = work_modes[work_mode]
-if work_mode == "1":
+# Ask about label starting with
+if work_mode in ["1", "3"]:
     starts_with = input("What should the label start with (Press enter for nothing)? ")
     if starts_with == "":
         starts_with = None
 
-
+# Helper functions
 def relay_image_downloaded(id):
     url = base_url + "image_downloaded?id=" + id
     response = requests.get(url)
@@ -33,16 +35,35 @@ def relay_image_downloaded(id):
     if "success" not in response:
         print("Something bad happened > " + url)
 
+def get_img_path(label):
+    img_path = "./iNaturalist/images/" + label + "/"
+    # Create directories if needed
+    if not path.exists(img_path):
+        mkdir(img_path)
+    return(img_path)
 
+def download_image(img_url, img_path, img_id_str):
+    # Check if image path exists
+    file_extension = img_url.split(".")[-1].split("?")[0]
+    file_path = img_path + img_id_str.zfill(8) + "." + file_extension
+    if not path.exists(file_path):
+        print("Downloading " + label + " >> " + img_url)
+        img = requests.get(img_url)
+        with open(file_path, "wb") as f:
+            f.write(img.content)
+
+# Main workflow loop
 while has_work:
     if cooling_down:
+        # We pulled data too fast so we need to let the API cool down
         print(
             "========> COMPLETED " + str(job_count) + " JOBS BEFORE COOL DOWN <========"
         )
         time.sleep(60)
         cooling_down = False
         job_count = 0
-    if endpoint == work_modes["3"]:
+
+    if endpoint == work_modes["4"]:
         # Reset
         response = requests.get(base_url + "reset")
         print("Reset complete")
@@ -62,7 +83,8 @@ while has_work:
 
[…]

### 2020-09-22 DATA698/03 - iNaturalist Work Controller.py
Added in viz

@@ -5,10 +5,11 @@ Created on Sat Sep 19 07:15:13 2020
 @author: Michael Silva
 """
 
-import sqlite3
-from flask import Flask, jsonify, request, g
 import json
-from collections import OrderedDict 
+import sqlite3
+from collections import OrderedDict
+from flask import Flask, jsonify, request, g, render_template
+from flask.views import View
 
 DATABASE = "images.db"
 debug = False
@@ -31,6 +32,24 @@ def db_save():
     get_db().commit()
 
 
+def get_viz_data():
+    labels = ""
+    series = {}
+    cur = get_db().cursor()
+
+    for data in cur.execute("SELECT * FROM record_counts;").fetchall():
+        label = data["label"]
+        for k, v in data.items():
+            if k != "label":
+                k = k.title()
+                series_k = series.get(k, list())
+                series_k.append(v)
+                series[k] = series_k
+        labels += "'" + label + "', "
+    labels = labels[: len(labels) - 2]
+    return (labels, series)
+
+
 @app.teardown_appcontext
 def close_connection(exception):
     db = getattr(g, "_database", None)
@@ -39,8 +58,10 @@ def close_connection(exception):
 
 
 @app.route("/")
-def home():
-    return ""
+def index():
+    data = {}
+    data["categories"], data["series"] = get_viz_data()
+    return render_template("index.html", data=data)
 
 
 @app.route("/image_downloaded")
@@ -73,7 +94,7 @@ def data_intake():
     for k in data.keys():
         sql = sql + k + " = ?, "
     sql = sql[: len(sql) - 2] + " WHERE id = " + image_id + ";"
-    
+
     try:
         cur.execute(sql, val)
         db_save()

### 2020-09-30 DATA622/M04-lda-heart.RMD
Initial commit of M04 files

---
title: "lda-heart"
author: "Raman Kannan"
date: "8/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear Discriminant Analysis

LDA attempts to identify a dimension in which the separation between classes is maximized. It can be implemented using linear algebra or the ISLR formulation. In the accompanying chapter we illustrated using iris to show that LDA can handle multi-class problems. Here, we will customize it for heart dataset, a binary classification problem.

```{r }
heart<-read.csv('heart.csv',sep=',',head=T,
                stringsAsFactors=F)
unlist(lapply(names(heart),FUN=function(x,data=heart){c(cname=x,uvalfreq=length(unique(data[[x]])))}))

nvpairsdf3<-do.call('rbind',lapply(names(heart),FUN=function(x,data=heart){c(cname=x,uvalfreq=length(unique(data[[x]])))}))

categoricalFeatures<-function(dset,ncol=7)
{
  df4<-as.data.frame(do.call('rbind',lapply(names(dset),FUN=function(x,data=dset){c(cname=x,uvalfreq=length(unique(data[[x]])))})))
  df4$uvalfreq=as.numeric(df4$uvalfreq)
  dset[,df4$uvalfreq<=ncol]
}

X<-categoricalFeatures(heart)

table(heart$target)
dim(X)

heart0<-X[X$target==0,]
heart1<-X[X$target==1,]
names(heart0)
```

## Let us compute class means for each feature

data.frames and vector manipulation -- made for R --

```{r }
heart0mean<-apply(heart0[,1:(ncol(X)-1)],2,mean)
heart1mean<-apply(heart1[,1:(ncol(X)-1)],2,mean)
```

Let us compute the mean for the entire dataset
```{r}
allclassmean<-apply(X[,1:(ncol(X)-1)],2,mean)
```
And the covariances
```{r}
S1<-cov(heart0[,1:(ncol(X)-1)])
S2<-cov(heart1[,1:(ncol(X)-1)])
```
Class frequencies
```{r}
N1<-nrow(heart0)
N2<-nrow(heart1)
```
Let us compute pooled covariance
```{r}
SWpooled<-((N1-1)*S1+(N2-1)*S2)/(N1+N2-2)
```
For LDA we need scatter between (SB) for each class.
```{r}
SB1<-N1*(heart0mean-allclassmean)*(heart0mean-allclassmean)
SB2<-N2*(heart1mean-allclassmean)*(heart1mean-allclassmean)
SB<-SB1+SB2
```
Let us predict using linear algebra
```{r}
invSWpooled<-solve(SWpooled)
invSWSBpooled<-invSWpooled*SB
SWSBpooledeigendecomp<-eigen(invSWSBpooled)
lambdas<-SWSBpooledeigendecomp$vectors[,which.max(SWSBpooledeigendecomp$values)]
as.matrix(X[,1:(ncol(X)-1)])%*% lambdas->heart.projections.pooled

```

Convert probabilities into class projections...

```{r}
rounded.heart.projections.pooled<-round(heart.projections.pooled)
heart.lda.pooled<-cbind(X,classp=heart.projections.pooled,spectral=rounded.heart.projections.pooled)
heart.lda.pooled$predicted <-ifelse(heart.lda.pooled$spectral < -1,1,0)
heart.lda.confMat.pooled<-table(  heart.lda.pooled$target,heart.lda.pooled$predicted)
heart.lda.confMat.pooled
```
Not exactly impressive...
Let us calculate the metrics 
```{r}
print(paste("Accuracy is::",
            as.numeric(sum(diag(heart.lda.confMat.pooled))/sum(heart.lda.confMat.pooled)),sep= " "))
print(paste("Error is::",
            1-as.numeric(sum(diag(heart.lda.confMat.pooled))/sum(heart.lda.confMat.pooled)),
            sep= " "))
plot(heart.lda.pooled$classp,col=as.factor(heart.lda.pooled$predicted),pch=16)
```
Let us now reimplement ISLR chapter 4 - implemented by Professor Dalal (Columbia.edu)
for the heart dataset.
```{r}
default_pos<-X[X[,9]==0,]
default_neg<-X[X[,9]==1,]
dim(default_pos)
mu_pos<-apply(default_pos[,1:8],2,mean)
mu_neg<-apply(default_neg[,1:8],2,mean)
sigma_pos<-cov(default_pos[,1:8])
sigma_neg<-cov(default_neg[,1:8])
n_pos<-dim(default_pos)[1]
n_neg<-dim(default_neg)[1]
sigma_all<-((n_pos-1)*sigma_pos+(n_neg-1)*sigma_neg)/(n_pos+n_neg-2) #pooled Cov matrix-
mu<-cbind(mu_pos,mu_neg)
pi.vec <- rep(0,2) # why not just pi?
pi.vec[1] <- sum((heart$target==0))/nrow(heart)
pi.vec[2] <- sum((heart$target==1))/nrow(heart)

my.lda <- function(pi.vec,mu,Sigma,x){
x.dims <- dim(x)
n <- x.dims[1]
Sigma.inv <- solve(Sigma) #Find inverse of Sigma
out.prod <- rep(0,n) #all items initiated to be negative
# equation 4.19 from ISLR Sixth Printing Page 157
discrim.pos <- apply(x,1,function(y) y %*% Sigma.inv %*% mu[,1]
- 0.5*t(mu[,1]) %*% Sigma.inv %*% mu[,1] + log(pi.vec[1])) # target == 0
discrim.neg <- apply(x,1,function(y) y %*% Sigma.inv %*% mu[,2]
- 0.5*t(mu[,2]) %*% Sigma.inv %*% mu[,2] + log(pi.vec[2]))# target == 1 
out.prod[discrim.neg >= discrim.pos] <- 1
out.prod.df<-data.frame(posp=discrim.pos,negp=discrim.neg,prod=out.prod)
return(out.prod.df)
}
```
Let us run LDA on heart dataset using Sid's implementation.

```{r}
heart_default_all<-my.lda(pi.vec,mu,sigma_all,X[,1:8])
hda<- heart_default_all
hda$posprob<-hda$posp/(hda$posp+hda$negp)
hda$negprob<-hda$negp/(hda$posp+hda$negp)
#produce confusion matrices
table(X$target,heart_default_all$prod)

cfmx<-caret::confusionMatrix(table(X$target,heart_default_all$prod))
```
Now let us run LDA using standard implementation in the MASS package
```{r}
require(MASS)
lda_model<-lda(target~.,data=X,type='response')
lda_pred<-predict(lda_model,X[,1:8])
table(X$target,lda_pred$class)
cfmx_mass<-caret::confusionMatrix(table(X$target,lda_pred$class))

dfcomp<-data.frame(sid=cfmx$byClass,mass=cfmx_mass$byClass)
dfcomp
```
```{r}
require(ROCR)
pred <- prediction(lda_pred$posterior[,2], X$target)
AUC<-performance(pred,"auc")
perf <- performance(pred,"tpr","fpr")
plot(perf,colorize=TRUE)
text(0.8,0.2,paste("AUC=",round(AUC@y.values[[1]],4),sep=''))

```

### 2020-09-30 DATA622/M05-kNN.rmd
Initial commit

---
title: "kNN Implementation"
author: "Raman Kannan"
date: "8/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

We will continue with Instance based methods. These algorithms do NOT compute
probability distribution.

kNN works as follows

compute the distance between the given instance and all the known instance.
Sort by distance and assign the most frequent class of the k nearest neighbors.

```{r}
euclideanDist <- function(a, b){
  d = 0
  for(i in c(1:(length(a)) ))
  {
    d = d + (a[[i]]-b[[i]])^2
  }
  d = sqrt(d)
  return(d)
}
```
Let us write another function to iterate over the test set 
and over the training set for each observation in the test set.
```{r}
knn_predict2 <- function(test_data, train_data, k_value, labelcol){
  print(k_value)
  print(labelcol)
  pred <- c()  #empty pred vector 
  #LOOP-1
  for(i in c(1:nrow(test_data))){   #looping over each record of test data
    eu_dist =c()          #eu_dist & eu_char empty  vector
    eu_char = c()
    good = 0              #good & bad variable initialization with 0 value
    bad = 0

    #LOOP-2-looping over train data 
    for(j in c(1:nrow(train_data))){

      #adding euclidean distance b/w test data point and train data to eu_dist vector
      eu_dist <- c(eu_dist, euclideanDist(test_data[i,-c(labelcol)], train_data[j,-c(labelcol)]))

      #adding class variable of training data in eu_char
      eu_char <- c(eu_char, as.character(train_data[j,][[labelcol]]))
      #print(i,j,as.character(train_data[j,][[labelcol]])))
    }

    eu <- data.frame(eu_char, eu_dist) #eu dataframe created with eu_char & eu_dist columns

    eu <- eu[order(eu$eu_dist),]       #sorting eu dataframe to gettop K neighbors
   # print(k_value)
    eu <- eu[1:k_value,]               #eu dataframe with top K neighbors

    head(eu[1:k_value,])
    tbl.sm.df<-table(eu$eu_char)
    cl_label<-  names(tbl.sm.df)[[as.integer(which.max(tbl.sm.df))]]

    pred <- c(pred, cl_label)
  }
  return(pred) #return pred vector
}

```
Let us also write a utility function for accuracy.
```{r}
accuracy <- function(test_data,labelcol,predcol){
  correct = 0
  for(i in c(1:nrow(test_data))){
    if(test_data[i,labelcol] == test_data[i,predcol]){ 
      correct = correct+1
    }
  }
  accu = (correct/nrow(test_data)) * 100  
  return(accu)
}
```
Now let us load and prepare training/test set from the heart dataset.

We will call the predict function we wrote above.
```{r}
#load data
file<-'heart.csv'
heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)
head(heart)
catheart<-heart[,c(2,3,6,7,9,11,12,13,14)]

knn.df<-catheart
labelcol <- 9 # for iris it is the fifth col 
predictioncol<-labelcol+1
# create train/test partitions
set.seed(43)
n<-nrow(knn.df)
knn.df<- knn.df[sample(n),] # randomize

train.df <- knn.df[1:as.integer(0.7*n),]
test.df <- knn.df[as.integer(0.7*n +1):n,]

K<-3 # number of neighbors to determine the class

predictions <- knn_predict2(test.df, train.df, K,labelcol) #calling knn_predict()

```
Computing ROCR and AUC is some what non-trivial as kNN do not compute probabilities
and results in unreliable ROCR plots.
```{r}
test.df[,predictioncol] <- predictions #
print(paste('accuracy=',accuracy(test.df,labelcol,predictioncol)))
```
Now let us label the test set using standard kNN implementation from class
package.
```{r}
library(class)
library(ROCR)

train.df <- knn.df[1:as.integer(0.7*n),]
test.df <- knn.df[as.integer(0.7*n +1):n,]

cl<-factor(train.df[,9])
knnPred<-knn(train.df[,-c(labelcol)],test.df[,-c(labelcol)], cl, k = 3, prob=TRUE)
prob<-attr(knnPred,"prob")
kp<-prediction(prob,test.df[,labelcol])
AUC<-performance(kp,"auc")
#perf <- performance(pred,"tpr","fpr")
#plot(perf,colorize=TRUE)
pred_knn<-performance(kp,"tpr","fpr")
```
```{r echo=FALSE}
plot(pred_knn,avg="threshold",colorize=T,lwd=3,main="ROCR knn")
text(0.8,0.2,paste("AUC=",round(AUC@y.values[[1]],4),sep=''))
```
```{r echo=FALSE}
TBL<-table(test.df[,labelcol],knnPred)
print(paste("accuracy=",sum(diag(TBL))/sum(TBL)))
```

### 2020-10-05 DATA622/M05-kNN.rmd
Updates to KNN files

@@
 knn_predict2 <- function(test_data, train_data, k_value, labelcol){
-  print(k_value)
-  print(labelcol)
+  #print(k_value)
+  #print(labelcol)
   pred <- c()  #empty pred vector 
@@
 
-    head(eu[1:k_value,])
+    #head(eu[1:k_value,])
     tbl.sm.df<-table(eu$eu_char)
@@
 }
-
 ```
@@
 ```
+
+```{r}
+knn_predict3 <- function(test_data, train_data, k_value, labelcol){
+  pred <- c() # Initialize an empty vector to hold the predictions
+  eu_char <- train_data[,labelcol] # Get the class labels from the training data
+  b <- train_data[,-c(labelcol)] # Drop the class label from the training data
+  for(i in c(1:nrow(test_data))){
+    a <- test_data[i,-c(labelcol)] # Pull the row from the test data
+    eu_dist <- euclideanDist(a, b) # Calculate the euclidean distance between training data and test row
+    eu <- data.frame(eu_char, eu_dist) # Put the distance and the label in a data frame
+    eu <- eu[order(eu$eu_dist),] # Sort it by the euclidean distance
+    eu <- eu[1:k_value,] # Subset the first k rows
+    tbl.sm.df<-table(eu$eu_char) # Create frequency table
+    pred <-  c(pred, names(tbl.sm.df)[[as.integer(which.max(tbl.sm.df))]]) # Add in the predicted class (highest frequency) to the predictions vector
+  }
+  return(pred) # Return the predictions
+}
+predictions2 <- knn_predict3(test.df, train.df, K,labelcol)
+```
+
+```{r}
+start_time <- Sys.time()
+n_tests <- 100
+for (i in 1:n_tests){
+  knn_predict2(test.df, train.df, K,labelcol)
+}
+end_time <- Sys.time()
+difftime(end_time, start_time)
+```
+
+```{r}
+start_time <- Sys.time()
+for (i in 1:n_tests){
+  knn_predict3(test.df, train.df, K,labelcol)
+}
+end_time <- Sys.time()
+difftime(end_time, start_time)
+```
 Computing ROCR and AUC is some what non-trivial as kNN do not compute probabilities

### 2020-10-06 DATA622/M05-kNN.rmd
Updated KNN files

@@
 n<-nrow(knn.df)
-knn.df<- knn.df[sample(n),] # randomize
+index <- sample(n)
+knn.df<- knn.df[index,] # randomize
 
@@
 ```{r}
-start_time <- Sys.time()
-n_tests <- 100
-for (i in 1:n_tests){
-  knn_predict2(test.df, train.df, K,labelcol)
-}
-end_time <- Sys.time()
-difftime(end_time, start_time)
+# start_time <- Sys.time()
+# n_tests <- 100
+# for (i in 1:n_tests){
+#   knn_predict2(test.df, train.df, K,labelcol)
+# }
+# end_time <- Sys.time()
+# difftime(end_time, start_time)
 ```
@@
 ```{r}
-start_time <- Sys.time()
-for (i in 1:n_tests){
-  knn_predict3(test.df, train.df, K,labelcol)
-}
-end_time <- Sys.time()
-difftime(end_time, start_time)
+# start_time <- Sys.time()
+# for (i in 1:n_tests){
+#   knn_predict3(test.df, train.df, K,labelcol)
+# }
+# end_time <- Sys.time()
+# difftime(end_time, start_time)
 ```

### 2020-10-10 DATA622/HW1.rmd
Initial commit

---
title: "Homework 1"
author: "Mike Silva"
date: "10/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

### Load Packages
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(kableExtra)
library(caret)
library(e1071)
library(class)
library(ROCR)
options(knitr.kable.NA = '')
```

### Read in the data

```{r}
df <- read.csv("HW1.csv", stringsAsFactors = TRUE)
```

### EDA

Let's take a look at the data:

```{r}
head(df) %>% kable() %>% kable_styling()
```
There are `r ncol(df)` columns and  `r nrow(df)` observations in the data set.  Let's look at a quick summary of the data:

```{r}
summary(df) %>% kable() %>% kable_styling()
```

`X` is a numeric variable ranging from `r min(df$X)` to `r max(df$X)`, while `Y` and `label` are categorical variables.  There are 6 values for `Y` and the observations are evenly distributed among the variable.  The `label` is not balanced.  Let's look at the `Y` by `label`:

```{r}
table(df$Y, df$label) %>% kable() %>% kable_styling()
```

It looks like that could be used to divide the data into the two label classes as the preponderance of observations with a `Y` of c are blue, and those with a `Y` of e are black.  Let's look at `X` by `label`

```{r}
df %>%
  group_by(label) %>%
  summarise(min = min(X),
            mean = mean(X),
            median = median(X),
            max = max(X), .groups = "drop") %>%
  kable() %>%
  kable_styling()
```

The `X`s are slightly lower for the blue `label` than the black. Finally, let's look at all the data visually:

```{r}
ggplot(df, aes(x=X, y=Y, color=label)) + geom_point() + scale_color_manual(values=c("#000000", "#56B4E9")) +   theme(legend.position = "none")
```

One can see that the lower `X` values of the blues coupled with the preponderance of `Y` of c being blue leads to a large cluster of blue in the lower left part of the chart.  Black dominates most of the other parts of the chart except for some blues on the extremes of `X`.

## Data Prep

Since we will be using KNN we will need to normalize `X`.  In order to get a fair comparison, we will used the same processed data to see how the algorithms preform.

```{r}
normalize <- function(x){
  return ((x - min(x)) / (max(x) - min(x)))
}

adjusted_df <- df
adjusted_df$X <- normalize(adjusted_df$X)
```

We also need to transform `Y` into dummy variables.

```{r}
adjusted_df <- adjusted_df %>%
  mutate(Y_a = ifelse(Y == "a", 1, 0),
         Y_b = ifelse(Y == "b", 1, 0),
         Y_c = ifelse(Y == "c", 1, 0),
         Y_d = ifelse(Y == "d", 1, 0),
         Y_e = ifelse(Y == "e", 1, 0)) %>%
  select(-Y)
```

In order to evaluate the ability of the of the algorithms to generalize, the data needs to be divided into training and test sets.  Since there is only `r nrow(df)` observations, I will be reserving 20% for the evaluation.

```{r}
set.seed(123)
in_training_set <- createDataPartition(adjusted_df$label, p = 0.8, list = FALSE, times = 1)
training_df <- adjusted_df[in_training_set,]
test_df  <- adjusted_df[-in_training_set,]
# These subsets will help with training and evaluation
training_df_without_label <- training_df %>% select(-label)
test_df_without_label  <- test_df %>% select(-label)
training_df_label <- training_df$label
test_df_label <- test_df$label
```

## Helper Functions

I will use the following helper functions.  This first one takes a logistic regression model and returns predictions for the `data` data.frame

```{r}
get_lr_yhat <- function(lr_model, data, label_col_name, threshold = 0.5){
  data_levels <- levels(data[[label_col_name]])
  cols_to_keep <- label_col_name != names(data)
  data <- data[,cols_to_keep]
  lr_yhat <- predict(lr_model, data, type = "response")
  lr_yhat <- as.factor(ifelse(lr_yhat <= threshold, data_levels[1], data_levels[2]))
  return(lr_yhat)
}
```

The second will take the the ground truth labels and predicted labels and create metrics for evaluation.

```{r}
evaluate_algo <- function(ground_truth, yhat, algo){
  cm <- confusionMatrix(table(ground_truth, yhat))
  cm_table <- cm$table
  tpr <- cm_table[[1]] / (cm_table[[1]] + cm_table[[4]])
  fnr <- 1 - tpr
  fpr <- cm_table[[3]] / (cm_table[[3]] + cm_table[[4]])
  tnr <- 1 - fpr
  accuracy <- cm$overall[[1]]
  for_auc <- prediction(c(yhat), ground_truth)
  auc <- performance(for_auc, "auc")
  auc <- auc@y.values[[1]]
  return(data.frame(Algo = algo, AUC = auc, ACCURACY = accuracy, TPR = tpr, FPR = fpr, TNR = tnr, FNR = fnr))
}
```

## Gathering the Model's Capacity to Learn

In order to evaluate this capacity, we will gather metrics on the algorithm's performance on the training set.

### Logistic Regression

```{r}
lr_model <- glm(label ~ ., data = training_df, family = "binomial")
training_lr_yhat <- get_lr_yhat(lr_model, training_df, "label")
capacity_to_learn <- evaluate_algo(training_df_label, training_lr_yhat, "LR")
```
### Naive Bayes

```{r}
nb_model <- naiveBayes(label ~ ., data = training_df)
training_nb_yhat <- predict(nb_model, training_df_without_label)
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(training_df_label, training_nb_yhat, "NB"))
```

### KNN

```{r}
training_knn3_yhat <- knn(training_df_without_label, training_df_without_label, training_df_label, k = 3)
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(training_df_label, training_knn3_yhat, "KNN3"))

training_knn5_yhat <- knn(training_df_without_label, training_df_without_label, training_df_label, k = 5)
capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(training_df_label, training_knn5_yhat, "KNN5"))
```

## Gathering the Model's Capacity to Generalize

### Logistic Regression

```{r}
lr_yhat <- get_lr_yhat(lr_model, test_df, "label")
test_lr_cm <- confusionMatrix(table(test_df_label, lr_yhat))
capacity_to_generalize <- evaluate_algo(test_df_label, lr_yhat, "LR")
```

### Naive Bayes

```{r}
nb_yhat <- predict(nb_model, test_df_without_label)
capacity_to_generalize <- rbind(capacity_to_generalize, evaluate_algo(test_df_label, nb_yhat, "NB"))
```

### KNN

```{r}
knn3_yhat <- knn(training_df_without_label, test_df_without_label, training_df_label, k = 3)
capacity_to_generalize <- rbind(capacity_to_generalize, evaluate_algo(test_df_label, knn3_yhat, "KNN3"))

knn5_yhat <- knn(training_df_without_label, test_df_without_label, training_df_label, k = 5)
capacity_to_generalize <- rbind(capacity_to_generalize, evaluate_algo(test_df_label, knn5_yhat, "KNN5"))
```

## Summary and Analysis

We have evaluated `r nrow(capacity_to_learn)` algorithm's capacity to learn and generalize.  The following tables summarize their performace:

**Table 1. Capacity to Learn**
```{r, echo=FALSE}
capacity_to_learn %>%  kable() %>%  kable_styling()
```

**Table 2. Capacity to Generalize**
```{r, echo=FALSE}
capacity_to_generalize %>% kable() %>% kable_styling()
```

### Discussion

In formulating our recommendations, we assume that the accuracy is very important, especially the TNR.  Since the data is imbalanced (there are more blacks than blues), getting an accurate blue prediction is a good measure of how well the algorithm preforms.  With this in mind, here's our analysis of each algorithm's performance.

**Logistic Regression** showed a fairly strong ability to learn with an overall accuracy of 73%.  It struggled to identify black cases but did well at identifying blue cases in the test data set.  It's ability to generalize lead us not to recommend using this algorithm.  The accuracy dropped to 33%, indicating that this algorithm over fit the data.

**Naive Bayes** exhibited performance similar to logistic regression.  It also had a 73% accuracy, and outperformed logistic regression's TPR and TNR.  It also showed weakness in the ability to generalize with identical performance to logistic regression.  It is for that reason we cannot recommend this algorithm.

**KNN** was well suited to give accurate predictions.  KNN3 had the highest capacity to learn out of all the algorithms analyzed. It had an accuracy of 80%.  KNN3 also indicated strength in generalization, with a accuracy of 50%.  That is a 17 percentage point improvement over logistic regression and naive Bayes.

KNN5 indicated a similar ability to learn as logistic regression and naive Bayes.  It did preform the best at detecting blues as indicated by the TNR of 83%.  Where KNN5 shined was in the ability to generalize.  It had the highest accuracy and the highest TNR of all the algorithms.

```{r, echo=FALSE}
ggplot() + 
  geom_point(data = df[in_training_set,], aes(x = X, y = Y, color = label)) + 
  geom_point(data = df[-in_training_set,], aes(x = X, y = Y, color = label), shape = 2) +
  scale_color_manual(values=c("#000000", "#56B4E9")) +
  theme(legend.position = "none")
```

KNN preformed so well because the data exhibited clustering.  This is seen in the image above where the triangles are the points that needed to be predicted and the dots are the data used to make inferences.  As previously noted, there is a large cluster of blue centered around where `X` is 20 and `Y` is c (which coincidentally is a point needing to be predicted).  There is another patch of blues where `X` is greater than 60 and `Y` is not a.  The black points dominated the chart where `X` ranged from 30 to 50.  Given the points needed to be predicted (triangles) were surrounded by peers KNN preformed well.

### Recommendation

Based on our analysis of the performance of the algorithms on the data set, we would recommend using a KNN algorithm.  KNN is well positioned to make an accurate prediction.  Based strictly on the draw that we made, setting K = 5 seems to be the best choice.  This is heavily dependent on the draw that was made in this process.  A K = 3 is a strong alternative and should be explored using additional draws of data.

### 2020-10-12 DATA698/pytorch_to_tensorflow_lite.py
Initial commit

@@ -1,0 +1,79 @@
+# Originally from https://github.com/kalaspuffar/onnx-convert-example
+# Running the script python3 pytorch_to_tensorflow_lite.py models/mnist.pth models
+import sys
+import os
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import onnx
+from collections import OrderedDict
+import tensorflow as tf
+from torch.autograd import Variable
+from onnx_tf.backend import prepare
+
+class MLP(nn.Module):
+    def __init__(self, input_dims, n_hiddens, n_class):
+        super(MLP, self).__init__()
+        assert isinstance(input_dims, int), 'Please provide int for input_dims'
+        self.input_dims = input_dims
+        current_dims = input_dims
+        layers = OrderedDict()
+
+        if isinstance(n_hiddens, int):
+            n_hiddens = [n_hiddens]
+        else:
+            n_hiddens = list(n_hiddens)
+        for i, n_hidden in enumerate(n_hiddens):
+            layers['fc{}'.format(i+1)] = nn.Linear(current_dims, n_hidden)
+            layers['relu{}'.format(i+1)] = nn.ReLU()
+            layers['drop{}'.format(i+1)] = nn.Dropout(0.2)
+            current_dims = n_hidden
+        layers['out'] = nn.Linear(current_dims, n_class)
+
+        self.model= nn.Sequential(layers)
+        print(self.model)
+
+    def forward(self, input):
+        input = input.view(input.size(0), -1)
+        assert input.size(1) == self.input_dims
+        return self.model.forward(input)
+
+print("%s" % sys.argv[1])
+print("%s" % sys.argv[2])
+
+
+# Load the trained model from file
+trained_dict = torch.load(sys.argv[1], map_location={'cuda:0': 'cpu'})
+
+trained_model = MLP(784, [256, 256], 10) #input_dims, n_hiddens, n_class
+trained_model.load_state_dict(trained_dict)
+
+if not os.path.exists("%s" % sys.argv[2]):
+    os.makedirs("%s" % sys.argv[2])
+
+# Export the trained model to ONNX
+dummy_input = Variable(torch.randn(1, 1, 28, 28)) # one black and white 28 x 28 picture will be the input to the model
+torch.onnx.export(trained_model, dummy_input, "%s/mnist.onnx" % sys.argv[2])
+
[…]

### 2020-10-15 DATA622/M06-tree.RMD
Initial commit

---
title: "M06-DTree.RMD"
author: "Raman Kannan"
date: "8/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
heart<-read.csv('heart.csv',sep=',',head=T,
                stringsAsFactors=F)
unlist(lapply(names(heart),FUN=function(x,data=heart){c(cname=x,uvalfreq=length(unique(data[[x]])))}))

nvpairsdf3<-do.call('rbind',lapply(names(heart),FUN=function(x,data=heart){c(cname=x,uvalfreq=length(unique(data[[x]])))}))

categoricalFeatures<-function(dset,ncol=7)
{
  df4<-as.data.frame(do.call('rbind',lapply(names(dset),FUN=function(x,data=dset){c(cname=x,uvalfreq=length(unique(data[[x]])))})))
  df4$uvalfreq=as.numeric(df4$uvalfreq)
  dset[,df4$uvalfreq<=ncol]
}

X<-categoricalFeatures(heart)

table(heart$target)
dim(X)

```

```{r, echo=FALSE}
library(rpart)
library(rpart.plot)
library(ROCR)

isModelReliable<-function(cfmx)
{	cfmx$overall[['AccuracyPValue']]<0.05;
 }

isUnderfitting<-function(cfmx)
{	cfmx$overall[['Accuracy']]<0.60;
 }

isOverfitting<-function(tr,tst)
{	tst$overall[['Accuracy']]/tr$overall[['Accuracy']]<0.75;
 }

 getSensitivity<-function(cfmx) cfmx$byClass[['Sensitivity']]

getAccuracy<-function(cfmx) cfmx$overall[['Accuracy']]

getSpecificity<-function(cfmx) cfmx$byClass[['Specificity']]

getMetrics<-function(actual_class,predicted_response)
{
X=list()
if ( require(ROCR) ) {
auc_1=prediction(predicted_response,actual_class)
prf=performance(auc_1, measure="tpr",x.measure="fpr")
slot_fp=slot(auc_1,"fp")
slot_tp=slot(auc_1,"tp")

fpr=unlist(slot_fp)/unlist(slot(auc_1,"n.neg"))
tpr=unlist(slot_tp)/unlist(slot(auc_1,"n.pos"))

auc<-performance(auc_1,"auc")
AUC<-auc@y.values[[1]]
X=list(prf=prf,fpr=fpr,tpr=tpr,auc=AUC)
}
X
}
X$target=as.factor(X$target)

rpart.model<-rpart(target~.,data=X,minsplit=3)
rpart.plot(rpart.model)

```

That is a nice chart but several questions come to mind....
What are the numbers in each node?
What do they represent?
There are three numbers in the root node 1,0.54,100%

Here 1 is the class.
54% of the observations belong to class 1 and we can confirm this by 
```{r} table(X$target)/nrow(X)```
The third number 100% says, all observations are accounted for in this node, obviously so for the
root node.

the test at this node is cp < 1.

Let us take the cp<1 branch to the left and here we find 0,0.27,47%.
This node is class 0 node.And, 27% of the observations will end up being class 1.
The node id is always the majority class. And these are easily confirmed 
`r table(X[X$cp<1,]$target)/sum(table(X[X$cp<1,]$target)) ` and this node represents
`r nrow(X[X$cp<1,])/nrow(X)` of the dataset.
```{r}
table(X[X$cp<1,]$target)/sum(table(X[X$cp<1,]$target)) #class probabilities 
nrow(X[X$cp<1,])/nrow(X) # proportion of the dataset
#or we can follow the next node
table(X[X$cp<1&X$ca>=1,]$target)/sum(table(X[X$cp<1&X$ca>=1,]$target))
# and confirm the majority class for this node is 0 and proportion of class 1 to be

```

Let us go to the next node the test here is ca >= 0
If ca < 1 we should get `r table(X[X$cp<1&X$ca<1,]$target)/sum(table(X[X$cp<1&X$ca<1,]$target)) `
The majority class here is 1. And the proportion of the dataset in this node is
`r nrow(X[X$cp<1&X$ca<1,])/nrow(X) `
```{r}
table(X[X$cp<1&X$ca<1,]$target)/sum(table(X[X$cp<1&X$ca<1,]$target))
nrow(X[X$cp<1&X$ca<1,])/nrow(X)

# if we follow ca>0
table(X[X$cp>0,]$target)/nrow(X[X$cp>0,]) # we get 79% for class 1
nrow(X[X$cp>0,])/nrow(X) # for 52% of the dataset
```
Confirm the numbers are in agreement for 3 more nodes..
Node ID is always the majority class.
The second number is the probability for class 1.
and the third number is % of observations out of the whole dataset represented by the node.

##Why did the tree start with the feature 'cp'?
##How did it find that is the right strategy?

Let us reduce the number of features and run a small model to understand 
these things. With this large a dataset we may never get to the point.

```{r}
library(sqldf)

s1<-sqldf('select count(*) from X where cp < 1')
head(s1)

nrow(X[X$cp<1,])/nrow(X)

table(X[X$cp<1,]$target)

table(X[X$cp<1,]$target)/sum(table(X[X$cp<1,]$target))
```
There are 143 with cp<1 out of ```{r} nrow(X) ``` let us start with a smaller 3 feature dataset ...
```{r}
smh<-X[,c(2:4,9)]
smh$target<-as.factor(smh$target)
smh_rpart<-rpart(target~.,data=smh,minsplit=2)
rpart.plot(smh_rpart,main=' 3 feature dataset to illustrate gini_index')

```

```{r}

smhx<-sqldf('select * from smh where cp <1 and fbs =1')

smhx

gini<-function(dset,cname,label)
{
cnames<-unique(dset[[label]])
vprop<- as.matrix(as.numeric(table(dset[[cname]])/sum(table(dset[[cname]]))))
 T1<-table(dset[[cname]],as.vector(dset[[label]]))
 T1mx<-as.matrix(T1)
 T11<-as.matrix(t(apply(T1mx,1,FUN=function(v){v/sum(v)})))
 pT1<-as.data.frame(cbind(T1mx,vprop,T11))
nclass<-length(unique(dset[[label]]))
pT1x<-pT1[,(nclass+1+1):(1+2*nclass)]
 pT2<-cbind(pT1,t(apply(pT1[,(nclass+1+1):(1+2*nclass)],1,FUN=function(v) v*v)))
pT3<-cbind(pT2,apply(pT2,1,FUN=function(v)(v[nclass+1]*(1-sum(v[(1+2*nclass+1):(1+3*nclass)])))))
pT3.df<-as.data.frame(pT3)
colnames<-c(paste('f_',cnames,sep=''),'valprop',paste('p_',cnames,sep=''),
paste('sqp_',cnames,sep=''),'gini')
names(pT3.df)<-colnames
pT3.df$feature=cname
pT3.df$label=label
pT3.df$feature_value=row.names(T1)
pT3.df
 }

compute_gini<-function(dset,labelCol)

{

ginimx<-do.call('rbind',lapply(names(dset)[names(dset)!=labelCol],FUN=function(x,df=dset,label=labelCol)gini(df,x,label)))

ginimx
}

ginimx<-compute_gini(smh,'target')

ginimx

ginif<-sqldf('select feature,sum(gini) as tgini from ginimx group by feature')

ginif[which.min(ginif$tgini),1]

```
Does it work recursively?

let us check on the full dataset for the partition when cp>0

Recall the tree we rendered initially.

```{r}
#catheart<-categoricalFeatures(heart)
bigkahuna<-compute_gini(X[X$cp>0,],'target')
ginit<- sqldf('select feature, sum(gini) tgini from bigkahuna group by feature')
ginit[which.min(ginit$tgini),c('feature')]
```

So we now have a systematic method to partition a given dataset
by the feature that most isolate the class using GINI index.

### 2020-10-15 DATA622/heart-tree.R

@@ -1,0 +1,30 @@
+
+file<-'c:/Users/rk215/data/heart.csv' # change this to run in your env
+heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)
+head(heart)
+catheart<-heart[,c(2,3,6,7,9,11,12,13,14)]
+require(rpart)
+require(rpart.plot)
+head(catheart,20)
+catheart$target<-as.factor(catheart$target)
+set.seed(43)
+tridx<-sample(1:nrow(catheart),0.7*nrow(catheart),replace=F)
+trdata<-catheart[tridx,]
+tstdata<-catheart[-tridx,]
+trmodel.rpart<-rpart(target~.,data=trdata,minsplit=2)
+rpart.plot(trmodel.rpart)
+table(trdata$target)/nrow(trdata) # compare this to the root node lables
+
+predicted.trmodel.rpart<-predict(trmodel.rpart,trdata[,-9],type='class')
+TBL<-table(trdata[,9],predicted.trmodel.rpart)
+
+caret::confusionMatrix(TBL)
+
+predicted.tstmodel.rpart<-predict(trmodel.rpart,tstdata[,-9],type='class')
+TSTTBL<-table(tstdata[,9],predicted.tstmodel.rpart)
+
+caret::confusionMatrix(TSTTBL)
+
+#Summary
+#Individual trees are prone to over-fitting so performance degrades over never seen 
+# before data as we see here 

### 2020-10-15 DATA622/species-tree.R

@@ -1,0 +1,82 @@
+# create the birds dataset
+birds<-data.frame(nlegs=rep(2,5),can_fly=c(0,1,1,0,1),height=c(25,40,20,150,10),
+color=c('black','black','blue','black','brown')
+)
+birds2<-cbind(birds,c('chicken','vulture','parrot','ostrich','sparrow'))
+chickencolors<-c('black','white','red','mixed')
+vulturecolors<-c('grey','black','white')
+parrotcolors<-c('teal','green','blue','mixed','pink')
+ostrichcolors<-c('grey','black')
+sparrowcolors<-c('dark cement','brown')
+names(birds2)<-c('nlegs','can_fly','height','color','species')
+birds2
+#chicken heights mu=25 sd=6
+hchicken<-sample(rnorm(10,25,6),5)
+#vulture heights mu=40 sd=4
+hvulture<-sample(rnorm(10,40,4),5)
+hparrot<-sample(rnorm(10,20,2),5)
+hostrich<-sample(rnorm(10,150,20),5)
+hsparrow<-sample(rnorm(10,10,1),5)
+#bigger dataset
+
+#chicken heights mu=25 sd=6
+hchicken<-sample(rnorm(10,25,6),5)
+#vulture heights mu=40 sd=4
+hvulture<-sample(rnorm(10,40,4),5)
+hparrot<-sample(rnorm(10,20,2),5)
+hostrich<-sample(rnorm(10,150,20),5)
+hsparrow<-sample(rnorm(10,10,1),5)
+
+cdset<-rbind(birds2,data.frame(nlegs=rep(2,5),can_fly=rep(0,5), height=hchicken,
+color=sample(chickencolors,5,replace=T),species=rep('chicken',5)
+))
+cdset
+
+#bigger dataset
+cdset<-rbind(birds2,data.frame(nlegs=rep(2,5),can_fly=rep(0,5), height=hchicken,
+color=sample(chickencolors,5,replace=T),species=rep('chicken',5)),
+data.frame(nlegs=rep(2,5),can_fly=rep(1,5), height=hvulture,
+color=sample(vulturecolors,5,replace=T),species=rep('vulture',5)),
+data.frame(nlegs=rep(2,5),can_fly=rep(1,5), height=hparrot,
+color=sample(parrotcolors,5,replace=T),species=rep('parrot',5)),
+data.frame(nlegs=rep(2,5),can_fly=rep(0,5), height=hostrich,
[…]

### 2020-10-16 DATA622/M07-svm.rmd
Initial commit

---
title: "M07-SVM"
author: "Raman Kannan"
date: "9/28/2020"
output: 
html_document: M07-svm.html
df_print: paged
pdf_document: default
---

```{r Startup}

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

#Load input data heart.csv and ecoli.csv
```{r MainCode}
unixpath<-"/home/data/ML.Data/heart.csv"
unixecolipath<-"/home/data/ML.Data/ecoli.csv"
path<-"c://Users/rk215/Data/heart.csv"
ecolipath<-"c://Users/rk215/Data/ecoli.csv"
heart<-read.csv(path,header=TRUE,sep=',',stringsAsFactors=FALSE)
ecoli<-read.csv(ecolipath,head=T,sep=',',stringsAsFactors=F)

```

## Review the data, quick and easy EDA
View a few records (head/tail), dimension, number of rows and records
and the variable names.
Everything is an object and mostly a list. We can index into a list to 
access the elements in the list. R index runs from 1 to the length.
```{r}
head(heart)
```
### unix head and tail commands
R commands *head* and *tail* are like the Unix head and tail shell commands
These commands output a few records from the beginning or at the end of the file.
```{r}
tail(heart)
```

let us change the name of the first column

```{r}
names(heart)[[1]]<-'age'
```
Always trust but verify that our indexing worked...
```{r}
names(heart)
```
We have already performed EDA when we ran glm. 
```{r}
library(plotly)
library(e1071)
library(caret)
library(pROC)

```

## Let us run
Let us run SVM from e1071 package and try to
understand each variable and how important they are in 
determining the target variable.

we will **NOT** split the data as this is a trial run
and this run is to develop a rough draft script for
any classification exercise.
```{r}
heart$target<-factor(heart$target)
svm_model<-svm(target~.,data=heart, probability=TRUE)
svm_model

```

```{r}
svm_predict<-predict(svm_model,heart[,-c(14)],probability=TRUE)
probs<-attr(svm_predict,"prob")[,"1"]
predclass<-ifelse(probs>0.5,1,0)
table(heart$target==predclass)
(cfmx<-caret::confusionMatrix(table(heart$target,predclass)))
```
### Our model can learn ....
As evidenced by the F1 ```{cfmx$byClass[['F1']]}``` and the Balanced Accuracy ```{cfmx$byClass[['Balanced Accuracy']]}``` Now we should compute testing error -so called- generalization error...
To determine the generalization error, we split the data into training and testing dataset.
We dont tune as a matter of principle so we do not need the validation set.
Then we train on the training set and predict the target for the testing set.
That will provide us a window into the generalization capability for the learner.
This capability to generalize is all that matters for the business. So let us do it.
```{r}
set.seed(43)
tstidx<-sample(1:nrow(heart),0.30*nrow(heart),replace=F)
trdata<-heart[-tstidx,]
tsdata<-heart[tstidx,]

#svm(target~.,data=heart, probability=TRUE)

svm_train_model<-svm(target~.,data=trdata,probability=TRUE)
svm_train_predict<-predict(svm_train_model,trdata[,-c(14)],probability=TRUE)
train_probs<-attr(svm_train_predict,"prob")[,"1"]
train_predclass<-ifelse(train_probs>0.5,1,0)
tr_tbl<-table(trdata$target,train_predclass)
(tr_cfmx<-caret::confusionMatrix(tr_tbl))

cdf<-data.frame(all=cfmx$byClass,tr=tr_cfmx$byClass)
cdf
```
The performance measures are very similar therefore we can conclude the data partition has no impact on 
the distribution or the model performance and the model is stable. Now we can evaluate the performance on the test set.
```{r}
svm_tst_predict<-predict(svm_train_model,tsdata[,-c(14)],probability=TRUE)
tst_probs<-attr(svm_tst_predict,"prob")[,"1"]
tst_predclass<-ifelse(tst_probs>0.5,1,0)
tst_tbl<-table(tsdata$target,tst_predclass)
(tst_cfmx<-caret::confusionMatrix(tst_tbl))

cdf<-cbind(cdf,ts=tst_cfmx$byClass)
cdf
```
# Let us now run the SVM on the attributed we identified with VIF
```{r}
Y<-heart[[14]]
vif_col<-c(2,3,6,7,9,11,12,13,14)
vif_heart<-heart[,vif_col]
vif_trdata<-trdata[,vif_col]
vif_tsdata<-tsdata[,vif_col]
head(vif_heart)
head(Y)

svm_vif_tr_model<-svm(target~.,data=vif_trdata,probability=TRUE)
svm_vif_tr_predict<-predict(svm_vif_tr_model,vif_trdata[,-c(14)],probability=TRUE)
vif_train_probs<-attr(svm_vif_tr_predict,"prob")[,"1"]
vif_train_predclass<-ifelse(vif_train_probs>0.5,1,0)
vif_tr_tbl<-table(vif_trdata$target,vif_train_predclass)
(vif_tr_cfmx<-caret::confusionMatrix(vif_tr_tbl))
(cdf<-cbind(cdf,vif_tr=vif_tr_cfmx$byClass)) # this represents our model's ability to learn -- training error
# is interpreted as underfitting when < 0.75 I use 0.85 
# remedy is to consider more complex model 
```
**Why is the performance with VIF predictors slightly better?**  
Next we will predict the target for our test data.
And the performance (or error) calculated on the test data represents generalization error for this model.
```{r}
svm_vif_tst_predict<-predict(svm_vif_tr_model,vif_tsdata[,-c(9)], probability=TRUE)
svm_vif_tst_predict_probs=attr(svm_vif_tst_predict, "prob")[,"1"]
svm_roc_obj<-roc(vif_tsdata[,c(9)],svm_vif_tst_predict_probs)
svm_auc<-auc(svm_roc_obj)

```
Now let us plot ROC Curve
```{r}
plot.new()
plot.window(xlim=c(1,0), ylim=c(0,1), xaxs='i', yaxs='i')
axis(1, las=1)
axis(2, las=1)
abline(1,-1, col="black", lwd=0.5)
box()
title(main="ROC", xlab="Specificity", ylab="Sensitivity")
lines(svm_roc_obj, col="black", las=2)
text(x=0.2,y=0.2,paste("SVM_AUC=",round(svm_auc,6),sep=''))
```
Area Under the Curve (AUC) for ROC Curve is ```{svm_auc}```.
Given the AUC, SVM has outperformed all the classifiers we have evaluated so far.

SVM by far has some attractive properties.
Adding new observations to the left or right of the computed support vectors
does not change the support vectors and our inferences are stable as new
data arrives.
SVM needs to carry the vectors nothing more -- not the dataset as we did in
the case of kNN.
SVMs are most explainable determined by directly by the data.

The decision boundary for a SVM is linear (recall it is straight line)
SVMs are most known for its customizability using kernel trick.
Here let us look at an example from ESL book which demonstrates
the power of SVM when the boundary is not linear.
Download this data, 
https://web.stanford.edu/~hastie/ElemStatLearn/datasets/ESL.mixture.rda
from the ESL book site.

```{r}
load(url("https://web.stanford.edu/~hastie/ElemStatLearn/datasets/ESL.mixture.rda"))
plot(ESL.mixture$x,col=ESL.mixture$y+3)
data_svm<-data.frame(y=factor(ESL.mixture$y),ESL.mixture$x)

 data.svm<-svm(factor(y)~.,data=data_svm,scale=FALSE,kernel="radial",cost=15)
 plot(data.svm,data_svm)

```

Clearly this is not a linear boundary but SVM does this even though the data is not
separable. The performance of this model may be computed just as we have been
doing and that is left as an exercise, not for credit though.

For Software Engineers, kernel technique, just passing in 
method as a parameter, is not new.

### 2020-10-16 DATA622/M09-Bootstrap-CrossValidation-NaiveBayes.RMD
Initial commit

---
title: "Bootstrapping-naiveBayes"
author: "Raman Kannan"
date: "10/5/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bootstrapping

Let us look at bootstrap -- The metaphor is, when we fall into a ditch, we can climb if we are wearing a shoes with bootstrap.No can do if you are wearing a slip on or other kinds of footwear without a strap. We learned that classification exercise often suffers bias and variance. Together, bias and variance, contribute to errors in modeling.
From wiki--In general, bootstrapping usually refers to a self-starting process that is supposed to proceed without external input. In computer technology the term (usually shortened to booting) usually refers to the process of loading the basic software into the memory of a computer after power-on or general reset, especially the operating system which will then take care of loading other software as needed.

## Bias
The Bias can be one of the three different kinds  
* Bias in the statistics world. Here the expected value of sample statistics is different from population parameter. 
* In Linear Regression, the intercept is also known as bias, and 
* in M/L, when underfitting (model is too simple and is unable to learn the patterns in the data) occurs it is attributed to bias.

## Variance
The variance, reflects how a given observation is classified differently by the model when trained with different training sets.

## Bias/Variance: The Tradeoff
Bias and Variance cannot be simultaneously reduced to zero. If we reduce Bias, variance will increase and vice versa.

## Established Bias reduction strategies

### Resampling
Just as person in a ditch uses the available shoe-lace to climb out of the hole, a dsc with limited data, can use that limited data by resampling many times with replacement and overcome poor performance due to lack of data.

Again we need to justify why we are engaging in such trials by establishing base-line performance and demonstrating performance can be improved with such
experiments.

### Base line performance of NB and GLM

```{r}
library('e1071')
file<-'c://Users/rk215/Data/heart.csv'
heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)
head(heart)
catheart<-heart[,c(2,3,6,7,9,11,12,13,14)]

set.seed(43)
trdidx<-sample(1:nrow(catheart),0.7*nrow(catheart),replace=F)
trcatheart<-catheart[trdidx,]
tstcatheart<-catheart[-trdidx,]

nbtr.model<-naiveBayes(target~.,data=trcatheart)
#str(nbtr.model)
object.size(nbtr.model) #11096

nbtr.trpred<-predict(nbtr.model,trcatheart[,-c(9)],type='raw')
nbtr.trclass<-unlist(apply(round(nbtr.trpred),1,which.max))-1
nbtr.trtbl<-table(trcatheart[[9]], nbtr.trclass)
tr.cfm<-caret::confusionMatrix(nbtr.trtbl)
tr.cfm
#tr.cfm$overall['Accuracy']

start_tm <- proc.time() 
#bootstrap here
#library(parallel)

#  mclapply = system.time(
#      mclapply(f, mc.cores=4, read_clean_write)
# )
df<-trcatheart
runModel<-function(df) {naiveBayes(target~.,data=df[sample(1:nrow(df),nrow(df),replace=T),])}
lapplyrunmodel<-function(x)runModel(df)

system.time(models<-lapply(1:100,lapplyrunmodel))

object.size(models)
#system.time(
boot_preds<-lapply(models,FUN=function(M,D=trcatheart[,-c(9)])predict(M,D,type='raw'))

boot_trcfm<-lapply(boot_preds,FUN=function(P,A=trcatheart[[9]])
{pred_class<-unlist(apply(round(P),1,which.max))-1
  pred_tbl<-table(pred_class,A)
  pred_cfm<-caret::confusionMatrix(pred_tbl)
  pred_cfm
})

c(boot_trcfm[[1]]$overall,boot_trcfm[[1]]$byClass)

as.numeric(c(boot_trcfm[[1]]$overall,boot_trcfm[[1]]$byClass))

boot_trcfm[[1]]$overall

class(boot_trcfm[[1]]$overall)
names(boot_trcfm[[1]]$overall)

#boot.perf<-do.call('rbind',lapply(boot_trcfm,FUN=function(cfm)c(acc=cfm$overall[[1]],pvalue=cfm$overall[[6]],sens=cfm$byClass[[1]],spec=cfm$byClass[[2]])))

boot.perf<-as.data.frame(do.call('rbind',lapply(boot_trcfm,FUN=function(cfm)c(cfm$overall,cfm$byClass))))

boot.tr.perf<-apply(boot.perf[boot.perf$AccuracyPValue<0.01,-c(6:7)],2,mean)
boot.tr.perf.var<-apply(boot.perf[boot.perf$AccuracyPValue<0.01,-c(6:7)],2,sd)

boot.tr.perf.var/boot.tr.perf

boot_tm<-proc.time()-start_tm

nbtr.tspred<-predict(nbtr.model,tstcatheart[,-c(9)],type='raw')
#roc.nbtr.tspred<-unlist(apply(nbtr.tspred,1,which.max)) # this returns the position
#roc.nbtr.tspred<-apply(nbtr.tspred,1,FUN=function(v)v[which.max(v)]) # this returns the value
#https://stackoverflow.com/questions/47883541/how-can-i-implement-roc-curve-analysis-for-naive-bayes-classification-algorithm
roc.nbtr.tspred<-nbtr.tspred[,2]
nbtr.tsclass<-unlist(apply(round(nbtr.tspred),1,which.max))-1
nbtr.tstbl<-table(tstcatheart[[9]], nbtr.tsclass)
tst.cfm<-caret::confusionMatrix(nbtr.tstbl)
tst.cfm
#tst.cfm$overall['Accuracy']

tstboot_preds<-lapply(models,FUN=function(M,D=tstcatheart[,-c(9)])predict(M,D,type='raw'))

boot_tstcfm<-lapply(tstboot_preds,FUN=function(P,A=tstcatheart[[9]])
{pred_class<-unlist(apply(round(P),1,which.max))-1
  pred_tbl<-table(pred_class,A)
  pred_cfm<-caret::confusionMatrix(pred_tbl)
  pred_cfm
})

tstboot.perf<-as.data.frame(do.call('rbind',lapply(boot_tstcfm,FUN=function(cfm)c(cfm$overall,cfm$byClass))))

boot.tst.perf<-apply(boot.perf[tstboot.perf$AccuracyPValue<0.01,-c(6:7)],2,mean)
boot.tst.perf.var<-apply(tstboot.perf[tstboot.perf$AccuracyPValue<0.01,-c(6:7)],2,sd)
```

### Cross Validation

Splitting a dataset into two disjoint sets (70/30) and using 70% for training and 30% for testing is not very smart. Because, we had 100 observations to
train with and test with. We end up training with 70% and testing with a paltry 30%. We have to find a way to overcome this under-utilization of data, the most critical resource, in our line of work. Cross Validation is an attempt to over come this deficiency.

```{r creating_folds}

N<-nrow(trcatheart)
NF=10
folds<-split(1:N,cut(1:N, quantile(1:N, probs = seq(0, 1, by =1/NF))))
length(folds)
lapply(folds,length)
ridx<-sample(1:nrow(trcatheart),nrow(trcatheart),replace=FALSE)

cv_df<-do.call('rbind',lapply(folds,FUN=function(idx,data=trcatheart[ridx,]) {
  m<-naiveBayes(target~.,data=data[-idx,])
   p<-predict(m,data[idx,-c(9)],type='raw')
   pc<-unlist(apply(round(p),1,which.max))-1
  pred_tbl<-table(pc,data[idx,c(9)])
  pred_cfm<-caret::confusionMatrix(pred_tbl)
  list(fold=idx,m=m,cfm=pred_cfm)
  }
))

```
**cv_df** now has the folds, models and corresponding **caret::confusionMatrix** which has all the metrics.
We can extract the metrics into a data.frame and average them as we did before.
```{r averaging_the_cv}

cv_df<-as.data.frame(cv_df)

lapply(cv_df$cfm,FUN=function(cfm)c(cfm$overall,cfm$byClass))
tstcv.perf<-as.data.frame(do.call('rbind',lapply(cv_df$cfm,FUN=function(cfm)c(cfm$overall,cfm$byClass))))

cv.tst.perf<-apply(tstcv.perf[tstcv.perf$AccuracyPValue<0.01,-c(6:7)],2,mean)
cv.tst.perf
boot.tst.perf
tst.cfm
cv.tst.perf.var<-apply(tstcv.perf[tstcv.perf$AccuracyPValue<0.01,-c(6:7)],2,sd)
```
## Bagging vs CV

In comparison to the bootstrap metrics ```{boot.tst.perf}``` a 10-Fold cross validation yields the following metrics 
```{cv.tst.perf}```. Cross Validation has improved the base Accuracy and it decreased for bootstrap.

To be honest, I had expected bootstrap to outperform both cross-validation and base model.

In Data expeditions, you go where the data takes you. Double/triple validation is required but data leads you
whereever it leads you. Let there not be any temptation to second guess data. Data is verily ***The Light of Truth***

### 2020-10-19 DATA622/M07-svm.rmd
Updated M07-SVM

@@
 ```{r MainCode}
-unixpath<-"/home/data/ML.Data/heart.csv"
-unixecolipath<-"/home/data/ML.Data/ecoli.csv"
-path<-"c://Users/rk215/Data/heart.csv"
-ecolipath<-"c://Users/rk215/Data/ecoli.csv"
+unixpath<-"heart.csv"
+unixecolipath<-"ecoli.csv"
+path<-"heart.csv"
+ecolipath<-"ecoli.csv"
 heart<-read.csv(path,header=TRUE,sep=',',stringsAsFactors=FALSE)

### 2020-11-02 DATA622/M10-gd.RMD
Initial commit

---
title: "Estimation-by-Gradient-Descent"
author: "Raman Kannan"
date: "10/5/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
# there are producedNAN so I suppressed after confirming the estimates are same in R (in windows, same machine)
# and other hw platforms.
```
# Central Task in Machine Learning

In Supervised Learning, our goal has been to build a model given a dataset to determine class labels (in classification problem) or estimate a numerical value (in regression), without knowing the true distribution. The predictions are not always identical to the true distribution.
The resulting differences is known as cost and it is an important measure of how suitable the machine learning model is. In Least Squares, MSE is the cost and by minimizing the MSE we arrive at a suitable/performant model. In Classification, we define a Loss function, L, a function of (betas,weights), which aggregates the misclassifications over all observations. We seek to minimize the error by iteratively tweaking the parameters (betas and weights).

In calculus, if a function is differentiable, the minima or the maxima of the function can be found using the first derivative. The first derivative is zero at the extremas. And the second derivative is -ve (when it is max) and +ve (when it is min). In optimization, formulating a Loss function is an important task. We find the extremas of the Loss function **iteratively** by tweaking the betas and weights in the cost function. Gradient Descent is thus a first-order iterative optimization algorithmm for finding the min/max of a function. In each iteration, we compute the gradient and update the betas by -gradient to find the minimum and by +gradient to find the maximum.

In Machine Learning, supervised learning task we are seeking to minimize the loss, and therefore we apply -1*gradient in each step. In the abstract this takes the algebraic form,

θ = θ − η · ∇θJ(θ)

where θ are the parameters,  η is the learning rate (the hyperparameter for GD), ∇θ is the gradient operator and J(θ) is the Loss function.
Most important, we are differentiating w.r.t θ and not the independent variable. The dataset is constant. We are seeking to find the weights that minimize the Loss given that dataset.

## Numerical Optimization

When  **an analytical solution** is not available we have to consider other strategies. Such iterative optimization methods have been around much longer than machine learning. Gradient Descent is one such optimization method.

Classic example is the sigmoid function used logistic regression, it does not have an analytical solution.  
# function template to implement gradient descent

*  Formulate a cost function (J(θ))-- must be differentiable (a convex function is differentiable)
*  initialize random weights (θ), set them all to zero and choose a learning rate, η (0.01,0.001,0.0001 or even 0.00001)
*  evaluate cost function (J(θ))
*  update weights with gradient ∇θJ(θ), according to the equation θ = θ − η · ∇θJ(θ)
*  repeat evaluate-update steps until convergence

Note if the learning rate is too large the algorithm will jump over minima. One must choose sufficiently small learning rate, with care.

In general, given F, the mapping function (x->y), where x are independent variables and y is the dependent variable, for regression problems the Loss Function often is squared-error (y-F)**2 or the absolute 
error |y-F|. In the case of (-1,1) classification problems, L is often the log-likelihood, log(1+exp(-2yF))

### Variations of Gradient Descent (GD)

Updating parameters using the loss function over the entire training set can be compute intensive. This is called Batch GD.
This method is stable but very slow.
Updating the parameters using the loss function using a single randomly chosen observation is fast and it is called Stochastic GD. As expected this is very noisy and jumps around. And sometimes may result in sub-optimal solution.
A compromise is the mini-batch GD, in which Loss Function and the gradient are evaluated over a random subset, neither the entire set nor a single observation. The mini=batch is often used in Deep Learning.

## Cost Function for Logistic Regression

The log of a function behaves similar to the function. Where the function maximizes the log also maximizes. So to find the maxima we can work with the log of the function instead of the function, if it simplifies the task at hand. For the minima we would use the negative log.

The likelihood estimator is a product of the probabilities of all the observations. Taking the log, reduces it to a sum of probabilities.
Product of probabilities becomes a very small number. Therefore, the cost function is the negative log likelihood **l**.

```{r Logistic_cost_function}
cost<-function(predictors,labels,parameters)
{
    predictions=gdpredict(predictors,parameters)
predictions[predictions<0]<-1-1*(1/nrow(predictors)) * sum(y*log(predictions) + (1-labels) * log ( 1-predictions))
}

```

## GD Predict

```{r gdpredict}
gdpredict<-function(predictors,parameters)
{
mx=as.matrix(predictors)
vec=as.vector(parameters)
dp=mx %*% vec
1/(1+exp(-1*dp))-0.0000001
}

```

## Gradient Descent

```{r convergence} 
GD<-function(predictors,labels,parameters,lrate,iter)
{
m=nrow(predictors)
J=cost(predictors,labels,parameters)

for( i in 1:iter)
{
        h=gdpredict(predictors,parameters)
        for (j in 1:ncol(predictors) )
        {
           parameters[j]= parameters[j] - (lrate/m) * ( sum (( h-labels)*predictors[,j] ))
        }
        c1=cost(predictors,labels,parameters)
      J=unlist(c(J,c1))
}
list(J=J,parameters=parameters)
}
```
##  Utilities to plot and compute metrics

```{r gd_plot_metrics}
plot_metrics <- function(L,lrate,iter)
{
plot(L$J,main=paste("lrate:",lrate,"iter:",iter,sep=''))
    }

get_metrics<-function(X,L)
    {
      #  plot(L$J)
        est_beta<-L$parameters
        print(L$parameters)
        gd_pred<-gdpredict(X,est_beta)

        gd_pred_class<-ifelse(gd_pred<0.5,0,1)

        table(gd_pred_class)
        table(y)

        (cfm<-table(y,gd_pred_class))

        sum(diag(cfm))/sum(cfm)
        caret_cfm<-caret::confusionMatrix(cfm)
        list(J=L$J,p=L$parameters,tbl=cfm,caret_cfm=caret_cfm)
    }

```
## Running Gradient Descent over Heart

### Loading data

```{r running_gd}
runGD<-function(X,y,lrate,iter,p)
    {
        beta<-rep(0.5,p)
        L<-GD(X,y,beta,lrate,iter)
        # plot_metrics(L,lrate,iter)
        get_metrics(X,L)
    }

lrate<-0.0001
iter<-25000

file<-'c://Users/rk215/Data/heart.csv'
ibmfile<-'C:/data/heart.csv'
heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)

X<-cbind(err=1,heart[,-c(ncol(heart))])
y<-heart$target
p<-ncol(X)

#oldpar<-par(mfrow=c(3,3))
L1<-runGD(X,y,lrate,iter,p)

lrate<-0.001
iter<-20000
L2<-runGD(X,y,lrate,iter,p)

lrate<-0.01
iter<-10000
L3<-runGD(X,y,lrate,iter,p)

L4<-runGD(X,y,lrate=0.0007,iter=20000,p)
L5<-runGD(X,y,lrate=0.00003,iter=20000,p)
L6<-runGD(X,y,lrate=0.00007,iter=30000,p)

L1$tbl
L2$tbl
L3$tbl
L4$tbl
L5$tbl
L6$tbl

#old_par=par(mfrow=c(3,2))
plot_metrics(L1,0.0001,25000)
plot_metrics(L6,0.00007,30000)
plot_metrics(L2,0.001,20000)
plot_metrics(L3,0.01,10000) 
plot_metrics(L5,0.00003,20000)
plot_metrics(L4,0.0007,20000)

```
## Comparison with GLM
How different are these estimates?
let us run glm and get the estimates...
```{r glm_comparison}
glm.model<-glm(target~.,data=heart,family='binomial')
coef(glm.model)
L1$p
L1$caret_cfm
glm.pred<-predict(glm.model,X,type='response')
glm.class<-ifelse(glm.pred<0.50,0,1)
glm.tbl<-table(y,glm.class)
caret::confusionMatrix(glm.tbl)
```

### Conclusion
GD is a numerical solution. GD however, requires a differentiable **Loss function** and when a Loss function is available, GD is an effective solution to an otherwise unsolvable problem.

### 2020-11-02 DATA622/M11-bagging-loocv-cv.RMD
Initial commit

---
title: "Bootstrap Aggregation-LooCV-naiveBayes"
author: "Raman Kannan"
date: "10/5/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bootstrap Aggregation: Bagging

We have already seen Bagging in M09 to compare with 10-Fold CV.

### Sampling with replacement

This is a time-honored, re-sampling technique. If we are given a small dataset to begin with, we can generate many replicas, if we allow   replacement. We draw N samples out of N samples with replacement. All features are included. Most importantly, the same observation may be present more than once in the replica. Not all the observations may be represented.

## Established Error reduction strategies

### Resampling

First we have to overcome small datasets. Statisticians have had to deal with small dataset problems and they have figured out a solution, called bootstrap. Just as person in a ditch uses the available shoe-lace to climb out of the hole, a data scientist with limited data, can use that limited data by resampling many times with replacement and try to overcome poor performance due to lack of data.

Again we need to justify why we are engaging in such trials by establishing base-line performance and demonstrating performance can be improved with such
experiments. Bootstrapping is a technique to generate slightly varying but different datasets from a given dataset. Then we run a model repeatedly over all the bootstrapped models, determine the metric and take the average. This is called Bagging.

### Base line performance of NB 
Ensemble methods start with a weak learner (low variance or low bias), combine them in different ways and estimate aggregate performance metrics either using majority (voting) or averaging. Voting is used in the case of classification and averaging (mean) in the case of regression. This technique can be used to improve the performance of any weak learner.

So instead of doing the tree based ensemble again, let us indulge in another classifier. Let us use NB as our base learner.

```{r}
library('e1071')
file<-'c://Users/rk215/Data/heart.csv'
heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)
head(heart)
catheart<-heart[,c(2,3,6,7,9,11,12,13,14)]

set.seed(43)
trdidx<-sample(1:nrow(catheart),0.7*nrow(catheart),replace=F)
trcatheart<-catheart[trdidx,]
tstcatheart<-catheart[-trdidx,]

nb.model<-naiveBayes(target~.,data=trcatheart)
#str(nb.model)
object.size(nb.model) #11096

nb.tstpred<-predict(nb.model,tstcatheart[,-c(9)],type='raw')
nb.tstclass<-unlist(apply(round(nb.tstpred),1,which.max))-1
nb.tbl<-table(tstcatheart[[9]], nb.tstclass)
nb.cfm<-caret::confusionMatrix(nb.tbl)
nb.cfm

start_tm <- proc.time()

df<-trcatheart
runModel<-function(df) {naiveBayes(target~.,data=df[sample(1:nrow(df),nrow(df),replace=T),])}
lapplyrunmodel<-function(x)runModel(df)

system.time(models<-lapply(1:100,lapplyrunmodel))

object.size(models)
end_tm<-proc.time() 
print(paste("time taken to run 100 bootstrapps",(end_tm-start_tm),sep=":"))

bagging_preds<-lapply(models,FUN=function(M,D=tstcatheart[,-c(9)])predict(M,D,type='raw'))

bagging_cfm<-lapply(bagging_preds,FUN=function(P,A=tstcatheart[[9]])
{pred_class<-unlist(apply(round(P),1,which.max))-1
  pred_tbl<-table(A,pred_class)
  pred_cfm<-caret::confusionMatrix(pred_tbl)
  pred_cfm
})

bagging.perf<-as.data.frame(do.call('rbind',lapply(bagging_cfm,FUN=function(cfm)c(cfm$overall,cfm$byClass))))

bagging.perf.mean<-apply(bagging.perf[bagging.perf$AccuracyPValue<0.01,-c(6:7)],2,mean)
bagging.perf.var<-apply(bagging.perf[bagging.perf$AccuracyPValue<0.01,-c(6:7)],2,sd)

bagging.perf.var
bagging.perf.mean

(bagging_tm<-proc.time()-start_tm)

```

### Jacknife: Leave One Out (LOO) Cross Validation

In M09, we did 10-Fold CV. Here we will do LOO-CV. In this form of CV, we iterate through every observation, train with all the other observations, and use that one observation to validate the models. As this is a classification exercise, class label is assigned using majority vote.

```{r running_jacknife_loocv}

N<-nrow(trcatheart)

cv_df<-do.call('rbind',lapply(1:N,FUN=function(idx,data=trcatheart) { # For each observation
  m<-naiveBayes(target~.,data=data[-idx,]) # train with ALL other observations
   p<-predict(m,data[idx,-c(9)],type='raw') # predict that one observation
   # NB returns the probabilities of the classes, as per Bayesian Classifier,we take the classs with the higher probability
   pc<-unlist(apply(round(p),1,which.max))-1 # -1 to make class to be 0 or 1, which.max returns 1 or 2
  #pred_tbl<-table(data[idx,c(9)],pc)
  #pred_cfm<-caret::confusionMatrix(pred_tbl)
  list(fold=idx,m=m,predicted=pc,actual=data[idx,c(9)]) # store the idx, model, predicted class and actual class
  }
))

```
**cv_df** now has the folds, models and the instance that was held out and the predicted label for that held out observation.
We can extract the metrics into a data.frame and average them as we did before.
```{r averaging_the_cv}

cv_df<-as.data.frame(cv_df)
head(cv_df)
tail(cv_df)
table(as.numeric(cv_df$actual)==as.numeric(cv_df$predicted))

loocv_tbl<-table(as.numeric(cv_df$actual),as.numeric(cv_df$predicted))
sum(diag(loocv_tbl))/sum(loocv_tbl)
(loocv_caret_cfm<-caret::confusionMatrix(loocv_tbl))
# now we have to apply the training models to testdata and average them 
# since this is classification we will take the majority vote
# double loop

tstcv.perf<-as.data.frame(do.call('cbind',lapply(cv_df$m,FUN=function(m,data=tstcatheart)
{

  v<-predict(m,data[,-c(9)],type='raw')
  lbllist<-unlist(apply(round(v),1,which.max))-1

}
  )))
np<-ncol(tstcv.perf)
predclass<-unlist(apply(tstcv.perf,1,FUN=function(v){ ifelse(sum(v[2:length(v)])/np<0.5,0,1)}))
loocvtbl<-table(tstcatheart[,c(9)],predclass)
(loocv_cfm<-caret::confusionMatrix(loocvtbl))

```
```{r comparing-metrics}
print(paste('Bagging:',bagging.perf.mean[1]))
print(paste('LOO-CV:',loocv_cfm$overall[1]))
print(paste('Base NB',nb.cfm$overall[[1]]))
```

##Bagging vs CV

 Even though the performance did not improve all that much, we now have one more tool in our toolbag In Data expeditions, you go where the data takes you. Double/triple validation is required but data leads you wherever it leads you. Let there not be any temptation to second guess data. Data is verily ***The Light of Truth***

### 2020-11-03 DATA622/M09-Bootstrap-CrossValidation-NaiveBayes.RMD
Updated M09

@@
 library('e1071')
-file<-'c://Users/rk215/Data/heart.csv'
+file<-'heart.csv'
 heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)

### 2020-11-10 DATA622/Test1.rmd
Initial commit

--- 2020-10-10 DATA622/HW1.rmd
@@
 ---
-title: "Homework 1"
+title: "Test 1"
 author: "Mike Silva"
-date: "10/10/2020"
+date: "`r Sys.Date()`"
 output: html_document
@@
 knitr::opts_chunk$set(echo = TRUE)
+# Use the dataset you used for HW-1 (Blue/Black)
+# 
+# (A) Run Bagging (ipred package)   
+# 
+#   -- sample with replacement
+# 
+#   -- estimate metrics for a model
+# 
+#   -- repeat as many times as specied and report the average
+# 
+# (B) Run LOOCV (jacknife) for the same dataset
+# 
+# --- iterate over all points
+# 
+#      -- keep one observation as test
+# 
+#     -- train using the rest of the observations
+# 
+#     -- determine test metrics
+# 
+#     -- aggregate the test metrics
+# 
+# end of loop
+# 
+# find the average of the test metric(s)
+# 
+# Compare (A), (B) above with the results you obtained in HW-1  and write 3 sentences explaining the
+# 
+# observed difference.
+# 
+# You can implement it in java/spark, python or R.
+# 
+# Submit RMD or PDF and the python/R scripts.
+# 
+# --
+# 
+# i would recommend that you implement it from scratch but you may use any package. It is not a requirement.
+# 
+# bagging is available in ipred package in R.
 ```
+This test is an extension of the work done in Homework 1.
+I will be extending it with bagging and leave one out cross-validation (LOOCV), and adding in the results into the evaluation of the algorithms.
 
@@
 ### Load Packages
-```{r, message=FALSE, warning=FALSE}
-library(ggplot2)
+
+```{r}
 library(dplyr)
@@
 library(ROCR)
-options(knitr.kable.NA = '')
 ```
@@
 ```{r}
-df <- read.csv("HW1.csv", stringsAsFactors = TRUE)
+df <- read.csv("https://github.com/mikeasilva/CUNY-SPS/raw/master/DATA622/HW1.csv", stringsAsFactors = TRUE)
 ```
-
-### EDA
-
-Let's take a look at the data:
-
-```{r}
-head(df) %>% kable() %>% kable_styling()
-```
-There are `r ncol(df)` columns and  `r nrow(df)` observations in the data set.  Let's look at a quick summary of the data:
-
-```{r}
-summary(df) %>% kable() %>% kable_styling()
-```
-
-`X` is a numeric variable ranging from `r min(df$X)` to `r max(df$X)`, while `Y` and `label` are categorical variables.
-There are 6 values for `Y` and the observations are evenly distributed among the variable.
-The `label` is not balanced.
-Let's look at the `Y` by `label`:
-
-```{r}
-table(df$Y, df$label) %>% kable() %>% kable_styling()
-```
-
-It looks like that could be used to divide the data into the two label classes as the preponderance of observations with a `Y` of c are blue, and those with a `Y` of e are black.
-Let's look at `X` by `label`
-
-```{r}
-df %>%
-  group_by(label) %>%
-  summarise(min = min(X),
-            mean = mean(X),
-            median = median(X),
-            max = max(X), .groups = "drop") %>%
-  kable() %>%
-  kable_styling()
-```
-
-The `X`s are slightly lower for the blue `label` than the black. Finally, let's look at all the data visually:
-
-```{r}
-ggplot(df, aes(x=X, y=Y, color=label)) + geom_point() + scale_color_manual(values=c("#000000", "#56B4E9")) +   theme(legend.position = "none")
-```
-
-One can see that the lower `X` values of the blues coupled with the preponderance of `Y` of c being blue leads to a large cluster of blue in the lower left part of the chart.
-Black dominates most of the other parts of the chart except for some blues on the extremes of `X`.
 
@@
 
+Now for the extensions for the test.
+This third function will be used to create a bootstrap sample.
+I will also make the assumption that we will use 100 boostrap samples in evaluating the performance of bagging.
+
+```{r}
+get_bootstrap_sample <- function(data, bootstrap_proportion = 0.6, sample_with_replacement = TRUE){
+  n_bootstrap_observations <- round(nrow(data) * bootstrap_proportion, 0)
+  data[sample(nrow(data), n_bootstrap_observations, replace = sample_with_replacement),]
+}
+
+n_bags <- 100
+```
+
 ## Gathering the Model's Capacity to Learn
@@
 lr_model <- glm(label ~ ., data = training_df, family = "binomial")
-training_lr_yhat <- get_lr_yhat(lr_model, training_df, "label")
-capacity_to_learn <- evaluate_algo(training_df_label, training_lr_yhat, "LR")
+summary(lr_model)
+```
+*The sample data are too small (`r nrow(training_df)`) to have any confidence in the Logistic Regression model, so I will be throwing this one out.  Bagging and LOOCV will not resolve this issue.*
+
+```{r}
+#training_lr_yhat <- get_lr_yhat(lr_model, training_df, "label") This does not make sense to use
+capacity_to_learn <- data.frame("Algo" = c("LR", "LR with Bagging", "LR with LOOCV"), AUC=c(NA), ACCURACY=c(NA), TPR=c(NA), FPR=c(NA), TNR=c(NA), FNR=c(NA))
 ```
@@
 capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(training_df_label, training_nb_yhat, "NB"))
+```
+
+#### With Bagging
+
+Now to try bagging the Naive Bayes model.  I will save all the models to a list so that they can be used to evaluate how well it would generalize latter in the document.
+
+```{r}
+nb_bag <- list()
+for (i in 1:n_bags){
+  bag_df <- get_bootstrap_sample(training_df)
+  bag_df_without_label <- bag_df %>% select(-label)
+  bag_df_label <- bag_df$label
+  bag_nb_model <- naiveBayes(label ~ ., data = bag_df)
+  training_bag_nb_yhat <- predict(bag_nb_model, bag_df_without_label)
+  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_nb_yhat, paste("NB Bag", i))
+  nb_bag[[i]] <- list(model = bag_nb_model, yhat=training_bag_nb_yhat, capacity_to_learn = bag_capacity_to_learn)
+  if(exists("nb_bag_capacity_to_learn")){
+    nb_bag_capacity_to_learn <- rbind(nb_bag_capacity_to_learn, bag_capacity_to_learn)
+  } else {
+    nb_bag_capacity_to_learn <- bag_capacity_to_learn
+  }
+}
+```
+
+Now to average the capacity to learn metrics:
+
+```{r}
+nb_bag_capacity_to_learn <- nb_bag_capacity_to_learn %>%
+  select(-Algo) %>%
+  colMeans(na.rm=TRUE) %>%
+  t() %>%
+  data.frame("Algo"=c("NB with Bagging"), .)
+capacity_to_learn <- rbind(capacity_to_learn, nb_bag_capacity_to_learn)
+```
+
+#### LOOCV
+
+Now Naive Bayes with LOOCV
+
+```{r}
+for (i in 1:nrow(training_df)){
+  loocv_test <- training_df[i,]
+  loocv_test_without_label <- loocv_test %>% select(-label)
+  loocv_training_df <- training_df[-c(i),]
+  loocv_nb_model <- naiveBayes(label ~ ., data = loocv_training_df)
+  training_loocv_nb_yhat <- predict(loocv_nb_model, loocv_test_without_label)
+  loocv_capacity_to_learn <- evaluate_algo(loocv_nb_model, training_loocv_nb_yhat, paste("NB LOOCV", i))
+}
 ```
@@
 ```
-
-**Table 2. Capacity to Generalize**
-```{r, echo=FALSE}
-capacity_to_generalize %>% kable() %>% kable_styling()
-```
-
-### Discussion
-
-In formulating our recommendations, we assume that the accuracy is very important, especially the TNR.
-Since the data is imbalanced (there are more blacks than blues), getting an accurate blue prediction is a good measure of how well the algorithm preforms.
-With this in mind, here's our analysis of each algorithm's performance.
-
-**Logistic Regression** showed a fairly strong ability to learn with an overall accuracy of 73%.
-It struggled to identify black cases but did well at identifying blue cases in the test data set.
-It's ability to generalize lead us not to recommend using this algorithm.
-The accuracy dropped to 33%, indicating that this algorithm over fit the data.
-
-**Naive Bayes** exhibited performance similar to logistic regression.
-It also had a 73% accuracy, and outperformed logistic regression's TPR and TNR.
-It also showed weakness in the ability to generalize with identical performance to logistic regression.
-It is for that reason we cannot recommend this algorithm.
-
-**KNN** was well suited to give accurate predictions.
-KNN3 had the highest capacity to learn out of all the algorithms analyzed.
-It had an accuracy of 80%.
-KNN3 also indicated strength in generalization, with a accuracy of 50%.
-That is a 17 percentage point improvement over logistic regression and naive Bayes.
-
-KNN5 indicated a similar ability to learn as logistic regression and naive Bayes.
-It did preform the best at detecting blues as indicated by the TNR of 83%.
-Where KNN5 shined was in the ability to generalize.
-It had the highest accuracy and the highest TNR of all the algorithms.
-
-```{r, echo=FALSE}
-ggplot() + 
-  geom_point(data = df[in_training_set,], aes(x = X, y = Y, color = label)) + 
-  geom_point(data = df[-in_training_set,], aes(x = X, y = Y, color = label), shape = 2) +
-  scale_color_manual(values=c("#000000", "#56B4E9")) +
-  theme(legend.position = "none")
-```
-
-KNN preformed so well because the data exhibited clustering.
-This is seen in the image above where the triangles are the points that needed to be predicted and the dots are the data used to make inferences.
-As previously noted, there is a large cluster of blue centered around where `X` is 20 and `Y` is c (which coincidentally is a point needing to be predicted).
-There is another patch of blues where `X` is greater than 60 and `Y` is not a.
-The black points dominated the chart where `X` ranged from 30 to 50.
-Given the points needed to be predicted (triangles) were surrounded by peers KNN preformed well.
-
-### Recommendation
-
-Based on our analysis of the performance of the algorithms on the data set, we would recommend using a KNN algorithm.
-KNN is well positioned to make an accurate prediction.
-Based strictly on the draw that we made, setting K = 5 seems to be the best choice.
-This is heavily dependent on the draw that was made in this process.
-A K = 3 is a strong alternative and should be explored using additional draws of data.

### 2020-11-11 DATA622/Test1.rmd
More work on Test 1

@@
   n_bootstrap_observations <- round(nrow(data) * bootstrap_proportion, 0)
-  data[sample(nrow(data), n_bootstrap_observations, replace = sample_with_replacement),]
+  return(data[sample(nrow(data), n_bootstrap_observations, replace = sample_with_replacement),])
 }
@@
 
-## Gathering the Model's Capacity to Learn
+I will also need a way to aggregate/average the metrics
 
-In order to evaluate this capacity, we will gather metrics on the algorithm's performance on the training set.
+```{r}
+average_metrics <- function(metrics, algo){
+  avg_metrics <- metrics %>%
+  select(-Algo) %>%
+  colMeans(na.rm = TRUE) %>%
+  t() %>%
+  data.frame("Algo" = c(algo), .)
+  return(avg_metrics)
+}
+```
+
+## Baseline Models
+
+First I will train the models following what was previously done in HW1 and collect the metrics on both the training and test data sets
 
@@
 ```
-*The sample data are too small (`r nrow(training_df)`) to have any confidence in the Logistic Regression model, so I will be throwing this one out.  Bagging and LOOCV will not resolve this issue.*
+#### Capacity to Learn
+
+**The sample data are too small (`r nrow(training_df)`) to have any confidence in the Logistic Regression model, so I will be throwing this one out.  Bagging and LOOCV will not resolve this issue.**
 
 ```{r}
-#training_lr_yhat <- get_lr_yhat(lr_model, training_df, "label") This does not make sense to use
 capacity_to_learn <- data.frame("Algo" = c("LR", "LR with Bagging", "LR with LOOCV"), AUC=c(NA), ACCURACY=c(NA), TPR=c(NA), FPR=c(NA), TNR=c(NA), FNR=c(NA))
 ```
+
+#### Capacity to Generalize
+
+**Again the sample data are too small to use to make meaningful inferences using Logistic Regression.  Bagging and LOOCV will not help and will be excluded from the metrics.**
+
+```{r}
+capacity_to_generalize <- data.frame("Algo" = c("LR", "LR with Bagging", "LR with LOOCV"), AUC=c(NA), ACCURACY=c(NA), TPR=c(NA), FPR=c(NA), TNR=c(NA), FNR=c(NA))
+```
+
 ### Naive Bayes
+
+I will repeat the same process for Naive Bayes.
+
+#### Capacity to Learn
 
@@
 
-#### With Bagging
-
-Now to try bagging the Naive Bayes model.  I will save all the models to a list so that they can be used to evaluate how well it would generalize latter in the document.
+#### Capacity to Generalize
 
 ```{r}
-nb_bag <- list()
-for (i in 1:n_bags){
-  bag_df <- get_bootstrap_sample(training_df)
-  bag_df_without_label <- bag_df %>% select(-label)
-  bag_df_label <- bag_df$label
-  bag_nb_model <- naiveBayes(label ~ ., data = bag_df)
-  training_bag_nb_yhat <- predict(bag_nb_model, bag_df_without_label)
-  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_nb_yhat, paste("NB Bag", i))
-  nb_bag[[i]] <- list(model = bag_nb_model, yhat=training_bag_nb_yhat, capacity_to_learn = bag_capacity_to_learn)
-  if(exists("nb_bag_capacity_to_learn")){
-    nb_bag_capacity_to_learn <- rbind(nb_bag_capacity_to_learn, bag_capacity_to_learn)
-  } else {
-    nb_bag_capacity_to_learn <- bag_capacity_to_learn
-  }
-}
-```
-
-Now to average the capacity to learn metrics:
-
-```{r}
-nb_bag_capacity_to_learn <- nb_bag_capacity_to_learn %>%
-  select(-Algo) %>%
-  colMeans(na.rm=TRUE) %>%
-  t() %>%
-  data.frame("Algo"=c("NB with Bagging"), .)
-capacity_to_learn <- rbind(capacity_to_learn, nb_bag_capacity_to_learn)
-```
-
-#### LOOCV
-
-Now Naive Bayes with LOOCV
-
-```{r}
-for (i in 1:nrow(training_df)){
-  loocv_test <- training_df[i,]
-  loocv_test_without_label <- loocv_test %>% select(-label)
-  loocv_training_df <- training_df[-c(i),]
-  loocv_nb_model <- naiveBayes(label ~ ., data = loocv_training_df)
-  training_loocv_nb_yhat <- predict(loocv_nb_model, loocv_test_without_label)
-  loocv_capacity_to_learn <- evaluate_algo(loocv_nb_model, training_loocv_nb_yhat, paste("NB LOOCV", i))
-}
+nb_yhat <- predict(nb_model, test_df_without_label)
+capacity_to_generalize <- rbind(capacity_to_generalize, evaluate_algo(test_df_label, nb_yhat, "NB"))
 ```
@@
 ### KNN
+
+Finally I will repeat the process for KNN where k=3 and k=5
+
+#### Capacity to Learn
 
@@
 
-## Gathering the Model's Capacity to Generalize
-
-### Logistic Regression
-
-```{r}
-lr_yhat <- get_lr_yhat(lr_model, test_df, "label")
-test_lr_cm <- confusionMatrix(table(test_df_label, lr_yhat))
-capacity_to_generalize <- evaluate_algo(test_df_label, lr_yhat, "LR")
-```
-
-### Naive Bayes
-
-```{r}
-nb_yhat <- predict(nb_model, test_df_without_label)
-capacity_to_generalize <- rbind(capacity_to_generalize, evaluate_algo(test_df_label, nb_yhat, "NB"))
-```
-
-### KNN
+#### Capacity to Generalize
 
@@
 
+## Models With Bagging
+
+Now to try bagging the Naive Bayes, and KNN models.
+I will use the models trained on the bootstrap samples to predict the test data and collect metrics on the capacity to learn and generalize, which will be aggregated.
+
+```{r}
+nb_bag <- list()
+for (i in 1:n_bags){
+  bag_df <- get_bootstrap_sample(training_df)
+  bag_df_without_label <- bag_df %>% select(-label)
+  bag_df_label <- bag_df$label
+
+  # Train the Naive Bayes Model
+  bag_nb_model <- naiveBayes(label ~ ., data = bag_df)
+  training_bag_nb_yhat <- predict(bag_nb_model, bag_df_without_label)
+  ## Evaluate the capacity to learn
+  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_nb_yhat, paste("NB Bag", i))
+  if(exists("nb_bag_capacity_to_learn")){
+    nb_bag_capacity_to_learn <- rbind(nb_bag_capacity_to_learn, bag_capacity_to_learn)
+  } else {
+    nb_bag_capacity_to_learn <- bag_capacity_to_learn
+  }
+  ## Evaluate the capacity to generalize
+  bag_nb_test_yhat <- predict(bag_nb_model, test_df_without_label)
+  bag_capacity_to_generalize <- evaluate_algo(test_df_label, bag_nb_test_yhat, paste("NB Bag", i))
+  if(exists("nb_bag_capacity_to_generalize")){
+    nb_bag_capacity_to_generalize <- rbind(nb_bag_capacity_to_generalize, bag_capacity_to_generalize)
+  } else {
+    nb_bag_capacity_to_generalize <- bag_capacity_to_generalize
+  }
+
+  # Train the KNN3 Model
+  ## Evaluate the capacity to learn
+  training_bag_knn3_yhat <- knn(bag_df_without_label, bag_df_without_label, bag_df_label, k = 3)
+  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_knn3_yhat, paste("KNN3 Bag", i))
+  if(exists("knn3_bag_capacity_to_learn")){
+    knn3_bag_capacity_to_learn <- rbind(knn3_bag_capacity_to_learn, bag_capacity_to_learn)
+  } else {
+    knn3_bag_capacity_to_learn <- bag_capacity_to_learn
+  }
+  ## Evaluate the capacity to generalize
+  bag_knn3_test_yhat <- knn(bag_df_without_label, test_df_without_label, bag_df_label, k = 3)
+  bag_capacity_to_generalize <- evaluate_algo(test_df_label, bag_knn3_test_yhat, paste("KNN3 with Bagging", i))
+  if(exists("knn3_bag_capacity_to_generalize")){
+    knn3_bag_capacity_to_generalize <- rbind(knn3_bag_capacity_to_generalize, bag_capacity_to_generalize)
+  } else {
+    knn3_bag_capacity_to_generalize <- bag_capacity_to_generalize
+  }
+
+  # Train the KNN5 Model
+  ## Evaluate the capacity to learn
+  training_bag_knn5_yhat <- knn(bag_df_without_label, bag_df_without_label, bag_df_label, k = 5)
+  bag_capacity_to_learn <- evaluate_algo(bag_df_label, training_bag_knn5_yhat, paste("KNN5 with Bagging", i))
+  if(exists("knn5_bag_capacity_to_learn")){
+    knn5_bag_capacity_to_learn <- rbind(knn5_bag_capacity_to_learn, bag_capacity_to_learn)
+  } else {
+    knn5_bag_capacity_to_learn <- bag_capacity_to_learn
+  }
+  ## Evaluate the capacity to generalize
+  bag_knn5_test_yhat <- knn(bag_df_without_label, test_df_without_label, bag_df_label, k = 5)
+  bag_capacity_to_generalize <- evaluate_algo(test_df_label, bag_knn5_test_yhat, paste("KNN5 with Bagging", i))
+  if(exists("knn5_bag_capacity_to_generalize")){
+    knn5_bag_capacity_to_generalize <- rbind(knn5_bag_capacity_to_generalize, bag_capacity_to_generalize)
+  } else {
+    knn5_bag_capacity_to_generalize <- bag_capacity_to_generalize
+  }
+
+}
+```
+
+Now to average the metrics:
+
+```{r}
+# Naive Bayes
+capacity_to_learn <- rbind(capacity_to_learn, average_metrics(nb_bag_capacity_to_learn, "NB with Bagging"))
+capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(nb_bag_capacity_to_generalize, "NB with Bagging"))
+
+# KNN3
+capacity_to_learn <- rbind(capacity_to_learn, average_metrics(knn3_bag_capacity_to_learn, "KNN3 with Bagging"))
+capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn3_bag_capacity_to_generalize, "KNN3 with Bagging"))
+
+# KNN5
+capacity_to_learn <- rbind(capacity_to_learn, average_metrics(knn5_bag_capacity_to_learn, "KNN5 with Bagging"))
+capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn5_bag_capacity_to_generalize, "KNN5 with Bagging"))
+```
+
+## Models With LOOCV
+
+Now to examine how the models preform using LOOCV
+
+```{r}
+training_loocv_nb_yhat <- c()
+training_loocv_knn3_yhat <- c()
+training_loocv_knn5_yhat <- c()
+loocv_test_label <- c()
+
+for (i in 1:nrow(training_df)){
+  # Leave One Out
+  loocv_test <- training_df[i,]
+  loocv_test_without_label <- loocv_test %>% select(-label)
+  loocv_test_label <- c(loocv_test_label, loocv_test$label)
+  loocv_training_df <- training_df[-c(i),]
+  loocv_training_df_without_label <- loocv_training_df %>% select(-label)
+  loocv_training_df_label <- loocv_training_df$label
+
+  # Train the Naive Bayes Model
+  loocv_nb_model <- naiveBayes(label ~ ., data = loocv_training_df)
+  ## Evaluate the capacity to learn
+  ### Note: This will be done once we have gone over the whole training data set
+  training_loocv_nb_yhat <- c(training_loocv_nb_yhat, predict(loocv_nb_model, loocv_test_without_label))
+  ## Evaluate the capacity to generalize
+  loocv_nb_test_yhat <- predict(loocv_nb_model, test_df_without_label)
+  loocv_capacity_to_generalize <- evaluate_algo(test_df_label, loocv_nb_test_yhat, paste("NB with LOOCV", i))
+  if(exists("nb_loocv_capacity_to_generalize")){
+    nb_loocv_capacity_to_generalize <- rbind(nb_loocv_capacity_to_generalize, loocv_capacity_to_generalize)
+  } else {
+    nb_loocv_capacity_to_generalize <- loocv_capacity_to_generalize
+  }
+
+  # Train the KNN3 Model
+  ## Evaluate the capacity to learn
+  ### Note: This will be done once we have gone over the whole training data set
+  training_loocv_knn3_yhat <- c(training_loocv_knn3_yhat, knn(loocv_training_df_without_label, loocv_test_without_label, loocv_training_df_label, k = 3))
+  ## Evaluate the capacity to generalize
+  loocv_knn3_test_yhat <- knn(loocv_training_df_without_label, test_df_without_label, loocv_training_df_label, k = 3)
+  loocv_capacity_to_generalize <- evaluate_algo(test_df_label, loocv_knn3_test_yhat, paste("KNN3 with LOOCV", i))
+  if(exists("knn3_loocv_capacity_to_generalize")){
+    knn3_loocv_capacity_to_generalize <- rbind(knn3_loocv_capacity_to_generalize, loocv_capacity_to_generalize)
+  } else {
+    knn3_loocv_capacity_to_generalize <- loocv_capacity_to_generalize
+  }
+
+  # Train the KNN5 Model
+  ## Evaluate the capacity to learn
+  ### Note: This will be done once we have gone over the whole training data set
+  training_loocv_knn5_yhat <- c(training_loocv_knn5_yhat, knn(loocv_training_df_without_label, loocv_test_without_label, loocv_training_df_label, k = 5))
+  ## Evaluate the capacity to generalize
+  loocv_knn5_test_yhat <- knn(loocv_training_df_without_label, test_df_without_label, loocv_training_df_label, k = 5)
+  loocv_capacity_to_generalize <- evaluate_algo(test_df_label, loocv_knn5_test_yhat, paste("KNN5 with LOOCV", i))
+  if(exists("knn5_loocv_capacity_to_generalize")){
+    knn5_loocv_capacity_to_generalize <- rbind(knn5_loocv_capacity_to_generalize, loocv_capacity_to_generalize)
+  } else {
+    knn5_loocv_capacity_to_generalize <- loocv_capacity_to_generalize
+  }
+
+}
+```
+
+Now to average the metrics:
+
+```{r}
+# Naive Bayes
+capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(loocv_test_label, training_loocv_nb_yhat, "NB with LOOCV"))
+capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(nb_loocv_capacity_to_generalize, "NB with LOOCV"))
+
+# KNN3
+capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(loocv_test_label, training_loocv_knn3_yhat, "KNN3 with LOOCV"))
+capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn3_loocv_capacity_to_generalize, "KNN3 with LOOCV"))
+
+# KNN5
+capacity_to_learn <- rbind(capacity_to_learn, evaluate_algo(loocv_test_label, training_loocv_knn5_yhat, "KNN5 with LOOCV"))
+capacity_to_generalize <- rbind(capacity_to_generalize, average_metrics(knn5_loocv_capacity_to_generalize, "KNN5 with LOOCV"))
+```
+
 ## Summary and Analysis
 
-We have evaluated `r nrow(capacity_to_learn)` algorithm's capacity to learn and generalize.  The following tables summarize their performace:
+We have evaluated `r nrow(capacity_to_learn)` algorithm's capacity to learn and generalize.  The following tables summarize their performance:
 
@@
 ```{r, echo=FALSE}
-capacity_to_learn %>%  kable() %>%  kable_styling()
+capacity_to_learn %>% arrange(Algo) %>% kable() %>%  kable_styling()
 ```
+
+**Table 2. Capacity to Generalize**
+```{r, echo=FALSE}
+capacity_to_generalize %>% arrange(Algo) %>% kable() %>% kable_styling()
+```
+
+### Discussion
+
+```{r}
+ggplot() + geom_density(aes(x=ACCURACY, color="Bagging"), data=nb_bag_capacity_to_generalize) + geom_density(aes(x=ACCURACY, color="LOOCV"), data=nb_loocv_capacity_to_generalize) + geom_vline(xintercept = 1/3) + ggtitle("Naive Bayes Capacity to Generalize") + scale_color_brewer(palette = "Set1")
+```
+
+```{r}
+ggplot() + geom_density(aes(x=ACCURACY), data=knn3_bag_capacity_to_generalize) + geom_vline(xintercept = 1/2)
+```
+
+```{r}
+ggplot() + geom_density(aes(x=ACCURACY), data=knn5_bag_capacity_to_generalize) + geom_vline(xintercept = 2/3)
+```

### 2020-11-11 DATA622/Test1.rmd
More work on Test 1

@@
 
-```{r}
+```{r, message=FALSE, warning=FALSE}
 library(dplyr)
@@
 library(ROCR)
+library(stringr)
 ```
@@
 
-```{r}
+```{r, comment=NA}
 lr_model <- glm(label ~ ., data = training_df, family = "binomial")
@@
 ```{r}
-nb_bag <- list()
+the_bags <- list()
 for (i in 1:n_bags){
@@
   bag_df_label <- bag_df$label
+  the_bags[[i]] <- bag_df
 
@@
 loocv_test_label <- c()
+loocv_dfs <- list()
 
@@
   loocv_training_df_label <- loocv_training_df$label
+  loocv_dfs[[i]] <- loocv_training_df
 
@@
 
-We have evaluated `r nrow(capacity_to_learn)` algorithm's capacity to learn and generalize.  The following tables summarize their performance:
+We have evaluated `r nrow(na.omit(capacity_to_learn))` algorithm's capacity to learn and generalize.  The following tables summarize their performance:
 
@@
 
+There are `r nrow(training_df)` rows in the training set.
+When doing LOOCV the model is trained on `r nrow(training_df) - 1` rows of data.
+This subsample is going to be very similar to the full training set.
+The bagging on the other hand will be significantly different.
+There are `r nrow(bag_df)` rows of data and since it is sampled with replacement a same observation could be selected multiple times.
+Consequently the underlying data is different.
+We can see this by looking at the share of records that are labeled "BLACK."
+
 ```{r}
-ggplot() + geom_density(aes(x=ACCURACY, color="Bagging"), data=nb_bag_capacity_to_generalize) + geom_density(aes(x=ACCURACY, color="LOOCV"), data=nb_loocv_capacity_to_generalize) + geom_vline(xintercept = 1/3) + ggtitle("Naive Bayes Capacity to Generalize") + scale_color_brewer(palette = "Set1")
+get_share_black <- function(df){
+  df %>% filter(label == "BLACK") %>% nrow() / nrow(df)
+}
+```
+
+The share of the training data set that is labeled black is `r get_share_black(training_df)`. Let's compare this to the shares in the bags.
+
+```{r, comment=NA}
+share_black <- c()
+for(i in 1:n_bags){
+  share_black <- c(share_black, get_share_black(the_bags[[i]]))
+}
+
+summary(share_black)
+```
+
+You can see that the share of observations that are black range from `r round(min(share_black) * 100, 0)`% to `r round(max(share_black) * 100, 0)`%.
+This is very different than what we observe with the LOOCV data set:
+
+```{r, comment=NA}
+share_black <- c()
+for(i in 1:nrow(training_df)){
+  share_black <- c(share_black, get_share_black(loocv_dfs[[i]]))
+}
+
+summary(share_black)
+```
+
+This is a lot closer to the full training set.
+Consequently a model trained on the bootstrap sample is likely to have different results if it is a weak learner, than the full training set.
+Some of the models will do a better job at generalizing than others due to the composition of the data.
+This can be seen by looking at a metric like the accuracy of the model.
+The following images represent the mean accuracy on the test dataset as a dashed vertical line.
+For the bagging and LOOCV runs the distribution of the accuracy is also shown:
+
+```{r}
+vline_df <- capacity_to_generalize %>% filter(str_detect(Algo, "NB"))
+ggplot() + geom_vline(data=vline_df, aes(xintercept=ACCURACY, color=Algo), linetype="dashed") + geom_density(aes(x=ACCURACY, color="NB with Bagging"), data=nb_bag_capacity_to_generalize) + geom_density(aes(x=ACCURACY, color="NB with LOOCV"), data=nb_loocv_capacity_to_generalize) + ggtitle("Naive Bayes Capacity to Generalize") + scale_color_brewer(palette = "Set1")
 ```
@@
 ```{r}
-ggplot() + geom_density(aes(x=ACCURACY), data=knn3_bag_capacity_to_generalize) + geom_vline(xintercept = 1/2)
+vline_df <- capacity_to_generalize %>% filter(str_detect(Algo, "KNN3"))
+ggplot() + geom_vline(data=vline_df, aes(xintercept=ACCURACY, color=Algo), linetype="dashed") + geom_density(aes(x=ACCURACY, color="KNN3 with Bagging"), data=knn3_bag_capacity_to_generalize) + geom_density(aes(x=ACCURACY, color="KNN3 with LOOCV"), data=knn3_loocv_capacity_to_generalize) + ggtitle("KNN3 Capacity to Generalize") + scale_color_brewer(palette = "Set1")
 ```
@@
 ```{r}
-ggplot() + geom_density(aes(x=ACCURACY), data=knn5_bag_capacity_to_generalize) + geom_vline(xintercept = 2/3)
+vline_df <- capacity_to_generalize %>% filter(str_detect(Algo, "KNN5"))
+ggplot() + geom_vline(data=vline_df, aes(xintercept=ACCURACY, color=Algo), linetype="dashed") + geom_density(aes(x=ACCURACY, color="KNN5 with Bagging"), data=knn5_bag_capacity_to_generalize) + geom_density(aes(x=ACCURACY, color="KNN5 with LOOCV"), data=knn5_loocv_capacity_to_generalize) + ggtitle("KNN5 Capacity to Generalize") + scale_color_brewer(palette = "Set1")
 ```
+
+Now I hasten to note that all of these figures hare very specific to the initial split in the data which is controlled by the random number seed.
+So the statements I will make at this point are specific to this run.
+If I would have used another random number seed the results could be vastly different.
+
+When I completed HW-1 I noted the the KNN family of models preformed the best.
+KNN3 with bagging out preformed KNN3 on the whole training data.
+As previously discussed it is due to the variation in the bootstrap samples.
+When it comes to the ability to generalize KNN5 with LOOCV out preformed KNN5 and KNN3 with bagging out preformed KNN3.
+The accuracy increases above the "baseline" algorithm.
+NB improved the ability to generalize through both bagging and LOOCV.
+This suggests that the performance of these weak learners is highly dependent on the data, but when averaged together the group outperforms the single model.
+
+Again these results are highly dependent on the underling data, and additional runs should be done in order to ensure the conclusions are real and not soley based on idiosyncrasies in the data.
+The idea is inline with what has been taught in this class.

### 2020-11-12 DATA622/Test1.rmd
Minor changes
More work on Test 1

@@
 
+**Available Online at: [https://github.com/mikeasilva/CUNY-SPS/blob/master/DATA622/Test1.rmd](https://github.com/mikeasilva/CUNY-SPS/blob/master/DATA622/Test1.rmd)**
+
 ```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
-# Use the dataset you used for HW-1 (Blue/Black)
-# 
-# (A) Run Bagging (ipred package)   
-# 
-#   -- sample with replacement
-# 
-#   -- estimate metrics for a model
-# 
-#   -- repeat as many times as specied and report the average
-# 
-# (B) Run LOOCV (jacknife) for the same dataset
-# 
-# --- iterate over all points
-# 
-#      -- keep one observation as test
-# 
-#     -- train using the rest of the observations
-# 
-#     -- determine test metrics
-# 
-#     -- aggregate the test metrics
-# 
-# end of loop
-# 
-# find the average of the test metric(s)
-# 
-# Compare (A), (B) above with the results you obtained in HW-1  and write 3 sentences explaining the
-# 
-# observed difference.
-# 
-# You can implement it in java/spark, python or R.
-# 
-# Submit RMD or PDF and the python/R scripts.
-# 
-# --
-# 
-# i would recommend that you implement it from scratch but you may use any package. It is not a requirement.
-# 
-# bagging is available in ipred package in R.
+
 ```
+## Preliminaries
 This test is an extension of the work done in Homework 1.
 I will be extending it with bagging and leave one out cross-validation (LOOCV), and adding in the results into the evaluation of the algorithms.
-
-## Preliminaries
 
@@
 This suggests that the performance of these weak learners is highly dependent on the data, but when averaged together the group outperforms the single model.
-
-Again these results are highly dependent on the underling data, and additional runs should be done in order to ensure the conclusions are real and not soley based on idiosyncrasies in the data.
-The idea is inline with what has been taught in this class.

### 2020-11-12 DATA698/pytorch_to_tensorflow_lite.py
More work on final project

@@ -1,79 +1,0 @@
-# Originally from https://github.com/kalaspuffar/onnx-convert-example
-# Running the script python3 pytorch_to_tensorflow_lite.py models/mnist.pth models
-import sys
-import os
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import onnx
-from collections import OrderedDict
-import tensorflow as tf
-from torch.autograd import Variable
-from onnx_tf.backend import prepare
-
-class MLP(nn.Module):
-    def __init__(self, input_dims, n_hiddens, n_class):
-        super(MLP, self).__init__()
-        assert isinstance(input_dims, int), 'Please provide int for input_dims'
-        self.input_dims = input_dims
-        current_dims = input_dims
-        layers = OrderedDict()
-
-        if isinstance(n_hiddens, int):
-            n_hiddens = [n_hiddens]
-        else:
-            n_hiddens = list(n_hiddens)
-        for i, n_hidden in enumerate(n_hiddens):
-            layers['fc{}'.format(i+1)] = nn.Linear(current_dims, n_hidden)
-            layers['relu{}'.format(i+1)] = nn.ReLU()
-            layers['drop{}'.format(i+1)] = nn.Dropout(0.2)
-            current_dims = n_hidden
-        layers['out'] = nn.Linear(current_dims, n_class)
-
-        self.model= nn.Sequential(layers)
-        print(self.model)
-
-    def forward(self, input):
-        input = input.view(input.size(0), -1)
-        assert input.size(1) == self.input_dims
-        return self.model.forward(input)
-
-print("%s" % sys.argv[1])
-print("%s" % sys.argv[2])
-
-
-# Load the trained model from file
-trained_dict = torch.load(sys.argv[1], map_location={'cuda:0': 'cpu'})
-
-trained_model = MLP(784, [256, 256], 10) #input_dims, n_hiddens, n_class
-trained_model.load_state_dict(trained_dict)
-
-if not os.path.exists("%s" % sys.argv[2]):
-    os.makedirs("%s" % sys.argv[2])
-
-# Export the trained model to ONNX
-dummy_input = Variable(torch.randn(1, 1, 28, 28)) # one black and white 28 x 28 picture will be the input to the model
-torch.onnx.export(trained_model, dummy_input, "%s/mnist.onnx" % sys.argv[2])
-
[…]

### 2020-11-13 DATA698/08 - Report Visualizations.rmd
More work on project

---
title: "08 - Report Viz"
author: "Mike Silva"
date: "Fall 2020"
output: html_document
---

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

yolov3_results <- read.csv("report/yolov3_results.csv") %>%
  rowwise() %>%
  mutate(Epoch = as.numeric(strsplit(Epoch, "/")[[1]][1])+1)

yolov5_results <- read.csv("report/yolov5_results.csv") %>%
  rowwise() %>%
  mutate(Epoch = as.numeric(strsplit(Epoch, "/")[[1]][1])+1)

```

```{r}
yolo_results <- pivot_longer(yolov3_results, GIoU:Val.classification, names_to = "Measure")
yolo_results <- pivot_longer(yolov5_results, GIoU:Val.classification, names_to = "Measure") %>%
  rbind(yolo_results)
```

```{r}
viz_data <- yolov3_results %>%
  mutate(Model = as.factor(paste0(model,": ", n_images, " images")))
ggplot(viz_data, aes(x=Epoch, y=mAP.0.5, color=Model, group=Model)) + geom_line() + scale_color_brewer(palette = "Set1")
```

```{r}
viz_data <- yolov5_results %>%
  mutate(Model = as.factor(paste0(model,": ", n_images, " images")))
ggplot(viz_data, aes(x=Epoch, y=mAP.0.5, color=Model, group=Model)) + geom_line() + scale_color_brewer(palette = "Set1")
```

```{r}
for(m in unique(yolo_results$Measure)){
  viz_data <- yolo_results %>%
    filter(Measure == m) %>%
    mutate(Model = as.factor(paste0(model,": ", n_images, " images")))
    p <- ggplot(viz_data, aes(x=Epoch, y=value, color=Model, group=Model)) + 
      geom_line() +
      theme(axis.title.y = element_blank()) +
      ggtitle(m)
    print(p)
}
```

### 2020-11-16 DATA622/M12-RF.RMD
Initial commit

---
title: "M12-RandomForest"
author: "Raman Kannan"
date: "8/24/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Random Forest#

Recall a classification exercise is characterized by the algorithm we used, the dataset, characterized by (**p** the number of features) and (**N** number of observations).

We can vary these seeking to improve performance.
When we did **BAGGING** we varied the set of observations but we kept all the features. Because it is done with replacement, some observations had more weights and others had no representation. Observations without representation never made it to the training set. Observations with more weight were included in the training set more than once.

**Cross Validation**, on the other hand includes all features but by splitting the data into disjoint subsets and by iterating through all the subsets, ensures every observation is included in testing and training. Process of CV ensures when an observation is included in the training phase it is not also included in the training phase.

Now we want to vary the features and aggregate the results.
Aggregation for Classification is not average but some form of voting.

A **Random Forest** is a forest of trees. Each tree is distinct in that it uses a random bootstrap of the training set and varies the features randomly at every split. Recall that a tree uses information theoretic principles to determine which feature to split on. A tree splits on the feature that results in the highest information (GINI) gain. Key distinction it considers the entire feature set to make that determination.

In Random Forest, not all the features are considered. Only a subset is considered, and at every split.

First to understand the impact of varying features, let us run two individual trees considering different features.

Then, we will run a standard implementation of RandomForest.

```{r utility_functions}
file<-'c://Users/rk215/Data/heart.csv'
heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)

unlist(lapply(names(heart),FUN=function(x,data=heart){c(cname=x,uvalfreq=length(unique(data[[x]])))}))

nvpairsdf3<-do.call('rbind',lapply(names(heart),FUN=function(x,data=heart){c(cname=x,uvalfreq=length(unique(data[[x]])))}))

categoricalFeatures<-function(dset,ncol=7)
{
  df4<-as.data.frame(do.call('rbind',lapply(names(dset),FUN=function(x,data=dset){c(cname=x,uvalfreq=length(unique(data[[x]])))})))
  df4$uvalfreq=as.numeric(df4$uvalfreq)
  dset[,df4$uvalfreq<=ncol]
}

X<-categoricalFeatures(heart)

table(heart$target)
dim(X)
class_col<-which(names(X)=='target')

fvcnt<-ncol(X[,-c(class_col)])

#set.seed if you want to repeatability
#RF implementations consider sqrt(p) features
# to avoid too many common features
# here we are seeking to establish that it matters
# we are concerned about features being present in both
exp_fset1<-sample(1:fvcnt,fvcnt-1,replace=F)
exp_fset2<-sample(1:fvcnt,fvcnt-1,replace=F)
table(sort(exp_fset1)==sort(exp_fset2))

exp_fset1
exp_fset2

```

```{r, echo=FALSE}
library(rpart)
library(rpart.plot)
library(ROCR)

isModelReliable<-function(cfmx)
{	cfmx$overall[['AccuracyPValue']]<0.05;
 }

isUnderfitting<-function(cfmx)
{	cfmx$overall[['Accuracy']]<0.60;
 }

isOverfitting<-function(tr,tst)
{	tst$overall[['Accuracy']]/tr$overall[['Accuracy']]<0.75;
 }

getSensitivity<-function(cfmx) cfmx$byClass[['Sensitivity']]

getAccuracy<-function(cfmx) cfmx$overall[['Accuracy']]

getSpecificity<-function(cfmx) cfmx$byClass[['Specificity']]

getMetrics<-function(actual_class,predicted_response)
{
X=list()
if ( require(ROCR) ) {
auc_1=prediction(predicted_response,actual_class)
prf=performance(auc_1, measure="tpr",x.measure="fpr")
slot_fp=slot(auc_1,"fp")
slot_tp=slot(auc_1,"tp")

fpr=unlist(slot_fp)/unlist(slot(auc_1,"n.neg"))
tpr=unlist(slot_tp)/unlist(slot(auc_1,"n.pos"))

auc<-performance(auc_1,"auc")
AUC<-auc@y.values[[1]]
X=list(prf=prf,fpr=fpr,tpr=tpr,auc=AUC)
}
X
}
X$target=as.factor(X$target)

#par(mfrow=c(3,1))
rpart.model<-rpart(target~.,data=X,minsplit=3)
rpart.plot(rpart.model)

#predict

rcpALL<-predict(rpart.model,
                                  X[,-c(which(names(X)=="target"))],type="class")

(rpart_mtabAll<-table(rcpALL,X$target))

LALL<-getMetrics(X$target,as.numeric(rcpALL))
LALL$auc
LALL$fpr
LALL$tpr

```

```{r exp1}
X1<-X[,c(exp_fset1,9)]

rpart.model1<-rpart(target~.,data=X1,minsplit=3)
rpart.plot(rpart.model1)

rcp1<-predict(rpart.model1,
                                  X1[,-c(which(names(X1)=="target"))],type="class")

(rpart_mtab1<-table(rcp1,X$target))

L1<-getMetrics(X1$target,as.numeric(rcp1))
L1$auc
L1$fpr
L1$tpr

```

```{r exp2}

X2<-X[,c(exp_fset2,9)]

rpart.model2<-rpart(target~.,data=X2,minsplit=3)
rpart.plot(rpart.model2)

rcp2<-predict(rpart.model2,
                                  X2[,-c(which(names(X2)=="target"))],type="class")

(rpart_mtab2<-table(rcp2,X2$target))

L2<-getMetrics(X2$target,as.numeric(rcp2))
L2$auc
L2$fpr
L2$tpr

```
```{r exp3}
exp_fset3<-sample(1:fvcnt,fvcnt-1,replace=F)
X3<-X[,c(exp_fset3,9)]

rpart.model3<-rpart(target~.,data=X3,minsplit=3)
rpart.plot(rpart.model2)

rcp3<-predict(rpart.model3,
                                  X3[,-c(which(names(X3)=="target"))],type="class")

(rpart_mtab3<-table(rcp3,X3$target))

L3<-getMetrics(X3$target,as.numeric(rcp3))
L3$auc
L3$fpr
L3$tpr
```

We ran three different combinatons of feature vectors and
we got three different performance matrices.

```{r reivew_pred}
dfpred<-data.frame(actual=X$target,rcpALL,rcp1,rcp2,rcp3)

head(dfpred)
table(X$target,rcpALL)

table(X$target,rcp1)
table(X$target,rcp2)
table(X$target,rcp3)
```
So as swap one feature for another, performance of the tree varies ever so lightly.
Now what if we vary a 1000 times and not at the root level but
at every level where we split...
That is RandomForest...
```{r rforest}

if(!require(randomForest))require(randomForest)

rf_model<-randomForest(target~.,data=X)
rf_pred<-predict(rf_model,X[,-c(which(names(X)=="target"))])
rf_mtab<-table(X$target,rf_pred)
rf_cmx<-caret::confusionMatrix(rf_mtab)
rf_mtab
(rf_accuracy<-sum(diag(rf_mtab))/sum(rf_mtab))
rf_cmx$overall
rf_cmx$byClass

```

```{r rFOptions}

rf1k_model<-randomForest(target~.,data=X,ntree=1000)
rf1k_pred<-predict(rf1k_model,X[,-c(which(names(X)=="target"))])
rf1k_mtab<-table(X$target,rf1k_pred)
rf1k_cmx<-caret::confusionMatrix(rf1k_mtab)
rf1k_mtab
(rf1k_accuracy<-sum(diag(rf1k_mtab))/sum(rf1k_mtab))
rf1k_cmx$overall
rf1k_cmx$byClass

rf100_model<-randomForest(target~.,data=X,ntree=100)
rf100_pred<-predict(rf100_model,X[,-c(which(names(X)=="target"))])
rf100_mtab<-table(X$target,rf100_pred)
rf100_cmx<-caret::confusionMatrix(rf100_mtab)
rf100_mtab
(rf100_accuracy<-sum(diag(rf100_mtab))/sum(rf100_mtab))
rf100_cmx$overall
rf100_cmx$byClass

rf50_model<-randomForest(target~.,data=X,ntree=50)
rf50_pred<-predict(rf50_model,X[,-c(which(names(X)=="target"))])
rf50_mtab<-table(X$target,rf50_pred)
rf50_cmx<-caret::confusionMatrix(rf50_mtab)
rf50_mtab
(rf1k_accuracy<-sum(diag(rf50_mtab))/sum(rf50_mtab))
rf1k_cmx$overall
rf1k_cmx$byClass

rfgrow<-grow(rf50_model,50)

rfgrow_pred<-predict(rfgrow,X[,-c(which(names(X)=="target"))])
rfgrow_mtab<-table(X$target,rfgrow_pred)
rfgrow_cmx<-caret::confusionMatrix(rfgrow_mtab)
rfgrow_mtab
(rfgrow_accuracy<-sum(diag(rfgrow_mtab))/sum(rfgrow_mtab))
rfgrow_cmx$overall
rfgrow_cmx$byClass

```
The randomForest achieves 0.92% accuracy better than any of the other individual trees.

Running with ntree=100 or growing 50 at a time yields almost
identical accuracy. Given the random features at every split
this is somewhat circumspect.

Please note that I am repeating this experiment with the entire set, and therefore these are training error estimates.

I encourage you to evaluate the generalization error using test-set.

Notice also that randomForest can be parallelized. I could have run the three experiments in parallel. If the goal is to be big data proficient, then focus on distribution, parallelization.

We have few more options to improve performance. And that will be the topic of next two weeks.

### 2020-11-20 DATA698/09 - Process Field Test Results.py
Initial commit

@@ -1,0 +1,29 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Nov 20 09:35:33 2020
+
+@author: Michael
+"""
+
+import json
+
+yolov3_cells = {250: 19, 500: 22, 1000: 25, 2000: 28}
+yolov5_cells = {250: 14, 500: 17, 1000: 20, 2000: 23}
+
+
+# Process the YOLOv3 Results
+with open('./results/YOLOv3 Field Tests Results.ipynb') as json_file:
+    yolov3 = json.load(json_file)["cells"]
+    
+for n_images, key in yolov3_cells.items():
+    cell = yolov3[key]["outputs"]
+    print(n_images)
+    
+    
+# Process the YOLOv5 Results    
+with open('./results/YOLOv5 Field Tests Results.ipynb') as json_file:
+    yolov5 = json.load(json_file)["cells"]
+    
+for n_images, key in yolov5_cells.items():
+    cell = yolov5[key]["outputs"]
+    print(n_images)

### 2020-11-20 DATA698/09 - Process Field Test Results.py
Updated field test results processing script

@@ -4,20 +4,52 @@ Created on Fri Nov 20 09:35:33 2020
 
 @author: Michael
 """
-
+import pandas as pd
 import json
 
 yolov3_cells = {250: 19, 500: 22, 1000: 25, 2000: 28}
 yolov5_cells = {250: 14, 500: 17, 1000: 20, 2000: 23}
+classes = ["Virginia Creeper", "Poison Ivy", "Brambles", "Box Elder"]
+
+def extract_data_from_cell(cell, n_images, model):
+    data = []
+    for c in cell:
+        if "text" in c:
+            for line in c["text"]:
+                if "image 1/1" in line:
+                    img = line.split(":")[0].split("/")[-1]
+                    found = {"Virginia Creeper":0, "Poison Ivy": 0, "Brambles":0, "Box Elder":0, "Null":0}
+                    findings = line.split(":")[1]
+                    for cl in classes:
+                        if cl in findings:
+                            found[cl] = 1
+                    if len(findings.split(",")) == 1:
+                        found["Null"] = 1
+                    found["Image"] = img
+                    found["n_images"] = n_images
+                    found["Model"] = model
+                    data.append(found)
+    return data
+
+
+def extract_data_from_yolov5_cell(cell):
+    data = {}
+    return data
 
 
 # Process the YOLOv3 Results
 with open('./results/YOLOv3 Field Tests Results.ipynb') as json_file:
     yolov3 = json.load(json_file)["cells"]
     
+
+
 for n_images, key in yolov3_cells.items():
     cell = yolov3[key]["outputs"]
-    print(n_images)
+    data = extract_data_from_cell(cell, n_images, "YOLOv3")
+    try:
+        yolo_data = yolo_data + data.copy()
+    except:
+        yolo_data = data.copy()
     
     
 # Process the YOLOv5 Results    
@@ -26,4 +58,9 @@ with open('./results/YOLOv5 Field Tests Results.ipynb') as json_file:
     
 for n_images, key in yolov5_cells.items():
     cell = yolov5[key]["outputs"]
-    print(n_images)
+    data = extract_data_from_cell(cell, n_images, "YOLOv5")
+    yolo_data = yolo_data + data.copy()
+    
+    
+df = pd.DataFrame(yolo_data)
+df.to_csv("./report/field_test_results.csv", index = False)

### 2020-11-24 DATA622/M12-gbm.RMD
Initial commit

---
title: "GradientBoostedTrees"
author: "Raman Kannan"
date: "10/5/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)

```
# Boosting

In Boosting is an iterative method, similar to penalization, and in each iteration, we increase the weights of observations, for which the classifier fails to label correctly.

## Available boosting methods

* AdaBoost
* Extreme gradient boosting (XGBoost)
* gradient based boosting
* CatBoost
* LightGBM

# Boosting Trials

```{r running_gbm}
require(gbm)
require(ROCR)
require(pROC)

file<-'heart.csv'
heart<-read.csv(file,head=T,sep=',',stringsAsFactors=F)

tridx<-sample(1:nrow(heart),0.7*nrow(heart),replace=F)
tstdata<-heart[-tridx,]
trdata<-heart[tridx,]
names(trdata)[[14]]
gbm_heart_3<-gbm(target~.,data=trdata,
                 distribution="bernoulli",
                 n.trees=500,
                 shrinkage=0.01,
                 interaction.depth=3,
                 n.minobsinnode=10,
                 verbose=T, 
                 keep.data=F)

#Gradient tree boosting implementations often also #use regularization by limiting the minimum number #of observations in trees' terminal nodes. It is #used in the tree building process by ignoring any #splits that lead to nodes containing fewer than #this number of training set instances.

#Imposing this limit helps to reduce variance in #predictions at leaves.

gbm_predict<-predict(gbm_heart_3,
                     tstdata[,-c(14)],
                    type="response",
                    gbm_heart_3$n.trees)

gbm_predicted<-round(gbm_predict)

gbm_prediction<-prediction(gbm_predicted,
                           tstdata$target)

(gbm_tbl<-table(tstdata$target,gbm_predicted))
(gbm_cfm<-caret::confusionMatrix(gbm_tbl))
gbm_perf<-performance(gbm_prediction,
                      measure="tpr",
                      x.measure="fpr")
(gbm_auc<-performance(gbm_prediction,
                     measure="auc"))@y.values[[1]]

plot(gbm_perf,main="ROC GBM n.tree=500")
text(0.5,0.5,paste("AUC=",format(gbm_auc@y.values[[1]],digits=5, scientific=FALSE)))
```
## Business of plotting ROC
The above ROC Curve plot is not sufficient as it has
three probabilities, including, {0,1}. We can use pROC package which is far easier to work with, as shown below.
```{r ROC}
par(pty="s")
gbm_roc<-roc(tstdata$target~gbm_predict,
             plot=TRUE,
             print.auc=TRUE,
             col='blue',
             lwd=3,
             legacy.axes=TRUE,
             main='ROC Curves(GBM)')

gbm_auc<-auc(gbm_roc)

```
Now let us run xgboost

```{r xgboost}
require(xgboost)
require(Matrix)

trdatamx<-sparse.model.matrix(target~.-1,data=trdata)

tstdatamx<-sparse.model.matrix(target~.-1,data=tstdata)

xgb_model<-xgboost(data=trdatamx,
                   label=trdata$target,
                   max_depth = 2,
  eta = 1, 
  nrounds = 2,
  nthread = 2, 
  #objective = "multi:softmax",# to get classes
  objective = "multi:softprob",# to get probs
  num_class=2)
#xgb_pred <- predict(xgb_model,tstdatamx,type='response')

xgb_pred <- predict(xgb_model,tstdatamx)
xgb_pred_df<-data.frame(actual=tstdata$target,
prob.1=xgb_pred[seq(0,length(xgb_pred),by=2)],    prob.0=xgb_pred[seq(1,length(xgb_pred),by=2)])
xgb_pred_class<-ifelse(xgb_pred_df$prob.0>xgb_pred_df$prob.1,0,1)
TN<-xgb_pred[(xgb_pred_class==1)&(xgb_pred==tstdata$target)]
TP<-xgb_pred[(xgb_pred_class==0)&(xgb_pred==tstdata$target)]
(xgb_tbl<-table(tstdata$target,xgb_pred_class))
(xgb_cfm<-caret::confusionMatrix(xgb_tbl))

#str(xgb_pred)

xgb_roc<-roc(tstdata$target~xgb_pred[seq(0,length(xgb_pred),by=2)],
             plot=TRUE,
             print.auc=TRUE,
             col='red',
             lwd=3,
             legacy.axes=TRUE,
             main='ROC Curves(XGBoost)')

xgb_auc<-auc(xgb_roc)
```
## Comparison with GLM
How different are these estimates?
let us run glm and get the estimates...
```{r glm_comparison}
glm.model<-glm(target~.,data=trdata,family='binomial')

glm.pred<-predict(glm.model,
                  tstdata[,-c(14)],
                  type='response')
glm.class<-ifelse(glm.pred<0.50,0,1)
glm_tbl<-table(tstdata$target,glm.class)
glm_cfm<-caret::confusionMatrix(glm_tbl)

glm_roc<-roc(tstdata$target~glm.pred,
             plot=TRUE,
             print.auc=TRUE,
             col='black',
             lwd=3,
             legacy.axes=TRUE,
             main='ROC Curves(GLM-Logistic)')

glm_auc<-auc(glm_roc)

```
## Comparative Analysis

Let us review the performance (standard metrics) for GLM,GBM and XGBoost.
```{r Performance Analysis}
accuracy<-function(xt)sum(diag(xt))/sum(xt)

gbm_tbl

gbm_cfm

glm_tbl

glm_cfm

xgb_tbl

xgb_cfm
gbm_acc<-accuracy(gbm_tbl)
glm_acc<-accuracy(glm_tbl)
xgb_acc<-accuracy(xgb_tbl)
print(paste("gbm:",gbm_acc,sep=""))
print(paste("glm:",glm_acc,sep=""))
print(paste("xgb:",xgb_acc,sep=""))

perfdf<-data.frame(algo=c("glm","gbm","xgb"),
                   acc=c(glm_acc,gbm_acc,xgb_acc),
                   auc=c(glm_auc,gbm_auc,xgb_auc))
```

## Stacking
For stacking we can run another meta learner
where the dependent variable is actual and the independent variables are the predictions from
the different classifiers.
```{r stacking_setup}
stackeddf<-data.frame(actual=tstdata$target,
                      glmpred=glm.class,
                      gbmpred=gbm_predicted,
                      xgbpred=xgb_pred_class)

head(stackeddf)
tail(stackeddf)
```
Or we can take the majority vote as follows
```{r stacking_by_vote}
stacked_mean<-unlist(apply(stackeddf[,2:4],1,mean))
#round()
stacked_label<-ifelse(stacked_mean<0.5,0,1)
stackeddf<-cbind(stackeddf,stacked=stacked_label)
(stbl<-table(stackeddf$actual,stacked_label))
(scfm<-caret::confusionMatrix(stbl))
stk_acc<-sum(diag(stbl))/sum(stbl)

stk_pred_prob<-(xgb_pred[seq(0,length(xgb_pred),by=2)]+
                           glm.pred+gbm_predict)/3

stk_roc<-roc(tstdata$target~glm.pred,
             plot=TRUE,
             print.auc=TRUE,
             col='black',
             lwd=3,
             legacy.axes=TRUE,
             main='ROC Curves(Stacked(GLM,GBM,XGB))')

stk_auc<-auc(stk_roc)

perfdf<-rbind(perfdf,c("STK",stk_acc,stk_auc))

perfdf

```
The improvement is anemic. Giving all three learners equal weight is not useful and so we might entertain giving them weights. And, thus, there are many interesting experiments we can undertake.

For now, we have finally, reached the end of our ML trail. We have introduced numerous ideas from Machine Learning and we have see computational techniques, if only you can focus on the moon and not on the pointer.

# Conclusion

From a Big Data perspective, notice that boosting is iterative and parallelizing boosting is either non-trivial or impossible. 
In the next lecture, we will review Big Data and Machine Learning and reserve the rest for student feed back.

During  the course I have used absolute numbers and they are NOT rules. You can change them working with your stake-holders. That is what I use and I make that part of ALL my SOW.

### 2020-11-25 DATA698/08 - Report Visualizations.rmd
Minor changes
Updated visualization generating script

@@
   mutate(Epoch = as.numeric(strsplit(Epoch, "/")[[1]][1])+1)
+```
 
+## Images by Month Chart
+
+```{r}
+months_df <- read.csv("report/all_month_counts.csv") %>%
+  mutate(Images = ifelse(Images == 0, NA, Images)) %>%
+  mutate(Plant = ifelse(Plant.Type == "Poison Ivy", "A", "ERROR")) %>%
+  mutate(Plant = ifelse(Plant.Type == "Box Elder", "B", Plant)) %>%
+  mutate(Plant = ifelse(Plant.Type == "Bramble", "C", Plant)) %>%
+  mutate(Plant = ifelse(Plant.Type == "Virginia Creeper", "D", Plant)) %>%
+  mutate(Plant = as.factor(Plant)) %>%
+  mutate(Plant = recode(Plant, A = "Poison Ivy", B = "Box Elder", C = "Bramble", D = "Virginia Creeper")) %>%
+  select(-Plant.Type)
+
+ggplot(months_df, aes(x=Month, y=Images, color=Plant, fill=Plant)) +
+  geom_col() +
+  scale_color_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") + 
+  facet_wrap(~Plant) +
+  theme(legend.position = "none") +
+  #xlim(1, 12)
+  scale_x_discrete(name ="Month",  limits=c("1","2","3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))
+```
+
+## Model Performance Charts
+
+```{r}
+viz_data <- yolov3_results %>% select(model, n_images, Epoch, mAP.0.5)
+viz_data <- yolov5_results %>% select(model, n_images, Epoch, mAP.0.5) %>% rbind(viz_data) 
+viz_data$model <- as.factor(viz_data$model)
+viz_data$n_images <- as.factor(viz_data$n_images)
+
+ggplot(viz_data) + 
+  geom_line(aes(x=Epoch, y=mAP.0.5, color=n_images, group=n_images)) + 
+  facet_wrap(~model) + 
+  scale_color_brewer(palette = "Set1") +
+  labs(color = "Number of Images") +
+  ylab("mAP@0.5") +
+  theme(legend.position = "bottom")
+```
+
+## YOLOv3
+
+```{r}
+viz_data <- yolov3_results %>% select(n_images, Epoch, obj) %>% mutate(set = "Training") %>% rename(measure = obj)
+viz_data <- yolov3_results %>% select(n_images, Epoch, Val.objectness) %>% mutate(set = "Test") %>% rename(measure = Val.objectness) %>% rbind(viz_data) 
+viz_data$n_images <- as.factor(viz_data$n_images)
+
+ggplot(viz_data) + 
+  geom_line(aes(x=Epoch, y=measure, color=set, group=set)) + 
+  facet_wrap(~n_images) + 
+  scale_color_brewer(palette = "Set1") +
+  labs(color = "Data Set") +
+  theme(legend.position = "bottom") +
+  ylab("Object Detection Error")
+```
+
+```{r}
+viz_data <- yolov3_results %>% select(n_images, Epoch, cls) %>% mutate(set = "Training") %>% rename(measure = cls)
+viz_data <- yolov3_results %>% select(n_images, Epoch, Val.classification) %>% mutate(set = "Test") %>% rename(measure = Val.classification) %>% rbind(viz_data) 
+viz_data$n_images <- as.factor(viz_data$n_images)
+
+ggplot(viz_data) + 
+  geom_line(aes(x=Epoch, y=measure, color=set, group=set)) + 
+  facet_wrap(~n_images) + 
+  scale_color_brewer(palette = "Set1") +
+  labs(color = "Data Set") +
+  theme(legend.position = "bottom") +
+  ylab("Classification Error")
+```
+
+## YOLOv5
+
+```{r}
+viz_data <- yolov5_results %>% select(n_images, Epoch, obj) %>% mutate(set = "Training") %>% rename(measure = obj)
+viz_data <- yolov5_results %>% select(n_images, Epoch, Val.objectness) %>% mutate(set = "Test") %>% rename(measure = Val.objectness) %>% rbind(viz_data) 
+viz_data$n_images <- as.factor(viz_data$n_images)
+
+ggplot(viz_data) + 
+  geom_line(aes(x=Epoch, y=measure, color=set, group=set)) + 
+  facet_wrap(~n_images) + 
+  scale_color_brewer(palette = "Set1") +
+  labs(color = "Data Set") +
+  theme(legend.position = "bottom") +
+  ylab("Object Detection Error")
+```
+
+```{r}
+viz_data <- yolov5_results %>% select(n_images, Epoch, cls) %>% mutate(set = "Training") %>% rename(measure = cls)
+viz_data <- yolov5_results %>% select(n_images, Epoch, Val.classification) %>% mutate(set = "Test") %>% rename(measure = Val.classification) %>% rbind(viz_data) 
+viz_data$n_images <- as.factor(viz_data$n_images)
+
+ggplot(viz_data) + 
+  geom_line(aes(x=Epoch, y=measure, color=set, group=set)) + 
+  facet_wrap(~n_images) + 
+  scale_color_brewer(palette = "Set1") +
+  labs(color = "Data Set") +
+  theme(legend.position = "bottom") +
+  ylab("Classification Error")
 ```
@@
   rbind(yolo_results)
-```
 
-```{r}
-viz_data <- yolov3_results %>%
-  mutate(Model = as.factor(paste0(model,": ", n_images, " images")))
-ggplot(viz_data, aes(x=Epoch, y=mAP.0.5, color=Model, group=Model)) + geom_line() + scale_color_brewer(palette = "Set1")
-```
-
-```{r}
-viz_data <- yolov5_results %>%
-  mutate(Model = as.factor(paste0(model,": ", n_images, " images")))
-ggplot(viz_data, aes(x=Epoch, y=mAP.0.5, color=Model, group=Model)) + geom_line() + scale_color_brewer(palette = "Set1")
-```
-
-```{r}
 for(m in unique(yolo_results$Measure)){
@@
 ```
+
+```{r}
+ggplot(yolov3_results, aes(R, P, color=as.factor(paste(model, n_images)))) + geom_point()
+```

### 2020-12-15 DATA622/HW2.rmd
Initial commit of HW 2

---
title: "HW2"
author: "Mike Silva"
output:
  pdf_document: default
  html_document: default
---
```{r, message=FALSE, warning=FALSE}
library(caret)
library(tictoc)
library(dplyr)
library(tidyr)
library(ROCR)
library(e1071)
library(kableExtra)
library(randomForest)
```

For this homework I will use Logistic Regression and SVM on the heart data set.  Instructions for the homework are in *italics*

```{r}
path <- "heart.csv"
df <- read.csv(path, stringsAsFactors = FALSE)
# Clean up the first column
names(df)[[1]] <- "age"
```

## PART-A

### Base Metrics

*Do a 80/20 split and determine the Accuracy, AUC and as many metrics as returned by the Caret
package (confusionMatrix) Call this the base_metric. Note down as best as you can development
(engineering) cost as well as computing cost(elapsed time).*

```{r}
set.seed(43)
in_train <- createDataPartition(df$target, p = 0.8, list = FALSE, times = 1)
train <- df[in_train, ]
test <- df[-in_train, ]
```

These are some general purpose utility functions to get the evaluations metrics

```{r}
get_cm_metrics <- function(pred, truth){
  cm <- confusionMatrix(table(pred, truth))
  metrics <- c(cm$overall["Accuracy"], cm$byClass)
  return(data.frame(as.list(metrics)))
}

get_auc <- function(pred, truth){
  temp <- prediction(pred, truth)
  prf <- performance(temp, "auc")
  return(prf@y.values[[1]])
}

get_elapsed <- function(tictoc){
  return(tictoc$toc - tictoc$tic)
}
```

#### Logistic Regression Classifier

It took me 20 minutes to develop this logistic regression pipeline with the evaluation metrics.  I also developed this logistic regression version of predict that will return the classes instead of the probabilities.

```{r}
lr_predict <- function(lr_model, df){
  lr_probs <- predict(lr_model, df, type="response")
  lr_yhat <- ifelse(lr_probs < 0.5, 0, 1)
  return(lr_yhat)
}

lr_engineering_cost <- 20
```

Now to train the baseline logistics regression classifier.

```{r}
tic("Logistic Regression Training")
glm_model <- glm(target~., data = df, family = "binomial")
summary_glm_model <- summary(glm_model)
coef_summary_glm_model <- coef(summary_glm_model)
# I only want statistically significant variables
lr_formula <- paste("target~", paste(row.names(coef_summary_glm_model[coef_summary_glm_model[, 4] < 0.05, ]), collapse = "+"), sep = "")
lr_model <- glm(lr_formula, data = df, family = "binomial")
lr_time <- toc()
lr_time <- get_elapsed(lr_time)
lr_yhat <- lr_predict(lr_model, test)
lr_auc <- get_auc(lr_yhat, test$target)
base_metric <- get_cm_metrics(lr_yhat, test$target) %>%
  mutate(AUC = lr_auc,
         model = "LR",
         engineering_cost = lr_engineering_cost,
         computing_cost = lr_time)
```

#### SVM Classifier

The SVM  classifier did not take as much engineering time since most of it was developed for the LR model.  Here's the SVM's helper functions

```{r}
svm_predict <- function(svm_model, df){
  svm_probs <- predict(svm_model, df)
  svm_yhat <- ifelse(svm_probs < 0.5, 0, 1)
  return(svm_yhat)
}

svm_engineering_cost <- 5
```

And now we train...

```{r}
tic("SVM Training")
svm_model <- svm(target~., data = df, probability = TRUE)
svm_time <- toc()
svm_time <- get_elapsed(svm_time)
svm_yhat <- svm_predict(svm_model, test)
svm_auc <- get_auc(svm_yhat, test$target)
base_metric <- get_cm_metrics(svm_yhat, test$target) %>%
  mutate(AUC = svm_auc,
         model = "SVM",
         engineering_cost = svm_engineering_cost,
         computing_cost = svm_time) %>%
  bind_rows(base_metric)
```

### K-Fold Cross Validated Metrics

*Start with the original dataset and set a seed (43). Then run a cross validation of 5 and 10 of the model
on the training set. Determine the same set of metrics and compare the cv_metrics with the
base_metric. Note down as best as you can development (engineering) cost as well as computing
cost(elapsed time).*

#### Logistic Regression

I have built this out to run both cross validations.  It took about 20 minutes to build which cost I split between the models.

```{r}
n <- nrow(df)
ks <- c(5, 10)
cv_lr_engineering_cost <- 20 / length(ks)

for (k in ks){
  set.seed(43)
  folds <- split(1:n, cut(1:n, quantile(1:n, probs = seq(0, 1, by = 1 / k))))
  ridx <- sample(1:nrow(df), nrow(df), replace = FALSE)

  tic(paste(k, "Fold CV LR Training"))

  k_fold_metrics <- do.call("rbind", lapply(folds, FUN = function(idx, data = df[ridx, ]) {
    cv_train <- data[-idx, ]
    cv_test <- data[idx, ]
    lr_model <- glm(lr_formula, data = cv_train, family = "binomial")
    lr_yhat <- lr_predict(lr_model, cv_test)
    lr_auc <- get_auc(lr_yhat, cv_test$target)
    temp <- get_cm_metrics(lr_yhat, cv_test$target) %>%
      mutate(AUC = lr_auc) 
    return(temp)
  }))

  cv_lr_time <- toc()
  cv_lr_time <- get_elapsed(cv_lr_time)

  k_fold_metrics <- k_fold_metrics %>%
    summarise_all(list(mean)) %>%
    mutate(model = paste(k, "Fold LR"),
           engineering_cost = cv_lr_engineering_cost,
           computing_cost = cv_lr_time)
  if(exists("cv_metrics")){
    cv_metrics <- bind_rows(cv_metrics, k_fold_metrics)
  } else {
    cv_metrics <- k_fold_metrics
  }
}
```
#### SVM

Again the SVM benefited from the development of the workflow for the LR model.  I once again split the developmental costs between the models.

```{r}
cv_svm_engineering_cost <- 5 / length(ks)

for (k in ks){
  set.seed(43)
  folds <- split(1:n, cut(1:n, quantile(1:n, probs = seq(0, 1, by = 1 / k))))
  ridx <- sample(1:nrow(df), nrow(df), replace = FALSE)

  tic(paste(k, "Fold CV SVM Training"))

  k_fold_metrics <- do.call("rbind", lapply(folds, FUN = function(idx, data = df[ridx, ]) {
    cv_train <- data[-idx, ]
    cv_test <- data[idx, ]
    svm_model <- svm(target~., data = df, probability = TRUE)
    svm_yhat <- svm_predict(svm_model, cv_test)
    svm_auc <- get_auc(svm_yhat, cv_test$target)
    temp <- get_cm_metrics(svm_yhat, cv_test$target) %>%
      mutate(AUC = svm_auc) 
    return(temp)
  }))

  cv_svm_time <- toc()
  cv_svm_time <- get_elapsed(cv_svm_time)

  cv_metrics <- k_fold_metrics %>%
    summarise_all(list(mean)) %>%
    mutate(model = paste(k, "Fold SVM"),
           engineering_cost = cv_svm_engineering_cost,
           computing_cost = cv_svm_time) %>%
    bind_rows(cv_metrics)
}
```

#### Comparison

The performance of the model increases with the number of folds.  This is said by looking at the AUC which is a good summary measure of the models performance.  The Base LR model has an AUC of about 0.81.  The K-5 CV LR model's AUC increases to 0.83 and K-10 CV LR increases to 0.84.  The same could be said with SVM where the AUC increases from a base of 0.88 to 0.913 and 0.914 for K-5 and 10 respectively.  Cross validation improves the model performance.  There is some computational cost.  The computational time on the SVM increased by a factor of 3 for K-5 CV and a factor of 6 (roughly) for K-10 CV.  I don't find the performance gain from the K-5 SMV to K-10 to be "worth it".  There is a bit of an uptick for logistic regression for a K-10 model in computational cost.  The performance metrics are summarized in the two tables below:

**Table 1. Base Metrics**

```{r, echo=FALSE}
base_metric %>% 
  kable() %>%
  kable_styling()
```

**Table 2. CV Metrics**

```{r, echo=FALSE}
cv_metrics %>% 
  kable() %>%
  kable_styling()
```

### Bootstrapped Metrics

*Start with the original dataset and set a seed (43) Then run a bootstrap of 200 resamples and compute
the same set of metrics and for each of the two classifiers build a three column table for each
experiment (base, bootstrap, cross-validated). Note down as best as you can development (engineering)
cost as well as computing cost(elapsed time).*

#### Bootstrapped Logistics Regression

First I will develop the bootstrap workflow for LR.  This is not the most efficient way of doing it but it works.

```{r}
set.seed(43)
n_runs <- 200
bs_lr_engineering_cost <- 10

tic("Bootstrap LR Training")

for(i in 1:n_runs){
  in_train <- sample(1:nrow(df), nrow(df), replace=T)
  train <- df[in_train, ]
  lr_model <- glm(lr_formula, data = train, family = "binomial")
  lr_yhat <- lr_predict(lr_model, df)
  lr_auc <- get_auc(lr_yhat, df$target)
  temp <- get_cm_metrics(lr_yhat, df$target) %>%
    mutate(AUC = lr_auc)
  if(exists("bs_lr_metrics")){
    bs_lr_metrics <- bind_rows(bs_lr_metrics, temp)
  } else {
    bs_lr_metrics <- temp
  }
}

bs_lr_time <- toc()
bs_lr_time <- get_elapsed(bs_lr_time)

bs_lr_metrics <- bs_lr_metrics %>%
    summarise_all(list(mean)) %>%
    mutate(model = "Bootstrapped LR",
           engineering_cost = bs_lr_engineering_cost,
           computing_cost = bs_lr_time)
```

#### Bootstraped SVM

And here's the SVM which is basically the LR with some minor tweaks...

```{r}
set.seed(43)
bs_svm_engineering_cost <- 5

tic("Bootstrap SVM Training")

for(i in 1:n_runs){
  in_train <- sample(1:nrow(df), nrow(df), replace=T)
  train <- df[in_train, ]
  svm_model <- svm(target~., data = train, probability = TRUE)
  svm_yhat <- svm_predict(svm_model, df)
  svm_auc <- get_auc(svm_yhat, df$target)
  temp <- get_cm_metrics(svm_yhat, df$target) %>%
    mutate(AUC = svm_auc)
  if(exists("bs_svm_metrics")){
    bs_svm_metrics <- bind_rows(bs_svm_metrics, temp)
  } else {
    bs_svm_metrics <- temp
  }
}

bs_svm_time <- toc()
bs_svm_time <- get_elapsed(bs_svm_time)

bs_svm_metrics <- bs_svm_metrics %>%
    summarise_all(list(mean)) %>%
    mutate(model = "Bootstrapped SVM",
           engineering_cost = bs_svm_engineering_cost,
           computing_cost = bs_svm_time)

bs_metrics <- bind_rows(bs_lr_metrics, bs_svm_metrics)
```

Here's how the bootstrapped models preformed:

**Table 3. Bootstrapped Metrics**

```{r, echo=FALSE}
bs_metrics %>% 
  kable() %>%
  kable_styling()
```

### Comparison of Experiments

Now to compare the base, CV and bootstrap results.  I will only be looking at the AUC.

**Table 4. Experiment Results**

```{r, echo=FALSE}
experiment_results <- base_metric %>% 
  select(model, AUC) %>%
  mutate(experiment = "Base")

experiment_results <- cv_metrics %>% 
  filter(model %in% c("10 Fold LR", "10 Fold SVM")) %>% 
  mutate(model = ifelse(model == "10 Fold LR", "LR", "SVM")) %>%
  select(model, AUC) %>%
  mutate(experiment = "10-Fold CV") %>%
  bind_rows(experiment_results)

experiment_results <- bs_metrics %>%
  mutate(model = ifelse(model == "Bootstrapped LR", "LR", "SVM")) %>%
  select(model, AUC) %>%
  mutate(experiment = "Bootstrapped") %>%
  bind_rows(experiment_results)

experiment_results %>%
  pivot_wider(id_cols = model, names_from = experiment, values_from = AUC) %>%
  select(model, Base, `10-Fold CV`, Bootstrapped) %>%
  kable() %>%
  kable_styling()
```

The Highest AUC is found in the cross validated models.  The bootstrapped models preformed a little better than the base models but that improvement came at a considerable cost:

**Table 5. Computational Cost**

```{r, echo=FALSE}
experiment_results <- base_metric %>% 
  select(model, computing_cost) %>%
  mutate(experiment = "Base")

experiment_results <- cv_metrics %>% 
  filter(model %in% c("10 Fold LR", "10 Fold SVM")) %>% 
  mutate(model = ifelse(model == "10 Fold LR", "LR", "SVM")) %>%
  select(model, computing_cost) %>%
  mutate(experiment = "10-Fold CV") %>%
  bind_rows(experiment_results)

experiment_results <- bs_metrics %>%
  mutate(model = ifelse(model == "Bootstrapped LR", "LR", "SVM")) %>%
  select(model, computing_cost) %>%
  mutate(experiment = "Bootstrapped") %>%
  bind_rows(experiment_results)

experiment_results %>%
  pivot_wider(id_cols = model, names_from = experiment, values_from = computing_cost) %>%
  select(model, Base, `10-Fold CV`, Bootstrapped) %>%
  kable() %>%
  kable_styling()
```

These results are not parallelized so there could be some decrease in the computational costs, but CV gives a lot of bang for the buck.

## PART-B

*For the same dataset, set seed (43) split 80/20.

Using randomForest grow three different forests varying the number of trees at least three times. Start
with seeding and fresh split for each forest. Note down as best as you can development (engineering)
cost as well as computing cost(elapsed time) for each run. And compare these results with the
experiment in Part A.*

```{r}
set.seed(43)
in_train <- createDataPartition(df$target, p = 0.8, list = FALSE, times = 1)
train <- df[in_train, ]
test <- df[-in_train, ]
train$target <- as.factor(train$target)
test$target <- as.factor(test$target)

rf_engineering_cost <- 15 / 3

for (n in c(100, 250, 1000)) {
  set.seed(43)
  tic(paste("Training RF ", n))
  rf_1000_model <- randomForest(target~., data = train, ntree = n)
  rf_time <- toc()
  rf_time <- get_elapsed(rf_time)
  rf_yhat <- predict(rf_1000_model, test)
  rf_auc <- get_auc(as.numeric(as.character(rf_yhat)), as.numeric(as.character(test$target)))
  temp <- get_cm_metrics(rf_yhat, test$target) %>%
    mutate(AUC = rf_auc,
         model = paste("RF", n),
         engineering_cost = rf_engineering_cost,
         computing_cost = rf_time)
  if(exists(("rf_metrics"))){
    rf_metrics <- bind_rows(rf_metrics, temp)
  } else {
    rf_metrics <- temp
  }
}
```

How did the random forests preform?  Table 6 below summarizes them.  The performance is not particularly strong.  The base SVM did a better job than the largest random forest.  Both bootstrapped models out preformed the random forests, and the 10 fold CV models did a lot better.

**Table 6. Random Forest Metrics**

```{r, echo=FALSE}
rf_metrics %>% 
  kable() %>%
  kable_styling()
```

## PART-C

*Include a summary of your findings. Which of the two methods bootstrap vs cv do you recommend to
your customer? And why? Be elaborate. Including computing costs, engineering costs and model
performance. Did you incorporate Pareto's maxim or the Razor and how did these two heuristics
influence your decision?*

In this exercise I developed two classifiers. One was a logistic regression model and the other a SVM.  I ran some experiments with these models to see if I could improve the model's performance.  The two experiments were to use cross validation and bootstrapping.  In looking at the accuracy of the model as measured using AUC, and if I had to recommend only one model I would recommend the 5-fold cross-validated SVM model.  It showed considerable improvement from the base model.  While the 10-fold SVM did a little better, the computational costs don't justify it.  I also examined 3 random forests as possible solutions but the cross validated SVMs  preformed better than the random forests.

This analysis applies Akum's razor by focusing on AUC instead of all the other metrics.  I could have just as easily focused on positive predictive value if that was more important to the client but I assumed the overall accuracy was the most important.  I applied sort of a Pareto analysis in recommending the K-5 SVM.  Yes the K-10 did better but in terms of computational costs it cost a lot more to squeeze out a little more performance.  It had hit the point of diminishing returns and so it was easier for me to recommend something that got the most bang for the buck.
